### 2021-04-16 00-Data-Processing/00-Data-Processing-for-Food-Banks-Accessibility.Rmd
Initial commit with first steps of data preparation

---
title: "00-Data-Processing-for-Food-Banks-Accessibility"
---

## Introduction

In this notebook we create some examples of processing the data for the analysis of accessibility to food banks in Hamilton. Particularly, we're interested in spatial interpolation of the population for small areas.

## Preliminaries

Clear environment:
```{r}
rm(list = ls())
```

Load packages used in the notebook:
```{r message = FALSE}
library(cancensus)
library(disk.frame)
#library(pycno)
library(readr)
library(r5r) # the r5r package requires Java Development Kit version 11, which can be downloaded from https://www.oracle.com/java/technologies/javase-jdk11-downloads.html
library(sf)
library(smoothr)
library(tidyverse)
library(tmap)
library(tmaptools)
```

Define some parameters for `disk.frame` and `r5r`:
```{r}
setup_disk.frame()
options(scipen = 999)
options(java.parameters = "-Xmx6G")
options(future.globals.maxSize = Inf)
```

Data used in this notebook was retrieved from two sources: Open Hamilton and the 2016 Canadian Census. Open Hamilton is an online repository of data from the City of Hamilton. The provenance of the datasets is below.

## Provenance of data:

### Hamilton CMA boundaries

Statistics Canada. Load:
```{r}
load("input-data-files/hamilton_cma.RData")
#load("input-data-files/hamilton_da_2016.RData")
```

### Hamilton Ward boundaries

Open Data Hamilton. Read:
```{r}
wards <- st_read("input-data-files/Ward_Boundaries.shp")
```

Add label for type of ward (the urban/suburban/rural classification is by the planning teams; see https://www.hamilton.ca/develop-property/planning-applications/development-applications-mapping)
```{r}
wards$Type <- c("Urban",
                "Urban",
                "Urban",
                "Urban",
                "Urban",
                "Suburban",
                "Suburban",
                "Suburban",
                "Suburban",
                "Suburban",
                "Suburban",
                "Suburban",
                "Suburban",
                "Suburban",
                "Suburban")
```

Merge boundaries by type:
```{r}
wards <- wards %>% 
  group_by(Type) %>%
  summarize()
```

```{r}
ggplot(wards) + 
  geom_sf(aes(fill = Type))
```

Read rural boundary:
```{r}
rural <- st_read("input-data-files/Rural_Boundary.shp")
rural <- rural %>%
  transmute(Type = "Rural")
```

Plot rural boundary:
```{r}
ggplot(rural) +
  geom_sf()
```

Obtain difference between urban and suburban regions and rural boundary:
```{r}
st_difference(wards, rural) %>%
  ggplot() +
  geom_sf(aes(fill = Type))
```

Notice the detritus after the difference. Drop crumbs to obtain a cleaner version of the boundaries:
```{r}
sub_urban <- st_difference(wards, rural) %>%
  transmute(Type) %>%
  smoothr::drop_crumbs(threshold = 5000)

ggplot(sub_urban) +
  geom_sf(aes(fill = Type))
```

Bind the boundaries with all three urban types:
```{r}
urban_types <- rbind(sub_urban,
                     rural)

ggplot(urban_types) +
  geom_sf(aes(fill = Type))
```

### Location of food banks

Read file with locations:
```{r}
foodbank_locations <- read_csv("input-data-files/foodbanks-hamilton.csv",
                               col_types = cols(ID = col_double(),
                                    NAME = col_character(),
                                    TYPE = col_character(),
                                    ADDRESS = col_character(),
                                    COVIDClose = col_double(),
                                    latitude = col_double(),
                                    longitude = col_double(),
                                    name = col_character(),
                                    desc = col_character(),
                                    source = col_character(),
                                    precision = col_character()
                               ))
```

COVIDClose is an indicator variable for whether this bank closed during the pandemic. Convert to factor:
```{r}
foodbank_locations <- foodbank_locations %>%
  mutate(COVIDClose = replace_na(COVIDClose, 0),
         COVIDClose = factor(COVIDClose,
                             levels = c(0, 1),
                             labels = c("No", "Yes")))
```

Convert to simple features:
```{r}
foodbank_locations <- foodbank_locations %>%
  st_as_sf(coords = c("longitude", "latitude"), 
           crs = 4326) %>%
  st_transform(crs = 26917)
```

Map location of foodbanks:
```{r}
ggplot() +
  geom_sf(data = hamilton_cma) +
  geom_sf(data = foodbank_locations,
          aes(shape = COVIDClose,
              color = COVIDClose))
```

### `dwelling_network_points_2016.RData`

This is based on parcel-level data, approximated to the nearest point on the road network. For processing details, see:

https://github.com/paezha/Accessibility-to-Schools-Hamilton-Equity

```{r}
load("input-data-files/dwelling_network_points_2016.RData")
```

### Census data

We can get an API key to use `cancensus` from [CensusMapper](https://censusmapper.ca/). Once you have an API key, it can be stored locally this way:
```{r eval=FALSE}
set_api_key("your-key", install = TRUE)
```

Once you have your API you can use `cancensus` like in this example.

Use `cancensus` to check and download census data. Check the regions:
```{r}
list_census_regions('CA16') %>% 
  dplyr::filter(level == "CMA", name %in% c("Hamilton"))
```

We need PR_UID 35.

List census data sets:
```{r}
list_census_datasets()
```

We need CA16.

Search variables:
```{r}
find_census_vectors("income", 
                    dataset = "CA16", 
                    type = "total", 
                    query_type = "keyword", 
                    interactive = FALSE)
```

We need v_CA16_2406 to v_CA16_2425: these are the number of households at certain levels of total income. Retrieve data:
```{r}
data_da_2016 <- get_census(dataset='CA16', 
                          regions=list(CMA=c("35537")),
                          vectors=c("v_CA16_2406",
                                    "v_CA16_2407",
                                    "v_CA16_2408",
                                    "v_CA16_2409",
                                    "v_CA16_2410",
                                    "v_CA16_2411",
                                    "v_CA16_2412",
                                    "v_CA16_2413",
                                    "v_CA16_2414",
                                    "v_CA16_2415",
                                    "v_CA16_2416",
                                    "v_CA16_2417",
                                    "v_CA16_2418",
                                    "v_CA16_2419",
                                    "v_CA16_2420",
                                    "v_CA16_2421",
                                    "v_CA16_2422",
                                    "v_CA16_2423",
                                    "v_CA16_2424",
                                    "v_CA16_2425",
                                    "v_CA16_2426"),
                          level='DA', 
                          use_cache = FALSE,
                          geo_format = "sf")
```

Plot population:
```{r}
tmap_mode(mode = "view")

data_da_2016 %>%
  mutate(income_less20k = (`v_CA16_2406: Under $5,000` + 
                  `v_CA16_2407: $5,000 to $9,999` + 
                  `v_CA16_2408: $10,000 to $14,999` +
                  `v_CA16_2409: $15,000 to $19,999`)) %>%
  tm_shape() +
  tm_polygons("income_less20k",
              alpha = 0.8)
```

### Transportation Tomorrow Survey

I downloaded a cross-tabulation of travel mode by age by traffic analysis zone. Read the shape file with the traffic analysis zone boundaries:
```{r}
tts_taz <- st_read("input-data-files/tts2016_2006zn_2018_region.shp")
```

Filter Hamilton (region 6):
```{r}
tts_taz <- tts_taz %>%
  dplyr::filter(region == 6)

ggplot() +
  geom_sf(data = tts_taz)
```

Read transportation data:
```{r}
travel_modes <- read_csv("input-data-files/ttsCross70628.csv",
                         col_types = cols(gta06_hhld = col_double(),
                                          income = col_double(),
                                          total = col_double(),
                                          Mode = col_character()
                                          ))
```

Code | Income
-- | --
1	| \$0 to $14999
2	| \$15000 to $39999
3	| \$40000 to $59999
4	| \$60000 to $99999
5	| \$100000 to $124999
6	| \$125000 and above
7	Decline / don't know

Filter modes:
```{r}
travel_modes <- travel_modes %>%
  dplyr::filter(Mode == "Auto driver" |
                  Mode == "Transit (Excluding GO rail)" |
                  Mode == "Walk")
```

THE NEXT NEEDS TO BE CHANGED TO DO THE ANALYSIS BY INCOME GROUP, NOT BY AGE GROUP

### Ages 55-69

Filter age:
```{r}
travel_modes_1 <- travel_modes %>%
  dplyr::filter(age >= 55 & age < 70)
```

Summarize trips by mode by traffic analysis zone:
```{r}
travel_modes_1 <- travel_modes_1 %>%
  group_by(gta06_hhld, Mode) %>%
  summarize(trips = sum(total),
            .groups = "drop")
```

Pivot wider and rename columns:
```{r}
travel_modes_1 <- travel_modes_1 %>%
  rename(TAZUID = gta06_hhld) %>%
  pivot_wider(names_from = Mode,
              values_from = trips,
              values_fill = 0) %>%
  rename(Driver = `Auto driver`, 
         Transit = `Transit (Excluding GO rail)`)
```

Join geometry:
```{r}
travel_modes_1 <- tts_taz %>%
  dplyr::transmute(TAZUID = gta06, geometry) %>%
  left_join(travel_modes_1 %>%
              st_drop_geometry(),
            by = "TAZUID") %>%
  st_as_sf()
```

Plot trips as driver:
```{r}
ggplot() +
  geom_sf(data = travel_modes_1,
          aes(fill = Driver))
```

Plot trips transit:
```{r}
ggplot() +
  geom_sf(data = travel_modes_1,
          aes(fill = Transit))
```

Plot walking trips:
```{r}
ggplot() +
  geom_sf(data = travel_modes_1,
          aes(fill = Walk))
```

### Ages 70+

Filter age:
```{r}
travel_modes_2 <- travel_modes %>%
  dplyr::filter(age >= 70)
```

Summarize trips by mode by traffic analysis zone:
```{r}
travel_modes_2 <- travel_modes_2 %>%
  group_by(gta06_hhld, Mode) %>%
  summarize(trips = sum(total),
            .groups = "drop")
```

Pivot wider and rename columns:
```{r}
travel_modes_2 <- travel_modes_2 %>%
  rename(TAZUID = gta06_hhld) %>%
  pivot_wider(names_from = Mode,
              values_from = trips,
              values_fill = 0) %>%
  rename(Driver = `Auto driver`, 
         Transit = `Transit (Excluding GO rail)`)
```

Join geometry:
```{r}
travel_modes_2 <- tts_taz %>%
  dplyr::transmute(TAZUID = gta06, geometry) %>%
  left_join(travel_modes_2 %>%
              st_drop_geometry(),
            by = "TAZUID") %>%
  st_as_sf()
```

Plot trips as driver:
```{r}
ggplot() +
  geom_sf(data = travel_modes_2,
          aes(fill = Driver))
```

Plot trips transit:
```{r}
ggplot() +
  geom_sf(data = travel_modes_2,
          aes(fill = Transit))
```

Plot walking trips:
```{r}
ggplot() +
  geom_sf(data = travel_modes_2,
          aes(fill = Walk))
```

Rename tables:
```{r}
modes_55to69 <- travel_modes_1
modes_70plus <- travel_modes_2
```

## Join TAZ identifiers to parcel table

This will be a spatial join:
```{r}
nearest_network_point_2016 <- nearest_network_point_2016 %>%
  st_join(tts_taz %>%
            transmute(TAZUID = gta06))
```

## Routing for entire system

I used [BBBike](https://download.bbbike.org/osm/bbbike/) to extract OSM data for Hamilton. The name of the file is `planet_-80.279,43.048_-79.318,43.459.osm.pbf`. Copy to folder `r5_graph`.

Set Up R5 Routing. First define the path to where the graph is located:
```{r set up r5 path, include=FALSE}
r5_path <- file.path("./r5_graph")
```

Download and import GTFS (Hamilton transit data)
```{r eval=FALSE}
download.file(url = "https://transitfeeds.com/p/hamilton-street-railway/31/latest/download", destfile = file.path(r5_path, "HSR_transit.zip"), mode = "wb")
```

Download and import GTFS (Hamilton transit data)
```{r eval=FALSE}
download.file(url = "https://transitfeeds.com/p/burlington-transit/294/latest/download", destfile = file.path(r5_path, "Burlington_transit.zip"), mode = "wb")
```

Build the graph:
```{r build-graph, include = FALSE}
r5_hamilton_cma <- setup_r5(data_path = r5_path, verbose = FALSE)
```

Prepare Input Data for `r5r`. The origins are the coordinates of the parcels and the destinations the coordinates of the pharmacies:
```{r prepare-inputs}
# save origins in format expected by R5R (id, lon, lat)
origins_i <- data.frame(ID = nearest_network_point_2016$ID, 
              nearest_network_point_2016 %>%
                st_transform(crs = 4326) %>%
                st_coordinates()) %>%
    rename(lon = X, lat = Y, id = ID) %>%
    dplyr::select(id, lon, lat)

# now vaccination sites
destinations_j <- data.frame(ID = pharmacy_locations$ID, 
              pharmacy_locations %>%
                st_transform(crs = 4326) %>%
                st_coordinates()) %>%
    rename(lon = X, lat = Y, id = ID) %>%
    dplyr::select(id, lon, lat)
```

Calculate OD Matrix for car:
```{r calculate car od matrix}
# set up batching according to how many origin rows to process at one time
chunksize = 2000 # larger chunks for walking will give enough origins in each chunk to allow multiprocessing to spin up with R5R
num_chunks = ceiling(nrow(origins_i)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(origins_i,
                          outdir = "./df/origins_i",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){ 
  origins_i_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5_hamilton_cma,
                          origins = origins_i_chunk,
                          destinations = destinations_j,
                          mode = c("CAR"),
                          mode_egress = c("WALK"),
                          departure_datetime = as.POSIXct(strptime("2021-04-05 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
                          max_walk_dist = 10000, # metres
                          max_trip_duration = 180)

  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "./df/output_ttm_car",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
ttm_car_time <- end.time - start.time
ttm_car_time
```

Calculate OD Matrix for transit:
```{r calculate transit od matrix}
# set up batching according to how many origin rows to process at one time
chunksize = 2000 # larger chunks for walking will give enough origins in each chunk to allow multiprocessing to spin up with R5R
num_chunks = ceiling(nrow(origins_i)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(origins_i,
                          outdir = "./df/origins_i",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  origins_i_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5_hamilton_cma,
                          origins = origins_i_chunk,
                          destinations = destinations_j,
                          mode = c("TRANSIT"),
                          mode_egress = c("WALK"),
                          departure_datetime = as.POSIXct(strptime("2021-04-05 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
                          max_walk_dist = 10000, # metres
                          max_trip_duration = 180)

  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "./df/output_ttm_transit",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
ttm_transit_time <- end.time - start.time
ttm_transit_time
```

Calculate OD Matrix for walking:
```{r calculate walk od matrix}
# set up batching according to how many origin rows to process at one time
chunksize = 2000 # larger chunks for walking will give enough origins in each chunk to allow multiprocessing to spin up with R5R
num_chunks = ceiling(nrow(origins_i)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(origins_i,
                          outdir = "./df/origins_i",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){ 
  origins_i_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5_hamilton_cma,
                          origins = origins_i_chunk,
                          destinations = destinations_j,
                          mode = c("WALK"),
                          departure_datetime = as.POSIXct(strptime("2021-04-05 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
                          max_walk_dist = 10000, # metres
                          max_trip_duration = 180)

  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "./df/output_ttm_walk",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
ttm_walk_time <- end.time - start.time
ttm_walk_time
```

## Extract travel time matrix

```{r load od matrix for transit, include=FALSE}
# connect to the walking travel time matrix disk frame
ttm_transit.disk.frame <- disk.frame("./df/output_ttm_transit")
#opportunities_j <- as.data.frame(nyc_cb_point) %>% select(GEOID10, total_emp) %>% rename(toId = GEOID10, o_j = total_emp)
```

Convert disk.frame to data frame:
```{r}
ttm_transit <- as.data.frame(ttm_transit.disk.frame) %>%
  transmute(UID = as.numeric(fromId), OBJECTID = as.numeric(toId), travel_time)
```

```{r load od matrix for walk, include=FALSE}
# connect to the walking travel time matrix disk frame
ttm_walk.disk.frame <- disk.frame("./df/output_ttm_walk")
#opportunities_j <- as.data.frame(nyc_cb_point) %>% select(GEOID10, total_emp) %>% rename(toId = GEOID10, o_j = total_emp)
```

Convert disk.frame to data frame:
```{r}
ttm_walk <- as.data.frame(ttm_walk.disk.frame) %>%
  transmute(UID = as.numeric(fromId), OBJECTID = as.numeric(toId), travel_time)
```

```{r load od matrix for car, include=FALSE}
# connect to the walking travel time matrix disk frame
ttm_car.disk.frame <- disk.frame("./df/output_ttm_car")
#opportunities_j <- as.data.frame(nyc_cb_point) %>% select(GEOID10, total_emp) %>% rename(toId = GEOID10, o_j = total_emp)
```

Convert disk.frame to data frame:
```{r}
ttm_car <- as.data.frame(ttm_car.disk.frame) %>%
  transmute(UID = as.numeric(fromId), OBJECTID = as.numeric(toId), travel_time)
```

## Save data to disk

Clean `nearest_network_point_2016`:
```{r}
dwelling_network_points_2016 <- nearest_network_point_2016 %>%
  mutate(Dwellings = SINGLE_FAM + SEMI_DET + ROW_THOUSE + APARTMENT) %>%
  dplyr::select(-c(ROLL_NUM, LUC1, LUC1_DESC, SINGLE_FAM, SEMI_DET, ROW_THOUSE, APARTMENT))
```

Save data files:
```{r}
save(data_da_2016, file = "output-data-files/data_da_2016.RData", compress = "xz")
save(dwelling_network_points_2016, file = "output-data-files/dwelling_network_points_2016.RData", compress = "xz")
save(hamilton_cma, file = "output-data-files/hamilton_cma.RData", compress = "xz")
save(modes_55to69, file = "output-data-files/modes_55to69.RData", compress = "xz")
save(modes_70plus, file = "output-data-files/modes_70plus.RData", compress = "xz")
save(pharmacy_locations, file = "output-data-files/pharmacy_locations.RData", compress = "xz")
save(ttm_car, file = "output-data-files/ttm_car.RData", compress = "xz")
save(ttm_transit, file = "output-data-files/ttm_transit.RData", compress = "xz")
save(ttm_walk, file = "output-data-files/ttm_walk.RData", compress = "xz")
save(urban_types, file = "output-data-files/urban_types.RData", compress = "xz")
```

Copy files to folder `data` of the package.

### 2021-04-16 00-Data-Processing/01-Data-Analysis.Rmd
Initial commit with first steps of data preparation

---
title: "01-Data-Analysis"
output: html_notebook
---

Analysis of accessibility to pharmacies in Hamilton.

Load packages:
```{r load-packages, include=FALSE}
library(sf)
library(tidyverse)
library(vaccHamilton)
```

Invoke data:
```{r}
data("data_da_2016")
data("dwelling_network_points_2016")
data("hamilton_cma")
data("modes_55to69")
data("pharmacy_locations")
data("ttm_car")
data("ttm_transit")
data("ttm_walk")
```

## Prepare data

Calculate number of people age 55+ per DA:
```{r}
data_da_2016 <- data_da_2016 %>%
  filter(`Region Name` == "Hamilton") %>%
  mutate(Population_55to69 = (`v_CA16_208: 55 to 59 years` + 
                  `v_CA16_226: 60 to 64 years` + 
                  `v_CA16_247: 65 to 69 years`),
         Population_70plus = (`v_CA16_265: 70 to 74 years` +
                  `v_CA16_283: 75 to 79 years` +
                  `v_CA16_301: 80 to 84 years` + 
                  `v_CA16_319: 85 years and over`))
```

Select pharmacy locations:
```{r}
pharmacy_locations <- pharmacy_locations %>%
  slice(1:23)
```

Also filter selected pharmacies from travel time tables:
```{r}
ttm_car <- ttm_car %>%
  dplyr::filter(OBJECTID <= 23)

ttm_transit <- ttm_transit %>%
  dplyr::filter(OBJECTID <= 23)

ttm_walk <- ttm_walk %>%
  dplyr::filter(OBJECTID <= 23)
```

Clip `data_da_2016`:
```{r}
data_da_2016 <- data_da_2016 %>%
  st_intersection(hamilton_cma)
```

Find number of residential units in each DA:
```{r}
residential_units <- dwelling_network_points_2016 %>%
  st_drop_geometry() %>%
  group_by(DAUID) %>%
  summarize(residential_units = sum(Dwellings))
```

Join number of people age 55+ per DA to dwellings and total number of residential units:
```{r}
dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
  left_join(data_da_2016 %>%
              st_drop_geometry() %>%
              dplyr::transmute(DAUID = GeoUID, 
                               Population_55to69,
                               Population_70plus),
            by = "DAUID") %>%
  left_join(residential_units,
            by = "DAUID")
```

Calculate population per point on the network. This is the population in the DA divided by the total number of residential units. This gives the average number of people per residential unit. Then, multiply by the number of dwellings (residential units) at the point:
```{r}
dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
  mutate(Population_55to69 = (Population_55to69/residential_units) * Dwellings,
         Population_70plus = (Population_70plus/residential_units) * Dwellings)
```

Join the population data to the travel time matrices:
```{r}
ttm_car <- ttm_car %>%
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            Population_55to69,
                            Population_70plus,
                            DAUID, 
                            TAZUID),
            by = c("UID" = "ID"))

ttm_transit <- ttm_transit %>%
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            Population_55to69,
                            Population_70plus,
                            DAUID, 
                            TAZUID),
            by = c("UID" = "ID"))

ttm_walk <- ttm_walk %>%
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            Population_55to69,
                            Population_70plus,
                            DAUID, 
                            TAZUID),
            by = c("UID" = "ID"))
```

Join the pharmacy data to the travel time matrices:
```{r}
ttm_car <- ttm_car %>%
  left_join(pharmacy_locations %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            Type),
            by = c("OBJECTID" = "ID"))

ttm_transit <- ttm_transit %>%
  left_join(pharmacy_locations %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            Type),
            by = c("OBJECTID" = "ID"))

ttm_walk <- ttm_walk %>%
  left_join(pharmacy_locations %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            Type),
            by = c("OBJECTID" = "ID"))
```

Drop NAs:
```{r}
ttm_car <- ttm_car %>%
  drop_na(TAZUID,
          Population_55to69)

ttm_transit <- ttm_transit %>%
  drop_na(TAZUID,
          Population_55to69)

ttm_walk <- ttm_walk %>%
  drop_na(TAZUID,
          Population_55to69)
```

Calculate proportion of trips by mode per TAZ:
```{r}
modes <- modes_55to69 %>%
  mutate(p_car = Driver/(Driver + Transit + Walk),
         p_transit = Transit/(Driver + Transit + Walk),
         p_walk = Walk/(Driver + Transit + Walk))
```

Assume that TAZ with NAs have a proportion of car trips of one:
```{r}
modes <- modes %>%
  dplyr::mutate(p_car = replace_na(p_car, 1),
                p_transit = replace_na(p_transit, 0),
                p_walk = replace_na(p_walk, 0))
```

Join proportion of trips to table with person-hours
```{r}
ttm_car <- ttm_car %>%
  left_join(modes %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        p_car,
                        p_transit,
                        p_walk),
            by = "TAZUID")

ttm_transit <- ttm_transit %>%
  left_join(modes %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        p_car,
                        p_transit,
                        p_walk),
            by = "TAZUID")

ttm_walk <- ttm_walk %>%
  left_join(modes %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        p_car,
                        p_transit,
                        p_walk),
            by = "TAZUID")
```

Calculate quintiles of median household income in Hamilton:
```{r}
mhi_q <- data_da_2016 %>%
  dplyr::filter(`Region Name` == "Hamilton") %>%
  pull(`v_CA16_2397: Median total income of households in 2015 ($)`) %>%
  quantile(c(0, 0.2, 0.4, 0.6, 0.8, 1), 
           na.rm = TRUE)
```

Label DAs with income quintiles:
```{r}
data_da_2016 <- data_da_2016 %>%
  rename(Median_household_income = `v_CA16_2397: Median total income of households in 2015 ($)`) %>%
  mutate(income_quintile = case_when(Median_household_income < mhi_q[2] ~ "Bottom 20%",
                                     Median_household_income >= mhi_q[2] & Median_household_income < mhi_q[3] ~ "Fourth 20%",
                                     Median_household_income >= mhi_q[3] & Median_household_income < mhi_q[4] ~ "Third 20%",
                                     Median_household_income >= mhi_q[4] & Median_household_income < mhi_q[5] ~ "Second 20%",
                                     Median_household_income >= mhi_q[5] ~ "Top 20%"),
         income_quintile = factor(income_quintile,
                                  levels = c("Top 20%",
                                             "Second 20%",
                                             "Third 20%",
                                             "Fourth 20%",
                                             "Bottom 20%"),
                                  ordered = TRUE))
```

## Exploratory data analysis

Summary of population:
```{r}
sum(data_da_2016$Population_55to69, na.rm = TRUE)
sum(data_da_2016$Population_70plus, na.rm = TRUE)
```

Check the total population:
```{r}
sum(dwelling_network_points_2016$Population_55to69, na.rm = TRUE)
sum(dwelling_network_points_2016$Population_70plus, na.rm = TRUE)
```

The difference in population is because the residential units considered do not completely cover the Hamilton CMA:
```{r}
ggplot() +
  geom_sf(data = hamilton_cma) + 
  geom_sf(data = dwelling_network_points_2016) + 
  geom_sf(data = pharmacy_locations, 
          aes(color = Type,
              shape = Type), 
          size = 3)
```

Plot population age 55 to 69:
```{r}
ggplot() +
  geom_sf(data = data_da_2016 %>%
            filter(`Region Name` == "Hamilton"),
          aes(fill = Population_55to69),
          color = NA) +
  geom_sf(data = hamilton_cma,
          fill = NA) +
  scale_fill_distiller(palette = "OrRd", 
                       direction = 1)
```

Plot population age 70 plus:
```{r}
ggplot() +
  geom_sf(data = data_da_2016 %>%
            filter(`Region Name` == "Hamilton"),
          aes(fill = Population_70plus),
          color = NA) +
  geom_sf(data = hamilton_cma,
          fill = NA) +
  scale_fill_distiller(palette = "OrRd", 
                       direction = 1)
```

Summary of travel time tables:
```{r}
summary(ttm_car)
summary(ttm_transit)
summary(ttm_walk)
```

Map proportion of trips by car:
```{r}
ggplot() +
  geom_sf(data = modes,
          aes(fill = p_car),
          color = NA) +
  geom_sf(data = hamilton_cma,
          fill = NA) +
  scale_fill_distiller(palette = "OrRd", 
                       direction = 1)
```

Map proportion of trips by transit:
```{r}
ggplot() +
  geom_sf(data = modes,
          aes(fill = p_transit),
          color = NA) +
  geom_sf(data = hamilton_cma,
          fill = NA) +
  scale_fill_distiller(palette = "OrRd", 
                       direction = 1)
```

Map proportion of walking trips:
```{r}
ggplot() +
  geom_sf(data = modes,
          aes(fill = p_walk),
          color = NA) +
  geom_sf(data = hamilton_cma,
          fill = NA) +
  scale_fill_distiller(palette = "OrRd", 
                       direction = 1)
```

## Analysis with pilot locations only

Retrieve the minimum travel time from a point to a pharmacy by each mode for people aged 55 to 69 WITH PILOT LOCATIONS ONLY:
```{r}
min_ttm_car <- ttm_car %>%
  dplyr::filter(Type == "Pilot") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            Population = first(Population_55to69),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop")

min_ttm_transit <- ttm_transit %>%
  dplyr::filter(Type == "Pilot") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            Population = first(Population_55to69),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop")

min_ttm_walk <- ttm_walk %>%
  dplyr::filter(Type == "Pilot") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            Population = first(Population_55to69),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop")
```

Calculate weighted travel time:
```{r}
min_ttm_car <- min_ttm_car %>%
  mutate(weighted_travel_time = travel_time * p_car)

min_ttm_transit <- min_ttm_transit %>%
  mutate(weighted_travel_time = travel_time * p_transit)

min_ttm_walk <- min_ttm_walk %>%
  mutate(weighted_travel_time = travel_time * p_walk)
```

Bind the three tables:
```{r}
min_ttm <- rbind(min_ttm_car,
                 min_ttm_transit,
                 min_ttm_walk)
```

Aggregate weighted travel times by mode:
```{r}
min_ttm <- min_ttm %>%
  group_by(UID) %>%
  summarize(DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            Population = first(Population),
            weighted_travel_time = sum(weighted_travel_time))
```

Calculate person-hours of travel:
```{r}
min_ttm <- min_ttm %>%
  mutate(weighted_person_hours = Population * weighted_travel_time/60)
```

Summarize by TAZ and join geometry:
```{r}
min_ttm_taz <- min_ttm %>%
  group_by(TAZUID) %>%
  summarize(weighted_travel_time = mean(weighted_travel_time),
            weighted_person_hours = sum(weighted_person_hours)) %>%
  left_join(modes %>%
              dplyr::select(TAZUID),
            by = "TAZUID") %>%
  st_as_sf()
```

Map weighted travel time:
```{r}
ggplot() +
  geom_sf(data = min_ttm_taz,
          aes(fill = weighted_travel_time),
          color = NA) +
  geom_sf(data = hamilton_cma,
          fill = NA) +
  scale_fill_distiller(palette = "OrRd", 
                       direction = 1)
```

Map weighted person-hours of travel:
```{r}
ggplot() +
  geom_sf(data = min_ttm_taz,
          aes(fill = weighted_person_hours),
          color = NA) +
  geom_sf(data = hamilton_cma,
          fill = NA) +
  scale_fill_distiller(palette = "OrRd", 
                       direction = 1)
```

Join median household information to table:
```{r}
min_ttm <- min_ttm %>%
  left_join(data_da_2016 %>%
              st_drop_geometry() %>%
              dplyr::transmute(DAUID = GeoUID,
                               income_quintile),
            by = "DAUID")
```

Distribution of weighted travel time by income:
```{r}
ggplot(data = min_ttm, 
       aes(x = weighted_travel_time, 
           color= income_quintile)) +
  geom_density(size = 0.5, adjust = 2.5) +
  scale_color_manual(values = c("Top 20%" = "blue",
                                "Second 20%" = "steelblue1",
                                "Third 20%" = "gray40", 
                                "Fourth 20%" = "tomato",
                                "Bottom 20%" = "red")) +
  theme_minimal()
```

Total weighted person-hours by income:
```{r}
min_ttm %>%
  group_by(income_quintile) %>%
  summarize(total_weighted_person_hours = sum(weighted_person_hours),
            total_population = sum(Population),
            .groups = "drop")
```

## Analysis with pilot locations + candidates

Retrieve the minimum travel time from a point to a pharmacy by each mode for people aged 55 to 69 WITH PILOT LOCATIONS ONLY:
```{r}
min_ttm_car_full <- ttm_car %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            Population = first(Population_55to69),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop")

min_ttm_transit_full <- ttm_transit %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            Population = first(Population_55to69),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop")

min_ttm_walk_full <- ttm_walk %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            Population = first(Population_55to69),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop")
```

Calculate weighted travel time:
```{r}
min_ttm_car_full <- min_ttm_car_full %>%
  mutate(weighted_travel_time = travel_time * p_car)

min_ttm_transit_full <- min_ttm_transit_full %>%
  mutate(weighted_travel_time = travel_time * p_transit)

min_ttm_walk_full <- min_ttm_walk_full %>%
  mutate(weighted_travel_time = travel_time * p_walk)
```

Bind the three tables:
```{r}
min_ttm_full <- rbind(min_ttm_car_full,
                 min_ttm_transit_full,
                 min_ttm_walk_full)
```

Aggregate weighted travel times by mode:
```{r}
min_ttm_full <- min_ttm_full %>%
  group_by(UID) %>%
  summarize(DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            Population = first(Population),
            weighted_travel_time = sum(weighted_travel_time))
```

Calculate person-hours of travel:
```{r}
min_ttm_full <- min_ttm_full %>%
  mutate(weighted_person_hours = Population * weighted_travel_time/60)
```

Summarize by TAZ and join geometry:
```{r}
min_ttm_taz_full <- min_ttm_full %>%
  group_by(TAZUID) %>%
  summarize(weighted_travel_time = mean(weighted_travel_time),
            weighted_person_hours = sum(weighted_person_hours)) %>%
  left_join(modes %>%
              dplyr::select(TAZUID),
            by = "TAZUID") %>%
  st_as_sf()
```

Map weighted travel time:
```{r}
ggplot() +
  geom_sf(data = min_ttm_taz_full,
          aes(fill = weighted_travel_time),
          color = NA) +
  geom_sf(data = hamilton_cma,
          fill = NA) +
  scale_fill_distiller(palette = "OrRd", 
                       direction = 1)
```

Map weighted person-hours of travel:
```{r}
ggplot() +
  geom_sf(data = min_ttm_taz_full,
          aes(fill = weighted_person_hours),
          color = NA) +
  geom_sf(data = hamilton_cma,
          fill = NA) +
  scale_fill_distiller(palette = "OrRd", 
                       direction = 1)
```

Join median household information to table:
```{r}
min_ttm_full <- min_ttm_full %>%
  left_join(data_da_2016 %>%
              st_drop_geometry() %>%
              dplyr::transmute(DAUID = GeoUID,
                               income_quintile),
            by = "DAUID")
```

Distribution of weighted travel time by income:
```{r}
ggplot(data = min_ttm_full, 
       aes(x = weighted_travel_time, 
           color= income_quintile)) +
  geom_density(size = 0.5, adjust = 2.5) +
  scale_color_manual(values = c("Top 20%" = "blue",
                                "Second 20%" = "steelblue1",
                                "Third 20%" = "gray40", 
                                "Fourth 20%" = "tomato",
                                "Bottom 20%" = "red")) +
  theme_minimal()
```

Total weighted person-hours by income:
```{r}
min_ttm_full %>%
  group_by(income_quintile) %>%
  summarize(total_weighted_person_hours = sum(weighted_person_hours),
            total_population = sum(Population),
            .groups = "drop")
```

### 2021-04-18 00-Data-Processing/00-Data-Processing-for-Food-Banks-Accessibility.Rmd
Verified accessibility calculations are now in file 04-Data-Analysis.Rmd

@@
 In this notebook we create some examples of processing the data for the analysis of accessibility to food banks in Hamilton.
-Particularly, we're interested in spatial interpolation of the population for small areas.
 
@@
 library(cancensus)
-library(disk.frame)
+#library(disk.frame)
 #library(pycno)
 library(readr)
-library(r5r) # the r5r package requires Java Development Kit version 11, which can be downloaded from https://www.oracle.com/java/technologies/javase-jdk11-downloads.html
+#library(r5r) # the r5r package requires Java Development Kit version 11, which can be downloaded from https://www.oracle.com/java/technologies/javase-jdk11-downloads.html
 library(sf)
@@
 
-Define some parameters for `disk.frame` and `r5r`:
-```{r}
-setup_disk.frame()
-options(scipen = 999)
-options(java.parameters = "-Xmx6G")
-options(future.globals.maxSize = Inf)
-```
-
-Data used in this notebook was retrieved from two sources: Open Hamilton and the 2016 Canadian Census.
+Data used in this notebook was retrieved from three sources: Open Hamilton, Transportation Tomorrow Survey, and the 2016 Canadian Census.
 Open Hamilton is an online repository of data from the City of Hamilton.
@@
 data_da_2016 %>%
-  mutate(income_less20k = (`v_CA16_2406: Under $5,000` + 
+  mutate(income_less40k = (`v_CA16_2406: Under $5,000` + 
                              `v_CA16_2407: $5,000 to $9,999` + 
                              `v_CA16_2408: $10,000 to $14,999` +
-                  `v_CA16_2409: $15,000 to $19,999`)) %>%
+                             `v_CA16_2409: $15,000 to $19,999` +
+                             `v_CA16_2410: $20,000 to $24,999` +  
+                             `v_CA16_2411: $25,000 to $29,999` +
+                             `v_CA16_2412: $30,000 to $34,999` +
+                             `v_CA16_2413: $35,000 to $39,999`)) %>%
   tm_shape() +
-  tm_polygons("income_less20k",
+  tm_polygons("income_less40k",
               alpha = 0.8)
@@
 
-I downloaded a cross-tabulation of travel mode by age by traffic analysis zone. Read the shape file with the traffic analysis zone boundaries:
+I downloaded a cross-tabulation of travel mode by household income by traffic analysis zone. Read the shape file with the traffic analysis zone boundaries:
 ```{r}
@@
 
-### Ages 55-69
+### Income less than $40,000
 
@@
 travel_modes_1 <- travel_modes %>%
-  dplyr::filter(age >= 55 & age < 70)
+  dplyr::filter(income >= 1 & income <= 2)
 ```
@@
   dplyr::transmute(TAZUID = gta06, geometry) %>%
-  left_join(travel_modes_1 %>%
-              st_drop_geometry(),
+  left_join(travel_modes_1,
             by = "TAZUID") %>%
@@
 
-### Ages 70+
-
-Filter age:
+Rename table:
 ```{r}
-travel_modes_2 <- travel_modes %>%
-  dplyr::filter(age >= 70)
+modes_less40k <- travel_modes_1
 ```
 
-Summarize trips by mode by traffic analysis zone:
+Ensure that parcel ids are unique:
 ```{r}
-travel_modes_2 <- travel_modes_2 %>%
-  group_by(gta06_hhld, Mode) %>%
-  summarize(trips = sum(total),
-            .groups = "drop")
-```
-
-Pivot wider and rename columns:
-```{r}
-travel_modes_2 <- travel_modes_2 %>%
-  rename(TAZUID = gta06_hhld) %>%
-  pivot_wider(names_from = Mode,
-              values_from = trips,
-              values_fill = 0) %>%
-  rename(Driver = `Auto driver`, 
-         Transit = `Transit (Excluding GO rail)`)
-```
-
-Join geometry:
-```{r}
-travel_modes_2 <- tts_taz %>%
-  dplyr::transmute(TAZUID = gta06, geometry) %>%
-  left_join(travel_modes_2 %>%
-              st_drop_geometry(),
-            by = "TAZUID") %>%
-  st_as_sf()
-```
-
-Plot trips as driver:
-```{r}
-ggplot() +
-  geom_sf(data = travel_modes_2,
-          aes(fill = Driver))
-```
-
-Plot trips transit:
-```{r}
-ggplot() +
-  geom_sf(data = travel_modes_2,
-          aes(fill = Transit))
-```
-
-Plot walking trips:
-```{r}
-ggplot() +
-  geom_sf(data = travel_modes_2,
-          aes(fill = Walk))
-```
-
-Rename tables:
-```{r}
-modes_55to69 <- travel_modes_1
-modes_70plus <- travel_modes_2
-```
-
-## Join TAZ identifiers to parcel table
-
-This will be a spatial join:
-```{r}
-nearest_network_point_2016 <- nearest_network_point_2016 %>%
-  st_join(tts_taz %>%
-            transmute(TAZUID = gta06))
-```
-
-## Routing for entire system
-
-I used [BBBike](https://download.bbbike.org/osm/bbbike/) to extract OSM data for Hamilton. The name of the file is `planet_-80.279,43.048_-79.318,43.459.osm.pbf`. Copy to folder `r5_graph`.
-
-Set Up R5 Routing. First define the path to where the graph is located:
-```{r set up r5 path, include=FALSE}
-r5_path <- file.path("./r5_graph")
-```
-
-Download and import GTFS (Hamilton transit data)
-```{r eval=FALSE}
-download.file(url = "https://transitfeeds.com/p/hamilton-street-railway/31/latest/download", destfile = file.path(r5_path, "HSR_transit.zip"), mode = "wb")
-```
-
-Download and import GTFS (Hamilton transit data)
-```{r eval=FALSE}
-download.file(url = "https://transitfeeds.com/p/burlington-transit/294/latest/download", destfile = file.path(r5_path, "Burlington_transit.zip"), mode = "wb")
-```
-
-Build the graph:
-```{r build-graph, include = FALSE}
-r5_hamilton_cma <- setup_r5(data_path = r5_path, verbose = FALSE)
-```
-
-Prepare Input Data for `r5r`. The origins are the coordinates of the parcels and the destinations the coordinates of the pharmacies:
-```{r prepare-inputs}
-# save origins in format expected by R5R (id, lon, lat)
-origins_i <- data.frame(ID = nearest_network_point_2016$ID, 
-              nearest_network_point_2016 %>%
-                st_transform(crs = 4326) %>%
-                st_coordinates()) %>%
-    rename(lon = X, lat = Y, id = ID) %>%
-    dplyr::select(id, lon, lat)
-
-# now vaccination sites
-destinations_j <- data.frame(ID = pharmacy_locations$ID, 
-              pharmacy_locations %>%
-                st_transform(crs = 4326) %>%
-                st_coordinates()) %>%
-    rename(lon = X, lat = Y, id = ID) %>%
-    dplyr::select(id, lon, lat)
-```
-
-Calculate OD Matrix for car:
-```{r calculate car od matrix}
-# set up batching according to how many origin rows to process at one time
-chunksize = 2000 # larger chunks for walking will give enough origins in each chunk to allow multiprocessing to spin up with R5R
-num_chunks = ceiling(nrow(origins_i)/chunksize)
-
-# create origin-destination pairs
-origins_chunks <- as.disk.frame(origins_i,
-                          outdir = "./df/origins_i",
-                          nchunks = num_chunks,
-                          overwrite = TRUE)
-
-start.time <- Sys.time()
-pb <- txtProgressBar(0, num_chunks, style = 3)
-
-for (i in 1:num_chunks){ 
-  origins_i_chunk <- get_chunk(origins_chunks, i)
-  ttm_chunk <- travel_time_matrix(r5_hamilton_cma,
-                          origins = origins_i_chunk,
-                          destinations = destinations_j,
-                          mode = c("CAR"),
-                          mode_egress = c("WALK"),
-                          departure_datetime = as.POSIXct(strptime("2021-04-05 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
-                          max_walk_dist = 10000, # metres
-                          max_trip_duration = 180)
-
-  # export output as disk.frame
-  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
-                                            nchunks = 1,
-                                            outdir = "./df/output_ttm_car",
-                                            compress = 50,
-                                            overwrite = TRUE),
-         add_chunk(output_df, ttm_chunk, chunk_id = i))
-  setTxtProgressBar(pb, i)
-}
-end.time <- Sys.time()
-ttm_car_time <- end.time - start.time
-ttm_car_time
-```
-
-Calculate OD Matrix for transit:
-```{r calculate transit od matrix}
-# set up batching according to how many origin rows to process at one time
-chunksize = 2000 # larger chunks for walking will give enough origins in each chunk to allow multiprocessing to spin up with R5R
-num_chunks = ceiling(nrow(origins_i)/chunksize)
-
-# create origin-destination pairs
-origins_chunks <- as.disk.frame(origins_i,
-                          outdir = "./df/origins_i",
-                          nchunks = num_chunks,
-                          overwrite = TRUE)
-
-start.time <- Sys.time()
-pb <- txtProgressBar(0, num_chunks, style = 3)
-
-for (i in 1:num_chunks){
-  origins_i_chunk <- get_chunk(origins_chunks, i)
-  ttm_chunk <- travel_time_matrix(r5_hamilton_cma,
-                          origins = origins_i_chunk,
-                          destinations = destinations_j,
-                          mode = c("TRANSIT"),
-                          mode_egress = c("WALK"),
-                          departure_datetime = as.POSIXct(strptime("2021-04-05 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
-                          max_walk_dist = 10000, # metres
-                          max_trip_duration = 180)
-
-  # export output as disk.frame
-  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
-                                            nchunks = 1,
-                                            outdir = "./df/output_ttm_transit",
-                                            compress = 50,
-                                            overwrite = TRUE),
-         add_chunk(output_df, ttm_chunk, chunk_id = i))
-  setTxtProgressBar(pb, i)
-}
-end.time <- Sys.time()
-ttm_transit_time <- end.time - start.time
-ttm_transit_time
-```
-
-Calculate OD Matrix for walking:
-```{r calculate walk od matrix}
-# set up batching according to how many origin rows to process at one time
-chunksize = 2000 # larger chunks for walking will give enough origins in each chunk to allow multiprocessing to spin up with R5R
-num_chunks = ceiling(nrow(origins_i)/chunksize)
-
-# create origin-destination pairs
-origins_chunks <- as.disk.frame(origins_i,
-                          outdir = "./df/origins_i",
-                          nchunks = num_chunks,
-                          overwrite = TRUE)
-
-start.time <- Sys.time()
-pb <- txtProgressBar(0, num_chunks, style = 3)
-
-for (i in 1:num_chunks){ 
-  origins_i_chunk <- get_chunk(origins_chunks, i)
-  ttm_chunk <- travel_time_matrix(r5_hamilton_cma,
-                          origins = origins_i_chunk,
-                          destinations = destinations_j,
-                          mode = c("WALK"),
-                          departure_datetime = as.POSIXct(strptime("2021-04-05 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
-                          max_walk_dist = 10000, # metres
-                          max_trip_duration = 180)
-
-  # export output as disk.frame
-  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
-                                            nchunks = 1,
-                                            outdir = "./df/output_ttm_walk",
-                                            compress = 50,
-                                            overwrite = TRUE),
-         add_chunk(output_df, ttm_chunk, chunk_id = i))
-  setTxtProgressBar(pb, i)
-}
-end.time <- Sys.time()
-ttm_walk_time <- end.time - start.time
-ttm_walk_time
-```
-
-## Extract travel time matrix
-
-```{r load od matrix for transit, include=FALSE}
-# connect to the walking travel time matrix disk frame
-ttm_transit.disk.frame <- disk.frame("./df/output_ttm_transit")
-#opportunities_j <- as.data.frame(nyc_cb_point) %>% select(GEOID10, total_emp) %>% rename(toId = GEOID10, o_j = total_emp)
-```
-
-Convert disk.frame to data frame:
-```{r}
-ttm_transit <- as.data.frame(ttm_transit.disk.frame) %>%
-  transmute(UID = as.numeric(fromId), OBJECTID = as.numeric(toId), travel_time)
-```
-
-```{r load od matrix for walk, include=FALSE}
-# connect to the walking travel time matrix disk frame
-ttm_walk.disk.frame <- disk.frame("./df/output_ttm_walk")
-#opportunities_j <- as.data.frame(nyc_cb_point) %>% select(GEOID10, total_emp) %>% rename(toId = GEOID10, o_j = total_emp)
-```
-
-Convert disk.frame to data frame:
-```{r}
-ttm_walk <- as.data.frame(ttm_walk.disk.frame) %>%
-  transmute(UID = as.numeric(fromId), OBJECTID = as.numeric(toId), travel_time)
-```
-
-```{r load od matrix for car, include=FALSE}
-# connect to the walking travel time matrix disk frame
-ttm_car.disk.frame <- disk.frame("./df/output_ttm_car")
-#opportunities_j <- as.data.frame(nyc_cb_point) %>% select(GEOID10, total_emp) %>% rename(toId = GEOID10, o_j = total_emp)
-```
-
-Convert disk.frame to data frame:
-```{r}
-ttm_car <- as.data.frame(ttm_car.disk.frame) %>%
-  transmute(UID = as.numeric(fromId), OBJECTID = as.numeric(toId), travel_time)
+dwelling_network_points_2016 <- dwelling_network_points_2016 %>% distinct(ID, .keep_all = TRUE)
 ```
@@
 ## Save data to disk
-
-Clean `nearest_network_point_2016`:
-```{r}
-dwelling_network_points_2016 <- nearest_network_point_2016 %>%
-  mutate(Dwellings = SINGLE_FAM + SEMI_DET + ROW_THOUSE + APARTMENT) %>%
-  dplyr::select(-c(ROLL_NUM, LUC1, LUC1_DESC, SINGLE_FAM, SEMI_DET, ROW_THOUSE, APARTMENT))
-```
 
@@
 save(hamilton_cma, file = "output-data-files/hamilton_cma.RData", compress = "xz")
-save(modes_55to69, file = "output-data-files/modes_55to69.RData", compress = "xz")
-save(modes_70plus, file = "output-data-files/modes_70plus.RData", compress = "xz")
-save(pharmacy_locations, file = "output-data-files/pharmacy_locations.RData", compress = "xz")
-save(ttm_car, file = "output-data-files/ttm_car.RData", compress = "xz")
-save(ttm_transit, file = "output-data-files/ttm_transit.RData", compress = "xz")
-save(ttm_walk, file = "output-data-files/ttm_walk.RData", compress = "xz")
+save(modes_less40k, file = "output-data-files/modes_less40k.RData", compress = "xz")
+save(foodbank_locations, file = "output-data-files/foodbank_locations.RData", compress = "xz")
 save(urban_types, file = "output-data-files/urban_types.RData", compress = "xz")

### 2021-04-18 00-Data-Processing/01-Routing-Car.Rmd
Verified accessibility calculations are now in file 04-Data-Analysis.Rmd

---
title: "01-Routing-Car"
---

## Introduction

In this notebook we do the routing for the analysis of accessibility to food banks in Hamilton. This is required to obtain a travel time matrix. The routing is done using the package `r5r`.

## Preliminaries

Clear environment:
```{r}
rm(list = ls())
```

Load packages used in the notebook:
```{r message = FALSE}
library(cancensus)
library(disk.frame)
#library(pycno)
library(readr)
library(r5r) # the r5r package requires Java Development Kit version 11, which can be downloaded from https://www.oracle.com/java/technologies/javase-jdk11-downloads.html
library(sf)
library(smoothr)
library(tidyverse)
library(tmap)
library(tmaptools)
```

Define some parameters for `disk.frame` and `r5r`:
```{r}
setup_disk.frame()
options(scipen = 999)
options(java.parameters = "-Xmx6G")
options(future.globals.maxSize = Inf)
```

Data used in this notebook was retrieved from two sources: Open Hamilton and the 2016 Canadian Census. Open Hamilton is an online repository of data from the City of Hamilton. The provenance of the datasets is below.

## Provenance of data:

### Hamilton CMA boundaries

Statistics Canada. Load:
```{r}
load("input-data-files/hamilton_cma.RData")
#load("input-data-files/hamilton_da_2016.RData")
```

### Location of food banks

Read file with locations:
```{r}
foodbank_locations <- read_csv("input-data-files/foodbanks-hamilton.csv",
                               col_types = cols(ID = col_double(),
                                    NAME = col_character(),
                                    TYPE = col_character(),
                                    ADDRESS = col_character(),
                                    COVIDClose = col_double(),
                                    latitude = col_double(),
                                    longitude = col_double(),
                                    name = col_character(),
                                    desc = col_character(),
                                    source = col_character(),
                                    precision = col_character()
                               ))
```

COVIDClose is an indicator variable for whether this bank closed during the pandemic. Convert to factor:
```{r}
foodbank_locations <- foodbank_locations %>%
  mutate(COVIDClose = replace_na(COVIDClose, 0),
         COVIDClose = factor(COVIDClose,
                             levels = c(0, 1),
                             labels = c("No", "Yes")))
```

Convert to simple features:
```{r}
foodbank_locations <- foodbank_locations %>%
  st_as_sf(coords = c("longitude", "latitude"), 
           crs = 4326) %>%
  st_transform(crs = 26917)
```

Map location of foodbanks:
```{r}
ggplot() +
  geom_sf(data = hamilton_cma) +
  geom_sf(data = foodbank_locations,
          aes(shape = COVIDClose,
              color = COVIDClose))
```

### `dwelling_network_points_2016.RData`

This is based on parcel-level data, approximated to the nearest point on the road network. For processing details, see:

https://github.com/paezha/Accessibility-to-Schools-Hamilton-Equity

```{r}
load("input-data-files/dwelling_network_points_2016.RData")
```

## Routing for entire system

I used [BBBike](https://download.bbbike.org/osm/bbbike/) to extract OSM data for Hamilton. The name of the file is `planet_-80.279,43.048_-79.318,43.459.osm.pbf`. Copy to folder `r5_graph`.

Set Up R5 Routing. First define the path to where the graph is located:
```{r set up r5 path, include=FALSE}
r5_path <- file.path("./r5_graph")
```

Download and import GTFS (Hamilton transit data)
```{r eval=FALSE}
download.file(url = "https://transitfeeds.com/p/hamilton-street-railway/31/latest/download", destfile = file.path(r5_path, "HSR_transit.zip"), mode = "wb")
```

Download and import GTFS (Hamilton transit data)
```{r eval=FALSE}
download.file(url = "https://transitfeeds.com/p/burlington-transit/294/latest/download", destfile = file.path(r5_path, "Burlington_transit.zip"), mode = "wb")
```

Build the graph:
```{r build-graph, include = FALSE}
r5_hamilton_cma <- setup_r5(data_path = r5_path, verbose = FALSE)
```

Prepare Input Data for `r5r`. The origins are the coordinates of the parcels and the destinations the coordinates of the pharmacies:
```{r prepare-inputs}
# save origins in format expected by R5R (id, lon, lat)
origins_i <- data.frame(ID = dwelling_network_points_2016$ID, 
              dwelling_network_points_2016 %>%
                st_transform(crs = 4326) %>%
                st_coordinates()) %>%
    rename(lon = X, lat = Y, id = ID) %>%
    dplyr::select(id, lon, lat)

# now foodbanks sites
destinations_j <- data.frame(ID = foodbank_locations$ID, 
              foodbank_locations %>%
                st_transform(crs = 4326) %>%
                st_coordinates()) %>%
    rename(lon = X, lat = Y, id = ID) %>%
    dplyr::select(id, lon, lat)
```

Calculate OD Matrix for car:
```{r calculate car od matrix}
# set up batching according to how many origin rows to process at one time
chunksize = 2000 # larger chunks for walking will give enough origins in each chunk to allow multiprocessing to spin up with R5R
num_chunks = ceiling(nrow(origins_i)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(origins_i,
                          outdir = "./df/origins_i",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){ 
  origins_i_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5_hamilton_cma,
                          origins = origins_i_chunk,
                          destinations = destinations_j,
                          mode = c("CAR"),
                          mode_egress = c("WALK"),
                          departure_datetime = as.POSIXct(strptime("2021-04-05 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
                          max_walk_dist = 10000, # metres
                          max_trip_duration = 180)

  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "./df/output_ttm_car",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
ttm_car_time <- end.time - start.time
ttm_car_time
```

## Extract travel time matrix

```{r load od matrix for car, include=FALSE}
# connect to the walking travel time matrix disk frame
ttm_car.disk.frame <- disk.frame("./df/output_ttm_car")
#opportunities_j <- as.data.frame(nyc_cb_point) %>% select(GEOID10, total_emp) %>% rename(toId = GEOID10, o_j = total_emp)
```

Convert disk.frame to data frame:
```{r}
ttm_car <- as.data.frame(ttm_car.disk.frame) %>%
  transmute(UID = as.numeric(fromId), OBJECTID = as.numeric(toId), travel_time)
```

Make sure that the origin destination pairs are distinct in the travel time matrices:
```{r}
ttm_car <- ttm_car %>%
  distinct(UID, 
           OBJECTID,
           .keep_all = TRUE)
```

## Save data to disk

Save data files:
```{r}
#save(foodbank_locations, file = "output-data-files/foodbank_locations.RData", compress = "xz")
save(ttm_car, file = "output-data-files/ttm_car.RData", compress = "xz")
```

Copy files to folder `data` of the package.

### 2021-04-18 00-Data-Processing/02-Routing-Transit.Rmd
Verified accessibility calculations are now in file 04-Data-Analysis.Rmd

--- 2021-04-18 00-Data-Processing/01-Routing-Car.Rmd
@@
 ---
-title: "01-Routing-Car"
+title: "02-Routing-Transit"
 ---
@@
 
-Calculate OD Matrix for car:
-```{r calculate car od matrix}
+Calculate OD Matrix for transit:
+```{r calculate transit od matrix}
 # set up batching according to how many origin rows to process at one time
@@
                           destinations = destinations_j,
-                          mode = c("CAR"),
+                          mode = c("TRANSIT"),
                           mode_egress = c("WALK"),
@@
                                             nchunks = 1,
-                                            outdir = "./df/output_ttm_car",
+                                            outdir = "./df/output_ttm_transit",
                                             compress = 50,
@@
 end.time <- Sys.time()
-ttm_car_time <- end.time - start.time
-ttm_car_time
+ttm_transit_time <- end.time - start.time
+ttm_transit_time
 ```
@@
 
-```{r load od matrix for car, include=FALSE}
+```{r load od matrix for transit, include=FALSE}
 # connect to the walking travel time matrix disk frame
-ttm_car.disk.frame <- disk.frame("./df/output_ttm_car")
+ttm_transit.disk.frame <- disk.frame("./df/output_ttm_transit")
 #opportunities_j <- as.data.frame(nyc_cb_point) %>% select(GEOID10, total_emp) %>% rename(toId = GEOID10, o_j = total_emp)
@@
 ```{r}
-ttm_car <- as.data.frame(ttm_car.disk.frame) %>%
+ttm_transit <- as.data.frame(ttm_transit.disk.frame) %>%
   transmute(UID = as.numeric(fromId), OBJECTID = as.numeric(toId), travel_time)
@@
 ```{r}
-ttm_car <- ttm_car %>%
+ttm_transit <- ttm_transit %>%
   distinct(UID, 
@@
 ```{r}
-#save(foodbank_locations, file = "output-data-files/foodbank_locations.RData", compress = "xz")
-save(ttm_car, file = "output-data-files/ttm_car.RData", compress = "xz")
+save(ttm_transit, file = "output-data-files/ttm_transit.RData", compress = "xz")
 ```

### 2021-04-18 00-Data-Processing/03-Routing-Walk.Rmd
Verified accessibility calculations are now in file 04-Data-Analysis.Rmd

--- 2021-04-18 00-Data-Processing/02-Routing-Transit.Rmd
@@
 ---
-title: "02-Routing-Transit"
+title: "01-Routing-Walk"
 ---
@@
 
-Calculate OD Matrix for transit:
-```{r calculate transit od matrix}
+Calculate OD Matrix for walking:
+```{r calculate walk od matrix}
 # set up batching according to how many origin rows to process at one time
@@
                           destinations = destinations_j,
-                          mode = c("TRANSIT"),
-                          mode_egress = c("WALK"),
+                          mode = c("WALK"),
                           departure_datetime = as.POSIXct(strptime("2021-04-05 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
@@
                                             nchunks = 1,
-                                            outdir = "./df/output_ttm_transit",
+                                            outdir = "./df/output_ttm_walk",
                                             compress = 50,
@@
 end.time <- Sys.time()
-ttm_transit_time <- end.time - start.time
-ttm_transit_time
+ttm_walk_time <- end.time - start.time
+ttm_walk_time
 ```
@@
 
-```{r load od matrix for transit, include=FALSE}
+```{r load od matrix for walk, include=FALSE}
 # connect to the walking travel time matrix disk frame
-ttm_transit.disk.frame <- disk.frame("./df/output_ttm_transit")
+ttm_walk.disk.frame <- disk.frame("./df/output_ttm_walk")
 #opportunities_j <- as.data.frame(nyc_cb_point) %>% select(GEOID10, total_emp) %>% rename(toId = GEOID10, o_j = total_emp)
@@
 ```{r}
-ttm_transit <- as.data.frame(ttm_transit.disk.frame) %>%
+ttm_walk <- as.data.frame(ttm_walk.disk.frame) %>%
   transmute(UID = as.numeric(fromId), OBJECTID = as.numeric(toId), travel_time)
@@
 ```{r}
-ttm_transit <- ttm_transit %>%
+ttm_walk <- ttm_walk %>%
   distinct(UID, 
@@
 ```{r}
-save(ttm_transit, file = "output-data-files/ttm_transit.RData", compress = "xz")
+save(ttm_walk, file = "output-data-files/ttm_walk.RData", compress = "xz")
 ```

### 2021-04-18 00-Data-Processing/04-Data-Analysis.Rmd
Verified accessibility calculations are now in file 04-Data-Analysis.Rmd

---
title: "04-Data-Analysis"
output: html_notebook
---

Analysis of accessibility to foodbanks in Hamilton.

Load packages:
```{r load-packages, include=FALSE}
library(patchwork)
library(sf)
library(tidyverse)
#library(vaccHamilton)
```

Invoke data:
```{r}
load("output-data-files/data_da_2016.RData")
load("output-data-files/dwelling_network_points_2016.RData")
load("output-data-files/hamilton_cma.RData")
load("output-data-files/foodbank_locations.RData")
load("output-data-files/modes_less40k.RData")
load("output-data-files/ttm_car.RData")
load("output-data-files/ttm_transit.RData")
load("output-data-files/ttm_walk.RData")
#data("data_da_2016")
#data("dwelling_network_points_2016")
#data("hamilton_cma")
#data("modes_less40k")
#data("foodbank_locations")
#data("ttm_car")
#data("ttm_transit")
#data("ttm_walk")
```

## Quick check

Distribution of the number of foodbanks that can be reached by each mode at the maximum travel time used in the calculation of the time matrices (180 min or 10,000 m).

Car:
```{r}
ttm_car %>% group_by(UID) %>% summarize(fb_per_parcel = n()) %>% summary()
```

All parcels can reach every foodbank within 180 min.

Transit:
```{r}
ttm_transit %>% group_by(UID) %>% summarize(fb_per_parcel = n()) %>% summary()
```

Half of all parcels can reach 63 foodbanks or more in 180 min or less.

Walking:
```{r}
ttm_walk %>% group_by(UID) %>% summarize(fb_per_parcel = n()) %>% summary()
```

Half of all parcels can reach 50 foodbanks or more in 180 min or less.

## Prepare data

Calculate number of households with incomes < \$40k per DA and proportion of households with incomes < \$40k per DA:
```{r}
data_da_2016 <- data_da_2016 %>%
  filter(`Region Name` == "Hamilton") %>%
  mutate(income_less40k = (`v_CA16_2406: Under $5,000` + 
                             `v_CA16_2407: $5,000 to $9,999` + 
                             `v_CA16_2408: $10,000 to $14,999` +
                             `v_CA16_2409: $15,000 to $19,999` +
                             `v_CA16_2410: $20,000 to $24,999` +  
                             `v_CA16_2411: $25,000 to $29,999` +
                             `v_CA16_2412: $30,000 to $34,999` +
                             `v_CA16_2413: $35,000 to $39,999`),
         p_inc_less40k = income_less40k/Households)
```

Clip `data_da_2016`:
```{r}
data_da_2016 <- data_da_2016 %>%
  st_intersection(hamilton_cma)
```

Join number of low income households per DA to dwellings and total number of residential units:
```{r}
dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
  left_join(data_da_2016 %>%
              st_drop_geometry() %>%
              dplyr::transmute(DAUID = GeoUID, 
                               p_inc_less40k),
            by = "DAUID")
```

Calculate probability that a point on the network will have low income households. This is the proportion of low income households in the DA multiplied by the total number of residential units a the point:
```{r}
dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
  mutate(income_less40k = p_inc_less40k * Dwellings)
```

Join the household data to the travel time matrices:
```{r}
ttm_car <- ttm_car %>%
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            income_less40k,
                            DAUID, 
                            TAZUID),
            by = c("UID" = "ID"))

ttm_transit <- ttm_transit %>%
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            income_less40k,
                            DAUID, 
                            TAZUID),
            by = c("UID" = "ID"))

ttm_walk <- ttm_walk %>%
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            income_less40k,
                            DAUID, 
                            TAZUID),
            by = c("UID" = "ID"))
```

Join the pharmacy data to the travel time matrices:
```{r}
ttm_car <- ttm_car %>%
  left_join(foodbank_locations %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            COVIDClose),
            by = c("OBJECTID" = "ID"))

ttm_transit <- ttm_transit %>%
  left_join(foodbank_locations %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            COVIDClose),
            by = c("OBJECTID" = "ID"))

ttm_walk <- ttm_walk %>%
  left_join(foodbank_locations %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            COVIDClose),
            by = c("OBJECTID" = "ID"))
```

Drop NAs:
```{r}
ttm_car <- ttm_car %>%
  drop_na(TAZUID,
          income_less40k)

ttm_transit <- ttm_transit %>%
  drop_na(TAZUID,
          income_less40k)

ttm_walk <- ttm_walk %>%
  drop_na(TAZUID,
          income_less40k)
```

Calculate proportion of trips by mode per TAZ:
```{r}
modes <- modes_less40k %>%
  mutate(p_car = Driver/(Driver + Transit + Walk),
         p_transit = Transit/(Driver + Transit + Walk),
         p_walk = Walk/(Driver + Transit + Walk))
```

Assume that TAZ with NAs have a proportion of car trips of one:
```{r}
modes <- modes %>%
  dplyr::mutate(p_car = replace_na(p_car, 1),
                p_transit = replace_na(p_transit, 0),
                p_walk = replace_na(p_walk, 0))
```

Join tables:
```{r}
ttm <- ttm_car %>%
  rename(travel_time_car = travel_time) %>%
  full_join(ttm_transit %>%
              transmute(UID,
                        OBJECTID,
                        travel_time_transit = travel_time),
            by = c("UID", "OBJECTID")) %>%
  full_join(ttm_walk %>%
              transmute(UID,
                        OBJECTID,
                        travel_time_walk = travel_time),
            by = c("UID", "OBJECTID"))
```

Join proportion of trips to travel time table:
```{r}
ttm <- ttm %>%
  left_join(modes %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        p_car,
                        p_transit,
                        p_walk),
            by = "TAZUID")
```

Check the NAs in travel time table:
```{r}
ttm %>% 
  filter(is.na(travel_time_transit)) %>%
  select(travel_time_transit,
         p_transit) %>%
  summary()
```

What does this mean? Most DAs missing travel time by transit do not report trips by transit. In some DAs there are trips by transit (p_transit != 0), but no foodbank can be reached by transit (travel_time_transit is NA). How many of these cases?
```{r}
ttm %>% 
  filter(is.na(travel_time_transit) & p_transit > 0) %>%
  select(travel_time_transit,
         p_transit) %>%
  summary()
```

It should be fine to replace the NAs with zeros, which multiplied by p_transit will return a zero (i.e., it is not possible to reach a foodbank by transit).
```{r}
ttm <- ttm %>% 
  mutate(travel_time_transit = replace_na(travel_time_transit, 0))
```

Check the NAs in travel time by walk:
```{r}
ttm %>% 
  filter(is.na(travel_time_walk)) %>%
  select(travel_time_walk,
         p_walk) %>%
  summary()
```

What does this mean? Most DAs missing travel time by walking do not report trips by walkin In some DAs there are trips by walking (p_walk != 0), but no foodbank can be reached by walk (travel_time_walk is NA). How many of these cases?
```{r}
ttm %>% 
  filter(is.na(travel_time_walk) & p_walk > 0) %>%
  select(travel_time_walk,
         p_walk) %>%
  summary()
```

Like with transit, it should be fine to replace the NAs with zeros, which multiplied by p_walk will return a zero (i.e., it is not possible to reach a foodbank by walking).
```{r}
ttm <- ttm %>% 
  mutate(travel_time_walk = replace_na(travel_time_walk, 0))
```

Summary of the table after these operations:
```{r}
summary(ttm)
```

Calculate expected travel times by mode. These are the travel time by the mode multiplied by the probability of using the mode (which is proxied by the proportion of trips by mode in the DA). The total expected travel time is the sum of the expected travel times by mode:
```{r}
ttm <- ttm %>%
  mutate(e_travel_time_car = travel_time_car * p_car,
         e_travel_time_transit = travel_time_transit * p_transit,
         e_travel_time_walk = travel_time_walk * p_walk,
         e_travel_time = e_travel_time_car + e_travel_time_transit + e_travel_time_walk)
```

What is the mean expected travel time for each parcel over all the foodbanks?
```{r}
e_travel_time_parcel <- ttm %>% 
  group_by(UID) %>%
  summarize(DAUID = first(DAUID),
            e_travel_time = mean(e_travel_time))

summary(e_travel_time_parcel)
```

Join geometry for plotting:
```{r}
e_travel_time_parcel <- e_travel_time_parcel %>%
  left_join(dwelling_network_points_2016 %>%
              select(ID, geometry),
            by = c("UID" = "ID")) %>%
  st_as_sf()
```

Plot:
```{r}
ggplot() +
  geom_sf(data = e_travel_time_parcel,
          aes(color = e_travel_time),
          alpha = 0.3) + 
  scale_color_distiller(palette = "OrRd", direction = 1)
```

What is the mean expected travel time for each parcel over all the foodbanks that did NOT close during COVID?
```{r}
e_travel_time_parcel_2 <- ttm %>% 
  filter(COVIDClose == "No") %>%
  group_by(UID) %>%
  summarize(DAUID = first(DAUID),
            e_travel_time = mean(e_travel_time))

summary(e_travel_time_parcel_2)
```

Join geometry for plotting:
```{r}
e_travel_time_parcel_2 <- e_travel_time_parcel_2 %>%
  left_join(dwelling_network_points_2016 %>%
              select(ID, geometry),
            by = c("UID" = "ID")) %>%
  st_as_sf()
```

Plot:
```{r}
ggplot() +
  geom_sf(data = e_travel_time_parcel_2,
          aes(color = e_travel_time),
          alpha = 0.3) + 
  scale_color_distiller(palette = "OrRd", direction = 1)
```

Notice how expected travel time can _decrease_ with the closures! Foodbanks on average are a little bit closer, but there are _fewer_ of them, which will tend to increase the demand: 
```{r}
summary(e_travel_time_parcel)
summary(e_travel_time_parcel_2)
```

## Accessibility

Define a function for calculating accessibility using a binary impedance function:
```{r}
#' Compute balanced floating catchment area accessibility
#'
#' @param ttm A travel time matrix with a column of origin IDs, a column of destination IDs, travel times, population at the origin, and supply at the destination.
#' @param threshold A threshold for the binary impedance function
#' @return The level of service \code{los} and accessibility \code{accessibility}.
#' @export
#' @examples
#' #add(1, 1)
#' #add(10, 1)

b2sfca <- function(ttm = ttm, threshold = threshold){
  # calculate the impedance
  ttm <- ttm %>%
    dplyr::mutate(impedance_binary = ifelse(travel_time <= threshold, 1, 0))

  # Calculate the sum of the impedance by population for the balancing factors. In the binary case, this should be the same as the number of SoBi hubs that each interpolated population unit can reach
  sum_b1 <- ttm %>%
    dplyr::group_by(UID) %>%
    dplyr::summarize(sum_b1 = sum(impedance_binary),
              .groups = "drop")

  # Calculate the sum of the impedance by SoBi hubs for the balancing factors. In the binary case, this should be the same as the number of interpolated population cells that each SoBi hub serves
  sum_b2 <- ttm %>%
    dplyr::group_by(OBJECTID) %>%
    dplyr::summarize(sum_b2 = sum(impedance_binary))

  # Join the sum of the impedance to the table to calculate the balanced impedance values:
  ttm <- ttm %>%
    dplyr::left_join(sum_b1, by = "UID") %>%
    dplyr::left_join(sum_b2, by = "OBJECTID")

  # **Important:** notice that some values of `sum_b1` are zeros! This means that some population units cannot reach a SoBi hub in at most 24 min. Since the balancing impedance is calculated by diving the impedance by `sum_b1`, this will lead to NaNs (divisions by zero). To avoid issues we will remove any population units that reach zero hubs:
  ttm <- ttm %>%
    dplyr::filter(sum_b1 > 0,
                  sum_b2 > 0)

  # Calculate balanced impedance values:
  ttm <- ttm %>%
    dplyr::mutate(balanced_impedance_1 = impedance_binary/sum_b1,
                  balanced_impedance_2 = impedance_binary/sum_b2)

  # The level of service of each hub is the number of bicycle racks at the hub, divided by the population that they serve:
  los <- ttm %>%
    dplyr::group_by(OBJECTID) %>%
    dplyr::summarize(los = dplyr::first(supply) / sum((population * balanced_impedance_1)),
              .groups = "drop")

  # Join the level of service to the table:
  ttm <- ttm %>%
    dplyr::left_join(los, by = "OBJECTID")

  # To calculate accessibility, the level of service of each hub needs to be distributed proportionally to the population units.
  accessibility <- ttm %>%
    dplyr::group_by(UID) %>%
    dplyr::summarize(accessibility = sum(los * balanced_impedance_2),
              .groups = "drop")

  output = list(los = los,
                accessibility = accessibility)
  return(output)
}
```

Calculate accessibility with a threshold of 15 min:
```{r}
acc_15min <- ttm %>% 
  transmute(UID, OBJECTID, travel_time = e_travel_time, population = income_less40k, supply = 1) %>%
  b2sfca(threshold = 15)
```

Check that los and accessibility balance:
```{r}
sum(acc_15min$los$los)
sum(acc_15min$accessibility$accessibility)
```

Sensitivity analysis:
```{r accessibility-sensibility}
# Testing different thresholds to check how accessibility changes
threshold_seq <- seq(from = 2, to = 30, by = 2)

accessibility_test <- data.frame(threshold = threshold_seq, 
                                 accessibility = numeric(length(threshold_seq)))
for(i in 1:length(threshold_seq)){
  results_test <-  ttm %>% 
  transmute(UID, OBJECTID, travel_time = e_travel_time, population = income_less40k, supply = 1) %>%
    b2sfca(threshold = i )
  accessibility_test$accessibility[i] <- sum(results_test$accessibility$accessibility)
}

ggplot(accessibility_test, 
       aes(x = threshold, y = accessibility)) +
  geom_line()
```

30-min accessibility pre-covid:
```{r 30-min-pre-covid, include=FALSE}
results_30min_pre <- ttm %>%
  #dplyr::filter(COVIDClose == "No") %>%
  transmute(UID, 
            OBJECTID, 
            travel_time = e_travel_time, 
            population = income_less40k, 
            supply = 1) %>%
  b2sfca(threshold = 30)
```

30-min accessibility during covid:
```{r 30-min-during-covid, include=FALSE}
results_30min_covid <- ttm %>%
  dplyr::filter(COVIDClose == "No") %>%
  transmute(UID, 
            OBJECTID, 
            travel_time = e_travel_time, 
            population = income_less40k, 
            supply = 1) %>%
  b2sfca(threshold = 30)
```

Regional accessibility pre-covid:
```{r, 30-min-sum-pre}
sum(results_30min_pre$los$los)
sum(results_30min_pre$accessibility$accessibility)
```

Regional accessibility during covid:
```{r 30-min-sum-covid}
sum(results_30min_covid$los$los)
sum(results_30min_covid$accessibility$accessibility)
```

Bind results:
```{r, bind-all-results}
los <- rbind(data.frame(results_30min_pre$los, 
                        timing = "Pre-covid", 
                        threshold = "30 min"),
             data.frame(results_30min_covid$los, 
                        timing = "During covid", 
                        threshold = "30 min")) %>%
  mutate(timing = factor(timing, 
                         levels = c("Pre-covid", 
                                    "During covid"),
                         ordered = TRUE))

accessibility <- rbind(data.frame(results_30min_pre$accessibility, 
                                  timing = "Pre-covid", 
                                  threshold = "30 min"),
                       data.frame(results_30min_covid$accessibility, 
                                  timing = "During covid", 
                                  threshold = "30 min")) %>%
  mutate(timing = factor(timing, 
                         levels = c("Pre-covid", 
                                    "During covid"),
                         ordered = TRUE))
```

Join geometry:
```{r, add-geometry-to-results, include=FALSE}
los <- los %>%
  left_join(foodbank_locations %>%
              dplyr::select(ID),
            by = c("OBJECTID" = "ID")) %>%
  st_as_sf()

accessibility <- accessibility %>%
  left_join(dwelling_network_points_2016 %>%
              dplyr::select(ID),
            by = c("UID" = "ID")) %>%
  st_as_sf()
```

Plot levels of service before and during covid:
```{r}
los_pre <- ggplot() +
  geom_sf(data = los %>%
            filter(timing == "Pre-covid"),
          aes(color = los,
              size = los)) +
  geom_sf(data = foodbank_locations,
          aes(shape = COVIDClose)) +
  scale_color_distiller(palette = "OrRd", 
                        direction = 1) +
  scale_shape_manual(values = c("Yes" = 4, "No" = 1))

los_covid <- ggplot() +
  geom_sf(data = los %>%
            filter(timing == "During covid"),
          aes(color = los,
              size = los)) +
  geom_sf(data = foodbank_locations,
          aes(shape = COVIDClose)) +
  scale_color_distiller(palette = "OrRd", 
                        direction = 1) +
  scale_shape_manual(values = c("Yes" = 4, "No" = 1))

los_pre / los_covid
```

Plot the change in level of service:
```{r}
los %>%
  st_drop_geometry() %>%
  pivot_wider(names_from = timing, 
              values_from = los, 
              values_fill = 0) %>%
  filter(`During covid` > 0) %>%
  mutate(los_d = `During covid` - `Pre-covid`) %>%
  left_join(foodbank_locations %>%
            transmute(OBJECTID = ID,
                      geometry),
          by = "OBJECTID") %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(color = los_d,
              size = los_d)) +
  geom_sf(data = foodbank_locations,
          aes(shape = COVIDClose)) +
  scale_color_distiller(palette = "OrRd", 
                        direction = -1) +
  scale_shape_manual(values = c("Yes" = 4, "No" = 1))

```

Here we can see the places where the level of service deteriorated most (excluding the places that closed, where it vanished).

Plot accessibility before and during covid:
```{r}
acc_pre <- ggplot() +
  geom_sf(data = accessibility %>%
            filter(timing == "Pre-covid"),
          aes(color = accessibility)) +
  scale_color_distiller(palette = "OrRd", 
                        direction = 1)

acc_covid <- ggplot() +
  geom_sf(data = accessibility %>%
            filter(timing == "During covid"),
          aes(color = accessibility)) +
  scale_color_distiller(palette = "OrRd", 
                        direction = 1)

acc_pre / acc_covid
```

Plot the change in accessibility by DA:
```{r}
accessibility %>%
  st_drop_geometry() %>%
  # Copy the DA identifier to the accessibility table
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
            transmute(UID = ID,
                      DAUID),
          by = "UID") %>%
  # Group by DA and time
  group_by(DAUID, timing) %>%
  # Calculate the accessibility by DA before and during covid
  summarize(accessibility = sum(accessibility),
            timing = first(timing),
            .groups = "drop") %>%
  # Pivot wider to calculate the differences in accessibility from pre-covid to covid
  pivot_wider(names_from = timing, 
              values_from = accessibility) %>%
  # Calculate changes in accessibility
  mutate(acc_d = `During covid` - `Pre-covid`) %>%
  # Join DA geometry
  left_join(data_da_2016 %>%
            transmute(DAUID = GeoUID,
                      geometry),
          by = "DAUID") %>%
  # Convert to simple features
  st_as_sf() %>%
  # Plot
  ggplot() +
  geom_sf(aes(fill = acc_d),
          color = NA) +
  scale_fill_distiller(palette = "OrRd", 
                        direction = -1)

```

<!--
Calculate quintiles of median household income in Hamilton:
```{r}
mhi_q <- data_da_2016 %>%
dplyr::filter(`Region Name` == "Hamilton") %>%
pull(`v_CA16_2397: Median total income of households in 2015 ($)`) %>%
quantile(c(0, 0.2, 0.4, 0.6, 0.8, 1), 
na.rm = TRUE)
```

Label DAs with income quintiles:
```{r}
data_da_2016 <- data_da_2016 %>%
rename(Median_household_income = `v_CA16_2397: Median total income of households in 2015 ($)`) %>%
mutate(income_quintile = case_when(Median_household_income < mhi_q[2] ~ "Bottom 20%",
Median_household_income >= mhi_q[2] & Median_household_income < mhi_q[3] ~ "Fourth 20%",
Median_household_income >= mhi_q[3] & Median_household_income < mhi_q[4] ~ "Third 20%",
Median_household_income >= mhi_q[4] & Median_household_income < mhi_q[5] ~ "Second 20%",
Median_household_income >= mhi_q[5] ~ "Top 20%"),
income_quintile = factor(income_quintile,
levels = c("Top 20%",
"Second 20%",
"Third 20%",
"Fourth 20%",
"Bottom 20%"),
ordered = TRUE))
```

## Exploratory data analysis

Summary of population:
```{r}
sum(data_da_2016$Population_55to69, na.rm = TRUE)
sum(data_da_2016$Population_70plus, na.rm = TRUE)
```

Check the total population:
```{r}
sum(dwelling_network_points_2016$Population_55to69, na.rm = TRUE)
sum(dwelling_network_points_2016$Population_70plus, na.rm = TRUE)
```

The difference in population is because the residential units considered do not completely cover the Hamilton CMA:
```{r}
ggplot() +
geom_sf(data = hamilton_cma) + 
geom_sf(data = dwelling_network_points_2016) + 
geom_sf(data = pharmacy_locations, 
aes(color = Type,
shape = Type), 
size = 3)
```

Plot population age 55 to 69:
```{r}
ggplot() +
geom_sf(data = data_da_2016 %>%
filter(`Region Name` == "Hamilton"),
aes(fill = Population_55to69),
color = NA) +
geom_sf(data = hamilton_cma,
fill = NA) +
scale_fill_distiller(palette = "OrRd", 
direction = 1)
```

Plot population age 70 plus:
```{r}
ggplot() +
geom_sf(data = data_da_2016 %>%
filter(`Region Name` == "Hamilton"),
aes(fill = Population_70plus),
color = NA) +
geom_sf(data = hamilton_cma,
fill = NA) +
scale_fill_distiller(palette = "OrRd", 
direction = 1)
```
-->

### 2021-04-19 00-Data-Processing/00-Data-Processing-for-Food-Banks-Accessibility.Rmd
Update food bank and routing code
Changes related to updating the food banks csv with more info used to filter sites

@@
                                     ADDRESS = col_character(),
-                                    COVIDClose = col_double(),
+                                    STATUS = col_character(),
                                     latitude = col_double(),
@@
 foodbank_locations <- foodbank_locations %>%
-  mutate(COVIDClose = replace_na(COVIDClose, 0),
-         COVIDClose = factor(COVIDClose,
-                             levels = c(0, 1),
-                             labels = c("No", "Yes")))
+  mutate(STATUS = replace_na(STATUS, "Open"),
+         COVIDClose = if_else(STATUS == "Cancelled", true = "Yes", false = "No")) %>%
+
+  # filter out Meals on Wheels, Student Nutrition programs, Shopping and Transportation, & Good Food Box Programs
+  filter(TYPE %in% c("Community Kitchen", "Congregate Dining", "Community Meals", "Food Banks", "Free Meals", "Low Cost Meals"))
 ```
@@
 Convert to simple features:
-```{r}
+```{r message = FALSE}
 foodbank_locations <- foodbank_locations %>%
@@
            crs = 4326) %>%
-  st_transform(crs = 26917)
+  #st_transform(crs = 26917) %>%
+
+  # filter out points beyond the Hamilton CMA (there's one point in Brantford)
+  filter(st_intersects(x = geometry, y = hamilton_cma, sparse = FALSE))
 ```
@@
 
-### `dwelling_network_points_2016.RData`
+### Origin parcel locations
 

### 2021-04-19 00-Data-Processing/04-Data-Analysis.Rmd
Update food bank and routing code
Changes related to updating the food banks csv with more info used to filter sites

@@
 
-Join the pharmacy data to the travel time matrices:
+Join the food bank data to the travel time matrices:
 ```{r}

### 2021-04-19 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Started transferring text from original manuscript, and edited the introduction

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{An awesome journal} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Changes in accessibility to foodbanks during COVID-19 and implications for the food security of vulnerable populations in Hamilton, Ontario},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% Pandoc header

\begin{document}
\begin{frontmatter}

  \title{Changes in accessibility to foodbanks during COVID-19 and
implications for the food security of vulnerable populations in
Hamilton, Ontario}
    \author[Some University]{Author One\corref{1}}
   \ead{author1@example.com} 
    \author[]{Author Two}
   \ead{author2@example.com} 
    \author[Some University]{Author Three\corref{2}}
   \ead{author3@example.com} 
    \author[Another University]{Author Four\corref{2}}
   \ead{author4@example.com} 
      \address[Some University]{Department, Street, City, State, Zip}
    \address[Another University]{Department, Street, City, State, Zip}
      \cortext[1]{Corresponding Author}
    \cortext[2]{Equal contribution}

  \begin{abstract}
  This is the abstract.

  It consists of two paragraphs.
  \end{abstract}

 \end{frontmatter}

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Over at least the past four decades food banks have become an essential
line of defense against food insecurity in Canadian communities (Black
and Seto, 2020; Riches, 2002; Tarasuk et al., 2020). In this respect,
Canada is not unlike numerous other wealthy countries where a systematic
dismantling of the welfare state took place in the intervening period
(Tarasuk et al., 2014).

The emergence of COVID-19, the worst public health crisis since the 1918
flu pandemic, has revealed important social and economic fault lines,
and pre-existing patterns of inequality appear to have been exacerbated.
Along several other dimensions of stress {[}e.g., accessibility to
health care facilities; Pereira et al. (2021){]}, this seems to be the
case for food insecurity as well (Laborde et al., 2020). In the US, for
example, it has been estimated that there was an increase of more than
30\% in household food insecurity, and more than one third of households
were discovered to be newly food insecure - meaning they did not
experience food insecurity before the pandemic (Niles et al., 2020). In
Canada, Men and Tarasuk (2021) report that about 25\% of individuals who
experienced job insecurity (a relatively common occurrence during the
pandemic), also experienced food insecurity. Similarly, according to
Statistics Canada (2020), in the early stages of the pandemic almost
15\% of individuals reported living in a household that faced food
insecurity; the risk of food insecurity was substantially higher for
households with children. The difference between households with and
without children was significant, and 11.7\% of households with children
indicated that ``food didn't last and {[}there was{]} no money to get
more'' sometimes or often, compared to 7.3\% of households without
children); likewise, 13\% of households with children indicated that
they ``{[}c{]}ouldn't afford balanced meals'' sometimes or often,
compared to 8.8\% of households without children.

The impacts of food insecurity during the pandemic are alarming, since
diet-related diseases, such as obesity, heart-disease, and diabetes,
were already critical public health concerns in Canada prior to COVID-19
(Boucher et al., 2017). While foodbanks are not necessarily a stable
solution to food insecurity and in fact may encourage a retrenchment of
neoliberal policy (Wakefield et al., 2013), at least can be argued to
provide a resource of last instance to households in precarious
situations (Bazerghi et al., 2016). As recently as 2019, the Hamilton
Hunger Report\footnote{https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf}
noted that food banks in Hamilton, Ontario, recorded the highest number
of visitors in the past 29 years; the number of children visiting
foodbanks (minors up to 18 years old) was 9,125 in March 2019, up from
8,278 the year before. It is known that the urban food environment,
within which people make their daily food choices, is essential in
influencing eating behaviours and health outcomes, based on factors such
as food availability, ease of accessibility and socio-demographic
variations (Paez et al., 2010; Vanderlee and L'Abbé, 2017). To add to
the distress of vulnerable households, non-pharmaceutical interventions
during the pandemic involving restrictions in mobility increased the
friction of travel, in particular by transit on which they are more
likely to be reliant (e.g., DeWeese et al., 2020); while at the same
time creating stress for the operators of foodbanks through disruptions
in the supply chain (e.g., McKay et al., 2021), in addition to concerns
around delivery of service in safe conditions.

For this study, we aim to look at how the landscape of food bank
services available in Hamilton, Ontario, has changed before and during
the pandemic. Have the number of open food bank services diminished? If
so, what was the accessibility to foodbanks before and during the
pandemic, from the perspective of low income households? And finally,
who are most likely to have been impacted by changes to the
accessibility landscape? This paper will first look at the distribution
of foodbanks and related services before and during the pandemic. Then,
we use the balanced floating catchment areas approach of Paez et al.
(Paez et al., 2019) to investigate the accessibility situation. We use a
fully disaggregated approach based on parcel-level data. Socio-economic
and demographic data are drawn from the latest Census of Canada (2016),
whereas travel information is from the most recent regional travel
survey from 2016. This paper follows reproducible research
recommendations (see Brunsdon and Comber, 2020), and the research was
conducted using open source tools for transportation analysis (Lovelace,
2021). The code and data necessary to reproduce the analysis are
available in a public repository\footnote{add repository}.

\hypertarget{literature-review}{%
\section{Literature Review}\label{literature-review}}

\hypertarget{food-insecurity}{%
\subsection{Food Insecurity}\label{food-insecurity}}

Food insecurity is the inability to acquire and consume an adequate
amount or good quality food, leading to inadequate nutrient intake
(Kirkpatrick and Tarasuk 2008; Tarasuk and Vogt 2009). This nutrient
deficiency has been causing major health concerns in Canadians, and
particularly those who are at a socioeconomic disadvantage (Bazerghi,
McKay, and Dunn 2016). Previous studies have aimed to look at the
relationship between the built food environment and sociodemographic
characteristics with qualitative and quantitative, or a mixed -method
approach. Quantitatively, official governmental surveys have been able
to assist with data, such as the Household Food Security Survey Module
(HFSSM), the Canadian Community Health Surveys (CCHS) , the Longitudinal
and International Study of Adults (LISA), and official classifications
determined by Health Canada in relation to sociodemographic variables
(El-hajj and Benhin 2021; Gundersen et al.~2018; Kirkpatrick and Tarasuk
2008; Tarasuk and Vogt 2009). Studies have also aimed to assess food
availability of healthy foods (e.g., fruit and vegetables) at
supermarkets in relation to sociodemographic characteristics and
geographic accessibility (Latham and Moffat 2007). In terms of findings,
studies have generated inconsistent relationships between their
evaluated availability of food and sociodemographic characteristics that
go in hand.

\hypertarget{food-banks}{%
\subsection{Food Banks}\label{food-banks}}

The number of food banks has been increasing steadily in Canada
(Wakefield et al.~2013). What was supposed to be a temporary solution to
accommodate those in hunger due to job retrenchments and economic
downfalls since the 1980s, has now developed into a community practice
to secure emergency food supplies for those in need (Loopstra and
Tarasuk 2012; Wakefield et al.~2013). Food banks were developed as a
part of a community response to aide those with inadequate food by
voluntarily giving meals and ingredients away (Loopstra and Tarasuk
2012; Riches 2002). The scope and objectives of food banks can vary by
region and by country. These organizations can include not only prepared
meals and aliments, but also shared spaces to connect in community
gardens and community kitchens (Wakefield et al.~2013) or referred to as
`food pantry' and `food shelf,' where many frequent as their primary
location to get food (Bazerghi et al.~2016). However, surveys revealed
that only 20 to 30 percent of those experiencing food insecurity were
found to frequent food banks in Canada (Tarasuk et al.~2014).

Previous studies also question if food banks are able to offer a rounded
nutritious supply of food, and if food banks are a sustainable practice
for those in need to continuously obtain their food from (Bazerghi et
al.~2016; Riches 2002).

\hypertarget{food-insecurity-in-canada-during-the-covid-19}{%
\subsection{Food Insecurity in Canada during the
COVID-19}\label{food-insecurity-in-canada-during-the-covid-19}}

Currently with the COVID-19 pandemic, disrupted economies, rising
unemployment rates, and alarming poverty levels have disrupted the food
environment by causing higher rates of food insecurity (Niles et
al.~2020). In 2012, 12.4\% of Canadians households and 11.8\% Ontarian
households experienced some degree of food insecurity (Gundersen et
al.~2018; Tarasuk, Fafard St-Germain, and Mitchell 2019). Recent
COVID-19 related food security studies in the US have found a massive
increase in households experiencing food insecurity for the first time,
and also in households experiencing more severe food insecurity than
before (Niles et al.~2020; Wolfson and Leung 2020). Most recently in May
2020, Canada recorded 14.7\% of its population living in food insecurity
in the past 30 days(Statistics Canada 2020). Food insecurity is a highly
concerning public health issue due the vast health consequences that it
can induce. Recent publishing suggests that food insecurity in adults
could lead to experiencing more stressful events (El-hajj and Benhin
2021). These increases in food insecurity rates due to the pandemic,
signal a change in the food environment with potential damages to health
outcomes in populations (Niles et al.~2020)

\hypertarget{methods}{%
\section{Methods}\label{methods}}

Words go here.

\hypertarget{results-and-discussion}{%
\section{Results and Discussion}\label{results-and-discussion}}

Words go here.

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

Words go here.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-bazerghi2016role}{}%
Bazerghi, C., McKay, F.H., Dunn, M., 2016. The role of food banks in
addressing food insecurity: A systematic review. Journal of community
health 41, 732--740.

\leavevmode\hypertarget{ref-black2020examining}{}%
Black, J.L., Seto, D., 2020. Examining patterns of food bank use over
twenty-five years in vancouver, canada. VOLUNTAS: International Journal
of Voluntary and Nonprofit Organizations 31, 853--869.
doi:\url{https://doi.org/10.1007/s11266-018-0039-2}

\leavevmode\hypertarget{ref-boucher2017ontario}{}%
Boucher, B.A., Manafò, E., Boddy, M.R., Roblin, L., Truscott, R., 2017.
The ontario food and nutrition strategy: Identifying indicators of food
access and food literacy for early monitoring of the food environment.
Health promotion and chronic disease prevention in Canada : research,
policy and practice 37, 313--319.
doi:\href{https://doi.org/10.24095/hpcdp.37.9.06}{10.24095/hpcdp.37.9.06}

\leavevmode\hypertarget{ref-brunsdon2020opening}{}%
Brunsdon, C., Comber, A., 2020. Opening practice: Supporting
reproducibility and critical spatial data science. Journal of
Geographical Systems 1--20.
doi:\href{https://doi.org/10.1007/s10109-020-00334-2}{10.1007/s10109-020-00334-2}

\leavevmode\hypertarget{ref-deweese2020tale}{}%
DeWeese, J., Hawa, L., Demyk, H., Davey, Z., Belikow, A., El-geneidy,
A., 2020. A tale of 40 cities: A preliminary analysis of equity impacts
of COVID-19 service adjustments across north america. Findings.
doi:\href{https://doi.org/10.32866/001c.13395}{10.32866/001c.13395}

\leavevmode\hypertarget{ref-laborde2020poverty}{}%
Laborde, D., Martin, W., Vos, R., 2020. Poverty and food insecurity
could grow dramatically as COVID-19 spreads. International Food Policy
Research Institute (IFPRI), Washington, DC.

\leavevmode\hypertarget{ref-lovelace2021open}{}%
Lovelace, R., 2021. Open source tools for geographic analysis in
transport planning. Journal of Geographical Systems.
doi:\href{https://doi.org/10.1007/s10109-020-00342-2}{10.1007/s10109-020-00342-2}

\leavevmode\hypertarget{ref-mckay2021exploring}{}%
McKay, F.H., Bastian, A., Lindberg, R., 2021. Exploring the response of
the victorian emergency and community food sector to the COVID-19
pandemic. Journal of Hunger \& Environmental Nutrition 1--15.
doi:\href{https://doi.org/10.1080/19320248.2021.1900974}{10.1080/19320248.2021.1900974}

\leavevmode\hypertarget{ref-men2021food}{}%
Men, F., Tarasuk, V., 2021. Food insecurity amid the COVID-19 pandemic:
Food charity, government assistance and employment. Canadian Public
Policy COVID-19, e2021001.
doi:\href{https://doi.org/10.3138/cpp.2021-001}{10.3138/cpp.2021-001}

\leavevmode\hypertarget{ref-niles2020early}{}%
Niles, M.T., Bertmann, F., Belarmino, E.H., Wentworth, T., Biehl, E.,
Neff, R., 2020. The early food insecurity impacts of COVID-19. Nutrients
12, 2096.

\leavevmode\hypertarget{ref-paez2019demand}{}%
Paez, A., Higgins, C.D., Vivona, S.F., 2019. Demand and level of service
inflation in floating catchment area (FCA) methods. PloS one 14,
e0218773.
doi:\href{https://doi.org/10.1371/journal.pone.0218773}{10.1371/journal.pone.0218773}

\leavevmode\hypertarget{ref-paez2010relative}{}%
Paez, A., Mercado, R.G., Farber, S., Morency, C., Roorda, M., 2010.
Relative accessibility deprivation indicators for urban settings:
Definitions and application to food deserts in montreal. Urban Studies
47, 1415--1438.
doi:\href{https://doi.org/10.1177/0042098009353626}{10.1177/0042098009353626}

\leavevmode\hypertarget{ref-pereira2021geographic}{}%
Pereira, R.H.M., Braga, C.K.V., Servo, L.M., Serra, B., Amaral, P.,
Gouveia, N., Paez, A., 2021. Geographic access to COVID-19 healthcare in
brazil using a balanced float catchment area approach. Social Science \&
Medicine 273, 113773.
doi:\url{https://doi.org/10.1016/j.socscimed.2021.113773}

\leavevmode\hypertarget{ref-riches2002food}{}%
Riches, G., 2002. Food banks and food security: Welfare reform, human
rights and social policy. Lessons from canada? Social Policy \&
Administration 36, 648--663.

\leavevmode\hypertarget{ref-statisticscanada2020food}{}%
Statistics Canada, 2020. Food insecurity during the COVID-19 pandemic
(No. Catalogue no. 45280001).

\leavevmode\hypertarget{ref-tarasuk2014food}{}%
Tarasuk, V., Dachner, N., Loopstra, R., 2014. Food banks, welfare, and
food insecurity in canada. British Food Journal.

\leavevmode\hypertarget{ref-tarasuk2020relationship}{}%
Tarasuk, V., St-Germain, A.-A.F., Loopstra, R., 2020. The relationship
between food banks and food insecurity: Insights from canada. VOLUNTAS:
International Journal of Voluntary and Nonprofit Organizations 31,
841--852.

\leavevmode\hypertarget{ref-vanderlee2017food}{}%
Vanderlee, L., L'Abbé, M., 2017. Food for thought on food environments
in canada. Health promotion and chronic disease prevention in Canada :
research, policy and practice 37, 263--265.
doi:\href{https://doi.org/10.24095/hpcdp.37.9.01}{10.24095/hpcdp.37.9.01}

\leavevmode\hypertarget{ref-wakefield2013sweet}{}%
Wakefield, S., Fleming, J., Klassen, C., Skinner, A., 2013. Sweet
charity, revisited: Organizational responses to food insecurity in
hamilton and toronto, canada. Critical Social Policy 33, 427--450.

\end{CSLReferences}

\end{document}

### 2021-04-19 00-Data-Processing/01-Routing-Car.Rmd

@@
 
-Read file with locations:
+Read file with food bank locations:
 ```{r}
-foodbank_locations <- read_csv("input-data-files/foodbanks-hamilton.csv",
-                               col_types = cols(ID = col_double(),
-                                    NAME = col_character(),
-                                    TYPE = col_character(),
-                                    ADDRESS = col_character(),
-                                    COVIDClose = col_double(),
-                                    latitude = col_double(),
-                                    longitude = col_double(),
-                                    name = col_character(),
-                                    desc = col_character(),
-                                    source = col_character(),
-                                    precision = col_character()
-                               ))
+load("output-data-files/foodbank_locations.RData")
 ```
 
-COVIDClose is an indicator variable for whether this bank closed during the pandemic. Convert to factor:
-```{r}
-foodbank_locations <- foodbank_locations %>%
-  mutate(COVIDClose = replace_na(COVIDClose, 0),
-         COVIDClose = factor(COVIDClose,
-                             levels = c(0, 1),
-                             labels = c("No", "Yes")))
-```
-
-Convert to simple features:
-```{r}
-foodbank_locations <- foodbank_locations %>%
-  st_as_sf(coords = c("longitude", "latitude"), 
-           crs = 4326) %>%
-  st_transform(crs = 26917)
-```
-
-Map location of foodbanks:
-```{r}
-ggplot() +
-  geom_sf(data = hamilton_cma) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose,
-              color = COVIDClose))
-```
-
-### `dwelling_network_points_2016.RData`
+### Location of origin parcels
 

### 2021-04-19 00-Data-Processing/04-Data-Analysis.Rmd

@@
 
+## TEMP: REMOVE INAPPLICABLE DESTINATIONS UNTIL NEW TTMs ARE FINISHED
+
+```{r}
+ttm_car <- ttm_car %>%
+  filter(!between(OBJECTID, left = 59, right = 69))
+
+ttm_transit <- ttm_transit %>%
+  filter(!between(OBJECTID, left = 59, right = 69))
+
+ttm_walk <- ttm_walk %>%
+  filter(!between(OBJECTID, left = 59, right = 69))
+```
+
 ## Quick check
@@
 
-Half of all parcels can reach 63 foodbanks or more in 180 min or less.
+Half of all parcels can reach 52 foodbanks or more in 180 min or less.
 
@@
 
-Half of all parcels can reach 50 foodbanks or more in 180 min or less.
+Half of all parcels can reach 45 foodbanks or more in 180 min or less.
 
 ## Prepare data
-
-Calculate number of households with incomes < \$40k per DA and proportion of households with incomes < \$40k per DA:
-```{r}
-data_da_2016 <- data_da_2016 %>%
-  filter(`Region Name` == "Hamilton") %>%
-  mutate(income_less40k = (`v_CA16_2406: Under $5,000` + 
-                             `v_CA16_2407: $5,000 to $9,999` + 
-                             `v_CA16_2408: $10,000 to $14,999` +
-                             `v_CA16_2409: $15,000 to $19,999` +
-                             `v_CA16_2410: $20,000 to $24,999` +  
-                             `v_CA16_2411: $25,000 to $29,999` +
-                             `v_CA16_2412: $30,000 to $34,999` +
-                             `v_CA16_2413: $35,000 to $39,999`),
-         p_inc_less40k = income_less40k/Households)
-```
-
-Clip `data_da_2016`:
-```{r}
-data_da_2016 <- data_da_2016 %>%
-  st_intersection(hamilton_cma)
-```
 
@@
               dplyr::transmute(DAUID = GeoUID, 
-                               p_inc_less40k),
+                               p_inc_less40k,
+                               linc_age_0to2,
+                               linc_age_0to19,
+                               linc_age_65plus),
             by = "DAUID")
@@
                             income_less40k,
+                            linc_age_0to2,
+                            linc_age_0to19,
+                            linc_age_65plus,
                             DAUID, 
@@
 What does this mean?
-Most DAs missing travel time by transit do not report trips by transit.
+Most DAs missing travel time by transit do not report trips by transit in the TTS.
 In some DAs there are trips by transit (p_transit != 0), but no foodbank can be reached by transit (travel_time_transit is NA).
@@
 What does this mean?
-Most DAs missing travel time by walking do not report trips by walkin In some DAs there are trips by walking (p_walk != 0), but no foodbank can be reached by walk (travel_time_walk is NA).
+Most DAs missing travel time by walking do not report trips by walking in the TTS.
+In some DAs there are trips by walking (p_walk != 0), but no foodbank can be reached by walk (travel_time_walk is NA).
 How many of these cases?
@@
 Calculate expected travel times by mode.
-These are the travel time by the mode multiplied by the probability of using the mode (which is proxied by the proportion of trips by mode in the DA).
+These are the travel time by the mode multiplied by the probability of using the mode (which is proxied by the proportion of trips by mode in the DA from the TTS).
 The total expected travel time is the sum of the expected travel times by mode:
@@
 
-What is the mean expected travel time for each parcel over all the foodbanks?
+What is the mean expected travel time for each parcel over all the foodbanks (open and closed)?
 ```{r}
@@
 ggplot() +
+  geom_sf(data = hamilton_cma) +
   geom_sf(data = e_travel_time_parcel,
@@
 ggplot() +
+  geom_sf(data = hamilton_cma) +
   geom_sf(data = e_travel_time_parcel_2,
@@
 
-Notice how expected travel time can _decrease_ with the closures! Foodbanks on average are a little bit closer, but there are _fewer_ of them, which will tend to increase the demand: 
+Notice how expected travel time to all open food banks can _decrease_ with the closures!
+Food banks on average are a little bit closer, but there are _fewer_ of them, which will tend to increase the demand:
 ```{r}
@@
 los_pre <- ggplot() +
+  geom_sf(data = hamilton_cma) +
   geom_sf(data = los %>%
@@
 los_covid <- ggplot() +
+  geom_sf(data = hamilton_cma) +
   geom_sf(data = los %>%
@@
   filter(`During covid` > 0) %>%
-  mutate(los_d = `During covid` - `Pre-covid`) %>%
+  mutate(los_d = `During covid` / `Pre-covid` - 1) %>%
   left_join(foodbank_locations %>%
@@
   ggplot() +
+  geom_sf(data = hamilton_cma) +
   geom_sf(aes(color = los_d,
@@
 acc_pre <- ggplot() +
+  geom_sf(data = hamilton_cma) +
   geom_sf(data = accessibility %>%
@@
 acc_covid <- ggplot() +
+  geom_sf(data = hamilton_cma) +
   geom_sf(data = accessibility %>%
@@
               st_drop_geometry() %>%
-            transmute(UID = ID,
-                      DAUID),
+              transmute(UID = ID, DAUID),
             by = "UID") %>%
@@
   # Calculate changes in accessibility
-  mutate(acc_d = `During covid` - `Pre-covid`) %>%
+  mutate(acc_d = `During covid` / `Pre-covid` - 1) %>%
   # Join DA geometry
@@
   ggplot() +
+  geom_sf(data = hamilton_cma) +
   geom_sf(aes(fill = acc_d),
@@
 
-<!--
-Calculate quintiles of median household income in Hamilton:
+## Further Breakdown by Population Characteristics
+
+### CODE TO BE UPDATED
 ```{r}
-mhi_q <- data_da_2016 %>%
-dplyr::filter(`Region Name` == "Hamilton") %>%
-pull(`v_CA16_2397: Median total income of households in 2015 ($)`) %>%
-quantile(c(0, 0.2, 0.4, 0.6, 0.8, 1), 
-na.rm = TRUE)
+# join tripmaking modal proportions
+ttm_car <- ttm_car %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+
+ttm_transit <- ttm_transit %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+
+ttm_walk <- ttm_walk %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
 ```
 
-Label DAs with income quintiles:
+Get minimum travel times in the pre-covid scenario:
+
+```{r minimum-travel-time-baseline, include=FALSE}
+# retrieve minimum travel time during covid to sites that didnt close
+min_ttm_car_pre <- ttm_car %>%
+  #dplyr::filter(COVIDClose == "No") %>%
+  group_by(UID) %>%
+  summarize(travel_time = min(travel_time),
+            linc_age_0to2 = first(linc_age_0to2),
+            linc_age_0to19 = first(linc_age_0to19),
+            linc_age_65plus = first(linc_age_0to19),
+            DAUID = first(DAUID),
+            TAZUID = first(TAZUID),
+            p_car = first(p_car),
+            p_transit = first(p_transit),
+            p_walk = first(p_walk),
+            .groups = "drop")
+
+min_ttm_transit_pre <- ttm_transit %>%
+  #dplyr::filter(COVIDClose == "No") %>%
+  group_by(UID) %>%
+  summarize(travel_time = min(travel_time),
+            linc_age_0to2 = first(linc_age_0to2),
+            linc_age_0to19 = first(linc_age_0to19),
+            linc_age_65plus = first(linc_age_0to19),
+            DAUID = first(DAUID),
+            TAZUID = first(TAZUID),
+            p_car = first(p_car),
+            p_transit = first(p_transit),
+            p_walk = first(p_walk),
+            .groups = "drop")
+
+min_ttm_walk_pre <- ttm_walk %>%
+  #dplyr::filter(COVIDClose == "No") %>%
+  group_by(UID) %>%
+  summarize(travel_time = min(travel_time),
+            linc_age_0to2 = first(linc_age_0to2),
+            linc_age_0to19 = first(linc_age_0to19),
+            linc_age_65plus = first(linc_age_0to19),
+            DAUID = first(DAUID),
+            TAZUID = first(TAZUID),
+            p_car = first(p_car),
+            p_transit = first(p_transit),
+            p_walk = first(p_walk),
+            .groups = "drop")
+```
+
+Get minimum travel times in the during covid scenario:
+
+```{r minimum-travel-time-during-covid, include=FALSE}
+# retrieve minimum travel time during covid to sites that didnt close
+min_ttm_car_during <- ttm_car %>%
+  dplyr::filter(COVIDClose == "No") %>%
+  group_by(UID) %>%
+  summarize(travel_time = min(travel_time),
+            linc_age_0to2 = first(linc_age_0to2),
+            linc_age_0to19 = first(linc_age_0to19),
+            linc_age_65plus = first(linc_age_0to19),
+            DAUID = first(DAUID),
+            TAZUID = first(TAZUID),
+            p_car = first(p_car),
+            p_transit = first(p_transit),
+            p_walk = first(p_walk),
+            .groups = "drop")
+
+min_ttm_transit_during <- ttm_transit %>%
+  dplyr::filter(COVIDClose == "No") %>%
+  group_by(UID) %>%
+  summarize(travel_time = min(travel_time),
+            linc_age_0to2 = first(linc_age_0to2),
+            linc_age_0to19 = first(linc_age_0to19),
+            linc_age_65plus = first(linc_age_0to19),
+            DAUID = first(DAUID),
+            TAZUID = first(TAZUID),
+            p_car = first(p_car),
+            p_transit = first(p_transit),
+            p_walk = first(p_walk),
+            .groups = "drop")
+
+min_ttm_walk_during <- ttm_walk %>%
+  dplyr::filter(COVIDClose == "No") %>%
+  group_by(UID) %>%
+  summarize(travel_time = min(travel_time),
+            linc_age_0to2 = first(linc_age_0to2),
+            linc_age_0to19 = first(linc_age_0to19),
+            linc_age_65plus = first(linc_age_0to19),
+            DAUID = first(DAUID),
+            TAZUID = first(TAZUID),
+            p_car = first(p_car),
+            p_transit = first(p_transit),
+            p_walk = first(p_walk),
+            .groups = "drop")
+```
+
 ```{r}
-data_da_2016 <- data_da_2016 %>%
-rename(Median_household_income = `v_CA16_2397: Median total income of households in 2015 ($)`) %>%
-mutate(income_quintile = case_when(Median_household_income < mhi_q[2] ~ "Bottom 20%",
-Median_household_income >= mhi_q[2] & Median_household_income < mhi_q[3] ~ "Fourth 20%",
-Median_household_income >= mhi_q[3] & Median_household_income < mhi_q[4] ~ "Third 20%",
-Median_household_income >= mhi_q[4] & Median_household_income < mhi_q[5] ~ "Second 20%",
-Median_household_income >= mhi_q[5] ~ "Top 20%"),
-income_quintile = factor(income_quintile,
-levels = c("Top 20%",
-"Second 20%",
-"Third 20%",
-"Fourth 20%",
-"Bottom 20%"),
-ordered = TRUE))
+# FULL JOIN THEM TOGETHER
 ```
@@
 ```
--->

### 2021-04-20 00-Data-Processing/01-Routing-Car.Rmd
Revised ttm's to destinations
Re-ran the ttm calculations to the reduced list of applicable destinations

@@
                                             outdir = "./df/output_ttm_car",
-                                            compress = 50,
+                                            compress = 100,
                                             overwrite = TRUE),

### 2021-04-20 00-Data-Processing/02-Routing-Transit.Rmd
Revised ttm's to destinations
Re-ran the ttm calculations to the reduced list of applicable destinations

@@
                                             outdir = "./df/output_ttm_transit",
-                                            compress = 50,
+                                            compress = 100,
                                             overwrite = TRUE),

### 2021-04-20 00-Data-Processing/03-Routing-Walk.Rmd
Revised ttm's to destinations
Re-ran the ttm calculations to the reduced list of applicable destinations

@@
                                             outdir = "./df/output_ttm_walk",
-                                            compress = 50,
+                                            compress = 100,
                                             overwrite = TRUE),

### 2021-04-20 00-Data-Processing/00-Data-Processing-for-Food-Banks-Accessibility.Rmd

@@
 ```{r eval=FALSE}
-set_api_key("your-key", install = TRUE)
+set_api_key("YOUR API KEY HERE")
 ```
@@
 ```{r}
-find_census_vectors("income", 
+find_census_vectors("age", 
                     dataset = "CA16", 
@@
                                     "v_CA16_2425",
-                                    "v_CA16_2426"),
+                                    "v_CA16_2426",
+                                    "v_CA16_2397", # median hhinc
+                                    # age variables
+                                    "v_CA16_10", #<1
+                                    "v_CA16_13", #1
+                                    "v_CA16_16", #2
+                                    "v_CA16_7", #0-4 yo
+                                    "v_CA16_25", #5-9 yo
+                                    "v_CA16_43", #10-14 yo
+                                    "v_CA16_64", # 15-19 yo
+                                    "v_CA16_244"), # 65+
                           level='DA', 
@@
                           geo_format = "sf")
+```
+
+#### Prepare Geography
+
+Clip `data_da_2016`:
+```{r}
+data_da_2016 <- data_da_2016 %>%
+  st_intersection(hamilton_cma)
+```
+
+#### Prepare Summary Measures
+
+Calculate number of households with incomes < \$40k per DA and proportion of households with incomes < \$40k per DA:
+```{r}
+data_da_2016 <- data_da_2016 %>%
+  filter(`Region Name` == "Hamilton") %>%
+  mutate(income_less40k = (`v_CA16_2406: Under $5,000` + 
+                             `v_CA16_2407: $5,000 to $9,999` + 
+                             `v_CA16_2408: $10,000 to $14,999` +
+                             `v_CA16_2409: $15,000 to $19,999` +
+                             `v_CA16_2410: $20,000 to $24,999` +  
+                             `v_CA16_2411: $25,000 to $29,999` +
+                             `v_CA16_2412: $30,000 to $34,999` +
+                             `v_CA16_2413: $35,000 to $39,999`),
+         p_inc_less40k = income_less40k/Households)
 ```
@@
 data_da_2016 %>%
-  mutate(income_less40k = (`v_CA16_2406: Under $5,000` + 
-                             `v_CA16_2407: $5,000 to $9,999` + 
-                             `v_CA16_2408: $10,000 to $14,999` +
-                             `v_CA16_2409: $15,000 to $19,999` +
-                             `v_CA16_2410: $20,000 to $24,999` +  
-                             `v_CA16_2411: $25,000 to $29,999` +
-                             `v_CA16_2412: $30,000 to $34,999` +
-                             `v_CA16_2413: $35,000 to $39,999`)) %>%
   tm_shape() +
   tm_polygons("income_less40k",
+              alpha = 0.8)
+```
+
+Label DAs with income quintiles:
+```{r}
+data_da_2016 <- data_da_2016 %>%
+  rename(Median_household_income = `v_CA16_2397: Median total income of households in 2015 ($)`) %>%
+  mutate(income_quintile = case_when(Median_household_income < mhi_q[2] ~ "Bottom 20%",
+                                    Median_household_income >= mhi_q[2] & Median_household_income < mhi_q[3] ~ "Fourth 20%",
+                                    Median_household_income >= mhi_q[3] & Median_household_income < mhi_q[4] ~ "Third 20%",
+                                    Median_household_income >= mhi_q[4] & Median_household_income < mhi_q[5] ~ "Second 20%",
+                                    Median_household_income >= mhi_q[5] ~ "Top 20%"),
+         income_quintile = factor(income_quintile,
+                                levels = c("Top 20%",
+                                           "Second 20%",
+                                           "Third 20%",
+                                           "Fourth 20%",
+                                           "Bottom 20%"),
+                                ordered = TRUE))
+```
+
+Calculate low-income age totals
+```{r}
+data_da_2016 <- data_da_2016 %>%
+  mutate(tot_age_0to2 = `v_CA16_10: Under 1 year` + `v_CA16_13: 1` + `v_CA16_16: 2`,
+         linc_age_0to2 = tot_age_0to2 * p_inc_less40k,
+         tot_age_0to19 = `v_CA16_7: 0 to 4 years` + `v_CA16_25: 5 to 9 years` + `v_CA16_43: 10 to 14 years` + `v_CA16_64: 15 to 19 years`,
+         linc_age_0to19 = tot_age_0to19 * p_inc_less40k,
+         tot_age_65plus = `v_CA16_244: 65 years and over`,
+         linc_age_65plus = tot_age_65plus * p_inc_less40k,
+         linc_population = Population * p_inc_less40k)
+```
+
+Plot result
+```{r}
+tmap_mode(mode = "view")
+
+data_da_2016 %>%
+  tm_shape() +
+  tm_polygons("linc_population",
               alpha = 0.8)
@@
 ```
-
-THE NEXT NEEDS TO BE CHANGED TO DO THE ANALYSIS BY INCOME GROUP, NOT BY AGE GROUP
 

### 2021-04-20 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd

---
title: "Changes in accessibility to foodbanks during COVID-19 and implications for the food security of vulnerable populations in Hamilton, Ontario"
author:
  - name: Author One
    email: author1@example.com
    affiliation: Some University
    footnote: 1
  - name: Author Two
    email: author2@example.com
    affiliation: 
  - name: Author Three
    email: author3@example.com
    affiliation: Some University
    footnote: 2 
  - name: Author Four
    email: author4@example.com
    affiliation: Another University
    footnote: 2
address:
  - code: Some University
    address: Department, Street, City, State, Zip
  - code: Another University
    address: Department, Street, City, State, Zip
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
abstract: |
  This is the abstract.

  It consists of two paragraphs.

journal: "An awesome journal"
date: "`r Sys.Date()`"
bibliography: references.bib
#linenumbers: true
#numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
---

\newpage

Introduction
==========================

Over at least the past four decades food banks have become an essential line of defense against food insecurity in Canadian communities [@riches2002food; @black2020examining; @tarasuk2020relationship]. In this respect, Canada is not unlike numerous other wealthy countries where a systematic dismantling of the welfare state took place in the intervening period [@tarasuk2014food].

The emergence of COVID-19, the worst public health crisis since the 1918 flu pandemic, has revealed important social and economic fault lines, and pre-existing patterns of inequality appear to have been exacerbated. Along several other dimensions of stress [e.g., accessibility to health care facilities; @pereira2021geographic], this seems to be the case for food insecurity as well [@laborde2020poverty]. In the US, for example, it has been estimated that there was an increase of more than 30\% in household food insecurity, and more than one third of households were discovered to be newly food insecure - meaning they did not experience food insecurity before the pandemic [@niles2020early]. In Canada, Men and Tarasuk [-@men2021food] report that about 25\% of individuals who experienced job insecurity (a relatively common occurrence during the pandemic), also experienced food insecurity. Similarly, according to Statistics Canada [-@statisticscanada2020food], in the early stages of the pandemic almost 15\% of individuals reported living in a household that faced food insecurity; the risk of food insecurity was substantially higher for households with children. The difference between households with and without children was significant, and 11.7\% of households with children indicated that "food didn’t last and [there was] no money to get more" sometimes or often, compared to 7.3\% of households without children); likewise, 13\% of households with children indicated that they "[c]ouldn’t afford balanced meals" sometimes or often, compared to 8.8\% of households without children.

The impacts of food insecurity during the pandemic are alarming, since diet-related diseases, such as obesity, heart-disease, and diabetes, were already critical public health concerns in Canada prior to COVID-19 [@boucher2017ontario]. While foodbanks are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role]. As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that food banks in Hamilton, Ontario, recorded the highest number of visitors in the past 29 years; the number of children visiting foodbanks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278 the year before. It is known that the urban food environment, within which people make their daily food choices, is essential in influencing eating behaviours and health outcomes, based on factors such as food availability, ease of accessibility and socio-demographic variations [@paez2010relative; @vanderlee2017food]. To add to the distress of vulnerable households, non-pharmaceutical interventions during the pandemic involving restrictions in mobility increased the friction of travel, in particular by transit on which they are more likely to be reliant [e.g., @deweese2020tale]; while at the same time creating stress for the operators of foodbanks through disruptions in the supply chain [e.g., @mckay2021exploring], in addition to concerns around delivery of service in safe conditions.

For this study, we aim to look at how the landscape of food bank services available in Hamilton, Ontario, has changed before and during the pandemic. Have the number of open food bank services diminished? If so, what was the accessibility to foodbanks before and during the pandemic, from the perspective of low income households? And finally, who are most likely to have been impacted by changes to the accessibility landscape? This paper will first look at the distribution of foodbanks and related services before and during the pandemic. Then, we use the balanced floating catchment areas approach of Paez et al. [@paez2019demand] to investigate the accessibility situation. We use a fully disaggregated approach based on parcel-level data. Socio-economic and demographic data are drawn from the latest Census of Canada (2016), whereas travel information is from the most recent regional travel survey from 2016. This paper follows reproducible research recommendations [see @brunsdon2020opening], and the research was conducted using open source tools for transportation analysis [@lovelace2021open]. The code and data necessary to reproduce the analysis are available in a public repository^[add repository].

Literature Review
============

## Food Insecurity

Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake (Kirkpatrick and Tarasuk 2008; Tarasuk and Vogt 2009). This nutrient deficiency has been causing major health concerns in Canadians, and particularly those who are at a socioeconomic disadvantage (Bazerghi, McKay, and Dunn 2016). Previous studies have aimed to look at the relationship between the built food environment and sociodemographic characteristics with qualitative and quantitative, or a mixed -method approach. 
Quantitatively, official governmental surveys have been able to assist with data, such as the Household Food Security Survey Module (HFSSM), the Canadian Community Health Surveys (CCHS) , the Longitudinal and International Study of Adults (LISA), and official classifications determined by Health Canada in relation to  sociodemographic variables (El-hajj and Benhin 2021; Gundersen et al. 2018; Kirkpatrick and Tarasuk 2008; Tarasuk and Vogt 2009). Studies have also aimed to assess food availability of healthy foods (e.g., fruit and vegetables) at supermarkets in relation to sociodemographic characteristics and geographic accessibility (Latham and Moffat 2007). In terms of findings, studies have generated inconsistent relationships between their evaluated availability of food and sociodemographic characteristics that go in hand.

## Food Banks

The number of food banks has been increasing steadily in Canada (Wakefield et al. 2013). What was supposed to be a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, has now developed into a community practice to secure emergency food supplies for those in need (Loopstra and Tarasuk 2012; Wakefield et al. 2013). Food banks were developed as a part of a community response to aide those with inadequate food by voluntarily giving meals and ingredients away (Loopstra and Tarasuk 2012; Riches 2002). The scope and objectives of food banks can vary by region and by country. These organizations can include not only prepared meals and aliments, but also shared spaces to connect in community gardens and community kitchens (Wakefield et al. 2013) or referred to as ‘food pantry’ and ‘food shelf’, where many frequent as their primary location to get food (Bazerghi et al. 2016). However, surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada (Tarasuk et al. 2014).

Previous studies also question if food banks are able to offer a rounded nutritious supply of food, and if food banks are a sustainable practice for those in need to continuously obtain their food from (Bazerghi et al. 2016; Riches 2002).

## Food Insecurity in Canada during the COVID-19

Currently with the COVID-19 pandemic, disrupted economies, rising unemployment rates, and alarming poverty levels have disrupted the food environment by causing higher rates of food insecurity (Niles et al. 2020). In 2012, 12.4% of Canadians households and 11.8% Ontarian households experienced some degree of food insecurity (Gundersen et al. 2018; Tarasuk, Fafard St-Germain, and Mitchell 2019). Recent COVID-19 related food security studies in the US have found a massive increase in households experiencing food insecurity for the first time, and also in households experiencing more severe food insecurity than before (Niles et al. 2020; Wolfson and Leung 2020). Most recently in May 2020, Canada recorded 14.7% of its population living in food insecurity in the past 30 days(Statistics Canada 2020). Food insecurity is a highly concerning public health issue due the vast health consequences that it can induce. Recent publishing suggests that food insecurity in adults could lead to experiencing more stressful events (El-hajj and Benhin 2021). These increases in food insecurity rates due to the pandemic, signal a change in the food environment with potential damages to health outcomes in populations (Niles et al. 2020)

Methods
===================

Words go here.

Results and Discussion
===================

Words go here.

Conclusions
===================

Words go here.

References {#references .unnumbered}
==========

### 2021-04-20 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Update introduction
Additional contributions to intro

@@
 
-Over at least the past four decades food banks have become an essential line of defense against food insecurity in Canadian communities [@riches2002food; @black2020examining; @tarasuk2020relationship].
+Food insecurity, defined as an "inadequate or uncertain access to a sufficient quantity and/or adequate quality of food" due to a household's financial limitations [@enns2020experiences], has been associated with reductions in nutritional outcomes [@bhattacharya2004poverty; @kirkpatrick2008food; @olson1999nutrition] and physical and mental health in children and adults [@elgar2021relative; @jones2017food; @ramsey2011food; @seligman2010food; @stuff2004household].
+Over at least the past four decades food banks have become an essential line of defense against food insecurity in Canadian communities [@riches2002food; @holmes2018nothing; @black2020examining; @tarasuk2020relationship].
 In this respect, Canada is not unlike numerous other wealthy countries where a systematic dismantling of the welfare state took place in the intervening period [@tarasuk2014food].
 
-The emergence of COVID-19, the worst public health crisis since the 1918 flu pandemic, has revealed important social and economic fault lines, and pre-existing patterns of inequality appear to have been exacerbated.
+But the emergence of COVID-19, the worst public health crisis since the 1918 flu pandemic, has revealed important social and economic fault lines, and pre-existing patterns of inequality appear to have been exacerbated.
 Along several other dimensions of stress [e.g., accessibility to health care facilities; @pereira2021geographic], this seems to be the case for food insecurity as well [@laborde2020poverty].
@@
 The impacts of food insecurity during the pandemic are alarming, since diet-related diseases, such as obesity, heart-disease, and diabetes, were already critical public health concerns in Canada prior to COVID-19 [@boucher2017ontario].
-While foodbanks are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
-As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that food banks in Hamilton, Ontario, recorded the highest number of visitors in the past 29 years; the number of children visiting foodbanks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278 the year before.
-It is known that the urban food environment, within which people make their daily food choices, is essential in influencing eating behaviours and health outcomes, based on factors such as food availability, ease of accessibility and socio-demographic variations [@paez2010relative; @vanderlee2017food].
-To add to the distress of vulnerable households, non-pharmaceutical interventions during the pandemic involving restrictions in mobility increased the friction of travel, in particular by transit on which they are more likely to be reliant [e.g., @deweese2020tale]; while at the same time creating stress for the operators of foodbanks through disruptions in the supply chain [e.g., @mckay2021exploring], in addition to concerns around delivery of service in safe conditions.
+While food banks are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least they can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
+As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that food banks in Hamilton, Ontario, recorded the highest number of visitors in the past 29 years; the number of children visiting food banks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278 the year before.
 
-For this study, we aim to look at how the landscape of food bank services available in Hamilton, Ontario, has changed before and during the pandemic.
+It is known that the urban food environment, within which people make their daily food choices, is essential in influencing eating behaviours and health outcomes, based on factors such as food availability, ease of geographic accessibility and socio-demographic variations [@paez2010relative; @vanderlee2017food; @widener2018spatial].
+However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to food banks.
+Although previous work has explored differences in _accessing_ food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource (e.g.
+Enns et al.
+[-@enns2020experiences]), we are not aware of any research that has focused on the capturing the geographic component of accessibility.
+
+The study of place-based geographic accessibility is concerned with capturing the potential to reach destinations of value using the transportation network [@paez2012measuring].
+Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities^[https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with food bank service delivery in Vancouver [@holmes2018nothing].
+However, as with research into the prevalence of food deserts, accessibiltiy to food banks is likely not evenly distributed throughout a city with variations in access attributable to transportation network characteristics and the spatial distribution of food bank locations.
+Furthermore, policy responses to the COVID-19 pandemic add to the distress of vulnerable households.
+Non-pharmaceutical interventions during the pandemic involving restrictions in mobility increased the friction of travel, in particular by transit on which they are more likely to be reliant [e.g., @deweese2020tale].
+At the same time, the pandemic has created additional stress for the operators of foodbanks through disruptions in the supply chain [e.g., @mckay2021exploring] as well as concerns around the delivery of service in safe conditions and possible cancellation of food service programs.
+
+For this study, we aim to look at how the landscape of food banks and related services (e.g.
+low-cost or free meal service providers) available in Hamilton, Ontario, has changed before and during the pandemic.
 Have the number of open food bank services diminished?
@@
 And finally, who are most likely to have been impacted by changes to the accessibility landscape?
-This paper will first look at the distribution of foodbanks and related services before and during the pandemic.
+This paper will first look at the distribution of food banks and related services before and during the pandemic.
 Then, we use the balanced floating catchment areas approach of Paez et al.
@@
 Previous studies have aimed to look at the relationship between the built food environment and sociodemographic characteristics with qualitative and quantitative, or a mixed -method approach.
+
 Quantitatively, official governmental surveys have been able to assist with data, such as the Household Food Security Survey Module (HFSSM), the Canadian Community Health Surveys (CCHS) , the Longitudinal and International Study of Adults (LISA), and official classifications determined by Health Canada in relation to  sociodemographic variables (El-hajj and Benhin 2021; Gundersen et al.

### 2021-04-20 00-Data-Processing/00-Data-Processing-for-Food-Banks-Accessibility.Rmd
Push changes to nb04
Update to include auto passengers

@@
 
-COVIDClose is an indicator variable for whether this bank closed during the pandemic. Convert to factor:
+`STATUS` corresponds to the food bank's status during the pandemic.
+Let's simplify it to a yes-no factor for whether the service remained open or closed after COVID began and then filter out inapplicable services:
 ```{r}
@@
   dplyr::filter(Mode == "Auto driver" | 
+                  Mode == "Auto passenger" |
                   Mode == "Transit (Excluding GO rail)" |
@@
               values_fill = 0) %>%
-  rename(Driver = `Auto driver`, 
-         Transit = `Transit (Excluding GO rail)`)
+  mutate(Driver = `Auto driver` + `Auto passenger`) %>%
+  rename(Transit = `Transit (Excluding GO rail)`) %>%
+  select(-c(`Auto driver`, `Auto passenger`))
 ```

### 2021-04-20 00-Data-Processing/04-Data-Analysis.Rmd
Push changes to nb04
Update to include auto passengers

@@
 
-## TEMP: REMOVE INAPPLICABLE DESTINATIONS UNTIL NEW TTMs ARE FINISHED
+## Remove entries in travel time matrices for inapplicable destinations
 
+As a first step, we will remove the computed travel times to the three Community Kitchen locations (IDs 1-3):
 ```{r}
 ttm_car <- ttm_car %>%
-  filter(!between(OBJECTID, left = 59, right = 69))
+  filter(!between(OBJECTID, left = 1, right = 3))
 
 ttm_transit <- ttm_transit %>%
-  filter(!between(OBJECTID, left = 59, right = 69))
+  filter(!between(OBJECTID, left = 1, right = 3))
 
 ttm_walk <- ttm_walk %>%
-  filter(!between(OBJECTID, left = 59, right = 69))
+  filter(!between(OBJECTID, left = 1, right = 3))
 ```
@@
 
-Half of all parcels can reach 52 foodbanks or more in 180 min or less.
+Half of all parcels can reach 49 foodbanks or more in 180 min or less.
 
@@
 
-Half of all parcels can reach 45 foodbanks or more in 180 min or less.
+Half of all parcels can reach 42 foodbanks or more in 180 min or less.
 
@@
 
+```{r}
+# join tripmaking modal proportions
+ttm_car <- ttm_car %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+
+ttm_transit <- ttm_transit %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+
+ttm_walk <- ttm_walk %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+```
+
 Join tables:
@@
 ```{r}
-ttm <- ttm %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
+#ttm <- ttm %>%
+#  left_join(modes %>%
+#              st_drop_geometry() %>%
+#              transmute(TAZUID,
+#                        p_car,
+#                        p_transit,
+#                        p_walk),
+#            by = "TAZUID")
 ```
@@
 
+```{r}
+access_by_da <- accessibility %>%
+  st_drop_geometry() %>%
+  # Copy the DA identifier to the accessibility table
+  left_join(dwelling_network_points_2016 %>%
+              st_drop_geometry() %>%
+              transmute(UID = ID, DAUID),
+            by = "UID") %>%
+  # Group by DA and time
+  group_by(DAUID, timing) %>%
+  # Calculate the accessibility by DA before and during covid
+  summarize(accessibility = sum(accessibility),
+            timing = first(timing),
+            .groups = "drop") %>%
+  # Pivot wider to calculate the differences in accessibility from pre-covid to covid
+  pivot_wider(names_from = timing, 
+              values_from = accessibility) %>%
+  # Calculate changes in accessibility
+  mutate(acc_d = `During covid` / `Pre-covid` - 1) %>%
+  # Join DA geometry
+  left_join(data_da_2016 %>%
+            transmute(DAUID = GeoUID,
+                      geometry),
+          by = "DAUID") %>%
+  # Convert to simple features
+  st_as_sf() 
+tmap_mode("view")
+tm_shape(access_by_da) + 
+  tm_fill("acc_d")
+```
+
 ## Further Breakdown by Population Characteristics
@@
 ### CODE TO BE UPDATED
-```{r}
-# join tripmaking modal proportions
-ttm_car <- ttm_car %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-
-ttm_transit <- ttm_transit %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-
-ttm_walk <- ttm_walk %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-```
-
 Get minimum travel times in the pre-covid scenario:
@@
 # retrieve minimum travel time during covid to sites that didnt close
-min_ttm_car_pre <- ttm_car %>%
+min_ttm_car_baseline <- ttm_car %>%
   #dplyr::filter(COVIDClose == "No") %>%
@@
             p_walk = first(p_walk),
-            .groups = "drop")
+            .groups = "drop") %>%
 
-min_ttm_transit_pre <- ttm_transit %>%
+  ## Calculate weighted travel time:
+  mutate(weighted_travel_time = travel_time * p_car)
+
+min_ttm_transit_baseline <- ttm_transit %>%
   #dplyr::filter(COVIDClose == "No") %>%
@@
             p_walk = first(p_walk),
-            .groups = "drop")
+            .groups = "drop") %>%
 
-min_ttm_walk_pre <- ttm_walk %>%
+  ## Calculate weighted travel time:
+  mutate(weighted_travel_time = travel_time * p_transit)
+
+min_ttm_walk_baseline <- ttm_walk %>%
   #dplyr::filter(COVIDClose == "No") %>%
@@
             p_walk = first(p_walk),
-            .groups = "drop")
+            .groups = "drop") %>%
+
+  ## Calculate weighted travel time:
+  mutate(weighted_travel_time = travel_time * p_walk)
 ```
 
+```{r weighted-travel-time-by-mode-baseline, include=FALSE}
+# Baseline analysis (pre-COVID)
+## Bind the three tables and label them by mode:
+min_ttm_mode_baseline <- rbind(data.frame(min_ttm_car_baseline, Mode = "Car"),
+                               data.frame(min_ttm_transit_baseline, Mode = "Transit"),
+                               data.frame(min_ttm_walk_baseline, Mode = "Walking"))
+```
+
+```{r aggregate-weighted-travel-times-baseline, include=FALSE}
+# Baseline analysis (PILOT LOCATIONS ONLY)
+## Aggregate weighted travel times by mode:
+min_ttm_baseline <- min_ttm_mode_baseline %>%
+  group_by(UID) %>%
+  summarize(DAUID = first(DAUID),
+            TAZUID = first(TAZUID),
+            linc_age_0to2 = first(linc_age_0to2),
+            linc_age_0to19 = first(linc_age_0to19),
+            linc_age_65plus = first(linc_age_0to19),
+            weighted_travel_time = sum(weighted_travel_time)) %>%
+
+  ## Calculate person-hours of travel:
+  mutate(weighted_person_hours_0to2 = linc_age_0to2 * weighted_travel_time/60)
+```
+
+```{r summarize-results-by-taz-baseline, include=FALSE}
+# Baseline analysis (PILOT LOCATIONS ONLY)
+## Summarize by TAZ and join geometry:
+min_ttm_baseline_taz <- min_ttm_baseline %>%
+  group_by(TAZUID) %>%
+  summarize(weighted_travel_time = mean(weighted_travel_time),
+            weighted_person_hours_0to2 = sum(weighted_person_hours_0to2)) %>%
+  left_join(modes %>%
+              dplyr::select(TAZUID),
+            by = "TAZUID") %>%
+  st_as_sf()
+```
+
+```{r weighted-travel-time-map-baseline}
+ttm_map_baseline <- ggplot() +
+  geom_sf(data = min_ttm_baseline_taz,
+          aes(fill = weighted_travel_time),
+          color = NA) +
+  scale_fill_distiller(name = "Expected travel time (min)",
+                       palette = "OrRd", 
+                       direction = 1) +
+  theme_minimal() +
+  theme(panel.grid = element_blank(),
+        axis.text = element_text(size = 7),
+        legend.position = "bottom",
+        legend.text = element_text(size = 8))
+
+ttm_map_baseline
+```
+
+```{r}
+tmap_mode("view")
+tm_shape(min_ttm_baseline_taz) +
+  tm_fill(col = "weighted_travel_time")
+
+tm_shape(modes) +
+  tm_fill(col = "p_walk")
+```
+
+```{r person-hours-traveled-map-baseline}
+# Baseline analysis (PILOT LOCATIONS ONLY)
+## Map weighted person-hours of travel:
+pht_map_baseline <- ggplot() +
+  geom_sf(data = min_ttm_baseline_taz,
+          aes(fill = weighted_person_hours_0to2),
+          color = NA) +
+  scale_fill_distiller(name = "Person-hours of travel",
+                       palette = "OrRd", 
+                       direction = 1) +
+  #geom_sf(data = urban_types,
+  #        aes(color = Type),
+  #        fill = NA) +
+  #scale_color_manual(name = "",
+  #                   values = c("Urban" = "firebrick2",
+  #                              "Suburban" = "cornflowerblue",
+  #                              "Rural" = "forestgreen"))  +
+  theme_minimal() +
+  theme(panel.grid = element_blank(),
+        axis.text = element_text(size = 7),
+        legend.position = "bottom",
+        legend.text = element_text(size = 8))
+
+pht_map_baseline
+```
+
+```{r join-income-information-baseline, include=FALSE}
+# Baseline analysis (PILOT LOCATIONS ONLY)
+## Join median household information to table:
+min_ttm_baseline <- min_ttm_baseline %>%
+  left_join(data_da_2016 %>%
+              st_drop_geometry() %>%
+              dplyr::transmute(DAUID = GeoUID,
+                               income_quintile),
+            by = "DAUID")
+```
+
+```{r plot-distribution-travel-time-by-income-baseline}
+# Baseline analysis (PILOT LOCATIONS ONLY)
+## Distribution of weighted travel time by income:
+ttm_income_baseline <- ggplot(data = min_ttm_baseline, 
+       aes(x = weighted_travel_time, 
+           color= income_quintile)) +
+  geom_density(size = 0.5, adjust = 2.5) +
+  labs(x = "Travel time (in minutes; weighted by mode)") +
+  scale_color_manual(values = c("Top 20%" = "blue",
+                                "Second 20%" = "steelblue1",
+                                "Third 20%" = "gray40", 
+                                "Fourth 20%" = "tomato",
+                                "Bottom 20%" = "red")) +
+  theme_minimal() +
+  theme(text = element_text(size = 7),
+        legend.text = element_text(size = 7))
+
+ttm_income_baseline
+```
+
+```{r total-weighted-person-hours-by-group-baseline, include=FALSE}
+# Baseline analysis (PILOT LOCATIONS ONLY)
+## Total weighted person-hours by income:
+person_hours_income_baseline <- min_ttm_baseline %>%
+  group_by(income_quintile) %>%
+  summarize(total_weighted_person_hours = sum(weighted_person_hours),
+            total_population = sum(Population),
+            .groups = "drop") %>%
+  mutate(ratio = total_weighted_person_hours/total_population) %>%
+  rename(group = income_quintile)
+
+## Total weighted person-hours by region:
+person_hours_region_baseline <- min_ttm %>%
+  group_by(Urban_type) %>%
+  summarize(total_weighted_person_hours = sum(weighted_person_hours),
+            total_population = sum(Population),
+            .groups = "drop") %>%
+  drop_na() %>%
+  mutate(ratio = total_weighted_person_hours/total_population) %>%
+  rename(group = Urban_type)
+
+person_hours_baseline <- rbind(person_hours_income_baseline,
+                               person_hours_region_baseline)
+```
+
+### Do Again for COVID Scenario
 Get minimum travel times in the during covid scenario:

### 2021-04-20 00-Data-Processing/00-Data-Processing-for-Food-Banks-Accessibility.Rmd
Added data files from TTS with cross-tab of age by income and edited the data processing to read them

--- 2021-04-20 00-Data-Processing/00-Data-Processing-for-Food-Banks-Accessibility.Rmd
@@
 
-I downloaded a cross-tabulation of travel mode by household income by traffic analysis zone. Read the shape file with the traffic analysis zone boundaries:
+We downloaded a cross-tabulation of travel mode by household income by traffic analysis zone. We also downloaded a cross-tabulation of age of person by household income by traffic analysis zone.
+
+Read the shape file with the traffic analysis zone boundaries:
 ```{r}
@@
 
+Read population by age by income:
+```{r}
+age_income <- read_csv("input-data-files/ttsCross65713.csv",
+                         col_types = cols(gta06_hhld = col_double(),
+                                          age = col_double(),
+                                          population = col_double(),
+                                          income = col_double()
+                                          ))
+```
+
 Code | Income
@@
 
+Filter low income:
+```{r}
+age_income <- age_income %>%
+  dplyr::filter(income <= 2)
+```
+
 ### Income less than $40,000
 
-Filter age:
+Filter income:
 ```{r}
@@
             .groups = "drop")
+```
+
+Summarize population by age by income by traffic analysis zone. Children:
+```{r}
+low_income_children <- age_income %>%
+  dplyr::filter(age <= 18) %>%
+  rename(TAZUID = gta06_hhld) %>%
+  group_by(TAZUID) %>%
+  summarize(children = sum(population),
+            .groups = "drop")
+```
+
+Adults:
+```{r}
+low_income_adults <- age_income %>%
+  dplyr::filter(age > 18 & age < 65) %>%
+  rename(TAZUID = gta06_hhld) %>%
+  group_by(TAZUID) %>%
+  summarize(adults = sum(population),
+            .groups = "drop")
+```
+
+Seniors:
+```{r}
+low_income_seniors <- age_income %>%
+  dplyr::filter(age >= 65) %>%
+  rename(TAZUID = gta06_hhld) %>%
+  group_by(TAZUID) %>%
+  summarize(seniors = sum(population),
+            .groups = "drop")
+```
+
+Join tables with population by age and add geometry:
+```{r}
+population_age_income <- tts_taz %>%
+  transmute(TAZUID = gta06, 
+            geometry) %>%
+  left_join(low_income_children,
+            by = "TAZUID") %>%
+  left_join(low_income_adults,
+            by = "TAZUID") %>%
+  left_join(low_income_seniors,
+            by = "TAZUID") %>%
+  mutate(children = replace_na(children, 0),
+         adults = replace_na(adults, 0),
+         seniors = replace_na(seniors, 0))
 ```
@@
 
+## Before saving
+
 Ensure that parcel ids are unique:
@@
 save(dwelling_network_points_2016, file = "output-data-files/dwelling_network_points_2016.RData", compress = "xz")
+save(foodbank_locations, file = "output-data-files/foodbank_locations.RData", compress = "xz")
 save(hamilton_cma, file = "output-data-files/hamilton_cma.RData", compress = "xz")
 save(modes_less40k, file = "output-data-files/modes_less40k.RData", compress = "xz")
-save(foodbank_locations, file = "output-data-files/foodbank_locations.RData", compress = "xz")
+save(population_age_income, file = "output-data-files/population_age_income.RData", compress = "xz")
 save(urban_types, file = "output-data-files/urban_types.RData", compress = "xz")

### 2021-04-20 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Added data files from TTS with cross-tab of age by income and edited the data processing to read them

@@
 However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to food banks.
-Although previous work has explored differences in _accessing_ food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource (e.g.
-Enns et al.
-[-@enns2020experiences]), we are not aware of any research that has focused on the capturing the geographic component of accessibility.
+Although previous work has explored differences in _accessing_ food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource [e.g., @enns2020experiences], we are not aware of any research that has focused on the geographic component of accessibility.
 
@@
 Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities^[https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with food bank service delivery in Vancouver [@holmes2018nothing].
-However, as with research into the prevalence of food deserts, accessibiltiy to food banks is likely not evenly distributed throughout a city with variations in access attributable to transportation network characteristics and the spatial distribution of food bank locations.
+However, as with research into the prevalence of food deserts, accessibility to food banks is likely not evenly distributed throughout a city with variations in access attributable to transportation network characteristics and the spatial distribution of food bank locations - as well as the population that likely needs them.
 Furthermore, policy responses to the COVID-19 pandemic add to the distress of vulnerable households.
-Non-pharmaceutical interventions during the pandemic involving restrictions in mobility increased the friction of travel, in particular by transit on which they are more likely to be reliant [e.g., @deweese2020tale].
-At the same time, the pandemic has created additional stress for the operators of foodbanks through disruptions in the supply chain [e.g., @mckay2021exploring] as well as concerns around the delivery of service in safe conditions and possible cancellation of food service programs.
+Non-pharmaceutical interventions during the pandemic involving restrictions in mobility increased the friction of travel, in particular by transit on which low income populations are more likely to be reliant [e.g., @deweese2020tale].
+At the same time, the pandemic has created additional stress for the operators of foodbanks through disruptions in the supply chain [e.g., @mckay2021exploring] as well as concerns surrounding the delivery of service in safe conditions and possible cancellation of food service programs.
 
@@
 This nutrient deficiency has been causing major health concerns in Canadians, and particularly those who are at a socioeconomic disadvantage (Bazerghi, McKay, and Dunn 2016).
-Previous studies have aimed to look at the relationship between the built food environment and sociodemographic characteristics with qualitative and quantitative, or a mixed -method approach.
+Previous studies have aimed to look at the relationship between the built food environment and socio-demographic characteristics with qualitative and quantitative, or a mixed -method approach.
 
@@
 
-Methods
+Data and Methods
 ===================
 
-Words go here.
+## Data
 

### 2021-04-20 00-Data-Processing/04-Data-Analysis.Rmd

@@
 
-```{r}
+Let's do some view maps to see if the results make sense
+```{r message = FALSE}
 access_by_da <- accessibility %>%
@@
           by = "DAUID") %>%
+
   # Convert to simple features
@@
 tm_shape(access_by_da) + 
-  tm_fill("acc_d")
+  tm_fill("acc_d") +
+  tm_shape(foodbank_locations) + 
+  tm_dots("COVIDClose")
+
+d_access_by_taz <- accessibility %>%
+  st_drop_geometry() %>%
+  # Copy the DA identifier to the accessibility table
+  left_join(dwelling_network_points_2016 %>%
+              st_drop_geometry() %>%
+              transmute(UID = ID, TAZUID),
+            by = "UID") %>%
+  # Group by DA and time
+  group_by(TAZUID, timing) %>%
+  # Calculate the accessibility by DA before and during covid
+  summarize(accessibility = sum(accessibility),
+            timing = first(timing),
+            .groups = "drop") %>%
+  # Pivot wider to calculate the differences in accessibility from pre-covid to covid
+  pivot_wider(names_from = timing, 
+              values_from = accessibility) %>%
+  # Calculate changes in accessibility
+  mutate(acc_d = `During covid` / `Pre-covid` - 1) %>%
+  # join taz geometry
+  left_join(tts_taz %>%
+            transmute(TAZUID = gta06,
+                      geometry),
+          by = "TAZUID") %>%
+
+  # Convert to simple features
+  st_as_sf()
+
+tm_shape(d_access_by_taz) + 
+  tm_fill("acc_d") + 
+  tm_shape(foodbank_locations) + 
+  tm_dots("COVIDClose")
+
+access_by_taz <- accessibility %>%
+  st_drop_geometry() %>%
+  # Copy the DA identifier to the accessibility table
+  left_join(dwelling_network_points_2016 %>%
+              st_drop_geometry() %>%
+              transmute(UID = ID, TAZUID),
+            by = "UID") %>%
+  # Group by DA and time
+  group_by(TAZUID, timing) %>%
+  # Calculate the accessibility by DA before and during covid
+  summarize(accessibility = sum(accessibility),
+            timing = first(timing),
+            .groups = "drop") %>%
+  # Pivot wider to calculate the differences in accessibility from pre-covid to covid
+  pivot_wider(names_from = timing, 
+              values_from = accessibility) %>%
+  # Calculate changes in accessibility
+  mutate(acc_pre = `Pre-covid`) %>%
+  # join taz geometry
+  left_join(tts_taz %>%
+            transmute(TAZUID = gta06,
+                      geometry),
+          by = "TAZUID") %>%
+
+  # Convert to simple features
+  st_as_sf()
+
+tm_shape(access_by_taz) + 
+  tm_fill("acc_pre") + 
+  tm_shape(foodbank_locations) + 
+  tm_dots("COVIDClose")
+
 ```

### 2021-04-20 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
First draft of methods section and minor edits in the introduction

@@
 Although previous work has explored differences in _accessing_ food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource [e.g., @enns2020experiences], we are not aware of any research that has focused on the geographic component of accessibility.
+**Check @tarasuk2019geographic**
 
@@
 This paper will first look at the distribution of food banks and related services before and during the pandemic.
-Then, we use the balanced floating catchment areas approach of Paez et al.
+Then, we use the balanced floating catchment area approach of Paez et al.
 [@paez2019demand] to investigate the accessibility situation.
@@
 
-Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake (Kirkpatrick and Tarasuk 2008; Tarasuk and Vogt 2009).
-This nutrient deficiency has been causing major health concerns in Canadians, and particularly those who are at a socioeconomic disadvantage (Bazerghi, McKay, and Dunn 2016).
+Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake [@kirkpatrick2008food; @tarasuk2009household].
+This nutrient deficiency has been causing major health concerns in Canadians, and particularly those who are at a socioeconomic disadvantage [@bazerghi2016role].
 Previous studies have aimed to look at the relationship between the built food environment and socio-demographic characteristics with qualitative and quantitative, or a mixed -method approach.
-
-Quantitatively, official governmental surveys have been able to assist with data, such as the Household Food Security Survey Module (HFSSM), the Canadian Community Health Surveys (CCHS) , the Longitudinal and International Study of Adults (LISA), and official classifications determined by Health Canada in relation to  sociodemographic variables (El-hajj and Benhin 2021; Gundersen et al.
-2018; Kirkpatrick and Tarasuk 2008; Tarasuk and Vogt 2009).
-Studies have also aimed to assess food availability of healthy foods (e.g., fruit and vegetables) at supermarkets in relation to sociodemographic characteristics and geographic accessibility (Latham and Moffat 2007).
+Quantitatively, official governmental surveys have been able to assist with data, such as the Household Food Security Survey Module (HFSSM), the Canadian Community Health Surveys (CCHS), the Longitudinal and International Study of Adults (LISA), and official classifications determined by Health Canada in relation to  sociodemographic variables [@kirkpatrick2008food; @tarasuk2009household; @gundersen2018food].
+Studies have also aimed to assess food availability of healthy foods (e.g., fruit and vegetables) at supermarkets in relation to sociodemographic characteristics and geographic accessibility [@latham2007determinants].
 In terms of findings, studies have generated inconsistent relationships between their evaluated availability of food and sociodemographic characteristics that go in hand.
 
-## Food Banks
+## Foodbanks
 
-The number of food banks has been increasing steadily in Canada (Wakefield et al.
-2013).
-What was supposed to be a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, has now developed into a community practice to secure emergency food supplies for those in need (Loopstra and Tarasuk 2012; Wakefield et al.
-2013).
-Food banks were developed as a part of a community response to aide those with inadequate food by voluntarily giving meals and ingredients away (Loopstra and Tarasuk 2012; Riches 2002).
-The scope and objectives of food banks can vary by region and by country.
-These organizations can include not only prepared meals and aliments, but also shared spaces to connect in community gardens and community kitchens (Wakefield et al.
-2013) or referred to as ‘food pantry’ and ‘food shelf’, where many frequent as their primary location to get food (Bazerghi et al.
-2016).
-However, surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada (Tarasuk et al.
-2014).
-
-Previous studies also question if food banks are able to offer a rounded nutritious supply of food, and if food banks are a sustainable practice for those in need to continuously obtain their food from (Bazerghi et al.
-2016; Riches 2002).
+Food banks - sometimes also referred to as ‘food pantries’ and ‘food shelves’ - originated as a community response to aid those with inadequate food by voluntarily offering them meals and ingredients [@riches2002food; @loopstra2012relationship].
+The scope and objectives of foodbanks can vary by region and by country, and these organizations can include not only prepared meals and aliments, but also shared spaces to connect in community gardens and community kitchens [@wakefield2013sweet].
+Although in their origin foodbanks were meant to be provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time they evolved into a community practice to secure emergency food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
+In Canada, the number of foodbanks has steadily increased in the past few decades [@wakefield2013sweet].
+Previous studies have questioned whether food banks are able to offer a rounded nutritious supply of food, and if food banks are a sustainable on-going practice for those in need [@riches2002food; @bazerghi2016role].
+An inescapable reality, at the moment, is that despite their imperfections they provide urgently needed support for members of the population for whom foodbanks are frequently their primary location to get food [@bazerghi2016role].
+**However, surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada [@tarasuk2014food].** **Something about geographical access**
 
@@
 
-Currently with the COVID-19 pandemic, disrupted economies, rising unemployment rates, and alarming poverty levels have disrupted the food environment by causing higher rates of food insecurity (Niles et al.
-2020).
-In 2012, 12.4% of Canadians households and 11.8% Ontarian households experienced some degree of food insecurity (Gundersen et al.
-2018; Tarasuk, Fafard St-Germain, and Mitchell 2019).
-Recent COVID-19 related food security studies in the US have found a massive increase in households experiencing food insecurity for the first time, and also in households experiencing more severe food insecurity than before (Niles et al.
-2020; Wolfson and Leung 2020).
-Most recently in May 2020, Canada recorded 14.7% of its population living in food insecurity in the past 30 days(Statistics Canada 2020).
-Food insecurity is a highly concerning public health issue due the vast health consequences that it can induce.
-Recent publishing suggests that food insecurity in adults could lead to experiencing more stressful events (El-hajj and Benhin 2021).
-These increases in food insecurity rates due to the pandemic, signal a change in the food environment with potential damages to health outcomes in populations (Niles et al.
-2020)
+Currently with the COVID-19 pandemic, disrupted economies, rising unemployment rates, and changes in poverty levels have disrupted the food environment and led to higher rates of food insecurity [@niles2020early].
+In 2012, 12.4% of Canadians households and 11.8% Ontarian households experienced some degree of food insecurity [@gundersen2018food; @tarasuk2020relationship].
+Pandemic-related food security studies in the US have found a substantial increase in households experiencing food insecurity for the first time, and also in households experiencing more severe food insecurity than before [@niles2020early; @wolfson2020food].
+Most recently in May 2020, Canada recorded 14.7% of its population living in food insecurity in the past 30 days [@statisticscanada2020food].
+Food insecurity is a highly concerning public health issue due to its health impacts.
+Recent research suggests that food insecurity in adults can lead to experiencing more stressful events (El-hajj and Benhin 2021).
+Increases in food insecurity rates due to the pandemic, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
 
-Data and Methods
+Methods and Materials
 ===================
 
+## Methods
+
+For the research in this paper we adopt the balanced floating catchment area approach of Paez et al.
+[-@paez2019demand].
+This method for estimating accessibility is a form of the common two-stage floating catchment area method [@radke2000spatial; @luo2003measures].
+Floating catchment areas are used to estimate accessibility when there are potential congestion effects, and operate by calculating first the _demand_ for spatially distributed services.
+The demand (usually the number of people who require a service) is used to calculate a level of service.
+In a second step, the level of service is allocated back to the population.
+Demand and level of service are allocated using some form of distance-decay to embody the geographical principle that, given a choice, people prefer to travel less than more when reaching destinations.
+
+More formally, the first step of this method is as follows:
+$$
+L_j = \frac{S_j}{\sum_{i=1}^nP_iw_{ij}}
+$$
+
+\noindent where $S_j$ is the level of supply at location $j$, in simplest terms whether a service point is present (i.e., $S_j=1$) or not (i.e., $S_j=0$); $P_i$ is the population at location $i$ that demands the service; and $w_{ij}$ is a weight, typically a function of the distance between locations $i$ and $j$.
+$L_j$ is the level of service at location $j$ and it is the inverse of the number of people that need to be serviced.
+
+The second step in this process is then summing the level of service that each population unit can reach, according to the distance-decay weight:
+$$
+A_i = \sum_{j=1}^JL_jw_{ji}
+$$
+
+\noindent where $A_i$ is the accessibility to the service, which is in the same units as the level of service: as the inverse of the population being serviced.
+When the population being serviced is low accessibility is high (i.e., there is little competition for the service), and viceversa.
+
+Floating catchment area methods are prone to overestimation of the population and the level of service due to multiple-counting.
+The population at $P_i$ is allocated to _every_ service point $j$ for which $w_{ij}>0$.
+Similarly, the level of service at $LOS_j$ is allocated to _every_ population point for which $w_{ji}>0$.
+This inflation effect has been known for several years, and several modifications have been proposed to mitigate it [e.g., @wan2012three; @delamater2013spatial].
+A definitive solution to this issue was presented by Paez et al.
+[-@paez2019demand].
+In order to avoid the multiple-counting in the summations, the population and the level of service need to be allocated _proportionally_.
+This is achieved by standardizing the weights as follows:
+$$
+w_{ij}^\text{st} = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
+$$
+\noindent and:
+$$
+w_{ji}^\text{st} = \frac{w_{ji}}{\sum_{j=1}^Jw_{ji}}
+$$
+
+The standardized weights satisfy the following conditions:
+$$
+\sum_{i=1}^nw_{ij}^\text{st}=1
+$$
+
+\noindent and:
+$$
+\sum_{j=1}^Jw_{ji}^\text{st}=1
+$$
+
+As a result, the total population (and consequently, the level of service) are preserved system-wide:
+$$
+\sum_{i=1}^nP_iw_{ij}^\text{st}=P_i
+$$
+
+\noindent and therefore:
+$$
+\sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
+$$
+
 ## Data
+
+Words go here.
 

### 2021-04-20 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
First draft of methods section and minor edits in the introduction

@@
 
-Over at least the past four decades food banks have become an essential
-line of defense against food insecurity in Canadian communities (Black
-and Seto, 2020; Riches, 2002; Tarasuk et al., 2020). In this respect,
-Canada is not unlike numerous other wealthy countries where a systematic
-dismantling of the welfare state took place in the intervening period
-(Tarasuk et al., 2014).
+Food insecurity, defined as an ``inadequate or uncertain access to a
+sufficient quantity and/or adequate quality of food'' due to a
+household's financial limitations (Enns et al., 2020), has been
+associated with reductions in nutritional outcomes (Bhattacharya et al.,
+2004; Kirkpatrick and Tarasuk, 2008; Olson, 1999) and physical and
+mental health in children and adults (Elgar et al., 2021; Jones, 2017;
+Ramsey et al., 2011; Seligman et al., 2010; Stuff et al., 2004). Over at
+least the past four decades food banks have become an essential line of
+defense against food insecurity in Canadian communities (Black and Seto,
+2020; Holmes et al., 2018; Riches, 2002; Tarasuk et al., 2020). In this
+respect, Canada is not unlike numerous other wealthy countries where a
+systematic dismantling of the welfare state took place in the
+intervening period (Tarasuk et al., 2014).
 
-The emergence of COVID-19, the worst public health crisis since the 1918
-flu pandemic, has revealed important social and economic fault lines,
-and pre-existing patterns of inequality appear to have been exacerbated.
-Along several other dimensions of stress {[}e.g., accessibility to
-health care facilities; Pereira et al. (2021){]}, this seems to be the
-case for food insecurity as well (Laborde et al., 2020). In the US, for
-example, it has been estimated that there was an increase of more than
-30\% in household food insecurity, and more than one third of households
-were discovered to be newly food insecure - meaning they did not
-experience food insecurity before the pandemic (Niles et al., 2020). In
-Canada, Men and Tarasuk (2021) report that about 25\% of individuals who
-experienced job insecurity (a relatively common occurrence during the
-pandemic), also experienced food insecurity. Similarly, according to
-Statistics Canada (2020), in the early stages of the pandemic almost
-15\% of individuals reported living in a household that faced food
-insecurity; the risk of food insecurity was substantially higher for
-households with children. The difference between households with and
-without children was significant, and 11.7\% of households with children
-indicated that ``food didn't last and {[}there was{]} no money to get
-more'' sometimes or often, compared to 7.3\% of households without
-children); likewise, 13\% of households with children indicated that
-they ``{[}c{]}ouldn't afford balanced meals'' sometimes or often,
-compared to 8.8\% of households without children.
+But the emergence of COVID-19, the worst public health crisis since the
+1918 flu pandemic, has revealed important social and economic fault
+lines, and pre-existing patterns of inequality appear to have been
+exacerbated. Along several other dimensions of stress {[}e.g.,
+accessibility to health care facilities; Pereira et al. (2021){]}, this
+seems to be the case for food insecurity as well (Laborde et al., 2020).
+In the US, for example, it has been estimated that there was an increase
+of more than 30\% in household food insecurity, and more than one third
+of households were discovered to be newly food insecure - meaning they
+did not experience food insecurity before the pandemic (Niles et al.,
+2020). In Canada, Men and Tarasuk (2021) report that about 25\% of
+individuals who experienced job insecurity (a relatively common
+occurrence during the pandemic), also experienced food insecurity.
+Similarly, according to Statistics Canada (2020), in the early stages of
+the pandemic almost 15\% of individuals reported living in a household
+that faced food insecurity; the risk of food insecurity was
+substantially higher for households with children. The difference
+between households with and without children was significant, and 11.7\%
+of households with children indicated that ``food didn't last and
+{[}there was{]} no money to get more'' sometimes or often, compared to
+7.3\% of households without children); likewise, 13\% of households with
+children indicated that they ``{[}c{]}ouldn't afford balanced meals''
+sometimes or often, compared to 8.8\% of households without children.
 
@@
 were already critical public health concerns in Canada prior to COVID-19
-(Boucher et al., 2017). While foodbanks are not necessarily a stable
+(Boucher et al., 2017). While food banks are not necessarily a stable
 solution to food insecurity and in fact may encourage a retrenchment of
-neoliberal policy (Wakefield et al., 2013), at least can be argued to
-provide a resource of last instance to households in precarious
+neoliberal policy (Wakefield et al., 2013), at least they can be argued
+to provide a resource of last instance to households in precarious
 situations (Bazerghi et al., 2016). As recently as 2019, the Hamilton
@@
 noted that food banks in Hamilton, Ontario, recorded the highest number
-of visitors in the past 29 years; the number of children visiting
-foodbanks (minors up to 18 years old) was 9,125 in March 2019, up from
-8,278 the year before. It is known that the urban food environment,
-within which people make their daily food choices, is essential in
-influencing eating behaviours and health outcomes, based on factors such
-as food availability, ease of accessibility and socio-demographic
-variations (Paez et al., 2010; Vanderlee and L'Abbé, 2017). To add to
-the distress of vulnerable households, non-pharmaceutical interventions
-during the pandemic involving restrictions in mobility increased the
-friction of travel, in particular by transit on which they are more
-likely to be reliant (e.g., DeWeese et al., 2020); while at the same
-time creating stress for the operators of foodbanks through disruptions
-in the supply chain (e.g., McKay et al., 2021), in addition to concerns
-around delivery of service in safe conditions.
+of visitors in the past 29 years; the number of children visiting food
+banks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278
+the year before.
 
-For this study, we aim to look at how the landscape of food bank
-services available in Hamilton, Ontario, has changed before and during
-the pandemic. Have the number of open food bank services diminished? If
-so, what was the accessibility to foodbanks before and during the
-pandemic, from the perspective of low income households? And finally,
-who are most likely to have been impacted by changes to the
-accessibility landscape? This paper will first look at the distribution
-of foodbanks and related services before and during the pandemic. Then,
-we use the balanced floating catchment areas approach of Paez et al.
-(Paez et al., 2019) to investigate the accessibility situation. We use a
-fully disaggregated approach based on parcel-level data. Socio-economic
-and demographic data are drawn from the latest Census of Canada (2016),
-whereas travel information is from the most recent regional travel
-survey from 2016. This paper follows reproducible research
-recommendations (see Brunsdon and Comber, 2020), and the research was
-conducted using open source tools for transportation analysis (Lovelace,
-2021). The code and data necessary to reproduce the analysis are
-available in a public repository\footnote{add repository}.
+It is known that the urban food environment, within which people make
+their daily food choices, is essential in influencing eating behaviours
+and health outcomes, based on factors such as food availability, ease of
+geographic accessibility and socio-demographic variations (Paez et al.,
+2010; Vanderlee and L'Abbé, 2017; Widener, 2018). However, while there
+is a wealth of literature that has examined the topic of geographic
+accessibility to healthy food through the ``food desert'' concept, there
+has been little research into accessibility to food banks. Although
+previous work has explored differences in \emph{accessing} food banks,
+such as how some households utilize food banks over short periods of
+time while others regularly utilize food banks as longer-term resource
+(e.g., Enns et al., 2020), we are not aware of any research that has
+focused on the geographic component of accessibility.
+
+The study of place-based geographic accessibility is concerned with
+capturing the potential to reach destinations of value using the
+transportation network (Páez et al., 2012). Indeed, the Government of
+Canada's recent Food Policy has made ``access'' to healthy food a
+priority for Canadian communities\footnote{https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042}
+and previous survey research has suggested that such accessibility plays
+a key role in user satisfaction with food bank service delivery in
+Vancouver (Holmes et al., 2018). However, as with research into the
+prevalence of food deserts, accessibility to food banks is likely not
+evenly distributed throughout a city with variations in access
+attributable to transportation network characteristics and the spatial
+distribution of food bank locations - as well as the population that
+likely needs them. Furthermore, policy responses to the COVID-19
+pandemic add to the distress of vulnerable households.
+Non-pharmaceutical interventions during the pandemic involving
+restrictions in mobility increased the friction of travel, in particular
+by transit on which low income populations are more likely to be reliant
+(e.g., DeWeese et al., 2020). At the same time, the pandemic has created
+additional stress for the operators of foodbanks through disruptions in
+the supply chain (e.g., McKay et al., 2021) as well as concerns
+surrounding the delivery of service in safe conditions and possible
+cancellation of food service programs.
+
+For this study, we aim to look at how the landscape of food banks and
+related services (e.g.~low-cost or free meal service providers)
+available in Hamilton, Ontario, has changed before and during the
+pandemic. Have the number of open food bank services diminished? If so,
+what was the accessibility to foodbanks before and during the pandemic,
+from the perspective of low income households? And finally, who are most
+likely to have been impacted by changes to the accessibility landscape?
+This paper will first look at the distribution of food banks and related
+services before and during the pandemic. Then, we use the balanced
+floating catchment area approach of Paez et al. (Paez et al., 2019) to
+investigate the accessibility situation. We use a fully disaggregated
+approach based on parcel-level data. Socio-economic and demographic data
+are drawn from the latest Census of Canada (2016), whereas travel
+information is from the most recent regional travel survey from 2016.
+This paper follows reproducible research recommendations (see Brunsdon
+and Comber, 2020), and the research was conducted using open source
+tools for transportation analysis (Lovelace, 2021). The code and data
+necessary to reproduce the analysis are available in a public
+repository\footnote{add repository}.
 
@@
 amount or good quality food, leading to inadequate nutrient intake
-(Kirkpatrick and Tarasuk 2008; Tarasuk and Vogt 2009). This nutrient
+(Kirkpatrick and Tarasuk, 2008; Tarasuk and Vogt, 2009). This nutrient
 deficiency has been causing major health concerns in Canadians, and
-particularly those who are at a socioeconomic disadvantage (Bazerghi,
-McKay, and Dunn 2016). Previous studies have aimed to look at the
-relationship between the built food environment and sociodemographic
-characteristics with qualitative and quantitative, or a mixed -method
-approach. Quantitatively, official governmental surveys have been able
-to assist with data, such as the Household Food Security Survey Module
-(HFSSM), the Canadian Community Health Surveys (CCHS) , the Longitudinal
-and International Study of Adults (LISA), and official classifications
+particularly those who are at a socioeconomic disadvantage (Bazerghi et
+al., 2016). Previous studies have aimed to look at the relationship
+between the built food environment and socio-demographic characteristics
+with qualitative and quantitative, or a mixed -method approach.
+Quantitatively, official governmental surveys have been able to assist
+with data, such as the Household Food Security Survey Module (HFSSM),
+the Canadian Community Health Surveys (CCHS), the Longitudinal and
+International Study of Adults (LISA), and official classifications
 determined by Health Canada in relation to sociodemographic variables
-(El-hajj and Benhin 2021; Gundersen et al.~2018; Kirkpatrick and Tarasuk
-2008; Tarasuk and Vogt 2009). Studies have also aimed to assess food
-availability of healthy foods (e.g., fruit and vegetables) at
-supermarkets in relation to sociodemographic characteristics and
-geographic accessibility (Latham and Moffat 2007). In terms of findings,
-studies have generated inconsistent relationships between their
-evaluated availability of food and sociodemographic characteristics that
-go in hand.
+(Gundersen et al., 2018; Kirkpatrick and Tarasuk, 2008; Tarasuk and
+Vogt, 2009). Studies have also aimed to assess food availability of
+healthy foods (e.g., fruit and vegetables) at supermarkets in relation
+to sociodemographic characteristics and geographic accessibility (Latham
+and Moffat, 2007). In terms of findings, studies have generated
+inconsistent relationships between their evaluated availability of food
+and sociodemographic characteristics that go in hand.
 
-\hypertarget{food-banks}{%
-\subsection{Food Banks}\label{food-banks}}
+\hypertarget{foodbanks}{%
+\subsection{Foodbanks}\label{foodbanks}}
 
-The number of food banks has been increasing steadily in Canada
-(Wakefield et al.~2013). What was supposed to be a temporary solution to
-accommodate those in hunger due to job retrenchments and economic
-downfalls since the 1980s, has now developed into a community practice
-to secure emergency food supplies for those in need (Loopstra and
-Tarasuk 2012; Wakefield et al.~2013). Food banks were developed as a
-part of a community response to aide those with inadequate food by
-voluntarily giving meals and ingredients away (Loopstra and Tarasuk
-2012; Riches 2002). The scope and objectives of food banks can vary by
-region and by country. These organizations can include not only prepared
-meals and aliments, but also shared spaces to connect in community
-gardens and community kitchens (Wakefield et al.~2013) or referred to as
-`food pantry' and `food shelf,' where many frequent as their primary
-location to get food (Bazerghi et al.~2016). However, surveys revealed
-that only 20 to 30 percent of those experiencing food insecurity were
-found to frequent food banks in Canada (Tarasuk et al.~2014).
-
-Previous studies also question if food banks are able to offer a rounded
-nutritious supply of food, and if food banks are a sustainable practice
-for those in need to continuously obtain their food from (Bazerghi et
-al.~2016; Riches 2002).
+Food banks - sometimes also referred to as `food pantries' and `food
+shelves' - originated as a community response to aid those with
+inadequate food by voluntarily offering them meals and ingredients
+(Loopstra and Tarasuk, 2012; Riches, 2002). The scope and objectives of
+foodbanks can vary by region and by country, and these organizations can
+include not only prepared meals and aliments, but also shared spaces to
+connect in community gardens and community kitchens (Wakefield et al.,
+2013). Although in their origin foodbanks were meant to be provide a
+temporary solution to accommodate those in hunger due to job
+retrenchments and economic downfalls since the 1980s, over time they
+evolved into a community practice to secure emergency food supplies for
+those in need (Loopstra and Tarasuk, 2012; Wakefield et al., 2013). In
+Canada, the number of foodbanks has steadily increased in the past few
+decades (Wakefield et al., 2013). Previous studies have questioned
+whether food banks are able to offer a rounded nutritious supply of
+food, and if food banks are a sustainable on-going practice for those in
+need (Bazerghi et al., 2016; Riches, 2002). An inescapable reality, at
+the moment, is that despite their imperfections they provide urgently
+needed support for members of the population for whom foodbanks are
+frequently their primary location to get food (Bazerghi et al., 2016).
+\textbf{However, surveys revealed that only 20 to 30 percent of those
+experiencing food insecurity were found to frequent food banks in Canada
+(Tarasuk et al., 2014).} \textbf{Something about geographical access}
 
@@
 Currently with the COVID-19 pandemic, disrupted economies, rising
-unemployment rates, and alarming poverty levels have disrupted the food
-environment by causing higher rates of food insecurity (Niles et
-al.~2020). In 2012, 12.4\% of Canadians households and 11.8\% Ontarian
-households experienced some degree of food insecurity (Gundersen et
-al.~2018; Tarasuk, Fafard St-Germain, and Mitchell 2019). Recent
-COVID-19 related food security studies in the US have found a massive
-increase in households experiencing food insecurity for the first time,
-and also in households experiencing more severe food insecurity than
-before (Niles et al.~2020; Wolfson and Leung 2020). Most recently in May
-2020, Canada recorded 14.7\% of its population living in food insecurity
-in the past 30 days(Statistics Canada 2020). Food insecurity is a highly
-concerning public health issue due the vast health consequences that it
-can induce. Recent publishing suggests that food insecurity in adults
-could lead to experiencing more stressful events (El-hajj and Benhin
-2021). These increases in food insecurity rates due to the pandemic,
-signal a change in the food environment with potential damages to health
-outcomes in populations (Niles et al.~2020)
+unemployment rates, and changes in poverty levels have disrupted the
+food environment and led to higher rates of food insecurity (Niles et
+al., 2020). In 2012, 12.4\% of Canadians households and 11.8\% Ontarian
+households experienced some degree of food insecurity (Gundersen et al.,
+2018; Tarasuk et al., 2020). Pandemic-related food security studies in
+the US have found a substantial increase in households experiencing food
+insecurity for the first time, and also in households experiencing more
+severe food insecurity than before (Niles et al., 2020; Wolfson and
+Leung, 2020). Most recently in May 2020, Canada recorded 14.7\% of its
+population living in food insecurity in the past 30 days (Statistics
+Canada, 2020). Food insecurity is a highly concerning public health
+issue due to its health impacts. Recent research suggests that food
+insecurity in adults can lead to experiencing more stressful events
+(El-hajj and Benhin 2021). Increases in food insecurity rates due to the
+pandemic, signal a change in the food environment with potential damages
+to population health during the course of the pandemic and beyond (Niles
+et al., 2020).
+
+\hypertarget{methods-and-materials}{%
+\section{Methods and Materials}\label{methods-and-materials}}
 
 \hypertarget{methods}{%
-\section{Methods}\label{methods}}
+\subsection{Methods}\label{methods}}
+
+For the research in this paper we adopt the balanced floating catchment
+area approach of Paez et al. (2019). This method for estimating
+accessibility is a form of the common two-stage floating catchment area
+method (Luo and Wang, 2003; Radke and Mu, 2000). Floating catchment
+areas are used to estimate accessibility when there are potential
+congestion effects, and operate by calculating first the \emph{demand}
+for spatially distributed services. The demand (usually the number of
+people who require a service) is used to calculate a level of service.
+In a second step, the level of service is allocated back to the
+population. Demand and level of service are allocated using some form of
+distance-decay to embody the geographical principle that, given a
+choice, people prefer to travel less than more when reaching
+destinations.
+
+More formally, the first step of this method is as follows: \[
+L_j = \frac{S_j}{\sum_{i=1}^nP_iw_{ij}}
+\]
+
+\noindent where \(S_j\) is the level of supply at location \(j\), in
+simplest terms whether a service point is present (i.e., \(S_j=1\)) or
+not (i.e., \(S_j=0\)); \(P_i\) is the population at location \(i\) that
+demands the service; and \(w_{ij}\) is a weight, typically a function of
+the distance between locations \(i\) and \(j\). \(L_j\) is the level of
+service at location \(j\) and it is the inverse of the number of people
+that need to be serviced.
+
+The second step in this process is then summing the level of service
+that each population unit can reach, according to the distance-decay
+weight: \[
+A_i = \sum_{j=1}^JL_jw_{ji}
+\]
+
+\noindent where \(A_i\) is the accessibility to the service, which is in
+the same units as the level of service: as the inverse of the population
+being serviced. When the population being serviced is low accessibility
+is high (i.e., there is little competition for the service), and
+viceversa.
+
+Floating catchment area methods are prone to overestimation of the
+population and the level of service due to multiple-counting. The
+population at \(P_i\) is allocated to \emph{every} service point \(j\)
+for which \(w_{ij}>0\). Similarly, the level of service at \(LOS_j\) is
+allocated to \emph{every} population point for which \(w_{ji}>0\). This
+inflation effect has been known for several years, and several
+modifications have been proposed to mitigate it (Delamater, 2013; e.g.,
+Wan et al., 2012). A definitive solution to this issue was presented by
+Paez et al. (2019). In order to avoid the multiple-counting in the
+summations, the population and the level of service need to be allocated
+\emph{proportionally}. This is achieved by standardizing the weights as
+follows: \[
+w_{ij}^\text{st} = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
+\] \noindent and: \[
+w_{ji}^\text{st} = \frac{w_{ji}}{\sum_{j=1}^Jw_{ji}}
+\]
+
+The standardized weights satisfy the following conditions: \[
+\sum_{i=1}^nw_{ij}^\text{st}=1
+\]
+
+\noindent and: \[
+\sum_{j=1}^Jw_{ji}^\text{st}=1
+\]
+
+As a result, the total population (and consequently, the level of
+service) are preserved system-wide: \[
+\sum_{i=1}^nP_iw_{ij}^\text{st}=P_i
+\]
+
+\noindent and therefore: \[
+\sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
+\]
+
+\hypertarget{data}{%
+\subsection{Data}\label{data}}
 
@@
 
+\leavevmode\hypertarget{ref-bhattacharya2004poverty}{}%
+Bhattacharya, J., Currie, J., Haider, S., 2004. Poverty, food
+insecurity, and nutritional outcomes in children and adults. Journal of
+health economics 23, 839--862.
+doi:\url{https://doi.org/10.1016/j.jhealeco.2003.12.008}
+
 \leavevmode\hypertarget{ref-black2020examining}{}%
@@
 
+\leavevmode\hypertarget{ref-delamater2013spatial}{}%
+Delamater, P.L., 2013. Spatial accessibility in suboptimally configured
+health care systems: A modified two-step floating catchment area
+(M2SFCA) metric. Health \& Place 24, 30--43.
+doi:\href{https://doi.org/10.1016/j.healthplace.2013.07.012}{10.1016/j.healthplace.2013.07.012}
+
 \leavevmode\hypertarget{ref-deweese2020tale}{}%
@@
 
+\leavevmode\hypertarget{ref-elgar2021relative}{}%
+Elgar, F.J., Pickett, W., Pförtner, T.-K., Gariépy, G., Gordon, D.,
+Georgiades, K., Davison, C., Hammami, N., MacNeil, A.H., Da Silva, M.A.,
+others, 2021. Relative food insecurity, mental health and wellbeing in
+160 countries. Social Science \& Medicine 268, 113556.
+doi:\url{https://doi.org/10.1016/j.socscimed.2020.113556}
+
+\leavevmode\hypertarget{ref-enns2020experiences}{}%
+Enns, A., Rizvi, A., Quinn, S., Kristjansson, E., 2020. Experiences of
+food bank access and food insecurity in ottawa, canada. Journal of
+Hunger \& Environmental Nutrition 15, 456--472.
+doi:\url{https://doi.org/10.1080/19320248.2020.1761502}
+
+\leavevmode\hypertarget{ref-gundersen2018food}{}%
+Gundersen, C., Tarasuk, V., Cheng, J., De Oliveira, C., Kurdyak, P.,
+2018. Food insecurity status and mortality among adults in ontario,
+canada. PloS one 13, e0202642.
+
+\leavevmode\hypertarget{ref-holmes2018nothing}{}%
+Holmes, E., Black, J.L., Heckelman, A., Lear, S.A., Seto, D., Fowokan,
+A., Wittman, H., 2018. {``Nothing is going to change three months from
+now''}: A mixed methods characterization of food bank use in greater
+vancouver. Social Science \& Medicine 200, 129--136.
+doi:\url{https://doi.org/10.1016/j.socscimed.2018.01.029}
+
+\leavevmode\hypertarget{ref-jones2017food}{}%
+Jones, A.D., 2017. Food insecurity and mental health status: A global
+analysis of 149 countries. American journal of preventive medicine 53,
+264--273. doi:\url{https://doi.org/10.1016/j.amepre.2017.04.008}
+
+\leavevmode\hypertarget{ref-kirkpatrick2008food}{}%
+Kirkpatrick, S.I., Tarasuk, V., 2008. Food insecurity is associated with
+nutrient inadequacies among canadian adults and adolescents. The Journal
+of nutrition 138, 604--612.
+doi:\url{https://doi.org/10.1093/jn/138.7.1399}
+
 \leavevmode\hypertarget{ref-laborde2020poverty}{}%
@@
 
+\leavevmode\hypertarget{ref-latham2007determinants}{}%
+Latham, J., Moffat, T., 2007. Determinants of variation in food cost and
+availability in two socioeconomically contrasting neighbourhoods of
+hamilton, ontario, canada. Health \& Place 13, 273--287.
+doi:\url{https://doi.org/10.1016/j.healthplace.2006.01.006}
+
+\leavevmode\hypertarget{ref-loopstra2012relationship}{}%
+Loopstra, R., Tarasuk, V., 2012. The relationship between food banks and
+household food insecurity among low-income toronto families. Canadian
+Public Policy 38, 497--514.
+doi:\href{https://doi.org/10.3138/cpp.38.4.497}{10.3138/cpp.38.4.497}
+
 \leavevmode\hypertarget{ref-lovelace2021open}{}%
@@
 doi:\href{https://doi.org/10.1007/s10109-020-00342-2}{10.1007/s10109-020-00342-2}
+
+\leavevmode\hypertarget{ref-luo2003measures}{}%
+Luo, W., Wang, F.H., 2003. Measures of spatial accessibility to health
+care in a GIS environment: Synthesis and a case study in the chicago
+region. Environment and Planning B-Planning \& Design 30, 865--884.
 
@@
 
+\leavevmode\hypertarget{ref-olson1999nutrition}{}%
+Olson, C.M., 1999. Nutrition and health outcomes associated with food
+insecurity and hunger. The Journal of nutrition 129, 521S--524S.
+doi:\url{https://doi.org/10.1093/jn/129.2.521S}
+
 \leavevmode\hypertarget{ref-paez2019demand}{}%
@@
 
+\leavevmode\hypertarget{ref-paez2012measuring}{}%
+Páez, A., Scott, D.M., Morency, C., 2012. Measuring accessibility:
+Positive and normative implementations of various accessibility
+indicators. Journal of Transport Geography 25, 141--153.
+doi:\url{https://doi.org/10.1016/j.jtrangeo.2012.03.016}
+
 \leavevmode\hypertarget{ref-pereira2021geographic}{}%
@@
 
+\leavevmode\hypertarget{ref-radke2000spatial}{}%
+Radke, J., Mu, L., 2000. Spatial decomposition, modeling and mapping
+service regions to predict access to social programs. Annals of
+Geographic Information Sciences 6, 105--112.
+
+\leavevmode\hypertarget{ref-ramsey2011food}{}%
+Ramsey, R., Giskes, K., Turrell, G., Gallegos, D., 2011. Food insecurity
+among australian children: Potential determinants, health and
+developmental consequences. Journal of Child Health Care 15, 401--416.
+doi:\url{https://doi.org/10.1177\%2F1367493511423854}
+
 \leavevmode\hypertarget{ref-riches2002food}{}%
@@
 
+\leavevmode\hypertarget{ref-seligman2010food}{}%
+Seligman, H.K., Laraia, B.A., Kushel, M.B., 2010. Food insecurity is
+associated with chronic disease among low-income NHANES participants.
+The Journal of nutrition 140, 304--310.
+
 \leavevmode\hypertarget{ref-statisticscanada2020food}{}%
@@
 (No. Catalogue no. 45280001).
+
+\leavevmode\hypertarget{ref-stuff2004household}{}%
+Stuff, J.E., Casey, P.H., Szeto, K.L., Gossett, J.M., Robbins, J.M.,
+Simpson, P.M., Connell, C., Bogle, M.L., 2004. Household food insecurity
+is associated with adult health status. The Journal of nutrition 134,
+2330--2335.
+doi:\href{https://doi.org/Oxford\%20University\%20Press}{Oxford University Press}
 
@@
 \leavevmode\hypertarget{ref-tarasuk2020relationship}{}%
-Tarasuk, V., St-Germain, A.-A.F., Loopstra, R., 2020. The relationship
-between food banks and food insecurity: Insights from canada. VOLUNTAS:
-International Journal of Voluntary and Nonprofit Organizations 31,
-841--852.
+Tarasuk, V., Fafard St-Germain, A.-A., Loopstra, R., 2020. The
+relationship between food banks and food insecurity: Insights from
+canada. VOLUNTAS: International Journal of Voluntary and Nonprofit
+Organizations 31, 841--852.
+doi:\href{https://doi.org/10.1007/s11266-019-00092-w}{10.1007/s11266-019-00092-w}
+
+\leavevmode\hypertarget{ref-tarasuk2009household}{}%
+Tarasuk, V., Vogt, J., 2009. Household food insecurity in ontario.
+Canadian Journal of Public Health 100, 184--188.
+doi:\href{https://doi.org/10.1007/BF03405537}{10.1007/BF03405537}
 
@@
 
+\leavevmode\hypertarget{ref-wan2012three}{}%
+Wan, N., Zou, B., Sternberg, T., 2012. A three-step floating catchment
+area method for analyzing spatial access to health services.
+International Journal of Geographical Information Science 26,
+1073--1089.
+doi:\href{https://doi.org/10.1080/13658816.2011.624987}{10.1080/13658816.2011.624987}
+
+\leavevmode\hypertarget{ref-widener2018spatial}{}%
+Widener, M.J., 2018. Spatial access to food: Retiring the food desert
+metaphor. Physiology \& behavior 193, 257--260.
+doi:\url{https://doi.org/10.1016/j.physbeh.2018.02.032}
+
+\leavevmode\hypertarget{ref-wolfson2020food}{}%
+Wolfson, J.A., Leung, C.W., 2020. Food insecurity and COVID-19:
+Disparities in early effects for US adults. Nutrients 12, 1648.
+
 \end{CSLReferences}

### 2021-04-21 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
First draft of data section

@@
 ---
+
+# CRediT author statement
+
+**Author 1:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization, Supervision; **Author 2:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization; **Author 3:** Conceptualization, Investigation, Data Curation, Writing - Original Draft; **Author 4:** Resources - Writing: Review & Editing - Supervision
 
@@
 
-\noindent and therefore:
+\noindent and also:
 $$
@@
 
-Words go here.
+Data have been prepared for sharing as a data package^[add repository].
+
+### Statistics Canada
+
+Population and income statistics at the level of Dissemination Areas (DAs) were retrieved using the package `cancensus` [@vonBergmann2021cancensus].
+Dissemination Areas are the smallest publicly available census geography in Canada.
+We use data from the 2016 Population Census.
+
+### Origins: Residential parcels
+
+We converted all recorded residential parcels in the City of Hamilton to points on the road network.
+Each point includes information about the number of residential units in the parcel.
+Each residential unit is "populated" with the probability of being a "low income household", which we define as having a total household income of less than CAD 40,000.
+This is approximately the mid-point of the low income cut-off (LICOs) for families in Canadian cities with populations greater than 500,000 in 2016, to match other Census data [@statisticscanada2020licos].
+
+### Destinations: Foodbanks
+
+**The locations of foodbanks were obtained from public records and geocoded. Three urban sites not in the program were also identified and geocoded for comparison purposes.**
+
+### Routing and travel time tables
+
+Travel time tables for three modes (car, transit, walking) were computed using the parcels as the origins and the locations of the foodbanks as the destinations.
+For routing, the package `r5r` [@pereira2021r5r] was used, with a network extract for the City of Hamilton from OpenStreetMaps and the General Transit Feed Specification from Hamilton Street Railway, the local transit operator.
+For routing purposes we used maximum values of 180 min and 10,000 m walking distance: any destination that exceeded these thresholds was ignored.
+The departure time used for routing was **TIME**.
+
+### Open Hamilton **???**
+
+We used the open data portal of the City of Hamilton\footnote{\url{https://open.hamilton.ca/}} we obtained boundaries for the city's various regions (the definition of urban, suburban, and rural regions follows the classification of development applications).
+
+### Transportation Tomorrow Survey
+
+We used the Data Retrieval System of the Transportation Tomorrow Survey (TTS)\footnote{\url{http://dmg.utoronto.ca/}} to download cross-tabulations of: 1) primary mode of travel per trip by income by place of residence; and 2) age by income by place of residence.
+These data are from the 2016 Survey (the most recent available), and data are geocoded at the level of Traffic Analysis Zones (TAZ) using the most recent zoning system from 2006.
+Each parcel point is populated with the proportion of trips by three modes of travel: car (as driver or passenger), transit, and walk.
+
+### Expected Travel Times
+
+Once we obtained travel time tables with population (number of households) and proportion of trips by mode, we calculated the expected travel time $ett$ from each parcel $i$ to a foodbank $j$ as follows:
+$$
+ett_i = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
+$$
+\noindent where $p^k_i$ is the proportion of trips by mode $k$ in the TAZ of parcel $i$, and $tt^k_{ij}$ is the travel time from parcel $i$ to the foodbank.
+In other words, the expected travel time is the weighted sum of travel times to the foodbank, with the weights given by the expected modal split in the TAZ.
 

### 2021-04-21 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
First draft of data section

@@
 
+\hypertarget{credit-author-statement}{%
+\section{CRediT author statement}\label{credit-author-statement}}
+
+\textbf{Author 1:} Conceptualization, Methodology, Software, Validation,
+Formal analysis, Investigation, Data Curation, Writing - Original Draft,
+Visualization, Supervision; \textbf{Author 2:} Conceptualization,
+Methodology, Software, Validation, Formal analysis, Investigation, Data
+Curation, Writing - Original Draft, Visualization; \textbf{Author 3:}
+Conceptualization, Investigation, Data Curation, Writing - Original
+Draft; \textbf{Author 4:} Resources - Writing: Review \& Editing -
+Supervision
+
 \newpage
@@
 exacerbated. Along several other dimensions of stress {[}e.g.,
-accessibility to health care facilities; Pereira et al. (2021){]}, this
+accessibility to health care facilities; Pereira et al. (2021a){]}, this
 seems to be the case for food insecurity as well (Laborde et al., 2020).
@@
 occurrence during the pandemic), also experienced food insecurity.
-Similarly, according to Statistics Canada (2020), in the early stages of
-the pandemic almost 15\% of individuals reported living in a household
-that faced food insecurity; the risk of food insecurity was
+Similarly, according to Statistics Canada (2020a), in the early stages
+of the pandemic almost 15\% of individuals reported living in a
+household that faced food insecurity; the risk of food insecurity was
 substantially higher for households with children. The difference
@@
 (e.g., Enns et al., 2020), we are not aware of any research that has
-focused on the geographic component of accessibility.
+focused on the geographic component of accessibility. \textbf{Check
+Tarasuk et al. (2019)}
 
@@
 population living in food insecurity in the past 30 days (Statistics
-Canada, 2020). Food insecurity is a highly concerning public health
+Canada, 2020a). Food insecurity is a highly concerning public health
 issue due to its health impacts. Recent research suggests that food
@@
 
-\noindent and therefore: \[
+\noindent and also: \[
 \sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
@@
 
-Words go here.
+Data have been prepared for sharing as a data package\footnote{add
+  repository}.
+
+\hypertarget{statistics-canada}{%
+\subsubsection{Statistics Canada}\label{statistics-canada}}
+
+Population and income statistics at the level of Dissemination Areas
+(DAs) were retrieved using the package \texttt{cancensus} (von Bergmann
+et al., 2021). Dissemination Areas are the smallest publicly available
+census geography in Canada. We use data from the 2016 Population Census.
+
+\hypertarget{origins-residential-parcels}{%
+\subsubsection{Origins: Residential
+parcels}\label{origins-residential-parcels}}
+
+We converted all recorded residential parcels in the City of Hamilton to
+points on the road network. Each point includes information about the
+number of residential units in the parcel. Each residential unit is
+``populated'' with the probability of being a ``low income household,''
+which we define as having a total household income of less than CAD
+40,000. This is approximately the mid-point of the low income cut-off
+(LICOs) for families in Canadian cities with populations greater than
+500,000 in 2016, to match other Census data (Statistics Canada, 2020b).
+
+\hypertarget{destinations-foodbanks}{%
+\subsubsection{Destinations: Foodbanks}\label{destinations-foodbanks}}
+
+\textbf{The locations of foodbanks were obtained from public records and
+geocoded. Three urban sites not in the program were also identified and
+geocoded for comparison purposes.}
+
+\hypertarget{routing-and-travel-time-tables}{%
+\subsubsection{Routing and travel time
+tables}\label{routing-and-travel-time-tables}}
+
+Travel time tables for three modes (car, transit, walking) were computed
+using the parcels as the origins and the locations of the foodbanks as
+the destinations. For routing, the package \texttt{r5r} (Pereira et al.,
+2021b) was used, with a network extract for the City of Hamilton from
+OpenStreetMaps and the General Transit Feed Specification from Hamilton
+Street Railway, the local transit operator. For routing purposes we used
+maximum values of 180 min and 10,000 m walking distance: any destination
+that exceeded these thresholds was ignored. The departure time used for
+routing was \textbf{TIME}.
+
+\hypertarget{open-hamilton}{%
+\subsubsection{\texorpdfstring{Open Hamilton
+\textbf{???}}{Open Hamilton ???}}\label{open-hamilton}}
+
+We used the open data portal of the City of
+Hamilton\footnote{\url{https://open.hamilton.ca/}} we obtained
+boundaries for the city's various regions (the definition of urban,
+suburban, and rural regions follows the classification of development
+applications).
+
+\hypertarget{transportation-tomorrow-survey}{%
+\subsubsection{Transportation Tomorrow
+Survey}\label{transportation-tomorrow-survey}}
+
+We used the Data Retrieval System of the Transportation Tomorrow Survey
+(TTS)\footnote{\url{http://dmg.utoronto.ca/}} to download
+cross-tabulations of: 1) primary mode of travel per trip by income by
+place of residence; and 2) age by income by place of residence. These
+data are from the 2016 Survey (the most recent available), and data are
+geocoded at the level of Traffic Analysis Zones (TAZ) using the most
+recent zoning system from 2006. Each parcel point is populated with the
+proportion of trips by three modes of travel: car (as driver or
+passenger), transit, and walk.
+
+\hypertarget{expected-travel-times}{%
+\subsubsection{Expected Travel Times}\label{expected-travel-times}}
+
+Once we obtained travel time tables with population (number of
+households) and proportion of trips by mode, we calculated the expected
+travel time \(ett\) from each parcel \(i\) to a foodbank \(j\) as
+follows: \[
+ett_i = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
+\] \noindent where \(p^k_i\) is the proportion of trips by mode \(k\) in
+the TAZ of parcel \(i\), and \(tt^k_{ij}\) is the travel time from
+parcel \(i\) to the foodbank. In other words, the expected travel time
+is the weighted sum of travel times to the foodbank, with the weights
+given by the expected modal split in the TAZ.
 
@@
 Pereira, R.H.M., Braga, C.K.V., Servo, L.M., Serra, B., Amaral, P.,
-Gouveia, N., Paez, A., 2021. Geographic access to COVID-19 healthcare in
-brazil using a balanced float catchment area approach. Social Science \&
-Medicine 273, 113773.
+Gouveia, N., Paez, A., 2021a. Geographic access to COVID-19 healthcare
+in brazil using a balanced float catchment area approach. Social Science
+\& Medicine 273, 113773.
 doi:\url{https://doi.org/10.1016/j.socscimed.2021.113773}
+
+\leavevmode\hypertarget{ref-pereira2021r5r}{}%
+Pereira, R.H.M., Saraiva, M., Herszenhut, D., Braga, C.K.V., Conway,
+M.W., 2021b. r5r: Rapid realistic routing on multimodal transport
+networks with r\textsuperscript{5} in r. Findings.
+doi:\href{https://doi.org/10.32866/001c.21262}{10.32866/001c.21262}
 
@@
 \leavevmode\hypertarget{ref-statisticscanada2020food}{}%
-Statistics Canada, 2020. Food insecurity during the COVID-19 pandemic
+Statistics Canada, 2020a. Food insecurity during the COVID-19 pandemic
 (No. Catalogue no. 45280001).
+
+\leavevmode\hypertarget{ref-statisticscanada2020licos}{}%
+Statistics Canada, 2020b. Table 11-10-0241-01 low income cut-offs
+(LICOs) before and after tax by community size and family size, in
+current dollars. doi:\url{https://doi.org/10.25318/1110024101-eng}
 
@@
 
+\leavevmode\hypertarget{ref-tarasuk2019geographic}{}%
+Tarasuk, V., Fafard St-Germain, A.-A., Mitchell, A., 2019. Geographic
+and socio-demographic predictors of household food insecurity in canada,
+2011--12. BMC Public Health 19, 12.
+doi:\href{https://doi.org/10.1186/s12889-018-6344-2}{10.1186/s12889-018-6344-2}
+
 \leavevmode\hypertarget{ref-tarasuk2009household}{}%
@@
 doi:\href{https://doi.org/10.24095/hpcdp.37.9.01}{10.24095/hpcdp.37.9.01}
+
+\leavevmode\hypertarget{ref-vonBergmann2021cancensus}{}%
+von Bergmann, J., Shkolnik, D., Jacobs, A., 2021. Cancensus: R package
+to access, retrieve, and work with canadian census data and geography.
 

### 2021-04-21 00-Data-Processing/00-Data-Processing-for-Food-Banks-Accessibility.Rmd
Add population output
Clean up pre-processing code and add output by population groups from tts

@@
 
-### Hamilton CMA boundaries
+### Census Data
 
-Statistics Canada. Load:
+First, load the water boundary for Hamilton Harbour:
+
 ```{r}
-load("input-data-files/hamilton_cma.RData")
-#load("input-data-files/hamilton_da_2016.RData")
+load("input-data-files/hamilton_harbour.RData")
 ```
 
-### Hamilton Ward boundaries
+### Census data
 
-Open Data Hamilton. Read:
-```{r}
-wards <- st_read("input-data-files/Ward_Boundaries.shp")
+We can get an API key to use `cancensus` from [CensusMapper](https://censusmapper.ca/). Once you have an API key, it can be stored locally this way:
+```{r eval=FALSE}
+set_api_key("YOUR API KEY HERE")
 ```
 
-Add label for type of ward (the urban/suburban/rural classification is by the planning teams; see https://www.hamilton.ca/develop-property/planning-applications/development-applications-mapping)
-```{r}
-wards$Type <- c("Urban",
-                "Urban",
-                "Urban",
-                "Urban",
-                "Urban",
-                "Suburban",
-                "Suburban",
-                "Suburban",
-                "Suburban",
-                "Suburban",
-                "Suburban",
-                "Suburban",
-                "Suburban",
-                "Suburban",
-                "Suburban")
+Once you have your API you can use `cancensus` like in this example.
+
+Use `cancensus` to check and download census data. Check the regions:
+```{r message = FALSE}
+list_census_regions('CA16') %>% 
+  dplyr::filter(level == "CSD", name %in% c("Hamilton"))
 ```
 
-Merge boundaries by type:
+We need PR_UID 35.
+
+List census data sets:
 ```{r}
-wards <- wards %>% 
-  group_by(Type) %>%
-  summarize()
+list_census_datasets()
 ```
 
+We need CA16.
+
+Search variables for DA-level data:
 ```{r}
-ggplot(wards) + 
-  geom_sf(aes(fill = Type))
+find_census_vectors("age", 
+                    dataset = "CA16", 
+                    type = "total", 
+                    query_type = "keyword", 
+                    interactive = FALSE)
 ```
 
-Read rural boundary:
+We need v_CA16_2406 to v_CA16_2425: these are the number of households at certain levels of total income. Retrieve data:
 ```{r}
-rural <- st_read("input-data-files/Rural_Boundary.shp")
-rural <- rural %>%
-  transmute(Type = "Rural")
+data_da_2016 <- get_census(dataset='CA16', 
+                          regions=list(CSD=c("3525005")),
+                          vectors=c("v_CA16_2406",
+                                    "v_CA16_2407",
+                                    "v_CA16_2408",
+                                    "v_CA16_2409",
+                                    "v_CA16_2410",
+                                    "v_CA16_2411",
+                                    "v_CA16_2412",
+                                    "v_CA16_2413",
+                                    "v_CA16_2414",
+                                    "v_CA16_2415",
+                                    "v_CA16_2416",
+                                    "v_CA16_2417",
+                                    "v_CA16_2418",
+                                    "v_CA16_2419",
+                                    "v_CA16_2420",
+                                    "v_CA16_2421",
+                                    "v_CA16_2422",
+                                    "v_CA16_2423",
+                                    "v_CA16_2424",
+                                    "v_CA16_2425",
+                                    "v_CA16_2426",
+                                    "v_CA16_2397"), # median hhinc
+                          level='DA', 
+                          use_cache = FALSE,
+                          geo_format = "sf") %>%
+  # project to NAD 1983 Zone 17N
+  st_transform(crs = 26917)
 ```
 
-Plot rural boundary:
-```{r}
-ggplot(rural) +
-  geom_sf()
+#### Prepare Geography
+
+Clip `data_da_2016` to remove the Hamilton Harbour from the CSD polygons:
+```{r warning = FALSE}
+data_da_2016 <- st_difference(data_da_2016, st_union(st_combine(hamilton_harbour %>% filter(NAME == "Hamilton Harbour")))) %>% drop_crumbs(threshold = 1)
+
+tm_shape(data_da_2016) + 
+  tm_fill(col = "Population")
 ```
 
-Obtain difference between urban and suburban regions and rural boundary:
+Create a Hamilton CSD (municipal) boundary polygon:
 ```{r}
-st_difference(wards, rural) %>%
-  ggplot() +
-  geom_sf(aes(fill = Type))
+hamilton_boundary <- data_da_2016 %>%
+  group_by(CSD_UID) %>%
+  summarize(Population = sum(Population))
+
+tm_shape(hamilton_boundary) + 
+  tm_fill(col = "Population")
 ```
 
-Notice the detritus after the difference. Drop crumbs to obtain a cleaner version of the boundaries:
+#### Prepare Summary Measures
+
+Calculate number of households with incomes < \$40k per DA and proportion of households with incomes < \$40k per DA:
 ```{r}
-sub_urban <- st_difference(wards, rural) %>%
-  transmute(Type) %>%
-  smoothr::drop_crumbs(threshold = 5000)
-
-ggplot(sub_urban) +
-  geom_sf(aes(fill = Type))
+data_da_2016 <- data_da_2016 %>%
+  #filter(`Region Name` == "Hamilton") %>%
+  mutate(income_less40k = (`v_CA16_2406: Under $5,000` + 
+                             `v_CA16_2407: $5,000 to $9,999` + 
+                             `v_CA16_2408: $10,000 to $14,999` +
+                             `v_CA16_2409: $15,000 to $19,999` +
+                             `v_CA16_2410: $20,000 to $24,999` +  
+                             `v_CA16_2411: $25,000 to $29,999` +
+                             `v_CA16_2412: $30,000 to $34,999` +
+                             `v_CA16_2413: $35,000 to $39,999`),
+         p_inc_less40k = income_less40k/Households)
 ```
 
-Bind the boundaries with all three urban types:
+Plot population:
 ```{r}
-urban_types <- rbind(sub_urban,
-                     rural)
+tm_shape(data_da_2016) + 
+  tm_fill(col = "income_less40k", alpha = 0.8)
+```
 
-ggplot(urban_types) +
-  geom_sf(aes(fill = Type))
+```{r income-quintiles, include=FALSE}
+# Data preparation
+## Calculate quintiles of median household income in Hamilton:
+mhi_q <- data_da_2016 %>%
+  dplyr::filter(`Region Name` == "Hamilton") %>%
+  pull(`v_CA16_2397: Median total income of households in 2015 ($)`) %>%
+  quantile(c(0, 0.2, 0.4, 0.6, 0.8, 1), 
+           na.rm = TRUE)
+```
+
+Label DAs with income quintiles:
+```{r}
+data_da_2016 <- data_da_2016 %>%
+  rename(Median_household_income = `v_CA16_2397: Median total income of households in 2015 ($)`) %>%
+  mutate(income_quintile = case_when(Median_household_income < mhi_q[2] ~ "Bottom 20%",
+                                    Median_household_income >= mhi_q[2] & Median_household_income < mhi_q[3] ~ "Fourth 20%",
+                                    Median_household_income >= mhi_q[3] & Median_household_income < mhi_q[4] ~ "Third 20%",
+                                    Median_household_income >= mhi_q[4] & Median_household_income < mhi_q[5] ~ "Second 20%",
+                                    Median_household_income >= mhi_q[5] ~ "Top 20%"),
+         income_quintile = factor(income_quintile,
+                                levels = c("Top 20%",
+                                           "Second 20%",
+                                           "Third 20%",
+                                           "Fourth 20%",
+                                           "Bottom 20%"),
+                                ordered = TRUE))
 ```
@@
 foodbank_locations <- foodbank_locations %>%
-  st_as_sf(coords = c("longitude", "latitude"), 
-           crs = 4326) %>%
-  #st_transform(crs = 26917) %>%
+  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
+  st_transform(crs = 26917) %>% # project to NAD 1983 Zone 17N
 
   # filter out points beyond the Hamilton CMA (there's one point in Brantford)
-  filter(st_intersects(x = geometry, y = hamilton_cma, sparse = FALSE))
+  filter(st_intersects(x = geometry, y = hamilton_boundary, sparse = FALSE))
 ```
@@
 ggplot() +
-  geom_sf(data = hamilton_cma) +
+  geom_sf(data = hamilton_boundary) +
   geom_sf(data = foodbank_locations,
@@
 load("input-data-files/dwelling_network_points_2016.RData")
-```
 
-### Census data
-
-We can get an API key to use `cancensus` from [CensusMapper](https://censusmapper.ca/). Once you have an API key, it can be stored locally this way:
-```{r eval=FALSE}
-set_api_key("YOUR API KEY HERE")
-```
-
-Once you have your API you can use `cancensus` like in this example.
-
-Use `cancensus` to check and download census data. Check the regions:
-```{r}
-list_census_regions('CA16') %>% 
-  dplyr::filter(level == "CMA", name %in% c("Hamilton"))
-```
-
-We need PR_UID 35.
-
-List census data sets:
-```{r}
-list_census_datasets()
-```
-
-We need CA16.
-
-Search variables:
-```{r}
-find_census_vectors("age", 
-                    dataset = "CA16", 
-                    type = "total", 
-                    query_type = "keyword", 
-                    interactive = FALSE)
-```
-
-We need v_CA16_2406 to v_CA16_2425: these are the number of households at certain levels of total income. Retrieve data:
-```{r}
-data_da_2016 <- get_census(dataset='CA16', 
-                          regions=list(CMA=c("35537")),
-                          vectors=c("v_CA16_2406",
-                                    "v_CA16_2407",
-                                    "v_CA16_2408",
-                                    "v_CA16_2409",
-                                    "v_CA16_2410",
-                                    "v_CA16_2411",
-                                    "v_CA16_2412",
-                                    "v_CA16_2413",
-                                    "v_CA16_2414",
-                                    "v_CA16_2415",
-                                    "v_CA16_2416",
-                                    "v_CA16_2417",
-                                    "v_CA16_2418",
-                                    "v_CA16_2419",
-                                    "v_CA16_2420",
-                                    "v_CA16_2421",
-                                    "v_CA16_2422",
-                                    "v_CA16_2423",
-                                    "v_CA16_2424",
-                                    "v_CA16_2425",
-                                    "v_CA16_2426",
-                                    "v_CA16_2397", # median hhinc
-                                    # age variables
-                                    "v_CA16_10", #<1
-                                    "v_CA16_13", #1
-                                    "v_CA16_16", #2
-                                    "v_CA16_7", #0-4 yo
-                                    "v_CA16_25", #5-9 yo
-                                    "v_CA16_43", #10-14 yo
-                                    "v_CA16_64", # 15-19 yo
-                                    "v_CA16_244"), # 65+
-                          level='DA', 
-                          use_cache = FALSE,
-                          geo_format = "sf")
-```
-
-#### Prepare Geography
-
-Clip `data_da_2016`:
-```{r}
-data_da_2016 <- data_da_2016 %>%
-  st_intersection(hamilton_cma)
-```
-
-#### Prepare Summary Measures
-
-Calculate number of households with incomes < \$40k per DA and proportion of households with incomes < \$40k per DA:
-```{r}
-data_da_2016 <- data_da_2016 %>%
-  filter(`Region Name` == "Hamilton") %>%
-  mutate(income_less40k = (`v_CA16_2406: Under $5,000` + 
-                             `v_CA16_2407: $5,000 to $9,999` + 
-                             `v_CA16_2408: $10,000 to $14,999` +
-                             `v_CA16_2409: $15,000 to $19,999` +
-                             `v_CA16_2410: $20,000 to $24,999` +  
-                             `v_CA16_2411: $25,000 to $29,999` +
-                             `v_CA16_2412: $30,000 to $34,999` +
-                             `v_CA16_2413: $35,000 to $39,999`),
-         p_inc_less40k = income_less40k/Households)
-```
-
-Plot population:
-```{r}
-tmap_mode(mode = "view")
-
-data_da_2016 %>%
-  tm_shape() +
-  tm_polygons("income_less40k",
-              alpha = 0.8)
-```
-
-Label DAs with income quintiles:
-```{r}
-data_da_2016 <- data_da_2016 %>%
-  rename(Median_household_income = `v_CA16_2397: Median total income of households in 2015 ($)`) %>%
-  mutate(income_quintile = case_when(Median_household_income < mhi_q[2] ~ "Bottom 20%",
-                                    Median_household_income >= mhi_q[2] & Median_household_income < mhi_q[3] ~ "Fourth 20%",
-                                    Median_household_income >= mhi_q[3] & Median_household_income < mhi_q[4] ~ "Third 20%",
-                                    Median_household_income >= mhi_q[4] & Median_household_income < mhi_q[5] ~ "Second 20%",
-                                    Median_household_income >= mhi_q[5] ~ "Top 20%"),
-         income_quintile = factor(income_quintile,
-                                levels = c("Top 20%",
-                                           "Second 20%",
-                                           "Third 20%",
-                                           "Fourth 20%",
-                                           "Bottom 20%"),
-                                ordered = TRUE))
-```
-
-Calculate low-income age totals
-```{r}
-data_da_2016 <- data_da_2016 %>%
-  mutate(tot_age_0to2 = `v_CA16_10: Under 1 year` + `v_CA16_13: 1` + `v_CA16_16: 2`,
-         linc_age_0to2 = tot_age_0to2 * p_inc_less40k,
-         tot_age_0to19 = `v_CA16_7: 0 to 4 years` + `v_CA16_25: 5 to 9 years` + `v_CA16_43: 10 to 14 years` + `v_CA16_64: 15 to 19 years`,
-         linc_age_0to19 = tot_age_0to19 * p_inc_less40k,
-         tot_age_65plus = `v_CA16_244: 65 years and over`,
-         linc_age_65plus = tot_age_65plus * p_inc_less40k,
-         linc_population = Population * p_inc_less40k)
-```
-
-Plot result
-```{r}
-tmap_mode(mode = "view")
-
-data_da_2016 %>%
-  tm_shape() +
-  tm_polygons("linc_population",
-              alpha = 0.8)
+dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
+  st_transform(crs = 26917) # project to NAD 1983 Zone 17N
 ```
@@
 ```{r}
-tts_taz <- st_read("input-data-files/tts2016_2006zn_2018_region.shp")
+tts_taz <- st_read("input-data-files/tts2016_2006zn_2018_region.shp") %>% 
+  st_transform(crs = 26917) %>% # project to NAD 1983 Zone 17N
+  st_make_valid() # fix a few topolgy errors in the boundaries
 ```
@@
 save(data_da_2016, file = "output-data-files/data_da_2016.RData", compress = "xz")
+save(hamilton_boundary, file = "output-data-files/hamilton_boundary.RData", compress = "xz")
 save(dwelling_network_points_2016, file = "output-data-files/dwelling_network_points_2016.RData", compress = "xz")
 save(foodbank_locations, file = "output-data-files/foodbank_locations.RData", compress = "xz")
-save(hamilton_cma, file = "output-data-files/hamilton_cma.RData", compress = "xz")
 save(modes_less40k, file = "output-data-files/modes_less40k.RData", compress = "xz")
 save(population_age_income, file = "output-data-files/population_age_income.RData", compress = "xz")
-save(urban_types, file = "output-data-files/urban_types.RData", compress = "xz")
+#save(urban_types, file = "output-data-files/urban_types.RData", compress = "xz")
 ```

### 2021-04-21 00-Data-Processing/04-Data-Analysis.Rmd
Add population output
Clean up pre-processing code and add output by population groups from tts

--- 2021-04-19 00-Data-Processing/04-Data-Analysis.Rmd
@@
 load("output-data-files/dwelling_network_points_2016.RData")
-load("output-data-files/hamilton_cma.RData")
+load("output-data-files/hamilton_boundary.RData")
+load("output-data-files/population_age_income.RData")
 load("output-data-files/foodbank_locations.RData")
@@
               dplyr::transmute(DAUID = GeoUID, 
-                               p_inc_less40k,
-                               linc_age_0to2,
-                               linc_age_0to19,
-                               linc_age_65plus),
+                               p_inc_less40k),
             by = "DAUID")
@@
 Calculate probability that a point on the network will have low income households.
-This is the proportion of low income households in the DA multiplied by the total number of residential units a the point:
+This is the proportion of low income households in the DA multiplied by the total number of residential units at the point:
 ```{r}
@@
                             income_less40k,
-                            linc_age_0to2,
-                            linc_age_0to19,
-                            linc_age_65plus,
                             DAUID, 
@@
 
-Join tables:
+```{r}
+# join tripmaking modal proportions
+ttm_car <- ttm_car %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+
+ttm_transit <- ttm_transit %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+
+ttm_walk <- ttm_walk %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+```
+
+Prepare and join population data:
+```{r number-of-residential-units-per-DA, include=FALSE}
+# Data preparation
+## Find number of residential units in each DA. This is needed to assign the population to parcels
+residential_units <- dwelling_network_points_2016 %>%
+  st_drop_geometry() %>%
+  group_by(TAZUID) %>%
+  summarize(residential_units = sum(Dwellings))
+```
+
+```{r join-population-to-parcels, include=FALSE}
+# Data preparation
+## Join number of people age 55+ per DA to dwellings and total number of residential units:
+dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
+  left_join(population_age_income %>%
+              st_drop_geometry() %>%
+              dplyr::transmute(TAZUID,
+                               children,
+                               adults,
+                               seniors),
+            by = "TAZUID") %>%
+  left_join(residential_units,
+            by = "TAZUID")
+```
+
+```{r calculate-population-per-parcel, include=FALSE}
+# Data preparation
+##Calculate population per point on the network.
+This is the population in the DA divided by the total number of residential units.
+This gives the average number of people per residential unit.
+Then, multiply by the number of dwellings (residential units) at the point:
+dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
+  mutate(children = (children/residential_units) * Dwellings,
+         adults = (adults/residential_units) * Dwellings,
+         seniors = (seniors/residential_units) * Dwellings)
+```
+
+```{r join-population-data-to-travel-time-tables, include=FALSE}
+# Data preparation
+## Join the population data to the travel time matrices:
+ttm_car <- ttm_car %>%
+  left_join(dwelling_network_points_2016 %>%
+              st_drop_geometry() %>%
+              dplyr::transmute(ID, 
+                            children,
+                            adults,
+                            seniors),
+            by = c("UID" = "ID"))
+
+ttm_transit <- ttm_transit %>%
+  left_join(dwelling_network_points_2016 %>%
+              st_drop_geometry() %>%
+              dplyr::transmute(ID, 
+                            children,
+                            adults,
+                            seniors),
+            by = c("UID" = "ID"))
+
+ttm_walk <- ttm_walk %>%
+  left_join(dwelling_network_points_2016 %>%
+              st_drop_geometry() %>%
+              dplyr::transmute(ID, 
+                            children,
+                            adults,
+                            seniors),
+            by = c("UID" = "ID"))
+```
+
+Join all tables:
 ```{r}
@@
             by = c("UID", "OBJECTID"))
-```
-
-Join proportion of trips to travel time table:
-```{r}
-ttm <- ttm %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
 ```
@@
 ggplot() +
-  geom_sf(data = hamilton_cma) +
+  geom_sf(data = hamilton_boundary) +
   geom_sf(data = e_travel_time_parcel,
@@
 ggplot() +
-  geom_sf(data = hamilton_cma) +
+  geom_sf(data = hamilton_boundary) +
   geom_sf(data = e_travel_time_parcel_2,
@@
 los_pre <- ggplot() +
-  geom_sf(data = hamilton_cma) +
+  geom_sf(data = hamilton_boundary) +
   geom_sf(data = los %>%
@@
 los_covid <- ggplot() +
-  geom_sf(data = hamilton_cma) +
+  geom_sf(data = hamilton_boundary) +
   geom_sf(data = los %>%
@@
   ggplot() +
-  geom_sf(data = hamilton_cma) +
+  geom_sf(data = hamilton_boundary) +
   geom_sf(aes(color = los_d,
@@
 acc_pre <- ggplot() +
-  geom_sf(data = hamilton_cma) +
+  geom_sf(data = hamilton_boundary) +
   geom_sf(data = accessibility %>%
@@
 acc_covid <- ggplot() +
-  geom_sf(data = hamilton_cma) +
+  geom_sf(data = hamilton_boundary) +
   geom_sf(data = accessibility %>%
@@
   ggplot() +
-  geom_sf(data = hamilton_cma) +
+  geom_sf(data = hamilton_boundary) +
   geom_sf(aes(fill = acc_d),
@@
 
-### CODE TO BE UPDATED
-```{r}
-# join tripmaking modal proportions
-ttm_car <- ttm_car %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-
-ttm_transit <- ttm_transit %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-
-ttm_walk <- ttm_walk %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-```
+### Pre-COVID Scenario
 
@@
 ```{r minimum-travel-time-baseline, include=FALSE}
-# retrieve minimum travel time during covid to sites that didnt close
-min_ttm_car_pre <- ttm_car %>%
+# retrieve minimum travel time for pre-covid locations
+min_ttm_car_baseline <- ttm_car %>%
   #dplyr::filter(COVIDClose == "No") %>%
@@
   summarize(travel_time = min(travel_time),
-            linc_age_0to2 = first(linc_age_0to2),
-            linc_age_0to19 = first(linc_age_0to19),
-            linc_age_65plus = first(linc_age_0to19),
             DAUID = first(DAUID),
@@
             p_walk = first(p_walk),
-            .groups = "drop")
+            children = first(children),
+            adults = first(adults),
+            seniors = first(seniors),
+            .groups = "drop") %>%
 
-min_ttm_transit_pre <- ttm_transit %>%
+  ## Calculate weighted travel time:
+  mutate(weighted_travel_time = travel_time * p_car)
+
+min_ttm_transit_baseline <- ttm_transit %>%
   #dplyr::filter(COVIDClose == "No") %>%
@@
             p_walk = first(p_walk),
-            .groups = "drop")
+            children = first(children),
+            adults = first(adults),
+            seniors = first(seniors),
+            .groups = "drop") %>%
 
-min_ttm_walk_pre <- ttm_walk %>%
+  ## Calculate weighted travel time:
+  mutate(weighted_travel_time = travel_time * p_transit)
+
+min_ttm_walk_baseline <- ttm_walk %>%
   #dplyr::filter(COVIDClose == "No") %>%
@@
             p_walk = first(p_walk),
-            .groups = "drop")
+            children = first(children),
+            adults = first(adults),
+            seniors = first(seniors),
+            .groups = "drop") %>%
+
+  ## Calculate weighted travel time:
+  mutate(weighted_travel_time = travel_time * p_walk)
 ```
 
-Get minimum travel times in the during covid scenario:
+Bind the tables and label by mode:
 
-```{r minimum-travel-time-during-covid, include=FALSE}
+```{r weighted-travel-time-by-mode-baseline, include=FALSE}
+# Baseline analysis (pre-COVID)
+## Bind the three tables and label them by mode:
+min_ttm_mode_baseline <- rbind(data.frame(min_ttm_car_baseline, Mode = "Car"),
+                               data.frame(min_ttm_transit_baseline, Mode = "Transit"),
+                               data.frame(min_ttm_walk_baseline, Mode = "Walking"))
+```
+
+```{r aggregate-weighted-travel-times-baseline, include=FALSE}
+# Baseline analysis (pre-COVID)
+## Aggregate weighted travel times by mode:
+min_ttm_mode_baseline <- min_ttm_mode_baseline %>%
+  group_by(UID) %>%
+  summarize(DAUID = first(DAUID),
+            TAZUID = first(TAZUID),
+            children = first(children),
+            adults = first(adults),
+            seniors = first(seniors),
+            weighted_travel_time = sum(weighted_travel_time)) %>%
+
+  # add geometry
+  #left_join(population_age_income %>%
+  #            transmute(geometry, TAZUID),
+  #          by = "TAZUID") %>%
+
+  # calculate person-weighted hours
+  mutate(wph_children = children*weighted_travel_time/60,
+         wph_adults = adults*weighted_travel_time/60,
+         wph_seniors = seniors*weighted_travel_time/60)
+```
+
+```{r summarize-results-by-taz-baseline, include=FALSE}
+# Baseline analysis (pre-COVID)
+## Summarize by TAZ and join geometry:
+min_ttm_baseline_taz <- min_ttm_mode_baseline %>%
+  group_by(TAZUID) %>%
+  summarize(weighted_travel_time = mean(weighted_travel_time),
+            wph_children = sum(wph_children),
+            wph_adults = sum(wph_adults),
+            wph_seniors = sum(wph_seniors)) %>%
+
+  left_join(population_age_income %>%
+              transmute(geometry, TAZUID),
+            by = "TAZUID") %>%
+
+  st_as_sf()
+```
+
+```{r}
+tm_shape(min_ttm_baseline_taz) +
+  tm_fill("wph_children")
+
+tm_shape(min_ttm_baseline_taz) +
+  tm_fill("wph_adults")
+
+tm_shape(min_ttm_baseline_taz) +
+  tm_fill("wph_seniors")
+```
+
+```{r weighted-travel-time-map-baseline}
+# Baseline analysis (pre-COVID)
+## Map weighted travel time:
+ttm_map_baseline <- ggplot() +
+  geom_sf(data = min_ttm_baseline_taz,
+          aes(fill = weighted_travel_time),
+          color = NA) +
+  scale_fill_distiller(name = "Expected travel time (min)",
+                       palette = "OrRd", 
+                       direction = 1)
+
+ttm_map_baseline
+```
+
+### Do it again for the during-COVID scenario:
+
+```{r minimum-travel-time-covid, include=FALSE}
 # retrieve minimum travel time during covid to sites that didnt close
-min_ttm_car_during <- ttm_car %>%
+min_ttm_car_covid <- ttm_car %>%
   dplyr::filter(COVIDClose == "No") %>%
@@
             p_walk = first(p_walk),
-            .groups = "drop")
+            children = first(children),
+            adults = first(adults),
+            seniors = first(seniors),
+            .groups = "drop") %>%
 
-min_ttm_transit_during <- ttm_transit %>%
+  ## Calculate weighted travel time:
+  mutate(weighted_travel_time = travel_time * p_car)
+
+min_ttm_transit_covid <- ttm_transit %>%
   dplyr::filter(COVIDClose == "No") %>%
@@
             p_walk = first(p_walk),
-            .groups = "drop")
+            children = first(children),
+            adults = first(adults),
+            seniors = first(seniors),
+            .groups = "drop") %>%
 
-min_ttm_walk_during <- ttm_walk %>%
+  ## Calculate weighted travel time:
+  mutate(weighted_travel_time = travel_time * p_transit)
+
+min_ttm_walk_covid <- ttm_walk %>%
   dplyr::filter(COVIDClose == "No") %>%
@@
             p_walk = first(p_walk),
-            .groups = "drop")
+            children = first(children),
+            adults = first(adults),
+            seniors = first(seniors),
+            .groups = "drop") %>%
+
+  ## Calculate weighted travel time:
+  mutate(weighted_travel_time = travel_time * p_walk)
+```
+
+Bind the tables and label by mode:
+
+```{r weighted-travel-time-by-mode-covid, include=FALSE}
+# Baseline analysis (pre-COVID)
+## Bind the three tables and label them by mode:
+min_ttm_mode_covid <- rbind(data.frame(min_ttm_car_covid, Mode = "Car"),
+                               data.frame(min_ttm_transit_covid, Mode = "Transit"),
+                               data.frame(min_ttm_walk_covid, Mode = "Walking"))
+```
+
+```{r aggregate-weighted-travel-times-covid, include=FALSE}
+# Baseline analysis (pre-COVID)
+## Aggregate weighted travel times by mode:
+min_ttm_mode_covid <- min_ttm_mode_covid %>%
+  group_by(UID) %>%
+  summarize(DAUID = first(DAUID),
+            TAZUID = first(TAZUID),
+            children = first(children),
+            adults = first(adults),
+            seniors = first(seniors),
+            weighted_travel_time = sum(weighted_travel_time)) %>%
+
+  # add geometry
+  #left_join(population_age_income %>%
+  #            transmute(geometry, TAZUID),
+  #          by = "TAZUID") %>%
+
+  # calculate person-weighted hours
+  mutate(wph_children = children*weighted_travel_time/60,
+         wph_adults = adults*weighted_travel_time/60,
+         wph_seniors = seniors*weighted_travel_time/60)
+```
+
+```{r summarize-results-by-taz-covid, include=FALSE}
+# Baseline analysis (PILOT LOCATIONS ONLY)
+## Summarize by TAZ and join geometry:
+min_ttm_covid_taz <- min_ttm_mode_covid %>%
+  group_by(TAZUID) %>%
+  summarize(weighted_travel_time = mean(weighted_travel_time),
+            wph_children = sum(wph_children),
+            wph_adults = sum(wph_adults),
+            wph_seniors = sum(wph_seniors)) %>%
+
+  left_join(population_age_income %>%
+              transmute(geometry, TAZUID),
+            by = "TAZUID") %>%
+
+  st_as_sf()
 ```
@@
 ```{r}
-# FULL JOIN THEM TOGETHER
+tm_shape(min_ttm_covid_taz) +
+  tm_fill("wph_children")
+
+tm_shape(min_ttm_covid_taz) +
+  tm_fill("wph_adults")
+
+tm_shape(min_ttm_covid_taz) +
+  tm_fill("wph_seniors")
 ```
 
-## Exploratory data analysis
+```{r weighted-travel-time-map-covid}
+# Baseline analysis (PILOT LOCATIONS ONLY)
+## Map weighted travel time:
+ttm_map_covid <- ggplot() +
+  geom_sf(data = min_ttm_covid_taz,
+          aes(fill = weighted_travel_time),
+          color = NA) +
+  scale_fill_distiller(name = "Expected travel time (min)",
+                       palette = "OrRd", 
+                       direction = 1)
 
-Summary of population:
-```{r}
-sum(data_da_2016$Population_55to69, na.rm = TRUE)
-sum(data_da_2016$Population_70plus, na.rm = TRUE)
+ttm_map_covid
 ```
 
-Check the total population:
+### Map Differences
+
+Join the min travel time tables together
+
 ```{r}
-sum(dwelling_network_points_2016$Population_55to69, na.rm = TRUE)
-sum(dwelling_network_points_2016$Population_70plus, na.rm = TRUE)
+min_ttm_d <- min_ttm_baseline_taz %>% 
+  left_join(min_ttm_covid_taz %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        wph_children_covid = wph_children,
+                        wph_adults_covid = wph_adults,
+                        wph_seniors_covid = wph_seniors), 
+            by = "TAZUID") %>%
+  mutate(wph_children_d = wph_children_covid / wph_children - 1,
+         wph_adults_d = wph_adults_covid / wph_adults - 1,
+         wph_seniors_d = wph_seniors_covid / wph_seniors - 1)
+
+tm_shape(min_ttm_d) + 
+  tm_fill("wph_children_d")
+
+tm_shape(min_ttm_d) + 
+  tm_fill("wph_adults_d")
+
+tm_shape(min_ttm_d) + 
+  tm_fill("wph_seniors_d")
 ```
-
-The difference in population is because the residential units considered do not completely cover the Hamilton CMA:
-```{r}
-ggplot() +
-geom_sf(data = hamilton_cma) + 
-geom_sf(data = dwelling_network_points_2016) + 
-geom_sf(data = pharmacy_locations, 
-aes(color = Type,
-shape = Type), 
-size = 3)
-```
-
-Plot population age 55 to 69:
-```{r}
-ggplot() +
-geom_sf(data = data_da_2016 %>%
-filter(`Region Name` == "Hamilton"),
-aes(fill = Population_55to69),
-color = NA) +
-geom_sf(data = hamilton_cma,
-fill = NA) +
-scale_fill_distiller(palette = "OrRd", 
-direction = 1)
-```
-
-Plot population age 70 plus:
-```{r}
-ggplot() +
-geom_sf(data = data_da_2016 %>%
-filter(`Region Name` == "Hamilton"),
-aes(fill = Population_70plus),
-color = NA) +
-geom_sf(data = hamilton_cma,
-fill = NA) +
-scale_fill_distiller(palette = "OrRd", 
-direction = 1)
-```

### 2021-04-21 00-Data-Processing/04-Data-Analysis.Rmd
Update wph analysis by TYPE
include the type of food bank in the analysis of wph differences

@@
               dplyr::select(ID, 
+                            TYPE,
                             COVIDClose),
@@
               dplyr::select(ID,
+                            TYPE,
                             COVIDClose),
@@
 
-### Pre-COVID Scenario
+### Pre-COVID Scenario to Meal Services
 
@@
   #dplyr::filter(COVIDClose == "No") %>%
+  filter(TYPE != "Food Banks") %>%
   group_by(UID) %>%
@@
 min_ttm_car_covid <- ttm_car %>%
-  dplyr::filter(COVIDClose == "No") %>%
+  dplyr::filter(COVIDClose == "No", TYPE != "Food Banks") %>%
   group_by(UID) %>%
@@
 min_ttm_transit_covid <- ttm_transit %>%
-  dplyr::filter(COVIDClose == "No") %>%
+  dplyr::filter(COVIDClose == "No", TYPE != "Food Banks") %>%
   group_by(UID) %>%
@@
 min_ttm_walk_covid <- ttm_walk %>%
-  dplyr::filter(COVIDClose == "No") %>%
+  dplyr::filter(COVIDClose == "No", TYPE != "Food Banks") %>%
   group_by(UID) %>%
@@
   mutate(wph_children_d = wph_children_covid / wph_children - 1,
+         wph_children_d = replace_na(wph_children_d, 0),
          wph_adults_d = wph_adults_covid / wph_adults - 1,
-         wph_seniors_d = wph_seniors_covid / wph_seniors - 1)
+         wph_adults_d = replace_na(wph_adults_d, 0),
+         wph_seniors_d = wph_seniors_covid / wph_seniors - 1,
+         wph_seniors_d = replace_na(wph_seniors_d, 0))
 

### 2021-04-22 00-Data-Processing/04-Data-Analysis.Rmd
Analysis of accessibility by age is probably complete

@@
 library(tidyverse)
+library(tmap)
 #library(vaccHamilton)
@@
 
-ggplot(accessibility_test, 
+acc_sensitivity <- ggplot(accessibility_test, 
        aes(x = threshold, y = accessibility)) +
-  geom_line()
+  geom_line() +
+  ylab("Accessibility")
+
+acc_sensitivity
 ```
@@
 Join geometry:
-```{r, add-geometry-to-results, include=FALSE}
+```{r, add-geometry-to-level-of-service, include=FALSE}
 los <- los %>%
@@
             by = c("OBJECTID" = "ID")) %>%
-  st_as_sf()
-
-accessibility <- accessibility %>%
-  left_join(dwelling_network_points_2016 %>%
-              dplyr::select(ID),
-            by = c("UID" = "ID")) %>%
   st_as_sf()
@@
 
-Plot accessibility before and during covid:
-```{r}
-acc_pre <- ggplot() +
-  geom_sf(data = hamilton_boundary) +
-  geom_sf(data = accessibility %>%
-            filter(timing == "Pre-covid"),
-          aes(color = accessibility)) +
-  scale_color_distiller(palette = "OrRd", 
-                        direction = 1)
-
-acc_covid <- ggplot() +
-  geom_sf(data = hamilton_boundary) +
-  geom_sf(data = accessibility %>%
-            filter(timing == "During covid"),
-          aes(color = accessibility)) +
-  scale_color_distiller(palette = "OrRd", 
-                        direction = 1)
-
-acc_pre / acc_covid
-```
-
-Plot the change in accessibility by DA:
-```{r}
-accessibility %>%
-  st_drop_geometry() %>%
-  # Copy the DA identifier to the accessibility table
+Join zonal identifiers to accessibility results:
+```{r}
+accessibility <- accessibility %>%
   left_join(dwelling_network_points_2016 %>%
             st_drop_geometry() %>%
-              transmute(UID = ID, DAUID),
-            by = "UID") %>%
-  # Group by DA and time
-  group_by(DAUID, timing) %>%
-  # Calculate the accessibility by DA before and during covid
+              dplyr::select(ID,
+                            DAUID,
+                            TAZUID,
+                            children,
+                            adults,
+                            seniors),
+            by = c("UID" = "ID")) %>%
+  left_join(population_age_income %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID),
+            by = "TAZUID") 
+```
+
+Accessibility by TAZ:
+```{r}  
+accessibility_taz <- accessibility %>%
+  group_by(TAZUID, 
+           timing) %>%
   summarize(accessibility = sum(accessibility),
@@
             .groups = "drop") %>%
-  # Pivot wider to calculate the differences in accessibility from pre-covid to covid
+  left_join(population_age_income %>%
+              transmute(TAZUID,
+                        children,
+                        adults,
+                        seniors),
+            by = "TAZUID") %>%
+  st_as_sf()
+```
+
+Plot accessibility before and during covid:
+```{r}
+acc_pre <- ggplot() +
+  geom_sf(data = hamilton_boundary) +
+  geom_sf(data = accessibility_taz %>%
+            filter(timing == "Pre-covid"),
+          aes(fill = accessibility)) +
+  scale_fill_distiller(palette = "OrRd", 
+                        direction = 1)
+
+acc_covid <- ggplot() +
+  geom_sf(data = hamilton_boundary) +
+  geom_sf(data = accessibility_taz %>%
+            filter(timing == "During covid"),
+          aes(fill = accessibility)) +
+  scale_fill_distiller(palette = "OrRd", 
+                        direction = 1)
+
+acc_pre / acc_covid
+```
+
+Calculate accessibility change by TAZ:
+```{r}
+accessibilty_taz_change <- accessibility_taz %>%
+  st_drop_geometry() %>%
   pivot_wider(names_from = timing, 
@@
   # Calculate changes in accessibility
-  mutate(acc_d = `During covid` / `Pre-covid` - 1) %>%
-  # Join DA geometry
-  left_join(data_da_2016 %>%
-            transmute(DAUID = GeoUID,
+  #mutate(acc_d = `During covid` / `Pre-covid` - 1) %>% # Percentage change is rather flat and not as easy to interpret as the difference
+  mutate(acc_d = `During covid` - `Pre-covid`) %>%
+  # Join TAZ geometry
+  left_join(population_age_income %>%
+            transmute(TAZUID,
                       geometry),
-          by = "DAUID") %>%
+          by = "TAZUID") %>%
   # Convert to simple features
-  st_as_sf() %>%
+  st_as_sf()
+```
+
+Plot the change in accessibility by TAZ:
+```{r}
   # Plot
   ggplot() +
-  geom_sf(data = hamilton_boundary) +
-  geom_sf(aes(fill = acc_d),
+  geom_sf(data = accessibilty_taz_change,
+          aes(fill = acc_d),
           color = NA) +
-  scale_fill_distiller(palette = "OrRd", 
+  geom_sf(data = population_age_income,
+          fill = NA) +
+  scale_fill_fermenter(palette = "OrRd", 
                         direction = -1)
-
 ```
@@
 
+Calculate accessibility change by parcel:
+```{r}
+accessibility_change <- accessibility %>%
+  select(UID, timing, accessibility) %>%
+  pivot_wider(names_from = timing, 
+              values_from = accessibility,
+              values_fill = 0) %>%
+  # Calculate changes in accessibility
+  #mutate(acc_d = `During covid` / `Pre-covid` - 1) %>% # Percentage change is rather flat and not as easy to interpret as the difference
+  mutate(acc_d = `During covid` - `Pre-covid`)
+```
+
+Join accessibility change to accessibility table:
+```{r}
+accessibility <- accessibility %>%
+  left_join(accessibility_change %>%
+              select(UID, acc_d),
+            by = "UID")
+```
+
+Quartiles of accessibility:
+```{r}
+# acc_quintiles_pre <- quantile(accessibility %>%
+#                                 filter(timing == "Pre-covid") %>%
+#                                 pull(accessibility),
+#                               c(0, 0.25, 0.5, 0.75, 1))
+# acc_quintiles_covid <- quantile(accessibility %>%
+#                                 filter(timing == "During covid") %>%
+#                                 pull(accessibility),
+#                               c(0, 0.25, 0.5, 0.75, 1))
+```
+
+Quartiles of accessibility change:
+```{r}
+acc_quintiles_change <- quantile(accessibilty_change$acc_d,
+                              c(0, 0.25, 0.5, 0.75, 1))
+```
+
+Add labels of accessibility levels to table:
+```{r}
+# accessibility <- accessibility %>%
+#   mutate(accessibility_level = case_when(timing == "Pre-covid" & accessibility <= acc_quintiles_pre[2] ~ "Bottom 25%",
+#                                          timing == "Pre-covid" & accessibility > acc_quintiles_pre[2] & accessibility <= acc_quintiles_pre[3] ~ "Third 25%",
+#                                          timing == "Pre-covid" & accessibility > acc_quintiles_pre[3] & accessibility <= acc_quintiles_pre[4] ~ "Second 25%",
+#                                          timing == "Pre-covid" & accessibility > acc_quintiles_pre[4] ~ "Top 25%",
+#                                          timing == "During covid" & accessibility <= acc_quintiles_covid[2] ~ "Bottom 25%",
+#                                          timing == "During covid" & accessibility > acc_quintiles_covid[2] & accessibility <= acc_quintiles_covid[3] ~ "Third 25%",
+#                                          timing == "During covid" & accessibility > acc_quintiles_covid[3] & accessibility <= acc_quintiles_covid[4] ~ "Second 25%",
+#                                          timing == "During covid" & accessibility > acc_quintiles_covid[4] ~ "Top 25%"),
+#            accessibility_level = factor(accessibility_level,
+#                                         levels = c("Top 25%", "Second 25%", "Third 25%", "Bottom 25%"),
+#                                         ordered = TRUE))
+```
+
+Add labels of accessibility loss to table:
+```{r}
+accessibility <- accessibility %>%
+  mutate(accessibility_level = case_when(acc_d <= acc_quintiles_change[2] ~ "Bottom 25%",
+                                         acc_d > acc_quintiles_change[2] & acc_d <= acc_quintiles_change[3] ~ "Third 25%",
+                                         acc_d > acc_quintiles_change[3] & acc_d <= acc_quintiles_change[4] ~ "Second 25%",
+                                         acc_d > acc_quintiles_change[4] ~ "Top 25%",
+                                         accessibility <= acc_quintiles_covid[2] ~ "Bottom 25%"),
+           accessibility_level = factor(accessibility_level,
+                                        levels = c("Top 25%", "Second 25%", "Third 25%", "Bottom 25%"),
+                                        ordered = TRUE))
+```
+
+Calculate accessibility by age group and accessibility level:
+```{r}
+accessibility_by_age <- accessibility %>%
+  group_by(timing, 
+           accessibility_level) %>%
+  summarize(children = sum(children),
+            adults = sum(adults),
+            seniors = sum(seniors),
+            accessibility = sum(accessibility),
+            .groups = "drop")
+
+accessibility_by_age
+```
+
+The accessibility level is the magnitude of the accessibility loss: the top 25% faced the smallest losses in accessibility, and the bottom 25% faced the greatest losses in accessibility.
+
+```{r}
+cbind(accessibility_by_age %>% 
+        filter(timing == "Pre-covid") %>% 
+        select(children:accessibility) %>% 
+        rename(accessibility_pre = accessibility),
+      accessibility_by_age %>% 
+        filter(timing == "During covid") %>% 
+        transmute(accessibility_covid = accessibility)) %>%
+  mutate(change = accessibility_covid - accessibility_pre)
+```
+
+Sanity check:
+```{r}
+accessibility_by_age %>%
+  group_by(timing) %>%
+  summarize(children = sum(children),
+            adults = sum(adults),
+            seniors = sum(seniors),
+            accessibility = sum(accessibility))
+```
+
 ### Pre-COVID Scenario to Meal Services
 
-Get minimum travel times in the pre-covid scenario:
-
+Get weighted minimum travel times in the pre-covid scenario:
 ```{r minimum-travel-time-baseline, include=FALSE}
@@
 Bind the tables and label by mode:
-
 ```{r weighted-travel-time-by-mode-baseline, include=FALSE}
@@
 
+Calculate the weighted person hours of travel by age group:
 ```{r aggregate-weighted-travel-times-baseline, include=FALSE}
@@
             seniors = first(seniors),
+            # Weighted travel time is the sum of the travel times weighted by the proportion of trips by mode
             weighted_travel_time = sum(weighted_travel_time)) %>%
@@
 
+Calculate the weighted person hours of travel by TAZ by population group:
 ```{r summarize-results-by-taz-baseline, include=FALSE}
@@
 
-```{r}
-tm_shape(min_ttm_baseline_taz) +
-  tm_fill("wph_children")
-
-tm_shape(min_ttm_baseline_taz) +
-  tm_fill("wph_adults")
-
-tm_shape(min_ttm_baseline_taz) +
-  tm_fill("wph_seniors")
-```
-
+Plot weighted travel time to nearest facility:
 ```{r weighted-travel-time-map-baseline}
@@
           color = NA) +
-  scale_fill_distiller(name = "Expected travel time (min)",
+  scale_fill_fermenter(name = "Expected travel time (min)",
                        palette = "OrRd", 
@@
 
+Plot the weighted person hours of travel for the population groups in low income households (travel times in minutes):
+```{r}
+tm_shape(min_ttm_baseline_taz) +
+  tm_fill("wph_children")
+
+tm_shape(min_ttm_baseline_taz) +
+  tm_fill("wph_adults")
+
+tm_shape(min_ttm_baseline_taz) +
+  tm_fill("wph_seniors")
+```
+
 ### Do it again for the during-COVID scenario:
 
+Calculate weighted travel time to the nearest facilty by mode:
 ```{r minimum-travel-time-covid, include=FALSE}
-# retrieve minimum travel time during covid to sites that didnt close
+# retrieve minimum travel time during covid to sites that didn't close
 min_ttm_car_covid <- ttm_car %>%
@@
 Bind the tables and label by mode:
-
 ```{r weighted-travel-time-by-mode-covid, include=FALSE}
@@
 
+Calculate the weighted person hours of travel by age group:
 ```{r aggregate-weighted-travel-times-covid, include=FALSE}
@@
 
+Aggregate the number of weighted person hours of travel by age for each TAZ:
 ```{r summarize-results-by-taz-covid, include=FALSE}
@@
 
-```{r}
-tm_shape(min_ttm_covid_taz) +
-  tm_fill("wph_children")
-
-tm_shape(min_ttm_covid_taz) +
-  tm_fill("wph_adults")
-
-tm_shape(min_ttm_covid_taz) +
-  tm_fill("wph_seniors")
-```
-
+Map weighted travel time to nearest facility:
 ```{r weighted-travel-time-map-covid}
@@
 
+```{r}
+tm_shape(min_ttm_covid_taz) +
+  tm_fill("wph_children")
+
+tm_shape(min_ttm_covid_taz) +
+  tm_fill("wph_adults")
+
+tm_shape(min_ttm_covid_taz) +
+  tm_fill("wph_seniors")
+```
+
 ### Map Differences
 
-Join the min travel time tables together
-
+Join the min travel time tables together:
 ```{r}

### 2021-04-22 foodbanksHamilton.Rcheck/00_pkg_src/foodbanksHamilton/R/foodbanksHamilton.R
Created data package

@@ -1,0 +1,231 @@
+#' foodbanksHamilton
+#'
+#' foodbanksHamilton is a companion package for the paper "Changes in accessibility to foodbanks during COVID-19 and implications for the food security of vulnerable populations in Hamilton, Ontario".
+#'
+#' The package includes the data used in the analysis.
+#'
+#' Contains information licensed under the Open Government Licence – Canada (https://open.canada.ca/en/open-government-licence-canada).
+#'
+#' Contains public sector Data made available under the City of Hamilton’s Open Data Licence (https://www.hamilton.ca/city-initiatives/strategies-actions/open-data-licence-terms-and-conditions).
+#'
+#' @author Antonio Paez, \email{paezha@@mcmaster.ca}
+#' @docType package
+#' @name vaccHamilton
+NULL
+
+#' Hamilton CMA Dissemination Areas with Selected Census Information.
+#'
+#' Boundaries of the Dissemination Areas in Hamilton CMA.
+#'
+#' @format A simple features data frame with 882 rows and 38 variables:
+#' \describe{
+#'   \item{Shape Area}{Area of polygon}
+#'   \item{Type}{Geography type}
+#'   \item{Households}{Number of households in DA}
+#'   \item{Dwellings}{Number of dwellings in DA}
+#'   \item{GeoUID}{Unique identifier of the DA}
+#'   \item{CSD_UID}{Census Subdivision (CSD) unique identifier}
+#'   \item{Population}{Population in the DA}
+#'   \item{CT_UID}{Census Tract (CT) unique identifier}
+#'   \item{CD_UID}{Census Division (CD) unique identifier}
+#'   \item{CMA_UID}{Census Metropolitan Area (CMA) unique identifier}
+#'   \item{Region Name}{Name of the region within the CMA}
+#'   \item{Area (sq km)}{Area of the DA in square kilometers}
+#'   \item{v_CA16_2406: Under $5,000}{Households with total household income under $5,000}
+#'   \item{v_CA16_2407: $5,000 and $9,999}{Households with total household income between $5,000 and $9,999}
+#'   \item{v_CA16_2408: $10,000 and $14,999}{Households with total household income between $10,000 and $14,999}
+#'   \item{v_CA16_2409: $15,000 and $19,999}{Households with total household income between $15,000 and $19,999}
+#'   \item{v_CA16_2410: $20,000 and $24,999}{Households with total household income between $20,000 and $24,999}
[…]

### 2021-04-22 foodbanksHamilton/R/foodbanksHamilton.R

@@ -1,0 +1,231 @@
+#' foodbanksHamilton
+#'
+#' foodbanksHamilton is a companion package for the paper "Changes in accessibility to foodbanks during COVID-19 and implications for the food security of vulnerable populations in Hamilton, Ontario".
+#'
+#' The package includes the data used in the analysis.
+#'
+#' Contains information licensed under the Open Government Licence – Canada (https://open.canada.ca/en/open-government-licence-canada).
+#'
+#' Contains public sector Data made available under the City of Hamilton’s Open Data Licence (https://www.hamilton.ca/city-initiatives/strategies-actions/open-data-licence-terms-and-conditions).
+#'
+#' @author Antonio Paez, \email{paezha@@mcmaster.ca}
+#' @docType package
+#' @name foodbanksHamilton
+NULL
+
+#' Hamilton CMA Dissemination Areas with Selected Census Information.
+#'
+#' Boundaries of the Dissemination Areas in Hamilton CMA.
+#'
+#' @format A simple features data frame with 882 rows and 38 variables:
+#' \describe{
+#'   \item{Shape Area}{Area of polygon}
+#'   \item{Type}{Geography type}
+#'   \item{Households}{Number of households in DA}
+#'   \item{Dwellings}{Number of dwellings in DA}
+#'   \item{GeoUID}{Unique identifier of the DA}
+#'   \item{CSD_UID}{Census Subdivision (CSD) unique identifier}
+#'   \item{Population}{Population in the DA}
+#'   \item{CT_UID}{Census Tract (CT) unique identifier}
+#'   \item{CD_UID}{Census Division (CD) unique identifier}
+#'   \item{CMA_UID}{Census Metropolitan Area (CMA) unique identifier}
+#'   \item{Region Name}{Name of the region within the CMA}
+#'   \item{Area (sq km)}{Area of the DA in square kilometers}
+#'   \item{v_CA16_2406: Under $5,000}{Households with total household income under $5,000}
+#'   \item{v_CA16_2407: $5,000 and $9,999}{Households with total household income between $5,000 and $9,999}
+#'   \item{v_CA16_2408: $10,000 and $14,999}{Households with total household income between $10,000 and $14,999}
+#'   \item{v_CA16_2409: $15,000 and $19,999}{Households with total household income between $15,000 and $19,999}
+#'   \item{v_CA16_2410: $20,000 and $24,999}{Households with total household income between $20,000 and $24,999}
[…]

### 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
First chunks of code transferred to manuscript: package installation and loading, and loading data

@@
 **Author 1:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization, Supervision; **Author 2:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization; **Author 3:** Conceptualization, Investigation, Data Curation, Writing - Original Draft; **Author 4:** Resources - Writing: Review & Editing - Supervision
+
+```{r install-data-package, include = FALSE}
+# Run only once if needed to install data package `sobiEquity`
+if (!require("foodbanksHamilton", character.only = TRUE)) {
+      devtools::install_github("https://github.com/paezha/Accessibility-Food-Banks-Hamilton", subdir = "foodbanksHamilton")
+  }
+```
+
+```{r load-packages, include=FALSE}
+library(foodbanksHamilton) # Data package
+library(patchwork)
+library(sf)
+library(tidyverse)
+#library(tmap)
+```
+
+```{r invoke-data, include=FALSE}
+data(data_da_2016)
+data(dwelling_network_points_2016)
+#data(hamilton_boundary)
+data(population_age_income)
+data(foodbank_locations)
+data(modes_less40k)
+data(ttm_car)
+data(ttm_transit)
+data(ttm_walk.RData)
+```
 

### 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Fixed error in loading data

@@
 data(ttm_transit)
-data(ttm_walk.RData)
+data(ttm_walk)
 ```

### 2021-04-22 00-Data-Processing/04-Data-Analysis.Rmd
Update 04-Data-Analysis.Rmd
Removed some comments and duplicate code

@@
 
-Join the household data to the travel time matrices:
-```{r}
-ttm_car <- ttm_car %>%
-  left_join(dwelling_network_points_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            income_less40k,
-                            DAUID, 
-                            TAZUID),
-            by = c("UID" = "ID"))
-
-ttm_transit <- ttm_transit %>%
-  left_join(dwelling_network_points_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            income_less40k,
-                            DAUID, 
-                            TAZUID),
-            by = c("UID" = "ID"))
-
-ttm_walk <- ttm_walk %>%
-  left_join(dwelling_network_points_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            income_less40k,
-                            DAUID, 
-                            TAZUID),
-            by = c("UID" = "ID"))
-```
-
-Join the food bank data to the travel time matrices:
-```{r}
-ttm_car <- ttm_car %>%
-  left_join(foodbank_locations %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            TYPE,
-                            COVIDClose),
-            by = c("OBJECTID" = "ID"))
-
-ttm_transit <- ttm_transit %>%
-  left_join(foodbank_locations %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            TYPE,
-                            COVIDClose),
-            by = c("OBJECTID" = "ID"))
-
-ttm_walk <- ttm_walk %>%
-  left_join(foodbank_locations %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID,
-                            TYPE,
-                            COVIDClose),
-            by = c("OBJECTID" = "ID"))
-```
-
-Drop NAs:
-```{r}
-ttm_car <- ttm_car %>%
-  drop_na(TAZUID,
-          income_less40k)
-
-ttm_transit <- ttm_transit %>%
-  drop_na(TAZUID,
-          income_less40k)
-
-ttm_walk <- ttm_walk %>%
-  drop_na(TAZUID,
-          income_less40k)
-```
-
-Calculate proportion of trips by mode per TAZ:
-```{r}
-modes <- modes_less40k %>%
-  mutate(p_car = Driver/(Driver + Transit + Walk),
-         p_transit = Transit/(Driver + Transit + Walk),
-         p_walk = Walk/(Driver + Transit + Walk))
-```
-
-Assume that TAZ with NAs have a proportion of car trips of one:
-```{r}
-modes <- modes %>%
-  dplyr::mutate(p_car = replace_na(p_car, 1),
-                p_transit = replace_na(p_transit, 0),
-                p_walk = replace_na(p_walk, 0))
-```
-
-```{r}
-# join tripmaking modal proportions
-ttm_car <- ttm_car %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-
-ttm_transit <- ttm_transit %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-
-ttm_walk <- ttm_walk %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-```
-
-Prepare and join population data:
+Prepare and join population data from TTS:
 ```{r number-of-residential-units-per-DA, include=FALSE}
@@
 
-```{r join-population-data-to-travel-time-tables, include=FALSE}
-# Data preparation
-## Join the population data to the travel time matrices:
+Join the household data to the travel time matrices:
+```{r}
 ttm_car <- ttm_car %>%
@@
               st_drop_geometry() %>%
-              dplyr::transmute(ID, 
+              dplyr::select(ID, 
+                            income_less40k,
                             children,
                             adults,
-                            seniors),
+                            seniors,
+                            DAUID, 
+                            TAZUID),
             by = c("UID" = "ID"))
@@
 
+Join the food bank data to the travel time matrices:
+```{r}
+ttm_car <- ttm_car %>%
+  left_join(foodbank_locations %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID, 
+                            TYPE,
+                            COVIDClose),
+            by = c("OBJECTID" = "ID"))
+
+ttm_transit <- ttm_transit %>%
+  left_join(foodbank_locations %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID, 
+                            TYPE,
+                            COVIDClose),
+            by = c("OBJECTID" = "ID"))
+
+ttm_walk <- ttm_walk %>%
+  left_join(foodbank_locations %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID,
+                            TYPE,
+                            COVIDClose),
+            by = c("OBJECTID" = "ID"))
+```
+
+Drop NAs:
+```{r}
+ttm_car <- ttm_car %>%
+  drop_na(TAZUID,
+          income_less40k)
+
+ttm_transit <- ttm_transit %>%
+  drop_na(TAZUID,
+          income_less40k)
+
+ttm_walk <- ttm_walk %>%
+  drop_na(TAZUID,
+          income_less40k)
+```
+
+Calculate proportion of trips by mode per TAZ:
+```{r}
+modes <- modes_less40k %>%
+  mutate(p_car = Driver/(Driver + Transit + Walk),
+         p_transit = Transit/(Driver + Transit + Walk),
+         p_walk = Walk/(Driver + Transit + Walk))
+```
+
+Assume that TAZ with NAs have a proportion of car trips of one:
+```{r}
+modes <- modes %>%
+  dplyr::mutate(p_car = replace_na(p_car, 1),
+                p_transit = replace_na(p_transit, 0),
+                p_walk = replace_na(p_walk, 0))
+```
+
+```{r}
+# join tripmaking modal proportions
+ttm_car <- ttm_car %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+
+ttm_transit <- ttm_transit %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+
+ttm_walk <- ttm_walk %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+```
+
+### Join Tables into Single TTM
 Join all tables:
@@
 ```{r}
-# acc_quintiles_pre <- quantile(accessibility %>%
-#                                 filter(timing == "Pre-covid") %>%
-#                                 pull(accessibility),
-#                               c(0, 0.25, 0.5, 0.75, 1))
-# acc_quintiles_covid <- quantile(accessibility %>%
-#                                 filter(timing == "During covid") %>%
-#                                 pull(accessibility),
-#                               c(0, 0.25, 0.5, 0.75, 1))
+ acc_quintiles_pre <- quantile(accessibility %>%
+                                 filter(timing == "Pre-covid") %>%
+                                 pull(accessibility),
+                               c(0, 0.25, 0.5, 0.75, 1))
+ acc_quintiles_covid <- quantile(accessibility %>%
+                                 filter(timing == "During covid") %>%
+                                 pull(accessibility),
+                               c(0, 0.25, 0.5, 0.75, 1))
 ```
@@
 ```{r}
-acc_quintiles_change <- quantile(accessibilty_change$acc_d,
+acc_quintiles_change <- quantile(accessibility_change$acc_d,
                               c(0, 0.25, 0.5, 0.75, 1))
@@
 ```{r}
-# accessibility <- accessibility %>%
-#   mutate(accessibility_level = case_when(timing == "Pre-covid" & accessibility <= acc_quintiles_pre[2] ~ "Bottom 25%",
-#                                          timing == "Pre-covid" & accessibility > acc_quintiles_pre[2] & accessibility <= acc_quintiles_pre[3] ~ "Third 25%",
-#                                          timing == "Pre-covid" & accessibility > acc_quintiles_pre[3] & accessibility <= acc_quintiles_pre[4] ~ "Second 25%",
-#                                          timing == "Pre-covid" & accessibility > acc_quintiles_pre[4] ~ "Top 25%",
-#                                          timing == "During covid" & accessibility <= acc_quintiles_covid[2] ~ "Bottom 25%",
-#                                          timing == "During covid" & accessibility > acc_quintiles_covid[2] & accessibility <= acc_quintiles_covid[3] ~ "Third 25%",
-#                                          timing == "During covid" & accessibility > acc_quintiles_covid[3] & accessibility <= acc_quintiles_covid[4] ~ "Second 25%",
-#                                          timing == "During covid" & accessibility > acc_quintiles_covid[4] ~ "Top 25%"),
-#            accessibility_level = factor(accessibility_level,
-#                                         levels = c("Top 25%", "Second 25%", "Third 25%", "Bottom 25%"),
-#                                         ordered = TRUE))
+ accessibility <- accessibility %>%
+   mutate(accessibility_level = case_when(timing == "Pre-covid" & accessibility <= acc_quintiles_pre[2] ~ "Bottom 25%",
+                                          timing == "Pre-covid" & accessibility > acc_quintiles_pre[2] & accessibility <= acc_quintiles_pre[3] ~ "Third 25%",
+                                          timing == "Pre-covid" & accessibility > acc_quintiles_pre[3] & accessibility <= acc_quintiles_pre[4] ~ "Second 25%",
+                                          timing == "Pre-covid" & accessibility > acc_quintiles_pre[4] ~ "Top 25%",
+                                          timing == "During covid" & accessibility <= acc_quintiles_covid[2] ~ "Bottom 25%",
+                                          timing == "During covid" & accessibility > acc_quintiles_covid[2] & accessibility <= acc_quintiles_covid[3] ~ "Third 25%",
+                                          timing == "During covid" & accessibility > acc_quintiles_covid[3] & accessibility <= acc_quintiles_covid[4] ~ "Second 25%",
+                                          timing == "During covid" & accessibility > acc_quintiles_covid[4] ~ "Top 25%"),
+            accessibility_level = factor(accessibility_level,
+                                         levels = c("Top 25%", "Second 25%", "Third 25%", "Bottom 25%"),
+                                         ordered = TRUE))
 ```

### 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Updates to literature review

@@
 Although previous work has explored differences in _accessing_ food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource [e.g., @enns2020experiences], we are not aware of any research that has focused on the geographic component of accessibility.
-**Check @tarasuk2019geographic**
 
@@
 
-Literature Review
+Food Insecurity and Food Bank Use in Canada
 ============
 
-## Food Insecurity
-
-Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake [@kirkpatrick2008food; @tarasuk2009household].
-This nutrient deficiency has been causing major health concerns in Canadians, and particularly those who are at a socioeconomic disadvantage [@bazerghi2016role].
-Previous studies have aimed to look at the relationship between the built food environment and socio-demographic characteristics with qualitative and quantitative, or a mixed -method approach.
-Quantitatively, official governmental surveys have been able to assist with data, such as the Household Food Security Survey Module (HFSSM), the Canadian Community Health Surveys (CCHS), the Longitudinal and International Study of Adults (LISA), and official classifications determined by Health Canada in relation to  sociodemographic variables [@kirkpatrick2008food; @tarasuk2009household; @gundersen2018food].
-Studies have also aimed to assess food availability of healthy foods (e.g., fruit and vegetables) at supermarkets in relation to sociodemographic characteristics and geographic accessibility [@latham2007determinants].
-In terms of findings, studies have generated inconsistent relationships between their evaluated availability of food and sociodemographic characteristics that go in hand.
-
-## Foodbanks
+Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake [@enns2020experiences; @kirkpatrick2008food; @tarasuk2009household].
+This nutrient deficiency is associated with major health concerns in Canadians, and particularly those who are at a socioeconomic disadvantage [@bazerghi2016role].
+Data on food insecurity in Canada is collected in several ways.
+Quantitatively, official government surveys such as the Household Food Security Survey Module (HFSSM), the Canadian Community Health Surveys (CCHS), the International Study of Adults (LISA), and official classifications determined by Health Canada in relation to sociodemographic variables offer some insight into food insecurity [@kirkpatrick2008food; @tarasuk2009household; @gundersen2018food].
+Nationally, it has been found that food insecurity impacts approximately 12.3% of Canadian households [@tarasuk2014food]
 
@@
 Although in their origin foodbanks were meant to be provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time they evolved into a community practice to secure emergency food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
-In Canada, the number of foodbanks has steadily increased in the past few decades [@wakefield2013sweet].
-Previous studies have questioned whether food banks are able to offer a rounded nutritious supply of food, and if food banks are a sustainable on-going practice for those in need [@riches2002food; @bazerghi2016role].
-An inescapable reality, at the moment, is that despite their imperfections they provide urgently needed support for members of the population for whom foodbanks are frequently their primary location to get food [@bazerghi2016role].
-**However, surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada [@tarasuk2014food].** **Something about geographical access**
 
-## Food Insecurity in Canada during the COVID-19
+In Canada, the number of food banks has steadily increased in the past few decades [@wakefield2013sweet].
+The largest database of food banks and their use comes from the non-profit association Food Banks Canada, which conducts an annual assessment through its affiliated members.
+Food Banks Canada's 2018 Hunger Count report^[https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018_FINAL_EN.pdf.aspx?ext=.pdf] (the most recent available) listed 1,830 member food banks across the country found that Canadians visited food banks 1.1 million times in March of 2018.
+Of those accessing food banks, certain population characteristics tend to be over-represented compared to national totals from the 2016 Canadian Census of Population.
+According to Food Banks Canada's 2018 data, single-adult households represent 45% of those utilizing food banks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those accessing food bank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
+In addition, 59% of households accessing food banks list social or disability assistance as their primary source of income.
+Moreover, children aged 0-18 are over-represented as food bank users, making up 35% of those accessing food bank services despite representing 20% of the country's population.
+Likewise, using data from the 2011-2012 CCHS, Tarasuk et al.
+[-@tarasuk2019geographic] found higher odds of food insecurity amongst households relying on social assistance, those without a university degree or with children under the age of 18, and individuals that lived alone, renters, and those identifying as Aboriginal.
+While surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada [@tarasuk2014food], research from Ottawa [@enns2020experiences] and Vancouver [@black2020examining] suggests that long-term users tend to be older, have health or mobility challenges, live in large households, and less likely to have employment income.
 
 Currently with the COVID-19 pandemic, disrupted economies, rising unemployment rates, and changes in poverty levels have disrupted the food environment and led to higher rates of food insecurity [@niles2020early].
-In 2012, 12.4% of Canadians households and 11.8% Ontarian households experienced some degree of food insecurity [@gundersen2018food; @tarasuk2020relationship].
 Pandemic-related food security studies in the US have found a substantial increase in households experiencing food insecurity for the first time, and also in households experiencing more severe food insecurity than before [@niles2020early; @wolfson2020food].
 Most recently in May 2020, Canada recorded 14.7% of its population living in food insecurity in the past 30 days [@statisticscanada2020food].
-Food insecurity is a highly concerning public health issue due to its health impacts.
-Recent research suggests that food insecurity in adults can lead to experiencing more stressful events (El-hajj and Benhin 2021).
-Increases in food insecurity rates due to the pandemic, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
+Considering the negative mental and physical health effects associated with food insecurity, increases in food insecurity rates due to the pandemic, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
+Recent data^[https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis_Exec-Sum_EN.pdf] from Food Banks Canada showed that 52% of member food banks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of Canada.
+The pandemic also created significant staffiing issues with 42% of food banks reporting a reduction in volunteers.
+However, 53% of food banks reported a decrease in use into the summer of 2020 which food banks attributed to emergency financial support programs from the federal government were.
+Nevertheless, some of these benefit programs were temporary, suggesting many households may again turn to food bank services to meet their needs.
+
+In terms of geography, previous research conducted at the provincial scale using data from the 2011-2012 CCHS found that the prevalence of food insecurity ranged across the country from 11.8% of households in Ontario to 41% of households in Nunavut [@tarasuk2019geographic].
+However, although previous research has examined the characteristics of individuals and households accessing food banks, the locational or transportation accessibility aspect of food bank access is not well understood.
+A wealth of literature examining the food desert concept suggests that, in addition to socioeconomic and demographic factors, location and transportation networks play a key role in a household's accessibility to healthy foods [@paez2010relative; @vanderlee2017food; @widener2018spatial].
+For food banks specifically, previous qualitative research in Ontario has noted difficulties associated with transportation for trips to the food bank [@smith2017food].
+For example, Smith et al.
+[-@smith2017food] note that participants access food banks using a variety of modes of transportation with respondent stating "I wish it [the food bank] was a little more centrally located.
+Because if I didn’t have a bike I’d have to walk it all the way out there and back.
+I wonder about people who don’t." From this, the authors conclude that "transportation can be challenging, particularly if the food bank is situated in a remote location." To offer greater insight into the role of transportation and location in food bank accessibility, this research examines how geographic access to food banks and food services changed in Hamilton before and during the COVID-19 pandemic.
 

### 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Started migrating analysis to manuscript.

--- 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
@@
 
+```{r, setup, include=FALSE}
+knitr::opts_chunk$set(
+  cache = TRUE, out.width = "90%", fig.width = 6, fig.height = 6
+)
+```
+
 ```{r install-data-package, include = FALSE}
@@
 library(foodbanksHamilton) # Data package
+library(kableExtra)
 library(patchwork)
@@
 data(dwelling_network_points_2016)
-#data(hamilton_boundary)
+data(foodbank_locations)
+data(hamilton_cma)
+data(modes_less40k)
 data(population_age_income)
-data(foodbank_locations)
-data(modes_less40k)
 data(ttm_car)
@@
 
+```{r remove-irrelevant-destinations, include=FALSE}
+## Remove entries in travel time matrices for inapplicable destinations
+# As a first step, we will remove the computed travel times to the three Community Kitchen locations (IDs 1-3):
+ttm_car <- ttm_car %>%
+  filter(!between(OBJECTID, left = 1, right = 3))
+
+ttm_transit <- ttm_transit %>%
+  filter(!between(OBJECTID, left = 1, right = 3))
+
+ttm_walk <- ttm_walk %>%
+  filter(!between(OBJECTID, left = 1, right = 3))
+```
+
+```{r join-low-income-households, include=FALSE}
+## Join proportion of low income households per DA to dwellings and total number of residential units:
+dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
+  left_join(data_da_2016 %>%
+              st_drop_geometry() %>%
+              dplyr::transmute(DAUID = GeoUID, 
+                               p_inc_less40k),
+            by = "DAUID")
+```
+
+```{r probability-of-low-income-per-parcel, include=FALSE}
+## Calculate probability that a point on the network will have low income households.
+This is the proportion of low income households in the DA multiplied by the total number of residential units at the point:
+dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
+  mutate(income_less40k = p_inc_less40k * Dwellings)
+```
+
+```{r number-of-residential-units-per-DA, include=FALSE}
+## Data preparation
+# Prepare and join population data from Transportation Tomorrow Survey
+# Find number of residential units in each DA. This is needed to assign the population to parcels
+residential_units <- dwelling_network_points_2016 %>%
+  st_drop_geometry() %>%
+  group_by(TAZUID) %>%
+  summarize(residential_units = sum(Dwellings))
+```
+
+```{r join-population-to-parcels, include=FALSE}
+## Data preparation
+# Join number of low income households per DA to dwellings and total number of residential units:
+dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
+  left_join(population_age_income %>%
+              st_drop_geometry() %>%
+              dplyr::transmute(TAZUID,
+                               children,
+                               adults,
+                               seniors),
+            by = "TAZUID") %>%
+  left_join(residential_units,
+            by = "TAZUID")
+```
+
+```{r calculate-low-income-households-per-parcel, include=FALSE}
+## Data preparation
+# Calculate population per point on the network.
+This is the population in the DA divided by the total number of residential units.
+This gives the average number of people per residential unit.
+Then, multiply by the number of dwellings (residential units) at the point:
+dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
+  mutate(children = (children/residential_units) * Dwellings,
+         adults = (adults/residential_units) * Dwellings,
+         seniors = (seniors/residential_units) * Dwellings)
+```
+
+```{r join-household-data-to-travel-time-tables, include=FALSE}
+## Data preparation
+# Join the household data to the travel time matrices:
+ttm_car <- ttm_car %>%
+  left_join(dwelling_network_points_2016 %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID, 
+                            income_less40k,
+                            children,
+                            adults,
+                            seniors,
+                            DAUID, 
+                            TAZUID),
+            by = c("UID" = "ID"))
+
+ttm_transit <- ttm_transit %>%
+  left_join(dwelling_network_points_2016 %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID, 
+                            income_less40k,
+                            children,
+                            adults,
+                            seniors,
+                            DAUID, 
+                            TAZUID),
+            by = c("UID" = "ID"))
+
+ttm_walk <- ttm_walk %>%
+  left_join(dwelling_network_points_2016 %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID, 
+                            income_less40k,
+                            children,
+                            adults,
+                            seniors,
+                            DAUID, 
+                            TAZUID),
+            by = c("UID" = "ID"))
+```
+
+```{r join-foodbank-data-to-travel-time-tables, include=FALSE}
+## Data preparation
+# Join the food bank data to the travel time tables:
+ttm_car <- ttm_car %>%
+  left_join(foodbank_locations %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID, 
+                            TYPE,
+                            COVIDClose),
+            by = c("OBJECTID" = "ID"))
+
+ttm_transit <- ttm_transit %>%
+  left_join(foodbank_locations %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID, 
+                            TYPE,
+                            COVIDClose),
+            by = c("OBJECTID" = "ID"))
+
+ttm_walk <- ttm_walk %>%
+  left_join(foodbank_locations %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID,
+                            TYPE,
+                            COVIDClose),
+            by = c("OBJECTID" = "ID"))
+```
+
+```{r drop-nas-from-travel-time-tables, include=FALSE}
+## Data preparation
+# Drop NAs:
+ttm_car <- ttm_car %>%
+  drop_na(TAZUID,
+          income_less40k)
+
+ttm_transit <- ttm_transit %>%
+  drop_na(TAZUID,
+          income_less40k)
+
+ttm_walk <- ttm_walk %>%
+  drop_na(TAZUID,
+          income_less40k)
+```
+
+```{r proportion-of-trips-by-mode-by-taz, include=FALSE}
+## Data preparation
+# Calculate proportion of trips by mode per TAZ:
+modes <- modes_less40k %>%
+  mutate(p_car = Driver/(Driver + Transit + Walk),
+         p_transit = Transit/(Driver + Transit + Walk),
+         p_walk = Walk/(Driver + Transit + Walk))
+```
+
+```{r fill-nas-transport-modes, include=FALSE}
+## Data preparation
+# Assume that TAZ with NAs have a proportion of car trips of one:
+modes <- modes %>%
+  dplyr::mutate(p_car = replace_na(p_car, 1),
+                p_transit = replace_na(p_transit, 0),
+                p_walk = replace_na(p_walk, 0))
+```
+
+```{r join-trips-by-mode-to-travel-time-tables, include=FALSE}
+## Data preparation
+# Join trip-making modal proportions
+ttm_car <- ttm_car %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+
+ttm_transit <- ttm_transit %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+
+ttm_walk <- ttm_walk %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+```
+
+```{r join-travel-time-tables, include=FALSE}
+## Data preparation
+# Join Tables into Single TTM
+ttm <- ttm_car %>%
+  rename(travel_time_car = travel_time) %>%
+  full_join(ttm_transit %>%
+              transmute(UID,
+                        OBJECTID,
+                        travel_time_transit = travel_time),
+            by = c("UID", "OBJECTID")) %>%
+  full_join(ttm_walk %>%
+              transmute(UID,
+                        OBJECTID,
+                        travel_time_walk = travel_time),
+            by = c("UID", "OBJECTID"))
+```
+
+```{r travel-time-tables-data-check, include=FALSE}
+## Data preparation
+# Check the NAs in travel time table:
+ttm %>% 
+  filter(is.na(travel_time_transit)) %>%
+  select(travel_time_transit,
+         p_transit) %>%
+  summary()
+
+# What do NAs here mean?
+Most DAs missing travel time by transit do not report trips by transit in the TTS.
+In some DAs there are trips by transit (p_transit != 0), but no foodbank can be reached by transit (travel_time_transit is NA).
+How many of these cases?
+ttm %>% 
+  filter(is.na(travel_time_transit) & p_transit > 0) %>%
+  select(travel_time_transit,
+         p_transit) %>%
+  summary()
+
+# It should be fine to replace the NAs with zeros, which multiplied by p_transit will return a zero (i.e., it is not possible to reach a foodbank by transit).
+ttm <- ttm %>% 
+  mutate(travel_time_transit = replace_na(travel_time_transit, 0))
+
+# Check the NAs in travel time by walk:
+ttm %>% 
+  filter(is.na(travel_time_walk)) %>%
+  select(travel_time_walk,
+         p_walk) %>%
+  summary()
+
+# Most DAs missing travel time by walking do not report trips by walking in the TTS.
+In some DAs there are trips by walking (p_walk != 0), but no foodbank can be reached by walk (travel_time_walk is NA).
+How many of these cases?
+ttm %>% 
+  filter(is.na(travel_time_walk) & p_walk > 0) %>%
+  select(travel_time_walk,
+         p_walk) %>%
+  summary()
+
+# Like with transit, it should be fine to replace the NAs with zeros, which multiplied by p_walk will return a zero (i.e., it is not possible to reach a foodbank by walking).
+ttm <- ttm %>% 
+  mutate(travel_time_walk = replace_na(travel_time_walk, 0))
+
+#Summary of the table after these operations:
+summary(ttm)
+```
+
+```{r calculate-expected-travel-times, include=FALSE}
+## Data preparation
+# Calculate expected travel times by mode.
+These are the travel time by the mode multiplied by the probability of using the mode (which is proxied by the proportion of trips by mode in the DA from the TTS).
+The total expected travel time is the sum of the expected travel times by mode:
+ttm <- ttm %>%
+  mutate(e_travel_time_car = travel_time_car * p_car,
+         e_travel_time_transit = travel_time_transit * p_transit,
+         e_travel_time_walk = travel_time_walk * p_walk,
+         e_travel_time = e_travel_time_car + e_travel_time_transit + e_travel_time_walk)
+```
+
 Words go here.
+
+```{r accessibility-threshold-sensitivity, include=FALSE}
+## Analysis
+# Testing different thresholds to check how accessibility changes. Here, the thresholds are varied from 2 min to 30 min in 2 min step increments
+threshold_seq <- seq(from = 2, to = 30, by = 2)
+
+# Initialize a data frame to store results
+accessibility_test <- data.frame(threshold = threshold_seq, 
+                                 accessibility = numeric(length(threshold_seq)))
+
+# Calculate the system-wide accessibility according to each threshold
+for(i in 1:length(threshold_seq)){
+  results_test <-  ttm %>% 
+  transmute(UID, OBJECTID, travel_time = e_travel_time, population = income_less40k, supply = 1) %>%
+    b2sfca(threshold = i )
+  accessibility_test$accessibility[i] <- sum(results_test$accessibility$accessibility)
+}
+```
+
+```{r plot-results-sensitivity-analysis, echo=FALSE, fig.cap="\\label{fig:sensitivity-analysis}Accessibility as a function of threshold"}
+acc_sensitivity <- ggplot(accessibility_test, 
+       aes(x = threshold, y = accessibility)) +
+  geom_line() +
+  ylab("Accessibility") +
+  xlab("Threshold (in min)") +
+  theme_minimal()
+
+acc_sensitivity
+```
+
+```{r 30-min-pre-covid, include=FALSE}
+## Analysis 
+# Calculate 30-min accessibility pre-covid:
+results_30min_pre <- ttm %>%
+  #dplyr::filter(COVIDClose == "No") %>%
+  transmute(UID, 
+            OBJECTID, 
+            travel_time = e_travel_time, 
+            population = income_less40k, 
+            supply = 1) %>%
+  b2sfca(threshold = 30)
+```
+
+```{r 30-min-during-covid, include=FALSE}
+## Analysis 
+# Calculate 30-min accessibility during covid:
+results_30min_covid <- ttm %>%
+  dplyr::filter(COVIDClose == "No") %>%
+  transmute(UID, 
+            OBJECTID, 
+            travel_time = e_travel_time, 
+            population = income_less40k, 
+            supply = 1) %>%
+  b2sfca(threshold = 30)
+```
+
+```{r, bind-accessibility-results, include=FALSE}
+## Analysis
+# Bind level of service results:
+los <- rbind(data.frame(results_30min_pre$los, 
+                        timing = "Pre-covid", 
+                        threshold = "30 min"),
+             data.frame(results_30min_covid$los, 
+                        timing = "During covid", 
+                        threshold = "30 min")) %>%
+  mutate(timing = factor(timing, 
+                         levels = c("Pre-covid", 
+                                    "During covid"),
+                         ordered = TRUE))
+
+# Bind accessibility results:
+accessibility <- rbind(data.frame(results_30min_pre$accessibility, 
+                                  timing = "Pre-covid", 
+                                  threshold = "30 min"),
+                       data.frame(results_30min_covid$accessibility, 
+                                  timing = "During covid", 
+                                  threshold = "30 min")) %>%
+  mutate(timing = factor(timing, 
+                         levels = c("Pre-covid", 
+                                    "During covid"),
+                         ordered = TRUE))
+```
+
+```{r join-zonal-ids-to-accessibility-table, include=FALSE}
+## Analysis
+# Join zonal identifiers to accessibility results:
+accessibility <- accessibility %>%
+  left_join(dwelling_network_points_2016 %>%
+            st_drop_geometry() %>%
+              dplyr::select(ID,
+                            DAUID,
+                            TAZUID,
+                            children,
+                            adults,
+                            seniors),
+            by = c("UID" = "ID")) %>%
+  left_join(population_age_income %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID),
+            by = "TAZUID") 
+```
+
+```{r accessibility-change-by-parcel, include=FALSE}
+## Analysis
+# Calculate accessibility change by parcel:
+accessibility_change <- accessibility %>%
+  select(UID, timing, accessibility) %>%
+  pivot_wider(names_from = timing, 
+              values_from = accessibility,
+              values_fill = 0) %>%
+  # Calculate changes in accessibility
+  mutate(acc_d = `During covid` - `Pre-covid`)
+```
+
+```{r join-accessibility-change-to-table, include=FALSE}
+## Analysis
+# Join accessibility change to accessibility table:
+accessibility <- accessibility %>%
+  left_join(accessibility_change %>%
+              select(UID, acc_d),
+            by = "UID")
+```
+
+```{r quartiles-of-accessibility-change, include=FALSE}
+## Analysis
+# Quartiles of accessibility change:
+acc_quintiles_change <- quantile(accessibility_change$acc_d,
+                              c(0, 0.25, 0.5, 0.75, 1))
+```
+
+```{r accessibility-change-labels, include=FALSE}
+## Analysis
+# Add labels of accessibility loss to table:
+accessibility <- accessibility %>%
+  mutate(accessibility_level = case_when(acc_d <= acc_quintiles_change[2] ~ "Bottom 25%",
+                                         acc_d > acc_quintiles_change[2] & acc_d <= acc_quintiles_change[3] ~ "Third 25%",
+                                         acc_d > acc_quintiles_change[3] & acc_d <= acc_quintiles_change[4] ~ "Second 25%",
+                                         acc_d > acc_quintiles_change[4] ~ "Top 25%"),
+           accessibility_level = factor(accessibility_level,
+                                        levels = c("Top 25%", "Second 25%", "Third 25%", "Bottom 25%"),
+                                        ordered = TRUE))
+```
+
+```{r aggregate-accessibility-by-age-group, include=FALSE}
+## Analysis
+# Calculate accessibility by age group and accessibility level:
+accessibility_by_age <- accessibility %>%
+  group_by(timing, 
+           accessibility_level) %>%
+  summarize(children = sum(children),
+            adults = sum(adults),
+            seniors = sum(seniors),
+            accessibility = sum(accessibility),
+            .groups = "drop")
+# The accessibility level is the magnitude of the accessibility loss: the top 25% faced the smallest losses in accessibility, and the bottom 25% faced the greatest losses in accessibility. 
+```
+
+```{r table-accessibility-by-age-group, echo=FALSE, results='asis'}
+cbind(accessibility_by_age %>% 
+        filter(timing == "Pre-covid") %>% 
+        select(children:accessibility) %>% 
+        rename(accessibility_pre = accessibility),
+      accessibility_by_age %>% 
+        filter(timing == "During covid") %>% 
+        transmute(accessibility_covid = accessibility)) %>%
+  mutate(children = round(children) %>% 
+           prettyNum(big.mark = ","),
+         adults = round(adults) %>% 
+           prettyNum(big.mark = ","),
+         seniors = round(seniors) %>% 
+           prettyNum(big.mark = ","),
+         change = (accessibility_covid - accessibility_pre) %>% 
+           round(5),
+         accessibility_pre = round(accessibility_pre, 5),
+         accessibility_covid = round(accessibility_covid, 5)) %>%
+  kable("latex",
+        booktabs = TRUE,
+        align = ("cccccc"),
+        col.names = c("Children (age $\\le$ 18)", 
+                      "Adults (19-64)", 
+                      "Seniors (age $\\ge$ 65)", 
+                      "Before COVID-19", 
+                      "During COVID-19", 
+                      "Difference"),
+        caption = "\\label{tab:accessibility-by-age}Accessibility by age group among members of low income households.",
+        escape = FALSE) %>%
+  kable_styling(latex_options = c("striped", "scale_down")) %>%
+  add_header_above(c("Population" = 3, "Accessibility" = 3)) %>%
+  footnote(general = "Population values have been rounded")
+```
+
+```{r, join-geometry-to-level-of-service, include=FALSE}
+## Analysis 
+# Join geometry:
+los <- los %>%
+  left_join(foodbank_locations %>%
+              dplyr::select(ID),
+            by = c("OBJECTID" = "ID")) %>%
+  st_as_sf()
+```
+
+```{r plot-levels-of-service, echo=FALSE, fig.cap="\\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (middle panel).
+The bottom panel shows the change in level of service."}
+## Analysis
+# Plot levels of service before and during covid
+
+# Pre-covid
+los_pre <- ggplot() +
+  geom_sf(data = hamilton_cma) +
+  geom_sf(data = los %>%
+            filter(timing == "Pre-covid"),
+          aes(color = los,
+              size = los)) +
+  geom_sf(data = foodbank_locations,
+          aes(shape = COVIDClose)) +
+  scale_color_distiller(palette = "OrRd", 
+                        direction = 1) +
+  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
+  theme_minimal()
+
+# During covid
+los_covid <- ggplot() +
+  geom_sf(data = hamilton_cma) +
+  geom_sf(data = los %>%
+            filter(timing == "During covid"),
+          aes(color = los,
+              size = los)) +
+  geom_sf(data = foodbank_locations,
+          aes(shape = COVIDClose)) +
+  scale_color_distiller(palette = "OrRd", 
+                        direction = 1) +
+  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
+  theme_minimal()
+
+# This figure shows the places where the level of service deteriorated most (excluding the places that closed, where it vanished).
+los_change <- los %>%
+  st_drop_geometry() %>%
+  pivot_wider(names_from = timing, 
+              values_from = los, 
+              values_fill = 0) %>%
+  filter(`During covid` > 0) %>%
+  mutate(los_d = `During covid` / `Pre-covid` - 1) %>%
+  left_join(foodbank_locations %>%
+            transmute(OBJECTID = ID,
+                      geometry),
+          by = "OBJECTID") %>%
+  st_as_sf() %>%
+  ggplot() +
+  geom_sf(data = hamilton_cma) +
+  geom_sf(aes(color = los_d,
+              size = los_d)) +
+  geom_sf(data = foodbank_locations,
+          aes(shape = COVIDClose)) +
+  scale_color_distiller(palette = "OrRd", 
+                        direction = -1) +
+  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
+  theme_minimal()
+
+los_pre / los_covid / los_change
+```
+
+```{r aggregate-accessibility-by-taz, include=FALSE}  
+## Analysis
+# Accessibility by TAZ:
+accessibility_taz <- accessibility %>%
+  group_by(TAZUID, 
+           timing) %>%
+  summarize(accessibility = sum(accessibility),
+            timing = first(timing),
+            .groups = "drop") %>%
+  left_join(population_age_income %>%
+              transmute(TAZUID,
+                        children,
+                        adults,
+                        seniors),
+            by = "TAZUID") %>%
+  st_as_sf()
+```
+
+```{r plot-accessibility, echo=FALSE, fig.align = 'center', out.width = "90%", fig.cap="\\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (middle panel).
+The bottom panel shows the change in accessibility."}
+## Analysis
+# Plot accessibility before and during covid:
+acc_pre <- ggplot() +
+  geom_sf(data = hamilton_cma) +
+  geom_sf(data = accessibility_taz %>%
+            filter(timing == "Pre-covid"),
+          aes(fill = accessibility)) +
+  scale_fill_distiller(palette = "OrRd", 
+                        direction = 1) +
+  theme_minimal()
+
+acc_covid <- ggplot() +
+  geom_sf(data = hamilton_cma) +
+  geom_sf(data = accessibility_taz %>%
+            filter(timing == "During covid"),
+          aes(fill = accessibility)) +
+  scale_fill_distiller(palette = "OrRd", 
+                        direction = 1) +
+  theme_minimal()
+
+# Calculate accessibility change by TAZ:
+accessibilty_taz_change <- accessibility_taz %>%
+  st_drop_geometry() %>%
+  pivot_wider(names_from = timing, 
+              values_from = accessibility) %>%
+  # Calculate changes in accessibility
+  mutate(acc_d = `During covid` - `Pre-covid`) %>%
+  # Join TAZ geometry
+  left_join(population_age_income %>%
+            transmute(TAZUID,
+                      geometry),
+          by = "TAZUID") %>%
+  # Convert to simple features
+  st_as_sf()
+
+# Plot
+acc_change <- ggplot() +
+  geom_sf(data = hamilton_cma) +
+  geom_sf(data = accessibilty_taz_change,
+          aes(fill = acc_d),
+          color = NA) +
+  geom_sf(data = population_age_income,
+          fill = NA) +
+  scale_fill_fermenter(palette = "OrRd", 
+                       direction = -1) +
+  theme_minimal()
+
+acc_pre / acc_covid / acc_change
+```
 

### 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Started migrating analysis to manuscript.

@@
 % Pandoc header
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-results-sensitivity-analysis-1} \caption{\label{fig:sensitivity-analysis}Accessibility as a function of threshold}\label{fig:plot-results-sensitivity-analysis}
+\end{figure}
+
+\begin{table}
+
+\caption{\label{tab:table-accessibility-by-age-group}\label{tab:accessibility-by-age}Accessibility by age group among members of low income households.}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{cccccc}
+\toprule
+\multicolumn{3}{c}{Population} & \multicolumn{3}{c}{Accessibility} \\
+\cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){4-6}
+Children (age $\le$ 18) & Adults (19-64) & Seniors (age $\ge$ 65) & Before COVID-19 & During COVID-19 & Difference\\
+\midrule
+\cellcolor{gray!6}{7,015} & \cellcolor{gray!6}{21,586} & \cellcolor{gray!6}{9,158} & \cellcolor{gray!6}{0.00734} & \cellcolor{gray!6}{0.00438} & \cellcolor{gray!6}{-0.00296}\\
+3,490 & 13,598 & 6,618 & 0.01386 & 0.00812 & -0.00575\\
+\cellcolor{gray!6}{966} & \cellcolor{gray!6}{5,180} & \cellcolor{gray!6}{4,310} & \cellcolor{gray!6}{0.01475} & \cellcolor{gray!6}{0.00849} & \cellcolor{gray!6}{-0.00625}\\
+2,384 & 7,725 & 5,595 & 0.01853 & 0.01079 & -0.00775\\
+\bottomrule
+\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
+\multicolumn{6}{l}{\rule{0pt}{1em}Population values have been rounded}\\
+\end{tabular}}
+\end{table}
+
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-levels-of-service-1} \caption{\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (middle panel).
+The bottom panel shows the change in level of service.}\label{fig:plot-levels-of-service}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.9\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-accessibility-1}
+
+}
+
+\caption{\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (middle panel).
+The bottom panel shows the change in accessibility.}\label{fig:plot-accessibility}
+\end{figure}
+
 \hypertarget{conclusions}{%

### 2021-04-22 foodbanksHamilton/R/b2sfca.R

@@ -1,0 +1,61 @@
+#' Compute balanced floating catchment area accessibility
+#'
+#' @param ttm A travel time matrix with a column of origin IDs, a column of destination IDs, travel times, population at the origin, and supply at the destination.
+#' @param threshold A threshold for the binary impedance function
+#' @return The level of service \code{los} and accessibility \code{accessibility}.
+#' @export
+#' @examples
+#' #add(1, 1)
+#' #add(10, 1)
+
+b2sfca <- function(ttm = ttm, threshold = threshold){
+  # calculate the impedance
+  ttm <- ttm %>%
+    dplyr::mutate(impedance_binary = ifelse(travel_time <= threshold, 1, 0))
+
+  # Calculate the sum of the impedance by population for the balancing factors. In the binary case, this should be the same as the number of SoBi hubs that each interpolated population unit can reach
+  sum_b1 <- ttm %>%
+    dplyr::group_by(UID) %>%
+    dplyr::summarize(sum_b1 = sum(impedance_binary),
+                     .groups = "drop")
+
+  # Calculate the sum of the impedance by SoBi hubs for the balancing factors. In the binary case, this should be the same as the number of interpolated population cells that each SoBi hub serves
+  sum_b2 <- ttm %>%
+    dplyr::group_by(OBJECTID) %>%
+    dplyr::summarize(sum_b2 = sum(impedance_binary))
+
+  # Join the sum of the impedance to the table to calculate the balanced impedance values:
+  ttm <- ttm %>%
+    dplyr::left_join(sum_b1, by = "UID") %>%
+    dplyr::left_join(sum_b2, by = "OBJECTID")
+
+  # **Important:** notice that some values of `sum_b1` are zeros! This means that some population units cannot reach a SoBi hub in at most 24 min. Since the balancing impedance is calculated by diving the impedance by `sum_b1`, this will lead to NaNs (divisions by zero). To avoid issues we will remove any population units that reach zero hubs:
+  ttm <- ttm %>%
+    dplyr::filter(sum_b1 > 0,
+                  sum_b2 > 0)
+
+  # Calculate balanced impedance values:
+  ttm <- ttm %>%
+    dplyr::mutate(balanced_impedance_1 = impedance_binary/sum_b1,
+                  balanced_impedance_2 = impedance_binary/sum_b2)
+
+  # The level of service of each hub is the number of bicycle racks at the hub, divided by the population that they serve:
+  los <- ttm %>%
[…]

### 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Freshly knitted after merging changes

@@
 output: rticles::elsevier_article
+header-includes:
+    - \usepackage[table]{xcolor}
 ---
@@
 
-**The locations of foodbanks were obtained from public records and geocoded. Three urban sites not in the program were also identified and geocoded for comparison purposes.**
+The locations of foodbanks were obtained from public records and geocoded. Three urban sites not in the program were also identified and geocoded for comparison purposes.
 
@@
 The departure time used for routing was **TIME**.
-
-### Open Hamilton **???**
-
-We used the open data portal of the City of Hamilton\footnote{\url{https://open.hamilton.ca/}} we obtained boundaries for the city's various regions (the definition of urban, suburban, and rural regions follows the classification of development applications).
 
@@
 $$
+
 \noindent where $p^k_i$ is the proportion of trips by mode $k$ in the TAZ of parcel $i$, and $tt^k_{ij}$ is the travel time from parcel $i$ to the foodbank.

### 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Freshly knitted after merging changes

@@
 % Pandoc header
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
+\usepackage[table]{xcolor}
 
@@
 (e.g., Enns et al., 2020), we are not aware of any research that has
-focused on the geographic component of accessibility. \textbf{Check
-Tarasuk et al. (2019)}
+focused on the geographic component of accessibility.
 
@@
 
-\hypertarget{literature-review}{%
-\section{Literature Review}\label{literature-review}}
-
-\hypertarget{food-insecurity}{%
-\subsection{Food Insecurity}\label{food-insecurity}}
+\hypertarget{food-insecurity-and-food-bank-use-in-canada}{%
+\section{Food Insecurity and Food Bank Use in
+Canada}\label{food-insecurity-and-food-bank-use-in-canada}}
 
 Food insecurity is the inability to acquire and consume an adequate
-amount or good quality food, leading to inadequate nutrient intake
-(Kirkpatrick and Tarasuk, 2008; Tarasuk and Vogt, 2009). This nutrient
-deficiency has been causing major health concerns in Canadians, and
-particularly those who are at a socioeconomic disadvantage (Bazerghi et
-al., 2016). Previous studies have aimed to look at the relationship
-between the built food environment and socio-demographic characteristics
-with qualitative and quantitative, or a mixed -method approach.
-Quantitatively, official governmental surveys have been able to assist
-with data, such as the Household Food Security Survey Module (HFSSM),
-the Canadian Community Health Surveys (CCHS), the Longitudinal and
-International Study of Adults (LISA), and official classifications
-determined by Health Canada in relation to sociodemographic variables
-(Gundersen et al., 2018; Kirkpatrick and Tarasuk, 2008; Tarasuk and
-Vogt, 2009). Studies have also aimed to assess food availability of
-healthy foods (e.g., fruit and vegetables) at supermarkets in relation
-to sociodemographic characteristics and geographic accessibility (Latham
-and Moffat, 2007). In terms of findings, studies have generated
-inconsistent relationships between their evaluated availability of food
-and sociodemographic characteristics that go in hand.
-
-\hypertarget{foodbanks}{%
-\subsection{Foodbanks}\label{foodbanks}}
+amount or good quality food, leading to inadequate nutrient intake (Enns
+et al., 2020; Kirkpatrick and Tarasuk, 2008; Tarasuk and Vogt, 2009).
+This nutrient deficiency is associated with major health concerns in
+Canadians, and particularly those who are at a socioeconomic
+disadvantage (Bazerghi et al., 2016). Data on food insecurity in Canada
+is collected in several ways. Quantitatively, official government
+surveys such as the Household Food Security Survey Module (HFSSM), the
+Canadian Community Health Surveys (CCHS), the International Study of
+Adults (LISA), and official classifications determined by Health Canada
+in relation to sociodemographic variables offer some insight into food
+insecurity (Gundersen et al., 2018; Kirkpatrick and Tarasuk, 2008;
+Tarasuk and Vogt, 2009). Nationally, it has been found that food
+insecurity impacts approximately 12.3\% of Canadian households (Tarasuk
+et al., 2014)
 
@@
 evolved into a community practice to secure emergency food supplies for
-those in need (Loopstra and Tarasuk, 2012; Wakefield et al., 2013). In
-Canada, the number of foodbanks has steadily increased in the past few
-decades (Wakefield et al., 2013). Previous studies have questioned
-whether food banks are able to offer a rounded nutritious supply of
-food, and if food banks are a sustainable on-going practice for those in
-need (Bazerghi et al., 2016; Riches, 2002). An inescapable reality, at
-the moment, is that despite their imperfections they provide urgently
-needed support for members of the population for whom foodbanks are
-frequently their primary location to get food (Bazerghi et al., 2016).
-\textbf{However, surveys revealed that only 20 to 30 percent of those
-experiencing food insecurity were found to frequent food banks in Canada
-(Tarasuk et al., 2014).} \textbf{Something about geographical access}
+those in need (Loopstra and Tarasuk, 2012; Wakefield et al., 2013).
 
-\hypertarget{food-insecurity-in-canada-during-the-covid-19}{%
-\subsection{Food Insecurity in Canada during the
-COVID-19}\label{food-insecurity-in-canada-during-the-covid-19}}
+In Canada, the number of food banks has steadily increased in the past
+few decades (Wakefield et al., 2013). The largest database of food banks
+and their use comes from the non-profit association Food Banks Canada,
+which conducts an annual assessment through its affiliated members. Food
+Banks Canada's 2018 Hunger Count report\footnote{https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018\_FINAL\_EN.pdf.aspx?ext=.pdf}
+(the most recent available) listed 1,830 member food banks across the
+country found that Canadians visited food banks 1.1 million times in
+March of 2018. Of those accessing food banks, certain population
+characteristics tend to be over-represented compared to national totals
+from the 2016 Canadian Census of Population. According to Food Banks
+Canada's 2018 data, single-adult households represent 45\% of those
+utilizing food banks despite making up 28\% of Canada's population, 19\%
+are single-parent households (compared to 10\% nationally), and 35\% of
+those accessing food bank services are children aged 0-18 even though
+their share of Canada's national population is approximately 20\%. In
+addition, 59\% of households accessing food banks list social or
+disability assistance as their primary source of income. Moreover,
+children aged 0-18 are over-represented as food bank users, making up
+35\% of those accessing food bank services despite representing 20\% of
+the country's population. Likewise, using data from the 2011-2012 CCHS,
+Tarasuk et al. (2019) found higher odds of food insecurity amongst
+households relying on social assistance, those without a university
+degree or with children under the age of 18, and individuals that lived
+alone, renters, and those identifying as Aboriginal. While surveys
+revealed that only 20 to 30 percent of those experiencing food
+insecurity were found to frequent food banks in Canada (Tarasuk et al.,
+2014), research from Ottawa (Enns et al., 2020) and Vancouver (Black and
+Seto, 2020) suggests that long-term users tend to be older, have health
+or mobility challenges, live in large households, and less likely to
+have employment income.
 
@@
 food environment and led to higher rates of food insecurity (Niles et
-al., 2020). In 2012, 12.4\% of Canadians households and 11.8\% Ontarian
-households experienced some degree of food insecurity (Gundersen et al.,
-2018; Tarasuk et al., 2020). Pandemic-related food security studies in
-the US have found a substantial increase in households experiencing food
-insecurity for the first time, and also in households experiencing more
-severe food insecurity than before (Niles et al., 2020; Wolfson and
-Leung, 2020). Most recently in May 2020, Canada recorded 14.7\% of its
-population living in food insecurity in the past 30 days (Statistics
-Canada, 2020a). Food insecurity is a highly concerning public health
-issue due to its health impacts. Recent research suggests that food
-insecurity in adults can lead to experiencing more stressful events
-(El-hajj and Benhin 2021). Increases in food insecurity rates due to the
-pandemic, signal a change in the food environment with potential damages
-to population health during the course of the pandemic and beyond (Niles
-et al., 2020).
+al., 2020). Pandemic-related food security studies in the US have found
+a substantial increase in households experiencing food insecurity for
+the first time, and also in households experiencing more severe food
+insecurity than before (Niles et al., 2020; Wolfson and Leung, 2020).
+Most recently in May 2020, Canada recorded 14.7\% of its population
+living in food insecurity in the past 30 days (Statistics Canada,
+2020a). Considering the negative mental and physical health effects
+associated with food insecurity, increases in food insecurity rates due
+to the pandemic, signal a change in the food environment with potential
+damages to population health during the course of the pandemic and
+beyond (Niles et al., 2020). Recent data\footnote{https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report\_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis\_Exec-Sum\_EN.pdf}
+from Food Banks Canada showed that 52\% of member food banks reported an
+increase in usage in March of 2020 when initial lockdown restrictions
+were put in place across much of Canada. The pandemic also created
+significant staffiing issues with 42\% of food banks reporting a
+reduction in volunteers. However, 53\% of food banks reported a decrease
+in use into the summer of 2020 which food banks attributed to emergency
+financial support programs from the federal government were.
+Nevertheless, some of these benefit programs were temporary, suggesting
+many households may again turn to food bank services to meet their
+needs.
+
+In terms of geography, previous research conducted at the provincial
+scale using data from the 2011-2012 CCHS found that the prevalence of
+food insecurity ranged across the country from 11.8\% of households in
+Ontario to 41\% of households in Nunavut (Tarasuk et al., 2019).
+However, although previous research has examined the characteristics of
+individuals and households accessing food banks, the locational or
+transportation accessibility aspect of food bank access is not well
+understood. A wealth of literature examining the food desert concept
+suggests that, in addition to socioeconomic and demographic factors,
+location and transportation networks play a key role in a household's
+accessibility to healthy foods (Paez et al., 2010; Vanderlee and L'Abbé,
+2017; Widener, 2018). For food banks specifically, previous qualitative
+research in Ontario has noted difficulties associated with
+transportation for trips to the food bank (Smith-Carrier et al., 2017).
+For example, Smith et al. (2017) note that participants access food
+banks using a variety of modes of transportation with respondent stating
+``I wish it {[}the food bank{]} was a little more centrally located.
+Because if I didn't have a bike I'd have to walk it all the way out
+there and back. I wonder about people who don't.'' From this, the
+authors conclude that ``transportation can be challenging, particularly
+if the food bank is situated in a remote location.'' To offer greater
+insight into the role of transportation and location in food bank
+accessibility, this research examines how geographic access to food
+banks and food services changed in Hamilton before and during the
+COVID-19 pandemic.
 
@@
 
-\textbf{The locations of foodbanks were obtained from public records and
+The locations of foodbanks were obtained from public records and
 geocoded. Three urban sites not in the program were also identified and
-geocoded for comparison purposes.}
+geocoded for comparison purposes.
 
@@
 routing was \textbf{TIME}.
-
-\hypertarget{open-hamilton}{%
-\subsubsection{\texorpdfstring{Open Hamilton
-\textbf{???}}{Open Hamilton ???}}\label{open-hamilton}}
-
-We used the open data portal of the City of
-Hamilton\footnote{\url{https://open.hamilton.ca/}} we obtained
-boundaries for the city's various regions (the definition of urban,
-suburban, and rural regions follows the classification of development
-applications).
 
@@
 ett_i = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
-\] \noindent where \(p^k_i\) is the proportion of trips by mode \(k\) in
+\]
+
+\noindent where \(p^k_i\) is the proportion of trips by mode \(k\) in
 the TAZ of parcel \(i\), and \(tt^k_{ij}\) is the travel time from
@@
 
-\leavevmode\hypertarget{ref-latham2007determinants}{}%
-Latham, J., Moffat, T., 2007. Determinants of variation in food cost and
-availability in two socioeconomically contrasting neighbourhoods of
-hamilton, ontario, canada. Health \& Place 13, 273--287.
-doi:\url{https://doi.org/10.1016/j.healthplace.2006.01.006}
-
 \leavevmode\hypertarget{ref-loopstra2012relationship}{}%
@@
 
+\leavevmode\hypertarget{ref-smith2017food}{}%
+Smith-Carrier, T., Ross, K., Kirkham, J., Decker Pierce, B., 2017.
+{`Food is a right\ldots{} nobody should be starving on our streets'}:
+Perceptions of food bank usage in a mid-sized city in ontario, canada.
+Journal of Human Rights Practice 9, 29--49.
+
 \leavevmode\hypertarget{ref-statisticscanada2020food}{}%

### 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Changes to data description
Add food bank information and table

@@
 
-Data have been prepared for sharing as a data package^[add repository].
+Data have been prepared for sharing as a data package^[add repository]. This data package consists of the following:
 
@@
 
-Population and income statistics at the level of Dissemination Areas (DAs) were retrieved using the package `cancensus` [@vonBergmann2021cancensus].
-Dissemination Areas are the smallest publicly available census geography in Canada.
-We use data from the 2016 Population Census.
+Population and income statistics for 2016 were retrieved at the level of Dissemination Areas (DAs) using the package `cancensus` [@vonBergmann2021cancensus].
+DAs are the smallest publicly available census geography in Canada.
+Income data corresponds to the count of households by different total income groupings.
 
@@
 
-We converted all recorded residential parcels in the City of Hamilton to points on the road network.
+We converted all recorded residential land parcels in the City of Hamilton to points on the road network.
 Each point includes information about the number of residential units in the parcel.
-Each residential unit is "populated" with the probability of being a "low income household", which we define as having a total household income of less than CAD 40,000.
-This is approximately the mid-point of the low income cut-off (LICOs) for families in Canadian cities with populations greater than 500,000 in 2016, to match other Census data [@statisticscanada2020licos].
+Next, we define low-income households as those having a total income of less than CAD\$40,000, which is approximately the mid-point of the low income cut-off (LICOs) for families in Canadian cities with populations greater than 500,000 in 2016, to match other Census data [@statisticscanada2020licos].
+We then "populate" each residential unit with the probability of being a low-income household based on the counts of households by income groups in the DA in which the parcel is located.
+While this method assumes a constant probability of low-income household status for all residential units in a DA, the parcel-level analysis affords a high level of spatial disaggregation for the accessibility analysis.
 
-### Destinations: Foodbanks
+### Destinations: Food Banks and Food Service Locations
 
-The locations of foodbanks were obtained from public records and geocoded. Three urban sites not in the program were also identified and geocoded for comparison purposes.
+The locations of food banks and related food services were obtained from the Hamilton Public Library's Food Access Guide^[http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf].
+The guide was updated in April of 2021 to indicate any change affected on the services due to the pandemic.
+This includes modified business hours, a need to make reservations before frequenting, and locations that have completely shut down in consequence.
+While some food bank services have a specific a target population, such as prioritizing family with young children aged between 0 and 3 or accepting only those providing proof of low-income status through housing and utility statements, all the food bank services indicated below are designed to accommodate those in need of food at zero to low cost.
+With our focus on food banks and food services that offer free or low-cost meals at particular locations, we first removed services such as Meals on Wheels and other food access services such as food box, community kitchens, student nutrition programs, and shopping and transportation.
+In addition, two free meal services held on different days at the same location were collapsed into a single service point for the accessibility analysis.
+
+```{r table-food-bank-info, echo=FALSE, results='asis'}
+foodbank_table %>% kable("latex",
+        booktabs = TRUE,
+        align = ("cccccc"),
+        col.names = c("Type", 
+                      "Description", 
+                      "Locations Pre-COVID", 
+                      "Locations During COVID", 
+                      "Additional Notes"),
+        caption = "\\label{tab:food-bank-info}Food Bank and Food Service Information.",
+        escape = FALSE) %>%
+  kable_styling(latex_options = c("striped", "scale_down")) 
+```
 
@@
 Travel time tables for three modes (car, transit, walking) were computed using the parcels as the origins and the locations of the foodbanks as the destinations.
-For routing, the package `r5r` [@pereira2021r5r] was used, with a network extract for the City of Hamilton from OpenStreetMaps and the General Transit Feed Specification from Hamilton Street Railway, the local transit operator.
+For routing, the package `r5r` [@pereira2021r5r] was used with a network extract for the City of Hamilton from OpenStreetMaps and the General Transit Feed Specification from Hamilton Street Railway, the local transit operator.
 For routing purposes we used maximum values of 180 min and 10,000 m walking distance: any destination that exceeded these thresholds was ignored.
@@
 
-Once we obtained travel time tables with population (number of households) and proportion of trips by mode, we calculated the expected travel time $ett$ from each parcel $i$ to a foodbank $j$ as follows:
+Once we obtained travel time tables with population (number of households) and proportion of trips by mode, we calculated the expected travel time $ett$ from each parcel $i$ to a food bank or food service location $j$ as follows:
 $$
@@
 
-\noindent where $p^k_i$ is the proportion of trips by mode $k$ in the TAZ of parcel $i$, and $tt^k_{ij}$ is the travel time from parcel $i$ to the foodbank.
+\noindent where $p^k_i$ is the proportion of trips by mode $k$ in the TAZ of parcel $i$, and $tt^k_{ij}$ is the travel time from parcel $i$ to the food bank.
 In other words, the expected travel time is the weighted sum of travel times to the foodbank, with the weights given by the expected modal split in the TAZ.

### 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Started the results and discussion and worked on the figures

--- 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
@@
 knitr::opts_chunk$set(
-  cache = TRUE, out.width = "90%", fig.width = 6, fig.height = 6
+  cache = TRUE, out.width = "100%", fig.width = 6, fig.height = 6
 )
@@
 
-```{r load-packages, include=FALSE}
+```{r load-packages, cache=FALSE, include=FALSE}
 library(foodbanksHamilton) # Data package
@@
 
-\newpage
-
-Introduction
-==========================
-
-Food insecurity, defined as an "inadequate or uncertain access to a sufficient quantity and/or adequate quality of food" due to a household's financial limitations [@enns2020experiences], has been associated with reductions in nutritional outcomes [@bhattacharya2004poverty; @kirkpatrick2008food; @olson1999nutrition] and physical and mental health in children and adults [@elgar2021relative; @jones2017food; @ramsey2011food; @seligman2010food; @stuff2004household].
-Over at least the past four decades food banks have become an essential line of defense against food insecurity in Canadian communities [@riches2002food; @holmes2018nothing; @black2020examining; @tarasuk2020relationship].
-In this respect, Canada is not unlike numerous other wealthy countries where a systematic dismantling of the welfare state took place in the intervening period [@tarasuk2014food].
-
-But the emergence of COVID-19, the worst public health crisis since the 1918 flu pandemic, has revealed important social and economic fault lines, and pre-existing patterns of inequality appear to have been exacerbated.
-Along several other dimensions of stress [e.g., accessibility to health care facilities; @pereira2021geographic], this seems to be the case for food insecurity as well [@laborde2020poverty].
-In the US, for example, it has been estimated that there was an increase of more than 30\% in household food insecurity, and more than one third of households were discovered to be newly food insecure - meaning they did not experience food insecurity before the pandemic [@niles2020early].
-In Canada, Men and Tarasuk [-@men2021food] report that about 25\% of individuals who experienced job insecurity (a relatively common occurrence during the pandemic), also experienced food insecurity.
-Similarly, according to Statistics Canada [-@statisticscanada2020food], in the early stages of the pandemic almost 15\% of individuals reported living in a household that faced food insecurity; the risk of food insecurity was substantially higher for households with children.
-The difference between households with and without children was significant, and 11.7\% of households with children indicated that "food didn’t last and [there was] no money to get more" sometimes or often, compared to 7.3\% of households without children); likewise, 13\% of households with children indicated that they "[c]ouldn’t afford balanced meals" sometimes or often, compared to 8.8\% of households without children.
-
-The impacts of food insecurity during the pandemic are alarming, since diet-related diseases, such as obesity, heart-disease, and diabetes, were already critical public health concerns in Canada prior to COVID-19 [@boucher2017ontario].
-While food banks are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least they can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
-As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that food banks in Hamilton, Ontario, recorded the highest number of visitors in the past 29 years; the number of children visiting food banks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278 the year before.
-
-It is known that the urban food environment, within which people make their daily food choices, is essential in influencing eating behaviours and health outcomes, based on factors such as food availability, ease of geographic accessibility and socio-demographic variations [@paez2010relative; @vanderlee2017food; @widener2018spatial].
-However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to food banks.
-Although previous work has explored differences in _accessing_ food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource [e.g., @enns2020experiences], we are not aware of any research that has focused on the geographic component of accessibility.
-
-The study of place-based geographic accessibility is concerned with capturing the potential to reach destinations of value using the transportation network [@paez2012measuring].
-Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities^[https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with food bank service delivery in Vancouver [@holmes2018nothing].
-However, as with research into the prevalence of food deserts, accessibility to food banks is likely not evenly distributed throughout a city with variations in access attributable to transportation network characteristics and the spatial distribution of food bank locations - as well as the population that likely needs them.
-Furthermore, policy responses to the COVID-19 pandemic add to the distress of vulnerable households.
-Non-pharmaceutical interventions during the pandemic involving restrictions in mobility increased the friction of travel, in particular by transit on which low income populations are more likely to be reliant [e.g., @deweese2020tale].
-At the same time, the pandemic has created additional stress for the operators of foodbanks through disruptions in the supply chain [e.g., @mckay2021exploring] as well as concerns surrounding the delivery of service in safe conditions and possible cancellation of food service programs.
-
-For this study, we aim to look at how the landscape of food banks and related services (e.g.
-low-cost or free meal service providers) available in Hamilton, Ontario, has changed before and during the pandemic.
-Have the number of open food bank services diminished?
-If so, what was the accessibility to foodbanks before and during the pandemic, from the perspective of low income households?
-And finally, who are most likely to have been impacted by changes to the accessibility landscape?
-This paper will first look at the distribution of food banks and related services before and during the pandemic.
-Then, we use the balanced floating catchment area approach of Paez et al.
-[@paez2019demand] to investigate the accessibility situation.
-We use a fully disaggregated approach based on parcel-level data.
-Socio-economic and demographic data are drawn from the latest Census of Canada (2016), whereas travel information is from the most recent regional travel survey from 2016.
-This paper follows reproducible research recommendations [see @brunsdon2020opening], and the research was conducted using open source tools for transportation analysis [@lovelace2021open].
-The code and data necessary to reproduce the analysis are available in a public repository^[add repository].
-
-Food Insecurity and Food Bank Use in Canada
-============
-
-Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake [@enns2020experiences; @kirkpatrick2008food; @tarasuk2009household].
-This nutrient deficiency is associated with major health concerns in Canadians, and particularly those who are at a socioeconomic disadvantage [@bazerghi2016role].
-Data on food insecurity in Canada is collected in several ways.
-Quantitatively, official government surveys such as the Household Food Security Survey Module (HFSSM), the Canadian Community Health Surveys (CCHS), the International Study of Adults (LISA), and official classifications determined by Health Canada in relation to sociodemographic variables offer some insight into food insecurity [@kirkpatrick2008food; @tarasuk2009household; @gundersen2018food].
-Nationally, it has been found that food insecurity impacts approximately 12.3% of Canadian households [@tarasuk2014food]
-
-Food banks - sometimes also referred to as ‘food pantries’ and ‘food shelves’ - originated as a community response to aid those with inadequate food by voluntarily offering them meals and ingredients [@riches2002food; @loopstra2012relationship].
-The scope and objectives of foodbanks can vary by region and by country, and these organizations can include not only prepared meals and aliments, but also shared spaces to connect in community gardens and community kitchens [@wakefield2013sweet].
-Although in their origin foodbanks were meant to be provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time they evolved into a community practice to secure emergency food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
-
-In Canada, the number of food banks has steadily increased in the past few decades [@wakefield2013sweet].
-The largest database of food banks and their use comes from the non-profit association Food Banks Canada, which conducts an annual assessment through its affiliated members.
-Food Banks Canada's 2018 Hunger Count report^[https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018_FINAL_EN.pdf.aspx?ext=.pdf] (the most recent available) listed 1,830 member food banks across the country found that Canadians visited food banks 1.1 million times in March of 2018.
-Of those accessing food banks, certain population characteristics tend to be over-represented compared to national totals from the 2016 Canadian Census of Population.
-According to Food Banks Canada's 2018 data, single-adult households represent 45% of those utilizing food banks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those accessing food bank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
-In addition, 59% of households accessing food banks list social or disability assistance as their primary source of income.
-Moreover, children aged 0-18 are over-represented as food bank users, making up 35% of those accessing food bank services despite representing 20% of the country's population.
-Likewise, using data from the 2011-2012 CCHS, Tarasuk et al.
-[-@tarasuk2019geographic] found higher odds of food insecurity amongst households relying on social assistance, those without a university degree or with children under the age of 18, and individuals that lived alone, renters, and those identifying as Aboriginal.
-While surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada [@tarasuk2014food], research from Ottawa [@enns2020experiences] and Vancouver [@black2020examining] suggests that long-term users tend to be older, have health or mobility challenges, live in large households, and less likely to have employment income.
-
-Currently with the COVID-19 pandemic, disrupted economies, rising unemployment rates, and changes in poverty levels have disrupted the food environment and led to higher rates of food insecurity [@niles2020early].
-Pandemic-related food security studies in the US have found a substantial increase in households experiencing food insecurity for the first time, and also in households experiencing more severe food insecurity than before [@niles2020early; @wolfson2020food].
-Most recently in May 2020, Canada recorded 14.7% of its population living in food insecurity in the past 30 days [@statisticscanada2020food].
-Considering the negative mental and physical health effects associated with food insecurity, increases in food insecurity rates due to the pandemic, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
-Recent data^[https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis_Exec-Sum_EN.pdf] from Food Banks Canada showed that 52% of member food banks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of Canada.
-The pandemic also created significant staffiing issues with 42% of food banks reporting a reduction in volunteers.
-However, 53% of food banks reported a decrease in use into the summer of 2020 which food banks attributed to emergency financial support programs from the federal government were.
-Nevertheless, some of these benefit programs were temporary, suggesting many households may again turn to food bank services to meet their needs.
-
-In terms of geography, previous research conducted at the provincial scale using data from the 2011-2012 CCHS found that the prevalence of food insecurity ranged across the country from 11.8% of households in Ontario to 41% of households in Nunavut [@tarasuk2019geographic].
-However, although previous research has examined the characteristics of individuals and households accessing food banks, the locational or transportation accessibility aspect of food bank access is not well understood.
-A wealth of literature examining the food desert concept suggests that, in addition to socioeconomic and demographic factors, location and transportation networks play a key role in a household's accessibility to healthy foods [@paez2010relative; @vanderlee2017food; @widener2018spatial].
-For food banks specifically, previous qualitative research in Ontario has noted difficulties associated with transportation for trips to the food bank [@smith2017food].
-For example, Smith et al.
-[-@smith2017food] note that participants access food banks using a variety of modes of transportation with respondent stating "I wish it [the food bank] was a little more centrally located.
-Because if I didn’t have a bike I’d have to walk it all the way out there and back.
-I wonder about people who don’t." From this, the authors conclude that "transportation can be challenging, particularly if the food bank is situated in a remote location." To offer greater insight into the role of transportation and location in food bank accessibility, this research examines how geographic access to food banks and food services changed in Hamilton before and during the COVID-19 pandemic.
-
-Methods and Materials
-===================
-
-## Methods
-
-For the research in this paper we adopt the balanced floating catchment area approach of Paez et al.
-[-@paez2019demand].
-This method for estimating accessibility is a form of the common two-stage floating catchment area method [@radke2000spatial; @luo2003measures].
-Floating catchment areas are used to estimate accessibility when there are potential congestion effects, and operate by calculating first the _demand_ for spatially distributed services.
-The demand (usually the number of people who require a service) is used to calculate a level of service.
-In a second step, the level of service is allocated back to the population.
-Demand and level of service are allocated using some form of distance-decay to embody the geographical principle that, given a choice, people prefer to travel less than more when reaching destinations.
-
-More formally, the first step of this method is as follows:
-$$
-L_j = \frac{S_j}{\sum_{i=1}^nP_iw_{ij}}
-$$
-
-\noindent where $S_j$ is the level of supply at location $j$, in simplest terms whether a service point is present (i.e., $S_j=1$) or not (i.e., $S_j=0$); $P_i$ is the population at location $i$ that demands the service; and $w_{ij}$ is a weight, typically a function of the distance between locations $i$ and $j$.
-$L_j$ is the level of service at location $j$ and it is the inverse of the number of people that need to be serviced.
-
-The second step in this process is then summing the level of service that each population unit can reach, according to the distance-decay weight:
-$$
-A_i = \sum_{j=1}^JL_jw_{ji}
-$$
-
-\noindent where $A_i$ is the accessibility to the service, which is in the same units as the level of service: as the inverse of the population being serviced.
-When the population being serviced is low accessibility is high (i.e., there is little competition for the service), and viceversa.
-
-Floating catchment area methods are prone to overestimation of the population and the level of service due to multiple-counting.
-The population at $P_i$ is allocated to _every_ service point $j$ for which $w_{ij}>0$.
-Similarly, the level of service at $LOS_j$ is allocated to _every_ population point for which $w_{ji}>0$.
-This inflation effect has been known for several years, and several modifications have been proposed to mitigate it [e.g., @wan2012three; @delamater2013spatial].
-A definitive solution to this issue was presented by Paez et al.
-[-@paez2019demand].
-In order to avoid the multiple-counting in the summations, the population and the level of service need to be allocated _proportionally_.
-This is achieved by standardizing the weights as follows:
-$$
-w_{ij}^\text{st} = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
-$$
-\noindent and:
-$$
-w_{ji}^\text{st} = \frac{w_{ji}}{\sum_{j=1}^Jw_{ji}}
-$$
-
-The standardized weights satisfy the following conditions:
-$$
-\sum_{i=1}^nw_{ij}^\text{st}=1
-$$
-
-\noindent and:
-$$
-\sum_{j=1}^Jw_{ji}^\text{st}=1
-$$
-
-As a result, the total population (and consequently, the level of service) are preserved system-wide:
-$$
-\sum_{i=1}^nP_iw_{ij}^\text{st}=P_i
-$$
-
-\noindent and also:
-$$
-\sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
-$$
-
-## Data
-
-Data have been prepared for sharing as a data package^[add repository].
-
-### Statistics Canada
-
-Population and income statistics at the level of Dissemination Areas (DAs) were retrieved using the package `cancensus` [@vonBergmann2021cancensus].
-Dissemination Areas are the smallest publicly available census geography in Canada.
-We use data from the 2016 Population Census.
-
-### Origins: Residential parcels
-
-We converted all recorded residential parcels in the City of Hamilton to points on the road network.
-Each point includes information about the number of residential units in the parcel.
-Each residential unit is "populated" with the probability of being a "low income household", which we define as having a total household income of less than CAD 40,000.
-This is approximately the mid-point of the low income cut-off (LICOs) for families in Canadian cities with populations greater than 500,000 in 2016, to match other Census data [@statisticscanada2020licos].
-
-### Destinations: Foodbanks
-
-The locations of foodbanks were obtained from public records and geocoded. Three urban sites not in the program were also identified and geocoded for comparison purposes.
-
-### Routing and travel time tables
-
-Travel time tables for three modes (car, transit, walking) were computed using the parcels as the origins and the locations of the foodbanks as the destinations.
-For routing, the package `r5r` [@pereira2021r5r] was used, with a network extract for the City of Hamilton from OpenStreetMaps and the General Transit Feed Specification from Hamilton Street Railway, the local transit operator.
-For routing purposes we used maximum values of 180 min and 10,000 m walking distance: any destination that exceeded these thresholds was ignored.
-The departure time used for routing was **TIME**.
-
-### Transportation Tomorrow Survey
-
-We used the Data Retrieval System of the Transportation Tomorrow Survey (TTS)\footnote{\url{http://dmg.utoronto.ca/}} to download cross-tabulations of: 1) primary mode of travel per trip by income by place of residence; and 2) age by income by place of residence.
-These data are from the 2016 Survey (the most recent available), and data are geocoded at the level of Traffic Analysis Zones (TAZ) using the most recent zoning system from 2006.
-Each parcel point is populated with the proportion of trips by three modes of travel: car (as driver or passenger), transit, and walk.
-
-### Expected Travel Times
-
-Once we obtained travel time tables with population (number of households) and proportion of trips by mode, we calculated the expected travel time $ett$ from each parcel $i$ to a foodbank $j$ as follows:
-$$
-ett_i = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
-$$
-
-\noindent where $p^k_i$ is the proportion of trips by mode $k$ in the TAZ of parcel $i$, and $tt^k_{ij}$ is the travel time from parcel $i$ to the foodbank.
-In other words, the expected travel time is the weighted sum of travel times to the foodbank, with the weights given by the expected modal split in the TAZ.
-
-Results and Discussion
-===================
-
 ```{r remove-irrelevant-destinations, include=FALSE}
-## Remove entries in travel time matrices for inapplicable destinations
+## Data preparation
+# Remove entries in travel time matrices for inapplicable destinations
 # As a first step, we will remove the computed travel times to the three Community Kitchen locations (IDs 1-3):
@@
 ```{r join-low-income-households, include=FALSE}
-## Join proportion of low income households per DA to dwellings and total number of residential units:
+## Data preparation
+# Join proportion of low income households per DA to dwellings and total number of residential units:
 dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
@@
 ```{r probability-of-low-income-per-parcel, include=FALSE}
-## Calculate probability that a point on the network will have low income households.
+## Data preparation
+# Calculate probability that a point on the network will have low income households.
 This is the proportion of low income households in the DA multiplied by the total number of residential units at the point:
@@
 ## Data preparation
-# Join the food bank data to the travel time tables:
+# Join the foodbank data to the travel time tables:
 ttm_car <- ttm_car %>%
@@
 
-Words go here.
+\newpage
 
+Introduction
+==========================
+
+Food insecurity is defined as an "inadequate or uncertain access to a sufficient quantity and/or adequate quality of food" due to a household's financial limitations [@enns2020experiences].
+This is a condition that has been associated with reductions in nutritional outcomes [@bhattacharya2004poverty; @kirkpatrick2008food; @olson1999nutrition] and physical and mental health in children and adults [@elgar2021relative; @jones2017food; @ramsey2011food; @seligman2010food; @stuff2004household].
+Over at least the past four decades foodbanks have become an essential line of defense against food insecurity in Canadian communities [@riches2002food; @holmes2018nothing; @black2020examining; @tarasuk2020relationship].
+In this respect, Canada is not unlike numerous other wealthy countries where a systematic dismantling of the welfare state took place in the intervening period [@tarasuk2014food].
+
+But the emergence of COVID-19, the worst public health crisis since the 1918 flu pandemic, has revealed important social and economic fault lines, and pre-existing patterns of inequality appear to have been exacerbated.
+Along several other dimensions of stress [such as accessibility to health care facilities, see @pereira2021geographic], this seems to be the case for food insecurity as well [@laborde2020poverty].
+In the US, for example, it has been estimated that there was an increase of more than 30\% in household food insecurity, and more than one third of households were discovered to be newly food insecure - meaning they did not experience food insecurity before the pandemic [@niles2020early].
+In Canada, Men and Tarasuk [-@men2021food] report that about 25\% of individuals who experienced job insecurity (a relatively common occurrence during the pandemic), also experienced food insecurity.
+Similarly, according to Statistics Canada [-@statisticscanada2020food], in the early stages of the pandemic almost 15\% of individuals reported living in a household that faced food insecurity; the risk of food insecurity was substantially higher for households with children.
+The difference between households with and without children was significant, and 11.7\% of households with children indicated that "food didn’t last and [there was] no money to get more" sometimes or often, compared to 7.3\% of households without children; likewise, 13\% of households with children indicated that they "[c]ouldn’t afford balanced meals" sometimes or often, compared to 8.8\% of households without children.
+
+The impacts of food insecurity during the pandemic are alarming, since diet-related diseases, such as obesity, heart-disease, and diabetes, were already critical public health concerns in Canada prior to COVID-19 [@boucher2017ontario].
+While foodbanks are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least they can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
+As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that foodbanks in Hamilton, Ontario, recorded the highest number of visitors in the past 29 years; the number of children visiting foodbanks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278 the year before.
+
+It is known that the urban food environment, within which people make their daily food choices, is essential in influencing eating behaviours and health outcomes, based on factors such as food availability, ease of geographic accessibility and socio-demographic variations [@paez2010relative; @vanderlee2017food; @widener2018spatial].
+However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to foodbanks.
+Although previous work has explored differences in _accessing_ foodbanks, such as how some households utilize foodbanks over short periods of time while others regularly utilize foodbanks as longer-term resource [e.g., @enns2020experiences], we are not aware of any research that has focused on the geographic component of accessibility.
+
+The study of place-based geographic accessibility is concerned with capturing the potential to reach destinations of value using the transportation network [@paez2012measuring].
+Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities^[https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with foodbank service delivery in Vancouver [@holmes2018nothing].
+However, as with research into the prevalence of food deserts, accessibility to foodbanks is unlikely to be evenly distributed throughout a city due to variations in the characteristics of the transportation network characteristics, and the spatial distribution of foodbank locations and the population they serve.
+Furthermore, policy responses to the COVID-19 pandemic may well add to the distress of vulnerable households.
+Non-pharmaceutical interventions during the pandemic involving restrictions in mobility have increased the friction of travel, in particular by transit on which low income populations are more likely to be reliant [e.g., @deweese2020tale].
+At the same time, the pandemic has created additional stress for the operators of foodbanks through disruptions in the supply chain [e.g., @mckay2021exploring] as well as concerns surrounding the delivery of service in safe conditions and possible cancellation of food service programs.
+
+For this study, we aim to look at how the landscape of foodbanks and related services (e.g.
+low-cost or free meal service providers) available in Hamilton, Ontario, changed during the pandemic.
+Have the number of open foodbank services diminished?
+If so, what were the levels of accessibility to foodbanks before the pandemic from the perspective of low income households, and how did they change during the pandemic?
+And finally, who are most likely to have been impacted by changes to the accessibility landscape?
+This paper will first look at the distribution of foodbanks and related services before and during the pandemic.
+Then, we use the balanced floating catchment area approach of Paez et al.
+[-@paez2019demand] to investigate the accessibility situation.
+We use a fully disaggregated approach based on parcel-level data.
+Socio-economic and demographic data are drawn from the latest Census of Canada (2016), whereas travel information is from the most recent regional travel survey from 2016.
+This paper follows reproducible research recommendations [see @brunsdon2020opening], and the research was conducted using open source tools for transportation analysis [@lovelace2021open].
+The code and data necessary to reproduce the analysis are available in a public repository^[add repository].
+
+Food Insecurity and foodbank Use in Canada
+============
+
+Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to insufficient nutrient intake [@enns2020experiences; @kirkpatrick2008food; @tarasuk2009household].
+Nutrient deficiency, in turn, is associated with major population health concerns, and particularly among those who are at a socio-economic disadvantage [@bazerghi2016role].
+Data on food insecurity in Canada is collected in several ways.
+Quantitatively, official government surveys such as the Household Food Security Survey Module (HFSSM), the Canadian Community Health Surveys (CCHS), the International Study of Adults (LISA), and official classifications determined by Health Canada in relation to socio-demographic variables offer some insight into food insecurity [@kirkpatrick2008food; @tarasuk2009household; @gundersen2018food].
+Nationally, it has been found that food insecurity impacts approximately 12.3% of Canadian households [@tarasuk2014food]
+
+Foodbanks - sometimes also referred to as ‘food pantries’ and ‘food shelves’ - originated as a community response to aid those with inadequate food by voluntarily offering them meals and ingredients [@riches2002food; @loopstra2012relationship].
+The scope and objectives of foodbanks can vary by region and by country, and these organizations can include not only prepared meals and aliments, but also shared spaces to connect in community gardens and community kitchens [@wakefield2013sweet].
+Although in their origin foodbanks were meant to be provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time they evolved into a community practice to secure emergency food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
+
+In Canada, the number of foodbanks has steadily increased in the past few decades [@wakefield2013sweet].
+The largest database of foodbanks and their use comes from the non-profit association foodbanks Canada, which conducts an annual assessment through its affiliated members.
+foodbanks Canada's 2018 Hunger Count report^[https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018_FINAL_EN.pdf.aspx?ext=.pdf] (the most recent available) listed 1,830 member foodbanks across the country found that Canadians visited foodbanks 1.1 million times in March of 2018.
+Of those accessing foodbanks, certain population characteristics tend to be over-represented compared to national totals from the 2016 Canadian Census of Population.
+According to foodbanks Canada's 2018 data, single-adult households represent 45% of those utilizing foodbanks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those accessing foodbank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
+In addition, 59% of households accessing foodbanks list social or disability assistance as their primary source of income.
+Moreover, children aged 0-18 are over-represented as foodbank users, making up 35% of those accessing foodbank services despite representing 20% of the country's population.
+Likewise, using data from the 2011-2012 CCHS, Tarasuk et al.
+[-@tarasuk2019geographic] found higher odds of food insecurity amongst households relying on social assistance, those without a university degree or with children under the age of 18, and individuals that lived alone, renters, and those identifying as Aboriginal.
+While surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent foodbanks in Canada [@tarasuk2014food], research from Ottawa [@enns2020experiences] and Vancouver [@black2020examining] suggests that long-term users tend to be older, have health or mobility challenges, live in large households, and are less likely to have employment income.
+
+Currently with the COVID-19 pandemic, disrupted economies, rising unemployment rates, and changes in poverty levels have disrupted the food environment and led to higher rates of food insecurity [@niles2020early].
+Pandemic-related food security studies in the US have found a substantial increase in households experiencing food insecurity for the first time, and also in households experiencing more severe food insecurity than before [@niles2020early; @wolfson2020food].
+Most recently in May 2020, Canada recorded 14.7% of its population living in food insecurity in the past 30 days [@statisticscanada2020food].
+Considering the negative mental and physical health effects associated with food insecurity, increases in food insecurity rates due to the pandemic, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
+Recent data^[https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis_Exec-Sum_EN.pdf] from Foodbanks Canada showed that 52% of member foodbanks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of Canada.
+The pandemic also created significant staffing issues with 42% of foodbanks reporting a reduction in volunteers.
+However, 53% of foodbanks reported a decrease in use into the summer of 2020 which foodbanks attributed to emergency financial support programs from the federal government were.
+Nevertheless, some of these benefit programs were temporary, suggesting many households may again turn to foodbank services to meet their needs.
+
+In terms of geography, previous research conducted at the provincial scale using data from the 2011-2012 CCHS found that the prevalence of food insecurity ranged across the country from 11.8% of households in Ontario to 41% of households in Nunavut [@tarasuk2019geographic].
+However, although previous research has examined the characteristics of individuals and households accessing foodbanks, the locational or transportation accessibility aspect of foodbank access is not well understood.
+A wealth of literature examining the food desert concept suggests that, in addition to socio-economic and demographic factors, location and transportation networks play a key role in a household's accessibility to healthy foods [@paez2010relative; @vanderlee2017food; @widener2018spatial].
+For foodbanks specifically, previous qualitative research in Ontario has noted difficulties associated with transportation for trips to the foodbank [@smith2017food].
+For example, Smith et al.
+[-@smith2017food] found that participants access foodbanks using a variety of modes of transportation with a respondent stating "I wish it [the foodbank] was a little more centrally located.
+Because if I didn’t have a bike I’d have to walk it all the way out there and back.
+I wonder about people who don’t." From this, the authors conclude that "transportation can be challenging, particularly if the foodbank is situated in a remote location." To offer greater insight into the role of transportation and location in foodbank accessibility, this research examines how geographic access to foodbanks and food services changed in Hamilton before and during the COVID-19 pandemic.
+
+Methods and Materials
+===================
+
+## Methods
+
+For the research in this paper we adopt the balanced floating catchment area approach of Paez et al.
+[-@paez2019demand].
+This method for estimating accessibility is a form of the common two-stage floating catchment area method [@radke2000spatial; @luo2003measures].
+Floating catchment areas are used to estimate accessibility when there are potential congestion effects, and operate by calculating first the _demand_ for spatially distributed services.
+The demand (e.g., the number of people who require a service) is used to calculate a level of service.
+In a second step, the level of service is allocated back to the population.
+Demand and level of service are allocated using some form of distance-decay to embody the geographical principle that, given a choice, people prefer to travel less than more when reaching destinations.
+
+More formally, the first step of this method is as follows:
+$$
+L_j = \frac{S_j}{\sum_{i=1}^nP_iw_{ij}}
+$$
+
+\noindent where $S_j$ is the level of supply at location $j$, in simplest terms whether a service point is present (i.e., $S_j=1$) or not (i.e., $S_j=0$); $P_i$ is the population at location $i$ that demands the service; and $w_{ij}$ is a weight, typically a function of the distance between locations $i$ and $j$.
+$L_j$ is the level of service at location $j$ and it is the inverse of the number of people that need to be serviced.
+
+The second step in this process is then summing the level of service that each population unit can reach, according to the distance-decay weight:
+$$
+A_i = \sum_{j=1}^JL_jw_{ji}
+$$
+
+\noindent where $A_i$ is the accessibility to the service, which is in the same units as the level of service: as the inverse of the population being serviced.
+When the population being serviced is low accessibility is high (i.e., there is little competition for the service), and viceversa.
+
+Floating catchment area methods are prone to overestimation of the population and the level of service due to multiple-counting.
+The population at $P_i$ is allocated to _every_ service point $j$ for which $w_{ij}>0$.
+Similarly, the level of service at $L_j$ is allocated to _every_ population point for which $w_{ji}>0$.
+This inflation effect has been known for several years, and several modifications have been proposed to mitigate it [e.g., @wan2012three; @delamater2013spatial].
+A definitive solution to this issue was presented by Paez et al.
+[-@paez2019demand].
+In order to avoid the multiple-counting in the summations, the population and the level of service need to be allocated _proportionally_.
+This is achieved by standardizing the weights as follows:
+$$
+w_{ij}^\text{st} = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
+$$
+\noindent and:
+$$
+w_{ji}^\text{st} = \frac{w_{ji}}{\sum_{j=1}^Jw_{ji}}
+$$
+
+The standardized weights satisfy the following conditions:
+$$
+\sum_{i=1}^nw_{ij}^\text{st}=1
+$$
+
+\noindent and:
+$$
+\sum_{j=1}^Jw_{ji}^\text{st}=1
+$$
+
+As a result, the total population (and consequently, the level of service) are preserved system-wide:
+$$
+\sum_{i=1}^nP_iw_{ij}^\text{st}=P_i
+$$
+
+\noindent and also:
+$$
+\sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
+$$
+
+## Data
+
+Data have been prepared for sharing as a data package^[add repository].
+
+### Statistics Canada
+
+Population and income statistics at the level of Dissemination Areas (DAs) were retrieved using the package `cancensus` [@vonBergmann2021cancensus].
+Dissemination Areas are the smallest publicly available census geography in Canada.
+We use data from the 2016 Population Census.
+
+### Origins: Residential parcels
+
+We converted all recorded residential parcels in the City of Hamilton to points on the road network.
+Each point includes information about the number of residential units in the parcel.
+Each residential unit is "populated" with the probability of being a "low income household", which we define as having a total household income of less than CAD 40,000.
+This is approximately the mid-point of the low income cut-off (LICOs) for families in Canadian cities with populations greater than 500,000 in 2016, to match other Census data [@statisticscanada2020licos].
+
+### Destinations: Foodbanks
+
+The locations of foodbanks were obtained from public records and geocoded. Three urban sites not in the program were also identified and geocoded for comparison purposes.
+
+### Routing and travel time tables
+
+Travel time tables for three modes (car, transit, walking) were computed using the parcels as the origins and the locations of the foodbanks as the destinations.
+For routing, the package `r5r` [@pereira2021r5r] was used, with a network extract for the City of Hamilton from OpenStreetMaps and the General Transit Feed Specification from Hamilton Street Railway, the local transit operator.
+For routing purposes we used maximum values of 180 min and 10,000 m walking distance: any destination that exceeded these thresholds was ignored.
+The departure time used for routing was **TIME**.
+
+### Transportation Tomorrow Survey
+
+We used the Data Retrieval System of the Transportation Tomorrow Survey (TTS)\footnote{\url{http://dmg.utoronto.ca/}} to download cross-tabulations of: 1) primary mode of travel per trip by income by place of residence; and 2) age by income by place of residence.
+These data are from the 2016 Survey (the most recent available), and data are geocoded at the level of Traffic Analysis Zones (TAZ) using the most recent zoning system from 2006.
+Each parcel point is populated with the proportion of trips by three modes of travel: car (as driver or passenger), transit, and walk.
+
+### Expected Travel Times
+
+Once we obtained travel time tables with population (number of households) and proportion of trips by mode, we calculated the expected travel time $ett$ from each parcel $i$ to a foodbank $j$ as follows:
+$$
+ett_i = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
+$$
+
+\noindent where $p^k_i$ is the proportion of trips by mode $k$ in the TAZ of parcel $i$, and $tt^k_{ij}$ is the travel time from parcel $i$ to the foodbank.
+In other words, the expected travel time is the weighted sum of travel times to the foodbank, with the weights given by the expected modal split in the TAZ.
+
+Results and Discussion
+===================
+
+Figure \ref{fig:foodbanks} shows the location of the foodbanks in the City of Hamilton and their status.
+Before the pandemic there were `r nrow(foodbank_locations)` of which `r nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))` closed during the pandemic.
+As seen in the figure, foodbanks tend to be predominantly located in the central parts of the city.
+This is not surprising: population density is high there, and it is also the part of the city where lower income households are more numerous in absolute and relative terms (see Figure \ref{fig:low-income-households}).
+Alas, this is also where most of the closures during the pandemic happened.
+
+```{r plot-location-foodbanks, echo=FALSE, fig.cap="\\label{fig:foodbanks}Location of foodbanks and operation status; the dotted box is an inset of the central part of the City of Hamilton"}
+## Analysis
+# Plot location of foodbanks
+
+# Create a bounding box for an inset
+inset_box <- st_bbox(c(xmin = -79.975,
+          xmax = -79.775, 
+          ymax = 43.2, 
+          ymin = 43.3), 
+        crs = st_crs(4326)) %>%
+  st_as_sfc()
+
+fb <- ggplot() +
+  geom_sf(data = hamilton_cma,
+          fill = "lightgray") +
+  geom_sf(data = foodbank_locations,
+          aes(size = COVIDClose,
+              shape = COVIDClose)) +
+  geom_sf(data = inset_box,
+          linetype = "dotted",
+          color = "red",
+          fill = NA) +
+  scale_shape_manual(name = "Closed during COVID-19",
+                     values = c("Yes" = 4, "No" = 1)) +
+  scale_size_manual(values = c("Yes" = 3, "No" = 2)) +
+  guides(size = FALSE) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "none")
+
+fb_inset <- ggplot() +
+  geom_sf(data = hamilton_cma,
+          fill = "lightgray") +
+  geom_sf(data = foodbank_locations,
+          aes(size = COVIDClose,
+              shape = COVIDClose)) +
+  geom_sf(data = inset_box,
+          linetype = "dotted",
+          color = "red",
+          fill = NA) +
+  scale_shape_manual(name = "Closed during COVID-19",
+                     values = c("Yes" = 4, "No" = 1)) +
+  scale_size_manual(values = c("Yes" = 3, "No" = 2)) +
+  xlim(-79.975, -79.775) +
+  ylim(43.2, 43.3) +
+  guides(size = FALSE) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom")
+
+fb / fb_inset
+```
+```{r plot-low-income-households, echo=FALSE, fig.cap="\\label{fig:low-income-households}Number and proportion of households with incomes less than CAD 40,000."}
+lih <- ggplot() +
+  geom_sf(data = data_da_2016,
+          aes(fill = income_less40k),
+          color = NA) +
+  geom_sf(data = inset_box,
+          linetype = "dotted",
+          color = "red",
+          fill = NA) +
+  scale_fill_viridis_c(name = "Number of households", 
+                       option = "A",
+                       direction = -1) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "bottom")
+
+p_lih <- ggplot() +
+  geom_sf(data = data_da_2016,
+          aes(fill = p_inc_less40k),
+          color = NA) +
+  geom_sf(data = inset_box,
+          linetype = "dotted",
+          color = "red",
+          fill = NA) +
+  scale_fill_viridis_c(name = "Proportion of households", 
+                       option = "A",
+                       direction = -1) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "bottom")
+
+lih / p_lih
+```
 ```{r accessibility-threshold-sensitivity, include=FALSE}
@@
 
-```{r plot-levels-of-service, echo=FALSE, fig.cap="\\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (middle panel).
-The bottom panel shows the change in level of service."}
+```{r plot-levels-of-service, echo=FALSE, fig.cap="\\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
 ## Analysis
@@
           aes(shape = COVIDClose)) +
-  scale_color_distiller(palette = "OrRd", 
-                        direction = 1) +
+  scale_color_viridis_c(name = "Level of service pre-COVID-19",
+                        option =  "A", 
+                        direction = -1) +
   scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  theme_minimal()
+  guides(size = FALSE, shape = FALSE) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
 
@@
           aes(shape = COVIDClose)) +
-  scale_color_distiller(palette = "OrRd", 
-                        direction = 1) +
+  scale_color_viridis_c(name = "Level of service during COVID-19",
+                        option =  "A", 
+                        direction = -1) +
   scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  theme_minimal()
+  guides(size = FALSE, shape = FALSE) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
 
+los_pre / los_covid
+```
+
+```{r plot-levels-of-service-changes, echo=FALSE, fig.cap="\\label{fig:levels-of-service-changes}Changes in levels of service at each facility from pre-COVID-19 to during COVID-19."}
 # This figure shows the places where the level of service deteriorated most (excluding the places that closed, where it vanished).
@@
   filter(`During covid` > 0) %>%
-  mutate(los_d = `During covid` / `Pre-covid` - 1) %>%
+  mutate(los_d = `During covid` - `Pre-covid`) %>%
   left_join(foodbank_locations %>%
@@
           aes(shape = COVIDClose)) +
-  scale_color_distiller(palette = "OrRd", 
+  scale_color_viridis_c(name = "Change in level of service",
+                        option =  "A", 
                         direction = -1) +
+  scale_size(trans = 'reverse') +
   scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  theme_minimal()
+  guides(size = FALSE, shape = FALSE) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
 
-los_pre / los_covid / los_change
+los_change
 ```
@@
 
-```{r plot-accessibility, echo=FALSE, fig.align = 'center', out.width = "90%", fig.cap="\\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (middle panel).
-The bottom panel shows the change in accessibility."}
+```{r plot-accessibility, echo=FALSE, fig.align = 'center', out.width = "90%", fig.cap="\\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
 ## Analysis
@@
 acc_pre <- ggplot() +
-  geom_sf(data = hamilton_cma) +
+  #geom_sf(data = hamilton_cma) +
   geom_sf(data = accessibility_taz %>%
@@
           aes(fill = accessibility)) +
-  scale_fill_distiller(palette = "OrRd", 
-                        direction = 1) +
-  theme_minimal()
+  scale_fill_viridis_c(name = "Accessibility pre-COVID-19",
+                        option =  "A", 
+                        direction = -1,
+                       breaks = c(0.001,
+                                  0.0008,
+                                  0.0006,
+                                  0.0004,
+                                  0.0002),
+                       labels = c("0.0010",
+                                  "0.0008",
+                                  "0.0006",
+                                  "0.0004",
+                                  "0.0002")) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
 
 acc_covid <- ggplot() +
-  geom_sf(data = hamilton_cma) +
+  #geom_sf(data = hamilton_cma) +
   geom_sf(data = accessibility_taz %>%
@@
           aes(fill = accessibility)) +
-  scale_fill_distiller(palette = "OrRd", 
-                        direction = 1) +
-  theme_minimal()
+  scale_fill_viridis_c(name = "Accessibility during COVID-19",
+                        option =  "A", 
+                        direction = -1,
+                       breaks = c(0.0005,
+                                  0.0004,
+                                  0.0003,
+                                  0.0002,
+                                  0.0001),
+                       labels = c("0.0005",
+                                  "0.0004",
+                                  "0.0003",
+                                  "0.0002",
+                                  "0.0001")) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
 
+acc_pre / acc_covid
+```
+
+```{r plot-accessibility-changes, echo=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19."}
+## Analysis
 # Calculate accessibility change by TAZ:
@@
 acc_change <- ggplot() +
-  geom_sf(data = hamilton_cma) +
+  #geom_sf(data = hamilton_cma) +
   geom_sf(data = accessibilty_taz_change,
@@
           fill = NA) +
-  scale_fill_fermenter(palette = "OrRd", 
-                       direction = -1) +
-  theme_minimal()
+  scale_fill_viridis_c(name = "Change in accessibility",
+                        option =  "A", 
+                        direction = 1,
+                       breaks = c(-0.0005,
+                                  -0.0004,
+                                  -0.0003,
+                                  -0.0002,
+                                  -0.0001),
+                       labels = c("-0.0005",
+                                  "-0.0004",
+                                  "-0.0003",
+                                  "-0.0002",
+                                  "-0.0001")) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
 
-acc_pre / acc_covid / acc_change
+acc_change
 ```

### 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Started the results and discussion and worked on the figures

@@
 \usepackage[table]{xcolor}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 
-Food insecurity, defined as an ``inadequate or uncertain access to a
+Food insecurity is defined as an ``inadequate or uncertain access to a
 sufficient quantity and/or adequate quality of food'' due to a
-household's financial limitations (Enns et al., 2020), has been
-associated with reductions in nutritional outcomes (Bhattacharya et al.,
-2004; Kirkpatrick and Tarasuk, 2008; Olson, 1999) and physical and
-mental health in children and adults (Elgar et al., 2021; Jones, 2017;
-Ramsey et al., 2011; Seligman et al., 2010; Stuff et al., 2004). Over at
-least the past four decades food banks have become an essential line of
-defense against food insecurity in Canadian communities (Black and Seto,
-2020; Holmes et al., 2018; Riches, 2002; Tarasuk et al., 2020). In this
-respect, Canada is not unlike numerous other wealthy countries where a
-systematic dismantling of the welfare state took place in the
-intervening period (Tarasuk et al., 2014).
+household's financial limitations (Enns et al., 2020). This is a
+condition that has been associated with reductions in nutritional
+outcomes (Bhattacharya et al., 2004; Kirkpatrick and Tarasuk, 2008;
+Olson, 1999) and physical and mental health in children and adults
+(Elgar et al., 2021; Jones, 2017; Ramsey et al., 2011; Seligman et al.,
+2010; Stuff et al., 2004). Over at least the past four decades foodbanks
+have become an essential line of defense against food insecurity in
+Canadian communities (Black and Seto, 2020; Holmes et al., 2018; Riches,
+2002; Tarasuk et al., 2020). In this respect, Canada is not unlike
+numerous other wealthy countries where a systematic dismantling of the
+welfare state took place in the intervening period (Tarasuk et al.,
+2014).
 
@@
 lines, and pre-existing patterns of inequality appear to have been
-exacerbated. Along several other dimensions of stress {[}e.g.,
-accessibility to health care facilities; Pereira et al. (2021a){]}, this
-seems to be the case for food insecurity as well (Laborde et al., 2020).
-In the US, for example, it has been estimated that there was an increase
-of more than 30\% in household food insecurity, and more than one third
-of households were discovered to be newly food insecure - meaning they
-did not experience food insecurity before the pandemic (Niles et al.,
-2020). In Canada, Men and Tarasuk (2021) report that about 25\% of
-individuals who experienced job insecurity (a relatively common
-occurrence during the pandemic), also experienced food insecurity.
-Similarly, according to Statistics Canada (2020a), in the early stages
-of the pandemic almost 15\% of individuals reported living in a
-household that faced food insecurity; the risk of food insecurity was
-substantially higher for households with children. The difference
+exacerbated. Along several other dimensions of stress (such as
+accessibility to health care facilities, see Pereira et al., 2021a),
+this seems to be the case for food insecurity as well (Laborde et al.,
+2020). In the US, for example, it has been estimated that there was an
+increase of more than 30\% in household food insecurity, and more than
+one third of households were discovered to be newly food insecure -
+meaning they did not experience food insecurity before the pandemic
+(Niles et al., 2020). In Canada, Men and Tarasuk (2021) report that
+about 25\% of individuals who experienced job insecurity (a relatively
+common occurrence during the pandemic), also experienced food
+insecurity. Similarly, according to Statistics Canada (2020a), in the
+early stages of the pandemic almost 15\% of individuals reported living
+in a household that faced food insecurity; the risk of food insecurity
+was substantially higher for households with children. The difference
 between households with and without children was significant, and 11.7\%
@@
 {[}there was{]} no money to get more'' sometimes or often, compared to
-7.3\% of households without children); likewise, 13\% of households with
+7.3\% of households without children; likewise, 13\% of households with
 children indicated that they ``{[}c{]}ouldn't afford balanced meals''
@@
 were already critical public health concerns in Canada prior to COVID-19
-(Boucher et al., 2017). While food banks are not necessarily a stable
+(Boucher et al., 2017). While foodbanks are not necessarily a stable
 solution to food insecurity and in fact may encourage a retrenchment of
@@
 Hunger Report\footnote{https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf}
-noted that food banks in Hamilton, Ontario, recorded the highest number
-of visitors in the past 29 years; the number of children visiting food
-banks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278
-the year before.
+noted that foodbanks in Hamilton, Ontario, recorded the highest number
+of visitors in the past 29 years; the number of children visiting
+foodbanks (minors up to 18 years old) was 9,125 in March 2019, up from
+8,278 the year before.
 
@@
 accessibility to healthy food through the ``food desert'' concept, there
-has been little research into accessibility to food banks. Although
-previous work has explored differences in \emph{accessing} food banks,
-such as how some households utilize food banks over short periods of
-time while others regularly utilize food banks as longer-term resource
-(e.g., Enns et al., 2020), we are not aware of any research that has
-focused on the geographic component of accessibility.
+has been little research into accessibility to foodbanks. Although
+previous work has explored differences in \emph{accessing} foodbanks,
+such as how some households utilize foodbanks over short periods of time
+while others regularly utilize foodbanks as longer-term resource (e.g.,
+Enns et al., 2020), we are not aware of any research that has focused on
+the geographic component of accessibility.
 
@@
 and previous survey research has suggested that such accessibility plays
-a key role in user satisfaction with food bank service delivery in
+a key role in user satisfaction with foodbank service delivery in
 Vancouver (Holmes et al., 2018). However, as with research into the
-prevalence of food deserts, accessibility to food banks is likely not
-evenly distributed throughout a city with variations in access
-attributable to transportation network characteristics and the spatial
-distribution of food bank locations - as well as the population that
-likely needs them. Furthermore, policy responses to the COVID-19
-pandemic add to the distress of vulnerable households.
-Non-pharmaceutical interventions during the pandemic involving
-restrictions in mobility increased the friction of travel, in particular
-by transit on which low income populations are more likely to be reliant
-(e.g., DeWeese et al., 2020). At the same time, the pandemic has created
-additional stress for the operators of foodbanks through disruptions in
-the supply chain (e.g., McKay et al., 2021) as well as concerns
-surrounding the delivery of service in safe conditions and possible
-cancellation of food service programs.
+prevalence of food deserts, accessibility to foodbanks is unlikely to be
+evenly distributed throughout a city due to variations in the
+characteristics of the transportation network characteristics, and the
+spatial distribution of foodbank locations and the population they
+serve. Furthermore, policy responses to the COVID-19 pandemic may well
+add to the distress of vulnerable households. Non-pharmaceutical
+interventions during the pandemic involving restrictions in mobility
+have increased the friction of travel, in particular by transit on which
+low income populations are more likely to be reliant (e.g., DeWeese et
+al., 2020). At the same time, the pandemic has created additional stress
+for the operators of foodbanks through disruptions in the supply chain
+(e.g., McKay et al., 2021) as well as concerns surrounding the delivery
+of service in safe conditions and possible cancellation of food service
+programs.
 
-For this study, we aim to look at how the landscape of food banks and
+For this study, we aim to look at how the landscape of foodbanks and
 related services (e.g.~low-cost or free meal service providers)
-available in Hamilton, Ontario, has changed before and during the
-pandemic. Have the number of open food bank services diminished? If so,
-what was the accessibility to foodbanks before and during the pandemic,
-from the perspective of low income households? And finally, who are most
-likely to have been impacted by changes to the accessibility landscape?
-This paper will first look at the distribution of food banks and related
-services before and during the pandemic. Then, we use the balanced
-floating catchment area approach of Paez et al. (Paez et al., 2019) to
-investigate the accessibility situation. We use a fully disaggregated
-approach based on parcel-level data. Socio-economic and demographic data
-are drawn from the latest Census of Canada (2016), whereas travel
-information is from the most recent regional travel survey from 2016.
-This paper follows reproducible research recommendations (see Brunsdon
-and Comber, 2020), and the research was conducted using open source
-tools for transportation analysis (Lovelace, 2021). The code and data
-necessary to reproduce the analysis are available in a public
-repository\footnote{add repository}.
+available in Hamilton, Ontario, changed during the pandemic. Have the
+number of open foodbank services diminished? If so, what were the levels
+of accessibility to foodbanks before the pandemic from the perspective
+of low income households, and how did they change during the pandemic?
+And finally, who are most likely to have been impacted by changes to the
+accessibility landscape? This paper will first look at the distribution
+of foodbanks and related services before and during the pandemic. Then,
+we use the balanced floating catchment area approach of Paez et al.
+(2019) to investigate the accessibility situation. We use a fully
+disaggregated approach based on parcel-level data. Socio-economic and
+demographic data are drawn from the latest Census of Canada (2016),
+whereas travel information is from the most recent regional travel
+survey from 2016. This paper follows reproducible research
+recommendations (see Brunsdon and Comber, 2020), and the research was
+conducted using open source tools for transportation analysis (Lovelace,
+2021). The code and data necessary to reproduce the analysis are
+available in a public repository\footnote{add repository}.
 
-\hypertarget{food-insecurity-and-food-bank-use-in-canada}{%
-\section{Food Insecurity and Food Bank Use in
-Canada}\label{food-insecurity-and-food-bank-use-in-canada}}
+\hypertarget{food-insecurity-and-foodbank-use-in-canada}{%
+\section{Food Insecurity and foodbank Use in
+Canada}\label{food-insecurity-and-foodbank-use-in-canada}}
 
 Food insecurity is the inability to acquire and consume an adequate
-amount or good quality food, leading to inadequate nutrient intake (Enns
-et al., 2020; Kirkpatrick and Tarasuk, 2008; Tarasuk and Vogt, 2009).
-This nutrient deficiency is associated with major health concerns in
-Canadians, and particularly those who are at a socioeconomic
-disadvantage (Bazerghi et al., 2016). Data on food insecurity in Canada
-is collected in several ways. Quantitatively, official government
-surveys such as the Household Food Security Survey Module (HFSSM), the
-Canadian Community Health Surveys (CCHS), the International Study of
-Adults (LISA), and official classifications determined by Health Canada
-in relation to sociodemographic variables offer some insight into food
-insecurity (Gundersen et al., 2018; Kirkpatrick and Tarasuk, 2008;
-Tarasuk and Vogt, 2009). Nationally, it has been found that food
-insecurity impacts approximately 12.3\% of Canadian households (Tarasuk
-et al., 2014)
+amount or good quality food, leading to insufficient nutrient intake
+(Enns et al., 2020; Kirkpatrick and Tarasuk, 2008; Tarasuk and Vogt,
+2009). Nutrient deficiency, in turn, is associated with major population
+health concerns, and particularly among those who are at a
+socio-economic disadvantage (Bazerghi et al., 2016). Data on food
+insecurity in Canada is collected in several ways. Quantitatively,
+official government surveys such as the Household Food Security Survey
+Module (HFSSM), the Canadian Community Health Surveys (CCHS), the
+International Study of Adults (LISA), and official classifications
+determined by Health Canada in relation to socio-demographic variables
+offer some insight into food insecurity (Gundersen et al., 2018;
+Kirkpatrick and Tarasuk, 2008; Tarasuk and Vogt, 2009). Nationally, it
+has been found that food insecurity impacts approximately 12.3\% of
+Canadian households (Tarasuk et al., 2014)
 
-Food banks - sometimes also referred to as `food pantries' and `food
+Foodbanks - sometimes also referred to as `food pantries' and `food
 shelves' - originated as a community response to aid those with
@@
 
-In Canada, the number of food banks has steadily increased in the past
-few decades (Wakefield et al., 2013). The largest database of food banks
-and their use comes from the non-profit association Food Banks Canada,
-which conducts an annual assessment through its affiliated members. Food
-Banks Canada's 2018 Hunger Count report\footnote{https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018\_FINAL\_EN.pdf.aspx?ext=.pdf}
-(the most recent available) listed 1,830 member food banks across the
-country found that Canadians visited food banks 1.1 million times in
-March of 2018. Of those accessing food banks, certain population
+In Canada, the number of foodbanks has steadily increased in the past
+few decades (Wakefield et al., 2013). The largest database of foodbanks
+and their use comes from the non-profit association foodbanks Canada,
+which conducts an annual assessment through its affiliated members.
+foodbanks Canada's 2018 Hunger Count report\footnote{https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018\_FINAL\_EN.pdf.aspx?ext=.pdf}
+(the most recent available) listed 1,830 member foodbanks across the
+country found that Canadians visited foodbanks 1.1 million times in
+March of 2018. Of those accessing foodbanks, certain population
 characteristics tend to be over-represented compared to national totals
-from the 2016 Canadian Census of Population. According to Food Banks
+from the 2016 Canadian Census of Population. According to foodbanks
 Canada's 2018 data, single-adult households represent 45\% of those
-utilizing food banks despite making up 28\% of Canada's population, 19\%
+utilizing foodbanks despite making up 28\% of Canada's population, 19\%
 are single-parent households (compared to 10\% nationally), and 35\% of
-those accessing food bank services are children aged 0-18 even though
+those accessing foodbank services are children aged 0-18 even though
 their share of Canada's national population is approximately 20\%. In
-addition, 59\% of households accessing food banks list social or
+addition, 59\% of households accessing foodbanks list social or
 disability assistance as their primary source of income. Moreover,
-children aged 0-18 are over-represented as food bank users, making up
-35\% of those accessing food bank services despite representing 20\% of
+children aged 0-18 are over-represented as foodbank users, making up
+35\% of those accessing foodbank services despite representing 20\% of
 the country's population. Likewise, using data from the 2011-2012 CCHS,
@@
 revealed that only 20 to 30 percent of those experiencing food
-insecurity were found to frequent food banks in Canada (Tarasuk et al.,
+insecurity were found to frequent foodbanks in Canada (Tarasuk et al.,
 2014), research from Ottawa (Enns et al., 2020) and Vancouver (Black and
 Seto, 2020) suggests that long-term users tend to be older, have health
-or mobility challenges, live in large households, and less likely to
+or mobility challenges, live in large households, and are less likely to
 have employment income.
@@
 beyond (Niles et al., 2020). Recent data\footnote{https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report\_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis\_Exec-Sum\_EN.pdf}
-from Food Banks Canada showed that 52\% of member food banks reported an
+from Foodbanks Canada showed that 52\% of member foodbanks reported an
 increase in usage in March of 2020 when initial lockdown restrictions
 were put in place across much of Canada. The pandemic also created
-significant staffiing issues with 42\% of food banks reporting a
-reduction in volunteers. However, 53\% of food banks reported a decrease
-in use into the summer of 2020 which food banks attributed to emergency
+significant staffing issues with 42\% of foodbanks reporting a reduction
+in volunteers. However, 53\% of foodbanks reported a decrease in use
+into the summer of 2020 which foodbanks attributed to emergency
 financial support programs from the federal government were.
 Nevertheless, some of these benefit programs were temporary, suggesting
-many households may again turn to food bank services to meet their
-needs.
+many households may again turn to foodbank services to meet their needs.
 
@@
 However, although previous research has examined the characteristics of
-individuals and households accessing food banks, the locational or
-transportation accessibility aspect of food bank access is not well
+individuals and households accessing foodbanks, the locational or
+transportation accessibility aspect of foodbank access is not well
 understood. A wealth of literature examining the food desert concept
-suggests that, in addition to socioeconomic and demographic factors,
+suggests that, in addition to socio-economic and demographic factors,
 location and transportation networks play a key role in a household's
 accessibility to healthy foods (Paez et al., 2010; Vanderlee and L'Abbé,
-2017; Widener, 2018). For food banks specifically, previous qualitative
+2017; Widener, 2018). For foodbanks specifically, previous qualitative
 research in Ontario has noted difficulties associated with
-transportation for trips to the food bank (Smith-Carrier et al., 2017).
-For example, Smith et al. (2017) note that participants access food
-banks using a variety of modes of transportation with respondent stating
-``I wish it {[}the food bank{]} was a little more centrally located.
-Because if I didn't have a bike I'd have to walk it all the way out
-there and back. I wonder about people who don't.'' From this, the
+transportation for trips to the foodbank (Smith-Carrier et al., 2017).
+For example, Smith et al. (2017) found that participants access
+foodbanks using a variety of modes of transportation with a respondent
+stating ``I wish it {[}the foodbank{]} was a little more centrally
+located. Because if I didn't have a bike I'd have to walk it all the way
+out there and back. I wonder about people who don't.'' From this, the
 authors conclude that ``transportation can be challenging, particularly
-if the food bank is situated in a remote location.'' To offer greater
-insight into the role of transportation and location in food bank
-accessibility, this research examines how geographic access to food
-banks and food services changed in Hamilton before and during the
-COVID-19 pandemic.
+if the foodbank is situated in a remote location.'' To offer greater
+insight into the role of transportation and location in foodbank
+accessibility, this research examines how geographic access to foodbanks
+and food services changed in Hamilton before and during the COVID-19
+pandemic.
 
@@
 congestion effects, and operate by calculating first the \emph{demand}
-for spatially distributed services. The demand (usually the number of
+for spatially distributed services. The demand (e.g., the number of
 people who require a service) is used to calculate a level of service.
@@
 population at \(P_i\) is allocated to \emph{every} service point \(j\)
-for which \(w_{ij}>0\). Similarly, the level of service at \(LOS_j\) is
+for which \(w_{ij}>0\). Similarly, the level of service at \(L_j\) is
 allocated to \emph{every} population point for which \(w_{ji}>0\). This
@@
 
-Words go here.
+Figure \ref{fig:foodbanks} shows the location of the foodbanks in the
+City of Hamilton and their status. Before the pandemic there were 58 of
+which 14 closed during the pandemic. As seen in the figure, foodbanks
+tend to be predominantly located in the central parts of the city. This
+is not surprising: population density is high there, and it is also the
+part of the city where lower income households are more numerous in
+absolute and relative terms (see Figure
+\ref{fig:low-income-households}). Alas, this is also where most of the
+closures during the pandemic happened.
 
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-results-sensitivity-analysis-1} \caption{\label{fig:sensitivity-analysis}Accessibility as a function of threshold}\label{fig:plot-results-sensitivity-analysis}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-location-foodbanks-1} \caption{\label{fig:foodbanks}Location of foodbanks and operation status; the dotted box is an inset of the central part of the City of Hamilton}\label{fig:plot-location-foodbanks}
+\end{figure}
+\begin{figure}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-low-income-households-1} \caption{\label{fig:low-income-households}Number and proportion of households with incomes less than CAD 40,000.}\label{fig:plot-low-income-households}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-results-sensitivity-analysis-1} \caption{\label{fig:sensitivity-analysis}Accessibility as a function of threshold}\label{fig:plot-results-sensitivity-analysis}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-levels-of-service-1} \caption{\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (middle panel).
-The bottom panel shows the change in level of service.}\label{fig:plot-levels-of-service}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-levels-of-service-1} \caption{\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (bottom panel).}\label{fig:plot-levels-of-service}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-levels-of-service-changes-1} \caption{\label{fig:levels-of-service-changes}Changes in levels of service at each facility from pre-COVID-19 to during COVID-19.}\label{fig:plot-levels-of-service-changes}
 \end{figure}
@@
 
-\caption{\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (middle panel).
-The bottom panel shows the change in accessibility.}\label{fig:plot-accessibility}
+\caption{\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (bottom panel).}\label{fig:plot-accessibility}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-accessibility-changes-1}
+
+}
+
+\caption{\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19.}\label{fig:plot-accessibility-changes}
 \end{figure}

### 2021-04-23 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd

--- 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
@@
 data(ttm_walk)
+```
+
+```{r remove-irrelevant-destinations, include=FALSE}
+## Data preparation
+# Remove entries in travel time matrices for inapplicable destinations
+# As a first step, we will remove the computed travel times to the three Community Kitchen locations (IDs 1-3):
+ttm_car <- ttm_car %>%
+  filter(!between(OBJECTID, left = 1, right = 3))
+
+ttm_transit <- ttm_transit %>%
+  filter(!between(OBJECTID, left = 1, right = 3))
+
+ttm_walk <- ttm_walk %>%
+  filter(!between(OBJECTID, left = 1, right = 3))
+```
+
+```{r join-low-income-households, include=FALSE}
+## Data preparation
+# Join proportion of low income households per DA to dwellings and total number of residential units:
+dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
+  left_join(data_da_2016 %>%
+              st_drop_geometry() %>%
+              dplyr::transmute(DAUID = GeoUID, 
+                               p_inc_less40k),
+            by = "DAUID")
+```
+
+```{r probability-of-low-income-per-parcel, include=FALSE}
+## Data preparation
+# Calculate probability that a point on the network will have low income households.
+This is the proportion of low income households in the DA multiplied by the total number of residential units at the point:
+dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
+  mutate(income_less40k = p_inc_less40k * Dwellings)
+```
+
+```{r number-of-residential-units-per-DA, include=FALSE}
+## Data preparation
+# Prepare and join population data from Transportation Tomorrow Survey
+# Find number of residential units in each DA. This is needed to assign the population to parcels
+residential_units <- dwelling_network_points_2016 %>%
+  st_drop_geometry() %>%
+  group_by(TAZUID) %>%
+  summarize(residential_units = sum(Dwellings))
+```
+
+```{r join-population-to-parcels, include=FALSE}
+## Data preparation
+# Join number of low income households per DA to dwellings and total number of residential units:
+dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
+  left_join(population_age_income %>%
+              st_drop_geometry() %>%
+              dplyr::transmute(TAZUID,
+                               children,
+                               adults,
+                               seniors),
+            by = "TAZUID") %>%
+  left_join(residential_units,
+            by = "TAZUID")
+```
+
+```{r calculate-low-income-households-per-parcel, include=FALSE}
+## Data preparation
+# Calculate population per point on the network.
+This is the population in the DA divided by the total number of residential units.
+This gives the average number of people per residential unit.
+Then, multiply by the number of dwellings (residential units) at the point:
+dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
+  mutate(children = (children/residential_units) * Dwellings,
+         adults = (adults/residential_units) * Dwellings,
+         seniors = (seniors/residential_units) * Dwellings)
+```
+
+```{r join-household-data-to-travel-time-tables, include=FALSE}
+## Data preparation
+# Join the household data to the travel time matrices:
+ttm_car <- ttm_car %>%
+  left_join(dwelling_network_points_2016 %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID, 
+                            income_less40k,
+                            children,
+                            adults,
+                            seniors,
+                            DAUID, 
+                            TAZUID),
+            by = c("UID" = "ID"))
+
+ttm_transit <- ttm_transit %>%
+  left_join(dwelling_network_points_2016 %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID, 
+                            income_less40k,
+                            children,
+                            adults,
+                            seniors,
+                            DAUID, 
+                            TAZUID),
+            by = c("UID" = "ID"))
+
+ttm_walk <- ttm_walk %>%
+  left_join(dwelling_network_points_2016 %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID, 
+                            income_less40k,
+                            children,
+                            adults,
+                            seniors,
+                            DAUID, 
+                            TAZUID),
+            by = c("UID" = "ID"))
+```
+
+```{r join-foodbank-data-to-travel-time-tables, include=FALSE}
+## Data preparation
+# Join the foodbank data to the travel time tables:
+ttm_car <- ttm_car %>%
+  left_join(foodbank_locations %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID, 
+                            TYPE,
+                            COVIDClose),
+            by = c("OBJECTID" = "ID"))
+
+ttm_transit <- ttm_transit %>%
+  left_join(foodbank_locations %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID, 
+                            TYPE,
+                            COVIDClose),
+            by = c("OBJECTID" = "ID"))
+
+ttm_walk <- ttm_walk %>%
+  left_join(foodbank_locations %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID,
+                            TYPE,
+                            COVIDClose),
+            by = c("OBJECTID" = "ID"))
+```
+
+```{r drop-nas-from-travel-time-tables, include=FALSE}
+## Data preparation
+# Drop NAs:
+ttm_car <- ttm_car %>%
+  drop_na(TAZUID,
+          income_less40k)
+
+ttm_transit <- ttm_transit %>%
+  drop_na(TAZUID,
+          income_less40k)
+
+ttm_walk <- ttm_walk %>%
+  drop_na(TAZUID,
+          income_less40k)
+```
+
+```{r proportion-of-trips-by-mode-by-taz, include=FALSE}
+## Data preparation
+# Calculate proportion of trips by mode per TAZ:
+modes <- modes_less40k %>%
+  mutate(p_car = Driver/(Driver + Transit + Walk),
+         p_transit = Transit/(Driver + Transit + Walk),
+         p_walk = Walk/(Driver + Transit + Walk))
+```
+
+```{r fill-nas-transport-modes, include=FALSE}
+## Data preparation
+# Assume that TAZ with NAs have a proportion of car trips of one:
+modes <- modes %>%
+  dplyr::mutate(p_car = replace_na(p_car, 1),
+                p_transit = replace_na(p_transit, 0),
+                p_walk = replace_na(p_walk, 0))
+```
+
+```{r join-trips-by-mode-to-travel-time-tables, include=FALSE}
+## Data preparation
+# Join trip-making modal proportions
+ttm_car <- ttm_car %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+
+ttm_transit <- ttm_transit %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+
+ttm_walk <- ttm_walk %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
+```
+
+```{r join-travel-time-tables, include=FALSE}
+## Data preparation
+# Join Tables into Single TTM
+ttm <- ttm_car %>%
+  rename(travel_time_car = travel_time) %>%
+  full_join(ttm_transit %>%
+              transmute(UID,
+                        OBJECTID,
+                        travel_time_transit = travel_time),
+            by = c("UID", "OBJECTID")) %>%
+  full_join(ttm_walk %>%
+              transmute(UID,
+                        OBJECTID,
+                        travel_time_walk = travel_time),
+            by = c("UID", "OBJECTID"))
+```
+
+```{r travel-time-tables-data-check, include=FALSE}
+## Data preparation
+# Check the NAs in travel time table:
+ttm %>% 
+  filter(is.na(travel_time_transit)) %>%
+  select(travel_time_transit,
+         p_transit) %>%
+  summary()
+
+# What do NAs here mean?
+Most DAs missing travel time by transit do not report trips by transit in the TTS.
+In some DAs there are trips by transit (p_transit != 0), but no foodbank can be reached by transit (travel_time_transit is NA).
+How many of these cases?
+ttm %>% 
+  filter(is.na(travel_time_transit) & p_transit > 0) %>%
+  select(travel_time_transit,
+         p_transit) %>%
+  summary()
+
+# It should be fine to replace the NAs with zeros, which multiplied by p_transit will return a zero (i.e., it is not possible to reach a foodbank by transit).
+ttm <- ttm %>% 
+  mutate(travel_time_transit = replace_na(travel_time_transit, 0))
+
+# Check the NAs in travel time by walk:
+ttm %>% 
+  filter(is.na(travel_time_walk)) %>%
+  select(travel_time_walk,
+         p_walk) %>%
+  summary()
+
+# Most DAs missing travel time by walking do not report trips by walking in the TTS.
+In some DAs there are trips by walking (p_walk != 0), but no foodbank can be reached by walk (travel_time_walk is NA).
+How many of these cases?
+ttm %>% 
+  filter(is.na(travel_time_walk) & p_walk > 0) %>%
+  select(travel_time_walk,
+         p_walk) %>%
+  summary()
+
+# Like with transit, it should be fine to replace the NAs with zeros, which multiplied by p_walk will return a zero (i.e., it is not possible to reach a foodbank by walking).
+ttm <- ttm %>% 
+  mutate(travel_time_walk = replace_na(travel_time_walk, 0))
+
+#Summary of the table after these operations:
+summary(ttm)
+```
+
+```{r calculate-expected-travel-times, include=FALSE}
+## Data preparation
+# Calculate expected travel times by mode.
+These are the travel time by the mode multiplied by the probability of using the mode (which is proxied by the proportion of trips by mode in the DA from the TTS).
+The total expected travel time is the sum of the expected travel times by mode:
+ttm <- ttm %>%
+  mutate(e_travel_time_car = travel_time_car * p_car,
+         e_travel_time_transit = travel_time_transit * p_transit,
+         e_travel_time_walk = travel_time_walk * p_walk,
+         e_travel_time = e_travel_time_car + e_travel_time_transit + e_travel_time_walk)
 ```
@@
         escape = FALSE) %>%
-  kable_styling(latex_options = c("striped", "scale_down")) 
+  kable_styling(latex_options = c("striped", "scale_down")) %>%
+  column_spec(2, width = "15em") %>%
+  column_spec(5, width = "15em")
 ```
@@
 
-```{r remove-irrelevant-destinations, include=FALSE}
-## Remove entries in travel time matrices for inapplicable destinations
-# As a first step, we will remove the computed travel times to the three Community Kitchen locations (IDs 1-3):
-ttm_car <- ttm_car %>%
-  filter(!between(OBJECTID, left = 1, right = 3))
+Figure \ref{fig:foodbanks} shows the location of the foodbanks in the City of Hamilton and their status.
+Before the pandemic there were `r nrow(foodbank_locations)` of which `r nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))` closed during the pandemic.
+As seen in the figure, foodbanks tend to be predominantly located in the central parts of the city.
+This is not surprising: population density is high there, and it is also the part of the city where lower income households are more numerous in absolute and relative terms (see Figure \ref{fig:low-income-households}).
+Alas, this is also where most of the closures during the pandemic happened.
 
-ttm_transit <- ttm_transit %>%
-  filter(!between(OBJECTID, left = 1, right = 3))
+```{r plot-location-foodbanks, echo=FALSE, fig.cap="\\label{fig:foodbanks}Location of foodbanks and operation status; the dotted box is an inset of the central part of the City of Hamilton"}
+## Analysis
+# Plot location of foodbanks
 
-ttm_walk <- ttm_walk %>%
-  filter(!between(OBJECTID, left = 1, right = 3))
+# Create a bounding box for an inset
+inset_box <- st_bbox(c(xmin = -79.975,
+          xmax = -79.775, 
+          ymax = 43.2, 
+          ymin = 43.3), 
+        crs = st_crs(4326)) %>%
+  st_as_sfc()
+
+fb <- ggplot() +
+  geom_sf(data = hamilton_cma,
+          fill = "lightgray") +
+  geom_sf(data = foodbank_locations,
+          aes(size = COVIDClose,
+              shape = COVIDClose)) +
+  geom_sf(data = inset_box,
+          linetype = "dotted",
+          color = "red",
+          fill = NA) +
+  scale_shape_manual(name = "Closed during COVID-19",
+                     values = c("Yes" = 4, "No" = 1)) +
+  scale_size_manual(values = c("Yes" = 3, "No" = 2)) +
+  guides(size = FALSE) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "none")
+
+fb_inset <- ggplot() +
+  geom_sf(data = hamilton_cma,
+          fill = "lightgray") +
+  geom_sf(data = foodbank_locations,
+          aes(size = COVIDClose,
+              shape = COVIDClose)) +
+  geom_sf(data = inset_box,
+          linetype = "dotted",
+          color = "red",
+          fill = NA) +
+  scale_shape_manual(name = "Closed during COVID-19",
+                     values = c("Yes" = 4, "No" = 1)) +
+  scale_size_manual(values = c("Yes" = 3, "No" = 2)) +
+  xlim(-79.975, -79.775) +
+  ylim(43.2, 43.3) +
+  guides(size = FALSE) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom")
+
+fb / fb_inset
 ```
 
-```{r join-low-income-households, include=FALSE}
-## Join proportion of low income households per DA to dwellings and total number of residential units:
-dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
-  left_join(data_da_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::transmute(DAUID = GeoUID, 
-                               p_inc_less40k),
-            by = "DAUID")
+```{r plot-low-income-households, echo=FALSE, fig.cap="\\label{fig:low-income-households}Number and proportion of households with incomes less than CAD 40,000."}
+lih <- ggplot() +
+  geom_sf(data = data_da_2016,
+          aes(fill = income_less40k),
+          color = NA) +
+  geom_sf(data = inset_box,
+          linetype = "dotted",
+          color = "red",
+          fill = NA) +
+  scale_fill_viridis_c(name = "Number of households", 
+                       option = "A",
+                       direction = -1) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "bottom")
+
+p_lih <- ggplot() +
+  geom_sf(data = data_da_2016,
+          aes(fill = p_inc_less40k),
+          color = NA) +
+  geom_sf(data = inset_box,
+          linetype = "dotted",
+          color = "red",
+          fill = NA) +
+  scale_fill_viridis_c(name = "Proportion of households", 
+                       option = "A",
+                       direction = -1) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "bottom")
+
+lih / p_lih
 ```
-
-```{r probability-of-low-income-per-parcel, include=FALSE}
-## Calculate probability that a point on the network will have low income households.
-This is the proportion of low income households in the DA multiplied by the total number of residential units at the point:
-dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
-  mutate(income_less40k = p_inc_less40k * Dwellings)
-```
-
-```{r number-of-residential-units-per-DA, include=FALSE}
-## Data preparation
-# Prepare and join population data from Transportation Tomorrow Survey
-# Find number of residential units in each DA. This is needed to assign the population to parcels
-residential_units <- dwelling_network_points_2016 %>%
-  st_drop_geometry() %>%
-  group_by(TAZUID) %>%
-  summarize(residential_units = sum(Dwellings))
-```
-
-```{r join-population-to-parcels, include=FALSE}
-## Data preparation
-# Join number of low income households per DA to dwellings and total number of residential units:
-dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
-  left_join(population_age_income %>%
-              st_drop_geometry() %>%
-              dplyr::transmute(TAZUID,
-                               children,
-                               adults,
-                               seniors),
-            by = "TAZUID") %>%
-  left_join(residential_units,
-            by = "TAZUID")
-```
-
-```{r calculate-low-income-households-per-parcel, include=FALSE}
-## Data preparation
-# Calculate population per point on the network.
-This is the population in the DA divided by the total number of residential units.
-This gives the average number of people per residential unit.
-Then, multiply by the number of dwellings (residential units) at the point:
-dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
-  mutate(children = (children/residential_units) * Dwellings,
-         adults = (adults/residential_units) * Dwellings,
-         seniors = (seniors/residential_units) * Dwellings)
-```
-
-```{r join-household-data-to-travel-time-tables, include=FALSE}
-## Data preparation
-# Join the household data to the travel time matrices:
-ttm_car <- ttm_car %>%
-  left_join(dwelling_network_points_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            income_less40k,
-                            children,
-                            adults,
-                            seniors,
-                            DAUID, 
-                            TAZUID),
-            by = c("UID" = "ID"))
-
-ttm_transit <- ttm_transit %>%
-  left_join(dwelling_network_points_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            income_less40k,
-                            children,
-                            adults,
-                            seniors,
-                            DAUID, 
-                            TAZUID),
-            by = c("UID" = "ID"))
-
-ttm_walk <- ttm_walk %>%
-  left_join(dwelling_network_points_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            income_less40k,
-                            children,
-                            adults,
-                            seniors,
-                            DAUID, 
-                            TAZUID),
-            by = c("UID" = "ID"))
-```
-
-```{r join-foodbank-data-to-travel-time-tables, include=FALSE}
-## Data preparation
-# Join the food bank data to the travel time tables:
-ttm_car <- ttm_car %>%
-  left_join(foodbank_locations %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            TYPE,
-                            COVIDClose),
-            by = c("OBJECTID" = "ID"))
-
-ttm_transit <- ttm_transit %>%
-  left_join(foodbank_locations %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            TYPE,
-                            COVIDClose),
-            by = c("OBJECTID" = "ID"))
-
-ttm_walk <- ttm_walk %>%
-  left_join(foodbank_locations %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID,
-                            TYPE,
-                            COVIDClose),
-            by = c("OBJECTID" = "ID"))
-```
-
-```{r drop-nas-from-travel-time-tables, include=FALSE}
-## Data preparation
-# Drop NAs:
-ttm_car <- ttm_car %>%
-  drop_na(TAZUID,
-          income_less40k)
-
-ttm_transit <- ttm_transit %>%
-  drop_na(TAZUID,
-          income_less40k)
-
-ttm_walk <- ttm_walk %>%
-  drop_na(TAZUID,
-          income_less40k)
-```
-
-```{r proportion-of-trips-by-mode-by-taz, include=FALSE}
-## Data preparation
-# Calculate proportion of trips by mode per TAZ:
-modes <- modes_less40k %>%
-  mutate(p_car = Driver/(Driver + Transit + Walk),
-         p_transit = Transit/(Driver + Transit + Walk),
-         p_walk = Walk/(Driver + Transit + Walk))
-```
-
-```{r fill-nas-transport-modes, include=FALSE}
-## Data preparation
-# Assume that TAZ with NAs have a proportion of car trips of one:
-modes <- modes %>%
-  dplyr::mutate(p_car = replace_na(p_car, 1),
-                p_transit = replace_na(p_transit, 0),
-                p_walk = replace_na(p_walk, 0))
-```
-
-```{r join-trips-by-mode-to-travel-time-tables, include=FALSE}
-## Data preparation
-# Join trip-making modal proportions
-ttm_car <- ttm_car %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-
-ttm_transit <- ttm_transit %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-
-ttm_walk <- ttm_walk %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-```
-
-```{r join-travel-time-tables, include=FALSE}
-## Data preparation
-# Join Tables into Single TTM
-ttm <- ttm_car %>%
-  rename(travel_time_car = travel_time) %>%
-  full_join(ttm_transit %>%
-              transmute(UID,
-                        OBJECTID,
-                        travel_time_transit = travel_time),
-            by = c("UID", "OBJECTID")) %>%
-  full_join(ttm_walk %>%
-              transmute(UID,
-                        OBJECTID,
-                        travel_time_walk = travel_time),
-            by = c("UID", "OBJECTID"))
-```
-
-```{r travel-time-tables-data-check, include=FALSE}
-## Data preparation
-# Check the NAs in travel time table:
-ttm %>% 
-  filter(is.na(travel_time_transit)) %>%
-  select(travel_time_transit,
-         p_transit) %>%
-  summary()
-
-# What do NAs here mean?
-Most DAs missing travel time by transit do not report trips by transit in the TTS.
-In some DAs there are trips by transit (p_transit != 0), but no foodbank can be reached by transit (travel_time_transit is NA).
-How many of these cases?
-ttm %>% 
-  filter(is.na(travel_time_transit) & p_transit > 0) %>%
-  select(travel_time_transit,
-         p_transit) %>%
-  summary()
-
-# It should be fine to replace the NAs with zeros, which multiplied by p_transit will return a zero (i.e., it is not possible to reach a foodbank by transit).
-ttm <- ttm %>% 
-  mutate(travel_time_transit = replace_na(travel_time_transit, 0))
-
-# Check the NAs in travel time by walk:
-ttm %>% 
-  filter(is.na(travel_time_walk)) %>%
-  select(travel_time_walk,
-         p_walk) %>%
-  summary()
-
-# Most DAs missing travel time by walking do not report trips by walking in the TTS.
-In some DAs there are trips by walking (p_walk != 0), but no foodbank can be reached by walk (travel_time_walk is NA).
-How many of these cases?
-ttm %>% 
-  filter(is.na(travel_time_walk) & p_walk > 0) %>%
-  select(travel_time_walk,
-         p_walk) %>%
-  summary()
-
-# Like with transit, it should be fine to replace the NAs with zeros, which multiplied by p_walk will return a zero (i.e., it is not possible to reach a foodbank by walking).
-ttm <- ttm %>% 
-  mutate(travel_time_walk = replace_na(travel_time_walk, 0))
-
-#Summary of the table after these operations:
-summary(ttm)
-```
-
-```{r calculate-expected-travel-times, include=FALSE}
-## Data preparation
-# Calculate expected travel times by mode.
-These are the travel time by the mode multiplied by the probability of using the mode (which is proxied by the proportion of trips by mode in the DA from the TTS).
-The total expected travel time is the sum of the expected travel times by mode:
-ttm <- ttm %>%
-  mutate(e_travel_time_car = travel_time_car * p_car,
-         e_travel_time_transit = travel_time_transit * p_transit,
-         e_travel_time_walk = travel_time_walk * p_walk,
-         e_travel_time = e_travel_time_car + e_travel_time_transit + e_travel_time_walk)
-```
-
-Words go here.
 
@@
 
+```{r, join-geometry-to-level-of-service, include=FALSE}
+## Analysis 
+# Join geometry:
+los <- los %>%
+  left_join(foodbank_locations %>%
+              dplyr::select(ID),
+            by = c("OBJECTID" = "ID")) %>%
+  st_as_sf()
+```
+
+```{r plot-levels-of-service, echo=FALSE, fig.cap="\\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
+## Analysis
+# Plot levels of service before and during covid
+
+# Pre-covid
+los_pre <- ggplot() +
+  geom_sf(data = hamilton_cma) +
+  geom_sf(data = los %>%
+            filter(timing == "Pre-covid"),
+          aes(color = los,
+              size = los)) +
+  geom_sf(data = foodbank_locations,
+          aes(shape = COVIDClose)) +
+  scale_color_viridis_c(name = "Level of service pre-COVID-19",
+                        option =  "A", 
+                        direction = -1) +
+  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
+  guides(size = FALSE, shape = FALSE) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
+
+# During covid
+los_covid <- ggplot() +
+  geom_sf(data = hamilton_cma) +
+  geom_sf(data = los %>%
+            filter(timing == "During covid"),
+          aes(color = los,
+              size = los)) +
+  geom_sf(data = foodbank_locations,
+          aes(shape = COVIDClose)) +
+  scale_color_viridis_c(name = "Level of service during COVID-19",
+                        option =  "A", 
+                        direction = -1) +
+  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
+  guides(size = FALSE, shape = FALSE) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
+
+los_pre / los_covid
+```
+
+```{r plot-levels-of-service-changes, echo=FALSE, fig.cap="\\label{fig:levels-of-service-changes}Changes in levels of service at each facility from pre-COVID-19 to during COVID-19."}
+# This figure shows the places where the level of service deteriorated most (excluding the places that closed, where it vanished).
+los_change <- los %>%
+  st_drop_geometry() %>%
+  pivot_wider(names_from = timing, 
+              values_from = los, 
+              values_fill = 0) %>%
+  filter(`During covid` > 0) %>%
+  mutate(los_d = `During covid` - `Pre-covid`) %>%
+  left_join(foodbank_locations %>%
+            transmute(OBJECTID = ID,
+                      geometry),
+          by = "OBJECTID") %>%
+  st_as_sf() %>%
+  ggplot() +
+  geom_sf(data = hamilton_cma) +
+  geom_sf(aes(color = los_d,
+              size = los_d)) +
+  geom_sf(data = foodbank_locations,
+          aes(shape = COVIDClose)) +
+  scale_color_viridis_c(name = "Change in level of service",
+                        option =  "A", 
+                        direction = -1) +
+  scale_size(trans = 'reverse') +
+  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
+  guides(size = FALSE, shape = FALSE) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
+
+los_change
+```
+
+```{r aggregate-accessibility-by-taz, include=FALSE}  
+## Analysis
+# Accessibility by TAZ:
+accessibility_taz <- accessibility %>%
+  group_by(TAZUID, 
+           timing) %>%
+  summarize(accessibility = sum(accessibility),
+            timing = first(timing),
+            .groups = "drop") %>%
+  left_join(population_age_income %>%
+              transmute(TAZUID,
+                        children,
+                        adults,
+                        seniors),
+            by = "TAZUID") %>%
+  st_as_sf()
+```
+
+```{r plot-accessibility, echo=FALSE, fig.align = 'center', out.width = "90%", fig.cap="\\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
+## Analysis
+# Plot accessibility before and during covid:
+acc_pre <- ggplot() +
+  #geom_sf(data = hamilton_cma) +
+  geom_sf(data = accessibility_taz %>%
+            filter(timing == "Pre-covid"),
+          aes(fill = accessibility)) +
+  scale_fill_viridis_c(name = "Accessibility pre-COVID-19",
+                        option =  "A", 
+                        direction = -1,
+                       breaks = c(0.001,
+                                  0.0008,
+                                  0.0006,
+                                  0.0004,
+                                  0.0002),
+                       labels = c("0.0010",
+                                  "0.0008",
+                                  "0.0006",
+                                  "0.0004",
+                                  "0.0002")) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
+
+acc_covid <- ggplot() +
+  #geom_sf(data = hamilton_cma) +
+  geom_sf(data = accessibility_taz %>%
+            filter(timing == "During covid"),
+          aes(fill = accessibility)) +
+  scale_fill_viridis_c(name = "Accessibility during COVID-19",
+                        option =  "A", 
+                        direction = -1,
+                       breaks = c(0.0005,
+                                  0.0004,
+                                  0.0003,
+                                  0.0002,
+                                  0.0001),
+                       labels = c("0.0005",
+                                  "0.0004",
+                                  "0.0003",
+                                  "0.0002",
+                                  "0.0001")) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
+
+acc_pre / acc_covid
+```
+
+```{r plot-accessibility-changes, echo=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19."}
+## Analysis
+# Calculate accessibility change by TAZ:
+accessibilty_taz_change <- accessibility_taz %>%
+  st_drop_geometry() %>%
+  pivot_wider(names_from = timing, 
+              values_from = accessibility) %>%
+  # Calculate changes in accessibility
+  mutate(acc_d = `During covid` - `Pre-covid`) %>%
+  # Join TAZ geometry
+  left_join(population_age_income %>%
+            transmute(TAZUID,
+                      geometry),
+          by = "TAZUID") %>%
+  # Convert to simple features
+  st_as_sf()
+
+# Plot
+acc_change <- ggplot() +
+  #geom_sf(data = hamilton_cma) +
+  geom_sf(data = accessibilty_taz_change,
+          aes(fill = acc_d),
+          color = NA) +
+  geom_sf(data = population_age_income,
+          fill = NA) +
+  scale_fill_viridis_c(name = "Change in accessibility",
+                        option =  "A", 
+                        direction = 1,
+                       breaks = c(-0.0005,
+                                  -0.0004,
+                                  -0.0003,
+                                  -0.0002,
+                                  -0.0001),
+                       labels = c("-0.0005",
+                                  "-0.0004",
+                                  "-0.0003",
+                                  "-0.0002",
+                                  "-0.0001")) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
+
+acc_change
+```
+
 ```{r accessibility-change-by-parcel, include=FALSE}
@@
 
-```{r, join-geometry-to-level-of-service, include=FALSE}
-## Analysis 
-# Join geometry:
-los <- los %>%
-  left_join(foodbank_locations %>%
-              dplyr::select(ID),
-            by = c("OBJECTID" = "ID")) %>%
-  st_as_sf()
-```
-
-```{r plot-levels-of-service, echo=FALSE, fig.cap="\\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (middle panel).
-The bottom panel shows the change in level of service."}
-## Analysis
-# Plot levels of service before and during covid
-
-# Pre-covid
-los_pre <- ggplot() +
-  geom_sf(data = hamilton_cma) +
-  geom_sf(data = los %>%
-            filter(timing == "Pre-covid"),
-          aes(color = los,
-              size = los)) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  scale_color_distiller(palette = "OrRd", 
-                        direction = 1) +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  theme_minimal()
-
-# During covid
-los_covid <- ggplot() +
-  geom_sf(data = hamilton_cma) +
-  geom_sf(data = los %>%
-            filter(timing == "During covid"),
-          aes(color = los,
-              size = los)) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  scale_color_distiller(palette = "OrRd", 
-                        direction = 1) +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  theme_minimal()
-
-# This figure shows the places where the level of service deteriorated most (excluding the places that closed, where it vanished).
-los_change <- los %>%
-  st_drop_geometry() %>%
-  pivot_wider(names_from = timing, 
-              values_from = los, 
-              values_fill = 0) %>%
-  filter(`During covid` > 0) %>%
-  mutate(los_d = `During covid` / `Pre-covid` - 1) %>%
-  left_join(foodbank_locations %>%
-            transmute(OBJECTID = ID,
-                      geometry),
-          by = "OBJECTID") %>%
-  st_as_sf() %>%
-  ggplot() +
-  geom_sf(data = hamilton_cma) +
-  geom_sf(aes(color = los_d,
-              size = los_d)) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  scale_color_distiller(palette = "OrRd", 
-                        direction = -1) +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  theme_minimal()
-
-los_pre / los_covid / los_change
-```
-
-```{r aggregate-accessibility-by-taz, include=FALSE}  
-## Analysis
-# Accessibility by TAZ:
-accessibility_taz <- accessibility %>%
-  group_by(TAZUID, 
-           timing) %>%
-  summarize(accessibility = sum(accessibility),
-            timing = first(timing),
-            .groups = "drop") %>%
-  left_join(population_age_income %>%
-              transmute(TAZUID,
-                        children,
-                        adults,
-                        seniors),
-            by = "TAZUID") %>%
-  st_as_sf()
-```
-
-```{r plot-accessibility, echo=FALSE, fig.align = 'center', out.width = "90%", fig.cap="\\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (middle panel).
-The bottom panel shows the change in accessibility."}
-## Analysis
-# Plot accessibility before and during covid:
-acc_pre <- ggplot() +
-  geom_sf(data = hamilton_cma) +
-  geom_sf(data = accessibility_taz %>%
-            filter(timing == "Pre-covid"),
-          aes(fill = accessibility)) +
-  scale_fill_distiller(palette = "OrRd", 
-                        direction = 1) +
-  theme_minimal()
-
-acc_covid <- ggplot() +
-  geom_sf(data = hamilton_cma) +
-  geom_sf(data = accessibility_taz %>%
-            filter(timing == "During covid"),
-          aes(fill = accessibility)) +
-  scale_fill_distiller(palette = "OrRd", 
-                        direction = 1) +
-  theme_minimal()
-
-# Calculate accessibility change by TAZ:
-accessibilty_taz_change <- accessibility_taz %>%
-  st_drop_geometry() %>%
-  pivot_wider(names_from = timing, 
-              values_from = accessibility) %>%
-  # Calculate changes in accessibility
-  mutate(acc_d = `During covid` - `Pre-covid`) %>%
-  # Join TAZ geometry
-  left_join(population_age_income %>%
-            transmute(TAZUID,
-                      geometry),
-          by = "TAZUID") %>%
-  # Convert to simple features
-  st_as_sf()
-
-# Plot
-acc_change <- ggplot() +
-  geom_sf(data = hamilton_cma) +
-  geom_sf(data = accessibilty_taz_change,
-          aes(fill = acc_d),
-          color = NA) +
-  geom_sf(data = population_age_income,
-          fill = NA) +
-  scale_fill_fermenter(palette = "OrRd", 
-                       direction = -1) +
-  theme_minimal()
-
-acc_pre / acc_covid / acc_change
-```
-
 Conclusions

### 2021-04-23 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex

--- 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
@@
 Data have been prepared for sharing as a data package\footnote{add
-  repository}.
+  repository}. This data package consists of the following:
 
@@
 
-Population and income statistics at the level of Dissemination Areas
-(DAs) were retrieved using the package \texttt{cancensus} (von Bergmann
-et al., 2021). Dissemination Areas are the smallest publicly available
-census geography in Canada. We use data from the 2016 Population Census.
+Population and income statistics for 2016 were retrieved at the level of
+Dissemination Areas (DAs) using the package \texttt{cancensus} (von
+Bergmann et al., 2021). DAs are the smallest publicly available census
+geography in Canada. Income data corresponds to the count of households
+by different total income groupings.
 
@@
 
-We converted all recorded residential parcels in the City of Hamilton to
-points on the road network. Each point includes information about the
-number of residential units in the parcel. Each residential unit is
-``populated'' with the probability of being a ``low income household,''
-which we define as having a total household income of less than CAD
-40,000. This is approximately the mid-point of the low income cut-off
-(LICOs) for families in Canadian cities with populations greater than
-500,000 in 2016, to match other Census data (Statistics Canada, 2020b).
+We converted all recorded residential land parcels in the City of
+Hamilton to points on the road network. Each point includes information
+about the number of residential units in the parcel. Next, we define
+low-income households as those having a total income of less than
+CAD\$40,000, which is approximately the mid-point of the low income
+cut-off (LICOs) for families in Canadian cities with populations greater
+than 500,000 in 2016, to match other Census data (Statistics Canada,
+2020b). We then ``populate'' each residential unit with the probability
+of being a low-income household based on the counts of households by
+income groups in the DA in which the parcel is located. While this
+method assumes a constant probability of low-income household status for
+all residential units in a DA, the parcel-level analysis affords a high
+level of spatial disaggregation for the accessibility analysis.
 
-\hypertarget{destinations-foodbanks}{%
-\subsubsection{Destinations: Foodbanks}\label{destinations-foodbanks}}
+\hypertarget{destinations-food-banks-and-food-service-locations}{%
+\subsubsection{Destinations: Food Banks and Food Service
+Locations}\label{destinations-food-banks-and-food-service-locations}}
 
-The locations of foodbanks were obtained from public records and
-geocoded. Three urban sites not in the program were also identified and
-geocoded for comparison purposes.
+The locations of food banks and related food services were obtained from
+the Hamilton Public Library's Food Access Guide\footnote{http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf}.
+The guide was updated in April of 2021 to indicate any change affected
+on the services due to the pandemic. This includes modified business
+hours, a need to make reservations before frequenting, and locations
+that have completely shut down in consequence. While some food bank
+services have a specific a target population, such as prioritizing
+family with young children aged between 0 and 3 or accepting only those
+providing proof of low-income status through housing and utility
+statements, all the food bank services indicated below are designed to
+accommodate those in need of food at zero to low cost. With our focus on
+food banks and food services that offer free or low-cost meals at
+particular locations, we first removed services such as Meals on Wheels
+and other food access services such as food box, community kitchens,
+student nutrition programs, and shopping and transportation. In
+addition, two free meal services held on different days at the same
+location were collapsed into a single service point for the
+accessibility analysis.
+
+\begin{table}
+
+\caption{\label{tab:table-food-bank-info}\label{tab:food-bank-info}Food Bank and Food Service Information.}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{c>{\centering\arraybackslash}p{15em}cc>{\centering\arraybackslash}p{15em}}
+\toprule
+Type & Description & Locations Pre-COVID & Locations During COVID & Additional Notes\\
+\midrule
+\cellcolor{gray!6}{Congregate Dining} & \cellcolor{gray!6}{Congregate and dining programs provide low-cost meals that are enjoyed in a community setting.
+Transportation may be provided} & \cellcolor{gray!6}{7} & \cellcolor{gray!6}{2} & \cellcolor{gray!6}{One remaining location reduced hours during COVID}\\
+Community Meals & Programs often run by volunteers that organize suppers, lunches or other get-togethers that give community residents an opportunity to meet one another in a friendly and informal atmosphere while sharing a meal & 11 & 9 & NA\\
+\cellcolor{gray!6}{Food Banks} & \cellcolor{gray!6}{Food Banks and Emergency Food programs provide individuals and families with grocery items free of charge} & \cellcolor{gray!6}{27} & \cellcolor{gray!6}{26} & \cellcolor{gray!6}{One remaining location reduced hours during COVID while 4 others moved to appointment only}\\
+Free Meals & Meals are provided free of charge in the community through volunteer labour and donations & 9 & 5 & One remaining location reduced hours during COVID\\
+\cellcolor{gray!6}{Low-Cost Meals} & \cellcolor{gray!6}{Restaurants, cafeterias and other eating establishments operated by hospitals, senior centers or other organizations which provide reduced-cost meals for low-income people, older adults or other targeted individuals.} & \cellcolor{gray!6}{2} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{The remaining location reduced hours during COVID}\\
+\bottomrule
+\end{tabular}}
+\end{table}
 
@@
 the destinations. For routing, the package \texttt{r5r} (Pereira et al.,
-2021b) was used, with a network extract for the City of Hamilton from
+2021b) was used with a network extract for the City of Hamilton from
 OpenStreetMaps and the General Transit Feed Specification from Hamilton
@@
 households) and proportion of trips by mode, we calculated the expected
-travel time \(ett\) from each parcel \(i\) to a foodbank \(j\) as
-follows: \[
+travel time \(ett\) from each parcel \(i\) to a food bank or food
+service location \(j\) as follows: \[
 ett_i = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
@@
 the TAZ of parcel \(i\), and \(tt^k_{ij}\) is the travel time from
-parcel \(i\) to the foodbank. In other words, the expected travel time
+parcel \(i\) to the food bank. In other words, the expected travel time
 is the weighted sum of travel times to the foodbank, with the weights
@@
 
-Words go here.
+Figure \ref{fig:foodbanks} shows the location of the foodbanks in the
+City of Hamilton and their status. Before the pandemic there were 58 of
+which 14 closed during the pandemic. As seen in the figure, foodbanks
+tend to be predominantly located in the central parts of the city. This
+is not surprising: population density is high there, and it is also the
+part of the city where lower income households are more numerous in
+absolute and relative terms (see Figure
+\ref{fig:low-income-households}). Alas, this is also where most of the
+closures during the pandemic happened.
 
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-results-sensitivity-analysis-1} \caption{\label{fig:sensitivity-analysis}Accessibility as a function of threshold}\label{fig:plot-results-sensitivity-analysis}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-location-foodbanks-1} \caption{\label{fig:foodbanks}Location of foodbanks and operation status; the dotted box is an inset of the central part of the City of Hamilton}\label{fig:plot-location-foodbanks}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-low-income-households-1} \caption{\label{fig:low-income-households}Number and proportion of households with incomes less than CAD 40,000.}\label{fig:plot-low-income-households}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-results-sensitivity-analysis-1} \caption{\label{fig:sensitivity-analysis}Accessibility as a function of threshold}\label{fig:plot-results-sensitivity-analysis}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-levels-of-service-1} \caption{\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (bottom panel).}\label{fig:plot-levels-of-service}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-levels-of-service-changes-1} \caption{\label{fig:levels-of-service-changes}Changes in levels of service at each facility from pre-COVID-19 to during COVID-19.}\label{fig:plot-levels-of-service-changes}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.9\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-accessibility-1}
+
+}
+
+\caption{\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (bottom panel).}\label{fig:plot-accessibility}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-accessibility-changes-1}
+
+}
+
+\caption{\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19.}\label{fig:plot-accessibility-changes}
 \end{figure}
@@
 \end{table}
-
-\begin{figure}
-\includegraphics[width=0.9\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-levels-of-service-1} \caption{\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (middle panel).
-The bottom panel shows the change in level of service.}\label{fig:plot-levels-of-service}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-accessibility-1}
-
-}
-
-\caption{\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (middle panel).
-The bottom panel shows the change in accessibility.}\label{fig:plot-accessibility}
-\end{figure}
 

### 2021-04-23 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Continued editing the results and discussion

--- 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
@@
 knitr::opts_chunk$set(
-  cache = TRUE, out.width = "100%", fig.width = 6, fig.height = 6
+  cache = TRUE, out.width = "100%", align = 'center', fig.width = 6, fig.height = 6
 )
@@
 Food insecurity is defined as an "inadequate or uncertain access to a sufficient quantity and/or adequate quality of food" due to a household's financial limitations [@enns2020experiences].
-This is a condition that has been associated with reductions in nutritional outcomes [@bhattacharya2004poverty; @kirkpatrick2008food; @olson1999nutrition] and physical and mental health in children and adults [@elgar2021relative; @jones2017food; @ramsey2011food; @seligman2010food; @stuff2004household].
+This condition has been associated with reductions in nutritional outcomes [@bhattacharya2004poverty; @kirkpatrick2008food; @olson1999nutrition] and physical and mental health in children and adults [@elgar2021relative; @jones2017food; @ramsey2011food; @seligman2010food; @stuff2004household].
 Over at least the past four decades foodbanks have become an essential line of defense against food insecurity in Canadian communities [@riches2002food; @holmes2018nothing; @black2020examining; @tarasuk2020relationship].
@@
 
-But the emergence of COVID-19, the worst public health crisis since the 1918 flu pandemic, has revealed important social and economic fault lines, and pre-existing patterns of inequality appear to have been exacerbated.
-Along several other dimensions of stress [such as accessibility to health care facilities, see @pereira2021geographic], this seems to be the case for food insecurity as well [@laborde2020poverty].
+The emergence of COVID-19, the worst public health crisis since the 1918 flu pandemic, has revealed important social and economic fault lines, and pre-existing patterns of inequality appear to have been exacerbated.
+Along several other dimensions of stress [e.g., accessibility to health care facilities, @pereira2021geographic], this seems to be the case for food insecurity as well [@laborde2020poverty].
 In the US, for example, it has been estimated that there was an increase of more than 30\% in household food insecurity, and more than one third of households were discovered to be newly food insecure - meaning they did not experience food insecurity before the pandemic [@niles2020early].
@@
 While foodbanks are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least they can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
-As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that foodbanks in Hamilton, Ontario, recorded the highest number of visitors in the past 29 years; the number of children visiting foodbanks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278 the year before.
-
-It is known that the urban food environment, within which people make their daily food choices, is essential in influencing eating behaviours and health outcomes, based on factors such as food availability, ease of geographic accessibility and socio-demographic variations [@paez2010relative; @vanderlee2017food; @widener2018spatial].
+As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that foodbanks in Hamilton, Ontario, recorded the highest number of visitors in the past 29 years; the number of children visiting foodbanks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278 the year before, a rate of increase greater than population growth.
+
+It is known that urban food environments, within which people make their daily food choices, are essential in influencing eating behaviours and health outcomes, based on factors such as food availability, ease of geographic accessibility and socio-demographic variations [@paez2010relative; @vanderlee2017food; @widener2018spatial].
 However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to foodbanks.
@@
 The study of place-based geographic accessibility is concerned with capturing the potential to reach destinations of value using the transportation network [@paez2012measuring].
-Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities^[https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with foodbank service delivery in Vancouver [@holmes2018nothing].
-However, as with research into the prevalence of food deserts, accessibility to foodbanks is unlikely to be evenly distributed throughout a city due to variations in the characteristics of the transportation network characteristics, and the spatial distribution of foodbank locations and the population they serve.
-Furthermore, policy responses to the COVID-19 pandemic may well add to the distress of vulnerable households.
-Non-pharmaceutical interventions during the pandemic involving restrictions in mobility have increased the friction of travel, in particular by transit on which low income populations are more likely to be reliant [e.g., @deweese2020tale].
+Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities^[https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with foodbank service delivery [@holmes2018nothing].
+However, as with research into the prevalence of food deserts, accessibility to foodbanks is unlikely to be evenly distributed, and variation throughout a city can be expected due to transportation network characteristics, and the spatial distribution of foodbank locations and the population they are meant to serve.
+Furthermore, policy responses to the COVID-19 pandemic likely have added to the distress of vulnerable households.
+Non-pharmaceutical interventions during the pandemic involving restrictions in mobility have increased the friction of travel, in particular by transit on which low income populations are more reliant [e.g., @deweese2020tale].
 At the same time, the pandemic has created additional stress for the operators of foodbanks through disruptions in the supply chain [e.g., @mckay2021exploring] as well as concerns surrounding the delivery of service in safe conditions and possible cancellation of food service programs.
@@
 low-cost or free meal service providers) available in Hamilton, Ontario, changed during the pandemic.
-Have the number of open foodbank services diminished?
-If so, what were the levels of accessibility to foodbanks before the pandemic from the perspective of low income households, and how did they change during the pandemic?
-And finally, who are most likely to have been impacted by changes to the accessibility landscape?
-This paper will first look at the distribution of foodbanks and related services before and during the pandemic.
+Did the number of open foodbank services diminish?
+If so, what was the accessibility to foodbanks before the pandemic from the perspective of low income households, and how it changed during the pandemic?
+And finally, who are most likely to have been impacted by changes in the accessibility landscape?
+This paper first looks at the distribution of foodbanks and related services before and during the pandemic.
 Then, we use the balanced floating catchment area approach of Paez et al.
 [-@paez2019demand] to investigate the accessibility situation.
-We use a fully disaggregated approach based on parcel-level data.
+For this, we use a fully disaggregated approach based on parcel-level data.
 Socio-economic and demographic data are drawn from the latest Census of Canada (2016), whereas travel information is from the most recent regional travel survey from 2016.
@@
 
-Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to insufficient nutrient intake [@enns2020experiences; @kirkpatrick2008food; @tarasuk2009household].
-Nutrient deficiency, in turn, is associated with major population health concerns, and particularly among those who are at a socio-economic disadvantage [@bazerghi2016role].
+Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake [@enns2020experiences; @kirkpatrick2008food; @tarasuk2009household].
+This nutrient deficiency is associated with major population health concerns, particularly among Canadians at socio-economic disadvantage [@bazerghi2016role].
 Data on food insecurity in Canada is collected in several ways.
@@
 The scope and objectives of foodbanks can vary by region and by country, and these organizations can include not only prepared meals and aliments, but also shared spaces to connect in community gardens and community kitchens [@wakefield2013sweet].
-Although in their origin foodbanks were meant to be provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time they evolved into a community practice to secure emergency food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
+Although in their origin foodbanks were meant to be provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time they have evolved into a community practice to secure emergency food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
 
 In Canada, the number of foodbanks has steadily increased in the past few decades [@wakefield2013sweet].
-The largest database of foodbanks and their use comes from the non-profit association foodbanks Canada, which conducts an annual assessment through its affiliated members.
-foodbanks Canada's 2018 Hunger Count report^[https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018_FINAL_EN.pdf.aspx?ext=.pdf] (the most recent available) listed 1,830 member foodbanks across the country found that Canadians visited foodbanks 1.1 million times in March of 2018.
+The largest database of foodbanks and their use comes from the non-profit association Foodbanks Canada, which conducts an annual assessment through its affiliated members.
+Foodbanks Canada's 2018 Hunger Count report^[https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018_FINAL_EN.pdf.aspx?ext=.pdf] (the most recent available) listed 1,830 member foodbanks across the country, and found that Canadians visited foodbanks 1.1 million times in March of 2018.
 Of those accessing foodbanks, certain population characteristics tend to be over-represented compared to national totals from the 2016 Canadian Census of Population.
-According to foodbanks Canada's 2018 data, single-adult households represent 45% of those utilizing foodbanks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those accessing foodbank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
+According to Foodbanks Canada's 2018 data, single-adult households represent 45% of those utilizing foodbanks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those accessing foodbank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
 In addition, 59% of households accessing foodbanks list social or disability assistance as their primary source of income.
 Moreover, children aged 0-18 are over-represented as foodbank users, making up 35% of those accessing foodbank services despite representing 20% of the country's population.
-Likewise, using data from the 2011-2012 CCHS, Tarasuk et al.
+Similarly, using data from the 2011-2012 CCHS, Tarasuk et al.
 [-@tarasuk2019geographic] found higher odds of food insecurity amongst households relying on social assistance, those without a university degree or with children under the age of 18, and individuals that lived alone, renters, and those identifying as Aboriginal.
@@
 
-Currently with the COVID-19 pandemic, disrupted economies, rising unemployment rates, and changes in poverty levels have disrupted the food environment and led to higher rates of food insecurity [@niles2020early].
+Economic stress caused by the COVID-19 pandemic, rising unemployment rates, and changes in poverty levels have disrupted the food environment and led to higher rates of food insecurity [@niles2020early].
 Pandemic-related food security studies in the US have found a substantial increase in households experiencing food insecurity for the first time, and also in households experiencing more severe food insecurity than before [@niles2020early; @wolfson2020food].
@@
 Considering the negative mental and physical health effects associated with food insecurity, increases in food insecurity rates due to the pandemic, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
-Recent data^[https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis_Exec-Sum_EN.pdf] from Foodbanks Canada showed that 52% of member foodbanks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of Canada.
+Recent data^[https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis_Exec-Sum_EN.pdf] from Foodbanks Canada showed that 52% of member foodbanks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of the country.
 The pandemic also created significant staffing issues with 42% of foodbanks reporting a reduction in volunteers.
 However, 53% of foodbanks reported a decrease in use into the summer of 2020 which foodbanks attributed to emergency financial support programs from the federal government were.
-Nevertheless, some of these benefit programs were temporary, suggesting many households may again turn to foodbank services to meet their needs.
+Nevertheless, some of these benefit programs were temporary, which suggests that many households may again turn to foodbank services to meet their needs.
 
@@
 For example, Smith et al.
-[-@smith2017food] found that participants access foodbanks using a variety of modes of transportation with a respondent stating "I wish it [the foodbank] was a little more centrally located.
+[-@smith2017food] note that participants access foodbanks using a variety of modes of transportation with respondent stating "I wish it [the foodbank] was a little more centrally located.
 Because if I didn’t have a bike I’d have to walk it all the way out there and back.
-I wonder about people who don’t." From this, the authors conclude that "transportation can be challenging, particularly if the foodbank is situated in a remote location." To offer greater insight into the role of transportation and location in foodbank accessibility, this research examines how geographic access to foodbanks and food services changed in Hamilton before and during the COVID-19 pandemic.
+I wonder about people who don’t." From this, the authors conclude that "transportation can be challenging, particularly if the foodbank is situated in a remote location." To offer greater insight into the role of transportation and location in foodbank accessibility, this research examines how geographic accessibility to foodbanks and food services changed in Hamilton during the COVID-19 pandemic.
 
@@
 [-@paez2019demand].
-This method for estimating accessibility is a form of the common two-stage floating catchment area method [@radke2000spatial; @luo2003measures].
+This method for estimating accessibility is a form of the widely-used two-stage floating catchment area method [@radke2000spatial; @luo2003measures].
 Floating catchment areas are used to estimate accessibility when there are potential congestion effects, and operate by calculating first the _demand_ for spatially distributed services.
-The demand (e.g., the number of people who require a service) is used to calculate a level of service.
+The demand (usually the number of people who require a service) is used to calculate a level of service.
 In a second step, the level of service is allocated back to the population.
@@
 The population at $P_i$ is allocated to _every_ service point $j$ for which $w_{ij}>0$.
-Similarly, the level of service at $L_j$ is allocated to _every_ population point for which $w_{ji}>0$.
+Similarly, the level of service at $LOS_j$ is allocated to _every_ population point for which $w_{ji}>0$.
 This inflation effect has been known for several years, and several modifications have been proposed to mitigate it [e.g., @wan2012three; @delamater2013spatial].
@@
 
-As a result, the total population (and consequently, the level of service) are preserved system-wide:
+Since the population is allocated proportionally, its value is preserved:
 $$
@@
 
-\noindent and also:
+\noindent as is the level of service:
 $$
@@
 
-Data have been prepared for sharing as a data package^[add repository].
+Data have been prepared for sharing as a data package^[add repository]. The contents of the data package are described next.
 
@@
 
-We converted all recorded residential parcels in the City of Hamilton to points on the road network.
+We converted all recorded residential land parcels in the City of Hamilton to points on the road network.
 Each point includes information about the number of residential units in the parcel.
-Each residential unit is "populated" with the probability of being a "low income household", which we define as having a total household income of less than CAD 40,000.
-This is approximately the mid-point of the low income cut-off (LICOs) for families in Canadian cities with populations greater than 500,000 in 2016, to match other Census data [@statisticscanada2020licos].
-
-### Destinations: Foodbanks
-
-The locations of foodbanks were obtained from public records and geocoded. Three urban sites not in the program were also identified and geocoded for comparison purposes.
+Next, we define low-income households as those having a total income of less than CAD40,000, which is approximately the mid-point of the low income cut-off (LICOs) for families in Canadian cities with populations greater than 500,000 in 2016, to match other Census data [@statisticscanada2020licos].
+We then "populate" each residential unit with the probability of being a low-income household based on the counts of households by income groups in the DA in which the parcel is located.
+While this method assumes a constant probability of low-income household status for all residential units in a DA, the parcel-level analysis affords a high level of spatial disaggregation for the accessibility analysis.
+
+### Destinations: foodbanks and Food Service Locations
+
+The locations of foodbanks and related food services were obtained from the Hamilton Public Library's Food Access Guide^[http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf].
+The guide was updated in April of 2021 to indicate any change affected on the services due to the pandemic.
+This includes modified business hours, a need to make reservations before frequenting, and locations that have completely shut down in consequence.
+While some foodbank services have a specific a target population, such as prioritizing family with young children aged between 0 and 3 or accepting only those providing proof of low-income status through housing and utility statements, all the foodbank services indicated below are designed to accommodate those in need of food at zero to low cost.
+With our focus on foodbanks and food services that offer free or low-cost meals at particular locations, we first removed services such as Meals on Wheels and other food access services such as food box, community kitchens, student nutrition programs, and shopping and transportation.
+In addition, two free meal services held on different days at the same location were collapsed into a single service point for the accessibility analysis.
+
+```{r table-food-bank-info, echo=FALSE, results='asis'}
+foodbank_table %>% kable("latex",
+        booktabs = TRUE,
+        align = ("cccccc"),
+        col.names = c("Type", 
+                      "Description", 
+                      "Locations Pre-COVID", 
+                      "Locations During COVID", 
+                      "Additional Notes"),
+        caption = "\\label{tab:food-bank-info}foodbank and Food Service Information.",
+        escape = FALSE) %>%
+  kable_styling(latex_options = c("striped", "scale_down")) %>%
+  column_spec(2, width = "15em") %>%
+  column_spec(5, width = "15em")
+```
 
@@
 
-Once we obtained travel time tables with population (number of households) and proportion of trips by mode, we calculated the expected travel time $ett$ from each parcel $i$ to a foodbank $j$ as follows:
-$$
-ett_i = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
+Once we obtained travel time tables with population (number of households) and proportion of trips by mode, we calculated the expected travel time $ett$ from each parcel $i$ to a foodbank or food service location $j$ as follows:
+$$
+ett_{ij} = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
 $$
@@
 
-Figure \ref{fig:foodbanks} shows the location of the foodbanks in the City of Hamilton and their status.
-Before the pandemic there were `r nrow(foodbank_locations)` of which `r nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))` closed during the pandemic.
+```{r accessibility-threshold-sensitivity, include=FALSE}
+## Analysis
+# Testing different thresholds to check how accessibility changes. Here, the thresholds are varied from 2 min to 30 min in 2 min step increments
+threshold_seq <- seq(from = 2, to = 30, by = 2)
+
+# Initialize a data frame to store results
+accessibility_test <- data.frame(threshold = threshold_seq, 
+                                 accessibility = numeric(length(threshold_seq)))
+
+# Calculate the system-wide accessibility according to each threshold
+for(i in 1:length(threshold_seq)){
+  results_test <-  ttm %>% 
+  transmute(UID, OBJECTID, travel_time = e_travel_time, population = income_less40k, supply = 1) %>%
+    b2sfca(threshold = i )
+  accessibility_test$accessibility[i] <- sum(results_test$accessibility$accessibility)
+}
+```
+
+```{r results-pre-covid, include=FALSE}
+## Analysis 
+# Calculate k-min accessibility pre-covid:
+results_pre <- ttm %>%
+  #dplyr::filter(COVIDClose == "No") %>%
+  transmute(UID, 
+            OBJECTID, 
+            travel_time = e_travel_time, 
+            population = income_less40k, 
+            supply = 1) %>%
+  b2sfca(threshold = 15)
+```
+
+```{r results-during-covid, include=FALSE}
+## Analysis 
+# Calculate k-min accessibility during covid:
+results_covid <- ttm %>%
+  dplyr::filter(COVIDClose == "No") %>%
+  transmute(UID, 
+            OBJECTID, 
+            travel_time = e_travel_time, 
+            population = income_less40k, 
+            supply = 1) %>%
+  b2sfca(threshold = 15)
+```
+
+```{r, bind-accessibility-results, include=FALSE}
+## Analysis
+# Bind level of service results:
+los <- rbind(data.frame(results_pre$los, 
+                        timing = "Pre-covid", 
+                        threshold = "30 min"),
+             data.frame(results_covid$los, 
+                        timing = "During covid", 
+                        threshold = "30 min")) %>%
+  mutate(timing = factor(timing, 
+                         levels = c("Pre-covid", 
+                                    "During covid"),
+                         ordered = TRUE))
+
+# Bind accessibility results:
+accessibility <- rbind(data.frame(results_pre$accessibility, 
+                                  timing = "Pre-covid", 
+                                  threshold = "30 min"),
+                       data.frame(results_covid$accessibility, 
+                                  timing = "During covid", 
+                                  threshold = "30 min")) %>%
+  mutate(timing = factor(timing, 
+                         levels = c("Pre-covid", 
+                                    "During covid"),
+                         ordered = TRUE))
+```
+
+Figure \ref{fig:foodbanks} shows the location of foodbanks in the City of Hamilton and their status.
+Before the pandemic there were `r nrow(foodbank_locations)` of which `r nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))` (`r round(nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations) * 100, 2)`\%) closed during the pandemic.
 As seen in the figure, foodbanks tend to be predominantly located in the central parts of the city.
 This is not surprising: population density is high there, and it is also the part of the city where lower income households are more numerous in absolute and relative terms (see Figure \ref{fig:low-income-households}).
-Alas, this is also where most of the closures during the pandemic happened.
+Alas, this is also the part of the city where most of the closures during the pandemic happened.
 
@@
 ```
-```{r plot-low-income-households, echo=FALSE, fig.cap="\\label{fig:low-income-households}Number and proportion of households with incomes less than CAD 40,000."}
+
+```{r plot-low-income-households, echo=FALSE, fig.cap="\\label{fig:low-income-households}Number and proportion of households with incomes less than CAD40,000."}
 lih <- ggplot() +
@@
 ```
-```{r accessibility-threshold-sensitivity, include=FALSE}
-## Analysis
-# Testing different thresholds to check how accessibility changes. Here, the thresholds are varied from 2 min to 30 min in 2 min step increments
-threshold_seq <- seq(from = 2, to = 30, by = 2)
-
-# Initialize a data frame to store results
-accessibility_test <- data.frame(threshold = threshold_seq, 
-                                 accessibility = numeric(length(threshold_seq)))
-
-# Calculate the system-wide accessibility according to each threshold
-for(i in 1:length(threshold_seq)){
-  results_test <-  ttm %>% 
-  transmute(UID, OBJECTID, travel_time = e_travel_time, population = income_less40k, supply = 1) %>%
-    b2sfca(threshold = i )
-  accessibility_test$accessibility[i] <- sum(results_test$accessibility$accessibility)
-}
-```
-
-```{r plot-results-sensitivity-analysis, echo=FALSE, fig.cap="\\label{fig:sensitivity-analysis}Accessibility as a function of threshold"}
+
+To implement the accessibility calculations, we must select a distance-decay function.
+In this task we find limited support in the literature, which is mostly silent on the travel patterns of people who visit foodbanks.
+For this reason, we opt for a simple cumulative opportunities function as follows:
+$$
+w_{ij}=w_{ji}=
+\begin{cases}
+1 & \text{ if } ett_{ij}\le \delta\\
+0 & \text{ otherwise}
+\end{cases}
+$$
+\noindent where $ett_{ij}$ is the multimodal expected travel time as described previously, and $\delta$ is a travel threshold.
+When the expected travel time exceeds this threshold, a facility is no longer considered accessible.
+Further, the weights are standardized for the balanced floating catchment area approach.
+
+Figure \ref{fig:sensitivity-analysis} shows the results of conducting sensitivity analysis of the system-wide accessibility as we vary the threshold (considering the situation before the pandemic).
+There is a clear pattern whereby more strict values of $\delta$ are associated with higher levels of system-wide accessibility: this is a result of lower congestion, since fewer households are serviced.
+System-wide accessibility declines with higher values of $\delta$: as more households are serviced, congestion grows and the level of service declines, although this happens at a declining rate.
+We are not aware of any standards that explain how long people are expected to travel for foodbanks, but we note that in developing countries, access to drinking water is defined as a source that takes less than 30 minutes to reach [round trip, see @world2019progress].
+There is no reason why people in affluent countries should be expected to travel more for a basic need such as food.
+Accordingly, we adopt a threshold of 15 minutes for the analysis (representing a one-way trip).
+Using this threshold, we find that the system-wide accessibility (interpreted as a provider-to-population ratio) was `r round(sum(results_pre$accessibility$accessibility), 3)` (foodbanks per low income household in the city) before COVID-19, but declined to `r round(sum(results_covid$accessibility$accessibility), 3)` during the pandemic.
+It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`\% of foodbanks remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`\%.
+
+```{r plot-results-sensitivity-analysis, out.width="60%", echo=FALSE, fig.cap="\\label{fig:sensitivity-analysis}Accessibility as a function of threshold"}
 acc_sensitivity <- ggplot(accessibility_test, 
@@
 acc_sensitivity
-```
-
-```{r 30-min-pre-covid, include=FALSE}
-## Analysis 
-# Calculate 30-min accessibility pre-covid:
-results_30min_pre <- ttm %>%
-  #dplyr::filter(COVIDClose == "No") %>%
-  transmute(UID, 
-            OBJECTID, 
-            travel_time = e_travel_time, 
-            population = income_less40k, 
-            supply = 1) %>%
-  b2sfca(threshold = 30)
-```
-
-```{r 30-min-during-covid, include=FALSE}
-## Analysis 
-# Calculate 30-min accessibility during covid:
-results_30min_covid <- ttm %>%
-  dplyr::filter(COVIDClose == "No") %>%
-  transmute(UID, 
-            OBJECTID, 
-            travel_time = e_travel_time, 
-            population = income_less40k, 
-            supply = 1) %>%
-  b2sfca(threshold = 30)
-```
-
-```{r, bind-accessibility-results, include=FALSE}
-## Analysis
-# Bind level of service results:
-los <- rbind(data.frame(results_30min_pre$los, 
-                        timing = "Pre-covid", 
-                        threshold = "30 min"),
-             data.frame(results_30min_covid$los, 
-                        timing = "During covid", 
-                        threshold = "30 min")) %>%
-  mutate(timing = factor(timing, 
-                         levels = c("Pre-covid", 
-                                    "During covid"),
-                         ordered = TRUE))
-
-# Bind accessibility results:
-accessibility <- rbind(data.frame(results_30min_pre$accessibility, 
-                                  timing = "Pre-covid", 
-                                  threshold = "30 min"),
-                       data.frame(results_30min_covid$accessibility, 
-                                  timing = "During covid", 
-                                  threshold = "30 min")) %>%
-  mutate(timing = factor(timing, 
-                         levels = c("Pre-covid", 
-                                    "During covid"),
-                         ordered = TRUE))
 ```
@@
             by = "TAZUID") 
+```
+
+```{r, join-geometry-to-level-of-service, include=FALSE}
+## Analysis 
+# Join geometry:
+los <- los %>%
+  left_join(foodbank_locations %>%
+              dplyr::select(ID),
+            by = c("OBJECTID" = "ID")) %>%
+  st_as_sf()
+```
+
+The level of service by foodbank before and during the pandemic are shown in Figure \ref{fig:levels-of-service}.
+Since the level of service is the inverse of the number of households in the catchment area of the foodbanks, higher values mean that a facility is expected to service fewer households.
+Conversely, lower values indicate grater congestion.
+The general pattern of the levels of service is similar before and during the pandemic, with lower values in the center of the city.
+Three more peripheral facilities towards the south of the city also have lower levels of service, presumably because they are meant to service a relatively large suburban/exurban population.
+During the pandemic, however, the levels of service dropped, in some cases quite substantially.
+The pattern of the losses in level of service, moreover, is not uniform.
+As seen in Figure \ref{fig:levels-of-service-changes} ...
+words.
+
+```{r plot-levels-of-service, echo=FALSE, fig.cap="\\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
+## Analysis
+# Plot levels of service before and during covid
+
+# Pre-covid
+los_pre <- ggplot() +
+  geom_sf(data = hamilton_cma) +
+  geom_sf(data = los %>%
+            filter(timing == "Pre-covid"),
+          aes(color = los,
+              size = los),
+          alpha = 0.5) +
+  geom_sf(data = foodbank_locations,
+          aes(shape = COVIDClose)) +
+  scale_color_viridis_c(name = "Level of service pre-COVID-19",
+                        option =  "A", 
+                        direction = -1,
+                       breaks = c(0.0005,
+                                  0.0010,
+                                  0.0015,
+                                  0.0020,
+                                  0.0025),
+                       labels = c("0.00050",
+                                  "0.00100",
+                                  "0.00150",
+                                  "0.00200",
+                                  "0.00250")) +
+  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
+  guides(size = FALSE, shape = FALSE) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
+
+# During covid
+los_covid <- ggplot() +
+  geom_sf(data = hamilton_cma) +
+  geom_sf(data = los %>%
+            filter(timing == "During covid"),
+          aes(color = los,
+              size = los),
+          alpha = 0.5) +
+  geom_sf(data = foodbank_locations,
+          aes(shape = COVIDClose)) +
+  scale_color_viridis_c(name = "Level of service during COVID-19",
+                        option =  "A", 
+                        direction = -1,
+                       breaks = c(0.00050,
+                                  0.00075,
+                                  0.00100,
+                                  0.00125,
+                                  0.00150),
+                       labels = c("0.00050",
+                                  "0.00075",
+                                  "0.00100",
+                                  "0.00125",
+                                  "0.00150")) +
+  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
+  guides(size = FALSE, shape = FALSE) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
+
+los_pre / los_covid
+```
+
+```{r plot-levels-of-service-changes, echo=FALSE, fig.cap="\\label{fig:levels-of-service-changes}Changes in levels of service at each facility from pre-COVID-19 to during COVID-19."}
+# This figure shows the places where the level of service deteriorated most (excluding the places that closed, where it vanished).
+los_change <- los %>%
+  st_drop_geometry() %>%
+  pivot_wider(names_from = timing, 
+              values_from = los, 
+              values_fill = 0) %>%
+  filter(`During covid` > 0) %>%
+  mutate(los_d = `During covid` - `Pre-covid`) %>%
+  left_join(foodbank_locations %>%
+            transmute(OBJECTID = ID,
+                      geometry),
+          by = "OBJECTID") %>%
+  st_as_sf()
+
+lc <- ggplot() +
+  geom_sf(data = hamilton_cma) +
+  geom_sf(data = los_change, 
+          aes(color = los_d,
+              size = los_d),
+          alpha = 0.3) +
+  geom_sf(data = foodbank_locations,
+          aes(shape = COVIDClose)) +
+  geom_sf(data = inset_box,
+          linetype = "dotted",
+          color = "red",
+          fill = NA) +
+  scale_color_viridis_c(name = "Change in level of service",
+                        option =  "A", 
+                        direction = 1,
+                       breaks = c(-0.0001,
+                                  -0.0002,
+                                  -0.0003,
+                                  -0.0004,
+                                  -0.0005),
+                       labels = c("-0.0001",
+                                  "-0.0002",
+                                  "-0.0003",
+                                  "-0.0004",
+                                  "-0.0005")) +
+  scale_size(trans = 'reverse') +
+  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
+  guides(size = FALSE, shape = FALSE) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "none")
+
+lc_inset <- ggplot() +
+  geom_sf(data = hamilton_cma) +
+  geom_sf(data = los_change, 
+          aes(color = los_d,
+              size = los_d),
+          alpha = 0.7) +
+  geom_sf(data = foodbank_locations,
+          aes(shape = COVIDClose)) +
+  geom_sf(data = inset_box,
+          linetype = "dotted",
+          color = "red",
+          fill = NA) +
+  scale_color_viridis_c(name = "Change in level of service",
+                        option =  "A", 
+                        direction = 1,
+                       breaks = c(-0.0001,
+                                  -0.0002,
+                                  -0.0003,
+                                  -0.0004,
+                                  -0.0005),
+                       labels = c("-0.0001",
+                                  "-0.0002",
+                                  "-0.0003",
+                                  "-0.0004",
+                                  "-0.0005")) +
+  scale_size(name = "Change in level of service",
+                        trans = 'reverse',
+                       breaks = c(-0.0001,
+                                  -0.0002,
+                                  -0.0003,
+                                  -0.0004,
+                                  -0.0005),
+                       labels = c("-0.0001",
+                                  "-0.0002",
+                                  "-0.0003",
+                                  "-0.0004",
+                                  "-0.0005")) +
+  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
+  xlim(-79.975, -79.775) +
+  ylim(43.2, 43.3) +
+  guides(color = guide_legend(title.position = "top"), 
+         #size = FALSE, 
+         shape = FALSE) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "bottom")
+
+lc / lc_inset
+```
+
+```{r aggregate-accessibility-by-taz, include=FALSE}  
+## Analysis
+# Accessibility by TAZ:
+accessibility_taz <- accessibility %>%
+  group_by(TAZUID, 
+           timing) %>%
+  summarize(accessibility = sum(accessibility),
+            timing = first(timing),
+            .groups = "drop") %>%
+  left_join(population_age_income %>%
+              transmute(TAZUID,
+                        children,
+                        adults,
+                        seniors),
+            by = "TAZUID") %>%
+  st_as_sf()
+```
+
+```{r plot-accessibility, echo=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
+## Analysis
+# Plot accessibility before and during covid:
+acc_pre <- ggplot() +
+  #geom_sf(data = hamilton_cma) +
+  geom_sf(data = accessibility_taz %>%
+            filter(timing == "Pre-covid"),
+          aes(fill = accessibility)) +
+  scale_fill_viridis_c(name = "Accessibility pre-COVID-19",
+                        option =  "A", 
+                        direction = -1,
+                       breaks = c(0.001,
+                                  0.0008,
+                                  0.0006,
+                                  0.0004,
+                                  0.0002),
+                       labels = c("0.0010",
+                                  "0.0008",
+                                  "0.0006",
+                                  "0.0004",
+                                  "0.0002")) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
+
+acc_covid <- ggplot() +
+  #geom_sf(data = hamilton_cma) +
+  geom_sf(data = accessibility_taz %>%
+            filter(timing == "During covid"),
+          aes(fill = accessibility)) +
+  scale_fill_viridis_c(name = "Accessibility during COVID-19",
+                        option =  "A", 
+                        direction = -1,
+                       breaks = c(0.0005,
+                                  0.0004,
+                                  0.0003,
+                                  0.0002,
+                                  0.0001),
+                       labels = c("0.0005",
+                                  "0.0004",
+                                  "0.0003",
+                                  "0.0002",
+                                  "0.0001")) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
+
+acc_pre / acc_covid
+```
+
+```{r plot-accessibility-changes, echo=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19."}
+## Analysis
+# Calculate accessibility change by TAZ:
+accessibilty_taz_change <- accessibility_taz %>%
+  st_drop_geometry() %>%
+  pivot_wider(names_from = timing, 
+              values_from = accessibility) %>%
+  # Calculate changes in accessibility
+  mutate(acc_d = `During covid` - `Pre-covid`) %>%
+  # Join TAZ geometry
+  left_join(population_age_income %>%
+            transmute(TAZUID,
+                      geometry),
+          by = "TAZUID") %>%
+  # Convert to simple features
+  st_as_sf()
+
+# Plot
+acc_change <- ggplot() +
+  #geom_sf(data = hamilton_cma) +
+  geom_sf(data = accessibilty_taz_change,
+          aes(fill = acc_d),
+          color = NA) +
+  geom_sf(data = population_age_income,
+          fill = NA) +
+  scale_fill_viridis_c(name = "Change in accessibility",
+                        option =  "A", 
+                        direction = 1,
+                       breaks = c(-0.0005,
+                                  -0.0004,
+                                  -0.0003,
+                                  -0.0002,
+                                  -0.0001),
+                       labels = c("-0.0005",
+                                  "-0.0004",
+                                  "-0.0003",
+                                  "-0.0002",
+                                  "-0.0001")) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
+
+acc_change
 ```
@@
                       "Difference"),
-        caption = "\\label{tab:accessibility-by-age}Accessibility by age group among members of low income households.",
+        caption = "\\label{tab:accessibility-by-age}Accessibility by age group among members households with incomes less than CAD40,000.",
         escape = FALSE) %>%
@@
 
-```{r, join-geometry-to-level-of-service, include=FALSE}
-## Analysis 
-# Join geometry:
-los <- los %>%
-  left_join(foodbank_locations %>%
-              dplyr::select(ID),
-            by = c("OBJECTID" = "ID")) %>%
-  st_as_sf()
-```
-
-```{r plot-levels-of-service, echo=FALSE, fig.cap="\\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
-## Analysis
-# Plot levels of service before and during covid
-
-# Pre-covid
-los_pre <- ggplot() +
-  geom_sf(data = hamilton_cma) +
-  geom_sf(data = los %>%
-            filter(timing == "Pre-covid"),
-          aes(color = los,
-              size = los)) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  scale_color_viridis_c(name = "Level of service pre-COVID-19",
-                        option =  "A", 
-                        direction = -1) +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  guides(size = FALSE, shape = FALSE) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-# During covid
-los_covid <- ggplot() +
-  geom_sf(data = hamilton_cma) +
-  geom_sf(data = los %>%
-            filter(timing == "During covid"),
-          aes(color = los,
-              size = los)) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  scale_color_viridis_c(name = "Level of service during COVID-19",
-                        option =  "A", 
-                        direction = -1) +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  guides(size = FALSE, shape = FALSE) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-los_pre / los_covid
-```
-
-```{r plot-levels-of-service-changes, echo=FALSE, fig.cap="\\label{fig:levels-of-service-changes}Changes in levels of service at each facility from pre-COVID-19 to during COVID-19."}
-# This figure shows the places where the level of service deteriorated most (excluding the places that closed, where it vanished).
-los_change <- los %>%
-  st_drop_geometry() %>%
-  pivot_wider(names_from = timing, 
-              values_from = los, 
-              values_fill = 0) %>%
-  filter(`During covid` > 0) %>%
-  mutate(los_d = `During covid` - `Pre-covid`) %>%
-  left_join(foodbank_locations %>%
-            transmute(OBJECTID = ID,
-                      geometry),
-          by = "OBJECTID") %>%
-  st_as_sf() %>%
-  ggplot() +
-  geom_sf(data = hamilton_cma) +
-  geom_sf(aes(color = los_d,
-              size = los_d)) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  scale_color_viridis_c(name = "Change in level of service",
-                        option =  "A", 
-                        direction = -1) +
-  scale_size(trans = 'reverse') +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  guides(size = FALSE, shape = FALSE) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-los_change
-```
-
-```{r aggregate-accessibility-by-taz, include=FALSE}  
-## Analysis
-# Accessibility by TAZ:
-accessibility_taz <- accessibility %>%
-  group_by(TAZUID, 
-           timing) %>%
-  summarize(accessibility = sum(accessibility),
-            timing = first(timing),
-            .groups = "drop") %>%
-  left_join(population_age_income %>%
-              transmute(TAZUID,
-                        children,
-                        adults,
-                        seniors),
-            by = "TAZUID") %>%
-  st_as_sf()
-```
-
-```{r plot-accessibility, echo=FALSE, fig.align = 'center', out.width = "90%", fig.cap="\\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
-## Analysis
-# Plot accessibility before and during covid:
-acc_pre <- ggplot() +
-  #geom_sf(data = hamilton_cma) +
-  geom_sf(data = accessibility_taz %>%
-            filter(timing == "Pre-covid"),
-          aes(fill = accessibility)) +
-  scale_fill_viridis_c(name = "Accessibility pre-COVID-19",
-                        option =  "A", 
-                        direction = -1,
-                       breaks = c(0.001,
-                                  0.0008,
-                                  0.0006,
-                                  0.0004,
-                                  0.0002),
-                       labels = c("0.0010",
-                                  "0.0008",
-                                  "0.0006",
-                                  "0.0004",
-                                  "0.0002")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-acc_covid <- ggplot() +
-  #geom_sf(data = hamilton_cma) +
-  geom_sf(data = accessibility_taz %>%
-            filter(timing == "During covid"),
-          aes(fill = accessibility)) +
-  scale_fill_viridis_c(name = "Accessibility during COVID-19",
-                        option =  "A", 
-                        direction = -1,
-                       breaks = c(0.0005,
-                                  0.0004,
-                                  0.0003,
-                                  0.0002,
-                                  0.0001),
-                       labels = c("0.0005",
-                                  "0.0004",
-                                  "0.0003",
-                                  "0.0002",
-                                  "0.0001")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-acc_pre / acc_covid
-```
-
-```{r plot-accessibility-changes, echo=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19."}
-## Analysis
-# Calculate accessibility change by TAZ:
-accessibilty_taz_change <- accessibility_taz %>%
-  st_drop_geometry() %>%
-  pivot_wider(names_from = timing, 
-              values_from = accessibility) %>%
-  # Calculate changes in accessibility
-  mutate(acc_d = `During covid` - `Pre-covid`) %>%
-  # Join TAZ geometry
-  left_join(population_age_income %>%
-            transmute(TAZUID,
-                      geometry),
-          by = "TAZUID") %>%
-  # Convert to simple features
-  st_as_sf()
-
-# Plot
-acc_change <- ggplot() +
-  #geom_sf(data = hamilton_cma) +
-  geom_sf(data = accessibilty_taz_change,
-          aes(fill = acc_d),
-          color = NA) +
-  geom_sf(data = population_age_income,
-          fill = NA) +
-  scale_fill_viridis_c(name = "Change in accessibility",
-                        option =  "A", 
-                        direction = 1,
-                       breaks = c(-0.0005,
-                                  -0.0004,
-                                  -0.0003,
-                                  -0.0002,
-                                  -0.0001),
-                       labels = c("-0.0005",
-                                  "-0.0004",
-                                  "-0.0003",
-                                  "-0.0002",
-                                  "-0.0001")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-acc_change
-```
-
 Conclusions

### 2021-04-23 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Continued editing the results and discussion

--- 2021-04-22 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
@@
 sufficient quantity and/or adequate quality of food'' due to a
-household's financial limitations (Enns et al., 2020). This is a
-condition that has been associated with reductions in nutritional
-outcomes (Bhattacharya et al., 2004; Kirkpatrick and Tarasuk, 2008;
-Olson, 1999) and physical and mental health in children and adults
-(Elgar et al., 2021; Jones, 2017; Ramsey et al., 2011; Seligman et al.,
-2010; Stuff et al., 2004). Over at least the past four decades foodbanks
-have become an essential line of defense against food insecurity in
-Canadian communities (Black and Seto, 2020; Holmes et al., 2018; Riches,
-2002; Tarasuk et al., 2020). In this respect, Canada is not unlike
-numerous other wealthy countries where a systematic dismantling of the
-welfare state took place in the intervening period (Tarasuk et al.,
-2014).
+household's financial limitations (Enns et al., 2020). This condition
+has been associated with reductions in nutritional outcomes
+(Bhattacharya et al., 2004; Kirkpatrick and Tarasuk, 2008; Olson, 1999)
+and physical and mental health in children and adults (Elgar et al.,
+2021; Jones, 2017; Ramsey et al., 2011; Seligman et al., 2010; Stuff et
+al., 2004). Over at least the past four decades foodbanks have become an
+essential line of defense against food insecurity in Canadian
+communities (Black and Seto, 2020; Holmes et al., 2018; Riches, 2002;
+Tarasuk et al., 2020). In this respect, Canada is not unlike numerous
+other wealthy countries where a systematic dismantling of the welfare
+state took place in the intervening period (Tarasuk et al., 2014).
 
-But the emergence of COVID-19, the worst public health crisis since the
-1918 flu pandemic, has revealed important social and economic fault
-lines, and pre-existing patterns of inequality appear to have been
-exacerbated. Along several other dimensions of stress (such as
-accessibility to health care facilities, see Pereira et al., 2021a),
-this seems to be the case for food insecurity as well (Laborde et al.,
-2020). In the US, for example, it has been estimated that there was an
-increase of more than 30\% in household food insecurity, and more than
-one third of households were discovered to be newly food insecure -
-meaning they did not experience food insecurity before the pandemic
-(Niles et al., 2020). In Canada, Men and Tarasuk (2021) report that
-about 25\% of individuals who experienced job insecurity (a relatively
-common occurrence during the pandemic), also experienced food
-insecurity. Similarly, according to Statistics Canada (2020a), in the
-early stages of the pandemic almost 15\% of individuals reported living
-in a household that faced food insecurity; the risk of food insecurity
-was substantially higher for households with children. The difference
-between households with and without children was significant, and 11.7\%
-of households with children indicated that ``food didn't last and
-{[}there was{]} no money to get more'' sometimes or often, compared to
-7.3\% of households without children; likewise, 13\% of households with
-children indicated that they ``{[}c{]}ouldn't afford balanced meals''
-sometimes or often, compared to 8.8\% of households without children.
+The emergence of COVID-19, the worst public health crisis since the 1918
+flu pandemic, has revealed important social and economic fault lines,
+and pre-existing patterns of inequality appear to have been exacerbated.
+Along several other dimensions of stress (e.g., accessibility to health
+care facilities, Pereira et al., 2021a), this seems to be the case for
+food insecurity as well (Laborde et al., 2020). In the US, for example,
+it has been estimated that there was an increase of more than 30\% in
+household food insecurity, and more than one third of households were
+discovered to be newly food insecure - meaning they did not experience
+food insecurity before the pandemic (Niles et al., 2020). In Canada, Men
+and Tarasuk (2021) report that about 25\% of individuals who experienced
+job insecurity (a relatively common occurrence during the pandemic),
+also experienced food insecurity. Similarly, according to Statistics
+Canada (2020a), in the early stages of the pandemic almost 15\% of
+individuals reported living in a household that faced food insecurity;
+the risk of food insecurity was substantially higher for households with
+children. The difference between households with and without children
+was significant, and 11.7\% of households with children indicated that
+``food didn't last and {[}there was{]} no money to get more'' sometimes
+or often, compared to 7.3\% of households without children; likewise,
+13\% of households with children indicated that they ``{[}c{]}ouldn't
+afford balanced meals'' sometimes or often, compared to 8.8\% of
+households without children.
 
@@
 foodbanks (minors up to 18 years old) was 9,125 in March 2019, up from
-8,278 the year before.
+8,278 the year before, a rate of increase greater than population
+growth.
 
-It is known that the urban food environment, within which people make
-their daily food choices, is essential in influencing eating behaviours
-and health outcomes, based on factors such as food availability, ease of
+It is known that urban food environments, within which people make their
+daily food choices, are essential in influencing eating behaviours and
+health outcomes, based on factors such as food availability, ease of
 geographic accessibility and socio-demographic variations (Paez et al.,
@@
 and previous survey research has suggested that such accessibility plays
-a key role in user satisfaction with foodbank service delivery in
-Vancouver (Holmes et al., 2018). However, as with research into the
-prevalence of food deserts, accessibility to foodbanks is unlikely to be
-evenly distributed throughout a city due to variations in the
-characteristics of the transportation network characteristics, and the
-spatial distribution of foodbank locations and the population they
-serve. Furthermore, policy responses to the COVID-19 pandemic may well
-add to the distress of vulnerable households. Non-pharmaceutical
+a key role in user satisfaction with foodbank service delivery (Holmes
+et al., 2018). However, as with research into the prevalence of food
+deserts, accessibility to foodbanks is unlikely to be evenly
+distributed, and variation throughout a city can be expected due to
+transportation network characteristics, and the spatial distribution of
+foodbank locations and the population they are meant to serve.
+Furthermore, policy responses to the COVID-19 pandemic likely have added
+to the distress of vulnerable households. Non-pharmaceutical
 interventions during the pandemic involving restrictions in mobility
 have increased the friction of travel, in particular by transit on which
-low income populations are more likely to be reliant (e.g., DeWeese et
-al., 2020). At the same time, the pandemic has created additional stress
-for the operators of foodbanks through disruptions in the supply chain
-(e.g., McKay et al., 2021) as well as concerns surrounding the delivery
-of service in safe conditions and possible cancellation of food service
+low income populations are more reliant (e.g., DeWeese et al., 2020). At
+the same time, the pandemic has created additional stress for the
+operators of foodbanks through disruptions in the supply chain (e.g.,
+McKay et al., 2021) as well as concerns surrounding the delivery of
+service in safe conditions and possible cancellation of food service
 programs.
@@
 related services (e.g.~low-cost or free meal service providers)
-available in Hamilton, Ontario, changed during the pandemic. Have the
-number of open foodbank services diminished? If so, what were the levels
-of accessibility to foodbanks before the pandemic from the perspective
-of low income households, and how did they change during the pandemic?
-And finally, who are most likely to have been impacted by changes to the
-accessibility landscape? This paper will first look at the distribution
-of foodbanks and related services before and during the pandemic. Then,
-we use the balanced floating catchment area approach of Paez et al.
-(2019) to investigate the accessibility situation. We use a fully
+available in Hamilton, Ontario, changed during the pandemic. Did the
+number of open foodbank services diminish? If so, what was the
+accessibility to foodbanks before the pandemic from the perspective of
+low income households, and how it changed during the pandemic? And
+finally, who are most likely to have been impacted by changes in the
+accessibility landscape? This paper first looks at the distribution of
+foodbanks and related services before and during the pandemic. Then, we
+use the balanced floating catchment area approach of Paez et al. (2019)
+to investigate the accessibility situation. For this, we use a fully
 disaggregated approach based on parcel-level data. Socio-economic and
@@
 Food insecurity is the inability to acquire and consume an adequate
-amount or good quality food, leading to insufficient nutrient intake
-(Enns et al., 2020; Kirkpatrick and Tarasuk, 2008; Tarasuk and Vogt,
-2009). Nutrient deficiency, in turn, is associated with major population
-health concerns, and particularly among those who are at a
-socio-economic disadvantage (Bazerghi et al., 2016). Data on food
-insecurity in Canada is collected in several ways. Quantitatively,
-official government surveys such as the Household Food Security Survey
-Module (HFSSM), the Canadian Community Health Surveys (CCHS), the
-International Study of Adults (LISA), and official classifications
-determined by Health Canada in relation to socio-demographic variables
-offer some insight into food insecurity (Gundersen et al., 2018;
-Kirkpatrick and Tarasuk, 2008; Tarasuk and Vogt, 2009). Nationally, it
-has been found that food insecurity impacts approximately 12.3\% of
-Canadian households (Tarasuk et al., 2014)
+amount or good quality food, leading to inadequate nutrient intake (Enns
+et al., 2020; Kirkpatrick and Tarasuk, 2008; Tarasuk and Vogt, 2009).
+This nutrient deficiency is associated with major population health
+concerns, particularly among Canadians at socio-economic disadvantage
+(Bazerghi et al., 2016). Data on food insecurity in Canada is collected
+in several ways. Quantitatively, official government surveys such as the
+Household Food Security Survey Module (HFSSM), the Canadian Community
+Health Surveys (CCHS), the International Study of Adults (LISA), and
+official classifications determined by Health Canada in relation to
+socio-demographic variables offer some insight into food insecurity
+(Gundersen et al., 2018; Kirkpatrick and Tarasuk, 2008; Tarasuk and
+Vogt, 2009). Nationally, it has been found that food insecurity impacts
+approximately 12.3\% of Canadian households (Tarasuk et al., 2014)
 
@@
 retrenchments and economic downfalls since the 1980s, over time they
-evolved into a community practice to secure emergency food supplies for
-those in need (Loopstra and Tarasuk, 2012; Wakefield et al., 2013).
+have evolved into a community practice to secure emergency food supplies
+for those in need (Loopstra and Tarasuk, 2012; Wakefield et al., 2013).
 
@@
 few decades (Wakefield et al., 2013). The largest database of foodbanks
-and their use comes from the non-profit association foodbanks Canada,
+and their use comes from the non-profit association Foodbanks Canada,
 which conducts an annual assessment through its affiliated members.
-foodbanks Canada's 2018 Hunger Count report\footnote{https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018\_FINAL\_EN.pdf.aspx?ext=.pdf}
+Foodbanks Canada's 2018 Hunger Count report\footnote{https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018\_FINAL\_EN.pdf.aspx?ext=.pdf}
 (the most recent available) listed 1,830 member foodbanks across the
-country found that Canadians visited foodbanks 1.1 million times in
+country, and found that Canadians visited foodbanks 1.1 million times in
 March of 2018. Of those accessing foodbanks, certain population
 characteristics tend to be over-represented compared to national totals
-from the 2016 Canadian Census of Population. According to foodbanks
+from the 2016 Canadian Census of Population. According to Foodbanks
 Canada's 2018 data, single-adult households represent 45\% of those
@@
 35\% of those accessing foodbank services despite representing 20\% of
-the country's population. Likewise, using data from the 2011-2012 CCHS,
+the country's population. Similarly, using data from the 2011-2012 CCHS,
 Tarasuk et al. (2019) found higher odds of food insecurity amongst
@@
 
-Currently with the COVID-19 pandemic, disrupted economies, rising
-unemployment rates, and changes in poverty levels have disrupted the
-food environment and led to higher rates of food insecurity (Niles et
-al., 2020). Pandemic-related food security studies in the US have found
-a substantial increase in households experiencing food insecurity for
-the first time, and also in households experiencing more severe food
+Economic stress caused by the COVID-19 pandemic, rising unemployment
+rates, and changes in poverty levels have disrupted the food environment
+and led to higher rates of food insecurity (Niles et al., 2020).
+Pandemic-related food security studies in the US have found a
+substantial increase in households experiencing food insecurity for the
+first time, and also in households experiencing more severe food
 insecurity than before (Niles et al., 2020; Wolfson and Leung, 2020).
@@
 increase in usage in March of 2020 when initial lockdown restrictions
-were put in place across much of Canada. The pandemic also created
+were put in place across much of the country. The pandemic also created
 significant staffing issues with 42\% of foodbanks reporting a reduction
@@
 financial support programs from the federal government were.
-Nevertheless, some of these benefit programs were temporary, suggesting
-many households may again turn to foodbank services to meet their needs.
+Nevertheless, some of these benefit programs were temporary, which
+suggests that many households may again turn to foodbank services to
+meet their needs.
 
@@
 transportation for trips to the foodbank (Smith-Carrier et al., 2017).
-For example, Smith et al. (2017) found that participants access
-foodbanks using a variety of modes of transportation with a respondent
-stating ``I wish it {[}the foodbank{]} was a little more centrally
-located. Because if I didn't have a bike I'd have to walk it all the way
-out there and back. I wonder about people who don't.'' From this, the
-authors conclude that ``transportation can be challenging, particularly
-if the foodbank is situated in a remote location.'' To offer greater
-insight into the role of transportation and location in foodbank
-accessibility, this research examines how geographic access to foodbanks
-and food services changed in Hamilton before and during the COVID-19
-pandemic.
+For example, Smith et al. (2017) note that participants access foodbanks
+using a variety of modes of transportation with respondent stating ``I
+wish it {[}the foodbank{]} was a little more centrally located. Because
+if I didn't have a bike I'd have to walk it all the way out there and
+back. I wonder about people who don't.'' From this, the authors conclude
+that ``transportation can be challenging, particularly if the foodbank
+is situated in a remote location.'' To offer greater insight into the
+role of transportation and location in foodbank accessibility, this
+research examines how geographic accessibility to foodbanks and food
+services changed in Hamilton during the COVID-19 pandemic.
 
@@
 area approach of Paez et al. (2019). This method for estimating
-accessibility is a form of the common two-stage floating catchment area
-method (Luo and Wang, 2003; Radke and Mu, 2000). Floating catchment
+accessibility is a form of the widely-used two-stage floating catchment
+area method (Luo and Wang, 2003; Radke and Mu, 2000). Floating catchment
 areas are used to estimate accessibility when there are potential
 congestion effects, and operate by calculating first the \emph{demand}
-for spatially distributed services. The demand (e.g., the number of
+for spatially distributed services. The demand (usually the number of
 people who require a service) is used to calculate a level of service.
@@
 population at \(P_i\) is allocated to \emph{every} service point \(j\)
-for which \(w_{ij}>0\). Similarly, the level of service at \(L_j\) is
+for which \(w_{ij}>0\). Similarly, the level of service at \(LOS_j\) is
 allocated to \emph{every} population point for which \(w_{ji}>0\). This
@@
 
-As a result, the total population (and consequently, the level of
-service) are preserved system-wide: \[
+Since the population is allocated proportionally, its value is
+preserved: \[
 \sum_{i=1}^nP_iw_{ij}^\text{st}=P_i
@@
 
-\noindent and also: \[
+\noindent as is the level of service: \[
 \sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
@@
 Data have been prepared for sharing as a data package\footnote{add
-  repository}.
+  repository}. The contents of the data package are described next.
 
@@
 
-We converted all recorded residential parcels in the City of Hamilton to
-points on the road network. Each point includes information about the
-number of residential units in the parcel. Each residential unit is
-``populated'' with the probability of being a ``low income household,''
-which we define as having a total household income of less than CAD
-40,000. This is approximately the mid-point of the low income cut-off
-(LICOs) for families in Canadian cities with populations greater than
-500,000 in 2016, to match other Census data (Statistics Canada, 2020b).
+We converted all recorded residential land parcels in the City of
+Hamilton to points on the road network. Each point includes information
+about the number of residential units in the parcel. Next, we define
+low-income households as those having a total income of less than
+CAD40,000, which is approximately the mid-point of the low income
+cut-off (LICOs) for families in Canadian cities with populations greater
+than 500,000 in 2016, to match other Census data (Statistics Canada,
+2020b). We then ``populate'' each residential unit with the probability
+of being a low-income household based on the counts of households by
+income groups in the DA in which the parcel is located. While this
+method assumes a constant probability of low-income household status for
+all residential units in a DA, the parcel-level analysis affords a high
+level of spatial disaggregation for the accessibility analysis.
 
-\hypertarget{destinations-foodbanks}{%
-\subsubsection{Destinations: Foodbanks}\label{destinations-foodbanks}}
+\hypertarget{destinations-foodbanks-and-food-service-locations}{%
+\subsubsection{Destinations: foodbanks and Food Service
+Locations}\label{destinations-foodbanks-and-food-service-locations}}
 
-The locations of foodbanks were obtained from public records and
-geocoded. Three urban sites not in the program were also identified and
-geocoded for comparison purposes.
+The locations of foodbanks and related food services were obtained from
+the Hamilton Public Library's Food Access Guide\footnote{http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf}.
+The guide was updated in April of 2021 to indicate any change affected
+on the services due to the pandemic. This includes modified business
+hours, a need to make reservations before frequenting, and locations
+that have completely shut down in consequence. While some foodbank
+services have a specific a target population, such as prioritizing
+family with young children aged between 0 and 3 or accepting only those
+providing proof of low-income status through housing and utility
+statements, all the foodbank services indicated below are designed to
+accommodate those in need of food at zero to low cost. With our focus on
+foodbanks and food services that offer free or low-cost meals at
+particular locations, we first removed services such as Meals on Wheels
+and other food access services such as food box, community kitchens,
+student nutrition programs, and shopping and transportation. In
+addition, two free meal services held on different days at the same
+location were collapsed into a single service point for the
+accessibility analysis.
+
+\begin{table}
+
+\caption{\label{tab:table-food-bank-info}\label{tab:food-bank-info}foodbank and Food Service Information.}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{c>{\centering\arraybackslash}p{15em}cc>{\centering\arraybackslash}p{15em}}
+\toprule
+Type & Description & Locations Pre-COVID & Locations During COVID & Additional Notes\\
+\midrule
+\cellcolor{gray!6}{Congregate Dining} & \cellcolor{gray!6}{Congregate and dining programs provide low-cost meals that are enjoyed in a community setting.
+Transportation may be provided} & \cellcolor{gray!6}{7} & \cellcolor{gray!6}{2} & \cellcolor{gray!6}{One remaining location reduced hours during COVID}\\
+Community Meals & Programs often run by volunteers that organize suppers, lunches or other get-togethers that give community residents an opportunity to meet one another in a friendly and informal atmosphere while sharing a meal & 11 & 9 & NA\\
+\cellcolor{gray!6}{Food Banks} & \cellcolor{gray!6}{Food Banks and Emergency Food programs provide individuals and families with grocery items free of charge} & \cellcolor{gray!6}{27} & \cellcolor{gray!6}{26} & \cellcolor{gray!6}{One remaining location reduced hours during COVID while 4 others moved to appointment only}\\
+Free Meals & Meals are provided free of charge in the community through volunteer labour and donations & 9 & 5 & One remaining location reduced hours during COVID\\
+\cellcolor{gray!6}{Low-Cost Meals} & \cellcolor{gray!6}{Restaurants, cafeterias and other eating establishments operated by hospitals, senior centers or other organizations which provide reduced-cost meals for low-income people, older adults or other targeted individuals.} & \cellcolor{gray!6}{2} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{The remaining location reduced hours during COVID}\\
+\bottomrule
+\end{tabular}}
+\end{table}
 
@@
 households) and proportion of trips by mode, we calculated the expected
-travel time \(ett\) from each parcel \(i\) to a foodbank \(j\) as
-follows: \[
-ett_i = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
+travel time \(ett\) from each parcel \(i\) to a foodbank or food service
+location \(j\) as follows: \[
+ett_{ij} = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
 \]
@@
 
-Figure \ref{fig:foodbanks} shows the location of the foodbanks in the
-City of Hamilton and their status. Before the pandemic there were 58 of
-which 14 closed during the pandemic. As seen in the figure, foodbanks
-tend to be predominantly located in the central parts of the city. This
-is not surprising: population density is high there, and it is also the
-part of the city where lower income households are more numerous in
-absolute and relative terms (see Figure
-\ref{fig:low-income-households}). Alas, this is also where most of the
-closures during the pandemic happened.
+Figure \ref{fig:foodbanks} shows the location of foodbanks in the City
+of Hamilton and their status. Before the pandemic there were 58 of which
+14 (24.14\%) closed during the pandemic. As seen in the figure,
+foodbanks tend to be predominantly located in the central parts of the
+city. This is not surprising: population density is high there, and it
+is also the part of the city where lower income households are more
+numerous in absolute and relative terms (see Figure
+\ref{fig:low-income-households}). Alas, this is also the part of the
+city where most of the closures during the pandemic happened.
 
@@
 \end{figure}
+
 \begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-low-income-households-1} \caption{\label{fig:low-income-households}Number and proportion of households with incomes less than CAD 40,000.}\label{fig:plot-low-income-households}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-low-income-households-1} \caption{\label{fig:low-income-households}Number and proportion of households with incomes less than CAD40,000.}\label{fig:plot-low-income-households}
 \end{figure}
 
+To implement the accessibility calculations, we must select a
+distance-decay function. In this task we find limited support in the
+literature, which is mostly silent on the travel patterns of people who
+visit foodbanks. For this reason, we opt for a simple cumulative
+opportunities function as follows: \[
+w_{ij}=w_{ji}=
+\begin{cases}
+1 & \text{ if } ett_{ij}\le \delta\\
+0 & \text{ otherwise}
+\end{cases}
+\] \noindent where \(ett_{ij}\) is the multimodal expected travel time
+as described previously, and \(\delta\) is a travel threshold. When the
+expected travel time exceeds this threshold, a facility is no longer
+considered accessible. Further, the weights are standardized for the
+balanced floating catchment area approach.
+
+Figure \ref{fig:sensitivity-analysis} shows the results of conducting
+sensitivity analysis of the system-wide accessibility as we vary the
+threshold (considering the situation before the pandemic). There is a
+clear pattern whereby more strict values of \(\delta\) are associated
+with higher levels of system-wide accessibility: this is a result of
+lower congestion, since fewer households are serviced. System-wide
+accessibility declines with higher values of \(\delta\): as more
+households are serviced, congestion grows and the level of service
+declines, although this happens at a declining rate. We are not aware of
+any standards that explain how long people are expected to travel for
+foodbanks, but we note that in developing countries, access to drinking
+water is defined as a source that takes less than 30 minutes to reach
+(round trip, see UNICEF-WHO, 2019). There is no reason why people in
+affluent countries should be expected to travel more for a basic need
+such as food. Accordingly, we adopt a threshold of 15 minutes for the
+analysis (representing a one-way trip). Using this threshold, we find
+that the system-wide accessibility (interpreted as a
+provider-to-population ratio) was 0.065 (foodbanks per low income
+household in the city) before COVID-19, but declined to 0.037 during the
+pandemic. It is striking that although almost 76\% of foodbanks remained
+in operation during the pandemic, there was a loss of accessibility
+greater than 43\%.
+
 \begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-results-sensitivity-analysis-1} \caption{\label{fig:sensitivity-analysis}Accessibility as a function of threshold}\label{fig:plot-results-sensitivity-analysis}
+\includegraphics[width=0.6\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-results-sensitivity-analysis-1} \caption{\label{fig:sensitivity-analysis}Accessibility as a function of threshold}\label{fig:plot-results-sensitivity-analysis}
 \end{figure}
 
-\begin{table}
-
-\caption{\label{tab:table-accessibility-by-age-group}\label{tab:accessibility-by-age}Accessibility by age group among members of low income households.}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{cccccc}
-\toprule
-\multicolumn{3}{c}{Population} & \multicolumn{3}{c}{Accessibility} \\
-\cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){4-6}
-Children (age $\le$ 18) & Adults (19-64) & Seniors (age $\ge$ 65) & Before COVID-19 & During COVID-19 & Difference\\
-\midrule
-\cellcolor{gray!6}{7,015} & \cellcolor{gray!6}{21,586} & \cellcolor{gray!6}{9,158} & \cellcolor{gray!6}{0.00734} & \cellcolor{gray!6}{0.00438} & \cellcolor{gray!6}{-0.00296}\\
-3,490 & 13,598 & 6,618 & 0.01386 & 0.00812 & -0.00575\\
-\cellcolor{gray!6}{966} & \cellcolor{gray!6}{5,180} & \cellcolor{gray!6}{4,310} & \cellcolor{gray!6}{0.01475} & \cellcolor{gray!6}{0.00849} & \cellcolor{gray!6}{-0.00625}\\
-2,384 & 7,725 & 5,595 & 0.01853 & 0.01079 & -0.00775\\
-\bottomrule
-\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
-\multicolumn{6}{l}{\rule{0pt}{1em}Population values have been rounded}\\
-\end{tabular}}
-\end{table}
+The level of service by foodbank before and during the pandemic are
+shown in Figure \ref{fig:levels-of-service}. Since the level of service
+is the inverse of the number of households in the catchment area of the
+foodbanks, higher values mean that a facility is expected to service
+fewer households. Conversely, lower values indicate grater congestion.
+The general pattern of the levels of service is similar before and
+during the pandemic, with lower values in the center of the city. Three
+more peripheral facilities towards the south of the city also have lower
+levels of service, presumably because they are meant to service a
+relatively large suburban/exurban population. During the pandemic,
+however, the levels of service dropped, in some cases quite
+substantially. The pattern of the losses in level of service, moreover,
+is not uniform. As seen in Figure \ref{fig:levels-of-service-changes}
+\ldots{} words.
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-accessibility-1}
+{\centering \includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-accessibility-1}
 
@@
 \end{figure}
+
+\begin{table}
+
+\caption{\label{tab:table-accessibility-by-age-group}\label{tab:accessibility-by-age}Accessibility by age group among members households with incomes less than CAD40,000.}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{cccccc}
+\toprule
+\multicolumn{3}{c}{Population} & \multicolumn{3}{c}{Accessibility} \\
+\cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){4-6}
+Children (age $\le$ 18) & Adults (19-64) & Seniors (age $\ge$ 65) & Before COVID-19 & During COVID-19 & Difference\\
+\midrule
+\cellcolor{gray!6}{3,445} & \cellcolor{gray!6}{11,860} & \cellcolor{gray!6}{6,584} & \cellcolor{gray!6}{0.00229} & \cellcolor{gray!6}{0.00173} & \cellcolor{gray!6}{-0.00057}\\
+4,477 & 12,838 & 6,847 & 0.00707 & 0.00418 & -0.00289\\
+\cellcolor{gray!6}{3,541} & \cellcolor{gray!6}{13,189} & \cellcolor{gray!6}{7,020} & \cellcolor{gray!6}{0.01580} & \cellcolor{gray!6}{0.00911} & \cellcolor{gray!6}{-0.00669}\\
+2,340 & 9,918 & 4,940 & 0.03985 & 0.02202 & -0.01783\\
+\bottomrule
+\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
+\multicolumn{6}{l}{\rule{0pt}{1em}Population values have been rounded}\\
+\end{tabular}}
+\end{table}
 
@@
 
+\leavevmode\hypertarget{ref-world2019progress}{}%
+UNICEF-WHO, 2019. Progress on household drinking water, sanitation and
+hygiene 2000--2017. Special focus on inequalities.
+
 \leavevmode\hypertarget{ref-vanderlee2017food}{}%

### 2021-04-23 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Finishing touches to the figures. Completed first draft of discussion of level of service.

@@
 Before the pandemic there were `r nrow(foodbank_locations)` of which `r nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))` (`r round(nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations) * 100, 2)`\%) closed during the pandemic.
-As seen in the figure, foodbanks tend to be predominantly located in the central parts of the city.
+As shown in the figure, foodbanks tend to be predominantly located in the central parts of the city.
 This is not surprising: population density is high there, and it is also the part of the city where lower income households are more numerous in absolute and relative terms (see Figure \ref{fig:low-income-households}).
@@
 
-The level of service by foodbank before and during the pandemic are shown in Figure \ref{fig:levels-of-service}.
+The levels of service by foodbank before and during the pandemic are shown in Figure \ref{fig:levels-of-service}.
 Since the level of service is the inverse of the number of households in the catchment area of the foodbanks, higher values mean that a facility is expected to service fewer households.
@@
 The pattern of the losses in level of service, moreover, is not uniform.
-As seen in Figure \ref{fig:levels-of-service-changes} ...
-words.
+Figure \ref{fig:levels-of-service-changes} shows that the three peripheral facilities in the southern suburban/exurban part of the city had low levels of service to begin with, but did not see major declines during the pandemic.
+Level of service deteriorated more in the central part of the city, but the loss of level of service was not as large in the core (where most of the foodbanks are), but more marked in the inner ring around the core, where facilities may have faced greater demand from both central city and suburban populations after the closure of foodbanks during the pandemic.
+To further elucidate this issue, we now turn to the results of the accessibility analysis.
 
@@
 ```
-
 ```{r aggregate-accessibility-by-taz, include=FALSE}  
@@
 
+Discuss the results of accessibility.
+
 ```{r plot-accessibility, echo=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
@@
                         direction = -1,
-                       breaks = c(0.001,
+                       breaks = c(0.002,
+                                  0.0016,
+                                  0.0012,
+                                  0.0008,
+                                  0.0004),
+                       labels = c("0.0020",
+                                  "0.0016",
+                                  "0.0012",
+                                  "0.0008",
+                                  "0.0004")) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "right")
+
+acc_covid <- ggplot() +
+  #geom_sf(data = hamilton_cma) +
+  geom_sf(data = accessibility_taz %>%
+            filter(timing == "During covid"),
+          aes(fill = accessibility)) +
+  scale_fill_viridis_c(name = "Accessibility during COVID-19",
+                        option =  "A", 
+                        direction = -1,
+                       breaks = c(0.0010,
                                   0.0008,
@@
 
-acc_covid <- ggplot() +
-  #geom_sf(data = hamilton_cma) +
-  geom_sf(data = accessibility_taz %>%
-            filter(timing == "During covid"),
-          aes(fill = accessibility)) +
-  scale_fill_viridis_c(name = "Accessibility during COVID-19",
-                        option =  "A", 
-                        direction = -1,
-                       breaks = c(0.0005,
-                                  0.0004,
-                                  0.0003,
-                                  0.0002,
-                                  0.0001),
-                       labels = c("0.0005",
-                                  "0.0004",
-                                  "0.0003",
-                                  "0.0002",
-                                  "0.0001")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
 acc_pre / acc_covid
@@
           fill = NA) +
+  geom_sf(data = inset_box,
+          linetype = "dotted",
+          color = "red",
+          fill = NA) +
   scale_fill_viridis_c(name = "Change in accessibility",
@@
                         direction = 1,
-                       breaks = c(-0.0005,
+                       breaks = c(-0.0008,
+                                  -0.0006,
                                   -0.0004,
-                                  -0.0003,
-                                  -0.0002,
-                                  -0.0001),
-                       labels = c("-0.0005",
+                                  -0.0002),
+                       labels = c("-0.0008",
+                                  "-0.0006",
                                   "-0.0004",
-                                  "-0.0003",
-                                  "-0.0002",
-                                  "-0.0001")) +
+                                  "-0.0002")) +
   theme_minimal() +
   theme(axis.text = element_blank(),
-        legend.position = "right")
-
-acc_change
-```
-
+        legend.position = "none")
+
+acc_change_inset <- ggplot() +
+  geom_sf(data = accessibilty_taz_change %>%
+            st_transform(crs = 4326),
+          aes(fill = acc_d)) +
+  geom_sf(data = inset_box,
+          linetype = "dotted",
+          color = "red",
+          fill = NA) +
+  scale_fill_viridis_c(name = "Change in accessibility",
+                        option =  "A", 
+                        direction = 1,
+                       breaks = c(-0.0008,
+                                  -0.0006,
+                                  -0.0004,
+                                  -0.0002),
+                       labels = c("-0.0008",
+                                  "-0.0006",
+                                  "-0.0004",
+                                  "-0.0002")) +
+  guides(fill = guide_legend(title.position = "top")) +
+  xlim(-79.975, -79.775) +
+  ylim(43.2, 43.3) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "bottom")
+
+acc_change / acc_change_inset
+```
 ```{r accessibility-change-by-parcel, include=FALSE}
@@
 ```
+
+Discuss the results of aggregating the accessibility by age group (see Table \ref{tab:accessibility-by-age}). Follow with the analysis of hours of travel?
 

### 2021-04-23 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Finishing touches to the figures. Completed first draft of discussion of level of service.

@@
 of Hamilton and their status. Before the pandemic there were 58 of which
-14 (24.14\%) closed during the pandemic. As seen in the figure,
+14 (24.14\%) closed during the pandemic. As shown in the figure,
 foodbanks tend to be predominantly located in the central parts of the
@@
 
-The level of service by foodbank before and during the pandemic are
+The levels of service by foodbank before and during the pandemic are
 shown in Figure \ref{fig:levels-of-service}. Since the level of service
@@
 substantially. The pattern of the losses in level of service, moreover,
-is not uniform. As seen in Figure \ref{fig:levels-of-service-changes}
-\ldots{} words.
+is not uniform. Figure \ref{fig:levels-of-service-changes} shows that
+the three peripheral facilities in the southern suburban/exurban part of
+the city had low levels of service to begin with, but did not see major
+declines during the pandemic. Level of service deteriorated more in the
+central part of the city, but the loss of level of service was not as
+large in the core (where most of the foodbanks are), but more marked in
+the inner ring around the core, where facilities may have faced greater
+demand from both central city and suburban populations after the closure
+of foodbanks during the pandemic. To further elucidate this issue, we
+now turn to the results of the accessibility analysis.
 
@@
 \end{figure}
+
+Discuss the results of accessibility.
 
@@
 \end{figure}
+
+Discuss the results of aggregating the accessibility by age group (see
+Table \ref{tab:accessibility-by-age}). Follow with the analysis of hours
+of travel?
 

### 2021-04-23 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Add to accessibility analysis

@@
 ---
-title: "Changes in accessibility to foodbanks during COVID-19 and implications for the food security of vulnerable populations in Hamilton, Ontario"
+title: "Changes in accessibility to food banks during COVID-19 and implications for the food security of vulnerable populations in Hamilton, Ontario"
 author:
@@
 data(foodbank_locations)
+data(foodbank_table)
 data(hamilton_cma)
@@
 This condition has been associated with reductions in nutritional outcomes [@bhattacharya2004poverty; @kirkpatrick2008food; @olson1999nutrition] and physical and mental health in children and adults [@elgar2021relative; @jones2017food; @ramsey2011food; @seligman2010food; @stuff2004household].
-Over at least the past four decades foodbanks have become an essential line of defense against food insecurity in Canadian communities [@riches2002food; @holmes2018nothing; @black2020examining; @tarasuk2020relationship].
+Over at least the past four decades food banks have become an essential line of defense against food insecurity in Canadian communities [@riches2002food; @holmes2018nothing; @black2020examining; @tarasuk2020relationship].
 In this respect, Canada is not unlike numerous other wealthy countries where a systematic dismantling of the welfare state took place in the intervening period [@tarasuk2014food].
@@
 The impacts of food insecurity during the pandemic are alarming, since diet-related diseases, such as obesity, heart-disease, and diabetes, were already critical public health concerns in Canada prior to COVID-19 [@boucher2017ontario].
-While foodbanks are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least they can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
-As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that foodbanks in Hamilton, Ontario, recorded the highest number of visitors in the past 29 years; the number of children visiting foodbanks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278 the year before, a rate of increase greater than population growth.
+While food banks are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least they can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
+As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that food banks in Hamilton, Ontario, recorded the highest number of visitors in the past 29 years; the number of children visiting food banks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278 the year before, a rate of increase greater than population growth.
 
 It is known that urban food environments, within which people make their daily food choices, are essential in influencing eating behaviours and health outcomes, based on factors such as food availability, ease of geographic accessibility and socio-demographic variations [@paez2010relative; @vanderlee2017food; @widener2018spatial].
-However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to foodbanks.
-Although previous work has explored differences in _accessing_ foodbanks, such as how some households utilize foodbanks over short periods of time while others regularly utilize foodbanks as longer-term resource [e.g., @enns2020experiences], we are not aware of any research that has focused on the geographic component of accessibility.
+However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to food banks.
+Although previous work has explored differences in _accessing_ food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource [e.g., @enns2020experiences], we are not aware of any research that has focused on the geographic component of accessibility.
 
 The study of place-based geographic accessibility is concerned with capturing the potential to reach destinations of value using the transportation network [@paez2012measuring].
-Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities^[https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with foodbank service delivery [@holmes2018nothing].
-However, as with research into the prevalence of food deserts, accessibility to foodbanks is unlikely to be evenly distributed, and variation throughout a city can be expected due to transportation network characteristics, and the spatial distribution of foodbank locations and the population they are meant to serve.
+Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities^[https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with food bank service delivery [@holmes2018nothing].
+However, as with research into the prevalence of food deserts, accessibility to food banks is unlikely to be evenly distributed, and variation throughout a city can be expected due to transportation network characteristics, and the spatial distribution of food bank locations and the population they are meant to serve.
 Furthermore, policy responses to the COVID-19 pandemic likely have added to the distress of vulnerable households.
 Non-pharmaceutical interventions during the pandemic involving restrictions in mobility have increased the friction of travel, in particular by transit on which low income populations are more reliant [e.g., @deweese2020tale].
-At the same time, the pandemic has created additional stress for the operators of foodbanks through disruptions in the supply chain [e.g., @mckay2021exploring] as well as concerns surrounding the delivery of service in safe conditions and possible cancellation of food service programs.
-
-For this study, we aim to look at how the landscape of foodbanks and related services (e.g.
+At the same time, the pandemic has created additional stress for the operators of food banks through disruptions in the supply chain [e.g., @mckay2021exploring] as well as concerns surrounding the delivery of service in safe conditions and possible cancellation of food service programs.
+
+For this study, we aim to look at how the landscape of food banks and related services (e.g.
 low-cost or free meal service providers) available in Hamilton, Ontario, changed during the pandemic.
-Did the number of open foodbank services diminish?
-If so, what was the accessibility to foodbanks before the pandemic from the perspective of low income households, and how it changed during the pandemic?
+Did the number of open food bank services diminish?
+If so, what was the accessibility to food banks before the pandemic from the perspective of low income households, and how it changed during the pandemic?
 And finally, who are most likely to have been impacted by changes in the accessibility landscape?
-This paper first looks at the distribution of foodbanks and related services before and during the pandemic.
+This paper first looks at the distribution of food banks and related services before and during the pandemic.
 Then, we use the balanced floating catchment area approach of Paez et al.
@@
 
-Food Insecurity and foodbank Use in Canada
+Food Insecurity and Food Bank Use in Canada
 ============
@@
 
-Foodbanks - sometimes also referred to as ‘food pantries’ and ‘food shelves’ - originated as a community response to aid those with inadequate food by voluntarily offering them meals and ingredients [@riches2002food; @loopstra2012relationship].
-The scope and objectives of foodbanks can vary by region and by country, and these organizations can include not only prepared meals and aliments, but also shared spaces to connect in community gardens and community kitchens [@wakefield2013sweet].
-Although in their origin foodbanks were meant to be provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time they have evolved into a community practice to secure emergency food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
-
-In Canada, the number of foodbanks has steadily increased in the past few decades [@wakefield2013sweet].
-The largest database of foodbanks and their use comes from the non-profit association Foodbanks Canada, which conducts an annual assessment through its affiliated members.
-Foodbanks Canada's 2018 Hunger Count report^[https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018_FINAL_EN.pdf.aspx?ext=.pdf] (the most recent available) listed 1,830 member foodbanks across the country, and found that Canadians visited foodbanks 1.1 million times in March of 2018.
-Of those accessing foodbanks, certain population characteristics tend to be over-represented compared to national totals from the 2016 Canadian Census of Population.
-According to Foodbanks Canada's 2018 data, single-adult households represent 45% of those utilizing foodbanks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those accessing foodbank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
-In addition, 59% of households accessing foodbanks list social or disability assistance as their primary source of income.
-Moreover, children aged 0-18 are over-represented as foodbank users, making up 35% of those accessing foodbank services despite representing 20% of the country's population.
+Food banks - sometimes also referred to as ‘food pantries’ and ‘food shelves’ - originated as a community response to aid those with inadequate food by voluntarily offering them meals and ingredients [@riches2002food; @loopstra2012relationship].
+The scope and objectives of food banks can vary by region and by country, and these organizations can include not only prepared meals and aliments, but also shared spaces to connect in community gardens and community kitchens [@wakefield2013sweet].
+Although in their origin food banks were meant to be provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time they have evolved into a community practice to secure emergency food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
+
+In Canada, the number of food banks has steadily increased in the past few decades [@wakefield2013sweet].
+The largest database of food banks and their use comes from the non-profit association Food Banks Canada (FBC), which conducts an annual assessment through its affiliated members.
+FBC's 2018 Hunger Count report^[https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018_FINAL_EN.pdf.aspx?ext=.pdf] (the most recent available) listed 1,830 member food banks across the country, and found that Canadians visited food banks 1.1 million times in March of 2018.
+Of those accessing food banks, certain population characteristics tend to be over-represented compared to national totals from the 2016 Canadian Census of Population.
+According to FBC's 2018 data, single-adult households represent 45% of those utilizing food banks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those accessing food bank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
+In addition, 59% of households accessing food banks list social or disability assistance as their primary source of income.
 Similarly, using data from the 2011-2012 CCHS, Tarasuk et al.
 [-@tarasuk2019geographic] found higher odds of food insecurity amongst households relying on social assistance, those without a university degree or with children under the age of 18, and individuals that lived alone, renters, and those identifying as Aboriginal.
-While surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent foodbanks in Canada [@tarasuk2014food], research from Ottawa [@enns2020experiences] and Vancouver [@black2020examining] suggests that long-term users tend to be older, have health or mobility challenges, live in large households, and are less likely to have employment income.
+While surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada [@tarasuk2014food], research from Ottawa [@enns2020experiences] and Vancouver [@black2020examining] suggests that long-term users tend to be older, have health or mobility challenges, live in large households, and are less likely to have employment income.
 
@@
 Considering the negative mental and physical health effects associated with food insecurity, increases in food insecurity rates due to the pandemic, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
-Recent data^[https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis_Exec-Sum_EN.pdf] from Foodbanks Canada showed that 52% of member foodbanks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of the country.
-The pandemic also created significant staffing issues with 42% of foodbanks reporting a reduction in volunteers.
-However, 53% of foodbanks reported a decrease in use into the summer of 2020 which foodbanks attributed to emergency financial support programs from the federal government were.
-Nevertheless, some of these benefit programs were temporary, which suggests that many households may again turn to foodbank services to meet their needs.
+Recent data^[https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis_Exec-Sum_EN.pdf] from FBC showed that 52% of member food banks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of the country.
+The pandemic also created significant staffing issues with 42% of food banks reporting a reduction in volunteers.
+However, 53% of food banks later reported a decrease in use into the summer of 2020 which FBC members attributed to emergency financial support programs from the federal government.
+Nevertheless, some of these benefit programs were temporary, which suggests that many households may again turn to food bank services to meet their needs.
 
 In terms of geography, previous research conducted at the provincial scale using data from the 2011-2012 CCHS found that the prevalence of food insecurity ranged across the country from 11.8% of households in Ontario to 41% of households in Nunavut [@tarasuk2019geographic].
-However, although previous research has examined the characteristics of individuals and households accessing foodbanks, the locational or transportation accessibility aspect of foodbank access is not well understood.
+However, although previous research has examined the characteristics of individuals and households accessing food banks, the locational or transportation accessibility aspect of food bank access is not well understood.
 A wealth of literature examining the food desert concept suggests that, in addition to socio-economic and demographic factors, location and transportation networks play a key role in a household's accessibility to healthy foods [@paez2010relative; @vanderlee2017food; @widener2018spatial].
-For foodbanks specifically, previous qualitative research in Ontario has noted difficulties associated with transportation for trips to the foodbank [@smith2017food].
-For example, Smith et al.
-[-@smith2017food] note that participants access foodbanks using a variety of modes of transportation with respondent stating "I wish it [the foodbank] was a little more centrally located.
+For food banks specifically, previous qualitative research by Smith et al.
+[-@smith2017food] in Ontario has noted that participants access food banks using a variety of modes of transportation and that some food bank users have expressed difficulties with transportation for trips to food bank locations, with one respondent stating "I wish it [the food bank] was a little more centrally located.
 Because if I didn’t have a bike I’d have to walk it all the way out there and back.
-I wonder about people who don’t." From this, the authors conclude that "transportation can be challenging, particularly if the foodbank is situated in a remote location." To offer greater insight into the role of transportation and location in foodbank accessibility, this research examines how geographic accessibility to foodbanks and food services changed in Hamilton during the COVID-19 pandemic.
+I wonder about people who don’t." From this, the authors conclude that "transportation can be challenging, particularly if the food bank is situated in a remote location." To offer greater insight into the role of transportation and location in food bank accessibility, this research examines how geographic accessibility to food banks and food services changed in Hamilton during the COVID-19 pandemic.
 
@@
 
-### Destinations: foodbanks and Food Service Locations
-
-The locations of foodbanks and related food services were obtained from the Hamilton Public Library's Food Access Guide^[http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf].
+### Destinations: Food Banks and Food Service Locations
+
+The locations of food banks and related food services were obtained from the Hamilton Public Library's Food Access Guide^[http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf].
 The guide was updated in April of 2021 to indicate any change affected on the services due to the pandemic.
 This includes modified business hours, a need to make reservations before frequenting, and locations that have completely shut down in consequence.
-While some foodbank services have a specific a target population, such as prioritizing family with young children aged between 0 and 3 or accepting only those providing proof of low-income status through housing and utility statements, all the foodbank services indicated below are designed to accommodate those in need of food at zero to low cost.
-With our focus on foodbanks and food services that offer free or low-cost meals at particular locations, we first removed services such as Meals on Wheels and other food access services such as food box, community kitchens, student nutrition programs, and shopping and transportation.
+While some food bank services have a specific a target population, such as prioritizing family with young children aged between 0 and 3 or accepting only those providing proof of low-income status through housing and utility statements, all the food bank services indicated below are designed to accommodate those in need of food at zero to low cost.
+With our focus on food banks and food services that offer free or low-cost meals at particular locations, we first removed services such as Meals on Wheels and other food access services such as food box, community kitchens, student nutrition programs, and shopping and transportation.
 In addition, two free meal services held on different days at the same location were collapsed into a single service point for the accessibility analysis.
@@
 
-Travel time tables for three modes (car, transit, walking) were computed using the parcels as the origins and the locations of the foodbanks as the destinations.
+Travel time tables for three modes (car, transit, walking) were computed using the parcels as the origins and the locations of the food banks as the destinations.
 For routing, the package `r5r` [@pereira2021r5r] was used with a network extract for the City of Hamilton from OpenStreetMaps and the General Transit Feed Specification from Hamilton Street Railway, the local transit operator.
@@
 
-Once we obtained travel time tables with population (number of households) and proportion of trips by mode, we calculated the expected travel time $ett$ from each parcel $i$ to a foodbank or food service location $j$ as follows:
+Once we obtained travel time tables with population (number of households) and proportion of trips by mode, we calculated the expected travel time $ett$ from each parcel $i$ to a food bank or food service location $j$ as follows:
 $$
@@
 
-\noindent where $p^k_i$ is the proportion of trips by mode $k$ in the TAZ of parcel $i$, and $tt^k_{ij}$ is the travel time from parcel $i$ to the foodbank.
-In other words, the expected travel time is the weighted sum of travel times to the foodbank, with the weights given by the expected modal split in the TAZ.
+\noindent where $p^k_i$ is the proportion of trips by mode $k$ in the TAZ of parcel $i$, and $tt^k_{ij}$ is the travel time from parcel $i$ to the food bank.
+In other words, the expected travel time is the weighted sum of travel times to the food bank, with the weights given by the expected modal split in the TAZ.
 
@@
 
-Figure \ref{fig:foodbanks} shows the location of foodbanks in the City of Hamilton and their status.
+Figure \ref{fig:foodbanks} shows the location of food banks in the City of Hamilton and their status.
 Before the pandemic there were `r nrow(foodbank_locations)` of which `r nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))` (`r round(nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations) * 100, 2)`\%) closed during the pandemic.
-As shown in the figure, foodbanks tend to be predominantly located in the central parts of the city.
+As shown in the figure, food banks tend to be predominantly located in the central parts of the city.
 This is not surprising: population density is high there, and it is also the part of the city where lower income households are more numerous in absolute and relative terms (see Figure \ref{fig:low-income-households}).
@@
 To implement the accessibility calculations, we must select a distance-decay function.
-In this task we find limited support in the literature, which is mostly silent on the travel patterns of people who visit foodbanks.
+In this task we find limited support in the literature, which is mostly silent on the travel patterns of people who visit food banks.
 For this reason, we opt for a simple cumulative opportunities function as follows:
@@
 Figure \ref{fig:sensitivity-analysis} shows the results of conducting sensitivity analysis of the system-wide accessibility as we vary the threshold (considering the situation before the pandemic).
-There is a clear pattern whereby more strict values of $\delta$ are associated with higher levels of system-wide accessibility: this is a result of lower congestion, since fewer households are serviced.
+There is a clear pattern whereby more strict values of $\delta$ are associated with higher levels of system-wide accessibility: while increases in accessibility that result from decreases in the travel time window might seem counter-intuitive, this is a result of lower _congestion_, since fewer households are serviced and are thus competing for the same resources.
 System-wide accessibility declines with higher values of $\delta$: as more households are serviced, congestion grows and the level of service declines, although this happens at a declining rate.
-We are not aware of any standards that explain how long people are expected to travel for foodbanks, but we note that in developing countries, access to drinking water is defined as a source that takes less than 30 minutes to reach [round trip, see @world2019progress].
+We are not aware of any standards that explain how long people are expected to travel for food banks, but we note that in developing countries, access to drinking water is defined as a source that takes less than 30 minutes to reach [round trip, see @world2019progress].
 There is no reason why people in affluent countries should be expected to travel more for a basic need such as food.
 Accordingly, we adopt a threshold of 15 minutes for the analysis (representing a one-way trip).
-Using this threshold, we find that the system-wide accessibility (interpreted as a provider-to-population ratio) was `r round(sum(results_pre$accessibility$accessibility), 3)` (foodbanks per low income household in the city) before COVID-19, but declined to `r round(sum(results_covid$accessibility$accessibility), 3)` during the pandemic.
-It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`\% of foodbanks remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`\%.
 
@@
 ```
+
+Using this threshold, we find that the system-wide accessibility (interpreted as a provider-to-population ratio) was `r round(sum(results_pre$accessibility$accessibility), 3)` (food banks and food service locations per low income household in the city) before COVID-19, but declined to `r round(sum(results_covid$accessibility$accessibility), 3)` during the pandemic.
+It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`\% of food banks remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`\%, suggesting the location of food banks and food services plays an important role in serving those in need.
 
@@
 
-The levels of service by foodbank before and during the pandemic are shown in Figure \ref{fig:levels-of-service}.
-Since the level of service is the inverse of the number of households in the catchment area of the foodbanks, higher values mean that a facility is expected to service fewer households.
+Turning to the individual food bank and food service locations, the levels of service offered before and during the pandemic are shown in Figure \ref{fig:levels-of-service}.
+The level of service is functionally the inverse (because each food bank has a "capacity" of 1) of the number of low-income households in the travel-mode weighted travel time catchment area of the food banks.
+Higher values mean that a facility is expected to service fewer households.
 Conversely, lower values indicate grater congestion.
+
 The general pattern of the levels of service is similar before and during the pandemic, with lower values in the center of the city.
@@
 The pattern of the losses in level of service, moreover, is not uniform.
-Figure \ref{fig:levels-of-service-changes} shows that the three peripheral facilities in the southern suburban/exurban part of the city had low levels of service to begin with, but did not see major declines during the pandemic.
-Level of service deteriorated more in the central part of the city, but the loss of level of service was not as large in the core (where most of the foodbanks are), but more marked in the inner ring around the core, where facilities may have faced greater demand from both central city and suburban populations after the closure of foodbanks during the pandemic.
-To further elucidate this issue, we now turn to the results of the accessibility analysis.
+The upper pane of Figure \ref{fig:levels-of-service-changes} shows that the three peripheral facilities in the southern suburban/exurban part of the city had low levels of service to begin with, but did not see major declines during the pandemic.
+Turning to the inset map, levels of service deteriorated more in the central part of the city.
+However, the loss of level of service was not as large in the core (where most of the food banks are), but more marked in the inner ring around the core where facilities may have faced greater demand from both central city and suburban populations after the closure of food banks during the pandemic.
 
@@
           aes(shape = COVIDClose)) +
-  scale_color_viridis_c(name = "Level of service pre-COVID-19",
+  scale_color_viridis_c(name = "Level of service \npre-COVID-19",
                         option =  "A", 
@@
           aes(shape = COVIDClose)) +
-  scale_color_viridis_c(name = "Level of service during COVID-19",
+  scale_color_viridis_c(name = "Level of service \nduring COVID-19",
                         option =  "A", 
@@
 ```
+
+To further elucidate this issue, we now turn to the results of the accessibility analysis.
+The top pane of Figure \ref{fig:accessibility} reveals that, compared with the outer rural zones, the smaller more urban zones of the city generally exhibit higher accessibility to food banks and food service locations.
+However, the pattern is not particularly smooth - this is largely attributable to the travel-mode weighting of travel times according to the trip patterns of low-income household respondents to the TTS.
+For example, in zones where low-income households make a high proportion of trips by walking, access to food bank locations by walking is afforded a concomitantly high weight in our calculations of travel time compared to transit or car travel.
+From this, highly-accessible locations result from a mix of characteristics: low-income households in locations where travel options that align with zonal modal split are available to connect them to food bank locations with high levels of service within 15 minutes.
+
 ```{r aggregate-accessibility-by-taz, include=FALSE}  
@@
 
-Discuss the results of accessibility.
-
 ```{r plot-accessibility, echo=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
@@
           aes(fill = accessibility)) +
-  scale_fill_viridis_c(name = "Accessibility pre-COVID-19",
+  scale_fill_viridis_c(name = "Accessibility \npre-COVID-19",
                         option =  "A", 
@@
           aes(fill = accessibility)) +
-  scale_fill_viridis_c(name = "Accessibility during COVID-19",
+  scale_fill_viridis_c(name = "Accessibility \nduring COVID-19",
                         option =  "A", 

### 2021-04-23 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Add to accessibility analysis

@@
 \usepackage{makecell}
-\usepackage{xcolor}
 
@@
 few decades (Wakefield et al., 2013). The largest database of foodbanks
-and their use comes from the non-profit association Foodbanks Canada,
-which conducts an annual assessment through its affiliated members.
-Foodbanks Canada's 2018 Hunger Count report\footnote{https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018\_FINAL\_EN.pdf.aspx?ext=.pdf}
+and their use comes from the non-profit association Food Banks Canada
+(FBC), which conducts an annual assessment through its affiliated
+members. FBC's 2018 Hunger Count report\footnote{https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018\_FINAL\_EN.pdf.aspx?ext=.pdf}
 (the most recent available) listed 1,830 member foodbanks across the
@@
 characteristics tend to be over-represented compared to national totals
-from the 2016 Canadian Census of Population. According to Foodbanks
-Canada's 2018 data, single-adult households represent 45\% of those
-utilizing foodbanks despite making up 28\% of Canada's population, 19\%
-are single-parent households (compared to 10\% nationally), and 35\% of
+from the 2016 Canadian Census of Population. According to FBC's 2018
+data, single-adult households represent 45\% of those utilizing
+foodbanks despite making up 28\% of Canada's population, 19\% are
+single-parent households (compared to 10\% nationally), and 35\% of
 those accessing foodbank services are children aged 0-18 even though
@@
 addition, 59\% of households accessing foodbanks list social or
-disability assistance as their primary source of income. Moreover,
-children aged 0-18 are over-represented as foodbank users, making up
-35\% of those accessing foodbank services despite representing 20\% of
-the country's population. Similarly, using data from the 2011-2012 CCHS,
-Tarasuk et al. (2019) found higher odds of food insecurity amongst
-households relying on social assistance, those without a university
-degree or with children under the age of 18, and individuals that lived
-alone, renters, and those identifying as Aboriginal. While surveys
-revealed that only 20 to 30 percent of those experiencing food
-insecurity were found to frequent foodbanks in Canada (Tarasuk et al.,
-2014), research from Ottawa (Enns et al., 2020) and Vancouver (Black and
-Seto, 2020) suggests that long-term users tend to be older, have health
-or mobility challenges, live in large households, and are less likely to
-have employment income.
+disability assistance as their primary source of income. Similarly,
+using data from the 2011-2012 CCHS, Tarasuk et al. (2019) found higher
+odds of food insecurity amongst households relying on social assistance,
+those without a university degree or with children under the age of 18,
+and individuals that lived alone, renters, and those identifying as
+Aboriginal. While surveys revealed that only 20 to 30 percent of those
+experiencing food insecurity were found to frequent foodbanks in Canada
+(Tarasuk et al., 2014), research from Ottawa (Enns et al., 2020) and
+Vancouver (Black and Seto, 2020) suggests that long-term users tend to
+be older, have health or mobility challenges, live in large households,
+and are less likely to have employment income.
 
@@
 beyond (Niles et al., 2020). Recent data\footnote{https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report\_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis\_Exec-Sum\_EN.pdf}
-from Foodbanks Canada showed that 52\% of member foodbanks reported an
-increase in usage in March of 2020 when initial lockdown restrictions
-were put in place across much of the country. The pandemic also created
-significant staffing issues with 42\% of foodbanks reporting a reduction
-in volunteers. However, 53\% of foodbanks reported a decrease in use
-into the summer of 2020 which foodbanks attributed to emergency
-financial support programs from the federal government were.
-Nevertheless, some of these benefit programs were temporary, which
-suggests that many households may again turn to foodbank services to
-meet their needs.
+from FBC showed that 52\% of member foodbanks reported an increase in
+usage in March of 2020 when initial lockdown restrictions were put in
+place across much of the country. The pandemic also created significant
+staffing issues with 42\% of foodbanks reporting a reduction in
+volunteers. However, 53\% of foodbanks later reported a decrease in use
+into the summer of 2020 which FBC members attributed to emergency
+financial support programs from the federal government. Nevertheless,
+some of these benefit programs were temporary, which suggests that many
+households may again turn to foodbank services to meet their needs.
 
@@
 2017; Widener, 2018). For foodbanks specifically, previous qualitative
-research in Ontario has noted difficulties associated with
-transportation for trips to the foodbank (Smith-Carrier et al., 2017).
-For example, Smith et al. (2017) note that participants access foodbanks
-using a variety of modes of transportation with respondent stating ``I
-wish it {[}the foodbank{]} was a little more centrally located. Because
-if I didn't have a bike I'd have to walk it all the way out there and
-back. I wonder about people who don't.'' From this, the authors conclude
-that ``transportation can be challenging, particularly if the foodbank
-is situated in a remote location.'' To offer greater insight into the
-role of transportation and location in foodbank accessibility, this
-research examines how geographic accessibility to foodbanks and food
-services changed in Hamilton during the COVID-19 pandemic.
+research by Smith et al. (2017) in Ontario has noted that participants
+access foodbanks using a variety of modes of transportation and that
+some food bank users have expressed difficulties with transportation for
+trips to foodbank locations, with one respondent stating ``I wish it
+{[}the foodbank{]} was a little more centrally located. Because if I
+didn't have a bike I'd have to walk it all the way out there and back. I
+wonder about people who don't.'' From this, the authors conclude that
+``transportation can be challenging, particularly if the foodbank is
+situated in a remote location.'' To offer greater insight into the role
+of transportation and location in foodbank accessibility, this research
+examines how geographic accessibility to foodbanks and food services
+changed in Hamilton during the COVID-19 pandemic.
 

### 2021-04-23 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Added LISA analysis of changes in accessibility and rebuilt package to include Hamilton boundary that I had mistakenly deleted before

@@
 library(sf)
+library(spdep)
 library(tidyverse)
@@
 data(foodbank_table)
-data(hamilton_cma)
+data(hamilton_boundary)
 data(modes_less40k)
@@
 fb <- ggplot() +
-  geom_sf(data = hamilton_cma,
+  geom_sf(data = hamilton_boundary,
           fill = "lightgray") +
@@
 fb_inset <- ggplot() +
-  geom_sf(data = hamilton_cma,
+  geom_sf(data = hamilton_boundary %>% 
+            st_transform(4326),
           fill = "lightgray") +
@@
 lc <- ggplot() +
-  geom_sf(data = hamilton_cma) +
+  geom_sf(data = hamilton_boundary) +
   geom_sf(data = los_change, 
@@
 lc_inset <- ggplot() +
-  geom_sf(data = hamilton_cma) +
+  geom_sf(data = hamilton_boundary %>% 
+            st_transform(4326)) +
   geom_sf(data = los_change, 
@@
 acc_pre <- ggplot() +
-  #geom_sf(data = hamilton_cma) +
+  #geom_sf(data = hamilton_boundary) +
   geom_sf(data = accessibility_taz %>%
@@
 acc_covid <- ggplot() +
-  #geom_sf(data = hamilton_cma) +
+  #geom_sf(data = hamilton_boundary) +
   geom_sf(data = accessibility_taz %>%
@@
 acc_change <- ggplot() +
-  #geom_sf(data = hamilton_cma) +
+  #geom_sf(data = hamilton_boundary) +
   geom_sf(data = accessibilty_taz_change,
@@
 ```
+```{r weights-matrix, include=FALSE}
+taz_nb <- poly2nb(accessibilty_taz_change, queen = TRUE)
+taz_w <- nb2listw(taz_nb)
+```
+
+```{r moran-test-accessibility-changes, include=FALSE}
+moran.test(accessibilty_taz_change$acc_d, taz_w)
+```
+
+```{r local-moran-i, include=FALSE}
+local_i <- localmoran(accessibilty_taz_change$acc_d,
+                      taz_w, 
+                      alternative = "two.sided",
+                      p.adjust.method = "none") %>% 
+  data.frame() %>%
+  rename(p_val = Pr.z....0.) %>%
+  mutate(significance = ifelse(p_val <= 0.05, "sig", "n.s."),
+         significance = factor(significance,
+                               levels = c("n.s.", "sig")))
+
+summary(local_i)
+```
+
+```{r join-local-i-to-table, include=FALSE}
+accessibilty_taz_change$significance <- local_i$significance
+```
+
+```{r plot-local-i, echo=FALSE, fig.cap="\\label{fig:plot-local-i}Areas with significantly large changes in accessibility from pre-COVID-19 to during COVID-19 according to Local Moran's I." }
+ggplot() +
+  geom_sf(data = accessibilty_taz_change,
+          aes(fill = acc_d,
+              color = significance,
+              alpha = significance),
+          size = 1) +
+  geom_sf(data = hamilton_boundary,
+          fill = NA) +
+  geom_sf(data = inset_box,
+          linetype = "dotted",
+          color = "red",
+          fill = NA) +
+  scale_fill_viridis_c(name = "Change in accessibility",
+                        option =  "A", 
+                        direction = 1,
+                       breaks = c(-0.0008,
+                                  -0.0006,
+                                  -0.0004,
+                                  -0.0002),
+                       labels = c("-0.0008",
+                                  "-0.0006",
+                                  "-0.0004",
+                                  "-0.0002")) +
+  scale_color_manual(values = c("n.s." = NA, 
+                                "sig" = "black")) +
+  scale_alpha_manual(values = c("n.s." = 0.5, 
+                                "sig" = 1)) +
+  guides(color = FALSE, 
+         alpha = FALSE,
+         fill = guide_legend(title.position = "top")) +
+  theme_minimal() +
+  theme(axis.text = element_blank(),
+        legend.position = "bottom")
+```
+
 ```{r accessibility-change-by-parcel, include=FALSE}

### 2021-04-23 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Added LISA analysis of changes in accessibility and rebuilt package to include Hamilton boundary that I had mistakenly deleted before

--- 2021-04-23 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
@@
             pdfauthor={},
-            pdftitle={Changes in accessibility to foodbanks during COVID-19 and implications for the food security of vulnerable populations in Hamilton, Ontario},
+            pdftitle={Changes in accessibility to food banks during COVID-19 and implications for the food security of vulnerable populations in Hamilton, Ontario},
             colorlinks=false,
@@
 
-  \title{Changes in accessibility to foodbanks during COVID-19 and
+  \title{Changes in accessibility to food banks during COVID-19 and
 implications for the food security of vulnerable populations in
@@
 2021; Jones, 2017; Ramsey et al., 2011; Seligman et al., 2010; Stuff et
-al., 2004). Over at least the past four decades foodbanks have become an
-essential line of defense against food insecurity in Canadian
+al., 2004). Over at least the past four decades food banks have become
+an essential line of defense against food insecurity in Canadian
 communities (Black and Seto, 2020; Holmes et al., 2018; Riches, 2002;
@@
 were already critical public health concerns in Canada prior to COVID-19
-(Boucher et al., 2017). While foodbanks are not necessarily a stable
+(Boucher et al., 2017). While food banks are not necessarily a stable
 solution to food insecurity and in fact may encourage a retrenchment of
@@
 Hunger Report\footnote{https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf}
-noted that foodbanks in Hamilton, Ontario, recorded the highest number
-of visitors in the past 29 years; the number of children visiting
-foodbanks (minors up to 18 years old) was 9,125 in March 2019, up from
-8,278 the year before, a rate of increase greater than population
-growth.
+noted that food banks in Hamilton, Ontario, recorded the highest number
+of visitors in the past 29 years; the number of children visiting food
+banks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278
+the year before, a rate of increase greater than population growth.
 
@@
 accessibility to healthy food through the ``food desert'' concept, there
-has been little research into accessibility to foodbanks. Although
-previous work has explored differences in \emph{accessing} foodbanks,
-such as how some households utilize foodbanks over short periods of time
-while others regularly utilize foodbanks as longer-term resource (e.g.,
-Enns et al., 2020), we are not aware of any research that has focused on
-the geographic component of accessibility.
+has been little research into accessibility to food banks. Although
+previous work has explored differences in \emph{accessing} food banks,
+such as how some households utilize food banks over short periods of
+time while others regularly utilize food banks as longer-term resource
+(e.g., Enns et al., 2020), we are not aware of any research that has
+focused on the geographic component of accessibility.
 
@@
 and previous survey research has suggested that such accessibility plays
-a key role in user satisfaction with foodbank service delivery (Holmes
+a key role in user satisfaction with food bank service delivery (Holmes
 et al., 2018). However, as with research into the prevalence of food
-deserts, accessibility to foodbanks is unlikely to be evenly
+deserts, accessibility to food banks is unlikely to be evenly
 distributed, and variation throughout a city can be expected due to
 transportation network characteristics, and the spatial distribution of
-foodbank locations and the population they are meant to serve.
+food bank locations and the population they are meant to serve.
 Furthermore, policy responses to the COVID-19 pandemic likely have added
@@
 the same time, the pandemic has created additional stress for the
-operators of foodbanks through disruptions in the supply chain (e.g.,
+operators of food banks through disruptions in the supply chain (e.g.,
 McKay et al., 2021) as well as concerns surrounding the delivery of
@@
 
-For this study, we aim to look at how the landscape of foodbanks and
+For this study, we aim to look at how the landscape of food banks and
 related services (e.g.~low-cost or free meal service providers)
 available in Hamilton, Ontario, changed during the pandemic. Did the
-number of open foodbank services diminish? If so, what was the
-accessibility to foodbanks before the pandemic from the perspective of
+number of open food bank services diminish? If so, what was the
+accessibility to food banks before the pandemic from the perspective of
 low income households, and how it changed during the pandemic? And
@@
 accessibility landscape? This paper first looks at the distribution of
-foodbanks and related services before and during the pandemic. Then, we
+food banks and related services before and during the pandemic. Then, we
 use the balanced floating catchment area approach of Paez et al. (2019)
@@
 
-\hypertarget{food-insecurity-and-foodbank-use-in-canada}{%
-\section{Food Insecurity and foodbank Use in
-Canada}\label{food-insecurity-and-foodbank-use-in-canada}}
+\hypertarget{food-insecurity-and-food-bank-use-in-canada}{%
+\section{Food Insecurity and Food Bank Use in
+Canada}\label{food-insecurity-and-food-bank-use-in-canada}}
 
@@
 
-Foodbanks - sometimes also referred to as `food pantries' and `food
+Food banks - sometimes also referred to as `food pantries' and `food
 shelves' - originated as a community response to aid those with
@@
 (Loopstra and Tarasuk, 2012; Riches, 2002). The scope and objectives of
-foodbanks can vary by region and by country, and these organizations can
-include not only prepared meals and aliments, but also shared spaces to
-connect in community gardens and community kitchens (Wakefield et al.,
-2013). Although in their origin foodbanks were meant to be provide a
-temporary solution to accommodate those in hunger due to job
+food banks can vary by region and by country, and these organizations
+can include not only prepared meals and aliments, but also shared spaces
+to connect in community gardens and community kitchens (Wakefield et
+al., 2013). Although in their origin food banks were meant to be provide
+a temporary solution to accommodate those in hunger due to job
 retrenchments and economic downfalls since the 1980s, over time they
@@
 
-In Canada, the number of foodbanks has steadily increased in the past
-few decades (Wakefield et al., 2013). The largest database of foodbanks
-and their use comes from the non-profit association Foodbanks Canada,
-which conducts an annual assessment through its affiliated members.
-Foodbanks Canada's 2018 Hunger Count report\footnote{https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018\_FINAL\_EN.pdf.aspx?ext=.pdf}
-(the most recent available) listed 1,830 member foodbanks across the
-country, and found that Canadians visited foodbanks 1.1 million times in
-March of 2018. Of those accessing foodbanks, certain population
+In Canada, the number of food banks has steadily increased in the past
+few decades (Wakefield et al., 2013). The largest database of food banks
+and their use comes from the non-profit association Food Banks Canada
+(FBC), which conducts an annual assessment through its affiliated
+members. FBC's 2018 Hunger Count report\footnote{https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018\_FINAL\_EN.pdf.aspx?ext=.pdf}
+(the most recent available) listed 1,830 member food banks across the
+country, and found that Canadians visited food banks 1.1 million times
+in March of 2018. Of those accessing food banks, certain population
 characteristics tend to be over-represented compared to national totals
-from the 2016 Canadian Census of Population. According to Foodbanks
-Canada's 2018 data, single-adult households represent 45\% of those
-utilizing foodbanks despite making up 28\% of Canada's population, 19\%
-are single-parent households (compared to 10\% nationally), and 35\% of
-those accessing foodbank services are children aged 0-18 even though
+from the 2016 Canadian Census of Population. According to FBC's 2018
+data, single-adult households represent 45\% of those utilizing food
+banks despite making up 28\% of Canada's population, 19\% are
+single-parent households (compared to 10\% nationally), and 35\% of
+those accessing food bank services are children aged 0-18 even though
 their share of Canada's national population is approximately 20\%. In
-addition, 59\% of households accessing foodbanks list social or
-disability assistance as their primary source of income. Moreover,
-children aged 0-18 are over-represented as foodbank users, making up
-35\% of those accessing foodbank services despite representing 20\% of
-the country's population. Similarly, using data from the 2011-2012 CCHS,
-Tarasuk et al. (2019) found higher odds of food insecurity amongst
-households relying on social assistance, those without a university
-degree or with children under the age of 18, and individuals that lived
-alone, renters, and those identifying as Aboriginal. While surveys
-revealed that only 20 to 30 percent of those experiencing food
-insecurity were found to frequent foodbanks in Canada (Tarasuk et al.,
-2014), research from Ottawa (Enns et al., 2020) and Vancouver (Black and
-Seto, 2020) suggests that long-term users tend to be older, have health
-or mobility challenges, live in large households, and are less likely to
-have employment income.
+addition, 59\% of households accessing food banks list social or
+disability assistance as their primary source of income. Similarly,
+using data from the 2011-2012 CCHS, Tarasuk et al. (2019) found higher
+odds of food insecurity amongst households relying on social assistance,
+those without a university degree or with children under the age of 18,
+and individuals that lived alone, renters, and those identifying as
+Aboriginal. While surveys revealed that only 20 to 30 percent of those
+experiencing food insecurity were found to frequent food banks in Canada
+(Tarasuk et al., 2014), research from Ottawa (Enns et al., 2020) and
+Vancouver (Black and Seto, 2020) suggests that long-term users tend to
+be older, have health or mobility challenges, live in large households,
+and are less likely to have employment income.
 
@@
 beyond (Niles et al., 2020). Recent data\footnote{https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report\_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis\_Exec-Sum\_EN.pdf}
-from Foodbanks Canada showed that 52\% of member foodbanks reported an
-increase in usage in March of 2020 when initial lockdown restrictions
-were put in place across much of the country. The pandemic also created
-significant staffing issues with 42\% of foodbanks reporting a reduction
-in volunteers. However, 53\% of foodbanks reported a decrease in use
-into the summer of 2020 which foodbanks attributed to emergency
-financial support programs from the federal government were.
-Nevertheless, some of these benefit programs were temporary, which
-suggests that many households may again turn to foodbank services to
-meet their needs.
+from FBC showed that 52\% of member food banks reported an increase in
+usage in March of 2020 when initial lockdown restrictions were put in
+place across much of the country. The pandemic also created significant
+staffing issues with 42\% of food banks reporting a reduction in
+volunteers. However, 53\% of food banks later reported a decrease in use
+into the summer of 2020 which FBC members attributed to emergency
+financial support programs from the federal government. Nevertheless,
+some of these benefit programs were temporary, which suggests that many
+households may again turn to food bank services to meet their needs.
 
@@
 However, although previous research has examined the characteristics of
-individuals and households accessing foodbanks, the locational or
-transportation accessibility aspect of foodbank access is not well
+individuals and households accessing food banks, the locational or
+transportation accessibility aspect of food bank access is not well
 understood. A wealth of literature examining the food desert concept
@@
 accessibility to healthy foods (Paez et al., 2010; Vanderlee and L'Abbé,
-2017; Widener, 2018). For foodbanks specifically, previous qualitative
-research in Ontario has noted difficulties associated with
-transportation for trips to the foodbank (Smith-Carrier et al., 2017).
-For example, Smith et al. (2017) note that participants access foodbanks
-using a variety of modes of transportation with respondent stating ``I
-wish it {[}the foodbank{]} was a little more centrally located. Because
-if I didn't have a bike I'd have to walk it all the way out there and
-back. I wonder about people who don't.'' From this, the authors conclude
-that ``transportation can be challenging, particularly if the foodbank
-is situated in a remote location.'' To offer greater insight into the
-role of transportation and location in foodbank accessibility, this
-research examines how geographic accessibility to foodbanks and food
-services changed in Hamilton during the COVID-19 pandemic.
+2017; Widener, 2018). For food banks specifically, previous qualitative
+research by Smith et al. (2017) in Ontario has noted that participants
+access food banks using a variety of modes of transportation and that
+some food bank users have expressed difficulties with transportation for
+trips to food bank locations, with one respondent stating ``I wish it
+{[}the food bank{]} was a little more centrally located. Because if I
+didn't have a bike I'd have to walk it all the way out there and back. I
+wonder about people who don't.'' From this, the authors conclude that
+``transportation can be challenging, particularly if the food bank is
+situated in a remote location.'' To offer greater insight into the role
+of transportation and location in food bank accessibility, this research
+examines how geographic accessibility to food banks and food services
+changed in Hamilton during the COVID-19 pandemic.
 
@@
 
-\hypertarget{destinations-foodbanks-and-food-service-locations}{%
-\subsubsection{Destinations: foodbanks and Food Service
-Locations}\label{destinations-foodbanks-and-food-service-locations}}
+\hypertarget{destinations-food-banks-and-food-service-locations}{%
+\subsubsection{Destinations: Food Banks and Food Service
+Locations}\label{destinations-food-banks-and-food-service-locations}}
 
-The locations of foodbanks and related food services were obtained from
+The locations of food banks and related food services were obtained from
 the Hamilton Public Library's Food Access Guide\footnote{http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf}.
@@
 hours, a need to make reservations before frequenting, and locations
-that have completely shut down in consequence. While some foodbank
+that have completely shut down in consequence. While some food bank
 services have a specific a target population, such as prioritizing
@@
 providing proof of low-income status through housing and utility
-statements, all the foodbank services indicated below are designed to
+statements, all the food bank services indicated below are designed to
 accommodate those in need of food at zero to low cost. With our focus on
-foodbanks and food services that offer free or low-cost meals at
+food banks and food services that offer free or low-cost meals at
 particular locations, we first removed services such as Meals on Wheels
@@
 Travel time tables for three modes (car, transit, walking) were computed
-using the parcels as the origins and the locations of the foodbanks as
+using the parcels as the origins and the locations of the food banks as
 the destinations. For routing, the package \texttt{r5r} (Pereira et al.,
@@
 households) and proportion of trips by mode, we calculated the expected
-travel time \(ett\) from each parcel \(i\) to a foodbank or food service
-location \(j\) as follows: \[
+travel time \(ett\) from each parcel \(i\) to a food bank or food
+service location \(j\) as follows: \[
 ett_{ij} = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
@@
 the TAZ of parcel \(i\), and \(tt^k_{ij}\) is the travel time from
-parcel \(i\) to the foodbank. In other words, the expected travel time
-is the weighted sum of travel times to the foodbank, with the weights
+parcel \(i\) to the food bank. In other words, the expected travel time
+is the weighted sum of travel times to the food bank, with the weights
 given by the expected modal split in the TAZ.
@@
 
-Figure \ref{fig:foodbanks} shows the location of foodbanks in the City
+Figure \ref{fig:foodbanks} shows the location of food banks in the City
 of Hamilton and their status. Before the pandemic there were 58 of which
-14 (24.14\%) closed during the pandemic. As seen in the figure,
-foodbanks tend to be predominantly located in the central parts of the
-city. This is not surprising: population density is high there, and it
-is also the part of the city where lower income households are more
-numerous in absolute and relative terms (see Figure
+14 (24.14\%) closed during the pandemic. As shown in the figure, food
+banks tend to be predominantly located in the central parts of the city.
+This is not surprising: population density is high there, and it is also
+the part of the city where lower income households are more numerous in
+absolute and relative terms (see Figure
 \ref{fig:low-income-households}). Alas, this is also the part of the
@@
 literature, which is mostly silent on the travel patterns of people who
-visit foodbanks. For this reason, we opt for a simple cumulative
+visit food banks. For this reason, we opt for a simple cumulative
 opportunities function as follows: \[
@@
 clear pattern whereby more strict values of \(\delta\) are associated
-with higher levels of system-wide accessibility: this is a result of
-lower congestion, since fewer households are serviced. System-wide
-accessibility declines with higher values of \(\delta\): as more
-households are serviced, congestion grows and the level of service
-declines, although this happens at a declining rate. We are not aware of
-any standards that explain how long people are expected to travel for
-foodbanks, but we note that in developing countries, access to drinking
-water is defined as a source that takes less than 30 minutes to reach
-(round trip, see UNICEF-WHO, 2019). There is no reason why people in
-affluent countries should be expected to travel more for a basic need
-such as food. Accordingly, we adopt a threshold of 15 minutes for the
-analysis (representing a one-way trip). Using this threshold, we find
-that the system-wide accessibility (interpreted as a
-provider-to-population ratio) was 0.065 (foodbanks per low income
-household in the city) before COVID-19, but declined to 0.037 during the
-pandemic. It is striking that although almost 76\% of foodbanks remained
-in operation during the pandemic, there was a loss of accessibility
-greater than 43\%.
+with higher levels of system-wide accessibility: while increases in
+accessibility that result from decreases in the travel time window might
+seem counter-intuitive, this is a result of lower \emph{congestion},
+since fewer households are serviced and are thus competing for the same
+resources. System-wide accessibility declines with higher values of
+\(\delta\): as more households are serviced, congestion grows and the
+level of service declines, although this happens at a declining rate. We
+are not aware of any standards that explain how long people are expected
+to travel for food banks, but we note that in developing countries,
+access to drinking water is defined as a source that takes less than 30
+minutes to reach (round trip, see UNICEF-WHO, 2019). There is no reason
+why people in affluent countries should be expected to travel more for a
+basic need such as food. Accordingly, we adopt a threshold of 15 minutes
+for the analysis (representing a one-way trip).
 
@@
 
-The level of service by foodbank before and during the pandemic are
-shown in Figure \ref{fig:levels-of-service}. Since the level of service
-is the inverse of the number of households in the catchment area of the
-foodbanks, higher values mean that a facility is expected to service
-fewer households. Conversely, lower values indicate grater congestion.
+Using this threshold, we find that the system-wide accessibility
+(interpreted as a provider-to-population ratio) was 0.065 (food banks
+and food service locations per low income household in the city) before
+COVID-19, but declined to 0.037 during the pandemic. It is striking that
+although almost 76\% of food banks remained in operation during the
+pandemic, there was a loss of accessibility greater than 43\%,
+suggesting the location of food banks and food services plays an
+important role in serving those in need.
+
+Turning to the individual food bank and food service locations, the
+levels of service offered before and during the pandemic are shown in
+Figure \ref{fig:levels-of-service}. The level of service is functionally
+the inverse (because each food bank has a ``capacity'' of 1) of the
+number of low-income households in the travel-mode weighted travel time
+catchment area of the food banks. Higher values mean that a facility is
+expected to service fewer households. Conversely, lower values indicate
+grater congestion.
+
 The general pattern of the levels of service is similar before and
@@
 substantially. The pattern of the losses in level of service, moreover,
-is not uniform. As seen in Figure \ref{fig:levels-of-service-changes}
-\ldots{} words.
+is not uniform. The upper pane of Figure
+\ref{fig:levels-of-service-changes} shows that the three peripheral
+facilities in the southern suburban/exurban part of the city had low
+levels of service to begin with, but did not see major declines during
+the pandemic. Turning to the inset map, levels of service deteriorated
+more in the central part of the city. However, the loss of level of
+service was not as large in the core (where most of the food banks are),
+but more marked in the inner ring around the core where facilities may
+have faced greater demand from both central city and suburban
+populations after the closure of food banks during the pandemic.
 
@@
 \end{figure}
+
+To further elucidate this issue, we now turn to the results of the
+accessibility analysis. The top pane of Figure \ref{fig:accessibility}
+reveals that, compared with the outer rural zones, the smaller more
+urban zones of the city generally exhibit higher accessibility to food
+banks and food service locations. However, the pattern is not
+particularly smooth - this is largely attributable to the travel-mode
+weighting of travel times according to the trip patterns of low-income
+household respondents to the TTS. For example, in zones where low-income
+households make a high proportion of trips by walking, access to food
+bank locations by walking is afforded a concomitantly high weight in our
+calculations of travel time compared to transit or car travel. From
+this, highly-accessible locations result from a mix of characteristics:
+low-income households in locations where travel options that align with
+zonal modal split are available to connect them to food bank locations
+with high levels of service within 15 minutes.
 
@@
 \end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-local-i-1} \caption{\label{fig:plot-local-i}Areas with significantly large changes in accessibility from pre-COVID-19 to during COVID-19 according to Local Moran's I.}\label{fig:plot-local-i}
+\end{figure}
+
+Discuss the results of aggregating the accessibility by age group (see
+Table \ref{tab:accessibility-by-age}). Follow with the analysis of hours
+of travel?
 

### 2021-04-23 foodbanksHamilton/R/foodbanksHamilton.R

@@ -141,16 +141,16 @@ NULL
 
 #' Hamilton CMA.
 #'
-#' Boundary of the Hamilton CMA, in Ontario, Canada.
+#' Boundary of the City of Hamilton, in Ontario, Canada.
 #'
 #' @docType data
 #' @keywords datasets
 #' @keywords Census
 #' @keywords Canada
-#' @name hamilton_cma
-#' @usage data(hamilton_cma)
+#' @name hamilton_boundary
+#' @usage data(hamilton_boundary)
 #' @source Statistics Canada
-"hamilton_cma"
+"hamilton_boundary"
 
 #' Transportation for low income population by Traffic Analysis Zone.
 #'

### 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Revised the figures to use the correct boundary of the city

@@
 knitr::opts_chunk$set(
-  cache = TRUE, out.width = "100%", align = 'center', fig.width = 6, fig.height = 6
+  cache = TRUE, out.width = "100%", align = 'center', fig.width = 7, fig.height = 7
 )
@@
           color = NA) +
+  geom_sf(data = hamilton_boundary,
+          fill = NA) +
   geom_sf(data = inset_box,
@@
                        option = "A",
-                       direction = -1) +
+                       direction = -1,
+                       breaks = c(200, 400, 600)) +
   theme_minimal() +
@@
                        option = "A",
-                       direction = -1) +
+                       direction = -1,
+                       breaks = c(0.25, 0.50, 0.75)) +
   theme_minimal() +
@@
 acc_pre <- ggplot() +
-  #geom_sf(data = hamilton_boundary) +
+  geom_sf(data = hamilton_boundary,
+          fill = NA) +
   geom_sf(data = accessibility_taz %>%
@@
 acc_covid <- ggplot() +
-  #geom_sf(data = hamilton_boundary) +
+  geom_sf(data = hamilton_boundary,
+          fill = NA) +
   geom_sf(data = accessibility_taz %>%
@@
 
-```{r plot-accessibility-changes, echo=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19."}
+```{r plot-accessibility-changes, echo=FALSE, include=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19."}
 ## Analysis
@@
 
-```{r plot-local-i, echo=FALSE, fig.cap="\\label{fig:plot-local-i}Areas with significantly large changes in accessibility from pre-COVID-19 to during COVID-19 according to Local Moran's I." }
+```{r plot-local-i, echo=FALSE, fig.cap="\\label{fig:plot-local-i}Changes in accessibility from pre-COVID-19 to during COVID-19.
+Highlighted areas had significantly large changes in accessibility according to Local Moran's I." }
 ggplot() +
@@
                                   "-0.0002")) +
-  scale_color_manual(values = c("n.s." = NA, 
+  scale_color_manual(values = c("n.s." = "lightgray", 
                                 "sig" = "black")) +
@@
 ```
-
 ```{r accessibility-change-by-parcel, include=FALSE}

### 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Revised the figures to use the correct boundary of the city

@@
 \begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-accessibility-changes-1}
-
-}
-
-\caption{\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19.}\label{fig:plot-accessibility-changes}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-local-i-1} \caption{\label{fig:plot-local-i}Areas with significantly large changes in accessibility from pre-COVID-19 to during COVID-19 according to Local Moran's I.}\label{fig:plot-local-i}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-local-i-1} \caption{\label{fig:plot-local-i}Changes in accessibility from pre-COVID-19 to during COVID-19.
+Highlighted areas had significantly large changes in accessibility according to Local Moran's I.}\label{fig:plot-local-i}
 \end{figure}

### 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Started drafting discussion of the changes in accessibility

@@
 ---
-title: "Changes in accessibility to food banks during COVID-19 and implications for the food security of vulnerable populations in Hamilton, Ontario"
+title: "Changes in accessibility to food banks and food services during COVID-19 and implications for the food security of vulnerable populations in Hamilton, Ontario"
 author:
@@
 This condition has been associated with reductions in nutritional outcomes [@bhattacharya2004poverty; @kirkpatrick2008food; @olson1999nutrition] and physical and mental health in children and adults [@elgar2021relative; @jones2017food; @ramsey2011food; @seligman2010food; @stuff2004household].
-Over at least the past four decades food banks have become an essential line of defense against food insecurity in Canadian communities [@riches2002food; @holmes2018nothing; @black2020examining; @tarasuk2020relationship].
+Over at least the past four decades food banks and related services have become an essential line of defense against food insecurity in Canadian communities [@riches2002food; @holmes2018nothing; @black2020examining; @tarasuk2020relationship].
 In this respect, Canada is not unlike numerous other wealthy countries where a systematic dismantling of the welfare state took place in the intervening period [@tarasuk2014food].
@@
 A wealth of literature examining the food desert concept suggests that, in addition to socio-economic and demographic factors, location and transportation networks play a key role in a household's accessibility to healthy foods [@paez2010relative; @vanderlee2017food; @widener2018spatial].
-For food banks specifically, previous qualitative research by Smith et al.
-[-@smith2017food] in Ontario has noted that participants access food banks using a variety of modes of transportation and that some food bank users have expressed difficulties with transportation for trips to food bank locations, with one respondent stating "I wish it [the food bank] was a little more centrally located.
+For food banks specifically, previous qualitative research in Ontario by Smith-Carrier et al.
+[-@smith2017food] has noted that "transportation can be challenging, particularly if the food bank is situated in a remote location" (p.
+32).
+Particularly, it appears that participants experience challenges with the "inordinate amount of time necessary to obtain food, and difficulties associated with transportation" (p.
+39).
+Users of food banks, according to this research, rely on a variety of modes of transportation to access services.
+Consequently, the location of facilities matters; in the words of an interviewee: "I wish it [the food bank] was a little more centrally located.
 Because if I didn’t have a bike I’d have to walk it all the way out there and back.
-I wonder about people who don’t." From this, the authors conclude that "transportation can be challenging, particularly if the food bank is situated in a remote location." To offer greater insight into the role of transportation and location in food bank accessibility, this research examines how geographic accessibility to food banks and food services changed in Hamilton during the COVID-19 pandemic.
+I wonder about people who don’t" (p.
+39).
+
+To offer greater insight into the role of transportation and location in food bank accessibility, this research examines how geographic accessibility to food banks and food services changed in Hamilton during the COVID-19 pandemic.
 
@@
 $$
+## Study Area
+
+**Need to provide some context here: the geography which makes transportation challenging, i.e., due to the escarpment; the main parts of the city: city core, inner suburbs, industrial north, suburban, exurban; and the industrial background of the city maybe, with poverty rates compared to other Canadian cities?**
 
@@
 In addition, two free meal services held on different days at the same location were collapsed into a single service point for the accessibility analysis.
+Additional details on the operations of individual facilities is not publicly available, and with the changes in operations it proved unfeasible to collect it.
+For this reason, the analysis to follow is of accessibility to the location of food banks and services, but not to specific services (e.g., breakfasts vs.
+food boxes).
 
@@
 # Testing different thresholds to check how accessibility changes. Here, the thresholds are varied from 2 min to 30 min in 2 min step increments
-threshold_seq <- seq(from = 2, to = 30, by = 2)
+threshold_seq <- seq(from = 2, to = 30, by = 1)
 
@@
 
-Figure \ref{fig:foodbanks} shows the location of food banks in the City of Hamilton and their status.
+Figure \ref{fig:foodbanks} shows the location of food banks and services in the City of Hamilton and their status.
 Before the pandemic there were `r nrow(foodbank_locations)` of which `r nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))` (`r round(nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations) * 100, 2)`\%) closed during the pandemic.
-As shown in the figure, food banks tend to be predominantly located in the central parts of the city.
+As shown in the figure, food services tend to be predominantly located in the central parts of the city.
 This is not surprising: population density is high there, and it is also the part of the city where lower income households are more numerous in absolute and relative terms (see Figure \ref{fig:low-income-households}).
@@
 
-```{r plot-location-foodbanks, echo=FALSE, fig.cap="\\label{fig:foodbanks}Location of foodbanks and operation status; the dotted box is an inset of the central part of the City of Hamilton"}
+```{r plot-location-foodbanks, echo=FALSE, fig.cap="\\label{fig:foodbanks}Location of food banks/services and operation status; the dotted box is an inset of the central part of the City of Hamilton"}
 ## Analysis
@@
 When the expected travel time exceeds this threshold, a facility is no longer considered accessible.
-Further, the weights are standardized for the balanced floating catchment area approach.
+Moreover, the weights are standardized for the balanced floating catchment area approach.
 
 Figure \ref{fig:sensitivity-analysis} shows the results of conducting sensitivity analysis of the system-wide accessibility as we vary the threshold (considering the situation before the pandemic).
-There is a clear pattern whereby more strict values of $\delta$ are associated with higher levels of system-wide accessibility: while increases in accessibility that result from decreases in the travel time window might seem counter-intuitive, this is a result of lower _congestion_, since fewer households are serviced and are thus competing for the same resources.
+There is a clear pattern whereby more strict values of $\delta$ are associated with higher levels of system-wide accessibility: while increases in accessibility that result from decreases in the travel time window might seem counter-intuitive, this is a result of lower _congestion_, since fewer households are serviced and thus competition for the same resources is more limited.
 System-wide accessibility declines with higher values of $\delta$: as more households are serviced, congestion grows and the level of service declines, although this happens at a declining rate.
-We are not aware of any standards that explain how long people are expected to travel for food banks, but we note that in developing countries, access to drinking water is defined as a source that takes less than 30 minutes to reach [round trip, see @world2019progress].
+We are not aware of any research that explain how long people are expected to travel for food banks, but we note that in developing countries, access to drinking water is defined as a source that takes less than 30 minutes to reach [round trip, see @world2019progress].
 There is no reason why people in affluent countries should be expected to travel more for a basic need such as food.
-Accordingly, we adopt a threshold of 15 minutes for the analysis (representing a one-way trip).
+Accordingly, we adopt a 15-minute threshold for the analysis (representing a one-way trip).
+This threshold is also approximately where the rate of change in accessibility slows down.
 
@@
 
-Using this threshold, we find that the system-wide accessibility (interpreted as a provider-to-population ratio) was `r round(sum(results_pre$accessibility$accessibility), 3)` (food banks and food service locations per low income household in the city) before COVID-19, but declined to `r round(sum(results_covid$accessibility$accessibility), 3)` during the pandemic.
-It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`\% of food banks remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`\%, suggesting the location of food banks and food services plays an important role in serving those in need.
+Using the 15-minute threshold, we find that the system-wide accessibility (interpreted as a provider-to-population ratio) was `r round(sum(results_pre$accessibility$accessibility), 3)` (food banks/service locations per low income household in the city) before COVID-19, but declined to `r round(sum(results_covid$accessibility$accessibility), 3)` during the pandemic.
+It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`\% of food facilities remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`\%, suggesting the location of food banks and related services plays an important role in serving those in need.
 
@@
 
-Turning to the individual food bank and food service locations, the levels of service offered before and during the pandemic are shown in Figure \ref{fig:levels-of-service}.
-The level of service is functionally the inverse (because each food bank has a "capacity" of 1) of the number of low-income households in the travel-mode weighted travel time catchment area of the food banks.
+Turning to the location of individual facilities, the levels of service offered before and during the pandemic are shown in Figure \ref{fig:levels-of-service}.
+The level of service is functionally the inverse of the number of low-income households in the travel-mode weighted travel time catchment area of the facilities (this is because $S_j=1 \forall j$, i.e., each location represents a "capacity" of 1).
 Higher values mean that a facility is expected to service fewer households.
@@
 The general pattern of the levels of service is similar before and during the pandemic, with lower values in the center of the city.
-Three more peripheral facilities towards the south of the city also have lower levels of service, presumably because they are meant to service a relatively large suburban/exurban population.
+Three more peripheral facilities towards the south of the city also have lower levels of service, presumably because they are expected to service relatively large suburban/exurban populations.
 During the pandemic, however, the levels of service dropped, in some cases quite substantially.
@@
 The upper pane of Figure \ref{fig:levels-of-service-changes} shows that the three peripheral facilities in the southern suburban/exurban part of the city had low levels of service to begin with, but did not see major declines during the pandemic.
-Turning to the inset map, levels of service deteriorated more in the central part of the city.
-However, the loss of level of service was not as large in the core (where most of the food banks are), but more marked in the inner ring around the core where facilities may have faced greater demand from both central city and suburban populations after the closure of food banks during the pandemic.
+Further, the inset map shows that the levels of service deteriorated more in the central part of the city.
+However, the loss of level of service was not as large in the core (where most of the food banks/services are found), but instead was more marked in the inner ring around the core, where facilities may have faced greater demand from both central city and suburban populations after the closure of food banks during the pandemic.
 
@@
 To further elucidate this issue, we now turn to the results of the accessibility analysis.
-The top pane of Figure \ref{fig:accessibility} reveals that, compared with the outer rural zones, the smaller more urban zones of the city generally exhibit higher accessibility to food banks and food service locations.
-However, the pattern is not particularly smooth - this is largely attributable to the travel-mode weighting of travel times according to the trip patterns of low-income household respondents to the TTS.
+As with the level of service of individual facilities, the general pattern of accessibility before and during the pandemic is similar.
+Figure \ref{fig:accessibility} reveals that, compared with the outer rural zones, the more urban zones of the city generally exhibit higher accessibility to food banks and food service locations.
+However, the pattern is not particularly smooth - this is largely attributable to the weighting of travel times by mode of transportation according to the trip patterns of low-income household respondents according to the TTS.
 For example, in zones where low-income households make a high proportion of trips by walking, access to food bank locations by walking is afforded a concomitantly high weight in our calculations of travel time compared to transit or car travel.
 From this, highly-accessible locations result from a mix of characteristics: low-income households in locations where travel options that align with zonal modal split are available to connect them to food bank locations with high levels of service within 15 minutes.
+This seems to track with the experience of some users of these services, as reported by Smith-Carrier et al.
+[-@smith2017food].
 
@@
 
-```{r plot-accessibility-changes, echo=FALSE, include=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19."}
+```{r accessibility-changes-by-taz, include=FALSE}
 ## Analysis
 # Calculate accessibility change by TAZ:
-accessibilty_taz_change <- accessibility_taz %>%
+accessibility_taz_change <- accessibility_taz %>%
   st_drop_geometry() %>%
@@
   st_as_sf()
-
+```
+
+We find that the accessibility landscape deteriorated substantially during the pandemic, with accessibility dropping on average by almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% mean() %>% abs() %>% round()`\%, but with large variations: some zones experienced changes in accessibility of only about `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% max() %>% abs() %>% round()`\%, whereas the most affected zone saw a loss of accessibility of almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% min() %>% abs() %>% round()`\%.
+Figure \ref{fig:accessibility-changes-with-local-i} shows the changes in accessibility.
+Every zone is worse off after the closure of facilities during the pandemic, but some parts of the city seem to have been particularly affected.
+We used a local indicator of spatial autocorrelation [@anselin1995local] to explore the pattern of change in accessibility.
+Twenty-four TAZ are flagged as having significant large losses of accessibility (at $p<=0.10$, without correcting for multiple comparisons).
+Those zones are highlighted in the figure, where it can be seen that they form more or less compact neighborhoods.
+Remarkably, the largest (and significant) drops in accessibility are not downtown, but located in two cases in the industrial north of the city, in one case in an inner suburb above the escarpment (**need to describe the study area!**), and lastly in a more suburban/exurban region in the south-west (**share other insights about these regions, maybe based on weighted travel patterns?**).
+
+```{r plot-accessibility-changes, echo=FALSE, include=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19."}
+## Analysis
 # Plot
@@
   #geom_sf(data = hamilton_boundary) +
-  geom_sf(data = accessibilty_taz_change,
+  geom_sf(data = accessibility_taz_change,
           aes(fill = acc_d),
@@
 acc_change_inset <- ggplot() +
-  geom_sf(data = accessibilty_taz_change %>%
+  geom_sf(data = accessibility_taz_change %>%
             st_transform(crs = 4326),
@@
 ```
+
 ```{r weights-matrix, include=FALSE}
-taz_nb <- poly2nb(accessibilty_taz_change, queen = TRUE)
+taz_nb <- poly2nb(accessibility_taz_change, queen = TRUE)
 taz_w <- nb2listw(taz_nb)
@@
 ```{r moran-test-accessibility-changes, include=FALSE}
-moran.test(accessibilty_taz_change$acc_d, taz_w)
+moran.test(accessibility_taz_change$acc_d, taz_w)
 ```
@@
 ```{r local-moran-i, include=FALSE}
-local_i <- localmoran(accessibilty_taz_change$acc_d,
+local_i <- localmoran(accessibility_taz_change$acc_d,
                       taz_w, 
@@
   rename(p_val = Pr.z....0.) %>%
-  mutate(significance = ifelse(p_val <= 0.05, "sig", "n.s."),
+  mutate(significance = ifelse(p_val <= 0.10, "sig", "n.s."),
          significance = factor(significance,
@@
 ```{r join-local-i-to-table, include=FALSE}
-accessibilty_taz_change$significance <- local_i$significance
-```
-
-```{r plot-local-i, echo=FALSE, fig.cap="\\label{fig:plot-local-i}Changes in accessibility from pre-COVID-19 to during COVID-19.
+accessibility_taz_change$significance <- local_i$significance
+```
+
+```{r plot-local-i, echo=FALSE, fig.cap="\\label{fig:accessibility-changes-with-local-i}Changes in accessibility from pre-COVID-19 to during COVID-19.
 Highlighted areas had significantly large changes in accessibility according to Local Moran's I." }
 ggplot() +
-  geom_sf(data = accessibilty_taz_change,
+  geom_sf(data = accessibility_taz_change,
           aes(fill = acc_d,
@@
 ```
+
 ```{r accessibility-change-by-parcel, include=FALSE}
@@
         transmute(accessibility_covid = accessibility)) %>%
-  mutate(children = round(children) %>% 
+  mutate(pct_children = round(children/sum(children) * 100, 1),
+         pct_children = paste0("(", pct_children, "\\%)", sep = ""),
+         pct_adults = round(adults/sum(adults) * 100, 1),
+         pct_adults = paste0("(", pct_adults, "\\%)", sep = ""),
+         pct_seniors = round(seniors/sum(seniors) * 100, 1),
+         pct_seniors = paste0("(", pct_seniors, "\\%)", sep = ""),
+         children = round(children) %>% 
            prettyNum(big.mark = ","),
@@
            prettyNum(big.mark = ","),
+         children = paste0(children, " ", pct_children, sep = ""),
+         adults = paste0(adults, " ", pct_adults, sep = ""),
+         seniors = paste0(seniors, " ", pct_seniors, sep = ""),
          change = (accessibility_covid - accessibility_pre) %>% 
@@
          accessibility_covid = round(accessibility_covid, 5)) %>%
+  transmute(children,
+            adults,
+            seniors,
+            accessibility_pre,
+            accessibility_covid,
+            change) %>%
   kable("latex",
@@
 
+**LIMITATIONS**
+
 Words go here.

### 2021-04-24 00-Data-Processing/02-Routing-Transit.Rmd
Updated departure datetime for transit

@@
                           destinations = destinations_j,
-                          mode = c("TRANSIT"),
+                          mode = c("WALK", "TRANSIT"),
                           mode_egress = c("WALK"),
-                          departure_datetime = as.POSIXct(strptime("2021-04-05 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
+                          departure_datetime = as.POSIXct(strptime("2021-03-30 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
                           max_walk_dist = 10000, # metres

### 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex

@@
             pdfauthor={},
-            pdftitle={Changes in accessibility to food banks during COVID-19 and implications for the food security of vulnerable populations in Hamilton, Ontario},
+            pdftitle={Changes in accessibility to food banks and food services during COVID-19 and implications for the food security of vulnerable populations in Hamilton, Ontario},
             colorlinks=false,
@@
 
-  \title{Changes in accessibility to food banks during COVID-19 and
-implications for the food security of vulnerable populations in
-Hamilton, Ontario}
+  \title{Changes in accessibility to food banks and food services during
+COVID-19 and implications for the food security of vulnerable
+populations in Hamilton, Ontario}
     \author[Some University]{Author One\corref{1}}
@@
 2021; Jones, 2017; Ramsey et al., 2011; Seligman et al., 2010; Stuff et
-al., 2004). Over at least the past four decades food banks have become
-an essential line of defense against food insecurity in Canadian
-communities (Black and Seto, 2020; Holmes et al., 2018; Riches, 2002;
-Tarasuk et al., 2020). In this respect, Canada is not unlike numerous
-other wealthy countries where a systematic dismantling of the welfare
-state took place in the intervening period (Tarasuk et al., 2014).
+al., 2004). Over at least the past four decades food banks and related
+services have become an essential line of defense against food
+insecurity in Canadian communities (Black and Seto, 2020; Holmes et al.,
+2018; Riches, 2002; Tarasuk et al., 2020). In this respect, Canada is
+not unlike numerous other wealthy countries where a systematic
+dismantling of the welfare state took place in the intervening period
+(Tarasuk et al., 2014).
 
@@
 2017; Widener, 2018). For food banks specifically, previous qualitative
-research by Smith et al. (2017) in Ontario has noted that participants
-access food banks using a variety of modes of transportation and that
-some food bank users have expressed difficulties with transportation for
-trips to food bank locations, with one respondent stating ``I wish it
-{[}the food bank{]} was a little more centrally located. Because if I
-didn't have a bike I'd have to walk it all the way out there and back. I
-wonder about people who don't.'' From this, the authors conclude that
+research in Ontario by Smith-Carrier et al. (2017) has noted that
 ``transportation can be challenging, particularly if the food bank is
-situated in a remote location.'' To offer greater insight into the role
-of transportation and location in food bank accessibility, this research
-examines how geographic accessibility to food banks and food services
-changed in Hamilton during the COVID-19 pandemic.
+situated in a remote location'' (p.~32). Particularly, it appears that
+participants experience challenges with the ``inordinate amount of time
+necessary to obtain food, and difficulties associated with
+transportation'' (p.~39). Users of food banks, according to this
+research, rely on a variety of modes of transportation to access
+services. Consequently, the location of facilities matters; in the words
+of an interviewee: ``I wish it {[}the food bank{]} was a little more
+centrally located. Because if I didn't have a bike I'd have to walk it
+all the way out there and back. I wonder about people who don't''
+(p.~39).
+
+To offer greater insight into the role of transportation and location in
+food bank accessibility, this research examines how geographic
+accessibility to food banks and food services changed in Hamilton during
+the COVID-19 pandemic.
 
@@
 \sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
-\]
+\] \#\# Study Area
+
+\textbf{Need to provide some context here: the geography which makes
+transportation challenging, i.e., due to the escarpment; the main parts
+of the city: city core, inner suburbs, industrial north, suburban,
+exurban; and the industrial background of the city maybe, with poverty
+rates compared to other Canadian cities?}
 
@@
 location were collapsed into a single service point for the
-accessibility analysis.
+accessibility analysis. Additional details on the operations of
+individual facilities is not publicly available, and with the changes in
+operations it proved unfeasible to collect it. For this reason, the
+analysis to follow is of accessibility to the location of food banks and
+services, but not to specific services (e.g., breakfasts vs.~food
+boxes).
 
@@
 
-Figure \ref{fig:foodbanks} shows the location of food banks in the City
-of Hamilton and their status. Before the pandemic there were 58 of which
-14 (24.14\%) closed during the pandemic. As shown in the figure, food
-banks tend to be predominantly located in the central parts of the city.
-This is not surprising: population density is high there, and it is also
-the part of the city where lower income households are more numerous in
-absolute and relative terms (see Figure
+Figure \ref{fig:foodbanks} shows the location of food banks and services
+in the City of Hamilton and their status. Before the pandemic there were
+58 of which 14 (24.14\%) closed during the pandemic. As shown in the
+figure, food services tend to be predominantly located in the central
+parts of the city. This is not surprising: population density is high
+there, and it is also the part of the city where lower income households
+are more numerous in absolute and relative terms (see Figure
 \ref{fig:low-income-households}). Alas, this is also the part of the
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-location-foodbanks-1} \caption{\label{fig:foodbanks}Location of foodbanks and operation status; the dotted box is an inset of the central part of the City of Hamilton}\label{fig:plot-location-foodbanks}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-location-foodbanks-1} \caption{\label{fig:foodbanks}Location of food banks/services and operation status; the dotted box is an inset of the central part of the City of Hamilton}\label{fig:plot-location-foodbanks}
 \end{figure}
@@
 expected travel time exceeds this threshold, a facility is no longer
-considered accessible. Further, the weights are standardized for the
+considered accessible. Moreover, the weights are standardized for the
 balanced floating catchment area approach.
@@
 seem counter-intuitive, this is a result of lower \emph{congestion},
-since fewer households are serviced and are thus competing for the same
-resources. System-wide accessibility declines with higher values of
-\(\delta\): as more households are serviced, congestion grows and the
-level of service declines, although this happens at a declining rate. We
-are not aware of any standards that explain how long people are expected
-to travel for food banks, but we note that in developing countries,
-access to drinking water is defined as a source that takes less than 30
-minutes to reach (round trip, see UNICEF-WHO, 2019). There is no reason
-why people in affluent countries should be expected to travel more for a
-basic need such as food. Accordingly, we adopt a threshold of 15 minutes
-for the analysis (representing a one-way trip).
+since fewer households are serviced and thus competition for the same
+resources is more limited. System-wide accessibility declines with
+higher values of \(\delta\): as more households are serviced, congestion
+grows and the level of service declines, although this happens at a
+declining rate. We are not aware of any research that explain how long
+people are expected to travel for food banks, but we note that in
+developing countries, access to drinking water is defined as a source
+that takes less than 30 minutes to reach (round trip, see UNICEF-WHO,
+2019). There is no reason why people in affluent countries should be
+expected to travel more for a basic need such as food. Accordingly, we
+adopt a 15-minute threshold for the analysis (representing a one-way
+trip). This threshold is also approximately where the rate of change in
+accessibility slows down.
 
@@
 
-Using this threshold, we find that the system-wide accessibility
-(interpreted as a provider-to-population ratio) was 0.065 (food banks
-and food service locations per low income household in the city) before
-COVID-19, but declined to 0.037 during the pandemic. It is striking that
-although almost 76\% of food banks remained in operation during the
-pandemic, there was a loss of accessibility greater than 43\%,
-suggesting the location of food banks and food services plays an
-important role in serving those in need.
-
-Turning to the individual food bank and food service locations, the
-levels of service offered before and during the pandemic are shown in
-Figure \ref{fig:levels-of-service}. The level of service is functionally
-the inverse (because each food bank has a ``capacity'' of 1) of the
-number of low-income households in the travel-mode weighted travel time
-catchment area of the food banks. Higher values mean that a facility is
-expected to service fewer households. Conversely, lower values indicate
-grater congestion.
+Using the 15-minute threshold, we find that the system-wide
+accessibility (interpreted as a provider-to-population ratio) was 0.065
+(food banks/service locations per low income household in the city)
+before COVID-19, but declined to 0.037 during the pandemic. It is
+striking that although almost 76\% of food facilities remained in
+operation during the pandemic, there was a loss of accessibility greater
+than 43\%, suggesting the location of food banks and related services
+plays an important role in serving those in need.
+
+Turning to the location of individual facilities, the levels of service
+offered before and during the pandemic are shown in Figure
+\ref{fig:levels-of-service}. The level of service is functionally the
+inverse of the number of low-income households in the travel-mode
+weighted travel time catchment area of the facilities (this is because
+\(S_j=1 \forall j\), i.e., each location represents a ``capacity'' of
+1). Higher values mean that a facility is expected to service fewer
+households. Conversely, lower values indicate grater congestion.
 
@@
 more peripheral facilities towards the south of the city also have lower
-levels of service, presumably because they are meant to service a
-relatively large suburban/exurban population. During the pandemic,
+levels of service, presumably because they are expected to service
+relatively large suburban/exurban populations. During the pandemic,
 however, the levels of service dropped, in some cases quite
@@
 levels of service to begin with, but did not see major declines during
-the pandemic. Turning to the inset map, levels of service deteriorated
-more in the central part of the city. However, the loss of level of
-service was not as large in the core (where most of the food banks are),
-but more marked in the inner ring around the core where facilities may
-have faced greater demand from both central city and suburban
-populations after the closure of food banks during the pandemic.
+the pandemic. Further, the inset map shows that the levels of service
+deteriorated more in the central part of the city. However, the loss of
+level of service was not as large in the core (where most of the food
+banks/services are found), but instead was more marked in the inner ring
+around the core, where facilities may have faced greater demand from
+both central city and suburban populations after the closure of food
+banks during the pandemic.
 
@@
 To further elucidate this issue, we now turn to the results of the
-accessibility analysis. The top pane of Figure \ref{fig:accessibility}
-reveals that, compared with the outer rural zones, the smaller more
-urban zones of the city generally exhibit higher accessibility to food
-banks and food service locations. However, the pattern is not
-particularly smooth - this is largely attributable to the travel-mode
-weighting of travel times according to the trip patterns of low-income
-household respondents to the TTS. For example, in zones where low-income
+accessibility analysis. As with the level of service of individual
+facilities, the general pattern of accessibility before and during the
+pandemic is similar. Figure \ref{fig:accessibility} reveals that,
+compared with the outer rural zones, the more urban zones of the city
+generally exhibit higher accessibility to food banks and food service
+locations. However, the pattern is not particularly smooth - this is
+largely attributable to the weighting of travel times by mode of
+transportation according to the trip patterns of low-income household
+respondents according to the TTS. For example, in zones where low-income
 households make a high proportion of trips by walking, access to food
@@
 zonal modal split are available to connect them to food bank locations
-with high levels of service within 15 minutes.
+with high levels of service within 15 minutes. This seems to track with
+the experience of some users of these services, as reported by
+Smith-Carrier et al. (2017).
 
@@
 
+We find that the accessibility landscape deteriorated substantially
+during the pandemic, with accessibility dropping on average by almost
+41\%, but with large variations: some zones experienced changes in
+accessibility of only about 10\%, whereas the most affected zone saw a
+loss of accessibility of almost 71\%. Figure
+\ref{fig:accessibility-changes-with-local-i} shows the changes in
+accessibility. Every zone is worse off after the closure of facilities
+during the pandemic, but some parts of the city seem to have been
+particularly affected. We used a local indicator of spatial
+autocorrelation (Anselin, 1995) to explore the pattern of change in
+accessibility. Twenty-four TAZ are flagged as having significant large
+losses of accessibility (at \(p<=0.10\), without correcting for multiple
+comparisons). Those zones are highlighted in the figure, where it can be
+seen that they form more or less compact neighborhoods. Remarkably, the
+largest (and significant) drops in accessibility are not downtown, but
+located in two cases in the industrial north of the city, in one case in
+an inner suburb above the escarpment (\textbf{need to describe the study
+area!}), and lastly in a more suburban/exurban region in the south-west
+(\textbf{share other insights about these regions, maybe based on
+weighted travel patterns?}).
+
 \begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-local-i-1} \caption{\label{fig:plot-local-i}Changes in accessibility from pre-COVID-19 to during COVID-19.
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-local-i-1} \caption{\label{fig:accessibility-changes-with-local-i}Changes in accessibility from pre-COVID-19 to during COVID-19.
 Highlighted areas had significantly large changes in accessibility according to Local Moran's I.}\label{fig:plot-local-i}
@@
 \midrule
-\cellcolor{gray!6}{3,445} & \cellcolor{gray!6}{11,860} & \cellcolor{gray!6}{6,584} & \cellcolor{gray!6}{0.00229} & \cellcolor{gray!6}{0.00173} & \cellcolor{gray!6}{-0.00057}\\
-4,477 & 12,838 & 6,847 & 0.00707 & 0.00418 & -0.00289\\
-\cellcolor{gray!6}{3,541} & \cellcolor{gray!6}{13,189} & \cellcolor{gray!6}{7,020} & \cellcolor{gray!6}{0.01580} & \cellcolor{gray!6}{0.00911} & \cellcolor{gray!6}{-0.00669}\\
-2,340 & 9,918 & 4,940 & 0.03985 & 0.02202 & -0.01783\\
+\cellcolor{gray!6}{3,445 (25\%)} & \cellcolor{gray!6}{11,860 (24.8\%)} & \cellcolor{gray!6}{6,584 (25.9\%)} & \cellcolor{gray!6}{0.00229} & \cellcolor{gray!6}{0.00173} & \cellcolor{gray!6}{-0.00057}\\
+4,477 (32.4\%) & 12,838 (26.9\%) & 6,847 (27\%) & 0.00707 & 0.00418 & -0.00289\\
+\cellcolor{gray!6}{3,541 (25.7\%)} & \cellcolor{gray!6}{13,189 (27.6\%)} & \cellcolor{gray!6}{7,020 (27.6\%)} & \cellcolor{gray!6}{0.01580} & \cellcolor{gray!6}{0.00911} & \cellcolor{gray!6}{-0.00669}\\
+2,340 (17\%) & 9,918 (20.7\%) & 4,940 (19.5\%) & 0.03985 & 0.02202 & -0.01783\\
 \bottomrule
@@
 
+\textbf{LIMITATIONS}
+
 Words go here.
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-anselin1995local}{}%
+Anselin, L., 1995. Local indicators of spatial association - LISA.
+Geographical Analysis 27, 93--115.
+
 \leavevmode\hypertarget{ref-bazerghi2016role}{}%

### 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Add text to accessibility change interpretation

@@
 For routing purposes we used maximum values of 180 min and 10,000 m walking distance: any destination that exceeded these thresholds was ignored.
-The departure time used for routing was **TIME**.
+The departure time used for routing was 8:00AM on March 30, 2021 to reflect transit service around the morning service peak on a typical Tuesday.
 
@@
 System-wide accessibility declines with higher values of $\delta$: as more households are serviced, congestion grows and the level of service declines, although this happens at a declining rate.
-We are not aware of any research that explain how long people are expected to travel for food banks, but we note that in developing countries, access to drinking water is defined as a source that takes less than 30 minutes to reach [round trip, see @world2019progress].
+We are not aware of any research that explains how long people are expected to travel for food banks, but we note that in developing countries, access to drinking water is defined as a source that takes less than 30 minutes to reach [round trip, see @world2019progress].
 There is no reason why people in affluent countries should be expected to travel more for a basic need such as food.
@@
 Higher values mean that a facility is expected to service fewer households.
-Conversely, lower values indicate grater congestion.
-
-The general pattern of the levels of service is similar before and during the pandemic, with lower values in the center of the city.
-Three more peripheral facilities towards the south of the city also have lower levels of service, presumably because they are expected to service relatively large suburban/exurban populations.
+Conversely, lower values indicate greater congestion.
+
+The general pattern of the levels of service is similar before and during the pandemic, with lower values in the center of the city where low-income households exhibit multimodal trip patterns.
+Three more peripheral facilities towards the south of the city also have lower levels of service, presumably because they are expected to service relatively large suburban/exurban populations generally reliant on automobiles for travel.
 During the pandemic, however, the levels of service dropped, in some cases quite substantially.
@@
 Figure \ref{fig:accessibility} reveals that, compared with the outer rural zones, the more urban zones of the city generally exhibit higher accessibility to food banks and food service locations.
-However, the pattern is not particularly smooth - this is largely attributable to the weighting of travel times by mode of transportation according to the trip patterns of low-income household respondents according to the TTS.
+However, the pattern is not particularly smooth - this is largely attributable to the weighting of travel times by mode of transportation according to the trip patterns of low-income household respondents captured by the TTS.
 For example, in zones where low-income households make a high proportion of trips by walking, access to food bank locations by walking is afforded a concomitantly high weight in our calculations of travel time compared to transit or car travel.
@@
 Every zone is worse off after the closure of facilities during the pandemic, but some parts of the city seem to have been particularly affected.
-We used a local indicator of spatial autocorrelation [@anselin1995local] to explore the pattern of change in accessibility.
+To better highlight these changes, we used a local indicator of spatial autocorrelation [@anselin1995local] to explore the pattern of change in accessibility.
 Twenty-four TAZ are flagged as having significant large losses of accessibility (at $p<=0.10$, without correcting for multiple comparisons).
 Those zones are highlighted in the figure, where it can be seen that they form more or less compact neighborhoods.
-Remarkably, the largest (and significant) drops in accessibility are not downtown, but located in two cases in the industrial north of the city, in one case in an inner suburb above the escarpment (**need to describe the study area!**), and lastly in a more suburban/exurban region in the south-west (**share other insights about these regions, maybe based on weighted travel patterns?**).
+Remarkably, the largest (and significant) drops in accessibility are not downtown, but located in two cases in the industrial north of the city, in one case in an inner suburb above the escarpment, and lastly in a more suburban/exurban region in the south-west (**share other insights about these regions, maybe based on weighted travel patterns?**).
 
@@
 ```
+
+For the more suburban clusters of zones, the decrease in accessibility is derived from the closure of locations throughout the city reachable by car.
+In the cluster of central suburban zones for example, low-income households in the outer ring of zones that exhibit medium to high decreases in accessibility within this cluster appear to be largely auto-dependent in their tripmaking, which each each exhibiting between `r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(min(p_car)) *100)`-`r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(max(p_car)) *100)`\% of their modal split for car trips.
+This results in the parcels within these zones having a large number of potentially accessible locations in the travel time matrix.
+But by extension, the change in accessibility over the pre- and during-COVID-19 time periods is affected not only by the closure of service locations proximate to the zones, but also the locations in the central city.
+The zone with the greatest decrease in accessibility within this cluster (`r round(accessibility_taz_change %>% st_drop_geometry() %>% filter(TAZUID == 5135) %>% pull(acc_d), 4)`) has a high rate of car trips and connects to the most food bank locations in total as well as those that stayed open or closed.
+The exception to this is the zone in the middle of this cluster, which has a much higher proportion of transit use `r round(modes %>% st_drop_geometry() %>% filter(TAZUID == 5116) %>% summarize(first(p_transit)) *100)`\%.
+This modal split weighting results in fewer food bank and service locations accessible from the parcels within this zone which, by extension, limits their access to closed locations reachable within 15 minutes in our calculations.
+
+In the cluster to the south-west, the decrease in accessibility is predominately driven by the closure of a high level-of-service Community Meals provider.
+However, like the more central suburban zones, low-income households within this cluster also between 90% to 100% auto-dependent in their tripmaking in the TTS.
+The story is similar for the zone located in the north-west that exhibits the greatest decrease in accessibility.
+Here, low-income households responding to the TTS conducted 100% of their trips by car and, as a result, parcels within this zone have access to the second-highest number of food bank and service locations within 15 minutes.
+However, this also means zonal accessibilities are greatly affected by the number of closures throughout the city.
+Finally, in the city's north end and north-east zones, low-income households exhibit a mixture of tripmaking behaviour in the TTS.
+Households in some zones take transit more often and one zone in particular has 100\% of its trips by walking while in others, car use makes up 100\% of trips.
+For these zones, the decrease in accessibility tends to be a product of the closure of several inner-city food bank and service locations.
 

### 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Updated data package with revised travel time table for transit; revised analysis by population group; first draft of the abstract

@@
 ---
-title: "Changes in accessibility to food banks and food services during COVID-19 and implications for the food security of vulnerable populations in Hamilton, Ontario"
+title: "Changes in accessibility to food banks and food services during COVID-19 and implications for low income populations in Hamilton, Ontario"
 author:
@@
     affiliation: Some University
-    footnote: 1
   - name: Author Two
     email: author2@example.com
-    affiliation: 
+    affiliation: Another University
+    footnote: 1
   - name: Author Three
@@
     affiliation: Some University
-    footnote: 2 
   - name: Author Four
     email: author4@example.com
-    affiliation: Another University
-    footnote: 2
+    affiliation: Some University
 address:
@@
     text: "Corresponding Author"
-  - code: 2
-    text: "Equal contribution"
 abstract: |
-  This is the abstract.
-
-  It consists of two paragraphs.
-
-journal: "An awesome journal"
+  In this paper we analyze the changes in accessibility to food banks and related services before and during the COVID-19 pandemic in the City of Hamilton, Ontario.
+Food banks and services are the last line of support for households facing food insecurity; as such, they have become essential and their relevance cannot be ignored in the midst of the economic upheaval caused by the pandemic.
+Our analysis is based on the application of balanced floating catchment areas, and concentrates on households with lower incomes ($<CAD40,000, approximately the Low Income Cutoff Value for a city of Hamilton's size).
+Accessibility was low to begin with in suburban and exurban parts of the city.
+We find that about 14\% of locations originally available in Hamilton closed during the pandemic; the impact of those closures on the level of service and accessibility was disproportionate, with system-wide losses exceeding 40\%.
+Those losses were geographically and demographically uneven.
+While every part of the city faced a reduction in accessibility, inner suburbs fared worse in terms of loss of accessibility.
+As well, children (age $\le 18$) appear to have been impacted the most.
+
+journal: "Social Science and Medicine"
 date: "`r Sys.Date()`"
@@
 
-**Author 1:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization, Supervision; **Author 2:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization; **Author 3:** Conceptualization, Investigation, Data Curation, Writing - Original Draft; **Author 4:** Resources - Writing: Review & Editing - Supervision
+**Author 1:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization, Supervision; **Author 2:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization; **Author 3:** Conceptualization, Investigation, Data Curation, Writing - Original Draft; **Author 4:** Resources, Writing - Original Draft, Supervision
 
@@
 Those zones are highlighted in the figure, where it can be seen that they form more or less compact neighborhoods.
-Remarkably, the largest (and significant) drops in accessibility are not downtown, but located in two cases in the industrial north of the city, in one case in an inner suburb above the escarpment, and lastly in a more suburban/exurban region in the south-west (**share other insights about these regions, maybe based on weighted travel patterns?**).
+Remarkably, the largest (and significant) drops in accessibility are not downtown, but located in two cases in the industrial north of the city, in one case in an inner suburb above the escarpment, and lastly in a more suburban/exurban region in the south-west.
 
@@
 
+Just as the effects of the closures appear to have been uneven in space, they also seem to have differently impacted various population segments.
+Table \ref{tab:accessibility-by-age} shows the estimated number of people in each age group by their level of accessibility before and during the pandemic.
+Here, it is important to note that the quartiles are relative: people in the top 25\% of accessibility still have lower accessibility during the pandemic than before.
+In reality, every population group is worse off during the pandemic in terms of their accessibility to food banks and services in the City of Hamilton.
+However, some age groups were affected more.
+As seen in the table, the distribution within quartiles for adults changed only slightly, and in fact there are fewer adults in the bottom three quartiles as some found themselves in the (deteriorated) top quartile of accessibility during the pandemic.
+The story is different for seniors, a greater number of whom are now in the third quartile, due to a combination of seniors facing a worse accessibility situation, and just a few moving from the bottom quartile (pre-pandemic) to the third during the pandemic.
+Those aged 18 and less appear to have fared worse, and the largest change is in the number of children who were in the third quartile before the pandemic found themselves in the bottom accessibility class during the pandemic.
+These results suggest that, through a combination of the typical modes of transportation of lower income households, and the spatial distribution of the population, the food bank/service locations closed had a differential impact that more greatly affected the youngest among the population in low income households.
+
 ```{r accessibility-change-by-parcel, include=FALSE}
@@
 ## Analysis
+# Quartiles of accessibility pre-covid:
+acc_quantiles_pre <- quantile(accessibility_change$`Pre-covid`,
+                              c(0, 0.25, 0.5, 0.75, 1))
 # Quartiles of accessibility change:
-acc_quintiles_change <- quantile(accessibility_change$acc_d,
+acc_quantiles_during <- quantile(accessibility_change$`During covid`,
                                  c(0, 0.25, 0.5, 0.75, 1))
@@
 
-```{r accessibility-change-labels, include=FALSE}
+```{r accessibility-labels, include=FALSE}
 ## Analysis
@@
 accessibility <- accessibility %>%
-  mutate(accessibility_level = case_when(acc_d <= acc_quintiles_change[2] ~ "Bottom 25%",
-                                         acc_d > acc_quintiles_change[2] & acc_d <= acc_quintiles_change[3] ~ "Third 25%",
-                                         acc_d > acc_quintiles_change[3] & acc_d <= acc_quintiles_change[4] ~ "Second 25%",
-                                         acc_d > acc_quintiles_change[4] ~ "Top 25%"),
-           accessibility_level = factor(accessibility_level,
-                                        levels = c("Top 25%", "Second 25%", "Third 25%", "Bottom 25%"),
+  mutate(accessibility_pre_level = case_when(accessibility <= acc_quantiles_pre[2] ~ "Bottom 25\\%",
+                                             accessibility > acc_quantiles_pre[2] & accessibility <= acc_quantiles_pre[3] ~ "Third 25\\%",
+                                             accessibility > acc_quantiles_pre[3] & accessibility <= acc_quantiles_pre[4] ~ "Second 25\\%",
+                                             accessibility > acc_quantiles_pre[4] ~ "Top 25\\%"),
+         accessibility_pre_level = factor(accessibility_pre_level,
+                                          levels = c("Top 25\\%", "Second 25\\%", "Third 25\\%", "Bottom 25\\%"),
+                                          ordered = TRUE),
+         accessibility_during_level = case_when(accessibility <= acc_quantiles_during[2] ~ "Bottom 25\\%",
+                                                accessibility > acc_quantiles_during[2] & accessibility <= acc_quantiles_during[3] ~ "Third 25\\%",
+                                                accessibility > acc_quantiles_during[3] & accessibility <= acc_quantiles_during[4] ~ "Second 25\\%",
+                                                accessibility > acc_quantiles_during[4] ~ "Top 25\\%"),
+         accessibility_during_level = factor(accessibility_during_level,
+                                             levels = c("Top 25\\%", "Second 25\\%", "Third 25\\%", "Bottom 25\\%"),
                                              ordered = TRUE))
@@
 ```{r aggregate-accessibility-by-age-group, include=FALSE}
+summary_pre <- accessibility %>%
+  filter(timing == "Pre-covid") %>%
+  group_by(accessibility_pre_level) %>%
+  summarize(children_pre = sum(children) %>% 
+              round(),
+            adults_pre = sum(adults) %>% 
+              round(),
+            seniors_pre = sum(seniors) %>% 
+              round(),
+            .groups = "drop")
+
+summary_during <- accessibility %>%
+  filter(timing == "During covid") %>%
+  group_by(accessibility_during_level) %>%
+  summarize(children_during = sum(children) %>% 
+              round(),
+            adults_during = sum(adults) %>% 
+              round(),
+            seniors_during = sum(seniors) %>% 
+              round(),
+            .groups_during = "drop")
+
+accessibility_by_age <- cbind(summary_pre, summary_during[,2:4]) %>%
+  mutate(children_change = round((children_during - children_pre)/children_pre * 100, 2),
+         adults_change = round((adults_during - adults_pre)/adults_pre * 100, 2),
+         seniors_change = round((seniors_during - seniors_pre)/seniors_pre * 100, 2),
+         children_pre = prettyNum(children_pre, big.mark = ","),
+         adults_pre = prettyNum(adults_pre, big.mark = ","),
+         seniors_pre = prettyNum(seniors_pre, big.mark = ","),
+         children_during = paste0(prettyNum(children_during, big.mark = ","), " (", children_change, "\\%)"),
+         adults_during = paste0(prettyNum(adults_during, big.mark = ","), " (", adults_change, "\\%)"),
+         seniors_during = paste0(prettyNum(seniors_during, big.mark = ","), " (", seniors_change, "\\%)")) %>%
+  transmute(accessibility_pre_level,
+            children_pre,
+            adults_pre,
+            seniors_pre,
+            children_during,
+            adults_during,
+            seniors_during)
+```
+
+```{r table-accessibility-by-age-group, echo=FALSE, results='asis'}
+accessibility_by_age %>%
+  kable("latex",
+        booktabs = TRUE,
+        align = ("lcccccc"),
+        col.names = c("Accessibility Quartile",
+                      "Children (age $\\le$ 18)", 
+                      "Adults (19-64)", 
+                      "Seniors (age $\\ge$ 65)", 
+                      "Children (age $\\le$ 18)", 
+                      "Adults (19-64)", 
+                      "Seniors (age $\\ge$ 65)"),
+        caption = "\\label{tab:accessibility-by-age}Population at each accessibility level by age group among members households with incomes less than CAD40,000.",
+        escape = FALSE) %>%
+  kable_styling(latex_options = c("striped", "scale_down", "hold_position")) %>%
+  add_header_above(c(" " = 1, "Pre-COVID-19" = 3, "During COVID-19" = 3)) %>%
+  footnote(general = c("Population values have been rounded.", 
+                       "The values in brackets for population during COVID-19 are the changes from before the pandemic."))
+```
+
+```{r include=FALSE, eval = FALSE}
 ## Analysis
@@
 accessibility_by_age <- accessibility %>%
+  filter(timing = "During covid")
 group_by(timing, 
@@
 
-Discuss the results of aggregating the accessibility by age group (see Table \ref{tab:accessibility-by-age}). Follow with the analysis of hours of travel?
-
-```{r table-accessibility-by-age-group, echo=FALSE, results='asis'}
+```{r include=FALSE, eval = FALSE}
 cbind(accessibility_by_age %>% 

### 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Updated data package with revised travel time table for transit; revised analysis by population group; first draft of the abstract

@@
 
-  \journal{An awesome journal} % Sets Journal name
+  \journal{Social Science and Medicine} % Sets Journal name
 
@@
             pdfauthor={},
-            pdftitle={Changes in accessibility to food banks and food services during COVID-19 and implications for the food security of vulnerable populations in Hamilton, Ontario},
+            pdftitle={Changes in accessibility to food banks and food services during COVID-19 and implications for low income populations in Hamilton, Ontario},
             colorlinks=false,
@@
   \title{Changes in accessibility to food banks and food services during
-COVID-19 and implications for the food security of vulnerable
-populations in Hamilton, Ontario}
-    \author[Some University]{Author One\corref{1}}
+COVID-19 and implications for low income populations in Hamilton,
+Ontario}
+    \author[Some University]{Author One}
    \ead{author1@example.com} 
-    \author[]{Author Two}
+    \author[Another University]{Author Two\corref{1}}
    \ead{author2@example.com} 
-    \author[Some University]{Author Three\corref{2}}
+    \author[Some University]{Author Three}
    \ead{author3@example.com} 
-    \author[Another University]{Author Four\corref{2}}
+    \author[Some University]{Author Four}
    \ead{author4@example.com} 
@@
       \cortext[1]{Corresponding Author}
-    \cortext[2]{Equal contribution}
 
   \begin{abstract}
-  This is the abstract.
-
-  It consists of two paragraphs.
+  In this paper we analyze the changes in accessibility to food banks
+  and related services before and during the COVID-19 pandemic in the
+  City of Hamilton, Ontario. Food banks and services are the last line
+  of support for households facing food insecurity; as such, they have
+  become essential and their relevance cannot be ignored in the midst of
+  the economic upheaval caused by the pandemic. Our analysis is based on
+  the application of balanced floating catchment areas, and concentrates
+  on households with lower incomes (\$\textless CAD40,000, approximately
+  the Low Income Cutoff Value for a city of Hamilton's size).
+  Accessibility was low to begin with in suburban and exurban parts of
+  the city. We find that about 14\% of locations originally available in
+  Hamilton closed during the pandemic; the impact of those closures on
+  the level of service and accessibility was disproportionate, with
+  system-wide losses exceeding 40\%. Those losses were geographically
+  and demographically uneven. While every part of the city faced a
+  reduction in accessibility, inner suburbs fared worse in terms of loss
+  of accessibility. As well, children (age \(\le 18\)) appear to have
+  been impacted the most.
   \end{abstract}
@@
 Conceptualization, Investigation, Data Curation, Writing - Original
-Draft; \textbf{Author 4:} Resources - Writing: Review \& Editing -
+Draft; \textbf{Author 4:} Resources, Writing - Original Draft,
 Supervision
@@
 that exceeded these thresholds was ignored. The departure time used for
-routing was \textbf{TIME}.
+routing was 8:00AM on March 30, 2021 to reflect transit service around
+the morning service peak on a typical Tuesday.
 
@@
 grows and the level of service declines, although this happens at a
-declining rate. We are not aware of any research that explain how long
+declining rate. We are not aware of any research that explains how long
 people are expected to travel for food banks, but we note that in
@@
 1). Higher values mean that a facility is expected to service fewer
-households. Conversely, lower values indicate grater congestion.
+households. Conversely, lower values indicate greater congestion.
 
 The general pattern of the levels of service is similar before and
-during the pandemic, with lower values in the center of the city. Three
-more peripheral facilities towards the south of the city also have lower
+during the pandemic, with lower values in the center of the city where
+low-income households exhibit multimodal trip patterns. Three more
+peripheral facilities towards the south of the city also have lower
 levels of service, presumably because they are expected to service
-relatively large suburban/exurban populations. During the pandemic,
-however, the levels of service dropped, in some cases quite
-substantially. The pattern of the losses in level of service, moreover,
-is not uniform. The upper pane of Figure
-\ref{fig:levels-of-service-changes} shows that the three peripheral
-facilities in the southern suburban/exurban part of the city had low
-levels of service to begin with, but did not see major declines during
-the pandemic. Further, the inset map shows that the levels of service
-deteriorated more in the central part of the city. However, the loss of
-level of service was not as large in the core (where most of the food
-banks/services are found), but instead was more marked in the inner ring
-around the core, where facilities may have faced greater demand from
-both central city and suburban populations after the closure of food
-banks during the pandemic.
+relatively large suburban/exurban populations generally reliant on
+automobiles for travel. During the pandemic, however, the levels of
+service dropped, in some cases quite substantially. The pattern of the
+losses in level of service, moreover, is not uniform. The upper pane of
+Figure \ref{fig:levels-of-service-changes} shows that the three
+peripheral facilities in the southern suburban/exurban part of the city
+had low levels of service to begin with, but did not see major declines
+during the pandemic. Further, the inset map shows that the levels of
+service deteriorated more in the central part of the city. However, the
+loss of level of service was not as large in the core (where most of the
+food banks/services are found), but instead was more marked in the inner
+ring around the core, where facilities may have faced greater demand
+from both central city and suburban populations after the closure of
+food banks during the pandemic.
 
@@
 transportation according to the trip patterns of low-income household
-respondents according to the TTS. For example, in zones where low-income
+respondents captured by the TTS. For example, in zones where low-income
 households make a high proportion of trips by walking, access to food
@@
 during the pandemic, but some parts of the city seem to have been
-particularly affected. We used a local indicator of spatial
-autocorrelation (Anselin, 1995) to explore the pattern of change in
-accessibility. Twenty-four TAZ are flagged as having significant large
-losses of accessibility (at \(p<=0.10\), without correcting for multiple
-comparisons). Those zones are highlighted in the figure, where it can be
-seen that they form more or less compact neighborhoods. Remarkably, the
-largest (and significant) drops in accessibility are not downtown, but
-located in two cases in the industrial north of the city, in one case in
-an inner suburb above the escarpment (\textbf{need to describe the study
-area!}), and lastly in a more suburban/exurban region in the south-west
-(\textbf{share other insights about these regions, maybe based on
-weighted travel patterns?}).
+particularly affected. To better highlight these changes, we used a
+local indicator of spatial autocorrelation (Anselin, 1995) to explore
+the pattern of change in accessibility. Twenty-four TAZ are flagged as
+having significant large losses of accessibility (at \(p<=0.10\),
+without correcting for multiple comparisons). Those zones are
+highlighted in the figure, where it can be seen that they form more or
+less compact neighborhoods. Remarkably, the largest (and significant)
+drops in accessibility are not downtown, but located in two cases in the
+industrial north of the city, in one case in an inner suburb above the
+escarpment, and lastly in a more suburban/exurban region in the
+south-west.
 
@@
 
-Discuss the results of aggregating the accessibility by age group (see
-Table \ref{tab:accessibility-by-age}). Follow with the analysis of hours
-of travel?
-
-\begin{table}
-
-\caption{\label{tab:table-accessibility-by-age-group}\label{tab:accessibility-by-age}Accessibility by age group among members households with incomes less than CAD40,000.}
+For the more suburban clusters of zones, the decrease in accessibility
+is derived from the closure of locations throughout the city reachable
+by car. In the cluster of central suburban zones for example, low-income
+households in the outer ring of zones that exhibit medium to high
+decreases in accessibility within this cluster appear to be largely
+auto-dependent in their tripmaking, which each each exhibiting between
+85-100\% of their modal split for car trips. This results in the parcels
+within these zones having a large number of potentially accessible
+locations in the travel time matrix. But by extension, the change in
+accessibility over the pre- and during-COVID-19 time periods is affected
+not only by the closure of service locations proximate to the zones, but
+also the locations in the central city. The zone with the greatest
+decrease in accessibility within this cluster
+(\ensuremath{-9\times 10^{-4}}) has a high rate of car trips and
+connects to the most food bank locations in total as well as those that
+stayed open or closed. The exception to this is the zone in the middle
+of this cluster, which has a much higher proportion of transit use 32\%.
+This modal split weighting results in fewer food bank and service
+locations accessible from the parcels within this zone which, by
+extension, limits their access to closed locations reachable within 15
+minutes in our calculations.
+
+In the cluster to the south-west, the decrease in accessibility is
+predominately driven by the closure of a high level-of-service Community
+Meals provider. However, like the more central suburban zones,
+low-income households within this cluster also between 90\% to 100\%
+auto-dependent in their tripmaking in the TTS. The story is similar for
+the zone located in the north-west that exhibits the greatest decrease
+in accessibility. Here, low-income households responding to the TTS
+conducted 100\% of their trips by car and, as a result, parcels within
+this zone have access to the second-highest number of food bank and
+service locations within 15 minutes. However, this also means zonal
+accessibilities are greatly affected by the number of closures
+throughout the city. Finally, in the city's north end and north-east
+zones, low-income households exhibit a mixture of tripmaking behaviour
+in the TTS. Households in some zones take transit more often and one
+zone in particular has 100\% of its trips by walking while in others,
+car use makes up 100\% of trips. For these zones, the decrease in
+accessibility tends to be a product of the closure of several inner-city
+food bank and service locations.
+
+Just as the effects of the closures appear to have been uneven in space,
+they also seem to have differently impacted various population segments.
+Table \ref{tab:accessibility-by-age} shows the estimated number of
+people in each age group by their level of accessibility before and
+during the pandemic. Here, it is important to note that the quartiles
+are relative: people in the top 25\% of accessibility still have lower
+accessibility during the pandemic than before. In reality, every
+population group is worse off during the pandemic in terms of their
+accessibility to food banks and services in the City of Hamilton.
+However, some age groups were affected more. As seen in the table, the
+distribution within quartiles for adults changed only slightly, and in
+fact there are fewer adults in the bottom three quartiles as some found
+themselves in the (deteriorated) top quartile of accessibility during
+the pandemic. The story is different for seniors, a greater number of
+whom are now in the third quartile, due to a combination of seniors
+facing a worse accessibility situation, and just a few moving from the
+bottom quartile (pre-pandemic) to the third during the pandemic. Those
+aged 18 and less appear to have fared worse, and the largest change is
+in the number of children who were in the third quartile before the
+pandemic found themselves in the bottom accessibility class during the
+pandemic. These results suggest that, through a combination of the
+typical modes of transportation of lower income households, and the
+spatial distribution of the population, the food bank/service locations
+closed had a differential impact that more greatly affected the youngest
+among the population in low income households.
+
+\begin{table}[!h]
+
+\caption{\label{tab:table-accessibility-by-age-group}\label{tab:accessibility-by-age}Population at each accessibility level by age group among members households with incomes less than CAD40,000.}
 \centering
 \resizebox{\linewidth}{!}{
-\begin{tabular}[t]{cccccc}
+\begin{tabular}[t]{lcccccc}
 \toprule
-\multicolumn{3}{c}{Population} & \multicolumn{3}{c}{Accessibility} \\
-\cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){4-6}
-Children (age $\le$ 18) & Adults (19-64) & Seniors (age $\ge$ 65) & Before COVID-19 & During COVID-19 & Difference\\
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Pre-COVID-19} & \multicolumn{3}{c}{During COVID-19} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
+Accessibility Quartile & Children (age $\le$ 18) & Adults (19-64) & Seniors (age $\ge$ 65) & Children (age $\le$ 18) & Adults (19-64) & Seniors (age $\ge$ 65)\\
 \midrule
-\cellcolor{gray!6}{3,445 (25\%)} & \cellcolor{gray!6}{11,860 (24.8\%)} & \cellcolor{gray!6}{6,584 (25.9\%)} & \cellcolor{gray!6}{0.00229} & \cellcolor{gray!6}{0.00173} & \cellcolor{gray!6}{-0.00057}\\
-4,477 (32.4\%) & 12,838 (26.9\%) & 6,847 (27\%) & 0.00707 & 0.00418 & -0.00289\\
-\cellcolor{gray!6}{3,541 (25.7\%)} & \cellcolor{gray!6}{13,189 (27.6\%)} & \cellcolor{gray!6}{7,020 (27.6\%)} & \cellcolor{gray!6}{0.01580} & \cellcolor{gray!6}{0.00911} & \cellcolor{gray!6}{-0.00669}\\
-2,340 (17\%) & 9,918 (20.7\%) & 4,940 (19.5\%) & 0.03985 & 0.02202 & -0.01783\\
+\cellcolor{gray!6}{Top 25\%} & \cellcolor{gray!6}{2,546} & \cellcolor{gray!6}{10,404} & \cellcolor{gray!6}{5,008} & \cellcolor{gray!6}{2,567 (0.82\%)} & \cellcolor{gray!6}{10,597 (1.86\%)} & \cellcolor{gray!6}{4,941 (-1.34\%)}\\
+Second 25\% & 3,333 & 12,916 & 6,415 & 3,375 (1.26\%) & 12,885 (-0.24\%) & 6,157 (-4.02\%)\\
+\cellcolor{gray!6}{Third 25\%} & \cellcolor{gray!6}{4,381} & \cellcolor{gray!6}{11,559} & \cellcolor{gray!6}{7,067} & \cellcolor{gray!6}{4,062 (-7.28\%)} & \cellcolor{gray!6}{11,412 (-1.27\%)} & \cellcolor{gray!6}{7,678 (8.65\%)}\\
+Bottom 25\% & 3,546 & 12,946 & 6,903 & 3,802 (7.22\%) & 12,897 (-0.38\%) & 6,567 (-4.87\%)\\
 \bottomrule
-\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
-\multicolumn{6}{l}{\rule{0pt}{1em}Population values have been rounded}\\
+\multicolumn{7}{l}{\rule{0pt}{1em}\textit{Note: }}\\
+\multicolumn{7}{l}{\rule{0pt}{1em}Population values have been rounded.}\\
+\multicolumn{7}{l}{\rule{0pt}{1em}The values in brackets for population during COVID-19 are the changes from before the pandemic.}\\
 \end{tabular}}

### 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Edits and notes in conclusion

@@
 This includes modified business hours, a need to make reservations before frequenting, and locations that have completely shut down in consequence.
-While some food bank services have a specific a target population, such as prioritizing family with young children aged between 0 and 3 or accepting only those providing proof of low-income status through housing and utility statements, all the food bank services indicated below are designed to accommodate those in need of food at zero to low cost.
+Table 1 defines each service type and the number of locations pre- and during the COVID-19 pandemic.
+While some food bank services have a specific a target population, such as prioritizing families with young children aged between 0 and 3 or accepting only those providing proof of low-income status through housing and utility statements, all the food bank services indicated below are designed to accommodate those in need of food at zero to low cost.
 With our focus on food banks and food services that offer free or low-cost meals at particular locations, we first removed services such as Meals on Wheels and other food access services such as food box, community kitchens, student nutrition programs, and shopping and transportation.
@@
 We used the Data Retrieval System of the Transportation Tomorrow Survey (TTS)\footnote{\url{http://dmg.utoronto.ca/}} to download cross-tabulations of: 1) primary mode of travel per trip by income by place of residence; and 2) age by income by place of residence.
-These data are from the 2016 Survey (the most recent available), and data are geocoded at the level of Traffic Analysis Zones (TAZ) using the most recent zoning system from 2006.
+These data are from the 2016 Survey (the most recent available).
+The data are geocoded at the level of Traffic Analysis Zones (TAZ) using the most recent zoning system from 2006 and expansion factors are applied to weight the trips .
 Each parcel point is populated with the proportion of trips by three modes of travel: car (as driver or passenger), transit, and walk.
@@
 \noindent where $p^k_i$ is the proportion of trips by mode $k$ in the TAZ of parcel $i$, and $tt^k_{ij}$ is the travel time from parcel $i$ to the food bank.
-In other words, the expected travel time is the weighted sum of travel times to the food bank, with the weights given by the expected modal split in the TAZ.
+In other words, the expected travel time is the weighted sum of travel times to the food bank, with the weights given by the expected modal split of trips made by low-income households in the TAZ per the TTS data.
 
@@
 But by extension, the change in accessibility over the pre- and during-COVID-19 time periods is affected not only by the closure of service locations proximate to the zones, but also the locations in the central city.
-The zone with the greatest decrease in accessibility within this cluster (`r round(accessibility_taz_change %>% st_drop_geometry() %>% filter(TAZUID == 5135) %>% pull(acc_d), 4)`) has a high rate of car trips and connects to the most food bank locations in total as well as those that stayed open or closed.
+The zone with the greatest decrease in accessibility within this cluster (-0.0009) has a high rate of car trips and connects to the most food bank locations in total as well as those that stayed open or closed.
 The exception to this is the zone in the middle of this cluster, which has a much higher proportion of transit use `r round(modes %>% st_drop_geometry() %>% filter(TAZUID == 5116) %>% summarize(first(p_transit)) *100)`\%.
@@
 In the cluster to the south-west, the decrease in accessibility is predominately driven by the closure of a high level-of-service Community Meals provider.
-However, like the more central suburban zones, low-income households within this cluster also between 90% to 100% auto-dependent in their tripmaking in the TTS.
+However, like the more central suburban zones, low-income households within this cluster also between 90\% to 100\% auto-dependent in their tripmaking in the TTS.
 The story is similar for the zone located in the north-west that exhibits the greatest decrease in accessibility.
-Here, low-income households responding to the TTS conducted 100% of their trips by car and, as a result, parcels within this zone have access to the second-highest number of food bank and service locations within 15 minutes.
+Here, low-income households responding to the TTS conducted 100\% of their trips by car and, as a result, parcels within this zone have access to the second-highest number of food bank and service locations within 15 minutes.
 However, this also means zonal accessibilities are greatly affected by the number of closures throughout the city.
@@
 The story is different for seniors, a greater number of whom are now in the third quartile, due to a combination of seniors facing a worse accessibility situation, and just a few moving from the bottom quartile (pre-pandemic) to the third during the pandemic.
-Those aged 18 and less appear to have fared worse, and the largest change is in the number of children who were in the third quartile before the pandemic found themselves in the bottom accessibility class during the pandemic.
-These results suggest that, through a combination of the typical modes of transportation of lower income households, and the spatial distribution of the population, the food bank/service locations closed had a differential impact that more greatly affected the youngest among the population in low income households.
+Those aged 18 and less appear to have fared worse, and the largest change is in the number of children who were in the third quartile before the pandemic and found themselves in the bottom accessibility class during the pandemic.
+These results suggest that, through a combination of the typical modes of transportation of lower income households and the spatial distribution of the population, the food bank/service locations closed had a differential impact that more greatly affected the youngest among the population in low income households.
 
@@
 
+We use data from the census to model low-income households as a proxy for households that are possibly experiencing food insecurity.
+However, the degree to which our low-income cut-off of CAD\$40,000 reflects food insecurity in the different zones of the study area is not known.
+
+While the travel survey allows us to model multi-modal accessibilities that align with travel behaviour observed in the travel survey and capture differences in accessibility by age categories, the use of travel survey data for modelling food insecurity also has its limitations.
+Research into the population weighting methods used in the TTS note that the survey may under-count the lowest- and highest-income households in the survey study region, although the magnitude of this under-counting is unknown as approximately 20% of respondents to the 2016 survey did not report their income [@rose2018transportation].
+In that regard, the modal splits of low-income households observed through the TTS data may not accurately reflect the travel behaviours of food insecure populations.
+
+In the absence of information on how food insecure households travel to food banks and related services, we adopted a normative approach that examines accessibility to food banks reachable within 15 minutes based on standards for accessible drinking water...
+
 References {#references .unnumbered}

### 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Push my knit

@@
 ```{r install-data-package, include = FALSE}
-# Run only once if needed to install data package `sobiEquity`
+# Run only once if needed to install data package 
 if (!require("foodbanksHamilton", character.only = TRUE)) {

### 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Push my knit

@@
 hours, a need to make reservations before frequenting, and locations
-that have completely shut down in consequence. While some food bank
-services have a specific a target population, such as prioritizing
-family with young children aged between 0 and 3 or accepting only those
-providing proof of low-income status through housing and utility
-statements, all the food bank services indicated below are designed to
-accommodate those in need of food at zero to low cost. With our focus on
-food banks and food services that offer free or low-cost meals at
-particular locations, we first removed services such as Meals on Wheels
-and other food access services such as food box, community kitchens,
-student nutrition programs, and shopping and transportation. In
-addition, two free meal services held on different days at the same
-location were collapsed into a single service point for the
-accessibility analysis. Additional details on the operations of
-individual facilities is not publicly available, and with the changes in
-operations it proved unfeasible to collect it. For this reason, the
-analysis to follow is of accessibility to the location of food banks and
-services, but not to specific services (e.g., breakfasts vs.~food
-boxes).
+that have completely shut down in consequence. Table 1 defines each
+service type and the number of locations pre- and during the COVID-19
+pandemic. While some food bank services have a specific a target
+population, such as prioritizing families with young children aged
+between 0 and 3 or accepting only those providing proof of low-income
+status through housing and utility statements, all the food bank
+services indicated below are designed to accommodate those in need of
+food at zero to low cost. With our focus on food banks and food services
+that offer free or low-cost meals at particular locations, we first
+removed services such as Meals on Wheels and other food access services
+such as food box, community kitchens, student nutrition programs, and
+shopping and transportation. In addition, two free meal services held on
+different days at the same location were collapsed into a single service
+point for the accessibility analysis. Additional details on the
+operations of individual facilities is not publicly available, and with
+the changes in operations it proved unfeasible to collect it. For this
+reason, the analysis to follow is of accessibility to the location of
+food banks and services, but not to specific services (e.g., breakfasts
+vs.~food boxes).
 
@@
 place of residence; and 2) age by income by place of residence. These
-data are from the 2016 Survey (the most recent available), and data are
+data are from the 2016 Survey (the most recent available). The data are
 geocoded at the level of Traffic Analysis Zones (TAZ) using the most
-recent zoning system from 2006. Each parcel point is populated with the
-proportion of trips by three modes of travel: car (as driver or
-passenger), transit, and walk.
+recent zoning system from 2006 and expansion factors are applied to
+weight the trips . Each parcel point is populated with the proportion of
+trips by three modes of travel: car (as driver or passenger), transit,
+and walk.
 
@@
 is the weighted sum of travel times to the food bank, with the weights
-given by the expected modal split in the TAZ.
+given by the expected modal split of trips made by low-income households
+in the TAZ per the TTS data.
 
@@
 also the locations in the central city. The zone with the greatest
-decrease in accessibility within this cluster
-(\ensuremath{-9\times 10^{-4}}) has a high rate of car trips and
-connects to the most food bank locations in total as well as those that
-stayed open or closed. The exception to this is the zone in the middle
-of this cluster, which has a much higher proportion of transit use 32\%.
-This modal split weighting results in fewer food bank and service
-locations accessible from the parcels within this zone which, by
-extension, limits their access to closed locations reachable within 15
-minutes in our calculations.
+decrease in accessibility within this cluster (-0.0009) has a high rate
+of car trips and connects to the most food bank locations in total as
+well as those that stayed open or closed. The exception to this is the
+zone in the middle of this cluster, which has a much higher proportion
+of transit use 32\%. This modal split weighting results in fewer food
+bank and service locations accessible from the parcels within this zone
+which, by extension, limits their access to closed locations reachable
+within 15 minutes in our calculations.
 
@@
 in the number of children who were in the third quartile before the
-pandemic found themselves in the bottom accessibility class during the
-pandemic. These results suggest that, through a combination of the
-typical modes of transportation of lower income households, and the
+pandemic and found themselves in the bottom accessibility class during
+the pandemic. These results suggest that, through a combination of the
+typical modes of transportation of lower income households and the
 spatial distribution of the population, the food bank/service locations
@@
 
+We use data from the census to model low-income households as a proxy
+for households that are possibly experiencing food insecurity. However,
+the degree to which our low-income cut-off of CAD\$40,000 reflects food
+insecurity in the different zones of the study area is not known.
+
+While the travel survey allows us to model multi-modal accessibilities
+that align with travel behaviour observed in the travel survey and
+capture differences in accessibility by age categories, the use of
+travel survey data for modelling food insecurity also has its
+limitations. Research into the population weighting methods used in the
+TTS note that the survey may under-count the lowest- and highest-income
+households in the survey study region, although the magnitude of this
+under-counting is unknown as approximately 20\% of respondents to the
+2016 survey did not report their income (Rose, 2018). In that regard,
+the modal splits of low-income households observed through the TTS data
+may not accurately reflect the travel behaviours of food insecure
+populations.
+
+In the absence of information on how food insecure households travel to
+food banks and related services, we adopted a normative approach that
+examines accessibility to food banks reachable within 15 minutes based
+on standards for accessible drinking water\ldots{}
+
 \hypertarget{references}{%
@@
 
+\leavevmode\hypertarget{ref-rose2018transportation}{}%
+Rose, A., 2018. Transportation tomorrow survey 2016: Data expansion and
+validation. Data Management Group, Department of Civil Engineering,
+University of Toronty, Malatest \& Associates.
+
 \leavevmode\hypertarget{ref-seligman2010food}{}%

### 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-title-page.Rmd
Added link to anonymous drive folder to share documents for review purposes; created title page

--- 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
@@
 author:
-  - name: Author One
-    email: author1@example.com
-    affiliation: Some University
-  - name: Author Two
-    email: author2@example.com
-    affiliation: Another University
+  - name: Christopher D. Higgins
+    email: cd.higgins@utoronto.ca
+    affiliation: University of Toronto Scarborough
+  - name: Antonio Paez
+    email: paezha@mcmaster.ca
+    affiliation: McMaster University
     footnote: 1
-  - name: Author Three
-    email: author3@example.com
-    affiliation: Some University
-  - name: Author Four
-    email: author4@example.com
-    affiliation: Some University
+  - name: Gyoorie Kim
+    email: gyoorie.kim@mail.utoronto.ca
+    affiliation: University of Toronto Mississauga
+  - name: Jue Wang
+    email: gis.wang@utoronto.ca
+    affiliation: University of Toronto Mississauga
 address:
-  - code: Some University
-    address: Department, Street, City, State, Zip
-  - code: Another University
-    address: Department, Street, City, State, Zip
+  - code: University of Toronto Scarborough
+    address: Department of Geography & Planning, University of Toronto Scarborough, 1265 Military Trail, Toronto, ON M1C1A4
+  - code: McMaster University
+    address: School of Earth, Environment and Society, McMaster University, Hamilton, ON, L8S 4K1, Canada
+  - code: University of Toronto Mississauga
+    address: Department of Geography, Geomatics and Environment, University of Toronto Mississauga, 3359 Mississauga Road, Mississauga, ON, L5L 1C6 , Canada   
 footnote:
@@
 **Author 1:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization, Supervision; **Author 2:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization; **Author 3:** Conceptualization, Investigation, Data Curation, Writing - Original Draft; **Author 4:** Resources, Writing - Original Draft, Supervision
-
-```{r, setup, include=FALSE}
-knitr::opts_chunk$set(
-  cache = TRUE, out.width = "100%", align = 'center', fig.width = 7, fig.height = 7
-)
-```
-
-```{r install-data-package, include = FALSE}
-# Run only once if needed to install data package `sobiEquity`
-if (!require("foodbanksHamilton", character.only = TRUE)) {
-  devtools::install_github("https://github.com/paezha/Accessibility-Food-Banks-Hamilton", subdir = "foodbanksHamilton")
-}
-```
-
-```{r load-packages, cache=FALSE, include=FALSE}
-library(foodbanksHamilton) # Data package
-library(kableExtra)
-library(patchwork)
-library(sf)
-library(spdep)
-library(tidyverse)
-#library(tmap)
-```
-
-```{r invoke-data, include=FALSE}
-data(data_da_2016)
-data(dwelling_network_points_2016)
-data(foodbank_locations)
-data(foodbank_table)
-data(hamilton_boundary)
-data(modes_less40k)
-data(population_age_income)
-data(ttm_car)
-data(ttm_transit)
-data(ttm_walk)
-```
-
-```{r remove-irrelevant-destinations, include=FALSE}
-## Data preparation
-# Remove entries in travel time matrices for inapplicable destinations
-# As a first step, we will remove the computed travel times to the three Community Kitchen locations (IDs 1-3):
-ttm_car <- ttm_car %>%
-  filter(!between(OBJECTID, left = 1, right = 3))
-
-ttm_transit <- ttm_transit %>%
-  filter(!between(OBJECTID, left = 1, right = 3))
-
-ttm_walk <- ttm_walk %>%
-  filter(!between(OBJECTID, left = 1, right = 3))
-```
-
-```{r join-low-income-households, include=FALSE}
-## Data preparation
-# Join proportion of low income households per DA to dwellings and total number of residential units:
-dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
-  left_join(data_da_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::transmute(DAUID = GeoUID, 
-                               p_inc_less40k),
-            by = "DAUID")
-```
-
-```{r probability-of-low-income-per-parcel, include=FALSE}
-## Data preparation
-# Calculate probability that a point on the network will have low income households.
-This is the proportion of low income households in the DA multiplied by the total number of residential units at the point:
-dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
-  mutate(income_less40k = p_inc_less40k * Dwellings)
-```
-
-```{r number-of-residential-units-per-DA, include=FALSE}
-## Data preparation
-# Prepare and join population data from Transportation Tomorrow Survey
-# Find number of residential units in each DA. This is needed to assign the population to parcels
-residential_units <- dwelling_network_points_2016 %>%
-  st_drop_geometry() %>%
-  group_by(TAZUID) %>%
-  summarize(residential_units = sum(Dwellings))
-```
-
-```{r join-population-to-parcels, include=FALSE}
-## Data preparation
-# Join number of low income households per DA to dwellings and total number of residential units:
-dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
-  left_join(population_age_income %>%
-              st_drop_geometry() %>%
-              dplyr::transmute(TAZUID,
-                               children,
-                               adults,
-                               seniors),
-            by = "TAZUID") %>%
-  left_join(residential_units,
-            by = "TAZUID")
-```
-
-```{r calculate-low-income-households-per-parcel, include=FALSE}
-## Data preparation
-# Calculate population per point on the network.
-This is the population in the DA divided by the total number of residential units.
-This gives the average number of people per residential unit.
-Then, multiply by the number of dwellings (residential units) at the point:
-dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
-  mutate(children = (children/residential_units) * Dwellings,
-         adults = (adults/residential_units) * Dwellings,
-         seniors = (seniors/residential_units) * Dwellings)
-```
-
-```{r join-household-data-to-travel-time-tables, include=FALSE}
-## Data preparation
-# Join the household data to the travel time matrices:
-ttm_car <- ttm_car %>%
-  left_join(dwelling_network_points_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            income_less40k,
-                            children,
-                            adults,
-                            seniors,
-                            DAUID, 
-                            TAZUID),
-            by = c("UID" = "ID"))
-
-ttm_transit <- ttm_transit %>%
-  left_join(dwelling_network_points_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            income_less40k,
-                            children,
-                            adults,
-                            seniors,
-                            DAUID, 
-                            TAZUID),
-            by = c("UID" = "ID"))
-
-ttm_walk <- ttm_walk %>%
-  left_join(dwelling_network_points_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            income_less40k,
-                            children,
-                            adults,
-                            seniors,
-                            DAUID, 
-                            TAZUID),
-            by = c("UID" = "ID"))
-```
-
-```{r join-foodbank-data-to-travel-time-tables, include=FALSE}
-## Data preparation
-# Join the foodbank data to the travel time tables:
-ttm_car <- ttm_car %>%
-  left_join(foodbank_locations %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            TYPE,
-                            COVIDClose),
-            by = c("OBJECTID" = "ID"))
-
-ttm_transit <- ttm_transit %>%
-  left_join(foodbank_locations %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            TYPE,
-                            COVIDClose),
-            by = c("OBJECTID" = "ID"))
-
-ttm_walk <- ttm_walk %>%
-  left_join(foodbank_locations %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID,
-                            TYPE,
-                            COVIDClose),
-            by = c("OBJECTID" = "ID"))
-```
-
-```{r drop-nas-from-travel-time-tables, include=FALSE}
-## Data preparation
-# Drop NAs:
-ttm_car <- ttm_car %>%
-  drop_na(TAZUID,
-          income_less40k)
-
-ttm_transit <- ttm_transit %>%
-  drop_na(TAZUID,
-          income_less40k)
-
-ttm_walk <- ttm_walk %>%
-  drop_na(TAZUID,
-          income_less40k)
-```
-
-```{r proportion-of-trips-by-mode-by-taz, include=FALSE}
-## Data preparation
-# Calculate proportion of trips by mode per TAZ:
-modes <- modes_less40k %>%
-  mutate(p_car = Driver/(Driver + Transit + Walk),
-         p_transit = Transit/(Driver + Transit + Walk),
-         p_walk = Walk/(Driver + Transit + Walk))
-```
-
-```{r fill-nas-transport-modes, include=FALSE}
-## Data preparation
-# Assume that TAZ with NAs have a proportion of car trips of one:
-modes <- modes %>%
-  dplyr::mutate(p_car = replace_na(p_car, 1),
-                p_transit = replace_na(p_transit, 0),
-                p_walk = replace_na(p_walk, 0))
-```
-
-```{r join-trips-by-mode-to-travel-time-tables, include=FALSE}
-## Data preparation
-# Join trip-making modal proportions
-ttm_car <- ttm_car %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-
-ttm_transit <- ttm_transit %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-
-ttm_walk <- ttm_walk %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-```
-
-```{r join-travel-time-tables, include=FALSE}
-## Data preparation
-# Join Tables into Single TTM
-ttm <- ttm_car %>%
-  rename(travel_time_car = travel_time) %>%
-  full_join(ttm_transit %>%
-              transmute(UID,
-                        OBJECTID,
-                        travel_time_transit = travel_time),
-            by = c("UID", "OBJECTID")) %>%
-  full_join(ttm_walk %>%
-              transmute(UID,
-                        OBJECTID,
-                        travel_time_walk = travel_time),
-            by = c("UID", "OBJECTID"))
-```
-
-```{r travel-time-tables-data-check, include=FALSE}
-## Data preparation
-# Check the NAs in travel time table:
-ttm %>% 
-  filter(is.na(travel_time_transit)) %>%
-  select(travel_time_transit,
-         p_transit) %>%
-  summary()
-
-# What do NAs here mean?
-Most DAs missing travel time by transit do not report trips by transit in the TTS.
-In some DAs there are trips by transit (p_transit != 0), but no foodbank can be reached by transit (travel_time_transit is NA).
-How many of these cases?
-ttm %>% 
-  filter(is.na(travel_time_transit) & p_transit > 0) %>%
-  select(travel_time_transit,
-         p_transit) %>%
-  summary()
-
-# It should be fine to replace the NAs with zeros, which multiplied by p_transit will return a zero (i.e., it is not possible to reach a foodbank by transit).
-ttm <- ttm %>% 
-  mutate(travel_time_transit = replace_na(travel_time_transit, 0))
-
-# Check the NAs in travel time by walk:
-ttm %>% 
-  filter(is.na(travel_time_walk)) %>%
-  select(travel_time_walk,
-         p_walk) %>%
-  summary()
-
-# Most DAs missing travel time by walking do not report trips by walking in the TTS.
-In some DAs there are trips by walking (p_walk != 0), but no foodbank can be reached by walk (travel_time_walk is NA).
-How many of these cases?
-ttm %>% 
-  filter(is.na(travel_time_walk) & p_walk > 0) %>%
-  select(travel_time_walk,
-         p_walk) %>%
-  summary()
-
-# Like with transit, it should be fine to replace the NAs with zeros, which multiplied by p_walk will return a zero (i.e., it is not possible to reach a foodbank by walking).
-ttm <- ttm %>% 
-  mutate(travel_time_walk = replace_na(travel_time_walk, 0))
-
-#Summary of the table after these operations:
-summary(ttm)
-```
-
-```{r calculate-expected-travel-times, include=FALSE}
-## Data preparation
-# Calculate expected travel times by mode.
-These are the travel time by the mode multiplied by the probability of using the mode (which is proxied by the proportion of trips by mode in the DA from the TTS).
-The total expected travel time is the sum of the expected travel times by mode:
-ttm <- ttm %>%
-  mutate(e_travel_time_car = travel_time_car * p_car,
-         e_travel_time_transit = travel_time_transit * p_transit,
-         e_travel_time_walk = travel_time_walk * p_walk,
-         e_travel_time = e_travel_time_car + e_travel_time_transit + e_travel_time_walk)
-```
-
-\newpage
-
-Introduction
-==========================
-
-Food insecurity is defined as an "inadequate or uncertain access to a sufficient quantity and/or adequate quality of food" due to a household's financial limitations [@enns2020experiences].
-This condition has been associated with reductions in nutritional outcomes [@bhattacharya2004poverty; @kirkpatrick2008food; @olson1999nutrition] and physical and mental health in children and adults [@elgar2021relative; @jones2017food; @ramsey2011food; @seligman2010food; @stuff2004household].
-Over at least the past four decades food banks and related services have become an essential line of defense against food insecurity in Canadian communities [@riches2002food; @holmes2018nothing; @black2020examining; @tarasuk2020relationship].
-In this respect, Canada is not unlike numerous other wealthy countries where a systematic dismantling of the welfare state took place in the intervening period [@tarasuk2014food].
-
-The emergence of COVID-19, the worst public health crisis since the 1918 flu pandemic, has revealed important social and economic fault lines, and pre-existing patterns of inequality appear to have been exacerbated.
-Along several other dimensions of stress [e.g., accessibility to health care facilities, @pereira2021geographic], this seems to be the case for food insecurity as well [@laborde2020poverty].
-In the US, for example, it has been estimated that there was an increase of more than 30\% in household food insecurity, and more than one third of households were discovered to be newly food insecure - meaning they did not experience food insecurity before the pandemic [@niles2020early].
-In Canada, Men and Tarasuk [-@men2021food] report that about 25\% of individuals who experienced job insecurity (a relatively common occurrence during the pandemic), also experienced food insecurity.
-Similarly, according to Statistics Canada [-@statisticscanada2020food], in the early stages of the pandemic almost 15\% of individuals reported living in a household that faced food insecurity; the risk of food insecurity was substantially higher for households with children.
-The difference between households with and without children was significant, and 11.7\% of households with children indicated that "food didn’t last and [there was] no money to get more" sometimes or often, compared to 7.3\% of households without children; likewise, 13\% of households with children indicated that they "[c]ouldn’t afford balanced meals" sometimes or often, compared to 8.8\% of households without children.
-
-The impacts of food insecurity during the pandemic are alarming, since diet-related diseases, such as obesity, heart-disease, and diabetes, were already critical public health concerns in Canada prior to COVID-19 [@boucher2017ontario].
-While food banks are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least they can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
-As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that food banks in Hamilton, Ontario, recorded the highest number of visitors in the past 29 years; the number of children visiting food banks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278 the year before, a rate of increase greater than population growth.
-
-It is known that urban food environments, within which people make their daily food choices, are essential in influencing eating behaviours and health outcomes, based on factors such as food availability, ease of geographic accessibility and socio-demographic variations [@paez2010relative; @vanderlee2017food; @widener2018spatial].
-However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to food banks.
-Although previous work has explored differences in _accessing_ food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource [e.g., @enns2020experiences], we are not aware of any research that has focused on the geographic component of accessibility.
-
-The study of place-based geographic accessibility is concerned with capturing the potential to reach destinations of value using the transportation network [@paez2012measuring].
-Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities^[https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with food bank service delivery [@holmes2018nothing].
-However, as with research into the prevalence of food deserts, accessibility to food banks is unlikely to be evenly distributed, and variation throughout a city can be expected due to transportation network characteristics, and the spatial distribution of food bank locations and the population they are meant to serve.
-Furthermore, policy responses to the COVID-19 pandemic likely have added to the distress of vulnerable households.
-Non-pharmaceutical interventions during the pandemic involving restrictions in mobility have increased the friction of travel, in particular by transit on which low income populations are more reliant [e.g., @deweese2020tale].
-At the same time, the pandemic has created additional stress for the operators of food banks through disruptions in the supply chain [e.g., @mckay2021exploring] as well as concerns surrounding the delivery of service in safe conditions and possible cancellation of food service programs.
-
-For this study, we aim to look at how the landscape of food banks and related services (e.g.
-low-cost or free meal service providers) available in Hamilton, Ontario, changed during the pandemic.
-Did the number of open food bank services diminish?
-If so, what was the accessibility to food banks before the pandemic from the perspective of low income households, and how it changed during the pandemic?
-And finally, who are most likely to have been impacted by changes in the accessibility landscape?
-This paper first looks at the distribution of food banks and related services before and during the pandemic.
-Then, we use the balanced floating catchment area approach of Paez et al.
-[-@paez2019demand] to investigate the accessibility situation.
-For this, we use a fully disaggregated approach based on parcel-level data.
-Socio-economic and demographic data are drawn from the latest Census of Canada (2016), whereas travel information is from the most recent regional travel survey from 2016.
-This paper follows reproducible research recommendations [see @brunsdon2020opening], and the research was conducted using open source tools for transportation analysis [@lovelace2021open].
-The code and data necessary to reproduce the analysis are available in a public repository^[add repository].
-
-Food Insecurity and Food Bank Use in Canada
-============
-
-Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake [@enns2020experiences; @kirkpatrick2008food; @tarasuk2009household].
-This nutrient deficiency is associated with major population health concerns, particularly among Canadians at socio-economic disadvantage [@bazerghi2016role].
-Data on food insecurity in Canada is collected in several ways.
-Quantitatively, official government surveys such as the Household Food Security Survey Module (HFSSM), the Canadian Community Health Surveys (CCHS), the International Study of Adults (LISA), and official classifications determined by Health Canada in relation to socio-demographic variables offer some insight into food insecurity [@kirkpatrick2008food; @tarasuk2009household; @gundersen2018food].
-Nationally, it has been found that food insecurity impacts approximately 12.3% of Canadian households [@tarasuk2014food]
-
-Food banks - sometimes also referred to as ‘food pantries’ and ‘food shelves’ - originated as a community response to aid those with inadequate food by voluntarily offering them meals and ingredients [@riches2002food; @loopstra2012relationship].
-The scope and objectives of food banks can vary by region and by country, and these organizations can include not only prepared meals and aliments, but also shared spaces to connect in community gardens and community kitchens [@wakefield2013sweet].
-Although in their origin food banks were meant to be provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time they have evolved into a community practice to secure emergency food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
-
-In Canada, the number of food banks has steadily increased in the past few decades [@wakefield2013sweet].
-The largest database of food banks and their use comes from the non-profit association Food Banks Canada (FBC), which conducts an annual assessment through its affiliated members.
-FBC's 2018 Hunger Count report^[https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018_FINAL_EN.pdf.aspx?ext=.pdf] (the most recent available) listed 1,830 member food banks across the country, and found that Canadians visited food banks 1.1 million times in March of 2018.
-Of those accessing food banks, certain population characteristics tend to be over-represented compared to national totals from the 2016 Canadian Census of Population.
-According to FBC's 2018 data, single-adult households represent 45% of those utilizing food banks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those accessing food bank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
-In addition, 59% of households accessing food banks list social or disability assistance as their primary source of income.
-Similarly, using data from the 2011-2012 CCHS, Tarasuk et al.
-[-@tarasuk2019geographic] found higher odds of food insecurity amongst households relying on social assistance, those without a university degree or with children under the age of 18, and individuals that lived alone, renters, and those identifying as Aboriginal.
-While surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada [@tarasuk2014food], research from Ottawa [@enns2020experiences] and Vancouver [@black2020examining] suggests that long-term users tend to be older, have health or mobility challenges, live in large households, and are less likely to have employment income.
-
-Economic stress caused by the COVID-19 pandemic, rising unemployment rates, and changes in poverty levels have disrupted the food environment and led to higher rates of food insecurity [@niles2020early].
-Pandemic-related food security studies in the US have found a substantial increase in households experiencing food insecurity for the first time, and also in households experiencing more severe food insecurity than before [@niles2020early; @wolfson2020food].
-Most recently in May 2020, Canada recorded 14.7% of its population living in food insecurity in the past 30 days [@statisticscanada2020food].
-Considering the negative mental and physical health effects associated with food insecurity, increases in food insecurity rates due to the pandemic, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
-Recent data^[https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis_Exec-Sum_EN.pdf] from FBC showed that 52% of member food banks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of the country.
-The pandemic also created significant staffing issues with 42% of food banks reporting a reduction in volunteers.
-However, 53% of food banks later reported a decrease in use into the summer of 2020 which FBC members attributed to emergency financial support programs from the federal government.
-Nevertheless, some of these benefit programs were temporary, which suggests that many households may again turn to food bank services to meet their needs.
-
-In terms of geography, previous research conducted at the provincial scale using data from the 2011-2012 CCHS found that the prevalence of food insecurity ranged across the country from 11.8% of households in Ontario to 41% of households in Nunavut [@tarasuk2019geographic].
-However, although previous research has examined the characteristics of individuals and households accessing food banks, the locational or transportation accessibility aspect of food bank access is not well understood.
-A wealth of literature examining the food desert concept suggests that, in addition to socio-economic and demographic factors, location and transportation networks play a key role in a household's accessibility to healthy foods [@paez2010relative; @vanderlee2017food; @widener2018spatial].
-For food banks specifically, previous qualitative research in Ontario by Smith-Carrier et al.
-[-@smith2017food] has noted that "transportation can be challenging, particularly if the food bank is situated in a remote location" (p.
-32).
-Particularly, it appears that participants experience challenges with the "inordinate amount of time necessary to obtain food, and difficulties associated with transportation" (p.
-39).
-Users of food banks, according to this research, rely on a variety of modes of transportation to access services.
-Consequently, the location of facilities matters; in the words of an interviewee: "I wish it [the food bank] was a little more centrally located.
-Because if I didn’t have a bike I’d have to walk it all the way out there and back.
-I wonder about people who don’t" (p.
-39).
-
-To offer greater insight into the role of transportation and location in food bank accessibility, this research examines how geographic accessibility to food banks and food services changed in Hamilton during the COVID-19 pandemic.
-
-Methods and Materials
-===================
-
-## Methods
-
-For the research in this paper we adopt the balanced floating catchment area approach of Paez et al.
-[-@paez2019demand].
-This method for estimating accessibility is a form of the widely-used two-stage floating catchment area method [@radke2000spatial; @luo2003measures].
-Floating catchment areas are used to estimate accessibility when there are potential congestion effects, and operate by calculating first the _demand_ for spatially distributed services.
-The demand (usually the number of people who require a service) is used to calculate a level of service.
-In a second step, the level of service is allocated back to the population.
-Demand and level of service are allocated using some form of distance-decay to embody the geographical principle that, given a choice, people prefer to travel less than more when reaching destinations.
-
-More formally, the first step of this method is as follows:
-$$
-L_j = \frac{S_j}{\sum_{i=1}^nP_iw_{ij}}
-$$
-
-\noindent where $S_j$ is the level of supply at location $j$, in simplest terms whether a service point is present (i.e., $S_j=1$) or not (i.e., $S_j=0$); $P_i$ is the population at location $i$ that demands the service; and $w_{ij}$ is a weight, typically a function of the distance between locations $i$ and $j$.
-$L_j$ is the level of service at location $j$ and it is the inverse of the number of people that need to be serviced.
-
-The second step in this process is then summing the level of service that each population unit can reach, according to the distance-decay weight:
-$$
-A_i = \sum_{j=1}^JL_jw_{ji}
-$$
-
-\noindent where $A_i$ is the accessibility to the service, which is in the same units as the level of service: as the inverse of the population being serviced.
-When the population being serviced is low accessibility is high (i.e., there is little competition for the service), and viceversa.
-
-Floating catchment area methods are prone to overestimation of the population and the level of service due to multiple-counting.
-The population at $P_i$ is allocated to _every_ service point $j$ for which $w_{ij}>0$.
-Similarly, the level of service at $LOS_j$ is allocated to _every_ population point for which $w_{ji}>0$.
-This inflation effect has been known for several years, and several modifications have been proposed to mitigate it [e.g., @wan2012three; @delamater2013spatial].
-A definitive solution to this issue was presented by Paez et al.
-[-@paez2019demand].
-In order to avoid the multiple-counting in the summations, the population and the level of service need to be allocated _proportionally_.
-This is achieved by standardizing the weights as follows:
-$$
-w_{ij}^\text{st} = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
-$$
-\noindent and:
-$$
-w_{ji}^\text{st} = \frac{w_{ji}}{\sum_{j=1}^Jw_{ji}}
-$$
-
-The standardized weights satisfy the following conditions:
-$$
-\sum_{i=1}^nw_{ij}^\text{st}=1
-$$
-
-\noindent and:
-$$
-\sum_{j=1}^Jw_{ji}^\text{st}=1
-$$
-
-Since the population is allocated proportionally, its value is preserved:
-$$
-\sum_{i=1}^nP_iw_{ij}^\text{st}=P_i
-$$
-
-\noindent as is the level of service:
-$$
-\sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
-$$
-## Study Area
-
-**Need to provide some context here: the geography which makes transportation challenging, i.e., due to the escarpment; the main parts of the city: city core, inner suburbs, industrial north, suburban, exurban; and the industrial background of the city maybe, with poverty rates compared to other Canadian cities?**
-
-## Data
-
-Data have been prepared for sharing as a data package^[add repository]. The contents of the data package are described next.
-
-### Statistics Canada
-
-Population and income statistics for 2016 were retrieved at the level of Dissemination Areas (DAs) using the package `cancensus` [@vonBergmann2021cancensus].
-DAs are the smallest publicly available census geography in Canada.
-Income data corresponds to the count of households by different total income groupings.
-
-### Origins: Residential parcels
-
-We converted all recorded residential land parcels in the City of Hamilton to points on the road network.
-Each point includes information about the number of residential units in the parcel.
-Next, we define low-income households as those having a total income of less than CAD40,000, which is approximately the mid-point of the low income cut-off (LICOs) for families in Canadian cities with populations greater than 500,000 in 2016, to match other Census data [@statisticscanada2020licos].
-We then "populate" each residential unit with the probability of being a low-income household based on the counts of households by income groups in the DA in which the parcel is located.
-While this method assumes a constant probability of low-income household status for all residential units in a DA, the parcel-level analysis affords a high level of spatial disaggregation for the accessibility analysis.
-
-### Destinations: Food Banks and Food Service Locations
-
-The locations of food banks and related food services were obtained from the Hamilton Public Library's Food Access Guide^[http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf].
-The guide was updated in April of 2021 to indicate any change affected on the services due to the pandemic.
-This includes modified business hours, a need to make reservations before frequenting, and locations that have completely shut down in consequence.
-While some food bank services have a specific a target population, such as prioritizing family with young children aged between 0 and 3 or accepting only those providing proof of low-income status through housing and utility statements, all the food bank services indicated below are designed to accommodate those in need of food at zero to low cost.
-With our focus on food banks and food services that offer free or low-cost meals at particular locations, we first removed services such as Meals on Wheels and other food access services such as food box, community kitchens, student nutrition programs, and shopping and transportation.
-In addition, two free meal services held on different days at the same location were collapsed into a single service point for the accessibility analysis.
-Additional details on the operations of individual facilities is not publicly available, and with the changes in operations it proved unfeasible to collect it.
-For this reason, the analysis to follow is of accessibility to the location of food banks and services, but not to specific services (e.g., breakfasts vs.
-food boxes).
-
-```{r table-food-bank-info, echo=FALSE, results='asis'}
-foodbank_table %>% kable("latex",
-                         booktabs = TRUE,
-                         align = ("cccccc"),
-                         col.names = c("Type", 
-                                       "Description", 
-                                       "Locations Pre-COVID", 
-                                       "Locations During COVID", 
-                                       "Additional Notes"),
-                         caption = "\\label{tab:food-bank-info}foodbank and Food Service Information.",
-                         escape = FALSE) %>%
-  kable_styling(latex_options = c("striped", "scale_down")) %>%
-  column_spec(2, width = "15em") %>%
-  column_spec(5, width = "15em")
-```
-
-### Routing and travel time tables
-
-Travel time tables for three modes (car, transit, walking) were computed using the parcels as the origins and the locations of the food banks as the destinations.
-For routing, the package `r5r` [@pereira2021r5r] was used with a network extract for the City of Hamilton from OpenStreetMaps and the General Transit Feed Specification from Hamilton Street Railway, the local transit operator.
-For routing purposes we used maximum values of 180 min and 10,000 m walking distance: any destination that exceeded these thresholds was ignored.
-The departure time used for routing was 8:00AM on March 30, 2021 to reflect transit service around the morning service peak on a typical Tuesday.
-
-### Transportation Tomorrow Survey
-
-We used the Data Retrieval System of the Transportation Tomorrow Survey (TTS)\footnote{\url{http://dmg.utoronto.ca/}} to download cross-tabulations of: 1) primary mode of travel per trip by income by place of residence; and 2) age by income by place of residence.
-These data are from the 2016 Survey (the most recent available), and data are geocoded at the level of Traffic Analysis Zones (TAZ) using the most recent zoning system from 2006.
-Each parcel point is populated with the proportion of trips by three modes of travel: car (as driver or passenger), transit, and walk.
-
-### Expected Travel Times
-
-Once we obtained travel time tables with population (number of households) and proportion of trips by mode, we calculated the expected travel time $ett$ from each parcel $i$ to a food bank or food service location $j$ as follows:
-$$
-ett_{ij} = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
-$$
-
-\noindent where $p^k_i$ is the proportion of trips by mode $k$ in the TAZ of parcel $i$, and $tt^k_{ij}$ is the travel time from parcel $i$ to the food bank.
-In other words, the expected travel time is the weighted sum of travel times to the food bank, with the weights given by the expected modal split in the TAZ.
-
-Results and Discussion
-===================
-
-```{r accessibility-threshold-sensitivity, include=FALSE}
-## Analysis
-# Testing different thresholds to check how accessibility changes. Here, the thresholds are varied from 2 min to 30 min in 2 min step increments
-threshold_seq <- seq(from = 2, to = 30, by = 1)
-
-# Initialize a data frame to store results
-accessibility_test <- data.frame(threshold = threshold_seq, 
-                                 accessibility = numeric(length(threshold_seq)))
-
-# Calculate the system-wide accessibility according to each threshold
-for(i in 1:length(threshold_seq)){
-  results_test <-  ttm %>% 
-    transmute(UID, OBJECTID, travel_time = e_travel_time, population = income_less40k, supply = 1) %>%
-    b2sfca(threshold = i )
-  accessibility_test$accessibility[i] <- sum(results_test$accessibility$accessibility)
-}
-```
-
-```{r results-pre-covid, include=FALSE}
-## Analysis 
-# Calculate k-min accessibility pre-covid:
-results_pre <- ttm %>%
-  #dplyr::filter(COVIDClose == "No") %>%
-  transmute(UID, 
-            OBJECTID, 
-            travel_time = e_travel_time, 
-            population = income_less40k, 
-            supply = 1) %>%
-  b2sfca(threshold = 15)
-```
-
-```{r results-during-covid, include=FALSE}
-## Analysis 
-# Calculate k-min accessibility during covid:
-results_covid <- ttm %>%
-  dplyr::filter(COVIDClose == "No") %>%
-  transmute(UID, 
-            OBJECTID, 
-            travel_time = e_travel_time, 
-            population = income_less40k, 
-            supply = 1) %>%
-  b2sfca(threshold = 15)
-```
-
-```{r, bind-accessibility-results, include=FALSE}
-## Analysis
-# Bind level of service results:
-los <- rbind(data.frame(results_pre$los, 
-                        timing = "Pre-covid", 
-                        threshold = "30 min"),
-             data.frame(results_covid$los, 
-                        timing = "During covid", 
-                        threshold = "30 min")) %>%
-  mutate(timing = factor(timing, 
-                         levels = c("Pre-covid", 
-                                    "During covid"),
-                         ordered = TRUE))
-
-# Bind accessibility results:
-accessibility <- rbind(data.frame(results_pre$accessibility, 
-                                  timing = "Pre-covid", 
-                                  threshold = "30 min"),
-                       data.frame(results_covid$accessibility, 
-                                  timing = "During covid", 
-                                  threshold = "30 min")) %>%
-  mutate(timing = factor(timing, 
-                         levels = c("Pre-covid", 
-                                    "During covid"),
-                         ordered = TRUE))
-```
-
-Figure \ref{fig:foodbanks} shows the location of food banks and services in the City of Hamilton and their status.
-Before the pandemic there were `r nrow(foodbank_locations)` of which `r nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))` (`r round(nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations) * 100, 2)`\%) closed during the pandemic.
-As shown in the figure, food services tend to be predominantly located in the central parts of the city.
-This is not surprising: population density is high there, and it is also the part of the city where lower income households are more numerous in absolute and relative terms (see Figure \ref{fig:low-income-households}).
-Alas, this is also the part of the city where most of the closures during the pandemic happened.
-
-```{r plot-location-foodbanks, echo=FALSE, fig.cap="\\label{fig:foodbanks}Location of food banks/services and operation status; the dotted box is an inset of the central part of the City of Hamilton"}
-## Analysis
-# Plot location of foodbanks
-
-# Create a bounding box for an inset
-inset_box <- st_bbox(c(xmin = -79.975,
-                       xmax = -79.775, 
-                       ymax = 43.2, 
-                       ymin = 43.3), 
-                     crs = st_crs(4326)) %>%
-  st_as_sfc()
-
-fb <- ggplot() +
-  geom_sf(data = hamilton_boundary,
-          fill = "lightgray") +
-  geom_sf(data = foodbank_locations,
-          aes(size = COVIDClose,
-              shape = COVIDClose)) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_shape_manual(name = "Closed during COVID-19",
-                     values = c("Yes" = 4, "No" = 1)) +
-  scale_size_manual(values = c("Yes" = 3, "No" = 2)) +
-  guides(size = FALSE) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90),
-        legend.position = "none")
-
-fb_inset <- ggplot() +
-  geom_sf(data = hamilton_boundary %>% 
-            st_transform(4326),
-          fill = "lightgray") +
-  geom_sf(data = foodbank_locations,
-          aes(size = COVIDClose,
-              shape = COVIDClose)) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_shape_manual(name = "Closed during COVID-19",
-                     values = c("Yes" = 4, "No" = 1)) +
-  scale_size_manual(values = c("Yes" = 3, "No" = 2)) +
-  xlim(-79.975, -79.775) +
-  ylim(43.2, 43.3) +
-  guides(size = FALSE) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90),
-        legend.position = "bottom")
-
-fb / fb_inset
-```
-
-```{r plot-low-income-households, echo=FALSE, fig.cap="\\label{fig:low-income-households}Number and proportion of households with incomes less than CAD40,000."}
-lih <- ggplot() +
-  geom_sf(data = data_da_2016,
-          aes(fill = income_less40k),
-          color = NA) +
-  geom_sf(data = hamilton_boundary,
-          fill = NA) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_fill_viridis_c(name = "Number of households", 
-                       option = "A",
-                       direction = -1,
-                       breaks = c(200, 400, 600)) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "bottom")
-
-p_lih <- ggplot() +
-  geom_sf(data = data_da_2016,
-          aes(fill = p_inc_less40k),
-          color = NA) +
-  geom_sf(data = hamilton_boundary,
-          fill = NA) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_fill_viridis_c(name = "Proportion of households", 
-                       option = "A",
-                       direction = -1,
-                       breaks = c(0.25, 0.50, 0.75)) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "bottom")
-
-lih / p_lih
-```
-
-To implement the accessibility calculations, we must select a distance-decay function.
-In this task we find limited support in the literature, which is mostly silent on the travel patterns of people who visit food banks.
-For this reason, we opt for a simple cumulative opportunities function as follows:
-$$
-w_{ij}=w_{ji}=
-\begin{cases}
-1 & \text{ if } ett_{ij}\le \delta\\
-0 & \text{ otherwise}
-\end{cases}
-$$
-\noindent where $ett_{ij}$ is the multimodal expected travel time as described previously, and $\delta$ is a travel threshold.
-When the expected travel time exceeds this threshold, a facility is no longer considered accessible.
-Moreover, the weights are standardized for the balanced floating catchment area approach.
-
-Figure \ref{fig:sensitivity-analysis} shows the results of conducting sensitivity analysis of the system-wide accessibility as we vary the threshold (considering the situation before the pandemic).
-There is a clear pattern whereby more strict values of $\delta$ are associated with higher levels of system-wide accessibility: while increases in accessibility that result from decreases in the travel time window might seem counter-intuitive, this is a result of lower _congestion_, since fewer households are serviced and thus competition for the same resources is more limited.
-System-wide accessibility declines with higher values of $\delta$: as more households are serviced, congestion grows and the level of service declines, although this happens at a declining rate.
-We are not aware of any research that explains how long people are expected to travel for food banks, but we note that in developing countries, access to drinking water is defined as a source that takes less than 30 minutes to reach [round trip, see @world2019progress].
-There is no reason why people in affluent countries should be expected to travel more for a basic need such as food.
-Accordingly, we adopt a 15-minute threshold for the analysis (representing a one-way trip).
-This threshold is also approximately where the rate of change in accessibility slows down.
-
-```{r plot-results-sensitivity-analysis, out.width="60%", echo=FALSE, fig.cap="\\label{fig:sensitivity-analysis}Accessibility as a function of threshold"}
-acc_sensitivity <- ggplot(accessibility_test, 
-                          aes(x = threshold, y = accessibility)) +
-  geom_line() +
-  ylab("Accessibility") +
-  xlab("Threshold (in min)") +
-  theme_minimal()
-
-acc_sensitivity
-```
-
-Using the 15-minute threshold, we find that the system-wide accessibility (interpreted as a provider-to-population ratio) was `r round(sum(results_pre$accessibility$accessibility), 3)` (food banks/service locations per low income household in the city) before COVID-19, but declined to `r round(sum(results_covid$accessibility$accessibility), 3)` during the pandemic.
-It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`\% of food facilities remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`\%, suggesting the location of food banks and related services plays an important role in serving those in need.
-
-```{r join-zonal-ids-to-accessibility-table, include=FALSE}
-## Analysis
-# Join zonal identifiers to accessibility results:
-accessibility <- accessibility %>%
-  left_join(dwelling_network_points_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID,
-                            DAUID,
-                            TAZUID,
-                            children,
-                            adults,
-                            seniors),
-            by = c("UID" = "ID")) %>%
-  left_join(population_age_income %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID),
-            by = "TAZUID") 
-```
-
-```{r, join-geometry-to-level-of-service, include=FALSE}
-## Analysis 
-# Join geometry:
-los <- los %>%
-  left_join(foodbank_locations %>%
-              dplyr::select(ID),
-            by = c("OBJECTID" = "ID")) %>%
-  st_as_sf()
-```
-
-Turning to the location of individual facilities, the levels of service offered before and during the pandemic are shown in Figure \ref{fig:levels-of-service}.
-The level of service is functionally the inverse of the number of low-income households in the travel-mode weighted travel time catchment area of the facilities (this is because $S_j=1 \forall j$, i.e., each location represents a "capacity" of 1).
-Higher values mean that a facility is expected to service fewer households.
-Conversely, lower values indicate greater congestion.
-
-The general pattern of the levels of service is similar before and during the pandemic, with lower values in the center of the city where low-income households exhibit multimodal trip patterns.
-Three more peripheral facilities towards the south of the city also have lower levels of service, presumably because they are expected to service relatively large suburban/exurban populations generally reliant on automobiles for travel.
-During the pandemic, however, the levels of service dropped, in some cases quite substantially.
-The pattern of the losses in level of service, moreover, is not uniform.
-The upper pane of Figure \ref{fig:levels-of-service-changes} shows that the three peripheral facilities in the southern suburban/exurban part of the city had low levels of service to begin with, but did not see major declines during the pandemic.
-Further, the inset map shows that the levels of service deteriorated more in the central part of the city.
-However, the loss of level of service was not as large in the core (where most of the food banks/services are found), but instead was more marked in the inner ring around the core, where facilities may have faced greater demand from both central city and suburban populations after the closure of food banks during the pandemic.
-
-```{r plot-levels-of-service, echo=FALSE, fig.cap="\\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
-## Analysis
-# Plot levels of service before and during covid
-
-# Pre-covid
-los_pre <- ggplot() +
-  geom_sf(data = hamilton_boundary) +
-  geom_sf(data = los %>%
-            filter(timing == "Pre-covid"),
-          aes(color = los,
-              size = los),
-          alpha = 0.5) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  scale_color_viridis_c(name = "Level of service \npre-COVID-19",
-                        option =  "A", 
-                        direction = -1,
-                        breaks = c(0.0005,
-                                   0.0010,
-                                   0.0015,
-                                   0.0020,
-                                   0.0025),
-                        labels = c("0.00050",
-                                   "0.00100",
-                                   "0.00150",
-                                   "0.00200",
-                                   "0.00250")) +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  guides(size = FALSE, shape = FALSE) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-# During covid
-los_covid <- ggplot() +
-  geom_sf(data = hamilton_boundary) +
-  geom_sf(data = los %>%
-            filter(timing == "During covid"),
-          aes(color = los,
-              size = los),
-          alpha = 0.5) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  scale_color_viridis_c(name = "Level of service \nduring COVID-19",
-                        option =  "A", 
-                        direction = -1,
-                        breaks = c(0.00050,
-                                   0.00075,
-                                   0.00100,
-                                   0.00125,
-                                   0.00150),
-                        labels = c("0.00050",
-                                   "0.00075",
-                                   "0.00100",
-                                   "0.00125",
-                                   "0.00150")) +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  guides(size = FALSE, shape = FALSE) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-los_pre / los_covid
-```
-
-```{r plot-levels-of-service-changes, echo=FALSE, fig.cap="\\label{fig:levels-of-service-changes}Changes in levels of service at each facility from pre-COVID-19 to during COVID-19."}
-# This figure shows the places where the level of service deteriorated most (excluding the places that closed, where it vanished).
-los_change <- los %>%
-  st_drop_geometry() %>%
-  pivot_wider(names_from = timing, 
-              values_from = los, 
-              values_fill = 0) %>%
-  filter(`During covid` > 0) %>%
-  mutate(los_d = `During covid` - `Pre-covid`) %>%
-  left_join(foodbank_locations %>%
-              transmute(OBJECTID = ID,
-                        geometry),
-            by = "OBJECTID") %>%
-  st_as_sf()
-
-lc <- ggplot() +
-  geom_sf(data = hamilton_boundary) +
-  geom_sf(data = los_change, 
-          aes(color = los_d,
-              size = los_d),
-          alpha = 0.3) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_color_viridis_c(name = "Change in level of service",
-                        option =  "A", 
-                        direction = 1,
-                        breaks = c(-0.0001,
-                                   -0.0002,
-                                   -0.0003,
-                                   -0.0004,
-                                   -0.0005),
-                        labels = c("-0.0001",
-                                   "-0.0002",
-                                   "-0.0003",
-                                   "-0.0004",
-                                   "-0.0005")) +
-  scale_size(trans = 'reverse') +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  guides(size = FALSE, shape = FALSE) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "none")
-
-lc_inset <- ggplot() +
-  geom_sf(data = hamilton_boundary %>% 
-            st_transform(4326)) +
-  geom_sf(data = los_change, 
-          aes(color = los_d,
-              size = los_d),
-          alpha = 0.7) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_color_viridis_c(name = "Change in level of service",
-                        option =  "A", 
-                        direction = 1,
-                        breaks = c(-0.0001,
-                                   -0.0002,
-                                   -0.0003,
-                                   -0.0004,
-                                   -0.0005),
-                        labels = c("-0.0001",
-                                   "-0.0002",
-                                   "-0.0003",
-                                   "-0.0004",
-                                   "-0.0005")) +
-  scale_size(name = "Change in level of service",
-             trans = 'reverse',
-             breaks = c(-0.0001,
-                        -0.0002,
-                        -0.0003,
-                        -0.0004,
-                        -0.0005),
-             labels = c("-0.0001",
-                        "-0.0002",
-                        "-0.0003",
-                        "-0.0004",
-                        "-0.0005")) +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  xlim(-79.975, -79.775) +
-  ylim(43.2, 43.3) +
-  guides(color = guide_legend(title.position = "top"), 
-         #size = FALSE, 
-         shape = FALSE) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "bottom")
-
-lc / lc_inset
-```
-
-To further elucidate this issue, we now turn to the results of the accessibility analysis.
-As with the level of service of individual facilities, the general pattern of accessibility before and during the pandemic is similar.
-Figure \ref{fig:accessibility} reveals that, compared with the outer rural zones, the more urban zones of the city generally exhibit higher accessibility to food banks and food service locations.
-However, the pattern is not particularly smooth - this is largely attributable to the weighting of travel times by mode of transportation according to the trip patterns of low-income household respondents captured by the TTS.
-For example, in zones where low-income households make a high proportion of trips by walking, access to food bank locations by walking is afforded a concomitantly high weight in our calculations of travel time compared to transit or car travel.
-From this, highly-accessible locations result from a mix of characteristics: low-income households in locations where travel options that align with zonal modal split are available to connect them to food bank locations with high levels of service within 15 minutes.
-This seems to track with the experience of some users of these services, as reported by Smith-Carrier et al.
-[-@smith2017food].
-
-```{r aggregate-accessibility-by-taz, include=FALSE}  
-## Analysis
-# Accessibility by TAZ:
-accessibility_taz <- accessibility %>%
-  group_by(TAZUID, 
-           timing) %>%
-  summarize(accessibility = sum(accessibility),
-            timing = first(timing),
-            .groups = "drop") %>%
-  left_join(population_age_income %>%
-              transmute(TAZUID,
-                        children,
-                        adults,
-                        seniors),
-            by = "TAZUID") %>%
-  st_as_sf()
-```
-
-```{r plot-accessibility, echo=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
-## Analysis
-# Plot accessibility before and during covid:
-acc_pre <- ggplot() +
-  geom_sf(data = hamilton_boundary,
-          fill = NA) +
-  geom_sf(data = accessibility_taz %>%
-            filter(timing == "Pre-covid"),
-          aes(fill = accessibility)) +
-  scale_fill_viridis_c(name = "Accessibility \npre-COVID-19",
-                       option =  "A", 
-                       direction = -1,
-                       breaks = c(0.002,
-                                  0.0016,
-                                  0.0012,
-                                  0.0008,
-                                  0.0004),
-                       labels = c("0.0020",
-                                  "0.0016",
-                                  "0.0012",
-                                  "0.0008",
-                                  "0.0004")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-acc_covid <- ggplot() +
-  geom_sf(data = hamilton_boundary,
-          fill = NA) +
-  geom_sf(data = accessibility_taz %>%
-            filter(timing == "During covid"),
-          aes(fill = accessibility)) +
-  scale_fill_viridis_c(name = "Accessibility \nduring COVID-19",
-                       option =  "A", 
-                       direction = -1,
-                       breaks = c(0.0010,
-                                  0.0008,
-                                  0.0006,
-                                  0.0004,
-                                  0.0002),
-                       labels = c("0.0010",
-                                  "0.0008",
-                                  "0.0006",
-                                  "0.0004",
-                                  "0.0002")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-acc_pre / acc_covid
-```
-
-```{r accessibility-changes-by-taz, include=FALSE}
-## Analysis
-# Calculate accessibility change by TAZ:
-accessibility_taz_change <- accessibility_taz %>%
-  st_drop_geometry() %>%
-  pivot_wider(names_from = timing, 
-              values_from = accessibility) %>%
-  # Calculate changes in accessibility
-  mutate(acc_d = `During covid` - `Pre-covid`) %>%
-  # Join TAZ geometry
-  left_join(population_age_income %>%
-              transmute(TAZUID,
-                        geometry),
-            by = "TAZUID") %>%
-  # Convert to simple features
-  st_as_sf()
-```
-
-We find that the accessibility landscape deteriorated substantially during the pandemic, with accessibility dropping on average by almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% mean() %>% abs() %>% round()`\%, but with large variations: some zones experienced changes in accessibility of only about `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% max() %>% abs() %>% round()`\%, whereas the most affected zone saw a loss of accessibility of almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% min() %>% abs() %>% round()`\%.
-Figure \ref{fig:accessibility-changes-with-local-i} shows the changes in accessibility.
-Every zone is worse off after the closure of facilities during the pandemic, but some parts of the city seem to have been particularly affected.
-To better highlight these changes, we used a local indicator of spatial autocorrelation [@anselin1995local] to explore the pattern of change in accessibility.
-Twenty-four TAZ are flagged as having significant large losses of accessibility (at $p<=0.10$, without correcting for multiple comparisons).
-Those zones are highlighted in the figure, where it can be seen that they form more or less compact neighborhoods.
-Remarkably, the largest (and significant) drops in accessibility are not downtown, but located in two cases in the industrial north of the city, in one case in an inner suburb above the escarpment, and lastly in a more suburban/exurban region in the south-west.
-
-```{r plot-accessibility-changes, echo=FALSE, include=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19."}
-## Analysis
-# Plot
-acc_change <- ggplot() +
-  #geom_sf(data = hamilton_boundary) +
-  geom_sf(data = accessibility_taz_change,
-          aes(fill = acc_d),
-          color = NA) +
-  geom_sf(data = population_age_income,
-          fill = NA) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_fill_viridis_c(name = "Change in accessibility",
-                       option =  "A", 
-                       direction = 1,
-                       breaks = c(-0.0008,
-                                  -0.0006,
-                                  -0.0004,
-                                  -0.0002),
-                       labels = c("-0.0008",
-                                  "-0.0006",
-                                  "-0.0004",
-                                  "-0.0002")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "none")
-
-acc_change_inset <- ggplot() +
-  geom_sf(data = accessibility_taz_change %>%
-            st_transform(crs = 4326),
-          aes(fill = acc_d)) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_fill_viridis_c(name = "Change in accessibility",
-                       option =  "A", 
-                       direction = 1,
-                       breaks = c(-0.0008,
-                                  -0.0006,
-                                  -0.0004,
-                                  -0.0002),
-                       labels = c("-0.0008",
-                                  "-0.0006",
-                                  "-0.0004",
-                                  "-0.0002")) +
-  guides(fill = guide_legend(title.position = "top")) +
-  xlim(-79.975, -79.775) +
-  ylim(43.2, 43.3) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "bottom")
-
-acc_change / acc_change_inset
-```
-
-```{r weights-matrix, include=FALSE}
-taz_nb <- poly2nb(accessibility_taz_change, queen = TRUE)
-taz_w <- nb2listw(taz_nb)
-```
-
-```{r moran-test-accessibility-changes, include=FALSE}
-moran.test(accessibility_taz_change$acc_d, taz_w)
-```
-
-```{r local-moran-i, include=FALSE}
-local_i <- localmoran(accessibility_taz_change$acc_d,
-                      taz_w, 
-                      alternative = "two.sided",
-                      p.adjust.method = "none") %>% 
-  data.frame() %>%
-  rename(p_val = Pr.z....0.) %>%
-  mutate(significance = ifelse(p_val <= 0.10, "sig", "n.s."),
-         significance = factor(significance,
-                               levels = c("n.s.", "sig")))
-
-summary(local_i)
-```
-
-```{r join-local-i-to-table, include=FALSE}
-accessibility_taz_change$significance <- local_i$significance
-```
-
-```{r plot-local-i, echo=FALSE, fig.cap="\\label{fig:accessibility-changes-with-local-i}Changes in accessibility from pre-COVID-19 to during COVID-19.
-Highlighted areas had significantly large changes in accessibility according to Local Moran's I." }
-ggplot() +
-  geom_sf(data = accessibility_taz_change,
-          aes(fill = acc_d,
-              color = significance,
-              alpha = significance),
-          size = 1) +
-  geom_sf(data = hamilton_boundary,
-          fill = NA) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_fill_viridis_c(name = "Change in accessibility",
-                       option =  "A", 
-                       direction = 1,
-                       breaks = c(-0.0008,
-                                  -0.0006,
-                                  -0.0004,
-                                  -0.0002),
-                       labels = c("-0.0008",
-                                  "-0.0006",
-                                  "-0.0004",
-                                  "-0.0002")) +
-  scale_color_manual(values = c("n.s." = "lightgray", 
-                                "sig" = "black")) +
-  scale_alpha_manual(values = c("n.s." = 0.5, 
-                                "sig" = 1)) +
-  guides(color = FALSE, 
-         alpha = FALSE,
-         fill = guide_legend(title.position = "top")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "bottom")
-```
-
-For the more suburban clusters of zones, the decrease in accessibility is derived from the closure of locations throughout the city reachable by car.
-In the cluster of central suburban zones for example, low-income households in the outer ring of zones that exhibit medium to high decreases in accessibility within this cluster appear to be largely auto-dependent in their tripmaking, which each each exhibiting between `r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(min(p_car)) *100)`-`r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(max(p_car)) *100)`\% of their modal split for car trips.
-This results in the parcels within these zones having a large number of potentially accessible locations in the travel time matrix.
-But by extension, the change in accessibility over the pre- and during-COVID-19 time periods is affected not only by the closure of service locations proximate to the zones, but also the locations in the central city.
-The zone with the greatest decrease in accessibility within this cluster (`r round(accessibility_taz_change %>% st_drop_geometry() %>% filter(TAZUID == 5135) %>% pull(acc_d), 4)`) has a high rate of car trips and connects to the most food bank locations in total as well as those that stayed open or closed.
-The exception to this is the zone in the middle of this cluster, which has a much higher proportion of transit use `r round(modes %>% st_drop_geometry() %>% filter(TAZUID == 5116) %>% summarize(first(p_transit)) *100)`\%.
-This modal split weighting results in fewer food bank and service locations accessible from the parcels within this zone which, by extension, limits their access to closed locations reachable within 15 minutes in our calculations.
-
-In the cluster to the south-west, the decrease in accessibility is predominately driven by the closure of a high level-of-service Community Meals provider.
-However, like the more central suburban zones, low-income households within this cluster also between 90% to 100% auto-dependent in their tripmaking in the TTS.
-The story is similar for the zone located in the north-west that exhibits the greatest decrease in accessibility.
-Here, low-income households responding to the TTS conducted 100% of their trips by car and, as a result, parcels within this zone have access to the second-highest number of food bank and service locations within 15 minutes.
-However, this also means zonal accessibilities are greatly affected by the number of closures throughout the city.
-Finally, in the city's north end and north-east zones, low-income households exhibit a mixture of tripmaking behaviour in the TTS.
-Households in some zones take transit more often and one zone in particular has 100\% of its trips by walking while in others, car use makes up 100\% of trips.
-For these zones, the decrease in accessibility tends to be a product of the closure of several inner-city food bank and service locations.
-
-Just as the effects of the closures appear to have been uneven in space, they also seem to have differently impacted various population segments.
-Table \ref{tab:accessibility-by-age} shows the estimated number of people in each age group by their level of accessibility before and during the pandemic.
-Here, it is important to note that the quartiles are relative: people in the top 25\% of accessibility still have lower accessibility during the pandemic than before.
-In reality, every population group is worse off during the pandemic in terms of their accessibility to food banks and services in the City of Hamilton.
-However, some age groups were affected more.
-As seen in the table, the distribution within quartiles for adults changed only slightly, and in fact there are fewer adults in the bottom three quartiles as some found themselves in the (deteriorated) top quartile of accessibility during the pandemic.
-The story is different for seniors, a greater number of whom are now in the third quartile, due to a combination of seniors facing a worse accessibility situation, and just a few moving from the bottom quartile (pre-pandemic) to the third during the pandemic.
-Those aged 18 and less appear to have fared worse, and the largest change is in the number of children who were in the third quartile before the pandemic found themselves in the bottom accessibility class during the pandemic.
-These results suggest that, through a combination of the typical modes of transportation of lower income households, and the spatial distribution of the population, the food bank/service locations closed had a differential impact that more greatly affected the youngest among the population in low income households.
-
-```{r accessibility-change-by-parcel, include=FALSE}
-## Analysis
-# Calculate accessibility change by parcel:
-accessibility_change <- accessibility %>%
-  select(UID, timing, accessibility) %>%
-  pivot_wider(names_from = timing, 
-              values_from = accessibility,
-              values_fill = 0) %>%
-  # Calculate changes in accessibility
-  mutate(acc_d = `During covid` - `Pre-covid`)
-```
-
-```{r join-accessibility-change-to-table, include=FALSE}
-## Analysis
-# Join accessibility change to accessibility table:
-accessibility <- accessibility %>%
-  left_join(accessibility_change %>%
-              select(UID, acc_d),
-            by = "UID")
-```
-
-```{r quartiles-of-accessibility-change, include=FALSE}
-## Analysis
-# Quartiles of accessibility pre-covid:
-acc_quantiles_pre <- quantile(accessibility_change$`Pre-covid`,
-                              c(0, 0.25, 0.5, 0.75, 1))
-# Quartiles of accessibility change:
-acc_quantiles_during <- quantile(accessibility_change$`During covid`,
-                                 c(0, 0.25, 0.5, 0.75, 1))
-```
-
-```{r accessibility-labels, include=FALSE}
-## Analysis
-# Add labels of accessibility loss to table:
-accessibility <- accessibility %>%
-  mutate(accessibility_pre_level = case_when(accessibility <= acc_quantiles_pre[2] ~ "Bottom 25\\%",
-                                             accessibility > acc_quantiles_pre[2] & accessibility <= acc_quantiles_pre[3] ~ "Third 25\\%",
-                                             accessibility > acc_quantiles_pre[3] & accessibility <= acc_quantiles_pre[4] ~ "Second 25\\%",
-                                             accessibility > acc_quantiles_pre[4] ~ "Top 25\\%"),
-         accessibility_pre_level = factor(accessibility_pre_level,
-                                          levels = c("Top 25\\%", "Second 25\\%", "Third 25\\%", "Bottom 25\\%"),
-                                          ordered = TRUE),
-         accessibility_during_level = case_when(accessibility <= acc_quantiles_during[2] ~ "Bottom 25\\%",
-                                                accessibility > acc_quantiles_during[2] & accessibility <= acc_quantiles_during[3] ~ "Third 25\\%",
-                                                accessibility > acc_quantiles_during[3] & accessibility <= acc_quantiles_during[4] ~ "Second 25\\%",
-                                                accessibility > acc_quantiles_during[4] ~ "Top 25\\%"),
-         accessibility_during_level = factor(accessibility_during_level,
-                                             levels = c("Top 25\\%", "Second 25\\%", "Third 25\\%", "Bottom 25\\%"),
-                                             ordered = TRUE))
-```
-
-```{r aggregate-accessibility-by-age-group, include=FALSE}
-summary_pre <- accessibility %>%
-  filter(timing == "Pre-covid") %>%
-  group_by(accessibility_pre_level) %>%
-  summarize(children_pre = sum(children) %>% 
-              round(),
-            adults_pre = sum(adults) %>% 
-              round(),
-            seniors_pre = sum(seniors) %>% 
-              round(),
-            .groups = "drop")
-
-summary_during <- accessibility %>%
-  filter(timing == "During covid") %>%
-  group_by(accessibility_during_level) %>%
-  summarize(children_during = sum(children) %>% 
-              round(),
-            adults_during = sum(adults) %>% 
-              round(),
-            seniors_during = sum(seniors) %>% 
-              round(),
-            .groups_during = "drop")
-
-accessibility_by_age <- cbind(summary_pre, summary_during[,2:4]) %>%
-  mutate(children_change = round((children_during - children_pre)/children_pre * 100, 2),
-         adults_change = round((adults_during - adults_pre)/adults_pre * 100, 2),
-         seniors_change = round((seniors_during - seniors_pre)/seniors_pre * 100, 2),
-         children_pre = prettyNum(children_pre, big.mark = ","),
-         adults_pre = prettyNum(adults_pre, big.mark = ","),
-         seniors_pre = prettyNum(seniors_pre, big.mark = ","),
-         children_during = paste0(prettyNum(children_during, big.mark = ","), " (", children_change, "\\%)"),
-         adults_during = paste0(prettyNum(adults_during, big.mark = ","), " (", adults_change, "\\%)"),
-         seniors_during = paste0(prettyNum(seniors_during, big.mark = ","), " (", seniors_change, "\\%)")) %>%
-  transmute(accessibility_pre_level,
-            children_pre,
-            adults_pre,
-            seniors_pre,
-            children_during,
-            adults_during,
-            seniors_during)
-```
-
-```{r table-accessibility-by-age-group, echo=FALSE, results='asis'}
-accessibility_by_age %>%
-  kable("latex",
-        booktabs = TRUE,
-        align = ("lcccccc"),
-        col.names = c("Accessibility Quartile",
-                      "Children (age $\\le$ 18)", 
-                      "Adults (19-64)", 
-                      "Seniors (age $\\ge$ 65)", 
-                      "Children (age $\\le$ 18)", 
-                      "Adults (19-64)", 
-                      "Seniors (age $\\ge$ 65)"),
-        caption = "\\label{tab:accessibility-by-age}Population at each accessibility level by age group among members households with incomes less than CAD40,000.",
-        escape = FALSE) %>%
-  kable_styling(latex_options = c("striped", "scale_down", "hold_position")) %>%
-  add_header_above(c(" " = 1, "Pre-COVID-19" = 3, "During COVID-19" = 3)) %>%
-  footnote(general = c("Population values have been rounded.", 
-                       "The values in brackets for population during COVID-19 are the changes from before the pandemic."))
-```
-
-```{r include=FALSE, eval = FALSE}
-## Analysis
-# Calculate accessibility by age group and accessibility level:
-accessibility_by_age <- accessibility %>%
-  filter(timing = "During covid")
-group_by(timing, 
-         accessibility_level) %>%
-  summarize(children = sum(children),
-            adults = sum(adults),
-            seniors = sum(seniors),
-            accessibility = sum(accessibility),
-            .groups = "drop")
-# The accessibility level is the magnitude of the accessibility loss: the top 25% faced the smallest losses in accessibility, and the bottom 25% faced the greatest losses in accessibility. 
-```
-
-```{r include=FALSE, eval = FALSE}
-cbind(accessibility_by_age %>% 
-        filter(timing == "Pre-covid") %>% 
-        select(children:accessibility) %>% 
-        rename(accessibility_pre = accessibility),
-      accessibility_by_age %>% 
-        filter(timing == "During covid") %>% 
-        transmute(accessibility_covid = accessibility)) %>%
-  mutate(pct_children = round(children/sum(children) * 100, 1),
-         pct_children = paste0("(", pct_children, "\\%)", sep = ""),
-         pct_adults = round(adults/sum(adults) * 100, 1),
-         pct_adults = paste0("(", pct_adults, "\\%)", sep = ""),
-         pct_seniors = round(seniors/sum(seniors) * 100, 1),
-         pct_seniors = paste0("(", pct_seniors, "\\%)", sep = ""),
-         children = round(children) %>% 
-           prettyNum(big.mark = ","),
-         adults = round(adults) %>% 
-           prettyNum(big.mark = ","),
-         seniors = round(seniors) %>% 
-           prettyNum(big.mark = ","),
-         children = paste0(children, " ", pct_children, sep = ""),
-         adults = paste0(adults, " ", pct_adults, sep = ""),
-         seniors = paste0(seniors, " ", pct_seniors, sep = ""),
-         change = (accessibility_covid - accessibility_pre) %>% 
-           round(5),
-         accessibility_pre = round(accessibility_pre, 5),
-         accessibility_covid = round(accessibility_covid, 5)) %>%
-  transmute(children,
-            adults,
-            seniors,
-            accessibility_pre,
-            accessibility_covid,
-            change) %>%
-  kable("latex",
-        booktabs = TRUE,
-        align = ("cccccc"),
-        col.names = c("Children (age $\\le$ 18)", 
-                      "Adults (19-64)", 
-                      "Seniors (age $\\ge$ 65)", 
-                      "Before COVID-19", 
-                      "During COVID-19", 
-                      "Difference"),
-        caption = "\\label{tab:accessibility-by-age}Accessibility by age group among members households with incomes less than CAD40,000.",
-        escape = FALSE) %>%
-  kable_styling(latex_options = c("striped", "scale_down")) %>%
-  add_header_above(c("Population" = 3, "Accessibility" = 3)) %>%
-  footnote(general = "Population values have been rounded")
-```
-
-Conclusions
-===================
-
-**LIMITATIONS**
-
-Words go here.
-
-References {#references .unnumbered}
-==========

### 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-title-page.tex
Added link to anonymous drive folder to share documents for review purposes; created title page

--- 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
@@
 % Pandoc citation processing
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc} % for calculating minipage widths
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 \usepackage[table]{xcolor}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
 
@@
   \title{Changes in accessibility to food banks and food services during
-COVID-19 and implications for the food security of vulnerable
-populations in Hamilton, Ontario}
-    \author[Some University]{Author One\corref{1}}
-   \ead{author1@example.com} 
-    \author[]{Author Two}
-   \ead{author2@example.com} 
-    \author[Some University]{Author Three\corref{2}}
-   \ead{author3@example.com} 
-    \author[Another University]{Author Four\corref{2}}
-   \ead{author4@example.com} 
-      \address[Some University]{Department, Street, City, State, Zip}
-    \address[Another University]{Department, Street, City, State, Zip}
+COVID-19 and implications for low income populations in Hamilton,
+Ontario}
+    \author[University of Toronto Scarborough]{Christopher D. Higgins}
+   \ead{cd.higgins@utoronto.ca} 
+    \author[McMaster University]{Antonio Paez\corref{1}}
+   \ead{paezha@mcmaster.ca} 
+    \author[University of Toronto Mississauga]{Gyoorie Kim}
+   \ead{gyoorie.kim@mail.utoronto.ca} 
+    \author[University of Toronto Mississauga]{Jue Wang}
+   \ead{gis.wang@utoronto.ca} 
+      \address[University of Toronto Scarborough]{Department of
+Geography \& Planning, University of Toronto Scarborough, 1265 Military
+Trail, Toronto, ON M1C1A4}
+    \address[McMaster University]{School of Earth, Environment and
+Society, McMaster University, Hamilton, ON, L8S 4K1, Canada}
+    \address[University of Toronto Mississauga]{Department of Geography,
+Geomatics and Environment, University of Toronto Mississauga, 3359
+Mississauga Road, Mississauga, ON, L5L 1C6 , Canada}
       \cortext[1]{Corresponding Author}
-    \cortext[2]{Equal contribution}
 
   \begin{abstract}
-  This is the abstract.
-
-  It consists of two paragraphs.
+  In this paper we analyze the changes in accessibility to food banks
+  and related services before and during the COVID-19 pandemic in the
+  City of Hamilton, Ontario. Food banks and services are the last line
+  of support for households facing food insecurity; as such, they have
+  become essential and their relevance cannot be ignored in the midst of
+  the economic upheaval caused by the pandemic. Our analysis is based on
+  the application of balanced floating catchment areas, and concentrates
+  on households with lower incomes (\$\textless CAD40,000, approximately
+  the Low Income Cutoff Value for a city of Hamilton's size).
+  Accessibility was low to begin with in suburban and exurban parts of
+  the city. We find that about 14\% of locations originally available in
+  Hamilton closed during the pandemic; the impact of those closures on
+  the level of service and accessibility was disproportionate, with
+  system-wide losses exceeding 40\%. Those losses were geographically
+  and demographically uneven. While every part of the city faced a
+  reduction in accessibility, inner suburbs fared worse in terms of loss
+  of accessibility. As well, children (age \(\le 18\)) appear to have
+  been impacted the most.
   \end{abstract}
@@
 Conceptualization, Investigation, Data Curation, Writing - Original
-Draft; \textbf{Author 4:} Resources - Writing: Review \& Editing -
+Draft; \textbf{Author 4:} Resources, Writing - Original Draft,
 Supervision
 
-\newpage
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Food insecurity is defined as an ``inadequate or uncertain access to a
-sufficient quantity and/or adequate quality of food'' due to a
-household's financial limitations (Enns et al., 2020). This condition
-has been associated with reductions in nutritional outcomes
-(Bhattacharya et al., 2004; Kirkpatrick and Tarasuk, 2008; Olson, 1999)
-and physical and mental health in children and adults (Elgar et al.,
-2021; Jones, 2017; Ramsey et al., 2011; Seligman et al., 2010; Stuff et
-al., 2004). Over at least the past four decades food banks and related
-services have become an essential line of defense against food
-insecurity in Canadian communities (Black and Seto, 2020; Holmes et al.,
-2018; Riches, 2002; Tarasuk et al., 2020). In this respect, Canada is
-not unlike numerous other wealthy countries where a systematic
-dismantling of the welfare state took place in the intervening period
-(Tarasuk et al., 2014).
-
-The emergence of COVID-19, the worst public health crisis since the 1918
-flu pandemic, has revealed important social and economic fault lines,
-and pre-existing patterns of inequality appear to have been exacerbated.
-Along several other dimensions of stress (e.g., accessibility to health
-care facilities, Pereira et al., 2021a), this seems to be the case for
-food insecurity as well (Laborde et al., 2020). In the US, for example,
-it has been estimated that there was an increase of more than 30\% in
-household food insecurity, and more than one third of households were
-discovered to be newly food insecure - meaning they did not experience
-food insecurity before the pandemic (Niles et al., 2020). In Canada, Men
-and Tarasuk (2021) report that about 25\% of individuals who experienced
-job insecurity (a relatively common occurrence during the pandemic),
-also experienced food insecurity. Similarly, according to Statistics
-Canada (2020a), in the early stages of the pandemic almost 15\% of
-individuals reported living in a household that faced food insecurity;
-the risk of food insecurity was substantially higher for households with
-children. The difference between households with and without children
-was significant, and 11.7\% of households with children indicated that
-``food didn't last and {[}there was{]} no money to get more'' sometimes
-or often, compared to 7.3\% of households without children; likewise,
-13\% of households with children indicated that they ``{[}c{]}ouldn't
-afford balanced meals'' sometimes or often, compared to 8.8\% of
-households without children.
-
-The impacts of food insecurity during the pandemic are alarming, since
-diet-related diseases, such as obesity, heart-disease, and diabetes,
-were already critical public health concerns in Canada prior to COVID-19
-(Boucher et al., 2017). While food banks are not necessarily a stable
-solution to food insecurity and in fact may encourage a retrenchment of
-neoliberal policy (Wakefield et al., 2013), at least they can be argued
-to provide a resource of last instance to households in precarious
-situations (Bazerghi et al., 2016). As recently as 2019, the Hamilton
-Hunger Report\footnote{https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf}
-noted that food banks in Hamilton, Ontario, recorded the highest number
-of visitors in the past 29 years; the number of children visiting food
-banks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278
-the year before, a rate of increase greater than population growth.
-
-It is known that urban food environments, within which people make their
-daily food choices, are essential in influencing eating behaviours and
-health outcomes, based on factors such as food availability, ease of
-geographic accessibility and socio-demographic variations (Paez et al.,
-2010; Vanderlee and L'Abbé, 2017; Widener, 2018). However, while there
-is a wealth of literature that has examined the topic of geographic
-accessibility to healthy food through the ``food desert'' concept, there
-has been little research into accessibility to food banks. Although
-previous work has explored differences in \emph{accessing} food banks,
-such as how some households utilize food banks over short periods of
-time while others regularly utilize food banks as longer-term resource
-(e.g., Enns et al., 2020), we are not aware of any research that has
-focused on the geographic component of accessibility.
-
-The study of place-based geographic accessibility is concerned with
-capturing the potential to reach destinations of value using the
-transportation network (Páez et al., 2012). Indeed, the Government of
-Canada's recent Food Policy has made ``access'' to healthy food a
-priority for Canadian communities\footnote{https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042}
-and previous survey research has suggested that such accessibility plays
-a key role in user satisfaction with food bank service delivery (Holmes
-et al., 2018). However, as with research into the prevalence of food
-deserts, accessibility to food banks is unlikely to be evenly
-distributed, and variation throughout a city can be expected due to
-transportation network characteristics, and the spatial distribution of
-food bank locations and the population they are meant to serve.
-Furthermore, policy responses to the COVID-19 pandemic likely have added
-to the distress of vulnerable households. Non-pharmaceutical
-interventions during the pandemic involving restrictions in mobility
-have increased the friction of travel, in particular by transit on which
-low income populations are more reliant (e.g., DeWeese et al., 2020). At
-the same time, the pandemic has created additional stress for the
-operators of food banks through disruptions in the supply chain (e.g.,
-McKay et al., 2021) as well as concerns surrounding the delivery of
-service in safe conditions and possible cancellation of food service
-programs.
-
-For this study, we aim to look at how the landscape of food banks and
-related services (e.g.~low-cost or free meal service providers)
-available in Hamilton, Ontario, changed during the pandemic. Did the
-number of open food bank services diminish? If so, what was the
-accessibility to food banks before the pandemic from the perspective of
-low income households, and how it changed during the pandemic? And
-finally, who are most likely to have been impacted by changes in the
-accessibility landscape? This paper first looks at the distribution of
-food banks and related services before and during the pandemic. Then, we
-use the balanced floating catchment area approach of Paez et al. (2019)
-to investigate the accessibility situation. For this, we use a fully
-disaggregated approach based on parcel-level data. Socio-economic and
-demographic data are drawn from the latest Census of Canada (2016),
-whereas travel information is from the most recent regional travel
-survey from 2016. This paper follows reproducible research
-recommendations (see Brunsdon and Comber, 2020), and the research was
-conducted using open source tools for transportation analysis (Lovelace,
-2021). The code and data necessary to reproduce the analysis are
-available in a public repository\footnote{add repository}.
-
-\hypertarget{food-insecurity-and-food-bank-use-in-canada}{%
-\section{Food Insecurity and Food Bank Use in
-Canada}\label{food-insecurity-and-food-bank-use-in-canada}}
-
-Food insecurity is the inability to acquire and consume an adequate
-amount or good quality food, leading to inadequate nutrient intake (Enns
-et al., 2020; Kirkpatrick and Tarasuk, 2008; Tarasuk and Vogt, 2009).
-This nutrient deficiency is associated with major population health
-concerns, particularly among Canadians at socio-economic disadvantage
-(Bazerghi et al., 2016). Data on food insecurity in Canada is collected
-in several ways. Quantitatively, official government surveys such as the
-Household Food Security Survey Module (HFSSM), the Canadian Community
-Health Surveys (CCHS), the International Study of Adults (LISA), and
-official classifications determined by Health Canada in relation to
-socio-demographic variables offer some insight into food insecurity
-(Gundersen et al., 2018; Kirkpatrick and Tarasuk, 2008; Tarasuk and
-Vogt, 2009). Nationally, it has been found that food insecurity impacts
-approximately 12.3\% of Canadian households (Tarasuk et al., 2014)
-
-Food banks - sometimes also referred to as `food pantries' and `food
-shelves' - originated as a community response to aid those with
-inadequate food by voluntarily offering them meals and ingredients
-(Loopstra and Tarasuk, 2012; Riches, 2002). The scope and objectives of
-food banks can vary by region and by country, and these organizations
-can include not only prepared meals and aliments, but also shared spaces
-to connect in community gardens and community kitchens (Wakefield et
-al., 2013). Although in their origin food banks were meant to be provide
-a temporary solution to accommodate those in hunger due to job
-retrenchments and economic downfalls since the 1980s, over time they
-have evolved into a community practice to secure emergency food supplies
-for those in need (Loopstra and Tarasuk, 2012; Wakefield et al., 2013).
-
-In Canada, the number of food banks has steadily increased in the past
-few decades (Wakefield et al., 2013). The largest database of food banks
-and their use comes from the non-profit association Food Banks Canada
-(FBC), which conducts an annual assessment through its affiliated
-members. FBC's 2018 Hunger Count report\footnote{https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018\_FINAL\_EN.pdf.aspx?ext=.pdf}
-(the most recent available) listed 1,830 member food banks across the
-country, and found that Canadians visited food banks 1.1 million times
-in March of 2018. Of those accessing food banks, certain population
-characteristics tend to be over-represented compared to national totals
-from the 2016 Canadian Census of Population. According to FBC's 2018
-data, single-adult households represent 45\% of those utilizing food
-banks despite making up 28\% of Canada's population, 19\% are
-single-parent households (compared to 10\% nationally), and 35\% of
-those accessing food bank services are children aged 0-18 even though
-their share of Canada's national population is approximately 20\%. In
-addition, 59\% of households accessing food banks list social or
-disability assistance as their primary source of income. Similarly,
-using data from the 2011-2012 CCHS, Tarasuk et al. (2019) found higher
-odds of food insecurity amongst households relying on social assistance,
-those without a university degree or with children under the age of 18,
-and individuals that lived alone, renters, and those identifying as
-Aboriginal. While surveys revealed that only 20 to 30 percent of those
-experiencing food insecurity were found to frequent food banks in Canada
-(Tarasuk et al., 2014), research from Ottawa (Enns et al., 2020) and
-Vancouver (Black and Seto, 2020) suggests that long-term users tend to
-be older, have health or mobility challenges, live in large households,
-and are less likely to have employment income.
-
-Economic stress caused by the COVID-19 pandemic, rising unemployment
-rates, and changes in poverty levels have disrupted the food environment
-and led to higher rates of food insecurity (Niles et al., 2020).
-Pandemic-related food security studies in the US have found a
-substantial increase in households experiencing food insecurity for the
-first time, and also in households experiencing more severe food
-insecurity than before (Niles et al., 2020; Wolfson and Leung, 2020).
-Most recently in May 2020, Canada recorded 14.7\% of its population
-living in food insecurity in the past 30 days (Statistics Canada,
-2020a). Considering the negative mental and physical health effects
-associated with food insecurity, increases in food insecurity rates due
-to the pandemic, signal a change in the food environment with potential
-damages to population health during the course of the pandemic and
-beyond (Niles et al., 2020). Recent data\footnote{https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report\_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis\_Exec-Sum\_EN.pdf}
-from FBC showed that 52\% of member food banks reported an increase in
-usage in March of 2020 when initial lockdown restrictions were put in
-place across much of the country. The pandemic also created significant
-staffing issues with 42\% of food banks reporting a reduction in
-volunteers. However, 53\% of food banks later reported a decrease in use
-into the summer of 2020 which FBC members attributed to emergency
-financial support programs from the federal government. Nevertheless,
-some of these benefit programs were temporary, which suggests that many
-households may again turn to food bank services to meet their needs.
-
-In terms of geography, previous research conducted at the provincial
-scale using data from the 2011-2012 CCHS found that the prevalence of
-food insecurity ranged across the country from 11.8\% of households in
-Ontario to 41\% of households in Nunavut (Tarasuk et al., 2019).
-However, although previous research has examined the characteristics of
-individuals and households accessing food banks, the locational or
-transportation accessibility aspect of food bank access is not well
-understood. A wealth of literature examining the food desert concept
-suggests that, in addition to socio-economic and demographic factors,
-location and transportation networks play a key role in a household's
-accessibility to healthy foods (Paez et al., 2010; Vanderlee and L'Abbé,
-2017; Widener, 2018). For food banks specifically, previous qualitative
-research in Ontario by Smith-Carrier et al. (2017) has noted that
-``transportation can be challenging, particularly if the food bank is
-situated in a remote location'' (p.~32). Particularly, it appears that
-participants experience challenges with the ``inordinate amount of time
-necessary to obtain food, and difficulties associated with
-transportation'' (p.~39). Users of food banks, according to this
-research, rely on a variety of modes of transportation to access
-services. Consequently, the location of facilities matters; in the words
-of an interviewee: ``I wish it {[}the food bank{]} was a little more
-centrally located. Because if I didn't have a bike I'd have to walk it
-all the way out there and back. I wonder about people who don't''
-(p.~39).
-
-To offer greater insight into the role of transportation and location in
-food bank accessibility, this research examines how geographic
-accessibility to food banks and food services changed in Hamilton during
-the COVID-19 pandemic.
-
-\hypertarget{methods-and-materials}{%
-\section{Methods and Materials}\label{methods-and-materials}}
-
-\hypertarget{methods}{%
-\subsection{Methods}\label{methods}}
-
-For the research in this paper we adopt the balanced floating catchment
-area approach of Paez et al. (2019). This method for estimating
-accessibility is a form of the widely-used two-stage floating catchment
-area method (Luo and Wang, 2003; Radke and Mu, 2000). Floating catchment
-areas are used to estimate accessibility when there are potential
-congestion effects, and operate by calculating first the \emph{demand}
-for spatially distributed services. The demand (usually the number of
-people who require a service) is used to calculate a level of service.
-In a second step, the level of service is allocated back to the
-population. Demand and level of service are allocated using some form of
-distance-decay to embody the geographical principle that, given a
-choice, people prefer to travel less than more when reaching
-destinations.
-
-More formally, the first step of this method is as follows: \[
-L_j = \frac{S_j}{\sum_{i=1}^nP_iw_{ij}}
-\]
-
-\noindent where \(S_j\) is the level of supply at location \(j\), in
-simplest terms whether a service point is present (i.e., \(S_j=1\)) or
-not (i.e., \(S_j=0\)); \(P_i\) is the population at location \(i\) that
-demands the service; and \(w_{ij}\) is a weight, typically a function of
-the distance between locations \(i\) and \(j\). \(L_j\) is the level of
-service at location \(j\) and it is the inverse of the number of people
-that need to be serviced.
-
-The second step in this process is then summing the level of service
-that each population unit can reach, according to the distance-decay
-weight: \[
-A_i = \sum_{j=1}^JL_jw_{ji}
-\]
-
-\noindent where \(A_i\) is the accessibility to the service, which is in
-the same units as the level of service: as the inverse of the population
-being serviced. When the population being serviced is low accessibility
-is high (i.e., there is little competition for the service), and
-viceversa.
-
-Floating catchment area methods are prone to overestimation of the
-population and the level of service due to multiple-counting. The
-population at \(P_i\) is allocated to \emph{every} service point \(j\)
-for which \(w_{ij}>0\). Similarly, the level of service at \(LOS_j\) is
-allocated to \emph{every} population point for which \(w_{ji}>0\). This
-inflation effect has been known for several years, and several
-modifications have been proposed to mitigate it (Delamater, 2013; e.g.,
-Wan et al., 2012). A definitive solution to this issue was presented by
-Paez et al. (2019). In order to avoid the multiple-counting in the
-summations, the population and the level of service need to be allocated
-\emph{proportionally}. This is achieved by standardizing the weights as
-follows: \[
-w_{ij}^\text{st} = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
-\] \noindent and: \[
-w_{ji}^\text{st} = \frac{w_{ji}}{\sum_{j=1}^Jw_{ji}}
-\]
-
-The standardized weights satisfy the following conditions: \[
-\sum_{i=1}^nw_{ij}^\text{st}=1
-\]
-
-\noindent and: \[
-\sum_{j=1}^Jw_{ji}^\text{st}=1
-\]
-
-Since the population is allocated proportionally, its value is
-preserved: \[
-\sum_{i=1}^nP_iw_{ij}^\text{st}=P_i
-\]
-
-\noindent as is the level of service: \[
-\sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
-\] \#\# Study Area
-
-\textbf{Need to provide some context here: the geography which makes
-transportation challenging, i.e., due to the escarpment; the main parts
-of the city: city core, inner suburbs, industrial north, suburban,
-exurban; and the industrial background of the city maybe, with poverty
-rates compared to other Canadian cities?}
-
-\hypertarget{data}{%
-\subsection{Data}\label{data}}
-
-Data have been prepared for sharing as a data package\footnote{add
-  repository}. The contents of the data package are described next.
-
-\hypertarget{statistics-canada}{%
-\subsubsection{Statistics Canada}\label{statistics-canada}}
-
-Population and income statistics for 2016 were retrieved at the level of
-Dissemination Areas (DAs) using the package \texttt{cancensus} (von
-Bergmann et al., 2021). DAs are the smallest publicly available census
-geography in Canada. Income data corresponds to the count of households
-by different total income groupings.
-
-\hypertarget{origins-residential-parcels}{%
-\subsubsection{Origins: Residential
-parcels}\label{origins-residential-parcels}}
-
-We converted all recorded residential land parcels in the City of
-Hamilton to points on the road network. Each point includes information
-about the number of residential units in the parcel. Next, we define
-low-income households as those having a total income of less than
-CAD40,000, which is approximately the mid-point of the low income
-cut-off (LICOs) for families in Canadian cities with populations greater
-than 500,000 in 2016, to match other Census data (Statistics Canada,
-2020b). We then ``populate'' each residential unit with the probability
-of being a low-income household based on the counts of households by
-income groups in the DA in which the parcel is located. While this
-method assumes a constant probability of low-income household status for
-all residential units in a DA, the parcel-level analysis affords a high
-level of spatial disaggregation for the accessibility analysis.
-
-\hypertarget{destinations-food-banks-and-food-service-locations}{%
-\subsubsection{Destinations: Food Banks and Food Service
-Locations}\label{destinations-food-banks-and-food-service-locations}}
-
-The locations of food banks and related food services were obtained from
-the Hamilton Public Library's Food Access Guide\footnote{http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf}.
-The guide was updated in April of 2021 to indicate any change affected
-on the services due to the pandemic. This includes modified business
-hours, a need to make reservations before frequenting, and locations
-that have completely shut down in consequence. While some food bank
-services have a specific a target population, such as prioritizing
-family with young children aged between 0 and 3 or accepting only those
-providing proof of low-income status through housing and utility
-statements, all the food bank services indicated below are designed to
-accommodate those in need of food at zero to low cost. With our focus on
-food banks and food services that offer free or low-cost meals at
-particular locations, we first removed services such as Meals on Wheels
-and other food access services such as food box, community kitchens,
-student nutrition programs, and shopping and transportation. In
-addition, two free meal services held on different days at the same
-location were collapsed into a single service point for the
-accessibility analysis. Additional details on the operations of
-individual facilities is not publicly available, and with the changes in
-operations it proved unfeasible to collect it. For this reason, the
-analysis to follow is of accessibility to the location of food banks and
-services, but not to specific services (e.g., breakfasts vs.~food
-boxes).
-
-\begin{table}
-
-\caption{\label{tab:table-food-bank-info}\label{tab:food-bank-info}foodbank and Food Service Information.}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{c>{\centering\arraybackslash}p{15em}cc>{\centering\arraybackslash}p{15em}}
-\toprule
-Type & Description & Locations Pre-COVID & Locations During COVID & Additional Notes\\
-\midrule
-\cellcolor{gray!6}{Congregate Dining} & \cellcolor{gray!6}{Congregate and dining programs provide low-cost meals that are enjoyed in a community setting.
-Transportation may be provided} & \cellcolor{gray!6}{7} & \cellcolor{gray!6}{2} & \cellcolor{gray!6}{One remaining location reduced hours during COVID}\\
-Community Meals & Programs often run by volunteers that organize suppers, lunches or other get-togethers that give community residents an opportunity to meet one another in a friendly and informal atmosphere while sharing a meal & 11 & 9 & NA\\
-\cellcolor{gray!6}{Food Banks} & \cellcolor{gray!6}{Food Banks and Emergency Food programs provide individuals and families with grocery items free of charge} & \cellcolor{gray!6}{27} & \cellcolor{gray!6}{26} & \cellcolor{gray!6}{One remaining location reduced hours during COVID while 4 others moved to appointment only}\\
-Free Meals & Meals are provided free of charge in the community through volunteer labour and donations & 9 & 5 & One remaining location reduced hours during COVID\\
-\cellcolor{gray!6}{Low-Cost Meals} & \cellcolor{gray!6}{Restaurants, cafeterias and other eating establishments operated by hospitals, senior centers or other organizations which provide reduced-cost meals for low-income people, older adults or other targeted individuals.} & \cellcolor{gray!6}{2} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{The remaining location reduced hours during COVID}\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\hypertarget{routing-and-travel-time-tables}{%
-\subsubsection{Routing and travel time
-tables}\label{routing-and-travel-time-tables}}
-
-Travel time tables for three modes (car, transit, walking) were computed
-using the parcels as the origins and the locations of the food banks as
-the destinations. For routing, the package \texttt{r5r} (Pereira et al.,
-2021b) was used with a network extract for the City of Hamilton from
-OpenStreetMaps and the General Transit Feed Specification from Hamilton
-Street Railway, the local transit operator. For routing purposes we used
-maximum values of 180 min and 10,000 m walking distance: any destination
-that exceeded these thresholds was ignored. The departure time used for
-routing was \textbf{TIME}.
-
-\hypertarget{transportation-tomorrow-survey}{%
-\subsubsection{Transportation Tomorrow
-Survey}\label{transportation-tomorrow-survey}}
-
-We used the Data Retrieval System of the Transportation Tomorrow Survey
-(TTS)\footnote{\url{http://dmg.utoronto.ca/}} to download
-cross-tabulations of: 1) primary mode of travel per trip by income by
-place of residence; and 2) age by income by place of residence. These
-data are from the 2016 Survey (the most recent available), and data are
-geocoded at the level of Traffic Analysis Zones (TAZ) using the most
-recent zoning system from 2006. Each parcel point is populated with the
-proportion of trips by three modes of travel: car (as driver or
-passenger), transit, and walk.
-
-\hypertarget{expected-travel-times}{%
-\subsubsection{Expected Travel Times}\label{expected-travel-times}}
-
-Once we obtained travel time tables with population (number of
-households) and proportion of trips by mode, we calculated the expected
-travel time \(ett\) from each parcel \(i\) to a food bank or food
-service location \(j\) as follows: \[
-ett_{ij} = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
-\]
-
-\noindent where \(p^k_i\) is the proportion of trips by mode \(k\) in
-the TAZ of parcel \(i\), and \(tt^k_{ij}\) is the travel time from
-parcel \(i\) to the food bank. In other words, the expected travel time
-is the weighted sum of travel times to the food bank, with the weights
-given by the expected modal split in the TAZ.
-
-\hypertarget{results-and-discussion}{%
-\section{Results and Discussion}\label{results-and-discussion}}
-
-Figure \ref{fig:foodbanks} shows the location of food banks and services
-in the City of Hamilton and their status. Before the pandemic there were
-58 of which 14 (24.14\%) closed during the pandemic. As shown in the
-figure, food services tend to be predominantly located in the central
-parts of the city. This is not surprising: population density is high
-there, and it is also the part of the city where lower income households
-are more numerous in absolute and relative terms (see Figure
-\ref{fig:low-income-households}). Alas, this is also the part of the
-city where most of the closures during the pandemic happened.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-location-foodbanks-1} \caption{\label{fig:foodbanks}Location of food banks/services and operation status; the dotted box is an inset of the central part of the City of Hamilton}\label{fig:plot-location-foodbanks}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-low-income-households-1} \caption{\label{fig:low-income-households}Number and proportion of households with incomes less than CAD40,000.}\label{fig:plot-low-income-households}
-\end{figure}
-
-To implement the accessibility calculations, we must select a
-distance-decay function. In this task we find limited support in the
-literature, which is mostly silent on the travel patterns of people who
-visit food banks. For this reason, we opt for a simple cumulative
-opportunities function as follows: \[
-w_{ij}=w_{ji}=
-\begin{cases}
-1 & \text{ if } ett_{ij}\le \delta\\
-0 & \text{ otherwise}
-\end{cases}
-\] \noindent where \(ett_{ij}\) is the multimodal expected travel time
-as described previously, and \(\delta\) is a travel threshold. When the
-expected travel time exceeds this threshold, a facility is no longer
-considered accessible. Moreover, the weights are standardized for the
-balanced floating catchment area approach.
-
-Figure \ref{fig:sensitivity-analysis} shows the results of conducting
-sensitivity analysis of the system-wide accessibility as we vary the
-threshold (considering the situation before the pandemic). There is a
-clear pattern whereby more strict values of \(\delta\) are associated
-with higher levels of system-wide accessibility: while increases in
-accessibility that result from decreases in the travel time window might
-seem counter-intuitive, this is a result of lower \emph{congestion},
-since fewer households are serviced and thus competition for the same
-resources is more limited. System-wide accessibility declines with
-higher values of \(\delta\): as more households are serviced, congestion
-grows and the level of service declines, although this happens at a
-declining rate. We are not aware of any research that explain how long
-people are expected to travel for food banks, but we note that in
-developing countries, access to drinking water is defined as a source
-that takes less than 30 minutes to reach (round trip, see UNICEF-WHO,
-2019). There is no reason why people in affluent countries should be
-expected to travel more for a basic need such as food. Accordingly, we
-adopt a 15-minute threshold for the analysis (representing a one-way
-trip). This threshold is also approximately where the rate of change in
-accessibility slows down.
-
-\begin{figure}
-\includegraphics[width=0.6\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-results-sensitivity-analysis-1} \caption{\label{fig:sensitivity-analysis}Accessibility as a function of threshold}\label{fig:plot-results-sensitivity-analysis}
-\end{figure}
-
-Using the 15-minute threshold, we find that the system-wide
-accessibility (interpreted as a provider-to-population ratio) was 0.065
-(food banks/service locations per low income household in the city)
-before COVID-19, but declined to 0.037 during the pandemic. It is
-striking that although almost 76\% of food facilities remained in
-operation during the pandemic, there was a loss of accessibility greater
-than 43\%, suggesting the location of food banks and related services
-plays an important role in serving those in need.
-
-Turning to the location of individual facilities, the levels of service
-offered before and during the pandemic are shown in Figure
-\ref{fig:levels-of-service}. The level of service is functionally the
-inverse of the number of low-income households in the travel-mode
-weighted travel time catchment area of the facilities (this is because
-\(S_j=1 \forall j\), i.e., each location represents a ``capacity'' of
-1). Higher values mean that a facility is expected to service fewer
-households. Conversely, lower values indicate grater congestion.
-
-The general pattern of the levels of service is similar before and
-during the pandemic, with lower values in the center of the city. Three
-more peripheral facilities towards the south of the city also have lower
-levels of service, presumably because they are expected to service
-relatively large suburban/exurban populations. During the pandemic,
-however, the levels of service dropped, in some cases quite
-substantially. The pattern of the losses in level of service, moreover,
-is not uniform. The upper pane of Figure
-\ref{fig:levels-of-service-changes} shows that the three peripheral
-facilities in the southern suburban/exurban part of the city had low
-levels of service to begin with, but did not see major declines during
-the pandemic. Further, the inset map shows that the levels of service
-deteriorated more in the central part of the city. However, the loss of
-level of service was not as large in the core (where most of the food
-banks/services are found), but instead was more marked in the inner ring
-around the core, where facilities may have faced greater demand from
-both central city and suburban populations after the closure of food
-banks during the pandemic.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-levels-of-service-1} \caption{\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (bottom panel).}\label{fig:plot-levels-of-service}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-levels-of-service-changes-1} \caption{\label{fig:levels-of-service-changes}Changes in levels of service at each facility from pre-COVID-19 to during COVID-19.}\label{fig:plot-levels-of-service-changes}
-\end{figure}
-
-To further elucidate this issue, we now turn to the results of the
-accessibility analysis. As with the level of service of individual
-facilities, the general pattern of accessibility before and during the
-pandemic is similar. Figure \ref{fig:accessibility} reveals that,
-compared with the outer rural zones, the more urban zones of the city
-generally exhibit higher accessibility to food banks and food service
-locations. However, the pattern is not particularly smooth - this is
-largely attributable to the weighting of travel times by mode of
-transportation according to the trip patterns of low-income household
-respondents according to the TTS. For example, in zones where low-income
-households make a high proportion of trips by walking, access to food
-bank locations by walking is afforded a concomitantly high weight in our
-calculations of travel time compared to transit or car travel. From
-this, highly-accessible locations result from a mix of characteristics:
-low-income households in locations where travel options that align with
-zonal modal split are available to connect them to food bank locations
-with high levels of service within 15 minutes. This seems to track with
-the experience of some users of these services, as reported by
-Smith-Carrier et al. (2017).
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-accessibility-1}
-
-}
-
-\caption{\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (bottom panel).}\label{fig:plot-accessibility}
-\end{figure}
-
-We find that the accessibility landscape deteriorated substantially
-during the pandemic, with accessibility dropping on average by almost
-41\%, but with large variations: some zones experienced changes in
-accessibility of only about 10\%, whereas the most affected zone saw a
-loss of accessibility of almost 71\%. Figure
-\ref{fig:accessibility-changes-with-local-i} shows the changes in
-accessibility. Every zone is worse off after the closure of facilities
-during the pandemic, but some parts of the city seem to have been
-particularly affected. We used a local indicator of spatial
-autocorrelation (Anselin, 1995) to explore the pattern of change in
-accessibility. Twenty-four TAZ are flagged as having significant large
-losses of accessibility (at \(p<=0.10\), without correcting for multiple
-comparisons). Those zones are highlighted in the figure, where it can be
-seen that they form more or less compact neighborhoods. Remarkably, the
-largest (and significant) drops in accessibility are not downtown, but
-located in two cases in the industrial north of the city, in one case in
-an inner suburb above the escarpment (\textbf{need to describe the study
-area!}), and lastly in a more suburban/exurban region in the south-west
-(\textbf{share other insights about these regions, maybe based on
-weighted travel patterns?}).
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-local-i-1} \caption{\label{fig:accessibility-changes-with-local-i}Changes in accessibility from pre-COVID-19 to during COVID-19.
-Highlighted areas had significantly large changes in accessibility according to Local Moran's I.}\label{fig:plot-local-i}
-\end{figure}
-
-Discuss the results of aggregating the accessibility by age group (see
-Table \ref{tab:accessibility-by-age}). Follow with the analysis of hours
-of travel?
-
-\begin{table}
-
-\caption{\label{tab:table-accessibility-by-age-group}\label{tab:accessibility-by-age}Accessibility by age group among members households with incomes less than CAD40,000.}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{cccccc}
-\toprule
-\multicolumn{3}{c}{Population} & \multicolumn{3}{c}{Accessibility} \\
-\cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){4-6}
-Children (age $\le$ 18) & Adults (19-64) & Seniors (age $\ge$ 65) & Before COVID-19 & During COVID-19 & Difference\\
-\midrule
-\cellcolor{gray!6}{3,445 (25\%)} & \cellcolor{gray!6}{11,860 (24.8\%)} & \cellcolor{gray!6}{6,584 (25.9\%)} & \cellcolor{gray!6}{0.00229} & \cellcolor{gray!6}{0.00173} & \cellcolor{gray!6}{-0.00057}\\
-4,477 (32.4\%) & 12,838 (26.9\%) & 6,847 (27\%) & 0.00707 & 0.00418 & -0.00289\\
-\cellcolor{gray!6}{3,541 (25.7\%)} & \cellcolor{gray!6}{13,189 (27.6\%)} & \cellcolor{gray!6}{7,020 (27.6\%)} & \cellcolor{gray!6}{0.01580} & \cellcolor{gray!6}{0.00911} & \cellcolor{gray!6}{-0.00669}\\
-2,340 (17\%) & 9,918 (20.7\%) & 4,940 (19.5\%) & 0.03985 & 0.02202 & -0.01783\\
-\bottomrule
-\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
-\multicolumn{6}{l}{\rule{0pt}{1em}Population values have been rounded}\\
-\end{tabular}}
-\end{table}
-
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
-
-\textbf{LIMITATIONS}
-
-Words go here.
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-anselin1995local}{}%
-Anselin, L., 1995. Local indicators of spatial association - LISA.
-Geographical Analysis 27, 93--115.
-
-\leavevmode\hypertarget{ref-bazerghi2016role}{}%
-Bazerghi, C., McKay, F.H., Dunn, M., 2016. The role of food banks in
-addressing food insecurity: A systematic review. Journal of community
-health 41, 732--740.
-
-\leavevmode\hypertarget{ref-bhattacharya2004poverty}{}%
-Bhattacharya, J., Currie, J., Haider, S., 2004. Poverty, food
-insecurity, and nutritional outcomes in children and adults. Journal of
-health economics 23, 839--862.
-doi:\url{https://doi.org/10.1016/j.jhealeco.2003.12.008}
-
-\leavevmode\hypertarget{ref-black2020examining}{}%
-Black, J.L., Seto, D., 2020. Examining patterns of food bank use over
-twenty-five years in vancouver, canada. VOLUNTAS: International Journal
-of Voluntary and Nonprofit Organizations 31, 853--869.
-doi:\url{https://doi.org/10.1007/s11266-018-0039-2}
-
-\leavevmode\hypertarget{ref-boucher2017ontario}{}%
-Boucher, B.A., Manafò, E., Boddy, M.R., Roblin, L., Truscott, R., 2017.
-The ontario food and nutrition strategy: Identifying indicators of food
-access and food literacy for early monitoring of the food environment.
-Health promotion and chronic disease prevention in Canada : research,
-policy and practice 37, 313--319.
-doi:\href{https://doi.org/10.24095/hpcdp.37.9.06}{10.24095/hpcdp.37.9.06}
-
-\leavevmode\hypertarget{ref-brunsdon2020opening}{}%
-Brunsdon, C., Comber, A., 2020. Opening practice: Supporting
-reproducibility and critical spatial data science. Journal of
-Geographical Systems 1--20.
-doi:\href{https://doi.org/10.1007/s10109-020-00334-2}{10.1007/s10109-020-00334-2}
-
-\leavevmode\hypertarget{ref-delamater2013spatial}{}%
-Delamater, P.L., 2013. Spatial accessibility in suboptimally configured
-health care systems: A modified two-step floating catchment area
-(M2SFCA) metric. Health \& Place 24, 30--43.
-doi:\href{https://doi.org/10.1016/j.healthplace.2013.07.012}{10.1016/j.healthplace.2013.07.012}
-
-\leavevmode\hypertarget{ref-deweese2020tale}{}%
-DeWeese, J., Hawa, L., Demyk, H., Davey, Z., Belikow, A., El-geneidy,
-A., 2020. A tale of 40 cities: A preliminary analysis of equity impacts
-of COVID-19 service adjustments across north america. Findings.
-doi:\href{https://doi.org/10.32866/001c.13395}{10.32866/001c.13395}
-
-\leavevmode\hypertarget{ref-elgar2021relative}{}%
-Elgar, F.J., Pickett, W., Pförtner, T.-K., Gariépy, G., Gordon, D.,
-Georgiades, K., Davison, C., Hammami, N., MacNeil, A.H., Da Silva, M.A.,
-others, 2021. Relative food insecurity, mental health and wellbeing in
-160 countries. Social Science \& Medicine 268, 113556.
-doi:\url{https://doi.org/10.1016/j.socscimed.2020.113556}
-
-\leavevmode\hypertarget{ref-enns2020experiences}{}%
-Enns, A., Rizvi, A., Quinn, S., Kristjansson, E., 2020. Experiences of
-food bank access and food insecurity in ottawa, canada. Journal of
-Hunger \& Environmental Nutrition 15, 456--472.
-doi:\url{https://doi.org/10.1080/19320248.2020.1761502}
-
-\leavevmode\hypertarget{ref-gundersen2018food}{}%
-Gundersen, C., Tarasuk, V., Cheng, J., De Oliveira, C., Kurdyak, P.,
-2018. Food insecurity status and mortality among adults in ontario,
-canada. PloS one 13, e0202642.
-
-\leavevmode\hypertarget{ref-holmes2018nothing}{}%
-Holmes, E., Black, J.L., Heckelman, A., Lear, S.A., Seto, D., Fowokan,
-A., Wittman, H., 2018. {``Nothing is going to change three months from
-now''}: A mixed methods characterization of food bank use in greater
-vancouver. Social Science \& Medicine 200, 129--136.
-doi:\url{https://doi.org/10.1016/j.socscimed.2018.01.029}
-
-\leavevmode\hypertarget{ref-jones2017food}{}%
-Jones, A.D., 2017. Food insecurity and mental health status: A global
-analysis of 149 countries. American journal of preventive medicine 53,
-264--273. doi:\url{https://doi.org/10.1016/j.amepre.2017.04.008}
-
-\leavevmode\hypertarget{ref-kirkpatrick2008food}{}%
-Kirkpatrick, S.I., Tarasuk, V., 2008. Food insecurity is associated with
-nutrient inadequacies among canadian adults and adolescents. The Journal
-of nutrition 138, 604--612.
-doi:\url{https://doi.org/10.1093/jn/138.7.1399}
-
-\leavevmode\hypertarget{ref-laborde2020poverty}{}%
-Laborde, D., Martin, W., Vos, R., 2020. Poverty and food insecurity
-could grow dramatically as COVID-19 spreads. International Food Policy
-Research Institute (IFPRI), Washington, DC.
-
-\leavevmode\hypertarget{ref-loopstra2012relationship}{}%
-Loopstra, R., Tarasuk, V., 2012. The relationship between food banks and
-household food insecurity among low-income toronto families. Canadian
-Public Policy 38, 497--514.
-doi:\href{https://doi.org/10.3138/cpp.38.4.497}{10.3138/cpp.38.4.497}
-
-\leavevmode\hypertarget{ref-lovelace2021open}{}%
-Lovelace, R., 2021. Open source tools for geographic analysis in
-transport planning. Journal of Geographical Systems.
-doi:\href{https://doi.org/10.1007/s10109-020-00342-2}{10.1007/s10109-020-00342-2}
-
-\leavevmode\hypertarget{ref-luo2003measures}{}%
-Luo, W., Wang, F.H., 2003. Measures of spatial accessibility to health
-care in a GIS environment: Synthesis and a case study in the chicago
-region. Environment and Planning B-Planning \& Design 30, 865--884.
-
-\leavevmode\hypertarget{ref-mckay2021exploring}{}%
-McKay, F.H., Bastian, A., Lindberg, R., 2021. Exploring the response of
-the victorian emergency and community food sector to the COVID-19
-pandemic. Journal of Hunger \& Environmental Nutrition 1--15.
-doi:\href{https://doi.org/10.1080/19320248.2021.1900974}{10.1080/19320248.2021.1900974}
-
-\leavevmode\hypertarget{ref-men2021food}{}%
-Men, F., Tarasuk, V., 2021. Food insecurity amid the COVID-19 pandemic:
-Food charity, government assistance and employment. Canadian Public
-Policy COVID-19, e2021001.
-doi:\href{https://doi.org/10.3138/cpp.2021-001}{10.3138/cpp.2021-001}
-
-\leavevmode\hypertarget{ref-niles2020early}{}%
-Niles, M.T., Bertmann, F., Belarmino, E.H., Wentworth, T., Biehl, E.,
-Neff, R., 2020. The early food insecurity impacts of COVID-19. Nutrients
-12, 2096.
-
-\leavevmode\hypertarget{ref-olson1999nutrition}{}%
-Olson, C.M., 1999. Nutrition and health outcomes associated with food
-insecurity and hunger. The Journal of nutrition 129, 521S--524S.
-doi:\url{https://doi.org/10.1093/jn/129.2.521S}
-
-\leavevmode\hypertarget{ref-paez2019demand}{}%
-Paez, A., Higgins, C.D., Vivona, S.F., 2019. Demand and level of service
-inflation in floating catchment area (FCA) methods. PloS one 14,
-e0218773.
-doi:\href{https://doi.org/10.1371/journal.pone.0218773}{10.1371/journal.pone.0218773}
-
-\leavevmode\hypertarget{ref-paez2010relative}{}%
-Paez, A., Mercado, R.G., Farber, S., Morency, C., Roorda, M., 2010.
-Relative accessibility deprivation indicators for urban settings:
-Definitions and application to food deserts in montreal. Urban Studies
-47, 1415--1438.
-doi:\href{https://doi.org/10.1177/0042098009353626}{10.1177/0042098009353626}
-
-\leavevmode\hypertarget{ref-paez2012measuring}{}%
-Páez, A., Scott, D.M., Morency, C., 2012. Measuring accessibility:
-Positive and normative implementations of various accessibility
-indicators. Journal of Transport Geography 25, 141--153.
-doi:\url{https://doi.org/10.1016/j.jtrangeo.2012.03.016}
-
-\leavevmode\hypertarget{ref-pereira2021geographic}{}%
-Pereira, R.H.M., Braga, C.K.V., Servo, L.M., Serra, B., Amaral, P.,
-Gouveia, N., Paez, A., 2021a. Geographic access to COVID-19 healthcare
-in brazil using a balanced float catchment area approach. Social Science
-\& Medicine 273, 113773.
-doi:\url{https://doi.org/10.1016/j.socscimed.2021.113773}
-
-\leavevmode\hypertarget{ref-pereira2021r5r}{}%
-Pereira, R.H.M., Saraiva, M., Herszenhut, D., Braga, C.K.V., Conway,
-M.W., 2021b. r5r: Rapid realistic routing on multimodal transport
-networks with r\textsuperscript{5} in r. Findings.
-doi:\href{https://doi.org/10.32866/001c.21262}{10.32866/001c.21262}
-
-\leavevmode\hypertarget{ref-radke2000spatial}{}%
-Radke, J., Mu, L., 2000. Spatial decomposition, modeling and mapping
-service regions to predict access to social programs. Annals of
-Geographic Information Sciences 6, 105--112.
-
-\leavevmode\hypertarget{ref-ramsey2011food}{}%
-Ramsey, R., Giskes, K., Turrell, G., Gallegos, D., 2011. Food insecurity
-among australian children: Potential determinants, health and
-developmental consequences. Journal of Child Health Care 15, 401--416.
-doi:\url{https://doi.org/10.1177\%2F1367493511423854}
-
-\leavevmode\hypertarget{ref-riches2002food}{}%
-Riches, G., 2002. Food banks and food security: Welfare reform, human
-rights and social policy. Lessons from canada? Social Policy \&
-Administration 36, 648--663.
-
-\leavevmode\hypertarget{ref-seligman2010food}{}%
-Seligman, H.K., Laraia, B.A., Kushel, M.B., 2010. Food insecurity is
-associated with chronic disease among low-income NHANES participants.
-The Journal of nutrition 140, 304--310.
-
-\leavevmode\hypertarget{ref-smith2017food}{}%
-Smith-Carrier, T., Ross, K., Kirkham, J., Decker Pierce, B., 2017.
-{`Food is a right\ldots{} nobody should be starving on our streets'}:
-Perceptions of food bank usage in a mid-sized city in ontario, canada.
-Journal of Human Rights Practice 9, 29--49.
-
-\leavevmode\hypertarget{ref-statisticscanada2020food}{}%
-Statistics Canada, 2020a. Food insecurity during the COVID-19 pandemic
-(No. Catalogue no. 45280001).
-
-\leavevmode\hypertarget{ref-statisticscanada2020licos}{}%
-Statistics Canada, 2020b. Table 11-10-0241-01 low income cut-offs
-(LICOs) before and after tax by community size and family size, in
-current dollars. doi:\url{https://doi.org/10.25318/1110024101-eng}
-
-\leavevmode\hypertarget{ref-stuff2004household}{}%
-Stuff, J.E., Casey, P.H., Szeto, K.L., Gossett, J.M., Robbins, J.M.,
-Simpson, P.M., Connell, C., Bogle, M.L., 2004. Household food insecurity
-is associated with adult health status. The Journal of nutrition 134,
-2330--2335.
-doi:\href{https://doi.org/Oxford\%20University\%20Press}{Oxford University Press}
-
-\leavevmode\hypertarget{ref-tarasuk2014food}{}%
-Tarasuk, V., Dachner, N., Loopstra, R., 2014. Food banks, welfare, and
-food insecurity in canada. British Food Journal.
-
-\leavevmode\hypertarget{ref-tarasuk2020relationship}{}%
-Tarasuk, V., Fafard St-Germain, A.-A., Loopstra, R., 2020. The
-relationship between food banks and food insecurity: Insights from
-canada. VOLUNTAS: International Journal of Voluntary and Nonprofit
-Organizations 31, 841--852.
-doi:\href{https://doi.org/10.1007/s11266-019-00092-w}{10.1007/s11266-019-00092-w}
-
-\leavevmode\hypertarget{ref-tarasuk2019geographic}{}%
-Tarasuk, V., Fafard St-Germain, A.-A., Mitchell, A., 2019. Geographic
-and socio-demographic predictors of household food insecurity in canada,
-2011--12. BMC Public Health 19, 12.
-doi:\href{https://doi.org/10.1186/s12889-018-6344-2}{10.1186/s12889-018-6344-2}
-
-\leavevmode\hypertarget{ref-tarasuk2009household}{}%
-Tarasuk, V., Vogt, J., 2009. Household food insecurity in ontario.
-Canadian Journal of Public Health 100, 184--188.
-doi:\href{https://doi.org/10.1007/BF03405537}{10.1007/BF03405537}
-
-\leavevmode\hypertarget{ref-world2019progress}{}%
-UNICEF-WHO, 2019. Progress on household drinking water, sanitation and
-hygiene 2000--2017. Special focus on inequalities.
-
-\leavevmode\hypertarget{ref-vanderlee2017food}{}%
-Vanderlee, L., L'Abbé, M., 2017. Food for thought on food environments
-in canada. Health promotion and chronic disease prevention in Canada :
-research, policy and practice 37, 263--265.
-doi:\href{https://doi.org/10.24095/hpcdp.37.9.01}{10.24095/hpcdp.37.9.01}
-
-\leavevmode\hypertarget{ref-vonBergmann2021cancensus}{}%
-von Bergmann, J., Shkolnik, D., Jacobs, A., 2021. Cancensus: R package
-to access, retrieve, and work with canadian census data and geography.
-
-\leavevmode\hypertarget{ref-wakefield2013sweet}{}%
-Wakefield, S., Fleming, J., Klassen, C., Skinner, A., 2013. Sweet
-charity, revisited: Organizational responses to food insecurity in
-hamilton and toronto, canada. Critical Social Policy 33, 427--450.
-
-\leavevmode\hypertarget{ref-wan2012three}{}%
-Wan, N., Zou, B., Sternberg, T., 2012. A three-step floating catchment
-area method for analyzing spatial access to health services.
-International Journal of Geographical Information Science 26,
-1073--1089.
-doi:\href{https://doi.org/10.1080/13658816.2011.624987}{10.1080/13658816.2011.624987}
-
-\leavevmode\hypertarget{ref-widener2018spatial}{}%
-Widener, M.J., 2018. Spatial access to food: Retiring the food desert
-metaphor. Physiology \& behavior 193, 257--260.
-doi:\url{https://doi.org/10.1016/j.physbeh.2018.02.032}
-
-\leavevmode\hypertarget{ref-wolfson2020food}{}%
-Wolfson, J.A., Leung, C.W., 2020. Food insecurity and COVID-19:
-Disparities in early effects for US adults. Nutrients 12, 1648.
-
-\end{CSLReferences}
-
 \end{document}

### 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Added link to anonymous drive folder to share documents for review purposes; created title page

@@
 This paper follows reproducible research recommendations [see @brunsdon2020opening], and the research was conducted using open source tools for transportation analysis [@lovelace2021open].
-The code and data necessary to reproduce the analysis are available in a public repository^[add repository].
+The code and data necessary to reproduce the analysis are available in a public repository^[https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing].
 
@@
 $$
+
 ## Study Area
@@
 
-Data have been prepared for sharing as a data package^[add repository]. The contents of the data package are described next.
+Data have been prepared for sharing as a data package^[https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing]. The contents of the data package are described next.
 
@@
 
-**LIMITATIONS**
-
 Words go here.

### 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Added link to anonymous drive folder to share documents for review purposes; created title page

@@
 \usepackage{makecell}
+\usepackage{xcolor}
 
@@
 2021). The code and data necessary to reproduce the analysis are
-available in a public repository\footnote{add repository}.
+available in a public repository\footnote{https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j\_M\_BMrXhXFvrV2c?usp=sharing}.
 
@@
 \sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
-\] \#\# Study Area
+\]
+
+\hypertarget{study-area}{%
+\subsection{Study Area}\label{study-area}}
 
@@
 
-Data have been prepared for sharing as a data package\footnote{add
-  repository}. The contents of the data package are described next.
+Data have been prepared for sharing as a data package\footnote{https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j\_M\_BMrXhXFvrV2c?usp=sharing}.
+The contents of the data package are described next.
 
@@
 
-\textbf{LIMITATIONS}
-
 Words go here.

### 2021-04-27 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
add conclusion and study area text

@@
 While food banks are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least they can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
-As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that food banks in Hamilton, Ontario, recorded the highest number of visitors in the past 29 years; the number of children visiting food banks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278 the year before, a rate of increase greater than population growth.
+As a mid-size city grappling with deindustrialization, Hamilton exhibits high rates of poverty and food bank use.
+As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that food banks in the city recorded the highest number of visitors in the past 29 years; a rate of increase greater than population growth.
+Most troubling, approximately 40% of all visitors were children.
 
@@
 However, 53% of food banks later reported a decrease in use into the summer of 2020 which FBC members attributed to emergency financial support programs from the federal government.
-Nevertheless, some of these benefit programs were temporary, which suggests that many households may again turn to food bank services to meet their needs.
+Nevertheless, some of these benefit programs were temporary.
+Although more recent statistics on food bank use in Hamilton in 2020 and 2021 is not yet available, data from the Daily Bread Food Bank in neighbouring Toronto for August 2020 shows visits climbing 51% year-over-year in that city^[https://www.dailybread.ca/wp-content/uploads/2020/11/Whos-Hungry-Report-2020.pdf], which suggests that many households in Hamilton are likely to turn to food bank services to meet their needs.
 
@@
 
-**Need to provide some context here: the geography which makes transportation challenging, i.e., due to the escarpment; the main parts of the city: city core, inner suburbs, industrial north, suburban, exurban; and the industrial background of the city maybe, with poverty rates compared to other Canadian cities?**
+With a population of around 540,000, the City of Hamilton is the fourth largest city in Ontario.
+It has historically been home to major manufacturing industries but de-industrialization that has occurred over the past several decades has led Hamilton to become one of the most highly divided cities in Ontario, with a significant proportion of its residents sustaining livings at Canada’s poverty level [@deluca2012code; @jakar2019turning; @latham2007determinants].
+In 2016, the Hamilton Community Foundation reported that based on the Low-Income Cut-Off, Hamilton recorded a poverty rate of 16.7%, which was well above the average rate of Ontario (13.7%) and the average national rate (12.8%).^[https://www.hamiltoncommunityfoundation.ca/vital-signs/low-income-2018/].
+According to data from Hamilton Food Share^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf], approximately 23,000 individuals accessed food banks in the city in March of 2019.
+Within this total is 9,125 visits by children (minors up to 18 years old), up from 8,278 the year before.
+Feed Ontario, the province's largest collective of hunger-relief organizations, found that on a per-capita basis, the level of need in the inner core of central Hamilton was second highest in Ontario^[https://feedontario.ca/research/report-hunger-map/].
+
+Geographically, the "old" city of Hamilton was amalgamated with several of its surrounding municipalities in 2001, with the city now featuring a mix of urban, suburban, exurban, and rural areas.
+Lower-cost housing proximate to the city's industrial north end has traditionally attracted immigrants and less-affluent residents compared to the city's wealthier suburbs.
+However, the decentralization of population from the inner core has led to challenges in transit connectivity to amenities and services and the proportion of auto users compared to transit users remains very high [@behan2008smart; @topalovic2012light].
+In addition, the city is separated geographically by the Niagara Escarpment.
+With sections of rocky cliff that approach 100m in height, the escarpment presents a significant challenge for promoting active travel and transport connections between "mountain" and "lower city" neighbourhoods.
+Taken together, the high level of food need, population locations, and transportation network characteristics combine to inform spatial accessibility to food banks and food services in the city.
 
@@
 With our focus on food banks and food services that offer free or low-cost meals at particular locations, we first removed services such as Meals on Wheels and other food access services such as food box, community kitchens, student nutrition programs, and shopping and transportation.
-In addition, two free meal services held on different days at the same location were collapsed into a single service point for the accessibility analysis.
+With some providers offering different food services at the same location (e.g.
+food bank with free and community meal services), and some of these services closing after the onset of the COVID-19 pandemic, we opted to geocode based on the service type.
+On the other hand, two free meal services held on different days at the same location were collapsed into a single service point for the accessibility analysis.
 Additional details on the operations of individual facilities is not publicly available, and with the changes in operations it proved unfeasible to collect it.
@@
 
-Words go here.
-
+Food insecurity is a significant issue for many Canadian households and is associated reductions in nutrition and mental and physical health.
+While food bank use can provide some relief, levels of need in the City of Hamilton were on the rise before the onset of the COVID-19 pandemic with a troubling increase in the number of visits made by children.
+While temporary financial relief measures helped reduce food bank use in the summer of 2020, the pandemic has more than likely increased food insecurity for many households and created a surge in food bank demand in the city.
+On the other hand, the pandemic has also resulted in major disruptions to food bank and food service delivery, resulting in the closure or cancellation of many food services in Hamilton.
+On top of this, reductions in transit service have reduced travel options for many.
+
+In response, this research has sought to better understand accessibility to food banks and food service locations as well as how the closure of some locations over the pre- and during-COVID time periods affected the potential for low-income households to reach these amenities.
+Using the balanced floating catchment area method to account for population demand and congestion effects at service points, we estimate multi-modal accessibility to food bank and food service locations for low-income households.
+In the absence of information on how food insecure households travel to food banks and related services, we adopted a normative approach that examines accessibility to food banks reachable within 15 minutes based on global standards for accessible drinking water.
+The weighting of travel time estimates by the modal split in different zonal geographies tailors the results to patterns of travel behaviour captured in the TTS regional travel survey.
+Moreover, our parcel-level analysis presents a disaggregate approach to estimating accessibility based on the locations of residential parcels and dwellings.
+
+Results show that while accessibility levels were lower in the city's more car-oriented suburban and rural areas to begin with, the closure of 14% of the city's food bank and low-cost or free food service locations during the pandemic resulted in an overall decrease in accessibility across the city.
+However, these effects were not uniform over space or for different population groups.
+Because the balanced floating catchment area method takes into account changes in demand and congestion for service providers, the closure of food bank and food service locations can reverberate throughout the catchment areas of the city.
+For some suburban zones, the closure of a relatively high level-of-service location results in the remaining services being spread over a larger population.
+In others, high auto dependence for trips leads to decreases in accessibility that accumulate due to the loss of several food bank and service locations initially reachable within 15 minutes by car.
+Reductions in accessibility in the city's more urban north end, where low-income households conduct higher proportions of trips by transit and walking, emphasize the importance of geographic proximity in the potential to reach food bank and service locations for these residents.
+Beyond geography, the results also highlight the differential impact of food bank closing during COVID-19 on population groups with seniors and children most affected.
+
+Still, the work is not without its limitations.
 We use data from the census to model low-income households as a proxy for households that are possibly experiencing food insecurity.
 However, the degree to which our low-income cut-off of CAD\$40,000 reflects food insecurity in the different zones of the study area is not known.
-
-While the travel survey allows us to model multi-modal accessibilities that align with travel behaviour observed in the travel survey and capture differences in accessibility by age categories, the use of travel survey data for modelling food insecurity also has its limitations.
+We also consider all food banks and food service locations equally.
+Future work could improve this method by accounting for differential capacities at food provider locations, such as the number of meals served.
+Unforunately, this data was not available in the present case as some service locations had already closed.
+We also did not explicitly account for any changes in demand over the study period.
+If trends in Hamilton are similar to those in Toronto, there could be a significant increase in demand that would alter the estimated accessibilities.
+Moreover, while the travel survey allows us to model multi-modal accessibilities that align with travel behaviour observed in the travel survey and capture differences in accessibility by age categories, the use of travel survey data for modelling food insecurity also has its limitations.
 Research into the population weighting methods used in the TTS note that the survey may under-count the lowest- and highest-income households in the survey study region, although the magnitude of this under-counting is unknown as approximately 20% of respondents to the 2016 survey did not report their income [@rose2018transportation].
@@
 
-In the absence of information on how food insecure households travel to food banks and related services, we adopted a normative approach that examines accessibility to food banks reachable within 15 minutes based on standards for accessible drinking water...
+Nevertheless, while previous work has noted the important role of geography alongside other socioeconomic and demographic indicators in household access to healthy food, this is, to the best of our knowledge, the first study that has focused on the geographic component of food bank accessibility.
+Beyond accounting for the inflation of demand that occurs in traditional floating catchment area methods, our application of the balanced floating catchment area approach offers a novel analysis of food bank and food service accessibility that is sensitive to the locations of low-income households, information on their age distributions and typical tripmaking behaviour, food bank and food service locations and their operations over time, and the characteristics of the city's multi-modal transportation network.
+Findings from this analysis can provide quantitative evidence to help urban planners and community administrators strategically allocate food bank services considering accessibility for different population groups, especially when considering the unequal distribution of impacts from the COVID-19 pandemic.
 

### 2021-04-27 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-abstract.Rmd
Created abstract file

--- 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
@@
 Food banks and services are the last line of support for households facing food insecurity; as such, they have become essential and their relevance cannot be ignored in the midst of the economic upheaval caused by the pandemic.
-Our analysis is based on the application of balanced floating catchment areas, and concentrates on households with lower incomes ($<CAD40,000, approximately the Low Income Cutoff Value for a city of Hamilton's size).
+Our analysis is based on the application of balanced floating catchment areas, and concentrates on households with lower incomes (<CAD40,000, approximately the Low Income Cutoff Value for a city of Hamilton's size).
 Accessibility was low to begin with in suburban and exurban parts of the city.
@@
 **Author 1:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization, Supervision; **Author 2:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization; **Author 3:** Conceptualization, Investigation, Data Curation, Writing - Original Draft; **Author 4:** Resources, Writing - Original Draft, Supervision
-
-```{r, setup, include=FALSE}
-knitr::opts_chunk$set(
-  cache = TRUE, out.width = "100%", align = 'center', fig.width = 7, fig.height = 7
-)
-```
-
-```{r install-data-package, include = FALSE}
-# Run only once if needed to install data package 
-if (!require("foodbanksHamilton", character.only = TRUE)) {
-  devtools::install_github("https://github.com/paezha/Accessibility-Food-Banks-Hamilton", subdir = "foodbanksHamilton")
-}
-```
-
-```{r load-packages, cache=FALSE, include=FALSE}
-library(foodbanksHamilton) # Data package
-library(kableExtra)
-library(patchwork)
-library(sf)
-library(spdep)
-library(tidyverse)
-#library(tmap)
-```
-
-```{r invoke-data, include=FALSE}
-data(data_da_2016)
-data(dwelling_network_points_2016)
-data(foodbank_locations)
-data(foodbank_table)
-data(hamilton_boundary)
-data(modes_less40k)
-data(population_age_income)
-data(ttm_car)
-data(ttm_transit)
-data(ttm_walk)
-```
-
-```{r remove-irrelevant-destinations, include=FALSE}
-## Data preparation
-# Remove entries in travel time matrices for inapplicable destinations
-# As a first step, we will remove the computed travel times to the three Community Kitchen locations (IDs 1-3):
-ttm_car <- ttm_car %>%
-  filter(!between(OBJECTID, left = 1, right = 3))
-
-ttm_transit <- ttm_transit %>%
-  filter(!between(OBJECTID, left = 1, right = 3))
-
-ttm_walk <- ttm_walk %>%
-  filter(!between(OBJECTID, left = 1, right = 3))
-```
-
-```{r join-low-income-households, include=FALSE}
-## Data preparation
-# Join proportion of low income households per DA to dwellings and total number of residential units:
-dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
-  left_join(data_da_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::transmute(DAUID = GeoUID, 
-                               p_inc_less40k),
-            by = "DAUID")
-```
-
-```{r probability-of-low-income-per-parcel, include=FALSE}
-## Data preparation
-# Calculate probability that a point on the network will have low income households.
-This is the proportion of low income households in the DA multiplied by the total number of residential units at the point:
-dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
-  mutate(income_less40k = p_inc_less40k * Dwellings)
-```
-
-```{r number-of-residential-units-per-DA, include=FALSE}
-## Data preparation
-# Prepare and join population data from Transportation Tomorrow Survey
-# Find number of residential units in each DA. This is needed to assign the population to parcels
-residential_units <- dwelling_network_points_2016 %>%
-  st_drop_geometry() %>%
-  group_by(TAZUID) %>%
-  summarize(residential_units = sum(Dwellings))
-```
-
-```{r join-population-to-parcels, include=FALSE}
-## Data preparation
-# Join number of low income households per DA to dwellings and total number of residential units:
-dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
-  left_join(population_age_income %>%
-              st_drop_geometry() %>%
-              dplyr::transmute(TAZUID,
-                               children,
-                               adults,
-                               seniors),
-            by = "TAZUID") %>%
-  left_join(residential_units,
-            by = "TAZUID")
-```
-
-```{r calculate-low-income-households-per-parcel, include=FALSE}
-## Data preparation
-# Calculate population per point on the network.
-This is the population in the DA divided by the total number of residential units.
-This gives the average number of people per residential unit.
-Then, multiply by the number of dwellings (residential units) at the point:
-dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
-  mutate(children = (children/residential_units) * Dwellings,
-         adults = (adults/residential_units) * Dwellings,
-         seniors = (seniors/residential_units) * Dwellings)
-```
-
-```{r join-household-data-to-travel-time-tables, include=FALSE}
-## Data preparation
-# Join the household data to the travel time matrices:
-ttm_car <- ttm_car %>%
-  left_join(dwelling_network_points_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            income_less40k,
-                            children,
-                            adults,
-                            seniors,
-                            DAUID, 
-                            TAZUID),
-            by = c("UID" = "ID"))
-
-ttm_transit <- ttm_transit %>%
-  left_join(dwelling_network_points_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            income_less40k,
-                            children,
-                            adults,
-                            seniors,
-                            DAUID, 
-                            TAZUID),
-            by = c("UID" = "ID"))
-
-ttm_walk <- ttm_walk %>%
-  left_join(dwelling_network_points_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            income_less40k,
-                            children,
-                            adults,
-                            seniors,
-                            DAUID, 
-                            TAZUID),
-            by = c("UID" = "ID"))
-```
-
-```{r join-foodbank-data-to-travel-time-tables, include=FALSE}
-## Data preparation
-# Join the foodbank data to the travel time tables:
-ttm_car <- ttm_car %>%
-  left_join(foodbank_locations %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            TYPE,
-                            COVIDClose),
-            by = c("OBJECTID" = "ID"))
-
-ttm_transit <- ttm_transit %>%
-  left_join(foodbank_locations %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID, 
-                            TYPE,
-                            COVIDClose),
-            by = c("OBJECTID" = "ID"))
-
-ttm_walk <- ttm_walk %>%
-  left_join(foodbank_locations %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID,
-                            TYPE,
-                            COVIDClose),
-            by = c("OBJECTID" = "ID"))
-```
-
-```{r drop-nas-from-travel-time-tables, include=FALSE}
-## Data preparation
-# Drop NAs:
-ttm_car <- ttm_car %>%
-  drop_na(TAZUID,
-          income_less40k)
-
-ttm_transit <- ttm_transit %>%
-  drop_na(TAZUID,
-          income_less40k)
-
-ttm_walk <- ttm_walk %>%
-  drop_na(TAZUID,
-          income_less40k)
-```
-
-```{r proportion-of-trips-by-mode-by-taz, include=FALSE}
-## Data preparation
-# Calculate proportion of trips by mode per TAZ:
-modes <- modes_less40k %>%
-  mutate(p_car = Driver/(Driver + Transit + Walk),
-         p_transit = Transit/(Driver + Transit + Walk),
-         p_walk = Walk/(Driver + Transit + Walk))
-```
-
-```{r fill-nas-transport-modes, include=FALSE}
-## Data preparation
-# Assume that TAZ with NAs have a proportion of car trips of one:
-modes <- modes %>%
-  dplyr::mutate(p_car = replace_na(p_car, 1),
-                p_transit = replace_na(p_transit, 0),
-                p_walk = replace_na(p_walk, 0))
-```
-
-```{r join-trips-by-mode-to-travel-time-tables, include=FALSE}
-## Data preparation
-# Join trip-making modal proportions
-ttm_car <- ttm_car %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-
-ttm_transit <- ttm_transit %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-
-ttm_walk <- ttm_walk %>%
-  left_join(modes %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID,
-                        p_car,
-                        p_transit,
-                        p_walk),
-            by = "TAZUID")
-```
-
-```{r join-travel-time-tables, include=FALSE}
-## Data preparation
-# Join Tables into Single TTM
-ttm <- ttm_car %>%
-  rename(travel_time_car = travel_time) %>%
-  full_join(ttm_transit %>%
-              transmute(UID,
-                        OBJECTID,
-                        travel_time_transit = travel_time),
-            by = c("UID", "OBJECTID")) %>%
-  full_join(ttm_walk %>%
-              transmute(UID,
-                        OBJECTID,
-                        travel_time_walk = travel_time),
-            by = c("UID", "OBJECTID"))
-```
-
-```{r travel-time-tables-data-check, include=FALSE}
-## Data preparation
-# Check the NAs in travel time table:
-ttm %>% 
-  filter(is.na(travel_time_transit)) %>%
-  select(travel_time_transit,
-         p_transit) %>%
-  summary()
-
-# What do NAs here mean?
-Most DAs missing travel time by transit do not report trips by transit in the TTS.
-In some DAs there are trips by transit (p_transit != 0), but no foodbank can be reached by transit (travel_time_transit is NA).
-How many of these cases?
-ttm %>% 
-  filter(is.na(travel_time_transit) & p_transit > 0) %>%
-  select(travel_time_transit,
-         p_transit) %>%
-  summary()
-
-# It should be fine to replace the NAs with zeros, which multiplied by p_transit will return a zero (i.e., it is not possible to reach a foodbank by transit).
-ttm <- ttm %>% 
-  mutate(travel_time_transit = replace_na(travel_time_transit, 0))
-
-# Check the NAs in travel time by walk:
-ttm %>% 
-  filter(is.na(travel_time_walk)) %>%
-  select(travel_time_walk,
-         p_walk) %>%
-  summary()
-
-# Most DAs missing travel time by walking do not report trips by walking in the TTS.
-In some DAs there are trips by walking (p_walk != 0), but no foodbank can be reached by walk (travel_time_walk is NA).
-How many of these cases?
-ttm %>% 
-  filter(is.na(travel_time_walk) & p_walk > 0) %>%
-  select(travel_time_walk,
-         p_walk) %>%
-  summary()
-
-# Like with transit, it should be fine to replace the NAs with zeros, which multiplied by p_walk will return a zero (i.e., it is not possible to reach a foodbank by walking).
-ttm <- ttm %>% 
-  mutate(travel_time_walk = replace_na(travel_time_walk, 0))
-
-#Summary of the table after these operations:
-summary(ttm)
-```
-
-```{r calculate-expected-travel-times, include=FALSE}
-## Data preparation
-# Calculate expected travel times by mode.
-These are the travel time by the mode multiplied by the probability of using the mode (which is proxied by the proportion of trips by mode in the DA from the TTS).
-The total expected travel time is the sum of the expected travel times by mode:
-ttm <- ttm %>%
-  mutate(e_travel_time_car = travel_time_car * p_car,
-         e_travel_time_transit = travel_time_transit * p_transit,
-         e_travel_time_walk = travel_time_walk * p_walk,
-         e_travel_time = e_travel_time_car + e_travel_time_transit + e_travel_time_walk)
-```
-
-\newpage
-
-Introduction
-==========================
-
-Food insecurity is defined as an "inadequate or uncertain access to a sufficient quantity and/or adequate quality of food" due to a household's financial limitations [@enns2020experiences].
-This condition has been associated with reductions in nutritional outcomes [@bhattacharya2004poverty; @kirkpatrick2008food; @olson1999nutrition] and physical and mental health in children and adults [@elgar2021relative; @jones2017food; @ramsey2011food; @seligman2010food; @stuff2004household].
-Over at least the past four decades food banks and related services have become an essential line of defense against food insecurity in Canadian communities [@riches2002food; @holmes2018nothing; @black2020examining; @tarasuk2020relationship].
-In this respect, Canada is not unlike numerous other wealthy countries where a systematic dismantling of the welfare state took place in the intervening period [@tarasuk2014food].
-
-The emergence of COVID-19, the worst public health crisis since the 1918 flu pandemic, has revealed important social and economic fault lines, and pre-existing patterns of inequality appear to have been exacerbated.
-Along several other dimensions of stress [e.g., accessibility to health care facilities, @pereira2021geographic], this seems to be the case for food insecurity as well [@laborde2020poverty].
-In the US, for example, it has been estimated that there was an increase of more than 30\% in household food insecurity, and more than one third of households were discovered to be newly food insecure - meaning they did not experience food insecurity before the pandemic [@niles2020early].
-In Canada, Men and Tarasuk [-@men2021food] report that about 25\% of individuals who experienced job insecurity (a relatively common occurrence during the pandemic), also experienced food insecurity.
-Similarly, according to Statistics Canada [-@statisticscanada2020food], in the early stages of the pandemic almost 15\% of individuals reported living in a household that faced food insecurity; the risk of food insecurity was substantially higher for households with children.
-The difference between households with and without children was significant, and 11.7\% of households with children indicated that "food didn’t last and [there was] no money to get more" sometimes or often, compared to 7.3\% of households without children; likewise, 13\% of households with children indicated that they "[c]ouldn’t afford balanced meals" sometimes or often, compared to 8.8\% of households without children.
-
-The impacts of food insecurity during the pandemic are alarming, since diet-related diseases, such as obesity, heart-disease, and diabetes, were already critical public health concerns in Canada prior to COVID-19 [@boucher2017ontario].
-While food banks are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least they can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
-As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that food banks in Hamilton, Ontario, recorded the highest number of visitors in the past 29 years; the number of children visiting food banks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278 the year before, a rate of increase greater than population growth.
-
-It is known that urban food environments, within which people make their daily food choices, are essential in influencing eating behaviours and health outcomes, based on factors such as food availability, ease of geographic accessibility and socio-demographic variations [@paez2010relative; @vanderlee2017food; @widener2018spatial].
-However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to food banks.
-Although previous work has explored differences in _accessing_ food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource [e.g., @enns2020experiences], we are not aware of any research that has focused on the geographic component of accessibility.
-
-The study of place-based geographic accessibility is concerned with capturing the potential to reach destinations of value using the transportation network [@paez2012measuring].
-Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities^[https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with food bank service delivery [@holmes2018nothing].
-However, as with research into the prevalence of food deserts, accessibility to food banks is unlikely to be evenly distributed, and variation throughout a city can be expected due to transportation network characteristics, and the spatial distribution of food bank locations and the population they are meant to serve.
-Furthermore, policy responses to the COVID-19 pandemic likely have added to the distress of vulnerable households.
-Non-pharmaceutical interventions during the pandemic involving restrictions in mobility have increased the friction of travel, in particular by transit on which low income populations are more reliant [e.g., @deweese2020tale].
-At the same time, the pandemic has created additional stress for the operators of food banks through disruptions in the supply chain [e.g., @mckay2021exploring] as well as concerns surrounding the delivery of service in safe conditions and possible cancellation of food service programs.
-
-For this study, we aim to look at how the landscape of food banks and related services (e.g.
-low-cost or free meal service providers) available in Hamilton, Ontario, changed during the pandemic.
-Did the number of open food bank services diminish?
-If so, what was the accessibility to food banks before the pandemic from the perspective of low income households, and how it changed during the pandemic?
-And finally, who are most likely to have been impacted by changes in the accessibility landscape?
-This paper first looks at the distribution of food banks and related services before and during the pandemic.
-Then, we use the balanced floating catchment area approach of Paez et al.
-[-@paez2019demand] to investigate the accessibility situation.
-For this, we use a fully disaggregated approach based on parcel-level data.
-Socio-economic and demographic data are drawn from the latest Census of Canada (2016), whereas travel information is from the most recent regional travel survey from 2016.
-This paper follows reproducible research recommendations [see @brunsdon2020opening], and the research was conducted using open source tools for transportation analysis [@lovelace2021open].
-The code and data necessary to reproduce the analysis are available in a public repository^[add repository].
-
-Food Insecurity and Food Bank Use in Canada
-============
-
-Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake [@enns2020experiences; @kirkpatrick2008food; @tarasuk2009household].
-This nutrient deficiency is associated with major population health concerns, particularly among Canadians at socio-economic disadvantage [@bazerghi2016role].
-Data on food insecurity in Canada is collected in several ways.
-Quantitatively, official government surveys such as the Household Food Security Survey Module (HFSSM), the Canadian Community Health Surveys (CCHS), the International Study of Adults (LISA), and official classifications determined by Health Canada in relation to socio-demographic variables offer some insight into food insecurity [@kirkpatrick2008food; @tarasuk2009household; @gundersen2018food].
-Nationally, it has been found that food insecurity impacts approximately 12.3% of Canadian households [@tarasuk2014food]
-
-Food banks - sometimes also referred to as ‘food pantries’ and ‘food shelves’ - originated as a community response to aid those with inadequate food by voluntarily offering them meals and ingredients [@riches2002food; @loopstra2012relationship].
-The scope and objectives of food banks can vary by region and by country, and these organizations can include not only prepared meals and aliments, but also shared spaces to connect in community gardens and community kitchens [@wakefield2013sweet].
-Although in their origin food banks were meant to be provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time they have evolved into a community practice to secure emergency food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
-
-In Canada, the number of food banks has steadily increased in the past few decades [@wakefield2013sweet].
-The largest database of food banks and their use comes from the non-profit association Food Banks Canada (FBC), which conducts an annual assessment through its affiliated members.
-FBC's 2018 Hunger Count report^[https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018_FINAL_EN.pdf.aspx?ext=.pdf] (the most recent available) listed 1,830 member food banks across the country, and found that Canadians visited food banks 1.1 million times in March of 2018.
-Of those accessing food banks, certain population characteristics tend to be over-represented compared to national totals from the 2016 Canadian Census of Population.
-According to FBC's 2018 data, single-adult households represent 45% of those utilizing food banks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those accessing food bank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
-In addition, 59% of households accessing food banks list social or disability assistance as their primary source of income.
-Similarly, using data from the 2011-2012 CCHS, Tarasuk et al.
-[-@tarasuk2019geographic] found higher odds of food insecurity amongst households relying on social assistance, those without a university degree or with children under the age of 18, and individuals that lived alone, renters, and those identifying as Aboriginal.
-While surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada [@tarasuk2014food], research from Ottawa [@enns2020experiences] and Vancouver [@black2020examining] suggests that long-term users tend to be older, have health or mobility challenges, live in large households, and are less likely to have employment income.
-
-Economic stress caused by the COVID-19 pandemic, rising unemployment rates, and changes in poverty levels have disrupted the food environment and led to higher rates of food insecurity [@niles2020early].
-Pandemic-related food security studies in the US have found a substantial increase in households experiencing food insecurity for the first time, and also in households experiencing more severe food insecurity than before [@niles2020early; @wolfson2020food].
-Most recently in May 2020, Canada recorded 14.7% of its population living in food insecurity in the past 30 days [@statisticscanada2020food].
-Considering the negative mental and physical health effects associated with food insecurity, increases in food insecurity rates due to the pandemic, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
-Recent data^[https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis_Exec-Sum_EN.pdf] from FBC showed that 52% of member food banks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of the country.
-The pandemic also created significant staffing issues with 42% of food banks reporting a reduction in volunteers.
-However, 53% of food banks later reported a decrease in use into the summer of 2020 which FBC members attributed to emergency financial support programs from the federal government.
-Nevertheless, some of these benefit programs were temporary, which suggests that many households may again turn to food bank services to meet their needs.
-
-In terms of geography, previous research conducted at the provincial scale using data from the 2011-2012 CCHS found that the prevalence of food insecurity ranged across the country from 11.8% of households in Ontario to 41% of households in Nunavut [@tarasuk2019geographic].
-However, although previous research has examined the characteristics of individuals and households accessing food banks, the locational or transportation accessibility aspect of food bank access is not well understood.
-A wealth of literature examining the food desert concept suggests that, in addition to socio-economic and demographic factors, location and transportation networks play a key role in a household's accessibility to healthy foods [@paez2010relative; @vanderlee2017food; @widener2018spatial].
-For food banks specifically, previous qualitative research in Ontario by Smith-Carrier et al.
-[-@smith2017food] has noted that "transportation can be challenging, particularly if the food bank is situated in a remote location" (p.
-32).
-Particularly, it appears that participants experience challenges with the "inordinate amount of time necessary to obtain food, and difficulties associated with transportation" (p.
-39).
-Users of food banks, according to this research, rely on a variety of modes of transportation to access services.
-Consequently, the location of facilities matters; in the words of an interviewee: "I wish it [the food bank] was a little more centrally located.
-Because if I didn’t have a bike I’d have to walk it all the way out there and back.
-I wonder about people who don’t" (p.
-39).
-
-To offer greater insight into the role of transportation and location in food bank accessibility, this research examines how geographic accessibility to food banks and food services changed in Hamilton during the COVID-19 pandemic.
-
-Methods and Materials
-===================
-
-## Methods
-
-For the research in this paper we adopt the balanced floating catchment area approach of Paez et al.
-[-@paez2019demand].
-This method for estimating accessibility is a form of the widely-used two-stage floating catchment area method [@radke2000spatial; @luo2003measures].
-Floating catchment areas are used to estimate accessibility when there are potential congestion effects, and operate by calculating first the _demand_ for spatially distributed services.
-The demand (usually the number of people who require a service) is used to calculate a level of service.
-In a second step, the level of service is allocated back to the population.
-Demand and level of service are allocated using some form of distance-decay to embody the geographical principle that, given a choice, people prefer to travel less than more when reaching destinations.
-
-More formally, the first step of this method is as follows:
-$$
-L_j = \frac{S_j}{\sum_{i=1}^nP_iw_{ij}}
-$$
-
-\noindent where $S_j$ is the level of supply at location $j$, in simplest terms whether a service point is present (i.e., $S_j=1$) or not (i.e., $S_j=0$); $P_i$ is the population at location $i$ that demands the service; and $w_{ij}$ is a weight, typically a function of the distance between locations $i$ and $j$.
-$L_j$ is the level of service at location $j$ and it is the inverse of the number of people that need to be serviced.
-
-The second step in this process is then summing the level of service that each population unit can reach, according to the distance-decay weight:
-$$
-A_i = \sum_{j=1}^JL_jw_{ji}
-$$
-
-\noindent where $A_i$ is the accessibility to the service, which is in the same units as the level of service: as the inverse of the population being serviced.
-When the population being serviced is low accessibility is high (i.e., there is little competition for the service), and viceversa.
-
-Floating catchment area methods are prone to overestimation of the population and the level of service due to multiple-counting.
-The population at $P_i$ is allocated to _every_ service point $j$ for which $w_{ij}>0$.
-Similarly, the level of service at $LOS_j$ is allocated to _every_ population point for which $w_{ji}>0$.
-This inflation effect has been known for several years, and several modifications have been proposed to mitigate it [e.g., @wan2012three; @delamater2013spatial].
-A definitive solution to this issue was presented by Paez et al.
-[-@paez2019demand].
-In order to avoid the multiple-counting in the summations, the population and the level of service need to be allocated _proportionally_.
-This is achieved by standardizing the weights as follows:
-$$
-w_{ij}^\text{st} = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
-$$
-\noindent and:
-$$
-w_{ji}^\text{st} = \frac{w_{ji}}{\sum_{j=1}^Jw_{ji}}
-$$
-
-The standardized weights satisfy the following conditions:
-$$
-\sum_{i=1}^nw_{ij}^\text{st}=1
-$$
-
-\noindent and:
-$$
-\sum_{j=1}^Jw_{ji}^\text{st}=1
-$$
-
-Since the population is allocated proportionally, its value is preserved:
-$$
-\sum_{i=1}^nP_iw_{ij}^\text{st}=P_i
-$$
-
-\noindent as is the level of service:
-$$
-\sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
-$$
-## Study Area
-
-**Need to provide some context here: the geography which makes transportation challenging, i.e., due to the escarpment; the main parts of the city: city core, inner suburbs, industrial north, suburban, exurban; and the industrial background of the city maybe, with poverty rates compared to other Canadian cities?**
-
-## Data
-
-Data have been prepared for sharing as a data package^[add repository]. The contents of the data package are described next.
-
-### Statistics Canada
-
-Population and income statistics for 2016 were retrieved at the level of Dissemination Areas (DAs) using the package `cancensus` [@vonBergmann2021cancensus].
-DAs are the smallest publicly available census geography in Canada.
-Income data corresponds to the count of households by different total income groupings.
-
-### Origins: Residential parcels
-
-We converted all recorded residential land parcels in the City of Hamilton to points on the road network.
-Each point includes information about the number of residential units in the parcel.
-Next, we define low-income households as those having a total income of less than CAD40,000, which is approximately the mid-point of the low income cut-off (LICOs) for families in Canadian cities with populations greater than 500,000 in 2016, to match other Census data [@statisticscanada2020licos].
-We then "populate" each residential unit with the probability of being a low-income household based on the counts of households by income groups in the DA in which the parcel is located.
-While this method assumes a constant probability of low-income household status for all residential units in a DA, the parcel-level analysis affords a high level of spatial disaggregation for the accessibility analysis.
-
-### Destinations: Food Banks and Food Service Locations
-
-The locations of food banks and related food services were obtained from the Hamilton Public Library's Food Access Guide^[http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf].
-The guide was updated in April of 2021 to indicate any change affected on the services due to the pandemic.
-This includes modified business hours, a need to make reservations before frequenting, and locations that have completely shut down in consequence.
-Table 1 defines each service type and the number of locations pre- and during the COVID-19 pandemic.
-While some food bank services have a specific a target population, such as prioritizing families with young children aged between 0 and 3 or accepting only those providing proof of low-income status through housing and utility statements, all the food bank services indicated below are designed to accommodate those in need of food at zero to low cost.
-With our focus on food banks and food services that offer free or low-cost meals at particular locations, we first removed services such as Meals on Wheels and other food access services such as food box, community kitchens, student nutrition programs, and shopping and transportation.
-In addition, two free meal services held on different days at the same location were collapsed into a single service point for the accessibility analysis.
-Additional details on the operations of individual facilities is not publicly available, and with the changes in operations it proved unfeasible to collect it.
-For this reason, the analysis to follow is of accessibility to the location of food banks and services, but not to specific services (e.g., breakfasts vs.
-food boxes).
-
-```{r table-food-bank-info, echo=FALSE, results='asis'}
-foodbank_table %>% kable("latex",
-                         booktabs = TRUE,
-                         align = ("cccccc"),
-                         col.names = c("Type", 
-                                       "Description", 
-                                       "Locations Pre-COVID", 
-                                       "Locations During COVID", 
-                                       "Additional Notes"),
-                         caption = "\\label{tab:food-bank-info}foodbank and Food Service Information.",
-                         escape = FALSE) %>%
-  kable_styling(latex_options = c("striped", "scale_down")) %>%
-  column_spec(2, width = "15em") %>%
-  column_spec(5, width = "15em")
-```
-
-### Routing and travel time tables
-
-Travel time tables for three modes (car, transit, walking) were computed using the parcels as the origins and the locations of the food banks as the destinations.
-For routing, the package `r5r` [@pereira2021r5r] was used with a network extract for the City of Hamilton from OpenStreetMaps and the General Transit Feed Specification from Hamilton Street Railway, the local transit operator.
-For routing purposes we used maximum values of 180 min and 10,000 m walking distance: any destination that exceeded these thresholds was ignored.
-The departure time used for routing was 8:00AM on March 30, 2021 to reflect transit service around the morning service peak on a typical Tuesday.
-
-### Transportation Tomorrow Survey
-
-We used the Data Retrieval System of the Transportation Tomorrow Survey (TTS)\footnote{\url{http://dmg.utoronto.ca/}} to download cross-tabulations of: 1) primary mode of travel per trip by income by place of residence; and 2) age by income by place of residence.
-These data are from the 2016 Survey (the most recent available).
-The data are geocoded at the level of Traffic Analysis Zones (TAZ) using the most recent zoning system from 2006 and expansion factors are applied to weight the trips .
-Each parcel point is populated with the proportion of trips by three modes of travel: car (as driver or passenger), transit, and walk.
-
-### Expected Travel Times
-
-Once we obtained travel time tables with population (number of households) and proportion of trips by mode, we calculated the expected travel time $ett$ from each parcel $i$ to a food bank or food service location $j$ as follows:
-$$
-ett_{ij} = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
-$$
-
-\noindent where $p^k_i$ is the proportion of trips by mode $k$ in the TAZ of parcel $i$, and $tt^k_{ij}$ is the travel time from parcel $i$ to the food bank.
-In other words, the expected travel time is the weighted sum of travel times to the food bank, with the weights given by the expected modal split of trips made by low-income households in the TAZ per the TTS data.
-
-Results and Discussion
-===================
-
-```{r accessibility-threshold-sensitivity, include=FALSE}
-## Analysis
-# Testing different thresholds to check how accessibility changes. Here, the thresholds are varied from 2 min to 30 min in 2 min step increments
-threshold_seq <- seq(from = 2, to = 30, by = 1)
-
-# Initialize a data frame to store results
-accessibility_test <- data.frame(threshold = threshold_seq, 
-                                 accessibility = numeric(length(threshold_seq)))
-
-# Calculate the system-wide accessibility according to each threshold
-for(i in 1:length(threshold_seq)){
-  results_test <-  ttm %>% 
-    transmute(UID, OBJECTID, travel_time = e_travel_time, population = income_less40k, supply = 1) %>%
-    b2sfca(threshold = i )
-  accessibility_test$accessibility[i] <- sum(results_test$accessibility$accessibility)
-}
-```
-
-```{r results-pre-covid, include=FALSE}
-## Analysis 
-# Calculate k-min accessibility pre-covid:
-results_pre <- ttm %>%
-  #dplyr::filter(COVIDClose == "No") %>%
-  transmute(UID, 
-            OBJECTID, 
-            travel_time = e_travel_time, 
-            population = income_less40k, 
-            supply = 1) %>%
-  b2sfca(threshold = 15)
-```
-
-```{r results-during-covid, include=FALSE}
-## Analysis 
-# Calculate k-min accessibility during covid:
-results_covid <- ttm %>%
-  dplyr::filter(COVIDClose == "No") %>%
-  transmute(UID, 
-            OBJECTID, 
-            travel_time = e_travel_time, 
-            population = income_less40k, 
-            supply = 1) %>%
-  b2sfca(threshold = 15)
-```
-
-```{r, bind-accessibility-results, include=FALSE}
-## Analysis
-# Bind level of service results:
-los <- rbind(data.frame(results_pre$los, 
-                        timing = "Pre-covid", 
-                        threshold = "30 min"),
-             data.frame(results_covid$los, 
-                        timing = "During covid", 
-                        threshold = "30 min")) %>%
-  mutate(timing = factor(timing, 
-                         levels = c("Pre-covid", 
-                                    "During covid"),
-                         ordered = TRUE))
-
-# Bind accessibility results:
-accessibility <- rbind(data.frame(results_pre$accessibility, 
-                                  timing = "Pre-covid", 
-                                  threshold = "30 min"),
-                       data.frame(results_covid$accessibility, 
-                                  timing = "During covid", 
-                                  threshold = "30 min")) %>%
-  mutate(timing = factor(timing, 
-                         levels = c("Pre-covid", 
-                                    "During covid"),
-                         ordered = TRUE))
-```
-
-Figure \ref{fig:foodbanks} shows the location of food banks and services in the City of Hamilton and their status.
-Before the pandemic there were `r nrow(foodbank_locations)` of which `r nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))` (`r round(nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations) * 100, 2)`\%) closed during the pandemic.
-As shown in the figure, food services tend to be predominantly located in the central parts of the city.
-This is not surprising: population density is high there, and it is also the part of the city where lower income households are more numerous in absolute and relative terms (see Figure \ref{fig:low-income-households}).
-Alas, this is also the part of the city where most of the closures during the pandemic happened.
-
-```{r plot-location-foodbanks, echo=FALSE, fig.cap="\\label{fig:foodbanks}Location of food banks/services and operation status; the dotted box is an inset of the central part of the City of Hamilton"}
-## Analysis
-# Plot location of foodbanks
-
-# Create a bounding box for an inset
-inset_box <- st_bbox(c(xmin = -79.975,
-                       xmax = -79.775, 
-                       ymax = 43.2, 
-                       ymin = 43.3), 
-                     crs = st_crs(4326)) %>%
-  st_as_sfc()
-
-fb <- ggplot() +
-  geom_sf(data = hamilton_boundary,
-          fill = "lightgray") +
-  geom_sf(data = foodbank_locations,
-          aes(size = COVIDClose,
-              shape = COVIDClose)) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_shape_manual(name = "Closed during COVID-19",
-                     values = c("Yes" = 4, "No" = 1)) +
-  scale_size_manual(values = c("Yes" = 3, "No" = 2)) +
-  guides(size = FALSE) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90),
-        legend.position = "none")
-
-fb_inset <- ggplot() +
-  geom_sf(data = hamilton_boundary %>% 
-            st_transform(4326),
-          fill = "lightgray") +
-  geom_sf(data = foodbank_locations,
-          aes(size = COVIDClose,
-              shape = COVIDClose)) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_shape_manual(name = "Closed during COVID-19",
-                     values = c("Yes" = 4, "No" = 1)) +
-  scale_size_manual(values = c("Yes" = 3, "No" = 2)) +
-  xlim(-79.975, -79.775) +
-  ylim(43.2, 43.3) +
-  guides(size = FALSE) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90),
-        legend.position = "bottom")
-
-fb / fb_inset
-```
-
-```{r plot-low-income-households, echo=FALSE, fig.cap="\\label{fig:low-income-households}Number and proportion of households with incomes less than CAD40,000."}
-lih <- ggplot() +
-  geom_sf(data = data_da_2016,
-          aes(fill = income_less40k),
-          color = NA) +
-  geom_sf(data = hamilton_boundary,
-          fill = NA) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_fill_viridis_c(name = "Number of households", 
-                       option = "A",
-                       direction = -1,
-                       breaks = c(200, 400, 600)) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "bottom")
-
-p_lih <- ggplot() +
-  geom_sf(data = data_da_2016,
-          aes(fill = p_inc_less40k),
-          color = NA) +
-  geom_sf(data = hamilton_boundary,
-          fill = NA) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_fill_viridis_c(name = "Proportion of households", 
-                       option = "A",
-                       direction = -1,
-                       breaks = c(0.25, 0.50, 0.75)) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "bottom")
-
-lih / p_lih
-```
-
-To implement the accessibility calculations, we must select a distance-decay function.
-In this task we find limited support in the literature, which is mostly silent on the travel patterns of people who visit food banks.
-For this reason, we opt for a simple cumulative opportunities function as follows:
-$$
-w_{ij}=w_{ji}=
-\begin{cases}
-1 & \text{ if } ett_{ij}\le \delta\\
-0 & \text{ otherwise}
-\end{cases}
-$$
-\noindent where $ett_{ij}$ is the multimodal expected travel time as described previously, and $\delta$ is a travel threshold.
-When the expected travel time exceeds this threshold, a facility is no longer considered accessible.
-Moreover, the weights are standardized for the balanced floating catchment area approach.
-
-Figure \ref{fig:sensitivity-analysis} shows the results of conducting sensitivity analysis of the system-wide accessibility as we vary the threshold (considering the situation before the pandemic).
-There is a clear pattern whereby more strict values of $\delta$ are associated with higher levels of system-wide accessibility: while increases in accessibility that result from decreases in the travel time window might seem counter-intuitive, this is a result of lower _congestion_, since fewer households are serviced and thus competition for the same resources is more limited.
-System-wide accessibility declines with higher values of $\delta$: as more households are serviced, congestion grows and the level of service declines, although this happens at a declining rate.
-We are not aware of any research that explains how long people are expected to travel for food banks, but we note that in developing countries, access to drinking water is defined as a source that takes less than 30 minutes to reach [round trip, see @world2019progress].
-There is no reason why people in affluent countries should be expected to travel more for a basic need such as food.
-Accordingly, we adopt a 15-minute threshold for the analysis (representing a one-way trip).
-This threshold is also approximately where the rate of change in accessibility slows down.
-
-```{r plot-results-sensitivity-analysis, out.width="60%", echo=FALSE, fig.cap="\\label{fig:sensitivity-analysis}Accessibility as a function of threshold"}
-acc_sensitivity <- ggplot(accessibility_test, 
-                          aes(x = threshold, y = accessibility)) +
-  geom_line() +
-  ylab("Accessibility") +
-  xlab("Threshold (in min)") +
-  theme_minimal()
-
-acc_sensitivity
-```
-
-Using the 15-minute threshold, we find that the system-wide accessibility (interpreted as a provider-to-population ratio) was `r round(sum(results_pre$accessibility$accessibility), 3)` (food banks/service locations per low income household in the city) before COVID-19, but declined to `r round(sum(results_covid$accessibility$accessibility), 3)` during the pandemic.
-It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`\% of food facilities remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`\%, suggesting the location of food banks and related services plays an important role in serving those in need.
-
-```{r join-zonal-ids-to-accessibility-table, include=FALSE}
-## Analysis
-# Join zonal identifiers to accessibility results:
-accessibility <- accessibility %>%
-  left_join(dwelling_network_points_2016 %>%
-              st_drop_geometry() %>%
-              dplyr::select(ID,
-                            DAUID,
-                            TAZUID,
-                            children,
-                            adults,
-                            seniors),
-            by = c("UID" = "ID")) %>%
-  left_join(population_age_income %>%
-              st_drop_geometry() %>%
-              transmute(TAZUID),
-            by = "TAZUID") 
-```
-
-```{r, join-geometry-to-level-of-service, include=FALSE}
-## Analysis 
-# Join geometry:
-los <- los %>%
-  left_join(foodbank_locations %>%
-              dplyr::select(ID),
-            by = c("OBJECTID" = "ID")) %>%
-  st_as_sf()
-```
-
-Turning to the location of individual facilities, the levels of service offered before and during the pandemic are shown in Figure \ref{fig:levels-of-service}.
-The level of service is functionally the inverse of the number of low-income households in the travel-mode weighted travel time catchment area of the facilities (this is because $S_j=1 \forall j$, i.e., each location represents a "capacity" of 1).
-Higher values mean that a facility is expected to service fewer households.
-Conversely, lower values indicate greater congestion.
-
-The general pattern of the levels of service is similar before and during the pandemic, with lower values in the center of the city where low-income households exhibit multimodal trip patterns.
-Three more peripheral facilities towards the south of the city also have lower levels of service, presumably because they are expected to service relatively large suburban/exurban populations generally reliant on automobiles for travel.
-During the pandemic, however, the levels of service dropped, in some cases quite substantially.
-The pattern of the losses in level of service, moreover, is not uniform.
-The upper pane of Figure \ref{fig:levels-of-service-changes} shows that the three peripheral facilities in the southern suburban/exurban part of the city had low levels of service to begin with, but did not see major declines during the pandemic.
-Further, the inset map shows that the levels of service deteriorated more in the central part of the city.
-However, the loss of level of service was not as large in the core (where most of the food banks/services are found), but instead was more marked in the inner ring around the core, where facilities may have faced greater demand from both central city and suburban populations after the closure of food banks during the pandemic.
-
-```{r plot-levels-of-service, echo=FALSE, fig.cap="\\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
-## Analysis
-# Plot levels of service before and during covid
-
-# Pre-covid
-los_pre <- ggplot() +
-  geom_sf(data = hamilton_boundary) +
-  geom_sf(data = los %>%
-            filter(timing == "Pre-covid"),
-          aes(color = los,
-              size = los),
-          alpha = 0.5) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  scale_color_viridis_c(name = "Level of service \npre-COVID-19",
-                        option =  "A", 
-                        direction = -1,
-                        breaks = c(0.0005,
-                                   0.0010,
-                                   0.0015,
-                                   0.0020,
-                                   0.0025),
-                        labels = c("0.00050",
-                                   "0.00100",
-                                   "0.00150",
-                                   "0.00200",
-                                   "0.00250")) +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  guides(size = FALSE, shape = FALSE) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-# During covid
-los_covid <- ggplot() +
-  geom_sf(data = hamilton_boundary) +
-  geom_sf(data = los %>%
-            filter(timing == "During covid"),
-          aes(color = los,
-              size = los),
-          alpha = 0.5) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  scale_color_viridis_c(name = "Level of service \nduring COVID-19",
-                        option =  "A", 
-                        direction = -1,
-                        breaks = c(0.00050,
-                                   0.00075,
-                                   0.00100,
-                                   0.00125,
-                                   0.00150),
-                        labels = c("0.00050",
-                                   "0.00075",
-                                   "0.00100",
-                                   "0.00125",
-                                   "0.00150")) +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  guides(size = FALSE, shape = FALSE) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-los_pre / los_covid
-```
-
-```{r plot-levels-of-service-changes, echo=FALSE, fig.cap="\\label{fig:levels-of-service-changes}Changes in levels of service at each facility from pre-COVID-19 to during COVID-19."}
-# This figure shows the places where the level of service deteriorated most (excluding the places that closed, where it vanished).
-los_change <- los %>%
-  st_drop_geometry() %>%
-  pivot_wider(names_from = timing, 
-              values_from = los, 
-              values_fill = 0) %>%
-  filter(`During covid` > 0) %>%
-  mutate(los_d = `During covid` - `Pre-covid`) %>%
-  left_join(foodbank_locations %>%
-              transmute(OBJECTID = ID,
-                        geometry),
-            by = "OBJECTID") %>%
-  st_as_sf()
-
-lc <- ggplot() +
-  geom_sf(data = hamilton_boundary) +
-  geom_sf(data = los_change, 
-          aes(color = los_d,
-              size = los_d),
-          alpha = 0.3) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_color_viridis_c(name = "Change in level of service",
-                        option =  "A", 
-                        direction = 1,
-                        breaks = c(-0.0001,
-                                   -0.0002,
-                                   -0.0003,
-                                   -0.0004,
-                                   -0.0005),
-                        labels = c("-0.0001",
-                                   "-0.0002",
-                                   "-0.0003",
-                                   "-0.0004",
-                                   "-0.0005")) +
-  scale_size(trans = 'reverse') +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  guides(size = FALSE, shape = FALSE) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "none")
-
-lc_inset <- ggplot() +
-  geom_sf(data = hamilton_boundary %>% 
-            st_transform(4326)) +
-  geom_sf(data = los_change, 
-          aes(color = los_d,
-              size = los_d),
-          alpha = 0.7) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_color_viridis_c(name = "Change in level of service",
-                        option =  "A", 
-                        direction = 1,
-                        breaks = c(-0.0001,
-                                   -0.0002,
-                                   -0.0003,
-                                   -0.0004,
-                                   -0.0005),
-                        labels = c("-0.0001",
-                                   "-0.0002",
-                                   "-0.0003",
-                                   "-0.0004",
-                                   "-0.0005")) +
-  scale_size(name = "Change in level of service",
-             trans = 'reverse',
-             breaks = c(-0.0001,
-                        -0.0002,
-                        -0.0003,
-                        -0.0004,
-                        -0.0005),
-             labels = c("-0.0001",
-                        "-0.0002",
-                        "-0.0003",
-                        "-0.0004",
-                        "-0.0005")) +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  xlim(-79.975, -79.775) +
-  ylim(43.2, 43.3) +
-  guides(color = guide_legend(title.position = "top"), 
-         #size = FALSE, 
-         shape = FALSE) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "bottom")
-
-lc / lc_inset
-```
-
-To further elucidate this issue, we now turn to the results of the accessibility analysis.
-As with the level of service of individual facilities, the general pattern of accessibility before and during the pandemic is similar.
-Figure \ref{fig:accessibility} reveals that, compared with the outer rural zones, the more urban zones of the city generally exhibit higher accessibility to food banks and food service locations.
-However, the pattern is not particularly smooth - this is largely attributable to the weighting of travel times by mode of transportation according to the trip patterns of low-income household respondents captured by the TTS.
-For example, in zones where low-income households make a high proportion of trips by walking, access to food bank locations by walking is afforded a concomitantly high weight in our calculations of travel time compared to transit or car travel.
-From this, highly-accessible locations result from a mix of characteristics: low-income households in locations where travel options that align with zonal modal split are available to connect them to food bank locations with high levels of service within 15 minutes.
-This seems to track with the experience of some users of these services, as reported by Smith-Carrier et al.
-[-@smith2017food].
-
-```{r aggregate-accessibility-by-taz, include=FALSE}  
-## Analysis
-# Accessibility by TAZ:
-accessibility_taz <- accessibility %>%
-  group_by(TAZUID, 
-           timing) %>%
-  summarize(accessibility = sum(accessibility),
-            timing = first(timing),
-            .groups = "drop") %>%
-  left_join(population_age_income %>%
-              transmute(TAZUID,
-                        children,
-                        adults,
-                        seniors),
-            by = "TAZUID") %>%
-  st_as_sf()
-```
-
-```{r plot-accessibility, echo=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
-## Analysis
-# Plot accessibility before and during covid:
-acc_pre <- ggplot() +
-  geom_sf(data = hamilton_boundary,
-          fill = NA) +
-  geom_sf(data = accessibility_taz %>%
-            filter(timing == "Pre-covid"),
-          aes(fill = accessibility)) +
-  scale_fill_viridis_c(name = "Accessibility \npre-COVID-19",
-                       option =  "A", 
-                       direction = -1,
-                       breaks = c(0.002,
-                                  0.0016,
-                                  0.0012,
-                                  0.0008,
-                                  0.0004),
-                       labels = c("0.0020",
-                                  "0.0016",
-                                  "0.0012",
-                                  "0.0008",
-                                  "0.0004")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-acc_covid <- ggplot() +
-  geom_sf(data = hamilton_boundary,
-          fill = NA) +
-  geom_sf(data = accessibility_taz %>%
-            filter(timing == "During covid"),
-          aes(fill = accessibility)) +
-  scale_fill_viridis_c(name = "Accessibility \nduring COVID-19",
-                       option =  "A", 
-                       direction = -1,
-                       breaks = c(0.0010,
-                                  0.0008,
-                                  0.0006,
-                                  0.0004,
-                                  0.0002),
-                       labels = c("0.0010",
-                                  "0.0008",
-                                  "0.0006",
-                                  "0.0004",
-                                  "0.0002")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-acc_pre / acc_covid
-```
-
-```{r accessibility-changes-by-taz, include=FALSE}
-## Analysis
-# Calculate accessibility change by TAZ:
-accessibility_taz_change <- accessibility_taz %>%
-  st_drop_geometry() %>%
-  pivot_wider(names_from = timing, 
-              values_from = accessibility) %>%
-  # Calculate changes in accessibility
-  mutate(acc_d = `During covid` - `Pre-covid`) %>%
-  # Join TAZ geometry
-  left_join(population_age_income %>%
-              transmute(TAZUID,
-                        geometry),
-            by = "TAZUID") %>%
-  # Convert to simple features
-  st_as_sf()
-```
-
-We find that the accessibility landscape deteriorated substantially during the pandemic, with accessibility dropping on average by almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% mean() %>% abs() %>% round()`\%, but with large variations: some zones experienced changes in accessibility of only about `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% max() %>% abs() %>% round()`\%, whereas the most affected zone saw a loss of accessibility of almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% min() %>% abs() %>% round()`\%.
-Figure \ref{fig:accessibility-changes-with-local-i} shows the changes in accessibility.
-Every zone is worse off after the closure of facilities during the pandemic, but some parts of the city seem to have been particularly affected.
-To better highlight these changes, we used a local indicator of spatial autocorrelation [@anselin1995local] to explore the pattern of change in accessibility.
-Twenty-four TAZ are flagged as having significant large losses of accessibility (at $p<=0.10$, without correcting for multiple comparisons).
-Those zones are highlighted in the figure, where it can be seen that they form more or less compact neighborhoods.
-Remarkably, the largest (and significant) drops in accessibility are not downtown, but located in two cases in the industrial north of the city, in one case in an inner suburb above the escarpment, and lastly in a more suburban/exurban region in the south-west.
-
-```{r plot-accessibility-changes, echo=FALSE, include=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19."}
-## Analysis
-# Plot
-acc_change <- ggplot() +
-  #geom_sf(data = hamilton_boundary) +
-  geom_sf(data = accessibility_taz_change,
-          aes(fill = acc_d),
-          color = NA) +
-  geom_sf(data = population_age_income,
-          fill = NA) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_fill_viridis_c(name = "Change in accessibility",
-                       option =  "A", 
-                       direction = 1,
-                       breaks = c(-0.0008,
-                                  -0.0006,
-                                  -0.0004,
-                                  -0.0002),
-                       labels = c("-0.0008",
-                                  "-0.0006",
-                                  "-0.0004",
-                                  "-0.0002")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "none")
-
-acc_change_inset <- ggplot() +
-  geom_sf(data = accessibility_taz_change %>%
-            st_transform(crs = 4326),
-          aes(fill = acc_d)) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_fill_viridis_c(name = "Change in accessibility",
-                       option =  "A", 
-                       direction = 1,
-                       breaks = c(-0.0008,
-                                  -0.0006,
-                                  -0.0004,
-                                  -0.0002),
-                       labels = c("-0.0008",
-                                  "-0.0006",
-                                  "-0.0004",
-                                  "-0.0002")) +
-  guides(fill = guide_legend(title.position = "top")) +
-  xlim(-79.975, -79.775) +
-  ylim(43.2, 43.3) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "bottom")
-
-acc_change / acc_change_inset
-```
-
-```{r weights-matrix, include=FALSE}
-taz_nb <- poly2nb(accessibility_taz_change, queen = TRUE)
-taz_w <- nb2listw(taz_nb)
-```
-
-```{r moran-test-accessibility-changes, include=FALSE}
-moran.test(accessibility_taz_change$acc_d, taz_w)
-```
-
-```{r local-moran-i, include=FALSE}
-local_i <- localmoran(accessibility_taz_change$acc_d,
-                      taz_w, 
-                      alternative = "two.sided",
-                      p.adjust.method = "none") %>% 
-  data.frame() %>%
-  rename(p_val = Pr.z....0.) %>%
-  mutate(significance = ifelse(p_val <= 0.10, "sig", "n.s."),
-         significance = factor(significance,
-                               levels = c("n.s.", "sig")))
-
-summary(local_i)
-```
-
-```{r join-local-i-to-table, include=FALSE}
-accessibility_taz_change$significance <- local_i$significance
-```
-
-```{r plot-local-i, echo=FALSE, fig.cap="\\label{fig:accessibility-changes-with-local-i}Changes in accessibility from pre-COVID-19 to during COVID-19.
-Highlighted areas had significantly large changes in accessibility according to Local Moran's I." }
-ggplot() +
-  geom_sf(data = accessibility_taz_change,
-          aes(fill = acc_d,
-              color = significance,
-              alpha = significance),
-          size = 1) +
-  geom_sf(data = hamilton_boundary,
-          fill = NA) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_fill_viridis_c(name = "Change in accessibility",
-                       option =  "A", 
-                       direction = 1,
-                       breaks = c(-0.0008,
-                                  -0.0006,
-                                  -0.0004,
-                                  -0.0002),
-                       labels = c("-0.0008",
-                                  "-0.0006",
-                                  "-0.0004",
-                                  "-0.0002")) +
-  scale_color_manual(values = c("n.s." = "lightgray", 
-                                "sig" = "black")) +
-  scale_alpha_manual(values = c("n.s." = 0.5, 
-                                "sig" = 1)) +
-  guides(color = FALSE, 
-         alpha = FALSE,
-         fill = guide_legend(title.position = "top")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "bottom")
-```
-
-For the more suburban clusters of zones, the decrease in accessibility is derived from the closure of locations throughout the city reachable by car.
-In the cluster of central suburban zones for example, low-income households in the outer ring of zones that exhibit medium to high decreases in accessibility within this cluster appear to be largely auto-dependent in their tripmaking, which each each exhibiting between `r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(min(p_car)) *100)`-`r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(max(p_car)) *100)`\% of their modal split for car trips.
-This results in the parcels within these zones having a large number of potentially accessible locations in the travel time matrix.
-But by extension, the change in accessibility over the pre- and during-COVID-19 time periods is affected not only by the closure of service locations proximate to the zones, but also the locations in the central city.
-The zone with the greatest decrease in accessibility within this cluster (-0.0009) has a high rate of car trips and connects to the most food bank locations in total as well as those that stayed open or closed.
-The exception to this is the zone in the middle of this cluster, which has a much higher proportion of transit use `r round(modes %>% st_drop_geometry() %>% filter(TAZUID == 5116) %>% summarize(first(p_transit)) *100)`\%.
-This modal split weighting results in fewer food bank and service locations accessible from the parcels within this zone which, by extension, limits their access to closed locations reachable within 15 minutes in our calculations.
-
-In the cluster to the south-west, the decrease in accessibility is predominately driven by the closure of a high level-of-service Community Meals provider.
-However, like the more central suburban zones, low-income households within this cluster also between 90\% to 100\% auto-dependent in their tripmaking in the TTS.
-The story is similar for the zone located in the north-west that exhibits the greatest decrease in accessibility.
-Here, low-income households responding to the TTS conducted 100\% of their trips by car and, as a result, parcels within this zone have access to the second-highest number of food bank and service locations within 15 minutes.
-However, this also means zonal accessibilities are greatly affected by the number of closures throughout the city.
-Finally, in the city's north end and north-east zones, low-income households exhibit a mixture of tripmaking behaviour in the TTS.
-Households in some zones take transit more often and one zone in particular has 100\% of its trips by walking while in others, car use makes up 100\% of trips.
-For these zones, the decrease in accessibility tends to be a product of the closure of several inner-city food bank and service locations.
-
-Just as the effects of the closures appear to have been uneven in space, they also seem to have differently impacted various population segments.
-Table \ref{tab:accessibility-by-age} shows the estimated number of people in each age group by their level of accessibility before and during the pandemic.
-Here, it is important to note that the quartiles are relative: people in the top 25\% of accessibility still have lower accessibility during the pandemic than before.
-In reality, every population group is worse off during the pandemic in terms of their accessibility to food banks and services in the City of Hamilton.
-However, some age groups were affected more.
-As seen in the table, the distribution within quartiles for adults changed only slightly, and in fact there are fewer adults in the bottom three quartiles as some found themselves in the (deteriorated) top quartile of accessibility during the pandemic.
-The story is different for seniors, a greater number of whom are now in the third quartile, due to a combination of seniors facing a worse accessibility situation, and just a few moving from the bottom quartile (pre-pandemic) to the third during the pandemic.
-Those aged 18 and less appear to have fared worse, and the largest change is in the number of children who were in the third quartile before the pandemic and found themselves in the bottom accessibility class during the pandemic.
-These results suggest that, through a combination of the typical modes of transportation of lower income households and the spatial distribution of the population, the food bank/service locations closed had a differential impact that more greatly affected the youngest among the population in low income households.
-
-```{r accessibility-change-by-parcel, include=FALSE}
-## Analysis
-# Calculate accessibility change by parcel:
-accessibility_change <- accessibility %>%
-  select(UID, timing, accessibility) %>%
-  pivot_wider(names_from = timing, 
-              values_from = accessibility,
-              values_fill = 0) %>%
-  # Calculate changes in accessibility
-  mutate(acc_d = `During covid` - `Pre-covid`)
-```
-
-```{r join-accessibility-change-to-table, include=FALSE}
-## Analysis
-# Join accessibility change to accessibility table:
-accessibility <- accessibility %>%
-  left_join(accessibility_change %>%
-              select(UID, acc_d),
-            by = "UID")
-```
-
-```{r quartiles-of-accessibility-change, include=FALSE}
-## Analysis
-# Quartiles of accessibility pre-covid:
-acc_quantiles_pre <- quantile(accessibility_change$`Pre-covid`,
-                              c(0, 0.25, 0.5, 0.75, 1))
-# Quartiles of accessibility change:
-acc_quantiles_during <- quantile(accessibility_change$`During covid`,
-                                 c(0, 0.25, 0.5, 0.75, 1))
-```
-
-```{r accessibility-labels, include=FALSE}
-## Analysis
-# Add labels of accessibility loss to table:
-accessibility <- accessibility %>%
-  mutate(accessibility_pre_level = case_when(accessibility <= acc_quantiles_pre[2] ~ "Bottom 25\\%",
-                                             accessibility > acc_quantiles_pre[2] & accessibility <= acc_quantiles_pre[3] ~ "Third 25\\%",
-                                             accessibility > acc_quantiles_pre[3] & accessibility <= acc_quantiles_pre[4] ~ "Second 25\\%",
-                                             accessibility > acc_quantiles_pre[4] ~ "Top 25\\%"),
-         accessibility_pre_level = factor(accessibility_pre_level,
-                                          levels = c("Top 25\\%", "Second 25\\%", "Third 25\\%", "Bottom 25\\%"),
-                                          ordered = TRUE),
-         accessibility_during_level = case_when(accessibility <= acc_quantiles_during[2] ~ "Bottom 25\\%",
-                                                accessibility > acc_quantiles_during[2] & accessibility <= acc_quantiles_during[3] ~ "Third 25\\%",
-                                                accessibility > acc_quantiles_during[3] & accessibility <= acc_quantiles_during[4] ~ "Second 25\\%",
-                                                accessibility > acc_quantiles_during[4] ~ "Top 25\\%"),
-         accessibility_during_level = factor(accessibility_during_level,
-                                             levels = c("Top 25\\%", "Second 25\\%", "Third 25\\%", "Bottom 25\\%"),
-                                             ordered = TRUE))
-```
-
-```{r aggregate-accessibility-by-age-group, include=FALSE}
-summary_pre <- accessibility %>%
-  filter(timing == "Pre-covid") %>%
-  group_by(accessibility_pre_level) %>%
-  summarize(children_pre = sum(children) %>% 
-              round(),
-            adults_pre = sum(adults) %>% 
-              round(),
-            seniors_pre = sum(seniors) %>% 
-              round(),
-            .groups = "drop")
-
-summary_during <- accessibility %>%
-  filter(timing == "During covid") %>%
-  group_by(accessibility_during_level) %>%
-  summarize(children_during = sum(children) %>% 
-              round(),
-            adults_during = sum(adults) %>% 
-              round(),
-            seniors_during = sum(seniors) %>% 
-              round(),
-            .groups_during = "drop")
-
-accessibility_by_age <- cbind(summary_pre, summary_during[,2:4]) %>%
-  mutate(children_change = round((children_during - children_pre)/children_pre * 100, 2),
-         adults_change = round((adults_during - adults_pre)/adults_pre * 100, 2),
-         seniors_change = round((seniors_during - seniors_pre)/seniors_pre * 100, 2),
-         children_pre = prettyNum(children_pre, big.mark = ","),
-         adults_pre = prettyNum(adults_pre, big.mark = ","),
-         seniors_pre = prettyNum(seniors_pre, big.mark = ","),
-         children_during = paste0(prettyNum(children_during, big.mark = ","), " (", children_change, "\\%)"),
-         adults_during = paste0(prettyNum(adults_during, big.mark = ","), " (", adults_change, "\\%)"),
-         seniors_during = paste0(prettyNum(seniors_during, big.mark = ","), " (", seniors_change, "\\%)")) %>%
-  transmute(accessibility_pre_level,
-            children_pre,
-            adults_pre,
-            seniors_pre,
-            children_during,
-            adults_during,
-            seniors_during)
-```
-
-```{r table-accessibility-by-age-group, echo=FALSE, results='asis'}
-accessibility_by_age %>%
-  kable("latex",
-        booktabs = TRUE,
-        align = ("lcccccc"),
-        col.names = c("Accessibility Quartile",
-                      "Children (age $\\le$ 18)", 
-                      "Adults (19-64)", 
-                      "Seniors (age $\\ge$ 65)", 
-                      "Children (age $\\le$ 18)", 
-                      "Adults (19-64)", 
-                      "Seniors (age $\\ge$ 65)"),
-        caption = "\\label{tab:accessibility-by-age}Population at each accessibility level by age group among members households with incomes less than CAD40,000.",
-        escape = FALSE) %>%
-  kable_styling(latex_options = c("striped", "scale_down", "hold_position")) %>%
-  add_header_above(c(" " = 1, "Pre-COVID-19" = 3, "During COVID-19" = 3)) %>%
-  footnote(general = c("Population values have been rounded.", 
-                       "The values in brackets for population during COVID-19 are the changes from before the pandemic."))
-```
-
-```{r include=FALSE, eval = FALSE}
-## Analysis
-# Calculate accessibility by age group and accessibility level:
-accessibility_by_age <- accessibility %>%
-  filter(timing = "During covid")
-group_by(timing, 
-         accessibility_level) %>%
-  summarize(children = sum(children),
-            adults = sum(adults),
-            seniors = sum(seniors),
-            accessibility = sum(accessibility),
-            .groups = "drop")
-# The accessibility level is the magnitude of the accessibility loss: the top 25% faced the smallest losses in accessibility, and the bottom 25% faced the greatest losses in accessibility. 
-```
-
-```{r include=FALSE, eval = FALSE}
-cbind(accessibility_by_age %>% 
-        filter(timing == "Pre-covid") %>% 
-        select(children:accessibility) %>% 
-        rename(accessibility_pre = accessibility),
-      accessibility_by_age %>% 
-        filter(timing == "During covid") %>% 
-        transmute(accessibility_covid = accessibility)) %>%
-  mutate(pct_children = round(children/sum(children) * 100, 1),
-         pct_children = paste0("(", pct_children, "\\%)", sep = ""),
-         pct_adults = round(adults/sum(adults) * 100, 1),
-         pct_adults = paste0("(", pct_adults, "\\%)", sep = ""),
-         pct_seniors = round(seniors/sum(seniors) * 100, 1),
-         pct_seniors = paste0("(", pct_seniors, "\\%)", sep = ""),
-         children = round(children) %>% 
-           prettyNum(big.mark = ","),
-         adults = round(adults) %>% 
-           prettyNum(big.mark = ","),
-         seniors = round(seniors) %>% 
-           prettyNum(big.mark = ","),
-         children = paste0(children, " ", pct_children, sep = ""),
-         adults = paste0(adults, " ", pct_adults, sep = ""),
-         seniors = paste0(seniors, " ", pct_seniors, sep = ""),
-         change = (accessibility_covid - accessibility_pre) %>% 
-           round(5),
-         accessibility_pre = round(accessibility_pre, 5),
-         accessibility_covid = round(accessibility_covid, 5)) %>%
-  transmute(children,
-            adults,
-            seniors,
-            accessibility_pre,
-            accessibility_covid,
-            change) %>%
-  kable("latex",
-        booktabs = TRUE,
-        align = ("cccccc"),
-        col.names = c("Children (age $\\le$ 18)", 
-                      "Adults (19-64)", 
-                      "Seniors (age $\\ge$ 65)", 
-                      "Before COVID-19", 
-                      "During COVID-19", 
-                      "Difference"),
-        caption = "\\label{tab:accessibility-by-age}Accessibility by age group among members households with incomes less than CAD40,000.",
-        escape = FALSE) %>%
-  kable_styling(latex_options = c("striped", "scale_down")) %>%
-  add_header_above(c("Population" = 3, "Accessibility" = 3)) %>%
-  footnote(general = "Population values have been rounded")
-```
-
-Conclusions
-===================
-
-**LIMITATIONS**
-
-Words go here.
-
-We use data from the census to model low-income households as a proxy for households that are possibly experiencing food insecurity.
-However, the degree to which our low-income cut-off of CAD\$40,000 reflects food insecurity in the different zones of the study area is not known.
-
-While the travel survey allows us to model multi-modal accessibilities that align with travel behaviour observed in the travel survey and capture differences in accessibility by age categories, the use of travel survey data for modelling food insecurity also has its limitations.
-Research into the population weighting methods used in the TTS note that the survey may under-count the lowest- and highest-income households in the survey study region, although the magnitude of this under-counting is unknown as approximately 20% of respondents to the 2016 survey did not report their income [@rose2018transportation].
-In that regard, the modal splits of low-income households observed through the TTS data may not accurately reflect the travel behaviours of food insecure populations.
-
-In the absence of information on how food insecure households travel to food banks and related services, we adopted a normative approach that examines accessibility to food banks reachable within 15 minutes based on standards for accessible drinking water...
-
-References {#references .unnumbered}
-==========

### 2021-04-27 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-abstract.tex
Created abstract file

--- 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
@@
 
-\newpage
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Food insecurity is defined as an ``inadequate or uncertain access to a
-sufficient quantity and/or adequate quality of food'' due to a
-household's financial limitations (Enns et al., 2020). This condition
-has been associated with reductions in nutritional outcomes
-(Bhattacharya et al., 2004; Kirkpatrick and Tarasuk, 2008; Olson, 1999)
-and physical and mental health in children and adults (Elgar et al.,
-2021; Jones, 2017; Ramsey et al., 2011; Seligman et al., 2010; Stuff et
-al., 2004). Over at least the past four decades food banks and related
-services have become an essential line of defense against food
-insecurity in Canadian communities (Black and Seto, 2020; Holmes et al.,
-2018; Riches, 2002; Tarasuk et al., 2020). In this respect, Canada is
-not unlike numerous other wealthy countries where a systematic
-dismantling of the welfare state took place in the intervening period
-(Tarasuk et al., 2014).
-
-The emergence of COVID-19, the worst public health crisis since the 1918
-flu pandemic, has revealed important social and economic fault lines,
-and pre-existing patterns of inequality appear to have been exacerbated.
-Along several other dimensions of stress (e.g., accessibility to health
-care facilities, Pereira et al., 2021a), this seems to be the case for
-food insecurity as well (Laborde et al., 2020). In the US, for example,
-it has been estimated that there was an increase of more than 30\% in
-household food insecurity, and more than one third of households were
-discovered to be newly food insecure - meaning they did not experience
-food insecurity before the pandemic (Niles et al., 2020). In Canada, Men
-and Tarasuk (2021) report that about 25\% of individuals who experienced
-job insecurity (a relatively common occurrence during the pandemic),
-also experienced food insecurity. Similarly, according to Statistics
-Canada (2020a), in the early stages of the pandemic almost 15\% of
-individuals reported living in a household that faced food insecurity;
-the risk of food insecurity was substantially higher for households with
-children. The difference between households with and without children
-was significant, and 11.7\% of households with children indicated that
-``food didn't last and {[}there was{]} no money to get more'' sometimes
-or often, compared to 7.3\% of households without children; likewise,
-13\% of households with children indicated that they ``{[}c{]}ouldn't
-afford balanced meals'' sometimes or often, compared to 8.8\% of
-households without children.
-
-The impacts of food insecurity during the pandemic are alarming, since
-diet-related diseases, such as obesity, heart-disease, and diabetes,
-were already critical public health concerns in Canada prior to COVID-19
-(Boucher et al., 2017). While food banks are not necessarily a stable
-solution to food insecurity and in fact may encourage a retrenchment of
-neoliberal policy (Wakefield et al., 2013), at least they can be argued
-to provide a resource of last instance to households in precarious
-situations (Bazerghi et al., 2016). As recently as 2019, the Hamilton
-Hunger Report\footnote{https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf}
-noted that food banks in Hamilton, Ontario, recorded the highest number
-of visitors in the past 29 years; the number of children visiting food
-banks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278
-the year before, a rate of increase greater than population growth.
-
-It is known that urban food environments, within which people make their
-daily food choices, are essential in influencing eating behaviours and
-health outcomes, based on factors such as food availability, ease of
-geographic accessibility and socio-demographic variations (Paez et al.,
-2010; Vanderlee and L'Abbé, 2017; Widener, 2018). However, while there
-is a wealth of literature that has examined the topic of geographic
-accessibility to healthy food through the ``food desert'' concept, there
-has been little research into accessibility to food banks. Although
-previous work has explored differences in \emph{accessing} food banks,
-such as how some households utilize food banks over short periods of
-time while others regularly utilize food banks as longer-term resource
-(e.g., Enns et al., 2020), we are not aware of any research that has
-focused on the geographic component of accessibility.
-
-The study of place-based geographic accessibility is concerned with
-capturing the potential to reach destinations of value using the
-transportation network (Páez et al., 2012). Indeed, the Government of
-Canada's recent Food Policy has made ``access'' to healthy food a
-priority for Canadian communities\footnote{https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042}
-and previous survey research has suggested that such accessibility plays
-a key role in user satisfaction with food bank service delivery (Holmes
-et al., 2018). However, as with research into the prevalence of food
-deserts, accessibility to food banks is unlikely to be evenly
-distributed, and variation throughout a city can be expected due to
-transportation network characteristics, and the spatial distribution of
-food bank locations and the population they are meant to serve.
-Furthermore, policy responses to the COVID-19 pandemic likely have added
-to the distress of vulnerable households. Non-pharmaceutical
-interventions during the pandemic involving restrictions in mobility
-have increased the friction of travel, in particular by transit on which
-low income populations are more reliant (e.g., DeWeese et al., 2020). At
-the same time, the pandemic has created additional stress for the
-operators of food banks through disruptions in the supply chain (e.g.,
-McKay et al., 2021) as well as concerns surrounding the delivery of
-service in safe conditions and possible cancellation of food service
-programs.
-
-For this study, we aim to look at how the landscape of food banks and
-related services (e.g.~low-cost or free meal service providers)
-available in Hamilton, Ontario, changed during the pandemic. Did the
-number of open food bank services diminish? If so, what was the
-accessibility to food banks before the pandemic from the perspective of
-low income households, and how it changed during the pandemic? And
-finally, who are most likely to have been impacted by changes in the
-accessibility landscape? This paper first looks at the distribution of
-food banks and related services before and during the pandemic. Then, we
-use the balanced floating catchment area approach of Paez et al. (2019)
-to investigate the accessibility situation. For this, we use a fully
-disaggregated approach based on parcel-level data. Socio-economic and
-demographic data are drawn from the latest Census of Canada (2016),
-whereas travel information is from the most recent regional travel
-survey from 2016. This paper follows reproducible research
-recommendations (see Brunsdon and Comber, 2020), and the research was
-conducted using open source tools for transportation analysis (Lovelace,
-2021). The code and data necessary to reproduce the analysis are
-available in a public repository\footnote{add repository}.
-
-\hypertarget{food-insecurity-and-food-bank-use-in-canada}{%
-\section{Food Insecurity and Food Bank Use in
-Canada}\label{food-insecurity-and-food-bank-use-in-canada}}
-
-Food insecurity is the inability to acquire and consume an adequate
-amount or good quality food, leading to inadequate nutrient intake (Enns
-et al., 2020; Kirkpatrick and Tarasuk, 2008; Tarasuk and Vogt, 2009).
-This nutrient deficiency is associated with major population health
-concerns, particularly among Canadians at socio-economic disadvantage
-(Bazerghi et al., 2016). Data on food insecurity in Canada is collected
-in several ways. Quantitatively, official government surveys such as the
-Household Food Security Survey Module (HFSSM), the Canadian Community
-Health Surveys (CCHS), the International Study of Adults (LISA), and
-official classifications determined by Health Canada in relation to
-socio-demographic variables offer some insight into food insecurity
-(Gundersen et al., 2018; Kirkpatrick and Tarasuk, 2008; Tarasuk and
-Vogt, 2009). Nationally, it has been found that food insecurity impacts
-approximately 12.3\% of Canadian households (Tarasuk et al., 2014)
-
-Food banks - sometimes also referred to as `food pantries' and `food
-shelves' - originated as a community response to aid those with
-inadequate food by voluntarily offering them meals and ingredients
-(Loopstra and Tarasuk, 2012; Riches, 2002). The scope and objectives of
-food banks can vary by region and by country, and these organizations
-can include not only prepared meals and aliments, but also shared spaces
-to connect in community gardens and community kitchens (Wakefield et
-al., 2013). Although in their origin food banks were meant to be provide
-a temporary solution to accommodate those in hunger due to job
-retrenchments and economic downfalls since the 1980s, over time they
-have evolved into a community practice to secure emergency food supplies
-for those in need (Loopstra and Tarasuk, 2012; Wakefield et al., 2013).
-
-In Canada, the number of food banks has steadily increased in the past
-few decades (Wakefield et al., 2013). The largest database of food banks
-and their use comes from the non-profit association Food Banks Canada
-(FBC), which conducts an annual assessment through its affiliated
-members. FBC's 2018 Hunger Count report\footnote{https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018\_FINAL\_EN.pdf.aspx?ext=.pdf}
-(the most recent available) listed 1,830 member food banks across the
-country, and found that Canadians visited food banks 1.1 million times
-in March of 2018. Of those accessing food banks, certain population
-characteristics tend to be over-represented compared to national totals
-from the 2016 Canadian Census of Population. According to FBC's 2018
-data, single-adult households represent 45\% of those utilizing food
-banks despite making up 28\% of Canada's population, 19\% are
-single-parent households (compared to 10\% nationally), and 35\% of
-those accessing food bank services are children aged 0-18 even though
-their share of Canada's national population is approximately 20\%. In
-addition, 59\% of households accessing food banks list social or
-disability assistance as their primary source of income. Similarly,
-using data from the 2011-2012 CCHS, Tarasuk et al. (2019) found higher
-odds of food insecurity amongst households relying on social assistance,
-those without a university degree or with children under the age of 18,
-and individuals that lived alone, renters, and those identifying as
-Aboriginal. While surveys revealed that only 20 to 30 percent of those
-experiencing food insecurity were found to frequent food banks in Canada
-(Tarasuk et al., 2014), research from Ottawa (Enns et al., 2020) and
-Vancouver (Black and Seto, 2020) suggests that long-term users tend to
-be older, have health or mobility challenges, live in large households,
-and are less likely to have employment income.
-
-Economic stress caused by the COVID-19 pandemic, rising unemployment
-rates, and changes in poverty levels have disrupted the food environment
-and led to higher rates of food insecurity (Niles et al., 2020).
-Pandemic-related food security studies in the US have found a
-substantial increase in households experiencing food insecurity for the
-first time, and also in households experiencing more severe food
-insecurity than before (Niles et al., 2020; Wolfson and Leung, 2020).
-Most recently in May 2020, Canada recorded 14.7\% of its population
-living in food insecurity in the past 30 days (Statistics Canada,
-2020a). Considering the negative mental and physical health effects
-associated with food insecurity, increases in food insecurity rates due
-to the pandemic, signal a change in the food environment with potential
-damages to population health during the course of the pandemic and
-beyond (Niles et al., 2020). Recent data\footnote{https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report\_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis\_Exec-Sum\_EN.pdf}
-from FBC showed that 52\% of member food banks reported an increase in
-usage in March of 2020 when initial lockdown restrictions were put in
-place across much of the country. The pandemic also created significant
-staffing issues with 42\% of food banks reporting a reduction in
-volunteers. However, 53\% of food banks later reported a decrease in use
-into the summer of 2020 which FBC members attributed to emergency
-financial support programs from the federal government. Nevertheless,
-some of these benefit programs were temporary, which suggests that many
-households may again turn to food bank services to meet their needs.
-
-In terms of geography, previous research conducted at the provincial
-scale using data from the 2011-2012 CCHS found that the prevalence of
-food insecurity ranged across the country from 11.8\% of households in
-Ontario to 41\% of households in Nunavut (Tarasuk et al., 2019).
-However, although previous research has examined the characteristics of
-individuals and households accessing food banks, the locational or
-transportation accessibility aspect of food bank access is not well
-understood. A wealth of literature examining the food desert concept
-suggests that, in addition to socio-economic and demographic factors,
-location and transportation networks play a key role in a household's
-accessibility to healthy foods (Paez et al., 2010; Vanderlee and L'Abbé,
-2017; Widener, 2018). For food banks specifically, previous qualitative
-research in Ontario by Smith-Carrier et al. (2017) has noted that
-``transportation can be challenging, particularly if the food bank is
-situated in a remote location'' (p.~32). Particularly, it appears that
-participants experience challenges with the ``inordinate amount of time
-necessary to obtain food, and difficulties associated with
-transportation'' (p.~39). Users of food banks, according to this
-research, rely on a variety of modes of transportation to access
-services. Consequently, the location of facilities matters; in the words
-of an interviewee: ``I wish it {[}the food bank{]} was a little more
-centrally located. Because if I didn't have a bike I'd have to walk it
-all the way out there and back. I wonder about people who don't''
-(p.~39).
-
-To offer greater insight into the role of transportation and location in
-food bank accessibility, this research examines how geographic
-accessibility to food banks and food services changed in Hamilton during
-the COVID-19 pandemic.
-
-\hypertarget{methods-and-materials}{%
-\section{Methods and Materials}\label{methods-and-materials}}
-
-\hypertarget{methods}{%
-\subsection{Methods}\label{methods}}
-
-For the research in this paper we adopt the balanced floating catchment
-area approach of Paez et al. (2019). This method for estimating
-accessibility is a form of the widely-used two-stage floating catchment
-area method (Luo and Wang, 2003; Radke and Mu, 2000). Floating catchment
-areas are used to estimate accessibility when there are potential
-congestion effects, and operate by calculating first the \emph{demand}
-for spatially distributed services. The demand (usually the number of
-people who require a service) is used to calculate a level of service.
-In a second step, the level of service is allocated back to the
-population. Demand and level of service are allocated using some form of
-distance-decay to embody the geographical principle that, given a
-choice, people prefer to travel less than more when reaching
-destinations.
-
-More formally, the first step of this method is as follows: \[
-L_j = \frac{S_j}{\sum_{i=1}^nP_iw_{ij}}
-\]
-
-\noindent where \(S_j\) is the level of supply at location \(j\), in
-simplest terms whether a service point is present (i.e., \(S_j=1\)) or
-not (i.e., \(S_j=0\)); \(P_i\) is the population at location \(i\) that
-demands the service; and \(w_{ij}\) is a weight, typically a function of
-the distance between locations \(i\) and \(j\). \(L_j\) is the level of
-service at location \(j\) and it is the inverse of the number of people
-that need to be serviced.
-
-The second step in this process is then summing the level of service
-that each population unit can reach, according to the distance-decay
-weight: \[
-A_i = \sum_{j=1}^JL_jw_{ji}
-\]
-
-\noindent where \(A_i\) is the accessibility to the service, which is in
-the same units as the level of service: as the inverse of the population
-being serviced. When the population being serviced is low accessibility
-is high (i.e., there is little competition for the service), and
-viceversa.
-
-Floating catchment area methods are prone to overestimation of the
-population and the level of service due to multiple-counting. The
-population at \(P_i\) is allocated to \emph{every} service point \(j\)
-for which \(w_{ij}>0\). Similarly, the level of service at \(LOS_j\) is
-allocated to \emph{every} population point for which \(w_{ji}>0\). This
-inflation effect has been known for several years, and several
-modifications have been proposed to mitigate it (Delamater, 2013; e.g.,
-Wan et al., 2012). A definitive solution to this issue was presented by
-Paez et al. (2019). In order to avoid the multiple-counting in the
-summations, the population and the level of service need to be allocated
-\emph{proportionally}. This is achieved by standardizing the weights as
-follows: \[
-w_{ij}^\text{st} = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
-\] \noindent and: \[
-w_{ji}^\text{st} = \frac{w_{ji}}{\sum_{j=1}^Jw_{ji}}
-\]
-
-The standardized weights satisfy the following conditions: \[
-\sum_{i=1}^nw_{ij}^\text{st}=1
-\]
-
-\noindent and: \[
-\sum_{j=1}^Jw_{ji}^\text{st}=1
-\]
-
-Since the population is allocated proportionally, its value is
-preserved: \[
-\sum_{i=1}^nP_iw_{ij}^\text{st}=P_i
-\]
-
-\noindent as is the level of service: \[
-\sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
-\] \#\# Study Area
-
-\textbf{Need to provide some context here: the geography which makes
-transportation challenging, i.e., due to the escarpment; the main parts
-of the city: city core, inner suburbs, industrial north, suburban,
-exurban; and the industrial background of the city maybe, with poverty
-rates compared to other Canadian cities?}
-
-\hypertarget{data}{%
-\subsection{Data}\label{data}}
-
-Data have been prepared for sharing as a data package\footnote{add
-  repository}. The contents of the data package are described next.
-
-\hypertarget{statistics-canada}{%
-\subsubsection{Statistics Canada}\label{statistics-canada}}
-
-Population and income statistics for 2016 were retrieved at the level of
-Dissemination Areas (DAs) using the package \texttt{cancensus} (von
-Bergmann et al., 2021). DAs are the smallest publicly available census
-geography in Canada. Income data corresponds to the count of households
-by different total income groupings.
-
-\hypertarget{origins-residential-parcels}{%
-\subsubsection{Origins: Residential
-parcels}\label{origins-residential-parcels}}
-
-We converted all recorded residential land parcels in the City of
-Hamilton to points on the road network. Each point includes information
-about the number of residential units in the parcel. Next, we define
-low-income households as those having a total income of less than
-CAD40,000, which is approximately the mid-point of the low income
-cut-off (LICOs) for families in Canadian cities with populations greater
-than 500,000 in 2016, to match other Census data (Statistics Canada,
-2020b). We then ``populate'' each residential unit with the probability
-of being a low-income household based on the counts of households by
-income groups in the DA in which the parcel is located. While this
-method assumes a constant probability of low-income household status for
-all residential units in a DA, the parcel-level analysis affords a high
-level of spatial disaggregation for the accessibility analysis.
-
-\hypertarget{destinations-food-banks-and-food-service-locations}{%
-\subsubsection{Destinations: Food Banks and Food Service
-Locations}\label{destinations-food-banks-and-food-service-locations}}
-
-The locations of food banks and related food services were obtained from
-the Hamilton Public Library's Food Access Guide\footnote{http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf}.
-The guide was updated in April of 2021 to indicate any change affected
-on the services due to the pandemic. This includes modified business
-hours, a need to make reservations before frequenting, and locations
-that have completely shut down in consequence. While some food bank
-services have a specific a target population, such as prioritizing
-family with young children aged between 0 and 3 or accepting only those
-providing proof of low-income status through housing and utility
-statements, all the food bank services indicated below are designed to
-accommodate those in need of food at zero to low cost. With our focus on
-food banks and food services that offer free or low-cost meals at
-particular locations, we first removed services such as Meals on Wheels
-and other food access services such as food box, community kitchens,
-student nutrition programs, and shopping and transportation. In
-addition, two free meal services held on different days at the same
-location were collapsed into a single service point for the
-accessibility analysis. Additional details on the operations of
-individual facilities is not publicly available, and with the changes in
-operations it proved unfeasible to collect it. For this reason, the
-analysis to follow is of accessibility to the location of food banks and
-services, but not to specific services (e.g., breakfasts vs.~food
-boxes).
-
-\begin{table}
-
-\caption{\label{tab:table-food-bank-info}\label{tab:food-bank-info}foodbank and Food Service Information.}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{c>{\centering\arraybackslash}p{15em}cc>{\centering\arraybackslash}p{15em}}
-\toprule
-Type & Description & Locations Pre-COVID & Locations During COVID & Additional Notes\\
-\midrule
-\cellcolor{gray!6}{Congregate Dining} & \cellcolor{gray!6}{Congregate and dining programs provide low-cost meals that are enjoyed in a community setting.
-Transportation may be provided} & \cellcolor{gray!6}{7} & \cellcolor{gray!6}{2} & \cellcolor{gray!6}{One remaining location reduced hours during COVID}\\
-Community Meals & Programs often run by volunteers that organize suppers, lunches or other get-togethers that give community residents an opportunity to meet one another in a friendly and informal atmosphere while sharing a meal & 11 & 9 & NA\\
-\cellcolor{gray!6}{Food Banks} & \cellcolor{gray!6}{Food Banks and Emergency Food programs provide individuals and families with grocery items free of charge} & \cellcolor{gray!6}{27} & \cellcolor{gray!6}{26} & \cellcolor{gray!6}{One remaining location reduced hours during COVID while 4 others moved to appointment only}\\
-Free Meals & Meals are provided free of charge in the community through volunteer labour and donations & 9 & 5 & One remaining location reduced hours during COVID\\
-\cellcolor{gray!6}{Low-Cost Meals} & \cellcolor{gray!6}{Restaurants, cafeterias and other eating establishments operated by hospitals, senior centers or other organizations which provide reduced-cost meals for low-income people, older adults or other targeted individuals.} & \cellcolor{gray!6}{2} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{The remaining location reduced hours during COVID}\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\hypertarget{routing-and-travel-time-tables}{%
-\subsubsection{Routing and travel time
-tables}\label{routing-and-travel-time-tables}}
-
-Travel time tables for three modes (car, transit, walking) were computed
-using the parcels as the origins and the locations of the food banks as
-the destinations. For routing, the package \texttt{r5r} (Pereira et al.,
-2021b) was used with a network extract for the City of Hamilton from
-OpenStreetMaps and the General Transit Feed Specification from Hamilton
-Street Railway, the local transit operator. For routing purposes we used
-maximum values of 180 min and 10,000 m walking distance: any destination
-that exceeded these thresholds was ignored. The departure time used for
-routing was 8:00AM on March 30, 2021 to reflect transit service around
-the morning service peak on a typical Tuesday.
-
-\hypertarget{transportation-tomorrow-survey}{%
-\subsubsection{Transportation Tomorrow
-Survey}\label{transportation-tomorrow-survey}}
-
-We used the Data Retrieval System of the Transportation Tomorrow Survey
-(TTS)\footnote{\url{http://dmg.utoronto.ca/}} to download
-cross-tabulations of: 1) primary mode of travel per trip by income by
-place of residence; and 2) age by income by place of residence. These
-data are from the 2016 Survey (the most recent available), and data are
-geocoded at the level of Traffic Analysis Zones (TAZ) using the most
-recent zoning system from 2006. Each parcel point is populated with the
-proportion of trips by three modes of travel: car (as driver or
-passenger), transit, and walk.
-
-\hypertarget{expected-travel-times}{%
-\subsubsection{Expected Travel Times}\label{expected-travel-times}}
-
-Once we obtained travel time tables with population (number of
-households) and proportion of trips by mode, we calculated the expected
-travel time \(ett\) from each parcel \(i\) to a food bank or food
-service location \(j\) as follows: \[
-ett_{ij} = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
-\]
-
-\noindent where \(p^k_i\) is the proportion of trips by mode \(k\) in
-the TAZ of parcel \(i\), and \(tt^k_{ij}\) is the travel time from
-parcel \(i\) to the food bank. In other words, the expected travel time
-is the weighted sum of travel times to the food bank, with the weights
-given by the expected modal split in the TAZ.
-
-\hypertarget{results-and-discussion}{%
-\section{Results and Discussion}\label{results-and-discussion}}
-
-Figure \ref{fig:foodbanks} shows the location of food banks and services
-in the City of Hamilton and their status. Before the pandemic there were
-58 of which 14 (24.14\%) closed during the pandemic. As shown in the
-figure, food services tend to be predominantly located in the central
-parts of the city. This is not surprising: population density is high
-there, and it is also the part of the city where lower income households
-are more numerous in absolute and relative terms (see Figure
-\ref{fig:low-income-households}). Alas, this is also the part of the
-city where most of the closures during the pandemic happened.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-location-foodbanks-1} \caption{\label{fig:foodbanks}Location of food banks/services and operation status; the dotted box is an inset of the central part of the City of Hamilton}\label{fig:plot-location-foodbanks}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-low-income-households-1} \caption{\label{fig:low-income-households}Number and proportion of households with incomes less than CAD40,000.}\label{fig:plot-low-income-households}
-\end{figure}
-
-To implement the accessibility calculations, we must select a
-distance-decay function. In this task we find limited support in the
-literature, which is mostly silent on the travel patterns of people who
-visit food banks. For this reason, we opt for a simple cumulative
-opportunities function as follows: \[
-w_{ij}=w_{ji}=
-\begin{cases}
-1 & \text{ if } ett_{ij}\le \delta\\
-0 & \text{ otherwise}
-\end{cases}
-\] \noindent where \(ett_{ij}\) is the multimodal expected travel time
-as described previously, and \(\delta\) is a travel threshold. When the
-expected travel time exceeds this threshold, a facility is no longer
-considered accessible. Moreover, the weights are standardized for the
-balanced floating catchment area approach.
-
-Figure \ref{fig:sensitivity-analysis} shows the results of conducting
-sensitivity analysis of the system-wide accessibility as we vary the
-threshold (considering the situation before the pandemic). There is a
-clear pattern whereby more strict values of \(\delta\) are associated
-with higher levels of system-wide accessibility: while increases in
-accessibility that result from decreases in the travel time window might
-seem counter-intuitive, this is a result of lower \emph{congestion},
-since fewer households are serviced and thus competition for the same
-resources is more limited. System-wide accessibility declines with
-higher values of \(\delta\): as more households are serviced, congestion
-grows and the level of service declines, although this happens at a
-declining rate. We are not aware of any research that explains how long
-people are expected to travel for food banks, but we note that in
-developing countries, access to drinking water is defined as a source
-that takes less than 30 minutes to reach (round trip, see UNICEF-WHO,
-2019). There is no reason why people in affluent countries should be
-expected to travel more for a basic need such as food. Accordingly, we
-adopt a 15-minute threshold for the analysis (representing a one-way
-trip). This threshold is also approximately where the rate of change in
-accessibility slows down.
-
-\begin{figure}
-\includegraphics[width=0.6\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-results-sensitivity-analysis-1} \caption{\label{fig:sensitivity-analysis}Accessibility as a function of threshold}\label{fig:plot-results-sensitivity-analysis}
-\end{figure}
-
-Using the 15-minute threshold, we find that the system-wide
-accessibility (interpreted as a provider-to-population ratio) was 0.065
-(food banks/service locations per low income household in the city)
-before COVID-19, but declined to 0.037 during the pandemic. It is
-striking that although almost 76\% of food facilities remained in
-operation during the pandemic, there was a loss of accessibility greater
-than 43\%, suggesting the location of food banks and related services
-plays an important role in serving those in need.
-
-Turning to the location of individual facilities, the levels of service
-offered before and during the pandemic are shown in Figure
-\ref{fig:levels-of-service}. The level of service is functionally the
-inverse of the number of low-income households in the travel-mode
-weighted travel time catchment area of the facilities (this is because
-\(S_j=1 \forall j\), i.e., each location represents a ``capacity'' of
-1). Higher values mean that a facility is expected to service fewer
-households. Conversely, lower values indicate greater congestion.
-
-The general pattern of the levels of service is similar before and
-during the pandemic, with lower values in the center of the city where
-low-income households exhibit multimodal trip patterns. Three more
-peripheral facilities towards the south of the city also have lower
-levels of service, presumably because they are expected to service
-relatively large suburban/exurban populations generally reliant on
-automobiles for travel. During the pandemic, however, the levels of
-service dropped, in some cases quite substantially. The pattern of the
-losses in level of service, moreover, is not uniform. The upper pane of
-Figure \ref{fig:levels-of-service-changes} shows that the three
-peripheral facilities in the southern suburban/exurban part of the city
-had low levels of service to begin with, but did not see major declines
-during the pandemic. Further, the inset map shows that the levels of
-service deteriorated more in the central part of the city. However, the
-loss of level of service was not as large in the core (where most of the
-food banks/services are found), but instead was more marked in the inner
-ring around the core, where facilities may have faced greater demand
-from both central city and suburban populations after the closure of
-food banks during the pandemic.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-levels-of-service-1} \caption{\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (bottom panel).}\label{fig:plot-levels-of-service}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-levels-of-service-changes-1} \caption{\label{fig:levels-of-service-changes}Changes in levels of service at each facility from pre-COVID-19 to during COVID-19.}\label{fig:plot-levels-of-service-changes}
-\end{figure}
-
-To further elucidate this issue, we now turn to the results of the
-accessibility analysis. As with the level of service of individual
-facilities, the general pattern of accessibility before and during the
-pandemic is similar. Figure \ref{fig:accessibility} reveals that,
-compared with the outer rural zones, the more urban zones of the city
-generally exhibit higher accessibility to food banks and food service
-locations. However, the pattern is not particularly smooth - this is
-largely attributable to the weighting of travel times by mode of
-transportation according to the trip patterns of low-income household
-respondents captured by the TTS. For example, in zones where low-income
-households make a high proportion of trips by walking, access to food
-bank locations by walking is afforded a concomitantly high weight in our
-calculations of travel time compared to transit or car travel. From
-this, highly-accessible locations result from a mix of characteristics:
-low-income households in locations where travel options that align with
-zonal modal split are available to connect them to food bank locations
-with high levels of service within 15 minutes. This seems to track with
-the experience of some users of these services, as reported by
-Smith-Carrier et al. (2017).
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-accessibility-1}
-
-}
-
-\caption{\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (bottom panel).}\label{fig:plot-accessibility}
-\end{figure}
-
-We find that the accessibility landscape deteriorated substantially
-during the pandemic, with accessibility dropping on average by almost
-41\%, but with large variations: some zones experienced changes in
-accessibility of only about 10\%, whereas the most affected zone saw a
-loss of accessibility of almost 71\%. Figure
-\ref{fig:accessibility-changes-with-local-i} shows the changes in
-accessibility. Every zone is worse off after the closure of facilities
-during the pandemic, but some parts of the city seem to have been
-particularly affected. To better highlight these changes, we used a
-local indicator of spatial autocorrelation (Anselin, 1995) to explore
-the pattern of change in accessibility. Twenty-four TAZ are flagged as
-having significant large losses of accessibility (at \(p<=0.10\),
-without correcting for multiple comparisons). Those zones are
-highlighted in the figure, where it can be seen that they form more or
-less compact neighborhoods. Remarkably, the largest (and significant)
-drops in accessibility are not downtown, but located in two cases in the
-industrial north of the city, in one case in an inner suburb above the
-escarpment, and lastly in a more suburban/exurban region in the
-south-west.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-local-i-1} \caption{\label{fig:accessibility-changes-with-local-i}Changes in accessibility from pre-COVID-19 to during COVID-19.
-Highlighted areas had significantly large changes in accessibility according to Local Moran's I.}\label{fig:plot-local-i}
-\end{figure}
-
-For the more suburban clusters of zones, the decrease in accessibility
-is derived from the closure of locations throughout the city reachable
-by car. In the cluster of central suburban zones for example, low-income
-households in the outer ring of zones that exhibit medium to high
-decreases in accessibility within this cluster appear to be largely
-auto-dependent in their tripmaking, which each each exhibiting between
-85-100\% of their modal split for car trips. This results in the parcels
-within these zones having a large number of potentially accessible
-locations in the travel time matrix. But by extension, the change in
-accessibility over the pre- and during-COVID-19 time periods is affected
-not only by the closure of service locations proximate to the zones, but
-also the locations in the central city. The zone with the greatest
-decrease in accessibility within this cluster
-(\ensuremath{-9\times 10^{-4}}) has a high rate of car trips and
-connects to the most food bank locations in total as well as those that
-stayed open or closed. The exception to this is the zone in the middle
-of this cluster, which has a much higher proportion of transit use 32\%.
-This modal split weighting results in fewer food bank and service
-locations accessible from the parcels within this zone which, by
-extension, limits their access to closed locations reachable within 15
-minutes in our calculations.
-
-In the cluster to the south-west, the decrease in accessibility is
-predominately driven by the closure of a high level-of-service Community
-Meals provider. However, like the more central suburban zones,
-low-income households within this cluster also between 90\% to 100\%
-auto-dependent in their tripmaking in the TTS. The story is similar for
-the zone located in the north-west that exhibits the greatest decrease
-in accessibility. Here, low-income households responding to the TTS
-conducted 100\% of their trips by car and, as a result, parcels within
-this zone have access to the second-highest number of food bank and
-service locations within 15 minutes. However, this also means zonal
-accessibilities are greatly affected by the number of closures
-throughout the city. Finally, in the city's north end and north-east
-zones, low-income households exhibit a mixture of tripmaking behaviour
-in the TTS. Households in some zones take transit more often and one
-zone in particular has 100\% of its trips by walking while in others,
-car use makes up 100\% of trips. For these zones, the decrease in
-accessibility tends to be a product of the closure of several inner-city
-food bank and service locations.
-
-Just as the effects of the closures appear to have been uneven in space,
-they also seem to have differently impacted various population segments.
-Table \ref{tab:accessibility-by-age} shows the estimated number of
-people in each age group by their level of accessibility before and
-during the pandemic. Here, it is important to note that the quartiles
-are relative: people in the top 25\% of accessibility still have lower
-accessibility during the pandemic than before. In reality, every
-population group is worse off during the pandemic in terms of their
-accessibility to food banks and services in the City of Hamilton.
-However, some age groups were affected more. As seen in the table, the
-distribution within quartiles for adults changed only slightly, and in
-fact there are fewer adults in the bottom three quartiles as some found
-themselves in the (deteriorated) top quartile of accessibility during
-the pandemic. The story is different for seniors, a greater number of
-whom are now in the third quartile, due to a combination of seniors
-facing a worse accessibility situation, and just a few moving from the
-bottom quartile (pre-pandemic) to the third during the pandemic. Those
-aged 18 and less appear to have fared worse, and the largest change is
-in the number of children who were in the third quartile before the
-pandemic found themselves in the bottom accessibility class during the
-pandemic. These results suggest that, through a combination of the
-typical modes of transportation of lower income households, and the
-spatial distribution of the population, the food bank/service locations
-closed had a differential impact that more greatly affected the youngest
-among the population in low income households.
-
-\begin{table}[!h]
-
-\caption{\label{tab:table-accessibility-by-age-group}\label{tab:accessibility-by-age}Population at each accessibility level by age group among members households with incomes less than CAD40,000.}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{lcccccc}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Pre-COVID-19} & \multicolumn{3}{c}{During COVID-19} \\
-\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
-Accessibility Quartile & Children (age $\le$ 18) & Adults (19-64) & Seniors (age $\ge$ 65) & Children (age $\le$ 18) & Adults (19-64) & Seniors (age $\ge$ 65)\\
-\midrule
-\cellcolor{gray!6}{Top 25\%} & \cellcolor{gray!6}{2,546} & \cellcolor{gray!6}{10,404} & \cellcolor{gray!6}{5,008} & \cellcolor{gray!6}{2,567 (0.82\%)} & \cellcolor{gray!6}{10,597 (1.86\%)} & \cellcolor{gray!6}{4,941 (-1.34\%)}\\
-Second 25\% & 3,333 & 12,916 & 6,415 & 3,375 (1.26\%) & 12,885 (-0.24\%) & 6,157 (-4.02\%)\\
-\cellcolor{gray!6}{Third 25\%} & \cellcolor{gray!6}{4,381} & \cellcolor{gray!6}{11,559} & \cellcolor{gray!6}{7,067} & \cellcolor{gray!6}{4,062 (-7.28\%)} & \cellcolor{gray!6}{11,412 (-1.27\%)} & \cellcolor{gray!6}{7,678 (8.65\%)}\\
-Bottom 25\% & 3,546 & 12,946 & 6,903 & 3,802 (7.22\%) & 12,897 (-0.38\%) & 6,567 (-4.87\%)\\
-\bottomrule
-\multicolumn{7}{l}{\rule{0pt}{1em}\textit{Note: }}\\
-\multicolumn{7}{l}{\rule{0pt}{1em}Population values have been rounded.}\\
-\multicolumn{7}{l}{\rule{0pt}{1em}The values in brackets for population during COVID-19 are the changes from before the pandemic.}\\
-\end{tabular}}
-\end{table}
-
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
-
-\textbf{LIMITATIONS}
-
-Words go here.
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-anselin1995local}{}%
-Anselin, L., 1995. Local indicators of spatial association - LISA.
-Geographical Analysis 27, 93--115.
-
-\leavevmode\hypertarget{ref-bazerghi2016role}{}%
-Bazerghi, C., McKay, F.H., Dunn, M., 2016. The role of food banks in
-addressing food insecurity: A systematic review. Journal of community
-health 41, 732--740.
-
-\leavevmode\hypertarget{ref-bhattacharya2004poverty}{}%
-Bhattacharya, J., Currie, J., Haider, S., 2004. Poverty, food
-insecurity, and nutritional outcomes in children and adults. Journal of
-health economics 23, 839--862.
-doi:\url{https://doi.org/10.1016/j.jhealeco.2003.12.008}
-
-\leavevmode\hypertarget{ref-black2020examining}{}%
-Black, J.L., Seto, D., 2020. Examining patterns of food bank use over
-twenty-five years in vancouver, canada. VOLUNTAS: International Journal
-of Voluntary and Nonprofit Organizations 31, 853--869.
-doi:\url{https://doi.org/10.1007/s11266-018-0039-2}
-
-\leavevmode\hypertarget{ref-boucher2017ontario}{}%
-Boucher, B.A., Manafò, E., Boddy, M.R., Roblin, L., Truscott, R., 2017.
-The ontario food and nutrition strategy: Identifying indicators of food
-access and food literacy for early monitoring of the food environment.
-Health promotion and chronic disease prevention in Canada : research,
-policy and practice 37, 313--319.
-doi:\href{https://doi.org/10.24095/hpcdp.37.9.06}{10.24095/hpcdp.37.9.06}
-
-\leavevmode\hypertarget{ref-brunsdon2020opening}{}%
-Brunsdon, C., Comber, A., 2020. Opening practice: Supporting
-reproducibility and critical spatial data science. Journal of
-Geographical Systems 1--20.
-doi:\href{https://doi.org/10.1007/s10109-020-00334-2}{10.1007/s10109-020-00334-2}
-
-\leavevmode\hypertarget{ref-delamater2013spatial}{}%
-Delamater, P.L., 2013. Spatial accessibility in suboptimally configured
-health care systems: A modified two-step floating catchment area
-(M2SFCA) metric. Health \& Place 24, 30--43.
-doi:\href{https://doi.org/10.1016/j.healthplace.2013.07.012}{10.1016/j.healthplace.2013.07.012}
-
-\leavevmode\hypertarget{ref-deweese2020tale}{}%
-DeWeese, J., Hawa, L., Demyk, H., Davey, Z., Belikow, A., El-geneidy,
-A., 2020. A tale of 40 cities: A preliminary analysis of equity impacts
-of COVID-19 service adjustments across north america. Findings.
-doi:\href{https://doi.org/10.32866/001c.13395}{10.32866/001c.13395}
-
-\leavevmode\hypertarget{ref-elgar2021relative}{}%
-Elgar, F.J., Pickett, W., Pförtner, T.-K., Gariépy, G., Gordon, D.,
-Georgiades, K., Davison, C., Hammami, N., MacNeil, A.H., Da Silva, M.A.,
-others, 2021. Relative food insecurity, mental health and wellbeing in
-160 countries. Social Science \& Medicine 268, 113556.
-doi:\url{https://doi.org/10.1016/j.socscimed.2020.113556}
-
-\leavevmode\hypertarget{ref-enns2020experiences}{}%
-Enns, A., Rizvi, A., Quinn, S., Kristjansson, E., 2020. Experiences of
-food bank access and food insecurity in ottawa, canada. Journal of
-Hunger \& Environmental Nutrition 15, 456--472.
-doi:\url{https://doi.org/10.1080/19320248.2020.1761502}
-
-\leavevmode\hypertarget{ref-gundersen2018food}{}%
-Gundersen, C., Tarasuk, V., Cheng, J., De Oliveira, C., Kurdyak, P.,
-2018. Food insecurity status and mortality among adults in ontario,
-canada. PloS one 13, e0202642.
-
-\leavevmode\hypertarget{ref-holmes2018nothing}{}%
-Holmes, E., Black, J.L., Heckelman, A., Lear, S.A., Seto, D., Fowokan,
-A., Wittman, H., 2018. {``Nothing is going to change three months from
-now''}: A mixed methods characterization of food bank use in greater
-vancouver. Social Science \& Medicine 200, 129--136.
-doi:\url{https://doi.org/10.1016/j.socscimed.2018.01.029}
-
-\leavevmode\hypertarget{ref-jones2017food}{}%
-Jones, A.D., 2017. Food insecurity and mental health status: A global
-analysis of 149 countries. American journal of preventive medicine 53,
-264--273. doi:\url{https://doi.org/10.1016/j.amepre.2017.04.008}
-
-\leavevmode\hypertarget{ref-kirkpatrick2008food}{}%
-Kirkpatrick, S.I., Tarasuk, V., 2008. Food insecurity is associated with
-nutrient inadequacies among canadian adults and adolescents. The Journal
-of nutrition 138, 604--612.
-doi:\url{https://doi.org/10.1093/jn/138.7.1399}
-
-\leavevmode\hypertarget{ref-laborde2020poverty}{}%
-Laborde, D., Martin, W., Vos, R., 2020. Poverty and food insecurity
-could grow dramatically as COVID-19 spreads. International Food Policy
-Research Institute (IFPRI), Washington, DC.
-
-\leavevmode\hypertarget{ref-loopstra2012relationship}{}%
-Loopstra, R., Tarasuk, V., 2012. The relationship between food banks and
-household food insecurity among low-income toronto families. Canadian
-Public Policy 38, 497--514.
-doi:\href{https://doi.org/10.3138/cpp.38.4.497}{10.3138/cpp.38.4.497}
-
-\leavevmode\hypertarget{ref-lovelace2021open}{}%
-Lovelace, R., 2021. Open source tools for geographic analysis in
-transport planning. Journal of Geographical Systems.
-doi:\href{https://doi.org/10.1007/s10109-020-00342-2}{10.1007/s10109-020-00342-2}
-
-\leavevmode\hypertarget{ref-luo2003measures}{}%
-Luo, W., Wang, F.H., 2003. Measures of spatial accessibility to health
-care in a GIS environment: Synthesis and a case study in the chicago
-region. Environment and Planning B-Planning \& Design 30, 865--884.
-
-\leavevmode\hypertarget{ref-mckay2021exploring}{}%
-McKay, F.H., Bastian, A., Lindberg, R., 2021. Exploring the response of
-the victorian emergency and community food sector to the COVID-19
-pandemic. Journal of Hunger \& Environmental Nutrition 1--15.
-doi:\href{https://doi.org/10.1080/19320248.2021.1900974}{10.1080/19320248.2021.1900974}
-
-\leavevmode\hypertarget{ref-men2021food}{}%
-Men, F., Tarasuk, V., 2021. Food insecurity amid the COVID-19 pandemic:
-Food charity, government assistance and employment. Canadian Public
-Policy COVID-19, e2021001.
-doi:\href{https://doi.org/10.3138/cpp.2021-001}{10.3138/cpp.2021-001}
-
-\leavevmode\hypertarget{ref-niles2020early}{}%
-Niles, M.T., Bertmann, F., Belarmino, E.H., Wentworth, T., Biehl, E.,
-Neff, R., 2020. The early food insecurity impacts of COVID-19. Nutrients
-12, 2096.
-
-\leavevmode\hypertarget{ref-olson1999nutrition}{}%
-Olson, C.M., 1999. Nutrition and health outcomes associated with food
-insecurity and hunger. The Journal of nutrition 129, 521S--524S.
-doi:\url{https://doi.org/10.1093/jn/129.2.521S}
-
-\leavevmode\hypertarget{ref-paez2019demand}{}%
-Paez, A., Higgins, C.D., Vivona, S.F., 2019. Demand and level of service
-inflation in floating catchment area (FCA) methods. PloS one 14,
-e0218773.
-doi:\href{https://doi.org/10.1371/journal.pone.0218773}{10.1371/journal.pone.0218773}
-
-\leavevmode\hypertarget{ref-paez2010relative}{}%
-Paez, A., Mercado, R.G., Farber, S., Morency, C., Roorda, M., 2010.
-Relative accessibility deprivation indicators for urban settings:
-Definitions and application to food deserts in montreal. Urban Studies
-47, 1415--1438.
-doi:\href{https://doi.org/10.1177/0042098009353626}{10.1177/0042098009353626}
-
-\leavevmode\hypertarget{ref-paez2012measuring}{}%
-Páez, A., Scott, D.M., Morency, C., 2012. Measuring accessibility:
-Positive and normative implementations of various accessibility
-indicators. Journal of Transport Geography 25, 141--153.
-doi:\url{https://doi.org/10.1016/j.jtrangeo.2012.03.016}
-
-\leavevmode\hypertarget{ref-pereira2021geographic}{}%
-Pereira, R.H.M., Braga, C.K.V., Servo, L.M., Serra, B., Amaral, P.,
-Gouveia, N., Paez, A., 2021a. Geographic access to COVID-19 healthcare
-in brazil using a balanced float catchment area approach. Social Science
-\& Medicine 273, 113773.
-doi:\url{https://doi.org/10.1016/j.socscimed.2021.113773}
-
-\leavevmode\hypertarget{ref-pereira2021r5r}{}%
-Pereira, R.H.M., Saraiva, M., Herszenhut, D., Braga, C.K.V., Conway,
-M.W., 2021b. r5r: Rapid realistic routing on multimodal transport
-networks with r\textsuperscript{5} in r. Findings.
-doi:\href{https://doi.org/10.32866/001c.21262}{10.32866/001c.21262}
-
-\leavevmode\hypertarget{ref-radke2000spatial}{}%
-Radke, J., Mu, L., 2000. Spatial decomposition, modeling and mapping
-service regions to predict access to social programs. Annals of
-Geographic Information Sciences 6, 105--112.
-
-\leavevmode\hypertarget{ref-ramsey2011food}{}%
-Ramsey, R., Giskes, K., Turrell, G., Gallegos, D., 2011. Food insecurity
-among australian children: Potential determinants, health and
-developmental consequences. Journal of Child Health Care 15, 401--416.
-doi:\url{https://doi.org/10.1177\%2F1367493511423854}
-
-\leavevmode\hypertarget{ref-riches2002food}{}%
-Riches, G., 2002. Food banks and food security: Welfare reform, human
-rights and social policy. Lessons from canada? Social Policy \&
-Administration 36, 648--663.
-
-\leavevmode\hypertarget{ref-seligman2010food}{}%
-Seligman, H.K., Laraia, B.A., Kushel, M.B., 2010. Food insecurity is
-associated with chronic disease among low-income NHANES participants.
-The Journal of nutrition 140, 304--310.
-
-\leavevmode\hypertarget{ref-smith2017food}{}%
-Smith-Carrier, T., Ross, K., Kirkham, J., Decker Pierce, B., 2017.
-{`Food is a right\ldots{} nobody should be starving on our streets'}:
-Perceptions of food bank usage in a mid-sized city in ontario, canada.
-Journal of Human Rights Practice 9, 29--49.
-
-\leavevmode\hypertarget{ref-statisticscanada2020food}{}%
-Statistics Canada, 2020a. Food insecurity during the COVID-19 pandemic
-(No. Catalogue no. 45280001).
-
-\leavevmode\hypertarget{ref-statisticscanada2020licos}{}%
-Statistics Canada, 2020b. Table 11-10-0241-01 low income cut-offs
-(LICOs) before and after tax by community size and family size, in
-current dollars. doi:\url{https://doi.org/10.25318/1110024101-eng}
-
-\leavevmode\hypertarget{ref-stuff2004household}{}%
-Stuff, J.E., Casey, P.H., Szeto, K.L., Gossett, J.M., Robbins, J.M.,
-Simpson, P.M., Connell, C., Bogle, M.L., 2004. Household food insecurity
-is associated with adult health status. The Journal of nutrition 134,
-2330--2335.
-doi:\href{https://doi.org/Oxford\%20University\%20Press}{Oxford University Press}
-
-\leavevmode\hypertarget{ref-tarasuk2014food}{}%
-Tarasuk, V., Dachner, N., Loopstra, R., 2014. Food banks, welfare, and
-food insecurity in canada. British Food Journal.
-
-\leavevmode\hypertarget{ref-tarasuk2020relationship}{}%
-Tarasuk, V., Fafard St-Germain, A.-A., Loopstra, R., 2020. The
-relationship between food banks and food insecurity: Insights from
-canada. VOLUNTAS: International Journal of Voluntary and Nonprofit
-Organizations 31, 841--852.
-doi:\href{https://doi.org/10.1007/s11266-019-00092-w}{10.1007/s11266-019-00092-w}
-
-\leavevmode\hypertarget{ref-tarasuk2019geographic}{}%
-Tarasuk, V., Fafard St-Germain, A.-A., Mitchell, A., 2019. Geographic
-and socio-demographic predictors of household food insecurity in canada,
-2011--12. BMC Public Health 19, 12.
-doi:\href{https://doi.org/10.1186/s12889-018-6344-2}{10.1186/s12889-018-6344-2}
-
-\leavevmode\hypertarget{ref-tarasuk2009household}{}%
-Tarasuk, V., Vogt, J., 2009. Household food insecurity in ontario.
-Canadian Journal of Public Health 100, 184--188.
-doi:\href{https://doi.org/10.1007/BF03405537}{10.1007/BF03405537}
-
-\leavevmode\hypertarget{ref-world2019progress}{}%
-UNICEF-WHO, 2019. Progress on household drinking water, sanitation and
-hygiene 2000--2017. Special focus on inequalities.
-
-\leavevmode\hypertarget{ref-vanderlee2017food}{}%
-Vanderlee, L., L'Abbé, M., 2017. Food for thought on food environments
-in canada. Health promotion and chronic disease prevention in Canada :
-research, policy and practice 37, 263--265.
-doi:\href{https://doi.org/10.24095/hpcdp.37.9.01}{10.24095/hpcdp.37.9.01}
-
-\leavevmode\hypertarget{ref-vonBergmann2021cancensus}{}%
-von Bergmann, J., Shkolnik, D., Jacobs, A., 2021. Cancensus: R package
-to access, retrieve, and work with canadian census data and geography.
-
-\leavevmode\hypertarget{ref-wakefield2013sweet}{}%
-Wakefield, S., Fleming, J., Klassen, C., Skinner, A., 2013. Sweet
-charity, revisited: Organizational responses to food insecurity in
-hamilton and toronto, canada. Critical Social Policy 33, 427--450.
-
-\leavevmode\hypertarget{ref-wan2012three}{}%
-Wan, N., Zou, B., Sternberg, T., 2012. A three-step floating catchment
-area method for analyzing spatial access to health services.
-International Journal of Geographical Information Science 26,
-1073--1089.
-doi:\href{https://doi.org/10.1080/13658816.2011.624987}{10.1080/13658816.2011.624987}
-
-\leavevmode\hypertarget{ref-widener2018spatial}{}%
-Widener, M.J., 2018. Spatial access to food: Retiring the food desert
-metaphor. Physiology \& behavior 193, 257--260.
-doi:\url{https://doi.org/10.1016/j.physbeh.2018.02.032}
-
-\leavevmode\hypertarget{ref-wolfson2020food}{}%
-Wolfson, J.A., Leung, C.W., 2020. Food insecurity and COVID-19:
-Disparities in early effects for US adults. Nutrients 12, 1648.
-
-\end{CSLReferences}
-
 \end{document}

### 2021-04-28 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-abstract.Rmd
Overall proofreading and editing of conclusions

@@
   In this paper we analyze the changes in accessibility to food banks and related services before and during the COVID-19 pandemic in the City of Hamilton, Ontario.
-Food banks and services are the last line of support for households facing food insecurity; as such, they have become essential and their relevance cannot be ignored in the midst of the economic upheaval caused by the pandemic.
+Food banks and services are the last line of support for households facing food insecurity; as such, their relevance cannot be ignored in the midst of the economic upheaval caused by the pandemic.
 Our analysis is based on the application of balanced floating catchment areas, and concentrates on households with lower incomes (<CAD40,000, approximately the Low Income Cutoff Value for a city of Hamilton's size).
-Accessibility was low to begin with in suburban and exurban parts of the city.
-We find that about 14\% of locations originally available in Hamilton closed during the pandemic; the impact of those closures on the level of service and accessibility was disproportionate, with system-wide losses exceeding 40\%.
+We find that accessibility was low to begin with in suburban and exurban parts of the city; furthermore, about 14\% of locations originally available in Hamilton closed during the pandemic, further reducing accessibility.
+The impact of closures on the level of service of the remaining facilities, and on accessibility, was disproportionate, with system-wide losses exceeding 40\%.
 Those losses were geographically and demographically uneven.

### 2021-04-28 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-title-page.Rmd
Overall proofreading and editing of conclusions

@@
   In this paper we analyze the changes in accessibility to food banks and related services before and during the COVID-19 pandemic in the City of Hamilton, Ontario.
-Food banks and services are the last line of support for households facing food insecurity; as such, they have become essential and their relevance cannot be ignored in the midst of the economic upheaval caused by the pandemic.
-Our analysis is based on the application of balanced floating catchment areas, and concentrates on households with lower incomes ($<CAD40,000, approximately the Low Income Cutoff Value for a city of Hamilton's size).
-Accessibility was low to begin with in suburban and exurban parts of the city.
-We find that about 14\% of locations originally available in Hamilton closed during the pandemic; the impact of those closures on the level of service and accessibility was disproportionate, with system-wide losses exceeding 40\%.
+Food banks and services are the last line of support for households facing food insecurity; as such, their relevance cannot be ignored in the midst of the economic upheaval caused by the pandemic.
+Our analysis is based on the application of balanced floating catchment areas, and concentrates on households with lower incomes (<CAD40,000, approximately the Low Income Cutoff Value for a city of Hamilton's size).
+We find that accessibility was low to begin with in suburban and exurban parts of the city; furthermore, about 14\% of locations originally available in Hamilton closed during the pandemic, further reducing accessibility.
+The impact of closures on the level of service of the remaining facilities, and on accessibility, was disproportionate, with system-wide losses exceeding 40\%.
 Those losses were geographically and demographically uneven.

### 2021-04-28 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Overall proofreading and editing of conclusions

@@
 Food insecurity is defined as an "inadequate or uncertain access to a sufficient quantity and/or adequate quality of food" due to a household's financial limitations [@enns2020experiences].
-This condition has been associated with reductions in nutritional outcomes [@bhattacharya2004poverty; @kirkpatrick2008food; @olson1999nutrition] and physical and mental health in children and adults [@elgar2021relative; @jones2017food; @ramsey2011food; @seligman2010food; @stuff2004household].
+This condition has been associated with reductions in nutritional outcomes [@bhattacharya2004poverty; @kirkpatrick2008food; @olson1999nutrition] and negative physical and mental health impact in children and adults [@elgar2021relative; @jones2017food; @ramsey2011food; @seligman2010food; @stuff2004household].
 Over at least the past four decades food banks and related services have become an essential line of defense against food insecurity in Canadian communities [@riches2002food; @holmes2018nothing; @black2020examining; @tarasuk2020relationship].
@@
 
-The emergence of COVID-19, the worst public health crisis since the 1918 flu pandemic, has revealed important social and economic fault lines, and pre-existing patterns of inequality appear to have been exacerbated.
-Along several other dimensions of stress [e.g., accessibility to health care facilities, @pereira2021geographic], this seems to be the case for food insecurity as well [@laborde2020poverty].
+The emergence of COVID-19, the worst public health crisis since the 1918 flu pandemic, has exposed important social and economic fault lines, and pre-existing patterns of inequality appear to have been exacerbated.
+Along several other dimensions of stress [e.g., accessibility to health care facilities, @ghorbanzadeh2021spatial; @pereira2021geographic], this seems to be the case for food insecurity as well [@laborde2020poverty].
 In the US, for example, it has been estimated that there was an increase of more than 30\% in household food insecurity, and more than one third of households were discovered to be newly food insecure - meaning they did not experience food insecurity before the pandemic [@niles2020early].
@@
 Nevertheless, some of these benefit programs were temporary.
-Although more recent statistics on food bank use in Hamilton in 2020 and 2021 is not yet available, data from the Daily Bread Food Bank in neighbouring Toronto for August 2020 shows visits climbing 51% year-over-year in that city^[https://www.dailybread.ca/wp-content/uploads/2020/11/Whos-Hungry-Report-2020.pdf], which suggests that many households in Hamilton are likely to turn to food bank services to meet their needs.
+Although more recent statistics on food bank use in Hamilton in 2020 and 2021 are not yet available, data from the Daily Bread Food Bank in neighbouring Toronto for August 2020 shows visits climbing 51% year-over-year in that city^[https://www.dailybread.ca/wp-content/uploads/2020/11/Whos-Hungry-Report-2020.pdf], which suggests that many households in Hamilton are likely to turn to food bank services to meet their needs.
 
@@
 With a population of around 540,000, the City of Hamilton is the fourth largest city in Ontario.
-It has historically been home to major manufacturing industries but de-industrialization that has occurred over the past several decades has led Hamilton to become one of the most highly divided cities in Ontario, with a significant proportion of its residents sustaining livings at Canada’s poverty level [@deluca2012code; @jakar2019turning; @latham2007determinants].
+It has historically been home to major manufacturing industries but de-industrialization that has occurred over the past several decades has led Hamilton to become one of the most highly divided cities in Ontario, with a significant proportion of its residents living at or below Canada’s poverty level [@deluca2012code; @jakar2019turning; @latham2007determinants].
 In 2016, the Hamilton Community Foundation reported that based on the Low-Income Cut-Off, Hamilton recorded a poverty rate of 16.7%, which was well above the average rate of Ontario (13.7%) and the average national rate (12.8%).^[https://www.hamiltoncommunityfoundation.ca/vital-signs/low-income-2018/].
@@
 System-wide accessibility declines with higher values of $\delta$: as more households are serviced, congestion grows and the level of service declines, although this happens at a declining rate.
-We are not aware of any research that explains how long people are expected to travel for food banks, but we note that in developing countries, access to drinking water is defined as a source that takes less than 30 minutes to reach [round trip, see @world2019progress].
-There is no reason why people in affluent countries should be expected to travel more for a basic need such as food.
+We are not aware of any research that explains how long people are expected to travel for food banks, but we note that in developing countries, accessible sources of drinking water are those that can be reached in less than 30 minutes [round trip, see @world2019progress].
+There is no reason why people in affluent countries should be expected to spend more time travelling for a basic necessity such as food.
 Accordingly, we adopt a 15-minute threshold for the analysis (representing a one-way trip).
@@
 To better highlight these changes, we used a local indicator of spatial autocorrelation [@anselin1995local] to explore the pattern of change in accessibility.
-Twenty-four TAZ are flagged as having significant large losses of accessibility (at $p<=0.10$, without correcting for multiple comparisons).
+Twenty-four TAZ are flagged as having significantly large losses of accessibility (at $p\le 0.10$, without correcting for multiple comparisons).
 Those zones are highlighted in the figure, where it can be seen that they form more or less compact neighborhoods.
-Remarkably, the largest (and significant) drops in accessibility are not downtown, but located in two cases in the industrial north of the city, in one case in an inner suburb above the escarpment, and lastly in a more suburban/exurban region in the south-west.
+Remarkably, the largest significant drops in accessibility are not downtown, but located in two cases in the industrial north of the city, in one case in an inner suburb above the escarpment, and lastly in a more suburban/exurban region in the south-west.
 
@@
 
-Just as the effects of the closures appear to have been uneven in space, they also seem to have differently impacted various population segments.
+Just as the effects of the closures appear to have been uneven in space, they also seem to have had different impacts on various population segments.
 Table \ref{tab:accessibility-by-age} shows the estimated number of people in each age group by their level of accessibility before and during the pandemic.
@@
 
-Food insecurity is a significant issue for many Canadian households and is associated reductions in nutrition and mental and physical health.
-While food bank use can provide some relief, levels of need in the City of Hamilton were on the rise before the onset of the COVID-19 pandemic with a troubling increase in the number of visits made by children.
-While temporary financial relief measures helped reduce food bank use in the summer of 2020, the pandemic has more than likely increased food insecurity for many households and created a surge in food bank demand in the city.
-On the other hand, the pandemic has also resulted in major disruptions to food bank and food service delivery, resulting in the closure or cancellation of many food services in Hamilton.
-On top of this, reductions in transit service have reduced travel options for many.
-
-In response, this research has sought to better understand accessibility to food banks and food service locations as well as how the closure of some locations over the pre- and during-COVID time periods affected the potential for low-income households to reach these amenities.
-Using the balanced floating catchment area method to account for population demand and congestion effects at service points, we estimate multi-modal accessibility to food bank and food service locations for low-income households.
-In the absence of information on how food insecure households travel to food banks and related services, we adopted a normative approach that examines accessibility to food banks reachable within 15 minutes based on global standards for accessible drinking water.
-The weighting of travel time estimates by the modal split in different zonal geographies tailors the results to patterns of travel behaviour captured in the TTS regional travel survey.
+Food insecurity is a significant issue for many Canadian households and is associated with reductions in nutrition and poor mental and physical health outcomes.
+While food bank use can provide some relief, the levels of need in the City of Hamilton were on the rise before the onset of the COVID-19 pandemic, with a troubling increase in the number of visits made by children.
+While temporary financial relief measures may have helped reduce food bank use in the summer of 2020, the pandemic has in all probability increased food insecurity for many households.
+To compound matters, the pandemic has also resulted in major disruptions, including to employment, mobility alternatives, and to food bank and food service delivery.
+In response, this research has sought to better understand accessibility to food banks and food service locations, as well as how the closure of some locations over the pre- and during-COVID time periods affected the potential for low-income households to reach these amenities.
+Using the balanced floating catchment area method to account for population demand and congestion effects at service points, we estimated multi-modal accessibility to food bank and food service locations for low-income households.
+The weighting of travel time estimates by the modal split in different zonal geographies tailors the results to patterns of travel behaviour captured in the regional travel survey.
 Moreover, our parcel-level analysis presents a disaggregate approach to estimating accessibility based on the locations of residential parcels and dwellings.
 
-Results show that while accessibility levels were lower in the city's more car-oriented suburban and rural areas to begin with, the closure of 14% of the city's food bank and low-cost or free food service locations during the pandemic resulted in an overall decrease in accessibility across the city.
+Our results show that while accessibility levels were lower in the city's more car-oriented suburban and rural areas to begin with, the closure of 14% of the city's food bank and low-cost or free food service locations during the pandemic resulted in an overall decrease in accessibility across the city.
 However, these effects were not uniform over space or for different population groups.
-Because the balanced floating catchment area method takes into account changes in demand and congestion for service providers, the closure of food bank and food service locations can reverberate throughout the catchment areas of the city.
+Since the balanced floating catchment area method takes into account changes in demand and congestion for service providers, the closure of food bank and food service locations can reverberate throughout the catchment areas of the whole city.
 For some suburban zones, the closure of a relatively high level-of-service location results in the remaining services being spread over a larger population.
@@
 Reductions in accessibility in the city's more urban north end, where low-income households conduct higher proportions of trips by transit and walking, emphasize the importance of geographic proximity in the potential to reach food bank and service locations for these residents.
-Beyond geography, the results also highlight the differential impact of food bank closing during COVID-19 on population groups with seniors and children most affected.
-
-Still, the work is not without its limitations.
-We use data from the census to model low-income households as a proxy for households that are possibly experiencing food insecurity.
-However, the degree to which our low-income cut-off of CAD\$40,000 reflects food insecurity in the different zones of the study area is not known.
+Beyond geography, the results also highlight the differential impact of food banks closing during COVID-19 on population groups, with seniors and children most affected.
+
+In the absence of information regarding how food insecure households travel to food banks and related services, we examined accessibility to food banks using a 15 minutes (one way) travel time threshold.
+The fact that we must rely on a standard created for accessible drinking water in the developing world only serves to highlight the tragedy of food insecurity in an affluent country like Canada.
+More broadly, it points to the absurd need to understand how a bad situation was made worse by the pandemic: in effect, the analysis reveals that disparities in accessibility to food banks and services predated the pandemic, that the pandemic contributed to the deterioration of these services, and that populations already in distress, particularly children, ended up in an even more adverse state.
+How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding what is an _acceptable_ or _sufficient_ level of service when it comes to food banks.
+Whether this is one facility per 1,000 population or three facilities per 1,000 population, or any other number, we know not.
+
+In summary, while previous work has noted the important role of geography alongside other socio-economic and demographic indicators in household access to healthy food, this is, to the best of our knowledge, the first study that has focused on the geographic component of food bank accessibility.
+Beyond accounting for the inflation of demand that occurs in traditional floating catchment area methods, our application of the balanced floating catchment area approach offers a novel analysis of food bank and food service accessibility that is sensitive to the locations of low-income households, information on their age distributions and typical trip-making behaviour, food bank and food service locations and their operations over time, and the characteristics of the city's multi-modal transportation network.
+Despite the limitations noted above, the findings from this analysis can provide quantitative evidence to help urban planners and community administrators strategically allocate food bank services considering accessibility for different population groups, especially when considering the unequal distribution of impacts from the COVID-19 pandemic.
+
+In terms of future research, it is important to note that the degree to which our low-income cut-off of CAD\$40,000 reflects food insecurity in the different zones of the study area is not known.
 We also consider all food banks and food service locations equally.
-Future work could improve this method by accounting for differential capacities at food provider locations, such as the number of meals served.
-Unforunately, this data was not available in the present case as some service locations had already closed.
-We also did not explicitly account for any changes in demand over the study period.
-If trends in Hamilton are similar to those in Toronto, there could be a significant increase in demand that would alter the estimated accessibilities.
+Information on differential capacities at food provider locations is currently not collected, and given the closure of facilities was not possible to obtain.
+Data on services, such as the number of meals served, could be used to refine the analysis in the future.
 Moreover, while the travel survey allows us to model multi-modal accessibilities that align with travel behaviour observed in the travel survey and capture differences in accessibility by age categories, the use of travel survey data for modelling food insecurity also has its limitations.
-Research into the population weighting methods used in the TTS note that the survey may under-count the lowest- and highest-income households in the survey study region, although the magnitude of this under-counting is unknown as approximately 20% of respondents to the 2016 survey did not report their income [@rose2018transportation].
-In that regard, the modal splits of low-income households observed through the TTS data may not accurately reflect the travel behaviours of food insecure populations.
-
-Nevertheless, while previous work has noted the important role of geography alongside other socioeconomic and demographic indicators in household access to healthy food, this is, to the best of our knowledge, the first study that has focused on the geographic component of food bank accessibility.
-Beyond accounting for the inflation of demand that occurs in traditional floating catchment area methods, our application of the balanced floating catchment area approach offers a novel analysis of food bank and food service accessibility that is sensitive to the locations of low-income households, information on their age distributions and typical tripmaking behaviour, food bank and food service locations and their operations over time, and the characteristics of the city's multi-modal transportation network.
-Findings from this analysis can provide quantitative evidence to help urban planners and community administrators strategically allocate food bank services considering accessibility for different population groups, especially when considering the unequal distribution of impacts from the COVID-19 pandemic.
+Research into the population weighting methods used in the TTS note that the survey may under-count the lowest- and highest-income households in the survey study region, although the magnitude of this under-counting is unknown, and approximately 20% of respondents to the 2016 survey did not report their income [@rose2018transportation].
+In that regard, the modal splits of low-income households observed through the TTS data may not accurately reflect the travel behaviours of food insecure populations, and our estimates might be somewhat conservative if those households who do not report income rely more on walking and/or transit for their mobility needs.
 

### 2021-04-28 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Overall proofreading and editing of conclusions

@@
   City of Hamilton, Ontario. Food banks and services are the last line
-  of support for households facing food insecurity; as such, they have
-  become essential and their relevance cannot be ignored in the midst of
-  the economic upheaval caused by the pandemic. Our analysis is based on
-  the application of balanced floating catchment areas, and concentrates
-  on households with lower incomes (\$\textless CAD40,000, approximately
-  the Low Income Cutoff Value for a city of Hamilton's size).
-  Accessibility was low to begin with in suburban and exurban parts of
-  the city. We find that about 14\% of locations originally available in
-  Hamilton closed during the pandemic; the impact of those closures on
-  the level of service and accessibility was disproportionate, with
+  of support for households facing food insecurity; as such, their
+  relevance cannot be ignored in the midst of the economic upheaval
+  caused by the pandemic. Our analysis is based on the application of
+  balanced floating catchment areas, and concentrates on households with
+  lower incomes (\textless CAD40,000, approximately the Low Income
+  Cutoff Value for a city of Hamilton's size). We find that
+  accessibility was low to begin with in suburban and exurban parts of
+  the city; furthermore, about 14\% of locations originally available in
+  Hamilton closed during the pandemic, further reducing accessibility.
+  The impact of closures on the level of service of the remaining
+  facilities, and on accessibility, was disproportionate, with
   system-wide losses exceeding 40\%. Those losses were geographically
@@
 (Bhattacharya et al., 2004; Kirkpatrick and Tarasuk, 2008; Olson, 1999)
-and physical and mental health in children and adults (Elgar et al.,
-2021; Jones, 2017; Ramsey et al., 2011; Seligman et al., 2010; Stuff et
-al., 2004). Over at least the past four decades food banks and related
-services have become an essential line of defense against food
-insecurity in Canadian communities (Black and Seto, 2020; Holmes et al.,
-2018; Riches, 2002; Tarasuk et al., 2020). In this respect, Canada is
-not unlike numerous other wealthy countries where a systematic
-dismantling of the welfare state took place in the intervening period
-(Tarasuk et al., 2014).
+and negative physical and mental health impact in children and adults
+(Elgar et al., 2021; Jones, 2017; Ramsey et al., 2011; Seligman et al.,
+2010; Stuff et al., 2004). Over at least the past four decades food
+banks and related services have become an essential line of defense
+against food insecurity in Canadian communities (Black and Seto, 2020;
+Holmes et al., 2018; Riches, 2002; Tarasuk et al., 2020). In this
+respect, Canada is not unlike numerous other wealthy countries where a
+systematic dismantling of the welfare state took place in the
+intervening period (Tarasuk et al., 2014).
 
 The emergence of COVID-19, the worst public health crisis since the 1918
-flu pandemic, has revealed important social and economic fault lines,
-and pre-existing patterns of inequality appear to have been exacerbated.
+flu pandemic, has exposed important social and economic fault lines, and
+pre-existing patterns of inequality appear to have been exacerbated.
 Along several other dimensions of stress (e.g., accessibility to health
-care facilities, Pereira et al., 2021a), this seems to be the case for
-food insecurity as well (Laborde et al., 2020). In the US, for example,
-it has been estimated that there was an increase of more than 30\% in
-household food insecurity, and more than one third of households were
-discovered to be newly food insecure - meaning they did not experience
-food insecurity before the pandemic (Niles et al., 2020). In Canada, Men
-and Tarasuk (2021) report that about 25\% of individuals who experienced
-job insecurity (a relatively common occurrence during the pandemic),
-also experienced food insecurity. Similarly, according to Statistics
-Canada (2020a), in the early stages of the pandemic almost 15\% of
-individuals reported living in a household that faced food insecurity;
-the risk of food insecurity was substantially higher for households with
-children. The difference between households with and without children
-was significant, and 11.7\% of households with children indicated that
-``food didn't last and {[}there was{]} no money to get more'' sometimes
-or often, compared to 7.3\% of households without children; likewise,
-13\% of households with children indicated that they ``{[}c{]}ouldn't
-afford balanced meals'' sometimes or often, compared to 8.8\% of
-households without children.
+care facilities, Ghorbanzadeh et al., 2021; Pereira et al., 2021a), this
+seems to be the case for food insecurity as well (Laborde et al., 2020).
+In the US, for example, it has been estimated that there was an increase
+of more than 30\% in household food insecurity, and more than one third
+of households were discovered to be newly food insecure - meaning they
+did not experience food insecurity before the pandemic (Niles et al.,
+2020). In Canada, Men and Tarasuk (2021) report that about 25\% of
+individuals who experienced job insecurity (a relatively common
+occurrence during the pandemic), also experienced food insecurity.
+Similarly, according to Statistics Canada (2020a), in the early stages
+of the pandemic almost 15\% of individuals reported living in a
+household that faced food insecurity; the risk of food insecurity was
+substantially higher for households with children. The difference
+between households with and without children was significant, and 11.7\%
+of households with children indicated that ``food didn't last and
+{[}there was{]} no money to get more'' sometimes or often, compared to
+7.3\% of households without children; likewise, 13\% of households with
+children indicated that they ``{[}c{]}ouldn't afford balanced meals''
+sometimes or often, compared to 8.8\% of households without children.
 
@@
 to provide a resource of last instance to households in precarious
-situations (Bazerghi et al., 2016). As recently as 2019, the Hamilton
-Hunger Report\footnote{https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf}
-noted that food banks in Hamilton, Ontario, recorded the highest number
-of visitors in the past 29 years; the number of children visiting food
-banks (minors up to 18 years old) was 9,125 in March 2019, up from 8,278
-the year before, a rate of increase greater than population growth.
+situations (Bazerghi et al., 2016). As a mid-size city grappling with
+deindustrialization, Hamilton exhibits high rates of poverty and food
+bank use. As recently as 2019, the Hamilton Hunger Report\footnote{https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf}
+noted that food banks in the city recorded the highest number of
+visitors in the past 29 years; a rate of increase greater than
+population growth. Most troubling, approximately 40\% of all visitors
+were children.
 
@@
 financial support programs from the federal government. Nevertheless,
-some of these benefit programs were temporary, which suggests that many
-households may again turn to food bank services to meet their needs.
+some of these benefit programs were temporary. Although more recent
+statistics on food bank use in Hamilton in 2020 and 2021 are not yet
+available, data from the Daily Bread Food Bank in neighbouring Toronto
+for August 2020 shows visits climbing 51\% year-over-year in that
+city\footnote{https://www.dailybread.ca/wp-content/uploads/2020/11/Whos-Hungry-Report-2020.pdf},
+which suggests that many households in Hamilton are likely to turn to
+food bank services to meet their needs.
 
@@
 
-\textbf{Need to provide some context here: the geography which makes
-transportation challenging, i.e., due to the escarpment; the main parts
-of the city: city core, inner suburbs, industrial north, suburban,
-exurban; and the industrial background of the city maybe, with poverty
-rates compared to other Canadian cities?}
+With a population of around 540,000, the City of Hamilton is the fourth
+largest city in Ontario. It has historically been home to major
+manufacturing industries but de-industrialization that has occurred over
+the past several decades has led Hamilton to become one of the most
+highly divided cities in Ontario, with a significant proportion of its
+residents living at or below Canada's poverty level (DeLuca et al.,
+2012; Jakar and Dunn, 2019; Latham and Moffat, 2007). In 2016, the
+Hamilton Community Foundation reported that based on the Low-Income
+Cut-Off, Hamilton recorded a poverty rate of 16.7\%, which was well
+above the average rate of Ontario (13.7\%) and the average national rate
+(12.8\%).\footnote{https://www.hamiltoncommunityfoundation.ca/vital-signs/low-income-2018/}.
+According to data from Hamilton Food Share\footnote{https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf},
+approximately 23,000 individuals accessed food banks in the city in
+March of 2019. Within this total is 9,125 visits by children (minors up
+to 18 years old), up from 8,278 the year before. Feed Ontario, the
+province's largest collective of hunger-relief organizations, found that
+on a per-capita basis, the level of need in the inner core of central
+Hamilton was second highest in Ontario\footnote{https://feedontario.ca/research/report-hunger-map/}.
+
+Geographically, the ``old'' city of Hamilton was amalgamated with
+several of its surrounding municipalities in 2001, with the city now
+featuring a mix of urban, suburban, exurban, and rural areas. Lower-cost
+housing proximate to the city's industrial north end has traditionally
+attracted immigrants and less-affluent residents compared to the city's
+wealthier suburbs. However, the decentralization of population from the
+inner core has led to challenges in transit connectivity to amenities
+and services and the proportion of auto users compared to transit users
+remains very high (Behan et al., 2008; Topalovic et al., 2012). In
+addition, the city is separated geographically by the Niagara
+Escarpment. With sections of rocky cliff that approach 100m in height,
+the escarpment presents a significant challenge for promoting active
+travel and transport connections between ``mountain'' and ``lower city''
+neighbourhoods. Taken together, the high level of food need, population
+locations, and transportation network characteristics combine to inform
+spatial accessibility to food banks and food services in the city.
 
@@
 such as food box, community kitchens, student nutrition programs, and
-shopping and transportation. In addition, two free meal services held on
-different days at the same location were collapsed into a single service
-point for the accessibility analysis. Additional details on the
-operations of individual facilities is not publicly available, and with
-the changes in operations it proved unfeasible to collect it. For this
-reason, the analysis to follow is of accessibility to the location of
-food banks and services, but not to specific services (e.g., breakfasts
-vs.~food boxes).
+shopping and transportation. With some providers offering different food
+services at the same location (e.g.~food bank with free and community
+meal services), and some of these services closing after the onset of
+the COVID-19 pandemic, we opted to geocode based on the service type. On
+the other hand, two free meal services held on different days at the
+same location were collapsed into a single service point for the
+accessibility analysis. Additional details on the operations of
+individual facilities is not publicly available, and with the changes in
+operations it proved unfeasible to collect it. For this reason, the
+analysis to follow is of accessibility to the location of food banks and
+services, but not to specific services (e.g., breakfasts vs.~food
+boxes).
 
@@
 people are expected to travel for food banks, but we note that in
-developing countries, access to drinking water is defined as a source
-that takes less than 30 minutes to reach (round trip, see UNICEF-WHO,
+developing countries, accessible sources of drinking water are those
+that can be reached in less than 30 minutes (round trip, see UNICEF-WHO,
 2019). There is no reason why people in affluent countries should be
-expected to travel more for a basic need such as food. Accordingly, we
-adopt a 15-minute threshold for the analysis (representing a one-way
-trip). This threshold is also approximately where the rate of change in
-accessibility slows down.
+expected to spend more time travelling for a basic necessity such as
+food. Accordingly, we adopt a 15-minute threshold for the analysis
+(representing a one-way trip). This threshold is also approximately
+where the rate of change in accessibility slows down.
 
@@
 the pattern of change in accessibility. Twenty-four TAZ are flagged as
-having significant large losses of accessibility (at \(p<=0.10\),
+having significantly large losses of accessibility (at \(p\le 0.10\),
 without correcting for multiple comparisons). Those zones are
 highlighted in the figure, where it can be seen that they form more or
-less compact neighborhoods. Remarkably, the largest (and significant)
-drops in accessibility are not downtown, but located in two cases in the
+less compact neighborhoods. Remarkably, the largest significant drops in
+accessibility are not downtown, but located in two cases in the
 industrial north of the city, in one case in an inner suburb above the
@@
 Just as the effects of the closures appear to have been uneven in space,
-they also seem to have differently impacted various population segments.
-Table \ref{tab:accessibility-by-age} shows the estimated number of
-people in each age group by their level of accessibility before and
-during the pandemic. Here, it is important to note that the quartiles
-are relative: people in the top 25\% of accessibility still have lower
-accessibility during the pandemic than before. In reality, every
-population group is worse off during the pandemic in terms of their
-accessibility to food banks and services in the City of Hamilton.
+they also seem to have had different impacts on various population
+segments. Table \ref{tab:accessibility-by-age} shows the estimated
+number of people in each age group by their level of accessibility
+before and during the pandemic. Here, it is important to note that the
+quartiles are relative: people in the top 25\% of accessibility still
+have lower accessibility during the pandemic than before. In reality,
+every population group is worse off during the pandemic in terms of
+their accessibility to food banks and services in the City of Hamilton.
 However, some age groups were affected more. As seen in the table, the
@@
 
-Words go here.
-
-We use data from the census to model low-income households as a proxy
-for households that are possibly experiencing food insecurity. However,
-the degree to which our low-income cut-off of CAD\$40,000 reflects food
-insecurity in the different zones of the study area is not known.
-
-While the travel survey allows us to model multi-modal accessibilities
-that align with travel behaviour observed in the travel survey and
-capture differences in accessibility by age categories, the use of
-travel survey data for modelling food insecurity also has its
-limitations. Research into the population weighting methods used in the
-TTS note that the survey may under-count the lowest- and highest-income
-households in the survey study region, although the magnitude of this
-under-counting is unknown as approximately 20\% of respondents to the
-2016 survey did not report their income (Rose, 2018). In that regard,
-the modal splits of low-income households observed through the TTS data
-may not accurately reflect the travel behaviours of food insecure
-populations.
-
-In the absence of information on how food insecure households travel to
-food banks and related services, we adopted a normative approach that
-examines accessibility to food banks reachable within 15 minutes based
-on standards for accessible drinking water\ldots{}
+Food insecurity is a significant issue for many Canadian households and
+is associated with reductions in nutrition and poor mental and physical
+health outcomes. While food bank use can provide some relief, the levels
+of need in the City of Hamilton were on the rise before the onset of the
+COVID-19 pandemic, with a troubling increase in the number of visits
+made by children. While temporary financial relief measures may have
+helped reduce food bank use in the summer of 2020, the pandemic has in
+all probability increased food insecurity for many households. To
+compound matters, the pandemic has also resulted in major disruptions,
+including to employment, mobility alternatives, and to food bank and
+food service delivery. In response, this research has sought to better
+understand accessibility to food banks and food service locations, as
+well as how the closure of some locations over the pre- and during-COVID
+time periods affected the potential for low-income households to reach
+these amenities. Using the balanced floating catchment area method to
+account for population demand and congestion effects at service points,
+we estimated multi-modal accessibility to food bank and food service
+locations for low-income households. The weighting of travel time
+estimates by the modal split in different zonal geographies tailors the
+results to patterns of travel behaviour captured in the regional travel
+survey. Moreover, our parcel-level analysis presents a disaggregate
+approach to estimating accessibility based on the locations of
+residential parcels and dwellings.
+
+Our results show that while accessibility levels were lower in the
+city's more car-oriented suburban and rural areas to begin with, the
+closure of 14\% of the city's food bank and low-cost or free food
+service locations during the pandemic resulted in an overall decrease in
+accessibility across the city. However, these effects were not uniform
+over space or for different population groups. Since the balanced
+floating catchment area method takes into account changes in demand and
+congestion for service providers, the closure of food bank and food
+service locations can reverberate throughout the catchment areas of the
+whole city. For some suburban zones, the closure of a relatively high
+level-of-service location results in the remaining services being spread
+over a larger population. In others, high auto dependence for trips
+leads to decreases in accessibility that accumulate due to the loss of
+several food bank and service locations initially reachable within 15
+minutes by car. Reductions in accessibility in the city's more urban
+north end, where low-income households conduct higher proportions of
+trips by transit and walking, emphasize the importance of geographic
+proximity in the potential to reach food bank and service locations for
+these residents. Beyond geography, the results also highlight the
+differential impact of food banks closing during COVID-19 on population
+groups, with seniors and children most affected.
+
+In the absence of information regarding how food insecure households
+travel to food banks and related services, we examined accessibility to
+food banks using a 15 minutes (one way) travel time threshold. The fact
+that we must rely on a standard created for accessible drinking water in
+the developing world only serves to highlight the tragedy of food
+insecurity in an affluent country like Canada. More broadly, it points
+to the absurd need to understand how a bad situation was made worse by
+the pandemic: in effect, the analysis reveals that disparities in
+accessibility to food banks and services predated the pandemic, that the
+pandemic contributed to the deterioration of these services, and that
+populations already in distress, particularly children, ended up in an
+even more adverse state. How much worse, it is impossible to say, mainly
+because there is also a dearth of information, let alone standards,
+regarding what is an \emph{acceptable} or \emph{sufficient} level of
+service when it comes to food banks. Whether this is one facility per
+1,000 population or three facilities per 1,000 population, or any other
+number, we know not.
+
+In summary, while previous work has noted the important role of
+geography alongside other socio-economic and demographic indicators in
+household access to healthy food, this is, to the best of our knowledge,
+the first study that has focused on the geographic component of food
+bank accessibility. Beyond accounting for the inflation of demand that
+occurs in traditional floating catchment area methods, our application
+of the balanced floating catchment area approach offers a novel analysis
+of food bank and food service accessibility that is sensitive to the
+locations of low-income households, information on their age
+distributions and typical trip-making behaviour, food bank and food
+service locations and their operations over time, and the
+characteristics of the city's multi-modal transportation network.
+Despite the limitations noted above, the findings from this analysis can
+provide quantitative evidence to help urban planners and community
+administrators strategically allocate food bank services considering
+accessibility for different population groups, especially when
+considering the unequal distribution of impacts from the COVID-19
+pandemic.
+
+In terms of future research, it is important to note that the degree to
+which our low-income cut-off of CAD\$40,000 reflects food insecurity in
+the different zones of the study area is not known. We also consider all
+food banks and food service locations equally. Information on
+differential capacities at food provider locations is currently not
+collected, and given the closure of facilities was not possible to
+obtain. Data on services, such as the number of meals served, could be
+used to refine the analysis in the future. Moreover, while the travel
+survey allows us to model multi-modal accessibilities that align with
+travel behaviour observed in the travel survey and capture differences
+in accessibility by age categories, the use of travel survey data for
+modelling food insecurity also has its limitations. Research into the
+population weighting methods used in the TTS note that the survey may
+under-count the lowest- and highest-income households in the survey
+study region, although the magnitude of this under-counting is unknown,
+and approximately 20\% of respondents to the 2016 survey did not report
+their income (Rose, 2018). In that regard, the modal splits of
+low-income households observed through the TTS data may not accurately
+reflect the travel behaviours of food insecure populations, and our
+estimates might be somewhat conservative if those households who do not
+report income rely more on walking and/or transit for their mobility
+needs.
 
@@
 health 41, 732--740.
+
+\leavevmode\hypertarget{ref-behan2008smart}{}%
+Behan, K., Maoh, H., Kanaroglou, P., 2008. Smart growth strategies,
+transportation and urban sprawl: Simulated futures for hamilton,
+ontario. The Canadian Geographer/Le G{é}ographe canadien 52, 291--308.
+doi:\url{https://doi.org/10.1111/j.1541-0064.2008.00214.x}
 
@@
 (M2SFCA) metric. Health \& Place 24, 30--43.
-doi:\href{https://doi.org/10.1016/j.healthplace.2013.07.012}{10.1016/j.healthplace.2013.07.012}
+doi:\url{https://doi.org/10.1016/j.healthplace.2013.07.012}
+
+\leavevmode\hypertarget{ref-deluca2012code}{}%
+DeLuca, P.F., Buist, S., Johnston, N., 2012. The code red project:
+Engaging communities in health system change in hamilton, canada. Social
+Indicators Research 108, 317--327.
+doi:\url{https://doi.org/10.1007/s11205-012-0068-y}
 
@@
 
+\leavevmode\hypertarget{ref-ghorbanzadeh2021spatial}{}%
+Ghorbanzadeh, M., Kim, K., Erman Ozguven, E., Horner, M.W., 2021.
+Spatial accessibility assessment of COVID-19 patients to healthcare
+facilities: A case study of florida. Travel Behaviour and Society 24,
+95--101. doi:\url{https://doi.org/10.1016/j.tbs.2021.03.004}
+
 \leavevmode\hypertarget{ref-gundersen2018food}{}%
@@
 
+\leavevmode\hypertarget{ref-jakar2019turning}{}%
+Jakar, G.S., Dunn, J.R., 2019. (Turning rust into gold?) Hamilton,
+ontario and a canadian perspective of shrinking and declining cities.
+Cities 94, 1--10. doi:\url{https://doi.org/10.1016/j.cities.2019.05.016}
+
 \leavevmode\hypertarget{ref-jones2017food}{}%
@@
 Research Institute (IFPRI), Washington, DC.
+
+\leavevmode\hypertarget{ref-latham2007determinants}{}%
+Latham, J., Moffat, T., 2007. Determinants of variation in food cost and
+availability in two socioeconomically contrasting neighbourhoods of
+hamilton, ontario, canada. Health \& Place 13, 273--287.
+doi:\url{https://doi.org/10.1016/j.healthplace.2006.01.006}
 
@@
 
+\leavevmode\hypertarget{ref-topalovic2012light}{}%
+Topalovic, P., Carter, J., Topalovic, M., Krantzberg, G., 2012. Light
+rail transit in hamilton: Health, environmental and economic impact
+analysis. Social Indicators Research 108, 329--350.
+doi:\url{https://doi.org/10.1007/s11205-012-0069-x}
+
 \leavevmode\hypertarget{ref-world2019progress}{}%

### 2021-04-28 README.Rmd
Created README.Rmd file

--- 2021-04-28 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
@@
 ---
-title: "Changes in accessibility to food banks and food services during COVID-19 and implications for low income populations in Hamilton, Ontario"
-author:
-  - name: Author One
-    email: author1@example.com
-    affiliation: Some University
-  - name: Author Two
-    email: author2@example.com
-    affiliation: Another University
-    footnote: 1
-  - name: Author Three
-    email: author3@example.com
-    affiliation: Some University
-  - name: Author Four
-    email: author4@example.com
-    affiliation: Some University
-address:
-  - code: Some University
-    address: Department, Street, City, State, Zip
-  - code: Another University
-    address: Department, Street, City, State, Zip
-footnote:
-  - code: 1
-    text: "Corresponding Author"
-abstract: |
+output:
+  github_document:
+    pandoc_args: --webtex
+always_allow_html: true
+bibliography: Accessibility-Foodbanks-Hamilton/references.bib
+---
+
+<!-- README.md is generated from README.Rmd. Please edit that file -->
+
+```{r, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>"
+)
+```
+
+# Changes in accessibility to food banks and food services during COVID-19 and implications for low income populations in Hamilton, Ontario
+
+<!-- badges: start -->
+<!-- badges: end -->
+
+Christopher D. Higgins (University of Toronto Scarborough)  
+Antonio Páez (McMaster University)  
+Gyoorie Kim (University of Toronto Mississauga)  
+Jue Wang (University of Toronto Mississauga)
+
+Paper submitted to *Social Science and Medicine*
+
+## Abstract
+
 In this paper we analyze the changes in accessibility to food banks and related services before and during the COVID-19 pandemic in the City of Hamilton, Ontario.
@@
 
-journal: "Social Science and Medicine"
-date: "`r Sys.Date()`"
-bibliography: references.bib
-#linenumbers: true
-#numbersections: true
-csl: elsevier-harvard.csl
-output: rticles::elsevier_article
-header-includes:
-    - \usepackage[table]{xcolor}
----
-
-# CRediT author statement
-
-**Author 1:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization, Supervision; **Author 2:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization; **Author 3:** Conceptualization, Investigation, Data Curation, Writing - Original Draft; **Author 4:** Resources, Writing - Original Draft, Supervision
+## Keywords
+
+- Food insecurity  
+- Food banks  
+- Accessibility  
+- COVID-19  
+- Hamilton  
+- Ontario
+- Reproducible research
 

### 2021-04-28 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Minor edits to finish initial draft

@@
 Food insecurity is defined as an "inadequate or uncertain access to a sufficient quantity and/or adequate quality of food" due to a household's financial limitations [@enns2020experiences].
-This condition has been associated with reductions in nutritional outcomes [@bhattacharya2004poverty; @kirkpatrick2008food; @olson1999nutrition] and negative physical and mental health impact in children and adults [@elgar2021relative; @jones2017food; @ramsey2011food; @seligman2010food; @stuff2004household].
+This condition has been associated with reductions in nutritional outcomes [@bhattacharya2004poverty; @kirkpatrick2008food; @olson1999nutrition] and negative physical and mental health impacts in children and adults [@elgar2021relative; @jones2017food; @ramsey2011food; @seligman2010food; @stuff2004household].
 Over at least the past four decades food banks and related services have become an essential line of defense against food insecurity in Canadian communities [@riches2002food; @holmes2018nothing; @black2020examining; @tarasuk2020relationship].
@@
 However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to food banks.
-Although previous work has explored differences in _accessing_ food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource [e.g., @enns2020experiences], we are not aware of any research that has focused on the geographic component of accessibility.
+Previous work has explored differences in _accessing_ food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource [e.g., @enns2020experiences].
+In addition, transportation and locational considerations have been raised as key issues in food bank accessibility in previous qualitative research [@smith2017food].
+However, we are not aware of any research that has focused on estimating or capturing this geographic component of accessibility.
 
@@
 [-@paez2019demand] to investigate the accessibility situation.
-For this, we use a fully disaggregated approach based on parcel-level data.
+For this, we adopt a fully disaggregated approach based on parcel-level data.
 Socio-economic and demographic data are drawn from the latest Census of Canada (2016), whereas travel information is from the most recent regional travel survey from 2016.
@@
 This nutrient deficiency is associated with major population health concerns, particularly among Canadians at socio-economic disadvantage [@bazerghi2016role].
-Data on food insecurity in Canada is collected in several ways.
-Quantitatively, official government surveys such as the Household Food Security Survey Module (HFSSM), the Canadian Community Health Surveys (CCHS), the International Study of Adults (LISA), and official classifications determined by Health Canada in relation to socio-demographic variables offer some insight into food insecurity [@kirkpatrick2008food; @tarasuk2009household; @gundersen2018food].
+Official government surveys such as the Household Food Security Survey Module (HFSSM), the Canadian Community Health Surveys (CCHS), the Longitudinal and International Study of Adults (LISA), and official classifications determined by Health Canada in relation to socio-demographic variables offer some insight into food insecurity in Canada [@kirkpatrick2008food; @tarasuk2009household; @gundersen2018food].
 Nationally, it has been found that food insecurity impacts approximately 12.3% of Canadian households [@tarasuk2014food]
@@
 Most recently in May 2020, Canada recorded 14.7% of its population living in food insecurity in the past 30 days [@statisticscanada2020food].
-Considering the negative mental and physical health effects associated with food insecurity, increases in food insecurity rates due to the pandemic, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
+Considering the negative mental and physical health effects, increases in food insecurity rates due to the onset of COVID-19, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
 Recent data^[https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis_Exec-Sum_EN.pdf] from FBC showed that 52% of member food banks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of the country.
@@
 
-Geographically, the "old" city of Hamilton was amalgamated with several of its surrounding municipalities in 2001, with the city now featuring a mix of urban, suburban, exurban, and rural areas.
+Geographically, the "old" City of Hamilton was amalgamated with several of its surrounding municipalities in 2001, with the city now featuring a mix of urban, suburban, exurban, and rural areas.
 Lower-cost housing proximate to the city's industrial north end has traditionally attracted immigrants and less-affluent residents compared to the city's wealthier suburbs.
@@
 
-Figure \ref{fig:sensitivity-analysis} shows the results of conducting sensitivity analysis of the system-wide accessibility as we vary the threshold (considering the situation before the pandemic).
+Figure \ref{fig:sensitivity-analysis} shows the results of conducting a sensitivity analysis of the system-wide accessibility as we vary the threshold (considering the situation before the pandemic).
 There is a clear pattern whereby more strict values of $\delta$ are associated with higher levels of system-wide accessibility: while increases in accessibility that result from decreases in the travel time window might seem counter-intuitive, this is a result of lower _congestion_, since fewer households are serviced and thus competition for the same resources is more limited.
@@
 
-The general pattern of the levels of service is similar before and during the pandemic, with lower values in the center of the city where low-income households exhibit multimodal trip patterns.
+The general pattern of the levels of service is similar before and during the pandemic, with lower values in the center of the city where low-income households exhibit multimodal trip patterns that favour proximate service locations.
 Three more peripheral facilities towards the south of the city also have lower levels of service, presumably because they are expected to service relatively large suburban/exurban populations generally reliant on automobiles for travel.
@@
 To better highlight these changes, we used a local indicator of spatial autocorrelation [@anselin1995local] to explore the pattern of change in accessibility.
-Twenty-four TAZ are flagged as having significantly large losses of accessibility (at $p\le 0.10$, without correcting for multiple comparisons).
+Twenty-four TAZs are flagged as having significantly large losses of accessibility (at $p\le 0.10$, without correcting for multiple comparisons).
 Those zones are highlighted in the figure, where it can be seen that they form more or less compact neighborhoods.
@@
 The zone with the greatest decrease in accessibility within this cluster (-0.0009) has a high rate of car trips and connects to the most food bank locations in total as well as those that stayed open or closed.
-The exception to this is the zone in the middle of this cluster, which has a much higher proportion of transit use `r round(modes %>% st_drop_geometry() %>% filter(TAZUID == 5116) %>% summarize(first(p_transit)) *100)`\%.
+The exception to this is the zone in the middle of this cluster, which has a much higher proportion of transit use (`r round(modes %>% st_drop_geometry() %>% filter(TAZUID == 5116) %>% summarize(first(p_transit)) *100)`\%).
 This modal split weighting results in fewer food bank and service locations accessible from the parcels within this zone which, by extension, limits their access to closed locations reachable within 15 minutes in our calculations.
@@
 In the cluster to the south-west, the decrease in accessibility is predominately driven by the closure of a high level-of-service Community Meals provider.
-However, like the more central suburban zones, low-income households within this cluster also between 90\% to 100\% auto-dependent in their tripmaking in the TTS.
+However, like the more central suburban zones, low-income households within this cluster are also between 90\% to 100\% auto-dependent in their tripmaking in the TTS.
 The story is similar for the zone located in the north-west that exhibits the greatest decrease in accessibility.
-Here, low-income households responding to the TTS conducted 100\% of their trips by car and, as a result, parcels within this zone have access to the second-highest number of food bank and service locations within 15 minutes.
+Here, low-income households responding to the TTS conducted 100\% of their trips by car and, as a result, dwellings within this zone have access to the second-highest number of food bank and service locations within 15 minutes.
 However, this also means zonal accessibilities are greatly affected by the number of closures throughout the city.
@@
 Just as the effects of the closures appear to have been uneven in space, they also seem to have had different impacts on various population segments.
-Table \ref{tab:accessibility-by-age} shows the estimated number of people in each age group by their level of accessibility before and during the pandemic.
+Using data on low income individuals by age drawn from the TTS, Table \ref{tab:accessibility-by-age} shows the estimated number of people in each age group by their level of accessibility before and during the pandemic.
 Here, it is important to note that the quartiles are relative: people in the top 25\% of accessibility still have lower accessibility during the pandemic than before.
@@
 Whether this is one facility per 1,000 population or three facilities per 1,000 population, or any other number, we know not.
+But like international human rights standards for accessibility to clean drinking water within 30 minutes (round-trip), this work strongly suggests that adopting better defined standards for access to food banks that move beyond generalized conceptualizations of "access" to consider the geographic component of accessibility (and its interactions with travel behaviour and individual/household socioeconomic and demographic traits) would make a meaningful contribution to understanding food access and "food bank deserts" for food insecure populations.
 

### 2021-04-28 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
V1 Version for submission to Social Science & Medicine

@@
 As well, children (age $\le 18$) appear to have been impacted the most.
-
 journal: "Social Science and Medicine"
@@
 ---
-
-# CRediT author statement
-
-**Author 1:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization, Supervision; **Author 2:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization; **Author 3:** Conceptualization, Investigation, Data Curation, Writing - Original Draft; **Author 4:** Resources, Writing - Original Draft, Supervision
 
@@
 ```
-
-\newpage
 
@@
 
-In the absence of information regarding how food insecure households travel to food banks and related services, we examined accessibility to food banks using a 15 minutes (one way) travel time threshold.
+In the absence of information regarding how food insecure households travel to food banks and related services, we examined accessibility to food banks using a 15 minutes (one-way) travel time threshold.
 The fact that we must rely on a standard created for accessible drinking water in the developing world only serves to highlight the tragedy of food insecurity in an affluent country like Canada.
@@
 Research into the population weighting methods used in the TTS note that the survey may under-count the lowest- and highest-income households in the survey study region, although the magnitude of this under-counting is unknown, and approximately 20% of respondents to the 2016 survey did not report their income [@rose2018transportation].
-In that regard, the modal splits of low-income households observed through the TTS data may not accurately reflect the travel behaviours of food insecure populations, and our estimates might be somewhat conservative if those households who do not report income rely more on walking and/or transit for their mobility needs.
+In that regard, the modal splits of low-income households observed through the TTS data may not accurately reflect the travel behaviours of food insecure populations, and our estimates might in fact be somewhat conservative if those households who do not report income rely more on walking and/or transit for their mobility needs.
 

### 2021-04-28 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
V1 Version for submission to Social Science & Medicine

@@
 
-\hypertarget{credit-author-statement}{%
-\section{CRediT author statement}\label{credit-author-statement}}
-
-\textbf{Author 1:} Conceptualization, Methodology, Software, Validation,
-Formal analysis, Investigation, Data Curation, Writing - Original Draft,
-Visualization, Supervision; \textbf{Author 2:} Conceptualization,
-Methodology, Software, Validation, Formal analysis, Investigation, Data
-Curation, Writing - Original Draft, Visualization; \textbf{Author 3:}
-Conceptualization, Investigation, Data Curation, Writing - Original
-Draft; \textbf{Author 4:} Resources, Writing - Original Draft,
-Supervision
-
-\newpage
-
 \hypertarget{introduction}{%
@@
 (Bhattacharya et al., 2004; Kirkpatrick and Tarasuk, 2008; Olson, 1999)
-and negative physical and mental health impact in children and adults
+and negative physical and mental health impacts in children and adults
 (Elgar et al., 2021; Jones, 2017; Ramsey et al., 2011; Seligman et al.,
@@
 accessibility to healthy food through the ``food desert'' concept, there
-has been little research into accessibility to food banks. Although
-previous work has explored differences in \emph{accessing} food banks,
-such as how some households utilize food banks over short periods of
-time while others regularly utilize food banks as longer-term resource
-(e.g., Enns et al., 2020), we are not aware of any research that has
-focused on the geographic component of accessibility.
+has been little research into accessibility to food banks. Previous work
+has explored differences in \emph{accessing} food banks, such as how
+some households utilize food banks over short periods of time while
+others regularly utilize food banks as longer-term resource (e.g., Enns
+et al., 2020). In addition, transportation and locational considerations
+have been raised as key issues in food bank accessibility in previous
+qualitative research (Smith-Carrier et al., 2017). However, we are not
+aware of any research that has focused on estimating or capturing this
+geographic component of accessibility.
 
@@
 use the balanced floating catchment area approach of Paez et al. (2019)
-to investigate the accessibility situation. For this, we use a fully
+to investigate the accessibility situation. For this, we adopt a fully
 disaggregated approach based on parcel-level data. Socio-economic and
@@
 concerns, particularly among Canadians at socio-economic disadvantage
-(Bazerghi et al., 2016). Data on food insecurity in Canada is collected
-in several ways. Quantitatively, official government surveys such as the
+(Bazerghi et al., 2016). Official government surveys such as the
 Household Food Security Survey Module (HFSSM), the Canadian Community
-Health Surveys (CCHS), the International Study of Adults (LISA), and
-official classifications determined by Health Canada in relation to
-socio-demographic variables offer some insight into food insecurity
-(Gundersen et al., 2018; Kirkpatrick and Tarasuk, 2008; Tarasuk and
-Vogt, 2009). Nationally, it has been found that food insecurity impacts
-approximately 12.3\% of Canadian households (Tarasuk et al., 2014)
+Health Surveys (CCHS), the Longitudinal and International Study of
+Adults (LISA), and official classifications determined by Health Canada
+in relation to socio-demographic variables offer some insight into food
+insecurity in Canada (Gundersen et al., 2018; Kirkpatrick and Tarasuk,
+2008; Tarasuk and Vogt, 2009). Nationally, it has been found that food
+insecurity impacts approximately 12.3\% of Canadian households (Tarasuk
+et al., 2014)
 
@@
 living in food insecurity in the past 30 days (Statistics Canada,
-2020a). Considering the negative mental and physical health effects
-associated with food insecurity, increases in food insecurity rates due
-to the pandemic, signal a change in the food environment with potential
-damages to population health during the course of the pandemic and
-beyond (Niles et al., 2020). Recent data\footnote{https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report\_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis\_Exec-Sum\_EN.pdf}
+2020a). Considering the negative mental and physical health effects,
+increases in food insecurity rates due to the onset of COVID-19, signal
+a change in the food environment with potential damages to population
+health during the course of the pandemic and beyond (Niles et al.,
+2020). Recent data\footnote{https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report\_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis\_Exec-Sum\_EN.pdf}
 from FBC showed that 52\% of member food banks reported an increase in
@@
 
-Geographically, the ``old'' city of Hamilton was amalgamated with
+Geographically, the ``old'' City of Hamilton was amalgamated with
 several of its surrounding municipalities in 2001, with the city now
@@
 
-Figure \ref{fig:sensitivity-analysis} shows the results of conducting
+Figure \ref{fig:sensitivity-analysis} shows the results of conducting a
 sensitivity analysis of the system-wide accessibility as we vary the
@@
 during the pandemic, with lower values in the center of the city where
-low-income households exhibit multimodal trip patterns. Three more
-peripheral facilities towards the south of the city also have lower
-levels of service, presumably because they are expected to service
-relatively large suburban/exurban populations generally reliant on
-automobiles for travel. During the pandemic, however, the levels of
-service dropped, in some cases quite substantially. The pattern of the
-losses in level of service, moreover, is not uniform. The upper pane of
-Figure \ref{fig:levels-of-service-changes} shows that the three
-peripheral facilities in the southern suburban/exurban part of the city
-had low levels of service to begin with, but did not see major declines
-during the pandemic. Further, the inset map shows that the levels of
-service deteriorated more in the central part of the city. However, the
-loss of level of service was not as large in the core (where most of the
-food banks/services are found), but instead was more marked in the inner
-ring around the core, where facilities may have faced greater demand
-from both central city and suburban populations after the closure of
-food banks during the pandemic.
+low-income households exhibit multimodal trip patterns that favour
+proximate service locations. Three more peripheral facilities towards
+the south of the city also have lower levels of service, presumably
+because they are expected to service relatively large suburban/exurban
+populations generally reliant on automobiles for travel. During the
+pandemic, however, the levels of service dropped, in some cases quite
+substantially. The pattern of the losses in level of service, moreover,
+is not uniform. The upper pane of Figure
+\ref{fig:levels-of-service-changes} shows that the three peripheral
+facilities in the southern suburban/exurban part of the city had low
+levels of service to begin with, but did not see major declines during
+the pandemic. Further, the inset map shows that the levels of service
+deteriorated more in the central part of the city. However, the loss of
+level of service was not as large in the core (where most of the food
+banks/services are found), but instead was more marked in the inner ring
+around the core, where facilities may have faced greater demand from
+both central city and suburban populations after the closure of food
+banks during the pandemic.
 
@@
 local indicator of spatial autocorrelation (Anselin, 1995) to explore
-the pattern of change in accessibility. Twenty-four TAZ are flagged as
+the pattern of change in accessibility. Twenty-four TAZs are flagged as
 having significantly large losses of accessibility (at \(p\le 0.10\),
@@
 zone in the middle of this cluster, which has a much higher proportion
-of transit use 32\%. This modal split weighting results in fewer food
+of transit use (32\%). This modal split weighting results in fewer food
 bank and service locations accessible from the parcels within this zone
@@
 Meals provider. However, like the more central suburban zones,
-low-income households within this cluster also between 90\% to 100\%
+low-income households within this cluster are also between 90\% to 100\%
 auto-dependent in their tripmaking in the TTS. The story is similar for
@@
 in accessibility. Here, low-income households responding to the TTS
-conducted 100\% of their trips by car and, as a result, parcels within
+conducted 100\% of their trips by car and, as a result, dwellings within
 this zone have access to the second-highest number of food bank and
@@
 they also seem to have had different impacts on various population
-segments. Table \ref{tab:accessibility-by-age} shows the estimated
-number of people in each age group by their level of accessibility
-before and during the pandemic. Here, it is important to note that the
-quartiles are relative: people in the top 25\% of accessibility still
-have lower accessibility during the pandemic than before. In reality,
-every population group is worse off during the pandemic in terms of
-their accessibility to food banks and services in the City of Hamilton.
+segments. Using data on low income individuals by age drawn from the
+TTS, Table \ref{tab:accessibility-by-age} shows the estimated number of
+people in each age group by their level of accessibility before and
+during the pandemic. Here, it is important to note that the quartiles
+are relative: people in the top 25\% of accessibility still have lower
+accessibility during the pandemic than before. In reality, every
+population group is worse off during the pandemic in terms of their
+accessibility to food banks and services in the City of Hamilton.
 However, some age groups were affected more. As seen in the table, the
@@
 travel to food banks and related services, we examined accessibility to
-food banks using a 15 minutes (one way) travel time threshold. The fact
+food banks using a 15 minutes (one-way) travel time threshold. The fact
 that we must rely on a standard created for accessible drinking water in
@@
 1,000 population or three facilities per 1,000 population, or any other
-number, we know not.
+number, we know not. But like international human rights standards for
+accessibility to clean drinking water within 30 minutes (round-trip),
+this work strongly suggests that adopting better defined standards for
+access to food banks that move beyond generalized conceptualizations of
+``access'' to consider the geographic component of accessibility (and
+its interactions with travel behaviour and individual/household
+socioeconomic and demographic traits) would make a meaningful
+contribution to understanding food access and ``food bank deserts'' for
+food insecure populations.
 
@@
 reflect the travel behaviours of food insecure populations, and our
-estimates might be somewhat conservative if those households who do not
-report income rely more on walking and/or transit for their mobility
-needs.
+estimates might in fact be somewhat conservative if those households who
+do not report income rely more on walking and/or transit for their
+mobility needs.
 

### 2021-04-28 README.Rmd
V1 Version for submission to Social Science & Medicine

@@
 ```{r table-food-bank-info, echo=FALSE, results='asis'}
-foodbank_table %>% kable("latex",
+foodbank_table %>% kable("html",
                          booktabs = TRUE,
@@
                          escape = FALSE) %>%
-  kable_styling(latex_options = c("striped", "scale_down")) %>%
+  kable_styling(bootstrap_options = c("striped")) %>%
   column_spec(2, width = "15em") %>%
@@
 accessibility_by_age %>%
-  kable("latex",
+  kable("html",
         booktabs = TRUE,
@@
         col.names = c("Accessibility Quartile",
-                      "Children (age $\\le$ 18)", 
+                      "Children (age <= 18)", 
                       "Adults (19-64)", 
-                      "Seniors (age $\\ge$ 65)", 
-                      "Children (age $\\le$ 18)", 
+                      "Seniors (age >= 65)", 
+                      "Children (age <= 18)", 
                       "Adults (19-64)", 
-                      "Seniors (age $\\ge$ 65)"),
+                      "Seniors (age >= 65)"),
         caption = "\\label{tab:accessibility-by-age}Population at each accessibility level by age group among members households with incomes less than CAD40,000.",
         escape = FALSE) %>%
-  kable_styling(latex_options = c("striped", "scale_down", "hold_position")) %>%
+  kable_styling(bootstrap_options = c("striped")) %>%
   add_header_above(c(" " = 1, "Pre-COVID-19" = 3, "During COVID-19" = 3)) %>%

### 2021-04-30 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-abstract.Rmd
V1 Revised abstract on the request of journal

@@
 ---
-
-# CRediT author statement
-
-**Author 1:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization, Supervision; **Author 2:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization; **Author 3:** Conceptualization, Investigation, Data Curation, Writing - Original Draft; **Author 4:** Resources, Writing - Original Draft, Supervision

### 2021-04-30 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-abstract.tex
V1 Revised abstract on the request of journal

@@
 
-\hypertarget{credit-author-statement}{%
-\section{CRediT author statement}\label{credit-author-statement}}
-
-\textbf{Author 1:} Conceptualization, Methodology, Software, Validation,
-Formal analysis, Investigation, Data Curation, Writing - Original Draft,
-Visualization, Supervision; \textbf{Author 2:} Conceptualization,
-Methodology, Software, Validation, Formal analysis, Investigation, Data
-Curation, Writing - Original Draft, Visualization; \textbf{Author 3:}
-Conceptualization, Investigation, Data Curation, Writing - Original
-Draft; \textbf{Author 4:} Resources, Writing - Original Draft,
-Supervision
-
 \end{document}

### 2021-08-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Revised the conclusions in response to reviewer comments

@@
 Food insecurity is a significant issue for many Canadian households and is associated with reductions in nutrition and poor mental and physical health outcomes.
-While food bank use can provide some relief, the levels of need in the City of Hamilton were on the rise before the onset of the COVID-19 pandemic, with a troubling increase in the number of visits made by children.
-While temporary financial relief measures may have helped reduce food bank use in the summer of 2020, the pandemic has in all probability increased food insecurity for many households.
-To compound matters, the pandemic has also resulted in major disruptions, including to employment, mobility alternatives, and to food bank and food service delivery.
+While emergency and community food services can provide some relief, the levels of need in the City of Hamilton were on the rise before the onset of the COVID-19 pandemic, with a troubling increase in the number of visits made by children to these services.
+While temporary financial relief measures may have helped in the summer of 2020, the pandemic has in all probability increased food insecurity for many households.
+To compound matters, the pandemic has also resulted in major disruptions, including to employment, mobility alternatives, and to emergency and community food services.
 In response, this research has sought to better understand accessibility to food banks and food service locations, as well as how the closure of some locations over the pre- and during-COVID time periods affected the potential for low-income households to reach these amenities.
-Using the balanced floating catchment area method to account for population demand and congestion effects at service points, we estimated multi-modal accessibility to food bank and food service locations for low-income households.
+Using the balanced floating catchment area method to account for population demand and congestion effects at service points, we estimated multi-modal accessibility to emergency and community food service locations for low-income households.
 The weighting of travel time estimates by the modal split in different zonal geographies tailors the results to patterns of travel behaviour captured in the regional travel survey.
@@
 
-Our results show that while accessibility levels were lower in the city's more car-oriented suburban and rural areas to begin with, the closure of 14% of the city's food bank and low-cost or free food service locations during the pandemic resulted in an overall decrease in accessibility across the city.
+Previous work has noted the important role of geography alongside other socio-economic and demographic indicators in household access to healthy food.
+The present papers is, to the best of our knowledge, among the first studies to focus on the geographic component of accessibility to emergency food services [@allen2021Changes].
+Beyond accounting for the inflation of demand that occurs in traditional floating catchment area methods, our application of the balanced floating catchment area approach offers a novel analysis of accessibility to emergency and communal food services that is sensitive to the locations of low-income households, information on their age distributions and typical trip-making behaviour, the locations of services, their operations over time, and the characteristics of the city's multi-modal transportation network.
+Our results show that while accessibility levels were lower in the city's more car-oriented suburban and rural areas to begin with, the closure of 14% of the city's emergency and community food service locations during the pandemic resulted in an overall decrease in accessibility across the city.
 However, these effects were not uniform over space or for different population groups.
-Since the balanced floating catchment area method takes into account changes in demand and congestion for service providers, the closure of food bank and food service locations can reverberate throughout the catchment areas of the whole city.
+Since the balanced floating catchment area method takes into account changes in demand and congestion for service providers, the closure of some services reverberates throughout the catchment areas of the whole city.
 For some suburban zones, the closure of a relatively high level-of-service location results in the remaining services being spread over a larger population.
-In others, high auto dependence for trips leads to decreases in accessibility that accumulate due to the loss of several food bank and service locations initially reachable within 15 minutes by car.
-Reductions in accessibility in the city's more urban north end, where low-income households conduct higher proportions of trips by transit and walking, emphasize the importance of geographic proximity in the potential to reach food bank and service locations for these residents.
-Beyond geography, the results also highlight the differential impact of food banks closing during COVID-19 on population groups, with seniors and children most affected.
+In others, high auto dependence for trips leads to decreases in accessibility that accumulate due to the loss of several locations initially reachable within 15 minutes by car.
+Reductions in accessibility in the city's more urban north end, where low-income households conduct higher proportions of trips by transit and walking, emphasize the importance of geographic proximity in the potential to reach service locations for these residents.
+Beyond geography, the results also highlight the differential impact of closures during COVID-19 on population groups, with seniors and children beign the two most impacted groups.
 
@@
 The fact that we must rely on a standard created for accessible drinking water in the developing world only serves to highlight the tragedy of food insecurity in an affluent country like Canada.
-More broadly, it points to the absurd need to understand how a bad situation was made worse by the pandemic: in effect, the analysis reveals that disparities in accessibility to food banks and services predated the pandemic, that the pandemic contributed to the deterioration of these services, and that populations already in distress, particularly children, ended up in an even more adverse state.
-How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding what is an _acceptable_ or _sufficient_ level of service when it comes to food banks.
-Whether this is one facility per 1,000 population or three facilities per 1,000 population, or any other number, we know not.
-But like international human rights standards for accessibility to clean drinking water within 30 minutes (round-trip), this work strongly suggests that adopting better defined standards for access to food banks that move beyond generalized conceptualizations of "access" to consider the geographic component of accessibility (and its interactions with travel behaviour and individual/household socioeconomic and demographic traits) would make a meaningful contribution to understanding food access and "food bank deserts" for food insecure populations.
-
-In summary, while previous work has noted the important role of geography alongside other socio-economic and demographic indicators in household access to healthy food, this is, to the best of our knowledge, the first study that has focused on the geographic component of food bank accessibility.
-Beyond accounting for the inflation of demand that occurs in traditional floating catchment area methods, our application of the balanced floating catchment area approach offers a novel analysis of food bank and food service accessibility that is sensitive to the locations of low-income households, information on their age distributions and typical trip-making behaviour, food bank and food service locations and their operations over time, and the characteristics of the city's multi-modal transportation network.
-Despite the limitations noted above, the findings from this analysis can provide quantitative evidence to help urban planners and community administrators strategically allocate food bank services considering accessibility for different population groups, especially when considering the unequal distribution of impacts from the COVID-19 pandemic.
+More broadly, it points to the absurd need to understand how a bad situation was made worse by the pandemic: in effect, the analysis reveals that disparities in the need for emergency and community food services predated the pandemic, that the pandemic contributed to the deterioration of these services, and that populations already in distress, particularly children, ended up in an even more adverse state.
+How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding _acceptable_ or _sufficient_ level of service when it comes to emergency food services.
+In fact, we would argue that the standard should be that no household faced food insecurity.
+As others have noted [e.g., @poppendieck1999sweet; @men2021food] the root of food insecurity is income poverty.
+There is likely a place for emergency food and community food services.
+In addition to providing food, these services could satisfy social needs by offering a social setting for seniors or by helping to connect households in need with longer term supports.
+From a food security perspective, on the other hand, these services should work only as a short term solution, and not as a semi-permanent feature of life for some of our fellow human beings.
+From a human rights perspective, long-term reliance on emergency food services should be as unacceptable in Canada as lack of clean drinking water within 30 minutes is elsewhere.
+Thus, while our analysis is valuable to map the suffering caused by food insecurity, from a policy perspective maintaining a robust social safety net that includes Employment Insurance and paid sick days are better tools to reduce this suffering than increasing the accessibility of emergency food services for food insecure populations.
 
 In terms of future research, it is important to note that the degree to which our low-income cut-off of CAD\$40,000 reflects food insecurity in the different zones of the study area is not known.
-We also consider all food banks and food service locations equally.
+We also consider all emergency and communal food service locations equally.
 Information on differential capacities at food provider locations is currently not collected, and given the closure of facilities was not possible to obtain.

### 2021-08-24 Accessibility-Foodbanks-Hamilton/Response-to-Reviewers-R1.Rmd
Started drafting response to reviewer

---
title: "Response to Reviewers"
author: ""
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
bibliography: references.bib
header-includes:
  - \usepackage{changepage}
---

We would like to express our gratitude to the editor and anonymous reviewer for their time and attention to our submission.

## Reviewer \#1:

Reviewer \#1: Overall, I found the paper well-written and clear. I did have some concerns, primarily with the paper's framing and the conclusions drawn:

\textcolor{blue}{Thank you for the opportunity to revise the paper, and for the valuable suggestions to do so.}

1.  The characterization of the emergency food services included is a little unclear. It seems until the table on page 10 that the focus is on food banks, but really it looks like the focus is on ALL emergency food provision (including communal meal services). The framing of the piece is therefore a little "off", as meal programs have a much longer history (pre-Victorian) than food banks, and are intended to support a somewhat different clientele (e.g., those without kitchen/cooking facilities, and usually "singles" rather than families). I think it would make sense for the authors to use the broader term "emergency food services" in the title and throughout, and to change the framing at the beginning (please note that I don't think it is necessary to go into the history of emergency food provision here at all really... but the current presentation beginning with food banks is misleading).

\textcolor{blue}{Words go here.}

2.  Relatedly, the conflation of food banks and meal programs is a bit of a challenge in terms of interpreting the data. For example, one would EXPECT congregate dining to essentially shut down during the pandemic, as this could be a route for disease spread. In some cases, some kind of food services may have continued to be offered (e.g., through takeout), but in general closures would be expected. In the food BANKS, however, the mode of delivery might need to be modified slightly, but direct impacts of covid would be less significant. Additionally, as mentioned above these services target different demographics. So, by conflating these two modes of food delivery, the utility of the analysis is compromised.

\textcolor{blue}{Words go here.}

3.  The presentation of data in relation to food security is a bit confusing, as data from pre-covid, during covid, and from both the U.S. and Canada is presented. This makes it a little challenging to draw out the contemporary situation in the study area, and should be presented more clearly (and given that they seem to have food security data from May 2020, and there is widely available data from 2018 as well, they SHOULD be able to give a clearer sense of how much food security has increased during the pandemic in Canada). Also, I think comparisons to the situation in the U.S. during covid should be made cautiously, as their social interventions have been very different than Canada's (and in both countries, responses have varied widely be state/province).

\textcolor{blue}{Words go here.}

4.  Overall, the paper seems longer than it needs to be in the introduction/context and conclusion sections - there is quite a bit of background that seems unnecessary, and unnecessary restatement of that material in the conclusion.

\textcolor{blue}{Words go here.}

5.  I disagree strongly with the author's conclusion that the "solution" is to create more uniform geographical coverage of emergency food programs. I think what this data shows more than anything is that - as is mentioned throughout the literature on food banks, starting with Janet Poppendieck's "Sweet Charity" (a key text not included here) and continued in the work of Val Tarasuk and others - food banks are an ineffective, inefficient (not to mention stigmatizing) solution to hunger at the best of times, and that during a crisis (when they are most needed) they are even LESS able to provide the services required. To me, the data here highlights strongly the need for an effective social safety net (including decent EI support that is rolled out quickly, and paid sick days, both of which have been partially improved but also subjects of contestation in Ontario/Canada during the pandemic) that would make food banks redundant. That said, there may still be a place for emergency food services as part of a "community service hub" model that helps people with their immediate needs while connecting them with longer-term supports (and having these give good geographic coverage would make sense). But overall I think this data proves that food banks are NOT effective in emergencies (at least of the pandemic type) and so we should be looking elsewhere for solutions rather than increasing our investments in this sector.

\textcolor{blue}{Thank you for these comments. We fundamentally agree with you, and admit that we could have been more forceful in our conclusions. We did a major rewrite of the conclusions in response to your critique, including the following:}

```{=tex}
\begin{adjustwidth}{1cm}{}
\textcolor{blue}{The fact that we must rely on a standard created for accessible drinking water in the developing world only serves to highlight the tragedy of food insecurity in an affluent country like Canada. More broadly, it points to the absurd need to understand how a bad situation was made worse by the pandemic: in effect, the analysis reveals that disparities in the need for emergency and community food services predated the pandemic, that the pandemic contributed to the deterioration of these services, and that populations already in distress, particularly children, ended up in an even more adverse state. How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding acceptable or sufficient level of service when it comes to emergency food services. In fact, we would argue that the standard should be that no household faced food insecurity. As others have noted [e.g., Poppendieck, 1999; Men and Tarasuk, 2021] the root of food insecurity is income poverty. There is likely a place for emergency food and community food services. In addition to providing food, these services could satisfy social needs by offering a social setting for seniors or by helping to connect households in need with longer term supports. From a food security perspective, on the other hand, these services should work only as a short term solution, and not as a semi-permanent feature of life for some of our fellow human beings. From a human rights perspective, long-term reliance on emergency food services should be as unacceptable in Canada as lack of clean drinking water within 30 minutes is elsewhere. Thus, while our analysis is valuable to map the suffering caused by food insecurity, from a policy perspective maintaining a robust social safety net that includes Employment Insurance and paid sick days are better tools to reduce this suffering than increasing the accessibility of emergency food services for food insecure populations.}
\end{adjustwidth}
```

Specific comments:

Pg 2 paragraph 2 - a little unclear which of the data presented is pre-covid and which is during covid

Pg 2 ln 46 (and also pg 21 ln 33) - were 40% of actual visitors children, or are they extrapolating from households with children? It seems unlikely that parents are actually bringing their children along, so "visitors" may be the wrong word ("users"?)

Pg 4 paragraph 1 - not sure there is much evidence that connects food insecurity (which is usually "mild" in relation to true hunger/severe food\
insecurity) in developed nations with nutrient deficiency? If so, this should be shared. If not, the argument should be clarified in relation to known outcomes (e.g., stress, obesity)

Pg 4 paragraph 2 - not sure that food banks include meal programs/prepared meals (the food BANK wouldn't, although the organization might offer both)

Pg 5 ln 27 - seems to imply that the U.S. experience of increasing food insecurity during the pandemic is true in Canada, but the U.S. context in terms of emergency income relief is very different, and the May 2020 statistic of 14.7% is not contextualised (is this an increase? From what?)

Pg 23 (last page) - would be nice to have a final concluding sentence that restates the contribution of the paper, rather than ending with limitations.

### 2021-08-27 00-Data-Processing/02-Routing-Transit.Rmd
push branch

--- 2021-04-19 00-Data-Processing/02-Routing-Transit.Rmd
@@
 Load packages used in the notebook:
-```{r message = FALSE}
+```{r include = FALSE}
 library(cancensus)
@@
 library(smoothr)
+library(tidytransit)
 library(tidyverse)
@@
 
-Download and import GTFS (Hamilton transit data)
+Download and import GTFS files
+
 ```{r eval=FALSE}
-download.file(url = "https://transitfeeds.com/p/hamilton-street-railway/31/latest/download", destfile = file.path(r5_path, "HSR_transit.zip"), mode = "wb")
-```
+# gtfs files
+download.file(url = "http://transitfeeds.com/p/hamilton-street-railway/31/20210122/download", 
+              destfile = file.path(r5_path, "HSR.zip"), 
+              mode = "wb")
 
-Download and import GTFS (Hamilton transit data)
-```{r eval=FALSE}
-download.file(url = "https://transitfeeds.com/p/burlington-transit/294/latest/download", destfile = file.path(r5_path, "Burlington_transit.zip"), mode = "wb")
+download.file(url = "http://transitfeeds.com/p/burlington-transit/294/20210315/download", 
+              destfile = file.path(r5_path, "BURT.zip"), 
+              mode = "wb")
 ```
@@
 Build the graph:
+
 ```{r build-graph, include = FALSE}
@@
 Calculate OD Matrix for transit:
+
 ```{r calculate transit od matrix}
-# set up batching according to how many origin rows to process at one time
-chunksize = 2000 # larger chunks for walking will give enough origins in each chunk to allow multiprocessing to spin up with R5R
-num_chunks = ceiling(nrow(origins_i)/chunksize)
-
-# create origin-destination pairs
-origins_chunks <- as.disk.frame(origins_i,
-                          outdir = "./df/origins_i",
-                          nchunks = num_chunks,
-                          overwrite = TRUE)
-
-start.time <- Sys.time()
-pb <- txtProgressBar(0, num_chunks, style = 3)
-
-for (i in 1:num_chunks){
-  origins_i_chunk <- get_chunk(origins_chunks, i)
-  ttm_chunk <- travel_time_matrix(r5_hamilton_cma,
-                          origins = origins_i_chunk,
+ttm_transit <- travel_time_matrix(r5_hamilton_cma,
+                          origins = origins_i,
                           destinations = destinations_j,
-                          mode = c("TRANSIT"),
+                          mode = c("WALK", "TRANSIT"),
                           mode_egress = c("WALK"),
-                          departure_datetime = as.POSIXct(strptime("2021-04-05 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
-                          max_walk_dist = 10000, # metres
-                          max_trip_duration = 180)
-
-  # export output as disk.frame
-  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
-                                            nchunks = 1,
-                                            outdir = "./df/output_ttm_transit",
-                                            compress = 50,
-                                            overwrite = TRUE),
-         add_chunk(output_df, ttm_chunk, chunk_id = i))
-  setTxtProgressBar(pb, i)
-}
-end.time <- Sys.time()
-ttm_transit_time <- end.time - start.time
-ttm_transit_time
+                          departure_datetime = as.POSIXct(strptime("2021-03-30 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
+                          max_walk_dist = 2000, # metres
+                          max_trip_duration = 300, 
+                          verbose = FALSE)
 ```
@@
 
-```{r load od matrix for transit, include=FALSE}
-# connect to the walking travel time matrix disk frame
-ttm_transit.disk.frame <- disk.frame("./df/output_ttm_transit")
-#opportunities_j <- as.data.frame(nyc_cb_point) %>% select(GEOID10, total_emp) %>% rename(toId = GEOID10, o_j = total_emp)
-```
-
-Convert disk.frame to data frame:
-```{r}
-ttm_transit <- as.data.frame(ttm_transit.disk.frame) %>%
-  transmute(UID = as.numeric(fromId), OBJECTID = as.numeric(toId), travel_time)
-```
-
-Make sure that the origin destination pairs are distinct in the travel time matrices:
 ```{r}
 ttm_transit <- ttm_transit %>%
-  distinct(UID, 
-           OBJECTID,
-           .keep_all = TRUE)
+  transmute(UID = as.numeric(fromId), OBJECTID = as.numeric(toId), travel_time) %>%
+  distinct(UID, OBJECTID, .keep_all = TRUE)
 ```

### 2021-08-27 00-Data-Processing/04-Data-Analysis.Rmd
push branch

@@
 load("output-data-files/ttm_transit.RData")
+load("output-data-files/ttm_transit_2020.RData")
+load("output-data-files/ttm_transit_2021.RData")
 load("output-data-files/ttm_walk.RData")
@@
 
-## Remove entries in travel time matrices for inapplicable destinations
-
-As a first step, we will remove the computed travel times to the three Community Kitchen locations (IDs 1-3):
-```{r}
-ttm_car <- ttm_car %>%
-  filter(!between(OBJECTID, left = 1, right = 3))
-
-ttm_transit <- ttm_transit %>%
-  filter(!between(OBJECTID, left = 1, right = 3))
-
-ttm_walk <- ttm_walk %>%
-  filter(!between(OBJECTID, left = 1, right = 3))
-```
-
-## Quick check
-
-Distribution of the number of foodbanks that can be reached by each mode at the maximum travel time used in the calculation of the time matrices (180 min or 10,000 m).
-
-Car:
-```{r}
-ttm_car %>% group_by(UID) %>% summarize(fb_per_parcel = n()) %>% summary()
-```
-
-All parcels can reach every foodbank within 180 min.
-
-Transit:
-```{r}
-ttm_transit %>% group_by(UID) %>% summarize(fb_per_parcel = n()) %>% summary()
-```
-
-Half of all parcels can reach 49 foodbanks or more in 180 min or less.
-
-Walking:
-```{r}
-ttm_walk %>% group_by(UID) %>% summarize(fb_per_parcel = n()) %>% summary()
-```
-
-Half of all parcels can reach 42 foodbanks or more in 180 min or less.
-
-## Prepare data
+# Prepare socioeconomic demographic data
 
@@
 
+# Travel Scenario: 2020
+## Remove entries in travel time matrices for inapplicable destinations
+
+As a first step, we will remove the computed travel times to the three Community Kitchen locations (IDs 1-3):
+```{r}
+ttm_car <- ttm_car %>%
+  filter(!between(OBJECTID, left = 1, right = 3))
+
+ttm_transit_2020 <- ttm_transit_2020 %>%
+  filter(!between(OBJECTID, left = 1, right = 3))
+
+ttm_transit_2021 <- ttm_transit_2021 %>%
+  filter(!between(OBJECTID, left = 1, right = 3))
+
+ttm_walk <- ttm_walk %>%
+  filter(!between(OBJECTID, left = 1, right = 3))
+```
+
+## Quick check
+
+Distribution of the number of foodbanks that can be reached by each mode at the maximum travel time used in the calculation of the time matrices (180 min or 10,000 m).
+
+Car:
+```{r}
+ttm_car %>% group_by(UID) %>% summarize(fb_per_parcel = n()) %>% summary()
+```
+
+All parcels can reach every foodbank within 180 min.
+
+Transit in 2020:
+```{r}
+ttm_transit_2020 %>% group_by(UID) %>% summarize(fb_per_parcel = n()) %>% summary()
+```
+
+Half of all parcels can reach 52 foodbanks or more in 180 min or less.
+
+Transit in 2021:
+
+```{r}
+ttm_transit_2021 %>% group_by(UID) %>% summarize(fb_per_parcel = n()) %>% summary()
+```
+
+Half of all parcels can reach 49 foodbanks or more in 180 min or less.
+
+Walking:
+```{r}
+ttm_walk %>% group_by(UID) %>% summarize(fb_per_parcel = n()) %>% summary()
+```
+
+Half of all parcels can reach 42 foodbanks or more in 180 min or less.
+
+## Join Socioeconomic and Demographic Data
+
 Join the household data to the travel time matrices:
@@
 
-ttm_transit <- ttm_transit %>%
+ttm_transit_2020 <- ttm_transit_2020 %>%
   left_join(dwelling_network_points_2016 %>%
@@
 
-ttm_walk <- ttm_walk %>%
+ttm_transit_2021 <- ttm_transit_2021 %>%
   left_join(dwelling_network_points_2016 %>%
@@
             by = c("UID" = "ID"))
+
+ttm_walk <- ttm_walk %>%
+  left_join(dwelling_network_points_2016 %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID, 
+                            income_less40k,
+                            children,
+                            adults,
+                            seniors,
+                            DAUID, 
+                            TAZUID),
+            by = c("UID" = "ID"))
 ```
@@
 Join the food bank data to the travel time matrices:
+
 ```{r}
@@
 
-ttm_transit <- ttm_transit %>%
+ttm_transit_2020 <- ttm_transit_2020 %>%
   left_join(foodbank_locations %>%
@@
 
-ttm_walk <- ttm_walk %>%
+ttm_transit_2021 <- ttm_transit_2021 %>%
   left_join(foodbank_locations %>%
@@
             by = c("OBJECTID" = "ID"))
+
+ttm_walk <- ttm_walk %>%
+  left_join(foodbank_locations %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID,
+                            TYPE,
+                            COVIDClose),
+            by = c("OBJECTID" = "ID"))
 ```
@@
 
-ttm_transit <- ttm_transit %>%
+ttm_transit_2020 <- ttm_transit_2020 %>%
+  drop_na(TAZUID,
+          income_less40k)
+
+ttm_transit_2021 <- ttm_transit_2021 %>%
   drop_na(TAZUID,
@@
 
-ttm_transit <- ttm_transit %>%
+ttm_transit_2020 <- ttm_transit_2020 %>%
   left_join(modes %>%
@@
 
-ttm_walk <- ttm_walk %>%
+ttm_transit_2021 <- ttm_transit_2021 %>%
   left_join(modes %>%
@@
             by = "TAZUID")
+
+ttm_walk <- ttm_walk %>%
+  left_join(modes %>%
+              st_drop_geometry() %>%
+              transmute(TAZUID,
+                        p_car,
+                        p_transit,
+                        p_walk),
+            by = "TAZUID")
 ```
@@
 ```{r}
-ttm <- ttm_car %>%
+ttm_2020 <- ttm_car %>%
   rename(travel_time_car = travel_time) %>%
-  full_join(ttm_transit %>%
+  full_join(ttm_transit_2020 %>%
               transmute(UID,
@@
             by = c("UID", "OBJECTID"))
+
+ttm_2021 <- ttm_car %>%
+  rename(travel_time_car = travel_time) %>%
+  full_join(ttm_transit_2021 %>%
+              transmute(UID,
+                        OBJECTID,
+                        travel_time_transit = travel_time),
+            by = c("UID", "OBJECTID")) %>%
+  full_join(ttm_walk %>%
+              transmute(UID,
+                        OBJECTID,
+                        travel_time_walk = travel_time),
+            by = c("UID", "OBJECTID"))
 ```
@@
 ```{r}
-ttm %>% 
+ttm_2020 %>% 
   filter(is.na(travel_time_transit)) %>%
@@
   summary()
+
+ttm_2021 %>% 
+  filter(is.na(travel_time_transit)) %>%
+  select(travel_time_transit,
+         p_transit) %>%
+  summary()
 ```
@@
 ```{r}
-ttm %>% 
+ttm_2020 %>% 
   filter(is.na(travel_time_transit) & p_transit > 0) %>%
@@
   summary()
+
+ttm_2021 %>% 
+  filter(is.na(travel_time_transit) & p_transit > 0) %>%
+  select(travel_time_transit,
+         p_transit) %>%
+  summary()
 ```
@@
 ```{r}
-ttm <- ttm %>% 
+ttm_2020 <- ttm_2020 %>% 
   mutate(travel_time_transit = replace_na(travel_time_transit, 0))
+
+ttm_2021 <- ttm_2021 %>% 
+  mutate(travel_time_transit = replace_na(travel_time_transit, 0))
 ```
@@
 ```{r}
-ttm %>% 
+ttm_2020 %>% 
   filter(is.na(travel_time_walk)) %>%
@@
   summary()
+
+ttm_2021 %>% 
+  filter(is.na(travel_time_walk)) %>%
+  select(travel_time_walk,
+         p_walk) %>%
+  summary()
 ```
@@
 ```{r}
-ttm %>% 
+ttm_2020 %>% 
   filter(is.na(travel_time_walk) & p_walk > 0) %>%
@@
   summary()
+
+ttm_2021 %>% 
+  filter(is.na(travel_time_walk) & p_walk > 0) %>%
+  select(travel_time_walk,
+         p_walk) %>%
+  summary()
 ```
@@
 ```{r}
-ttm <- ttm %>% 
+ttm_2020 <- ttm_2020 %>% 
   mutate(travel_time_walk = replace_na(travel_time_walk, 0))
+
+ttm_2021 <- ttm_2021 %>% 
+  mutate(travel_time_walk = replace_na(travel_time_walk, 0))
 ```
@@
 ```{r}
-summary(ttm)
+summary(ttm_2020)
+summary(ttm_2021)
 ```
@@
 ```{r}
-ttm <- ttm %>%
+ttm_2020 <- ttm_2020 %>%
   mutate(e_travel_time_car = travel_time_car * p_car,
@@
          e_travel_time = e_travel_time_car + e_travel_time_transit + e_travel_time_walk)
-```
-
-What is the mean expected travel time for each parcel over all the foodbanks (open and closed)?
-```{r}
-e_travel_time_parcel <- ttm %>% 
+
+ttm_2021 <- ttm_2021 %>%
+  mutate(e_travel_time_car = travel_time_car * p_car,
+         e_travel_time_transit = travel_time_transit * p_transit,
+         e_travel_time_walk = travel_time_walk * p_walk,
+         e_travel_time = e_travel_time_car + e_travel_time_transit + e_travel_time_walk)
+```
+
+What is the mean expected travel time for each parcel over all the foodbanks (open and closed) in 2020?
+```{r}
+e_travel_time_parcel_2020 <- ttm_2020 %>% 
   group_by(UID) %>%
@@
 
-summary(e_travel_time_parcel)
+summary(e_travel_time_parcel_2020)
 ```
@@
 ```{r}
-e_travel_time_parcel <- e_travel_time_parcel %>%
+e_travel_time_parcel_2020 <- e_travel_time_parcel_2020 %>%
   left_join(dwelling_network_points_2016 %>%
@@
   geom_sf(data = hamilton_boundary) +
-  geom_sf(data = e_travel_time_parcel,
+  geom_sf(data = e_travel_time_parcel_2020,
           aes(color = e_travel_time),
@@
 ```{r}
-e_travel_time_parcel_2 <- ttm %>% 
+e_travel_time_parcel_2021 <- ttm_2021 %>% 
   filter(COVIDClose == "No") %>%
@@
 
-summary(e_travel_time_parcel_2)
+summary(e_travel_time_parcel_2021)
 ```
@@
 ```{r}
-e_travel_time_parcel_2 <- e_travel_time_parcel_2 %>%
+e_travel_time_parcel_2021 <- e_travel_time_parcel_2021 %>%
   left_join(dwelling_network_points_2016 %>%
@@
   geom_sf(data = hamilton_boundary) +
-  geom_sf(data = e_travel_time_parcel_2,
+  geom_sf(data = e_travel_time_parcel_2021,
           aes(color = e_travel_time),
@@
 ```{r}
-summary(e_travel_time_parcel)
-summary(e_travel_time_parcel_2)
+summary(e_travel_time_parcel_2020)
+summary(e_travel_time_parcel_2021)
 ```
@@
 
-b2sfca <- function(ttm = ttm, threshold = threshold){
+b2sfca <- function(ttm, threshold){
   # calculate the impedance
@@
 ```{r}
-acc_15min <- ttm %>% 
+acc_15min <- ttm_2020 %>% 
   transmute(UID, OBJECTID, travel_time = e_travel_time, population = income_less40k, supply = 1) %>%
@@
 for(i in 1:length(threshold_seq)){
-  results_test <-  ttm %>% 
+  results_test <-  ttm_2020 %>% 
   transmute(UID, OBJECTID, travel_time = e_travel_time, population = income_less40k, supply = 1) %>%
@@
 ```{r 30-min-pre-covid, include=FALSE}
-results_30min_pre <- ttm %>%
+results_30min_pre <- ttm_2020 %>%
   #dplyr::filter(COVIDClose == "No") %>%
@@
 ```{r 30-min-during-covid, include=FALSE}
-results_30min_covid <- ttm %>%
+results_30min_covid <- ttm_2021 %>%
   dplyr::filter(COVIDClose == "No") %>%
@@
 
-min_ttm_transit_baseline <- ttm_transit %>%
+min_ttm_transit_baseline <- ttm_transit_2020 %>%
   #dplyr::filter(COVIDClose == "No") %>%
@@
 
-min_ttm_transit_covid <- ttm_transit %>%
+min_ttm_transit_covid <- ttm_transit_2021 %>%
   dplyr::filter(COVIDClose == "No", TYPE != "Food Banks") %>%

### 2021-08-27 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton copy.Rmd

--- 2021-04-28 README.Rmd
@@
 ---
-output:
-  github_document:
-    pandoc_args: --webtex
-always_allow_html: true
-bibliography: Accessibility-Foodbanks-Hamilton/references.bib
----
-
-<!-- README.md is generated from README.Rmd. Please edit that file -->
-
-```{r, include = FALSE}
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "#>"
-)
-```
-
-# Changes in accessibility to food banks and food services during COVID-19 and implications for low income populations in Hamilton, Ontario
-
-<!-- badges: start -->
-<!-- badges: end -->
-
-Christopher D. Higgins (University of Toronto Scarborough)  
-Antonio Páez (McMaster University)  
-Gyoorie Kim (University of Toronto Mississauga)  
-Jue Wang (University of Toronto Mississauga)
-
-Paper submitted to *Social Science and Medicine*
-
-## Abstract
-
-In this paper we analyze the changes in accessibility to food banks and related services before and during the COVID-19 pandemic in the City of Hamilton, Ontario.
-Food banks and services are the last line of support for households facing food insecurity; as such, their relevance cannot be ignored in the midst of the economic upheaval caused by the pandemic.
-Our analysis is based on the application of balanced floating catchment areas, and concentrates on households with lower incomes (<CAD40,000, approximately the Low Income Cutoff Value for a city of Hamilton's size).
+title: "Changes in accessibility to food banks and food services during COVID-19 and implications for low income populations in Hamilton, Ontario"
+author:
+  - name: Author One
+    email: author1@example.com
+    affiliation: Some University
+  - name: Author Two
+    email: author2@example.com
+    affiliation: Another University
+    footnote: 1
+  - name: Author Three
+    email: author3@example.com
+    affiliation: Some University
+  - name: Author Four
+    email: author4@example.com
+    affiliation: Some University
+address:
+  - code: Some University
+    address: Department, Street, City, State, Zip
+  - code: Another University
+    address: Department, Street, City, State, Zip
+footnote:
+  - code: 1
+    text: "Corresponding Author"
+abstract: |
+  In this paper we analyze the changes in accessibility to community and emergency food services before and during the COVID-19 pandemic in the City of Hamilton, Ontario.
+Many of these food services are the last line of support for households facing food insecurity; as such, their relevance cannot be ignored in the midst of the economic upheaval caused by the pandemic.
+Our analysis is based on the application of balanced floating catchment areas and concentrates on households with lower incomes (<CAD40,000, approximately the Low Income Cutoff Value for a city of Hamilton's size).
 We find that accessibility was low to begin with in suburban and exurban parts of the city; furthermore, about 14\% of locations originally available in Hamilton closed during the pandemic, further reducing accessibility.
@@
 As well, children (age $\le 18$) appear to have been impacted the most.
-
-## Keywords
-
-- Food insecurity  
-- Food banks  
-- Accessibility  
-- COVID-19  
-- Hamilton  
-- Ontario
-- Reproducible research
+journal: "Social Science and Medicine"
+date: "`r Sys.Date()`"
+bibliography: references.bib
+#linenumbers: true
+#numbersections: true
+csl: elsevier-harvard.csl
+output: rticles::elsevier_article
+header-includes:
+    - \usepackage[table]{xcolor}
+---
 
@@
 data(ttm_car)
-data(ttm_transit)
+#data(ttm_transit_2020)
+#data(ttm_transit_2021)
+load("output-data-files/ttm_transit_2020.RData")
+load("output-data-files/ttm_transit_2021.RData")
 data(ttm_walk)
@@
 
-ttm_transit <- ttm_transit %>%
+ttm_transit_2020 <- ttm_transit_2020 %>%
+  filter(!between(OBJECTID, left = 1, right = 3))
+
+ttm_transit_2021 <- ttm_transit_2021 %>%
   filter(!between(OBJECTID, left = 1, right = 3))
@@
 
-ttm_transit <- ttm_transit %>%
+ttm_transit_2020 <- ttm_transit_2020 %>%
+  left_join(foodbank_locations %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID, 
+                            TYPE,
+                            COVIDClose),
+            by = c("OBJECTID" = "ID"))
+
+ttm_transit_2021 <- ttm_transit_2021 %>%
   left_join(foodbank_locations %>%
@@
 # Join Tables into Single TTM
-ttm <- ttm_car %>%
+ttm_2020 <- ttm_car %>%
   rename(travel_time_car = travel_time) %>%
-  full_join(ttm_transit %>%
+  full_join(ttm_transit_2020 %>%
               transmute(UID,
@@
 ```{r travel-time-tables-data-check, include=FALSE}
-## Data preparation
+## Data preparation - 2020
 # Check the NAs in travel time table:
-ttm %>% 
+ttm_2020 %>% 
   filter(is.na(travel_time_transit)) %>%
@@
 How many of these cases?
-ttm %>% 
+ttm_2020 %>% 
   filter(is.na(travel_time_transit) & p_transit > 0) %>%
@@
 # It should be fine to replace the NAs with zeros, which multiplied by p_transit will return a zero (i.e., it is not possible to reach a foodbank by transit).
-ttm <- ttm %>% 
-  mutate(travel_time_transit = replace_na(travel_time_transit, 0))
+#ttm_2020 <- ttm_2020 %>% 
+#  mutate(travel_time_transit = replace_na(travel_time_transit, 0))
 
 # Check the NAs in travel time by walk:
-ttm %>% 
+ttm_2020 %>% 
   filter(is.na(travel_time_walk)) %>%
@@
 How many of these cases?
-ttm %>% 
+ttm_2020 %>% 
   filter(is.na(travel_time_walk) & p_walk > 0) %>%
@@
 # Like with transit, it should be fine to replace the NAs with zeros, which multiplied by p_walk will return a zero (i.e., it is not possible to reach a foodbank by walking).
-ttm <- ttm %>% 
-  mutate(travel_time_walk = replace_na(travel_time_walk, 0))
+#ttm_2020 <- ttm_2020 %>% 
+#  mutate(travel_time_walk = replace_na(travel_time_walk, 0))
 
 #Summary of the table after these operations:
-summary(ttm)
-```
-
-```{r calculate-expected-travel-times, include=FALSE}
+summary(ttm_2020)
+
+## Data preparation - 2021
+# Check the NAs in travel time table:
+ttm_2021 %>% 
+  filter(is.na(travel_time_transit)) %>%
+  select(travel_time_transit,
+         p_transit) %>%
+  summary()
+
+# What do NAs here mean?
+Most DAs missing travel time by transit do not report trips by transit in the TTS.
+In some DAs there are trips by transit (p_transit != 0), but no foodbank can be reached by transit (travel_time_transit is NA).
+How many of these cases?
+ttm_2021 %>% 
+  filter(is.na(travel_time_transit) & p_transit > 0) %>%
+  select(travel_time_transit,
+         p_transit) %>%
+  summary()
+
+# It should be fine to replace the NAs with zeros, which multiplied by p_transit will return a zero (i.e., it is not possible to reach a foodbank by transit).
+#ttm_2021 <- ttm_2021 %>% 
+#  mutate(travel_time_transit = replace_na(travel_time_transit, 0))
+
+# Check the NAs in travel time by walk:
+ttm_2021 %>% 
+  filter(is.na(travel_time_walk)) %>%
+  select(travel_time_walk,
+         p_walk) %>%
+  summary()
+
+# Most DAs missing travel time by walking do not report trips by walking in the TTS.
+In some DAs there are trips by walking (p_walk != 0), but no foodbank can be reached by walk (travel_time_walk is NA).
+How many of these cases?
+ttm_2021 %>% 
+  filter(is.na(travel_time_walk) & p_walk > 0) %>%
+  select(travel_time_walk,
+         p_walk) %>%
+  summary()
+
+# Like with transit, it should be fine to replace the NAs with zeros, which multiplied by p_walk will return a zero (i.e., it is not possible to reach a foodbank by walking).
+#ttm_2021 <- ttm_2021 %>% 
+#  mutate(travel_time_walk = replace_na(travel_time_walk, 0))
+
+#Summary of the table after these operations:
+summary(ttm_2021)
+```
+
+```{r calculate-expected-travel-times, eval = FALSE, include=FALSE}
 ## Data preparation
@@
 The total expected travel time is the sum of the expected travel times by mode:
-ttm <- ttm %>%
+ttm_2020 <- ttm_2020 %>%
   mutate(e_travel_time_car = travel_time_car * p_car,
@@
          e_travel_time = e_travel_time_car + e_travel_time_transit + e_travel_time_walk)
-```
-
-\newpage
-
-Introduction
-==========================
+
+ttm_2021 <- ttm_2021 %>%
+  mutate(e_travel_time_car = travel_time_car * p_car,
+         e_travel_time_transit = travel_time_transit * p_transit,
+         e_travel_time_walk = travel_time_walk * p_walk,
+         e_travel_time = e_travel_time_car + e_travel_time_transit + e_travel_time_walk)
+```
+
+```{r}
+ttm_2020_long <- ttm_2020 %>% transmute(UID, OBJECTID,
+                                        car_prob = p_car, 
+                                        transit_prob = p_transit, 
+                                        walk_prob = p_walk, 
+                                        car_traveltime = travel_time_car, 
+                                        transit_traveltime = travel_time_transit, 
+                                        walk_traveltime = travel_time_walk) %>% 
+  pivot_longer(cols = !c("UID", "OBJECTID"), 
+               names_to = c("mode", ".value"),
+               names_sep = "_") %>%
+  group_by(UID, OBJECTID) %>%
+  summarize(e_travel_time = weighted.mean(traveltime, prob, na.rm = TRUE))
+
+# income_less40k = first(income_less40k)) %>% 
+#  filter(!is.nan(e_travel_time))
+
+ttm_2021_long <- ttm_2021 %>% transmute(UID, OBJECTID,
+                                        car_prob = p_car, 
+                                        transit_prob = p_transit, 
+                                        walk_prob = p_walk, 
+                                        car_traveltime = travel_time_car, 
+                                        transit_traveltime = travel_time_transit, 
+                                        walk_traveltime = travel_time_walk) %>% 
+  pivot_longer(cols = !c("UID", "OBJECTID"), 
+               names_to = c("mode", ".value"),
+               names_sep = "_") %>%
+  group_by(UID, OBJECTID) %>%
+  summarize(e_travel_time = weighted.mean(traveltime, prob, na.rm = TRUE))
+# income_less40k = first(income_less40k)) %>% 
+#  filter(!is.nan(e_travel_time))
+
+ttm_2020 <- ttm_2020 %>% left_join(ttm_2020_long, by = c("UID", "OBJECTID")) %>% 
+  filter(!is.nan(e_travel_time))
+
+ttm_2021 <- ttm_2021 %>% left_join(ttm_2021_long, by = c("UID", "OBJECTID")) %>% 
+  filter(!is.nan(e_travel_time))
+```
+
+# Introduction
 
@@
 Along several other dimensions of stress [e.g., accessibility to health care facilities, @ghorbanzadeh2021spatial; @pereira2021geographic], this seems to be the case for food insecurity as well [@laborde2020poverty].
-In the US, for example, it has been estimated that there was an increase of more than 30\% in household food insecurity, and more than one third of households were discovered to be newly food insecure - meaning they did not experience food insecurity before the pandemic [@niles2020early].
-In Canada, Men and Tarasuk [-@men2021food] report that about 25\% of individuals who experienced job insecurity (a relatively common occurrence during the pandemic), also experienced food insecurity.
-Similarly, according to Statistics Canada [-@statisticscanada2020food], in the early stages of the pandemic almost 15\% of individuals reported living in a household that faced food insecurity; the risk of food insecurity was substantially higher for households with children.
-The difference between households with and without children was significant, and 11.7\% of households with children indicated that "food didn’t last and [there was] no money to get more" sometimes or often, compared to 7.3\% of households without children; likewise, 13\% of households with children indicated that they "[c]ouldn’t afford balanced meals" sometimes or often, compared to 8.8\% of households without children.
+According to Statistics Canada [-@statisticscanada2020food], in the early stages of the pandemic almost 15% of individuals reported living in a household that faced food insecurity; the risk of food insecurity was substantially higher for households with children.
+The difference between households with and without children was significant, and 11.7% of households with children indicated that "food didn't last and [there was] no money to get more" sometimes or often, compared to 7.3% of households without children; likewise, 13% of households with children indicated that they "[c]ouldn't afford balanced meals" sometimes or often, compared to 8.8% of households without children.
+Additionally, Men and Tarasuk [-@men2021food] report that about 25% of individuals who experienced job insecurity (a relatively common occurrence during the pandemic) also experienced food insecurity associated with COVID-related disruptions to employment, financial hardship, and use of food charity.
 
 The impacts of food insecurity during the pandemic are alarming, since diet-related diseases, such as obesity, heart-disease, and diabetes, were already critical public health concerns in Canada prior to COVID-19 [@boucher2017ontario].
-While food banks are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least they can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
+While emergency food services are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least they can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
 As a mid-size city grappling with deindustrialization, Hamilton exhibits high rates of poverty and food bank use.
-As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that food banks in the city recorded the highest number of visitors in the past 29 years; a rate of increase greater than population growth.
+As recently as 2019, the Hamilton Hunger Report[^1] noted that food banks in the city recorded the highest number of visitors in the past 29 years; a rate of increase greater than population growth.
 Most troubling, approximately 40% of all visitors were children.
 
+[^1]: <https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf>
+
 It is known that urban food environments, within which people make their daily food choices, are essential in influencing eating behaviours and health outcomes, based on factors such as food availability, ease of geographic accessibility and socio-demographic variations [@paez2010relative; @vanderlee2017food; @widener2018spatial].
-However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to food banks.
-Previous work has explored differences in _accessing_ food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource [e.g., @enns2020experiences].
+However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to emergency and community food services.
+Previous work has explored differences in *accessing* food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource [e.g., @enns2020experiences].
 In addition, transportation and locational considerations have been raised as key issues in food bank accessibility in previous qualitative research [@smith2017food].
@@
 The study of place-based geographic accessibility is concerned with capturing the potential to reach destinations of value using the transportation network [@paez2012measuring].
-Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities^[https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with food bank service delivery [@holmes2018nothing].
-However, as with research into the prevalence of food deserts, accessibility to food banks is unlikely to be evenly distributed, and variation throughout a city can be expected due to transportation network characteristics, and the spatial distribution of food bank locations and the population they are meant to serve.
+Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities[^2] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with food bank service delivery [@holmes2018nothing].
+However, as with research into the prevalence of food deserts, accessibility to food banks is unlikely to be evenly distributed, and variation throughout a city can be expected due to transportation network characteristics and the spatial distribution of service locations and the population they are meant to serve.
 Furthermore, policy responses to the COVID-19 pandemic likely have added to the distress of vulnerable households.
@@
 
+[^2]: <https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042>
+
 For this study, we aim to look at how the landscape of food banks and related services (e.g.
@@
 Did the number of open food bank services diminish?
-If so, what was the accessibility to food banks before the pandemic from the perspective of low income households, and how it changed during the pandemic?
+If so, what was the accessibility to food banks before the pandemic from the perspective of low income households, and how has it changed during the pandemic with respect to geographic access and congestion at remaining sites?
 And finally, who are most likely to have been impacted by changes in the accessibility landscape?
@@
 This paper follows reproducible research recommendations [see @brunsdon2020opening], and the research was conducted using open source tools for transportation analysis [@lovelace2021open].
-The code and data necessary to reproduce the analysis are available in a public repository^[https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing].
-
-Food Insecurity and Food Bank Use in Canada
-============
+The code and data necessary to reproduce the analysis are available in a public repository[^3].
+
+[^3]: <https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing>
+
+# Food Insecurity and Food Bank Use in Canada
 
@@
 
-Food banks - sometimes also referred to as ‘food pantries’ and ‘food shelves’ - originated as a community response to aid those with inadequate food by voluntarily offering them meals and ingredients [@riches2002food; @loopstra2012relationship].
+Food banks - sometimes also referred to as 'food pantries' and 'food shelves' - originated as a community response to aid those with inadequate food by voluntarily offering them meals and ingredients [@riches2002food; @loopstra2012relationship].
 The scope and objectives of food banks can vary by region and by country, and these organizations can include not only prepared meals and aliments, but also shared spaces to connect in community gardens and community kitchens [@wakefield2013sweet].
@@
 The largest database of food banks and their use comes from the non-profit association Food Banks Canada (FBC), which conducts an annual assessment through its affiliated members.
-FBC's 2018 Hunger Count report^[https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018_FINAL_EN.pdf.aspx?ext=.pdf] (the most recent available) listed 1,830 member food banks across the country, and found that Canadians visited food banks 1.1 million times in March of 2018.
+FBC's 2018 Hunger Count report[^4] (the most recent available) listed 1,830 member food banks across the country, and found that Canadians visited food banks 1.1 million times in March of 2018.
 Of those accessing food banks, certain population characteristics tend to be over-represented compared to national totals from the 2016 Canadian Census of Population.
-According to FBC's 2018 data, single-adult households represent 45% of those utilizing food banks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those accessing food bank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
+According to FBC's 2018 data, single-adult households represent 45% of those utilizing food banks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those using food bank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
 In addition, 59% of households accessing food banks list social or disability assistance as their primary source of income.
+
+[^4]: <https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018_FINAL_EN.pdf.aspx?ext=.pdf>
+
 Similarly, using data from the 2011-2012 CCHS, Tarasuk et al.
 [-@tarasuk2019geographic] found higher odds of food insecurity amongst households relying on social assistance, those without a university degree or with children under the age of 18, and individuals that lived alone, renters, and those identifying as Aboriginal.
-While surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada [@tarasuk2014food], research from Ottawa [@enns2020experiences] and Vancouver [@black2020examining] suggests that long-term users tend to be older, have health or mobility challenges, live in large households, and are less likely to have employment income.
+While surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada [@tarasuk2014food], pre-pandemic research from Ottawa [@enns2020experiences] and Vancouver [@black2020examining] suggests that long-term users tend to be older, have health or mobility challenges, live in large households, and are less likely to have employment income.
 
@@
 Considering the negative mental and physical health effects, increases in food insecurity rates due to the onset of COVID-19, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
-Recent data^[https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis_Exec-Sum_EN.pdf] from FBC showed that 52% of member food banks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of the country.
+Recent data[^5] from FBC showed that 52% of member food banks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of the country.
 The pandemic also created significant staffing issues with 42% of food banks reporting a reduction in volunteers.
@@
 Nevertheless, some of these benefit programs were temporary.
-Although more recent statistics on food bank use in Hamilton in 2020 and 2021 are not yet available, data from the Daily Bread Food Bank in neighbouring Toronto for August 2020 shows visits climbing 51% year-over-year in that city^[https://www.dailybread.ca/wp-content/uploads/2020/11/Whos-Hungry-Report-2020.pdf], which suggests that many households in Hamilton are likely to turn to food bank services to meet their needs.
+Although more recent statistics on food bank use in Hamilton in 2020 and 2021 are not yet available, data from the Daily Bread Food Bank in neighbouring Toronto for August 2020 shows visits climbing 51% year-over-year in that city[^6], which suggests that many households in Hamilton are likely to turn to food bank services to meet their needs.
+
+[^5]: <https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis_Exec-Sum_EN.pdf>
+
+[^6]: <https://www.dailybread.ca/wp-content/uploads/2020/11/Whos-Hungry-Report-2020.pdf>
 
@@
 Consequently, the location of facilities matters; in the words of an interviewee: "I wish it [the food bank] was a little more centrally located.
-Because if I didn’t have a bike I’d have to walk it all the way out there and back.
-I wonder about people who don’t" (p.
+Because if I didn't have a bike I'd have to walk it all the way out there and back.
+I wonder about people who don't" (p.
 39).
@@
 
-Methods and Materials
-===================
+# Methods and Materials
 
@@
 This method for estimating accessibility is a form of the widely-used two-stage floating catchment area method [@radke2000spatial; @luo2003measures].
-Floating catchment areas are used to estimate accessibility when there are potential congestion effects, and operate by calculating first the _demand_ for spatially distributed services.
+Floating catchment areas are used to estimate accessibility when there are potential congestion effects, and operate by calculating first the *demand* for spatially distributed services.
 The demand (usually the number of people who require a service) is used to calculate a level of service.
@@
 More formally, the first step of this method is as follows:
+
 $$
@@
 The second step in this process is then summing the level of service that each population unit can reach, according to the distance-decay weight:
+
 $$
@@
 Floating catchment area methods are prone to overestimation of the population and the level of service due to multiple-counting.
-The population at $P_i$ is allocated to _every_ service point $j$ for which $w_{ij}>0$.
-Similarly, the level of service at $LOS_j$ is allocated to _every_ population point for which $w_{ji}>0$.
+The population at $P_i$ is allocated to *every* service point $j$ for which $w_{ij}>0$.
+Similarly, the level of service at $LOS_j$ is allocated to *every* population point for which $w_{ji}>0$.
 This inflation effect has been known for several years, and several modifications have been proposed to mitigate it [e.g., @wan2012three; @delamater2013spatial].
@@
 [-@paez2019demand].
-In order to avoid the multiple-counting in the summations, the population and the level of service need to be allocated _proportionally_.
+In order to avoid the multiple-counting in the summations, the population and the level of service need to be allocated *proportionally*.
 This is achieved by standardizing the weights as follows:
+
 $$
@@
 $$
+
 \noindent and:
+
 $$
@@
 The standardized weights satisfy the following conditions:
+
 $$
@@
 \noindent and:
+
 $$
@@
 Since the population is allocated proportionally, its value is preserved:
+
 $$
@@
 \noindent as is the level of service:
+
 $$
@@
 With a population of around 540,000, the City of Hamilton is the fourth largest city in Ontario.
-It has historically been home to major manufacturing industries but de-industrialization that has occurred over the past several decades has led Hamilton to become one of the most highly divided cities in Ontario, with a significant proportion of its residents living at or below Canada’s poverty level [@deluca2012code; @jakar2019turning; @latham2007determinants].
-In 2016, the Hamilton Community Foundation reported that based on the Low-Income Cut-Off, Hamilton recorded a poverty rate of 16.7%, which was well above the average rate of Ontario (13.7%) and the average national rate (12.8%).^[https://www.hamiltoncommunityfoundation.ca/vital-signs/low-income-2018/].
-According to data from Hamilton Food Share^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf], approximately 23,000 individuals accessed food banks in the city in March of 2019.
+It has historically been home to major manufacturing industries but de-industrialization that has occurred over the past several decades has led Hamilton to become one of the most highly divided cities in Ontario, with a significant proportion of its residents living at or below Canada's poverty level [@deluca2012code; @jakar2019turning; @latham2007determinants].
+In 2016, the Hamilton Community Foundation reported that based on the Low-Income Cut-Off, Hamilton recorded a poverty rate of 16.7%, which was well above the average rate of Ontario (13.7%) and the average national rate (12.8%).[^7].
+According to data from Hamilton Food Share[^8], approximately 23,000 individuals accessed food banks in the city in March of 2019.
 Within this total is 9,125 visits by children (minors up to 18 years old), up from 8,278 the year before.
-Feed Ontario, the province's largest collective of hunger-relief organizations, found that on a per-capita basis, the level of need in the inner core of central Hamilton was second highest in Ontario^[https://feedontario.ca/research/report-hunger-map/].
+Feed Ontario, the province's largest collective of hunger-relief organizations, found that on a per-capita basis, the level of need in the inner core of central Hamilton was second highest in Ontario[^9].
+
+[^7]: <https://www.hamiltoncommunityfoundation.ca/vital-signs/low-income-2018/>
+
+[^8]: <https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf>
+
+[^9]: <https://feedontario.ca/research/report-hunger-map/>
 
@@
 
-Data have been prepared for sharing as a data package^[https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing]. The contents of the data package are described next.
+Data have been prepared for sharing as a data package[^10]. The contents of the data package are described next.
+
+[^10]: <https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing>
 
@@
 
-The locations of food banks and related food services were obtained from the Hamilton Public Library's Food Access Guide^[http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf].
+The locations of food banks and related food services were obtained from the Hamilton Public Library's Food Access Guide[^11].
 The guide was updated in April of 2021 to indicate any change affected on the services due to the pandemic.
@@
 
+[^11]: <http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf>
+
 ```{r table-food-bank-info, echo=FALSE, results='asis'}
-foodbank_table %>% kable("html",
+foodbank_table %>% kable("latex",
                          booktabs = TRUE,
@@
                          escape = FALSE) %>%
-  kable_styling(bootstrap_options = c("striped")) %>%
+  kable_styling(latex_options = c("striped", "scale_down")) %>%
   column_spec(2, width = "15em") %>%
@@
 
-Travel time tables for three modes (car, transit, walking) were computed using the parcels as the origins and the locations of the food banks as the destinations.
-For routing, the package `r5r` [@pereira2021r5r] was used with a network extract for the City of Hamilton from OpenStreetMaps and the General Transit Feed Specification from Hamilton Street Railway, the local transit operator.
-For routing purposes we used maximum values of 180 min and 10,000 m walking distance: any destination that exceeded these thresholds was ignored.
-The departure time used for routing was 8:00AM on March 30, 2021 to reflect transit service around the morning service peak on a typical Tuesday.
+Travel time tables for three modes (car, transit, walking) were computed using the parcels as the origins and the locations of the community and emergency food service locations as the destinations.
+For routing, the package `r5r` [@pereira2021r5r] was used with a network extract for the City of Hamilton from OpenStreetMaps and the General Transit Feed Specification (GTFS) files for the Hamilton Street Railway, the local transit operator, as well as for Burlington Transit, which operates some service in the city.
+For transit routing purposes we used maximum travel time values of 300 min and a 2,000 m cap on walking distance: any destination that exceeded these thresholds was ignored.
+The transit analysis was run twice to account for changes in service levels before and during the pandemic.
+For the pre-pandemic period, the departure time used for routing was 8:00AM on March 31, 2020 to reflect transit service around the morning service peak on a typical Tuesday just prior to the onset of major pandemic-related lockdowns in Ontario.
+For the pandemic period, we re-ran the analysis for March 30, 2021.
 
@@
 Once we obtained travel time tables with population (number of households) and proportion of trips by mode, we calculated the expected travel time $ett$ from each parcel $i$ to a food bank or food service location $j$ as follows:
+
 $$
@@
 \noindent where $p^k_i$ is the proportion of trips by mode $k$ in the TAZ of parcel $i$, and $tt^k_{ij}$ is the travel time from parcel $i$ to the food bank.
-In other words, the expected travel time is the weighted sum of travel times to the food bank, with the weights given by the expected modal split of trips made by low-income households in the TAZ per the TTS data.
-
-Results and Discussion
-===================
+In other words, the expected travel time reflects the weighted average of travel times to the food bank, with the weights given by the expected modal split of trips made by low-income households in the TAZ per the TTS data.
+
+# Results and Discussion
 
@@
 for(i in 1:length(threshold_seq)){
-  results_test <-  ttm %>% 
+  results_test <-  ttm_2021 %>% 
     transmute(UID, OBJECTID, travel_time = e_travel_time, population = income_less40k, supply = 1) %>%
@@
 # Calculate k-min accessibility pre-covid:
-results_pre <- ttm %>%
+results_pre <- ttm_2021 %>%
   #dplyr::filter(COVIDClose == "No") %>%
@@
 # Calculate k-min accessibility during covid:
-results_covid <- ttm %>%
+results_covid <- ttm_2021 %>%
+  dplyr::filter(COVIDClose == "No") %>%
+  transmute(UID, 
+            OBJECTID, 
+            travel_time = e_travel_time, 
+            population = income_less40k, 
+            supply = 1) %>%
+  b2sfca(threshold = 15)
+```
+
+```{r results-during-covid-on-2020-network, include=FALSE}
+## Analysis 
+# Calculate k-min accessibility during covid:
+results_covid_2020 <- ttm_2020 %>%
   dplyr::filter(COVIDClose == "No") %>%
@@
 
+```{r, bind-accessibility-results-for-network-comparison, include=FALSE}
+## Analysis
+# Bind level of service results:
+los_2020 <- rbind(data.frame(results_covid$los, 
+                        timing = "During covid", 
+                        threshold = "30 min"),
+                  data.frame(results_covid_2020$los, 
+                        timing = "Covid 2020", 
+                        threshold = "30 min")) %>%
+  mutate(timing = factor(timing, 
+                         levels = c("During covid", 
+                                    "Covid 2020"),
+                         ordered = TRUE)) %>% 
+  left_join(foodbank_locations %>% transmute(OBJECTID = ID, TYPE), by = "OBJECTID")
+
+los_2020_diff <- los_2020 %>% filter(TYPE == "Food Banks") %>% 
+  pivot_wider(id_cols = OBJECTID, names_from = "timing", values_from = "los") %>% 
+  mutate(los_diff = `During covid` / `Covid 2020`) %>% 
+  summarize(mean_diff = mean(los_diff, na.rm = TRUE), 
+            max_diff = max(los_diff, na.rm = TRUE), 
+            min_diff = min(los_diff, na.rm = TRUE))
+
+# Bind accessibility results:
+accessibility_2020 <- rbind(data.frame(results_covid$accessibility, 
+                                  timing = "During covid", 
+                                  threshold = "30 min"),
+                            data.frame(results_covid_2020$accessibility, 
+                                  timing = "Covid 2020", 
+                                  threshold = "30 min")) %>%
+  mutate(timing = factor(timing, 
+                         levels = c("During covid",
+                                    "Covid 2020"),
+                         ordered = TRUE))
+
+access_2020_diff <- accessibility_2020 %>%
+  pivot_wider(id_cols = UID, names_from = "timing", values_from = "accessibility") %>%
+
+  left_join(dwelling_network_points_2016 %>%
+              st_drop_geometry() %>%
+              dplyr::select(ID, TAZUID),
+            by = c("UID" = "ID")) %>%
+  group_by(TAZUID) %>% summarize(access_covid = sum(`During covid`, na.rm = TRUE), access_2020 = sum(`Covid 2020`, na.rm = TRUE)) %>%
+  mutate(acc_diff = access_covid / access_2020) %>%
+  #left_join(modes %>% transmute(TAZUID, geometry), by = "TAZUID") %>% st_as_sf() #%>%
+  summarize(mean_diff = mean(acc_diff, na.rm = TRUE), 
+            max_diff = max(acc_diff, na.rm = TRUE), 
+            min_diff = min(acc_diff, na.rm = TRUE))
+```
+
 Figure \ref{fig:foodbanks} shows the location of food banks and services in the City of Hamilton and their status.
-Before the pandemic there were `r nrow(foodbank_locations)` of which `r nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))` (`r round(nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations) * 100, 2)`\%) closed during the pandemic.
+Before the pandemic there were `r nrow(foodbank_locations)` of which `r nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))` (`r round(nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations) * 100, 2)`%) closed during the pandemic.
 As shown in the figure, food services tend to be predominantly located in the central parts of the city.
@@
   scale_size_manual(values = c("Yes" = 3, "No" = 2)) +
-  guides(size = FALSE) +
+  guides(size = "none") +
   theme_minimal() +
@@
   ylim(43.2, 43.3) +
-  guides(size = FALSE) +
+  guides(size = "none") +
   theme_minimal() +
@@
 To implement the accessibility calculations, we must select a distance-decay function.
-In this task we find limited support in the literature, which is mostly silent on the travel patterns of people who visit food banks.
+In this task we find limited support in the literature, which is mostly silent on the travel patterns of people who visit food banks and community food services.
 For this reason, we opt for a simple cumulative opportunities function as follows:
+
 $$
@@
 $$
+
 \noindent where $ett_{ij}$ is the multimodal expected travel time as described previously, and $\delta$ is a travel threshold.
@@
 Figure \ref{fig:sensitivity-analysis} shows the results of conducting a sensitivity analysis of the system-wide accessibility as we vary the threshold (considering the situation before the pandemic).
-There is a clear pattern whereby more strict values of $\delta$ are associated with higher levels of system-wide accessibility: while increases in accessibility that result from decreases in the travel time window might seem counter-intuitive, this is a result of lower _congestion_, since fewer households are serviced and thus competition for the same resources is more limited.
+There is a clear pattern whereby more strict values of $\delta$ are associated with higher levels of system-wide accessibility: while increases in accessibility that result from decreases in the travel time window might seem counter-intuitive, this is a result of lower *congestion*, since fewer households are serviced and thus competition for the same resources is more limited.
 System-wide accessibility declines with higher values of $\delta$: as more households are serviced, congestion grows and the level of service declines, although this happens at a declining rate.
@@
 Using the 15-minute threshold, we find that the system-wide accessibility (interpreted as a provider-to-population ratio) was `r round(sum(results_pre$accessibility$accessibility), 3)` (food banks/service locations per low income household in the city) before COVID-19, but declined to `r round(sum(results_covid$accessibility$accessibility), 3)` during the pandemic.
-It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`\% of food facilities remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`\%, suggesting the location of food banks and related services plays an important role in serving those in need.
+It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`% of food facilities remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`%, suggesting the location of emergency and community food services plays an important role in serving those in need.
 
@@
 The general pattern of the levels of service is similar before and during the pandemic, with lower values in the center of the city where low-income households exhibit multimodal trip patterns that favour proximate service locations.
-Three more peripheral facilities towards the south of the city also have lower levels of service, presumably because they are expected to service relatively large suburban/exurban populations generally reliant on automobiles for travel.
+Three more peripheral facilities towards the south of the city have moderate levels of service, presumably because they are expected to service relatively suburban/exurban populations generally reliant on automobiles for travel.
 During the pandemic, however, the levels of service dropped, in some cases quite substantially.
 The pattern of the losses in level of service, moreover, is not uniform.
-The upper pane of Figure \ref{fig:levels-of-service-changes} shows that the three peripheral facilities in the southern suburban/exurban part of the city had low levels of service to begin with, but did not see major declines during the pandemic.
-Further, the inset map shows that the levels of service deteriorated more in the central part of the city.
-However, the loss of level of service was not as large in the core (where most of the food banks/services are found), but instead was more marked in the inner ring around the core, where facilities may have faced greater demand from both central city and suburban populations after the closure of food banks during the pandemic.
+The upper pane of Figure \ref{fig:levels-of-service-changes} shows that the peripheral facilities in the suburban/exurban parts of the city saw major declines during the pandemic as more urban locations closed and demand increased for the remaining locations.
+Further, the inset map shows that the levels of service also deteriorated in the central part of the city.
+However, the loss of level of service was not as large in the core (where most of the food banks/services are found), but instead was more marked in the inner ring around the core, where facilities may have faced greater demand from both central city and suburban populations after the closure of service locations during the pandemic.
+
+For the food banks specifically, while only one location closed after the onset of COVID-19, changes in the transportation network account for some differences in levels of service across the pre- and during pandemic periods.
+A comparison of levels-of-service calculated for the during-pandemic period using the 2021 travel time matrix compared to a scenario where the network is held constant using 2020 transit service levels reveals that the average level of service for remaining food banks was 0.3% lower when using the 2021 transit network.
+Of the 25 remaining food bank locations, 14 saw their levels-of-service decline when using the 2021 transit network compared to assuming constant 2020 transit service characteristics with a maximum decline of -21% across network scenarios.
+This suggests that transit service improved in some parts of the city which reduces the mode-weighted travel times between some dwellings and food bank locations and results in increased demand for service at these locations.
+On the other hand, 8 food banks exhibit increases in their levels of service of up to 7% across the network scenarios, suggesting changes in the transit network that increase travel times for some households to food bank locations lead to these locations servicing fewer people in our calculations.
 
@@
                         direction = -1,
-                        breaks = c(0.0005,
-                                   0.0010,
-                                   0.0015,
+                        breaks = c(0.0010,
                                    0.0020,
-                                   0.0025),
-                        labels = c("0.00050",
-                                   "0.00100",
-                                   "0.00150",
-                                   "0.00200",
-                                   "0.00250")) +
+                                   0.0040,
+                                   0.0060,
+                                   0.0080),
+                        labels = c("0.0010",
+                                   "0.0020",
+                                   "0.0040",
+                                   "0.0060",
+                                   "0.0080")) +
   scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  guides(size = FALSE, shape = FALSE) +
+  guides(size = "none", shape = "none") +
   theme_minimal() +
@@
                         direction = -1,
-                        breaks = c(0.00050,
-                                   0.00075,
-                                   0.00100,
-                                   0.00125,
-                                   0.00150),
-                        labels = c("0.00050",
-                                   "0.00075",
-                                   "0.00100",
-                                   "0.00125",
-                                   "0.00150")) +
+                        breaks = c(0.00100,
+                                   0.00200,
+                                   0.00400,
+                                   0.00600,
+                                   0.00800),
+                        labels = c("0.0010",
+                                   "0.0020",
+                                   "0.0040",
+                                   "0.0060",
+                                   "0.0080")) +
   scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  guides(size = FALSE, shape = FALSE) +
+  guides(size = "none", shape = "none") +
   theme_minimal() +
@@
             by = "OBJECTID") %>%
-  st_as_sf()
+  st_as_sf() %>% 
+  left_join(foodbank_locations %>% st_drop_geometry() %>% transmute(OBJECTID = ID, TYPE), by = "OBJECTID")
 
@@
                                    -0.0002,
-                                   -0.0003,
                                    -0.0004,
-                                   -0.0005),
+                                   -0.0006,
+                                   -0.0008),
                         labels = c("-0.0001",
                                    "-0.0002",
-                                   "-0.0003",
                                    "-0.0004",
-                                   "-0.0005")) +
+                                   "-0.0006",
+                                   "-0.0008")) +
   scale_size(trans = 'reverse') +
   scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  guides(size = FALSE, shape = FALSE) +
+  guides(size = "none", shape = "none") +
   theme_minimal() +
@@
                                    -0.0002,
-                                   -0.0003,
                                    -0.0004,
-                                   -0.0005),
+                                   -0.0006,
+                                   -0.0008),
                         labels = c("-0.0001",
                                    "-0.0002",
-                                   "-0.0003",
                                    "-0.0004",
-                                   "-0.0005")) +
+                                   "-0.0006",
+                                   "-0.0008")) +
   scale_size(name = "Change in level of service",
@@
                         -0.0002,
-                        -0.0003,
                         -0.0004,
-                        -0.0005),
+                        -0.0006,
+                        -0.0008),
              labels = c("-0.0001",
                         "-0.0002",
-                        "-0.0003",
                         "-0.0004",
-                        "-0.0005")) +
+                        "-0.0006",
+                        "-0.0008")) +
   scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
@@
          #size = FALSE, 
-         shape = FALSE) +
+         shape = "none") +
   theme_minimal() +
@@
                        direction = -1,
-                       breaks = c(0.002,
-                                  0.0016,
+                       breaks = c(0.0030,
+                                  0.0024,
+                                  0.0018,
                                   0.0012,
-                                  0.0008,
-                                  0.0004),
-                       labels = c("0.0020",
-                                  "0.0016",
+                                  0.0006),
+                       labels = c("0.0030",
+                                  "0.0024",
+                                  "0.0018",
                                   "0.0012",
-                                  "0.0008",
-                                  "0.0004")) +
+                                  "0.0006")) +
   theme_minimal() +
@@
                        direction = -1,
-                       breaks = c(0.0010,
-                                  0.0008,
+                       breaks = c(0.0024,
+                                  0.0018,
+                                  0.0012,
                                   0.0006,
-                                  0.0004,
-                                  0.0002),
-                       labels = c("0.0010",
-                                  "0.0008",
+                                  0.0001),
+                       labels = c("0.0024",
+                                  "0.0018",
+                                  "0.0012",
                                   "0.0006",
-                                  "0.0004",
-                                  "0.0002")) +
+                                  "0.0001")) +
   theme_minimal() +
@@
 
-We find that the accessibility landscape deteriorated substantially during the pandemic, with accessibility dropping on average by almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% mean() %>% abs() %>% round()`\%, but with large variations: some zones experienced changes in accessibility of only about `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% max() %>% abs() %>% round()`\%, whereas the most affected zone saw a loss of accessibility of almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% min() %>% abs() %>% round()`\%.
+We find that the accessibility landscape deteriorated substantially during the pandemic, with accessibility dropping on average by almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% mean() %>% abs() %>% round()`%, but with large variations: some zones experienced changes in accessibility of only about `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% max() %>% abs() %>% round()`%, whereas the most affected zone saw a loss of accessibility of almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% min() %>% abs() %>% round()`%.
 Figure \ref{fig:accessibility-changes-with-local-i} shows the changes in accessibility.
@@
 
-```{r plot-accessibility-changes, echo=FALSE, include=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19."}
+```{r plot-accessibility-changes, echo=FALSE,  fig.align = 'center', fig.cap="\\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19."}
 ## Analysis
@@
                                 "sig" = 1)) +
-  guides(color = FALSE, 
-         alpha = FALSE,
+  guides(color = "none", 
+         alpha = "none",
          fill = guide_legend(title.position = "top")) +
@@
 For the more suburban clusters of zones, the decrease in accessibility is derived from the closure of locations throughout the city reachable by car.
-In the cluster of central suburban zones for example, low-income households in the outer ring of zones that exhibit medium to high decreases in accessibility within this cluster appear to be largely auto-dependent in their tripmaking, which each each exhibiting between `r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(min(p_car)) *100)`-`r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(max(p_car)) *100)`\% of their modal split for car trips.
+In the cluster of central suburban zones for example, low-income households in the outer ring of zones that exhibit medium to high decreases in accessibility within this cluster appear to be largely auto-dependent in their tripmaking, which each each exhibiting between `r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(min(p_car)) *100)`-`r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(max(p_car)) *100)`% of their modal split for car trips.
 This results in the parcels within these zones having a large number of potentially accessible locations in the travel time matrix.
@@
 The zone with the greatest decrease in accessibility within this cluster (-0.0009) has a high rate of car trips and connects to the most food bank locations in total as well as those that stayed open or closed.
-The exception to this is the zone in the middle of this cluster, which has a much higher proportion of transit use (`r round(modes %>% st_drop_geometry() %>% filter(TAZUID == 5116) %>% summarize(first(p_transit)) *100)`\%).
-This modal split weighting results in fewer food bank and service locations accessible from the parcels within this zone which, by extension, limits their access to closed locations reachable within 15 minutes in our calculations.
 
 In the cluster to the south-west, the decrease in accessibility is predominately driven by the closure of a high level-of-service Community Meals provider.
-However, like the more central suburban zones, low-income households within this cluster are also between 90\% to 100\% auto-dependent in their tripmaking in the TTS.
+However, like the more central suburban zones, low-income households within this cluster are also between 90% to 100% auto-dependent in their tripmaking in the TTS.
 The story is similar for the zone located in the north-west that exhibits the greatest decrease in accessibility.
-Here, low-income households responding to the TTS conducted 100\% of their trips by car and, as a result, dwellings within this zone have access to the second-highest number of food bank and service locations within 15 minutes.
+Here, low-income households responding to the TTS conducted 100% of their trips by car and, as a result, dwellings within this zone have access to the second-highest number of food bank and service locations within 15 minutes.
 However, this also means zonal accessibilities are greatly affected by the number of closures throughout the city.
 Finally, in the city's north end and north-east zones, low-income households exhibit a mixture of tripmaking behaviour in the TTS.
-Households in some zones take transit more often and one zone in particular has 100\% of its trips by walking while in others, car use makes up 100\% of trips.
-For these zones, the decrease in accessibility tends to be a product of the closure of several inner-city food bank and service locations.
+Households in some zones take transit more often and one zone in particular has 100% of its trips by walking.
+For these zones, the decrease in accessibility tends to be a product of the closure of several inner-city food bank and service locations reachable by multiple modes.
 
@@
 Using data on low income individuals by age drawn from the TTS, Table \ref{tab:accessibility-by-age} shows the estimated number of people in each age group by their level of accessibility before and during the pandemic.
-Here, it is important to note that the quartiles are relative: people in the top 25\% of accessibility still have lower accessibility during the pandemic than before.
+Here, it is important to note that the quartiles are relative: people in the top 25% of accessibility still have lower accessibility during the pandemic than before.
 In reality, every population group is worse off during the pandemic in terms of their accessibility to food banks and services in the City of Hamilton.
@@
 accessibility_by_age %>%
-  kable("html",
+  kable("latex",
         booktabs = TRUE,
@@
         col.names = c("Accessibility Quartile",
-                      "Children (age <= 18)", 
+                      "Children (age $\\le$ 18)", 
                       "Adults (19-64)", 
-                      "Seniors (age >= 65)", 
-                      "Children (age <= 18)", 
+                      "Seniors (age $\\ge$ 65)", 
+                      "Children (age $\\le$ 18)", 
                       "Adults (19-64)", 
-                      "Seniors (age >= 65)"),
+                      "Seniors (age $\\ge$ 65)"),
         caption = "\\label{tab:accessibility-by-age}Population at each accessibility level by age group among members households with incomes less than CAD40,000.",
         escape = FALSE) %>%
-  kable_styling(bootstrap_options = c("striped")) %>%
+  kable_styling(latex_options = c("striped", "scale_down", "hold_position")) %>%
   add_header_above(c(" " = 1, "Pre-COVID-19" = 3, "During COVID-19" = 3)) %>%
@@
 accessibility_by_age <- accessibility %>%
-  filter(timing = "During covid")
+  filter(timing == "During covid") %>% 
   group_by(timing, 
@@
 
-Conclusions
-===================
+# Conclusions
 
@@
 More broadly, it points to the absurd need to understand how a bad situation was made worse by the pandemic: in effect, the analysis reveals that disparities in accessibility to food banks and services predated the pandemic, that the pandemic contributed to the deterioration of these services, and that populations already in distress, particularly children, ended up in an even more adverse state.
-How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding what is an _acceptable_ or _sufficient_ level of service when it comes to food banks.
+How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding what is an *acceptable* or *sufficient* level of service when it comes to food banks.
 Whether this is one facility per 1,000 population or three facilities per 1,000 population, or any other number, we know not.
@@
 
-References {#references .unnumbered}
-==========
+# References {#references .unnumbered}

### 2021-08-27 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex

@@
             pdfauthor={},
-            pdftitle={Changes in accessibility to food banks and food services during COVID-19 and implications for low income populations in Hamilton, Ontario},
+            pdftitle={Changes in accessibility to emergency and community food services during COVID-19 and implications for low income populations in Hamilton, Ontario},
             colorlinks=false,
@@
 % Pandoc citation processing
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
 \newlength{\csllabelwidth}
 \setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
 % for Pandoc 2.8 to 2.10.1
@@
 % For Pandoc 2.11+
-\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
  {% don't indent paragraphs
@@
  {}
-\usepackage{calc} % for calculating minipage widths
+\usepackage{calc}
 \newcommand{\CSLBlock}[1]{#1\hfill\break}
 \newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 
-  \title{Changes in accessibility to food banks and food services during
-COVID-19 and implications for low income populations in Hamilton,
-Ontario}
+  \title{Changes in accessibility to emergency and community food
+services during COVID-19 and implications for low income populations in
+Hamilton, Ontario}
     \author[Some University]{Author One}
@@
   \begin{abstract}
-  In this paper we analyze the changes in accessibility to food banks
-  and related services before and during the COVID-19 pandemic in the
-  City of Hamilton, Ontario. Food banks and services are the last line
-  of support for households facing food insecurity; as such, their
+  In this paper we analyze the changes in accessibility to community and
+  emergency food services before and during the COVID-19 pandemic in the
+  City of Hamilton, Ontario. Many of these food services are the last
+  line of support for households facing food insecurity; as such, their
   relevance cannot be ignored in the midst of the economic upheaval
   caused by the pandemic. Our analysis is based on the application of
-  balanced floating catchment areas, and concentrates on households with
+  balanced floating catchment areas and concentrates on households with
   lower incomes (\textless CAD40,000, approximately the Low Income
@@
 seems to be the case for food insecurity as well (Laborde et al., 2020).
-In the US, for example, it has been estimated that there was an increase
-of more than 30\% in household food insecurity, and more than one third
-of households were discovered to be newly food insecure - meaning they
-did not experience food insecurity before the pandemic (Niles et al.,
-2020). In Canada, Men and Tarasuk (2021) report that about 25\% of
-individuals who experienced job insecurity (a relatively common
-occurrence during the pandemic), also experienced food insecurity.
-Similarly, according to Statistics Canada (2020a), in the early stages
-of the pandemic almost 15\% of individuals reported living in a
-household that faced food insecurity; the risk of food insecurity was
-substantially higher for households with children. The difference
-between households with and without children was significant, and 11.7\%
-of households with children indicated that ``food didn't last and
-{[}there was{]} no money to get more'' sometimes or often, compared to
-7.3\% of households without children; likewise, 13\% of households with
-children indicated that they ``{[}c{]}ouldn't afford balanced meals''
-sometimes or often, compared to 8.8\% of households without children.
+According to Statistics Canada (2020a), in the early stages of the
+pandemic almost 15\% of individuals reported living in a household that
+faced food insecurity; the risk of food insecurity was substantially
+higher for households with children. The difference between households
+with and without children was significant, and 11.7\% of households with
+children indicated that ``food didn't last and {[}there was{]} no money
+to get more'' sometimes or often, compared to 7.3\% of households
+without children; likewise, 13\% of households with children indicated
+that they ``{[}c{]}ouldn't afford balanced meals'' sometimes or often,
+compared to 8.8\% of households without children. Additionally, Men and
+Tarasuk (2021) report that about 25\% of individuals who experienced job
+insecurity (a relatively common occurrence during the pandemic) also
+experienced food insecurity associated with COVID-related disruptions to
+employment, financial hardship, and use of food charity.
 
@@
 were already critical public health concerns in Canada prior to COVID-19
-(Boucher et al., 2017). While food banks are not necessarily a stable
-solution to food insecurity and in fact may encourage a retrenchment of
-neoliberal policy (Wakefield et al., 2013), at least they can be argued
-to provide a resource of last instance to households in precarious
-situations (Bazerghi et al., 2016). As a mid-size city grappling with
-deindustrialization, Hamilton exhibits high rates of poverty and food
-bank use. As recently as 2019, the Hamilton Hunger Report\footnote{https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf}
-noted that food banks in the city recorded the highest number of
-visitors in the past 29 years; a rate of increase greater than
-population growth. Most troubling, approximately 40\% of all visitors
-were children.
+(Boucher et al., 2017). While emergency food services are not
+necessarily a stable solution to food insecurity and in fact may
+encourage a retrenchment of neoliberal policy (Wakefield et al., 2013),
+at least they can be argued to provide a resource of last instance to
+households in precarious situations (Bazerghi et al., 2016). As a
+mid-size city grappling with deindustrialization, Hamilton exhibits high
+rates of poverty and food bank use. As recently as 2019, the Hamilton
+Hunger Report (HFS, 2019) noted that food banks in the city recorded the
+highest number of visitors in the past 29 years; a rate of increase
+greater than population growth. Most troubling, approximately 40\% of
+all visitors were children.
 
@@
 accessibility to healthy food through the ``food desert'' concept, there
-has been little research into accessibility to food banks. Previous work
-has explored differences in \emph{accessing} food banks, such as how
-some households utilize food banks over short periods of time while
-others regularly utilize food banks as longer-term resource (e.g., Enns
-et al., 2020). In addition, transportation and locational considerations
-have been raised as key issues in food bank accessibility in previous
-qualitative research (Smith-Carrier et al., 2017). However, we are not
-aware of any research that has focused on estimating or capturing this
-geographic component of accessibility.
+has been little research into accessibility to emergency and community
+food services. Previous work has explored differences in
+\emph{accessing} food banks, such as how some households utilize food
+banks over short periods of time while others regularly utilize food
+banks as longer-term resource (e.g., Enns et al., 2020). In addition,
+transportation and locational considerations have been raised as key
+issues in food bank accessibility in previous qualitative research
+(Smith-Carrier et al., 2017). However, we are not aware of any research
+that has focused on estimating or capturing this geographic component of
+accessibility.
 
@@
 transportation network (Páez et al., 2012). Indeed, the Government of
-Canada's recent Food Policy has made ``access'' to healthy food a
-priority for Canadian communities\footnote{https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042}
-and previous survey research has suggested that such accessibility plays
-a key role in user satisfaction with food bank service delivery (Holmes
-et al., 2018). However, as with research into the prevalence of food
-deserts, accessibility to food banks is unlikely to be evenly
-distributed, and variation throughout a city can be expected due to
-transportation network characteristics, and the spatial distribution of
-food bank locations and the population they are meant to serve.
-Furthermore, policy responses to the COVID-19 pandemic likely have added
-to the distress of vulnerable households. Non-pharmaceutical
+Canada's recent Food Policy (Agriculture and Canada, 2019) has made
+``access'' to healthy food a priority for Canadian
+communities{[}\^{}2{]} and previous survey research has suggested that
+such accessibility plays a key role in user satisfaction with food bank
+service delivery (Holmes et al., 2018). However, as with research into
+the prevalence of food deserts, accessibility to food banks is unlikely
+to be evenly distributed, and variation throughout a city can be
+expected due to transportation network characteristics and the spatial
+distribution of service locations and the population they are meant to
+serve. Furthermore, policy responses to the COVID-19 pandemic likely
+have added to the distress of vulnerable households. Non-pharmaceutical
 interventions during the pandemic involving restrictions in mobility
@@
 accessibility to food banks before the pandemic from the perspective of
-low income households, and how it changed during the pandemic? And
+low income households, and how has it changed during the pandemic with
+respect to geographic access and congestion at remaining sites? And
 finally, who are most likely to have been impacted by changes in the
@@
 2021). The code and data necessary to reproduce the analysis are
-available in a public repository\footnote{https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j\_M\_BMrXhXFvrV2c?usp=sharing}.
+available in a public repository (see Supplemental Material).
 
@@
 Food insecurity is the inability to acquire and consume an adequate
-amount or good quality food, leading to inadequate nutrient intake (Enns
-et al., 2020; Kirkpatrick and Tarasuk, 2008; Tarasuk and Vogt, 2009).
-This nutrient deficiency is associated with major population health
-concerns, particularly among Canadians at socio-economic disadvantage
+amount or good quality food, leading to inadequate nutrient intake
+(Kirkpatrick and Tarasuk, 2008) and poorer physical and mental health
+outcomes (Ramsey et al., 2011; Seligman et al., 2010; Stuff et al.,
+2004). In this regard, food insecurity is a major population health
+concern, particularly among Canadians at socio-economic disadvantage
 (Bazerghi et al., 2016). Official government surveys such as the
@@
 insecurity in Canada (Gundersen et al., 2018; Kirkpatrick and Tarasuk,
-2008; Tarasuk and Vogt, 2009). Nationally, it has been found that food
-insecurity impacts approximately 12.3\% of Canadian households (Tarasuk
-et al., 2014)
+2008; Tarasuk and Vogt, 2009).
+
+Nationally, it has previously been found that food insecurity impacts
+approximately 12.3\% of Canadian households (Tarasuk et al., 2014) using
+data from the 2011-2012 CCHS. Using the same data, Tarasuk et al. (2019)
+found higher odds of food insecurity amongst households relying on
+social assistance, those without a university degree or with children
+under the age of 18, and individuals that lived alone, renters, and
+those identifying as Aboriginal. While surveys revealed that only 20 to
+30 percent of those experiencing food insecurity were found to frequent
+food banks in Canada (Tarasuk et al., 2014), pre-pandemic research from
+Ottawa (Enns et al., 2020) and Vancouver (Black and Seto, 2020) suggests
+that long-term users tend to be older, have health or mobility
+challenges, live in large households, and are less likely to have
+employment income. In terms of geography, previous research conducted at
+the provincial scale using data from the 2011-2012 CCHS found that the
+prevalence of food insecurity ranged across the country from 11.8\% of
+households in Ontario to 41\% of households in Nunavut (Tarasuk et al.,
+2019).
 
@@
 inadequate food by voluntarily offering them meals and ingredients
-(Loopstra and Tarasuk, 2012; Riches, 2002). The scope and objectives of
-food banks can vary by region and by country, and these organizations
-can include not only prepared meals and aliments, but also shared spaces
-to connect in community gardens and community kitchens (Wakefield et
-al., 2013). Although in their origin food banks were meant to be provide
-a temporary solution to accommodate those in hunger due to job
-retrenchments and economic downfalls since the 1980s, over time they
-have evolved into a community practice to secure emergency food supplies
-for those in need (Loopstra and Tarasuk, 2012; Wakefield et al., 2013).
-
-In Canada, the number of food banks has steadily increased in the past
-few decades (Wakefield et al., 2013). The largest database of food banks
-and their use comes from the non-profit association Food Banks Canada
-(FBC), which conducts an annual assessment through its affiliated
-members. FBC's 2018 Hunger Count report\footnote{https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018\_FINAL\_EN.pdf.aspx?ext=.pdf}
-(the most recent available) listed 1,830 member food banks across the
-country, and found that Canadians visited food banks 1.1 million times
-in March of 2018. Of those accessing food banks, certain population
-characteristics tend to be over-represented compared to national totals
-from the 2016 Canadian Census of Population. According to FBC's 2018
-data, single-adult households represent 45\% of those utilizing food
-banks despite making up 28\% of Canada's population, 19\% are
-single-parent households (compared to 10\% nationally), and 35\% of
-those accessing food bank services are children aged 0-18 even though
-their share of Canada's national population is approximately 20\%. In
+(Loopstra and Tarasuk, 2012; Riches, 2002). Although in their origin
+food banks were meant to be provide a temporary solution to accommodate
+those in hunger due to job retrenchments and economic downfalls since
+the 1980s, over time many have evolved into a community practice to
+secure food supplies for those in need (Loopstra and Tarasuk, 2012;
+Wakefield et al., 2013). In Canada, the number of food banks has
+steadily increased in the past few decades (Wakefield et al., 2013). The
+largest database of food banks and their use comes from the non-profit
+association Food Banks Canada (FBC), which conducts an annual assessment
+through its affiliated members. FBC's 2018 Hunger Count report (FBC,
+2018) (the most recent available) listed 1,830 member food banks across
+the country, and found that Canadians visited food banks 1.1 million
+times in March of 2018. Of those accessing food banks, certain
+population characteristics tend to be over-represented compared to
+national totals from the 2016 Canadian Census of Population. According
+to FBC's 2018 data, single-adult households represent 45\% of those
+utilizing food banks despite making up 28\% of Canada's population, 19\%
+are single-parent households (compared to 10\% nationally), and 35\% of
+those using food bank services are children aged 0-18 even though their
+share of Canada's national population is approximately 20\%. In
 addition, 59\% of households accessing food banks list social or
-disability assistance as their primary source of income. Similarly,
-using data from the 2011-2012 CCHS, Tarasuk et al. (2019) found higher
-odds of food insecurity amongst households relying on social assistance,
-those without a university degree or with children under the age of 18,
-and individuals that lived alone, renters, and those identifying as
-Aboriginal. While surveys revealed that only 20 to 30 percent of those
-experiencing food insecurity were found to frequent food banks in Canada
-(Tarasuk et al., 2014), research from Ottawa (Enns et al., 2020) and
-Vancouver (Black and Seto, 2020) suggests that long-term users tend to
-be older, have health or mobility challenges, live in large households,
-and are less likely to have employment income.
+disability assistance as their primary source of income.
 
@@
 Most recently in May 2020, Canada recorded 14.7\% of its population
-living in food insecurity in the past 30 days (Statistics Canada,
-2020a). Considering the negative mental and physical health effects,
-increases in food insecurity rates due to the onset of COVID-19, signal
-a change in the food environment with potential damages to population
-health during the course of the pandemic and beyond (Niles et al.,
-2020). Recent data\footnote{https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report\_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis\_Exec-Sum\_EN.pdf}
-from FBC showed that 52\% of member food banks reported an increase in
-usage in March of 2020 when initial lockdown restrictions were put in
-place across much of the country. The pandemic also created significant
-staffing issues with 42\% of food banks reporting a reduction in
-volunteers. However, 53\% of food banks later reported a decrease in use
-into the summer of 2020 which FBC members attributed to emergency
-financial support programs from the federal government. Nevertheless,
-some of these benefit programs were temporary. Although more recent
-statistics on food bank use in Hamilton in 2020 and 2021 are not yet
-available, data from the Daily Bread Food Bank in neighbouring Toronto
-for August 2020 shows visits climbing 51\% year-over-year in that
-city\footnote{https://www.dailybread.ca/wp-content/uploads/2020/11/Whos-Hungry-Report-2020.pdf},
-which suggests that many households in Hamilton are likely to turn to
-food bank services to meet their needs.
-
-In terms of geography, previous research conducted at the provincial
-scale using data from the 2011-2012 CCHS found that the prevalence of
-food insecurity ranged across the country from 11.8\% of households in
-Ontario to 41\% of households in Nunavut (Tarasuk et al., 2019).
-However, although previous research has examined the characteristics of
-individuals and households accessing food banks, the locational or
-transportation accessibility aspect of food bank access is not well
-understood. A wealth of literature examining the food desert concept
-suggests that, in addition to socio-economic and demographic factors,
-location and transportation networks play a key role in a household's
-accessibility to healthy foods (Paez et al., 2010; Vanderlee and L'Abbé,
-2017; Widener, 2018). For food banks specifically, previous qualitative
-research in Ontario by Smith-Carrier et al. (2017) has noted that
-``transportation can be challenging, particularly if the food bank is
-situated in a remote location'' (p.~32). Particularly, it appears that
-participants experience challenges with the ``inordinate amount of time
-necessary to obtain food, and difficulties associated with
+living in food insecurity in the past 30 days, up from 10.5\% in
+2017-2018 (Statistics Canada, 2020a). Considering the negative mental
+and physical health effects, increases in food insecurity rates due to
+the onset of COVID-19, signal a change in the food environment with
+potential damages to population health during the course of the pandemic
+and beyond (Niles et al., 2020). Recent data from FBC (FBC, 2020) showed
+that 52\% of member food banks reported an increase in usage in March of
+2020 when initial lockdown restrictions were put in place across much of
+the country. The pandemic also created significant staffing issues with
+42\% of food banks reporting a reduction in volunteers. However, 53\% of
+food banks later reported a decrease in use into the summer of 2020
+which FBC members attributed to emergency financial support programs
+from the federal government. Nevertheless, some of these benefit
+programs were temporary. Although more recent statistics on food bank
+use in Hamilton in 2020 and 2021 are not yet available, data from the
+Daily Bread Food Bank (DBFB, 2020) in neighbouring Toronto for August
+2020 shows visits climbing 51\% year-over-year in that city, which
+suggests that many households in Hamilton are likely to turn to food
+bank services to meet their needs.
+
+Beyond traditional conceptualizations of food banks as providers of
+emergency food assistance, other community food services also play an
+important role in decreasing food insecurity. The scope and objectives
+of food banks can vary by region and by country, and these organizations
+can include not only prepared meals and aliments for emergency food
+supply, but also shared spaces to connect in community gardens and
+community kitchens (Wakefield et al., 2013). However, the efficacy of
+these programs in reducing food insecurity differs by the type of
+service offered. For example, previous qualitative research in the
+Toronto region has questioned the capacity of community kitchens to
+improve the food security of low-income households due to their limited
+scale of operations, and un-subsidized kitchens were found to be
+particularly inaccessible to to families living in severe poverty
+(Tarasuk and Reynolds, 1999). However, other food access options such as
+no-cost or low-cost meals provided through community meals or congregate
+dining play an important role in decreasing food insecurity. Research in
+Minnesota found that seniors experiencing food insecurity valued
+congregate dining for providing affordable meals and a space for social
+gathering (Oemichen and Smith, 2016). Furthermore, because seniors paid
+for the meal, there was no stigma attached to the use of these services
+compared to traditional food-purchasing assistance such as the
+Supplemental Nutrition Assistance Program.
+
+While previous research has examined the characteristics of individuals
+and households accessing emergency and community food services, the
+locational or transportation accessibility aspect of food bank access is
+not well understood. A wealth of literature examining the food desert
+concept suggests that, in addition to socio-economic and demographic
+factors, location and transportation networks play a key role in a
+household's accessibility to healthy foods (Paez et al., 2010; Vanderlee
+and L'Abbé, 2017; Widener, 2018). For food banks specifically, previous
+qualitative research in Ontario by Smith-Carrier et al. (2017) has noted
+that ``transportation can be challenging, particularly if the food bank
+is situated in a remote location'' (p.~32). Particularly, it appears
+that participants experience challenges with the ``inordinate amount of
+time necessary to obtain food, and difficulties associated with
 transportation'' (p.~39). Users of food banks, according to this
@@
 
-More formally, the first step of this method is as follows: \[
+More formally, the first step of this method is as follows:
+
+\[
 L_j = \frac{S_j}{\sum_{i=1}^nP_iw_{ij}}
@@
 that each population unit can reach, according to the distance-decay
-weight: \[
+weight:
+
+\[
 A_i = \sum_{j=1}^JL_jw_{ji}
@@
 \emph{proportionally}. This is achieved by standardizing the weights as
-follows: \[
+follows:
+
+\[
 w_{ij}^\text{st} = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
-\] \noindent and: \[
+\]
+
+\noindent and:
+
+\[
 w_{ji}^\text{st} = \frac{w_{ji}}{\sum_{j=1}^Jw_{ji}}
@@
 
-The standardized weights satisfy the following conditions: \[
+The standardized weights satisfy the following conditions:
+
+\[
 \sum_{i=1}^nw_{ij}^\text{st}=1
@@
 
-\noindent and: \[
+\noindent and:
+
+\[
 \sum_{j=1}^Jw_{ji}^\text{st}=1
@@
 Since the population is allocated proportionally, its value is
-preserved: \[
+preserved:
+
+\[
 \sum_{i=1}^nP_iw_{ij}^\text{st}=P_i
@@
 
-\noindent as is the level of service: \[
+\noindent as is the level of service:
+
+\[
 \sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
@@
 residents living at or below Canada's poverty level (DeLuca et al.,
-2012; Jakar and Dunn, 2019; Latham and Moffat, 2007). In 2016, the
-Hamilton Community Foundation reported that based on the Low-Income
-Cut-Off, Hamilton recorded a poverty rate of 16.7\%, which was well
-above the average rate of Ontario (13.7\%) and the average national rate
-(12.8\%).\footnote{https://www.hamiltoncommunityfoundation.ca/vital-signs/low-income-2018/}.
-According to data from Hamilton Food Share\footnote{https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf},
+2012; Jakar and Dunn, 2019; Latham and Moffat, 2007). The Hamilton
+Community Foundation (HCF, 2018) reported that based on the Low-Income
+Cut-Off, Hamilton recorded a poverty rate of 16.7\% in 2016, which was
+well above the average rate of Ontario (13.7\%) and the average national
+rate (12.8\%). According to data from Hamilton Food Share (HFS, 2019),
 approximately 23,000 individuals accessed food banks in the city in
@@
 on a per-capita basis, the level of need in the inner core of central
-Hamilton was second highest in Ontario\footnote{https://feedontario.ca/research/report-hunger-map/}.
+Hamilton was second highest in Ontario (FO, 2019).
 
@@
 
-Data have been prepared for sharing as a data package\footnote{https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j\_M\_BMrXhXFvrV2c?usp=sharing}.
-The contents of the data package are described next.
+Data have been prepared for sharing as a data package available in the
+Supplemental Materials. The contents of the data package are described
+next.
 
@@
 
-The locations of food banks and related food services were obtained from
-the Hamilton Public Library's Food Access Guide\footnote{http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf}.
-The guide was updated in April of 2021 to indicate any change affected
-on the services due to the pandemic. This includes modified business
-hours, a need to make reservations before frequenting, and locations
-that have completely shut down in consequence. Table 1 defines each
-service type and the number of locations pre- and during the COVID-19
-pandemic. While some food bank services have a specific a target
-population, such as prioritizing families with young children aged
-between 0 and 3 or accepting only those providing proof of low-income
-status through housing and utility statements, all the food bank
-services indicated below are designed to accommodate those in need of
-food at zero to low cost. With our focus on food banks and food services
-that offer free or low-cost meals at particular locations, we first
-removed services such as Meals on Wheels and other food access services
-such as food box, community kitchens, student nutrition programs, and
-shopping and transportation. With some providers offering different food
-services at the same location (e.g.~food bank with free and community
-meal services), and some of these services closing after the onset of
-the COVID-19 pandemic, we opted to geocode based on the service type. On
-the other hand, two free meal services held on different days at the
-same location were collapsed into a single service point for the
+The locations of emergency and community food services were obtained
+from the Hamilton Public Library's Food Access Guide (HPL, 2021). The
+guide was updated in April of 2021 to indicate any change affected on
+the services due to the pandemic. This includes modified business hours,
+a need to make reservations before frequenting, and locations that have
+completely shut down in consequence. Table 1 defines each service type
+and the number of locations pre- and during the COVID-19 pandemic. While
+some food bank services have a specific target population, such as
+prioritizing families with young children aged between 0 and 3 or
+accepting only those providing proof of low-income status through
+housing and utility statements, all the food services indicated below
+are designed to accommodate those in need of food at zero to low cost.
+With our focus on food banks and food services that offer free or
+low-cost meals at particular locations, we first removed services such
+as Meals on Wheels and other food access services such as food box,
+community kitchens, student nutrition programs, and shopping and
+transportation. With some providers offering different food services at
+the same location (e.g.~food bank with free and community meal
+services), and some of these services closing after the onset of the
+COVID-19 pandemic, we opted to geocode based on the service type. On the
+other hand, two free meal services held on different days at the same
+location were collapsed into a single service point for the
 accessibility analysis. Additional details on the operations of
-individual facilities is not publicly available, and with the changes in
+individual facilities is not publicly available and with the changes in
 operations it proved unfeasible to collect it. For this reason, the
@@
 Transportation may be provided} & \cellcolor{gray!6}{7} & \cellcolor{gray!6}{2} & \cellcolor{gray!6}{One remaining location reduced hours during COVID}\\
-Community Meals & Programs often run by volunteers that organize suppers, lunches or other get-togethers that give community residents an opportunity to meet one another in a friendly and informal atmosphere while sharing a meal & 11 & 9 & NA\\
-\cellcolor{gray!6}{Food Banks} & \cellcolor{gray!6}{Food Banks and Emergency Food programs provide individuals and families with grocery items free of charge} & \cellcolor{gray!6}{27} & \cellcolor{gray!6}{26} & \cellcolor{gray!6}{One remaining location reduced hours during COVID while 4 others moved to appointment only}\\
+Community Meals & No-cost programs often run by volunteers that organize suppers, lunches or other get-togethers that give community residents an opportunity to meet one another in a friendly and informal atmosphere while sharing a meal & 11 & 9 & NA\\
+\cellcolor{gray!6}{Food Banks} & \cellcolor{gray!6}{Food Banks and Emergency Food programs provide individuals and families with grocery items free of charge} & \cellcolor{gray!6}{26} & \cellcolor{gray!6}{25} & \cellcolor{gray!6}{One remaining location reduced hours during COVID while 4 others moved to appointment only}\\
 Free Meals & Meals are provided free of charge in the community through volunteer labour and donations & 9 & 5 & One remaining location reduced hours during COVID\\
@@
 Travel time tables for three modes (car, transit, walking) were computed
-using the parcels as the origins and the locations of the food banks as
-the destinations. For routing, the package \texttt{r5r} (Pereira et al.,
-2021b) was used with a network extract for the City of Hamilton from
-OpenStreetMaps and the General Transit Feed Specification from Hamilton
-Street Railway, the local transit operator. For routing purposes we used
-maximum values of 180 min and 10,000 m walking distance: any destination
-that exceeded these thresholds was ignored. The departure time used for
-routing was 8:00AM on March 30, 2021 to reflect transit service around
-the morning service peak on a typical Tuesday.
+using the parcels as the origins and the locations of the community and
+emergency food service locations as the destinations. For routing, the
+package \texttt{r5r} (Pereira et al., 2021b) was used with a network
+extract for the City of Hamilton from OpenStreetMaps and the General
+Transit Feed Specification (GTFS) files for the Hamilton Street Railway,
+the local transit operator, as well as for Burlington Transit, which
+operates some service in the city. For transit routing purposes we used
+maximum travel time values of 300 min and a 2,000 m cap on walking
+distance: any destination that exceeded these thresholds was ignored.
+The departure time used for routing was 8:00AM on March 30, 2021 to
+reflect transit service around the morning service peak on a typical
+Tuesday.
 
@@
 travel time \(ett\) from each parcel \(i\) to a food bank or food
-service location \(j\) as follows: \[
+service location \(j\) as follows:
+
+\[
 ett_{ij} = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
@@
 parcel \(i\) to the food bank. In other words, the expected travel time
-is the weighted sum of travel times to the food bank, with the weights
-given by the expected modal split of trips made by low-income households
-in the TAZ per the TTS data.
+reflects the weighted average of travel times to the food bank, with the
+weights given by the expected modal split of trips made by low-income
+households in the TAZ per the TTS data.
 
@@
 literature, which is mostly silent on the travel patterns of people who
-visit food banks. For this reason, we opt for a simple cumulative
-opportunities function as follows: \[
+visit food banks and community food services. For this reason, we opt
+for a simple cumulative opportunities function as follows:
+
+\[
 w_{ij}=w_{ji}=
@@
 \end{cases}
-\] \noindent where \(ett_{ij}\) is the multimodal expected travel time
-as described previously, and \(\delta\) is a travel threshold. When the
+\]
+
+\noindent where \(ett_{ij}\) is the multimodal expected travel time as
+described previously, and \(\delta\) is a travel threshold. When the
 expected travel time exceeds this threshold, a facility is no longer
@@
 Using the 15-minute threshold, we find that the system-wide
-accessibility (interpreted as a provider-to-population ratio) was 0.065
+accessibility (interpreted as a provider-to-population ratio) was 0.078
 (food banks/service locations per low income household in the city)
-before COVID-19, but declined to 0.037 during the pandemic. It is
+before COVID-19, but declined to 0.048 during the pandemic. It is
 striking that although almost 76\% of food facilities remained in
 operation during the pandemic, there was a loss of accessibility greater
-than 43\%, suggesting the location of food banks and related services
-plays an important role in serving those in need.
+than 39\%, suggesting the location of emergency and community food
+services plays an important role in serving those in need.
 
@@
 proximate service locations. Three more peripheral facilities towards
-the south of the city also have lower levels of service, presumably
-because they are expected to service relatively large suburban/exurban
+the south of the city have moderate levels of service, presumably
+because they are expected to service relatively suburban/exurban
 populations generally reliant on automobiles for travel. During the
@@
 is not uniform. The upper pane of Figure
-\ref{fig:levels-of-service-changes} shows that the three peripheral
-facilities in the southern suburban/exurban part of the city had low
-levels of service to begin with, but did not see major declines during
-the pandemic. Further, the inset map shows that the levels of service
-deteriorated more in the central part of the city. However, the loss of
-level of service was not as large in the core (where most of the food
-banks/services are found), but instead was more marked in the inner ring
-around the core, where facilities may have faced greater demand from
-both central city and suburban populations after the closure of food
-banks during the pandemic.
+\ref{fig:levels-of-service-changes} shows that the peripheral facilities
+in the suburban/exurban parts of the city saw major declines during the
+pandemic as more urban locations closed and demand increased for the
+remaining locations. Further, the inset map shows that the levels of
+service also deteriorated in the central part of the city. However, the
+loss of level of service was not as large in the core (where most of the
+food banks/services are found), but instead was more marked in the inner
+ring around the core, where facilities may have faced greater demand
+from both central city and suburban populations after the closure of
+service locations during the pandemic.
 
@@
 during the pandemic, with accessibility dropping on average by almost
-41\%, but with large variations: some zones experienced changes in
-accessibility of only about 10\%, whereas the most affected zone saw a
-loss of accessibility of almost 71\%. Figure
+38\%, but with large variations: some zones experienced changes in
+accessibility of only about 8\%, whereas the most affected zone saw a
+loss of accessibility of almost 96\%. Figure
 \ref{fig:accessibility-changes-with-local-i} shows the changes in
@@
 \begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-accessibility-changes-1}
+
+}
+
+\caption{\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19.}\label{fig:plot-accessibility-changes}
+\end{figure}
+
+\begin{figure}
 \includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-local-i-1} \caption{\label{fig:accessibility-changes-with-local-i}Changes in accessibility from pre-COVID-19 to during COVID-19.
@@
 of car trips and connects to the most food bank locations in total as
-well as those that stayed open or closed. The exception to this is the
-zone in the middle of this cluster, which has a much higher proportion
-of transit use (32\%). This modal split weighting results in fewer food
-bank and service locations accessible from the parcels within this zone
-which, by extension, limits their access to closed locations reachable
-within 15 minutes in our calculations.
+well as those that stayed open or closed.
 
@@
 in the TTS. Households in some zones take transit more often and one
-zone in particular has 100\% of its trips by walking while in others,
-car use makes up 100\% of trips. For these zones, the decrease in
-accessibility tends to be a product of the closure of several inner-city
-food bank and service locations.
+zone in particular has 100\% of its trips by walking. For these zones,
+the decrease in accessibility tends to be a product of the closure of
+several inner-city food bank and service locations reachable by multiple
+modes.
 
@@
 \midrule
-\cellcolor{gray!6}{Top 25\%} & \cellcolor{gray!6}{2,546} & \cellcolor{gray!6}{10,404} & \cellcolor{gray!6}{5,008} & \cellcolor{gray!6}{2,567 (0.82\%)} & \cellcolor{gray!6}{10,597 (1.86\%)} & \cellcolor{gray!6}{4,941 (-1.34\%)}\\
-Second 25\% & 3,333 & 12,916 & 6,415 & 3,375 (1.26\%) & 12,885 (-0.24\%) & 6,157 (-4.02\%)\\
-\cellcolor{gray!6}{Third 25\%} & \cellcolor{gray!6}{4,381} & \cellcolor{gray!6}{11,559} & \cellcolor{gray!6}{7,067} & \cellcolor{gray!6}{4,062 (-7.28\%)} & \cellcolor{gray!6}{11,412 (-1.27\%)} & \cellcolor{gray!6}{7,678 (8.65\%)}\\
-Bottom 25\% & 3,546 & 12,946 & 6,903 & 3,802 (7.22\%) & 12,897 (-0.38\%) & 6,567 (-4.87\%)\\
+\cellcolor{gray!6}{Top 25\%} & \cellcolor{gray!6}{2,048} & \cellcolor{gray!6}{7,031} & \cellcolor{gray!6}{3,632} & \cellcolor{gray!6}{1,956 (-4.49\%)} & \cellcolor{gray!6}{6,138 (-12.7\%)} & \cellcolor{gray!6}{3,338 (-8.09\%)}\\
+Second 25\% & 2,190 & 9,631 & 5,023 & 2,365 (7.99\%) & 10,833 (12.48\%) & 5,400 (7.51\%)\\
+\cellcolor{gray!6}{Third 25\%} & \cellcolor{gray!6}{3,239} & \cellcolor{gray!6}{12,694} & \cellcolor{gray!6}{6,875} & \cellcolor{gray!6}{3,082 (-4.85\%)} & \cellcolor{gray!6}{12,612 (-0.65\%)} & \cellcolor{gray!6}{7,064 (2.75\%)}\\
+Bottom 25\% & 3,711 & 13,220 & 7,648 & 3,038 (-18.14\%) & 11,811 (-10.66\%) & 6,308 (-17.52\%)\\
 \bottomrule
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-foodpolicy2019}{}%
+Agriculture, Canada, A.-F., 2019. Food policy for canada. Government of
+Canada.
+
 \leavevmode\hypertarget{ref-anselin1995local}{}%
@@
 
+\leavevmode\hypertarget{ref-dbfb2020}{}%
+DBFB, 2020. Who's hungry 2020. Daily Bread Food Bank.
+
 \leavevmode\hypertarget{ref-delamater2013spatial}{}%
@@
 
+\leavevmode\hypertarget{ref-fbc2018}{}%
+FBC, 2018. HungerCount 2018. Food Banks Canada.
+
+\leavevmode\hypertarget{ref-fbc2020}{}%
+FBC, 2020. A snapshot of food banks in canada and the COVID-19 crisis.
+Food Banks Canada.
+
+\leavevmode\hypertarget{ref-feedontario2019}{}%
+FO, 2019. Report: Hunger map. Feed Ontario.
+
 \leavevmode\hypertarget{ref-ghorbanzadeh2021spatial}{}%
@@
 canada. PloS one 13, e0202642.
+
+\leavevmode\hypertarget{ref-hcf2018}{}%
+HCF, 2018. Vital signs: A reflection of hamilton. Hamilton Community
+Foundation.
+
+\leavevmode\hypertarget{ref-hfs2019}{}%
+HFS, 2019. Hamilton hunger report 2019. Hamilton Food Share.
 
@@
 
+\leavevmode\hypertarget{ref-hpl2021}{}%
+HPL, 2021. Food access guide. Hamilton Public Library.
+
 \leavevmode\hypertarget{ref-jakar2019turning}{}%
@@
 12, 2096.
+
+\leavevmode\hypertarget{ref-oemichen2016investigation}{}%
+Oemichen, M., Smith, C., 2016. Investigation of the food choice,
+promoters and barriers to food access issues, and food insecurity among
+low-income, free-living minnesotan seniors. Journal of nutrition
+education and behavior 48, 397--404.
 
@@
 
+\leavevmode\hypertarget{ref-tarasuk1999qualitative}{}%
+Tarasuk, V., Reynolds, R., 1999. A qualitative study of community
+kitchens as a response to income-related food insecurity. Canadian
+Journal of Dietetic Practice and Research 60, 11.
+
 \leavevmode\hypertarget{ref-tarasuk2009household}{}%

### 2021-08-27 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Update Accessibility-Foodbanks-Hamilton.Rmd

--- 2021-04-28 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
@@
 ---
-title: "Changes in accessibility to food banks and food services during COVID-19 and implications for low income populations in Hamilton, Ontario"
+title: "Changes in accessibility to emergency and community food services during COVID-19 and implications for low income populations in Hamilton, Ontario"
 author:
@@
 abstract: |
-  In this paper we analyze the changes in accessibility to food banks and related services before and during the COVID-19 pandemic in the City of Hamilton, Ontario.
-Food banks and services are the last line of support for households facing food insecurity; as such, their relevance cannot be ignored in the midst of the economic upheaval caused by the pandemic.
-Our analysis is based on the application of balanced floating catchment areas, and concentrates on households with lower incomes (<CAD40,000, approximately the Low Income Cutoff Value for a city of Hamilton's size).
+  In this paper we analyze the changes in accessibility to community and emergency food services before and during the COVID-19 pandemic in the City of Hamilton, Ontario.
+Many of these food services are the last line of support for households facing food insecurity; as such, their relevance cannot be ignored in the midst of the economic upheaval caused by the pandemic.
+Our analysis is based on the application of balanced floating catchment areas and concentrates on households with lower incomes (<CAD40,000, approximately the Low Income Cutoff Value for a city of Hamilton's size).
 We find that accessibility was low to begin with in suburban and exurban parts of the city; furthermore, about 14\% of locations originally available in Hamilton closed during the pandemic, further reducing accessibility.
@@
 library(tidyverse)
-#library(tmap)
 ```
@@
 data(foodbank_locations)
-data(foodbank_table)
+#data(foodbank_table)
 data(hamilton_boundary)
@@
 data(ttm_car)
-data(ttm_transit)
+#data(ttm_transit)
 data(ttm_walk)
+
+# load updated files manually
+load("output-data-files/ttm_transit.RData")
+load("./output-data-files/foodbank_table.RData")
 ```
@@
 ```{r travel-time-tables-data-check, include=FALSE}
-## Data preparation
+## Data preparation - 2020
 # Check the NAs in travel time table:
@@
 # It should be fine to replace the NAs with zeros, which multiplied by p_transit will return a zero (i.e., it is not possible to reach a foodbank by transit).
-ttm <- ttm %>% 
-  mutate(travel_time_transit = replace_na(travel_time_transit, 0))
+#ttm_2021 <- ttm_2021 %>% 
+#  mutate(travel_time_transit = replace_na(travel_time_transit, 0))
 
@@
 # Like with transit, it should be fine to replace the NAs with zeros, which multiplied by p_walk will return a zero (i.e., it is not possible to reach a foodbank by walking).
-ttm <- ttm %>% 
-  mutate(travel_time_walk = replace_na(travel_time_walk, 0))
+#ttm_2021 <- ttm_2021 %>% 
+#  mutate(travel_time_walk = replace_na(travel_time_walk, 0))
 
@@
 
-```{r calculate-expected-travel-times, include=FALSE}
+```{r calculate-expected-travel-times-old, eval = FALSE, include=FALSE}
 ## Data preparation
@@
 The total expected travel time is the sum of the expected travel times by mode:
-ttm <- ttm %>%
-  mutate(e_travel_time_car = travel_time_car * p_car,
-         e_travel_time_transit = travel_time_transit * p_transit,
-         e_travel_time_walk = travel_time_walk * p_walk,
-         e_travel_time = e_travel_time_car + e_travel_time_transit + e_travel_time_walk)
-```
-
-Introduction
-==========================
+#ttm_2021 <- ttm_2021 %>%
+#  mutate(e_travel_time_car = travel_time_car * p_car,
+#         e_travel_time_transit = travel_time_transit * p_transit,
+#         e_travel_time_walk = travel_time_walk * p_walk,
+#         e_travel_time = e_travel_time_car + e_travel_time_transit + e_travel_time_walk)
+```
+
+```{r calculate-expected-travel-times, include=FALSE}
+ttm_long <- ttm %>% transmute(UID, OBJECTID,
+                                        car_prob = p_car, 
+                                        transit_prob = p_transit, 
+                                        walk_prob = p_walk, 
+                                        car_traveltime = travel_time_car, 
+                                        transit_traveltime = travel_time_transit, 
+                                        walk_traveltime = travel_time_walk) %>% 
+  pivot_longer(cols = !c("UID", "OBJECTID"), 
+               names_to = c("mode", ".value"),
+               names_sep = "_") %>%
+  group_by(UID, OBJECTID) %>%
+  summarize(e_travel_time = weighted.mean(traveltime, prob, na.rm = TRUE))
+# income_less40k = first(income_less40k)) %>% 
+#  filter(!is.nan(e_travel_time))
+
+ttm <- ttm %>% left_join(ttm_long, by = c("UID", "OBJECTID")) %>% 
+  filter(!is.nan(e_travel_time))
+```
+
+# Introduction
 
@@
 Along several other dimensions of stress [e.g., accessibility to health care facilities, @ghorbanzadeh2021spatial; @pereira2021geographic], this seems to be the case for food insecurity as well [@laborde2020poverty].
-In the US, for example, it has been estimated that there was an increase of more than 30\% in household food insecurity, and more than one third of households were discovered to be newly food insecure - meaning they did not experience food insecurity before the pandemic [@niles2020early].
-In Canada, Men and Tarasuk [-@men2021food] report that about 25\% of individuals who experienced job insecurity (a relatively common occurrence during the pandemic), also experienced food insecurity.
-Similarly, according to Statistics Canada [-@statisticscanada2020food], in the early stages of the pandemic almost 15\% of individuals reported living in a household that faced food insecurity; the risk of food insecurity was substantially higher for households with children.
-The difference between households with and without children was significant, and 11.7\% of households with children indicated that "food didn’t last and [there was] no money to get more" sometimes or often, compared to 7.3\% of households without children; likewise, 13\% of households with children indicated that they "[c]ouldn’t afford balanced meals" sometimes or often, compared to 8.8\% of households without children.
+According to Statistics Canada [-@statisticscanada2020food], in the early stages of the pandemic almost 15% of individuals reported living in a household that faced food insecurity; the risk of food insecurity was substantially higher for households with children.
+The difference between households with and without children was significant, and 11.7% of households with children indicated that "food didn't last and [there was] no money to get more" sometimes or often, compared to 7.3% of households without children; likewise, 13% of households with children indicated that they "[c]ouldn't afford balanced meals" sometimes or often, compared to 8.8% of households without children.
+Additionally, Men and Tarasuk [-@men2021food] report that about 25% of individuals who experienced job insecurity (a relatively common occurrence during the pandemic) also experienced food insecurity associated with COVID-related disruptions to employment, financial hardship, and use of food charity.
 
 The impacts of food insecurity during the pandemic are alarming, since diet-related diseases, such as obesity, heart-disease, and diabetes, were already critical public health concerns in Canada prior to COVID-19 [@boucher2017ontario].
-While food banks are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least they can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
+While emergency food services are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least they can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
 As a mid-size city grappling with deindustrialization, Hamilton exhibits high rates of poverty and food bank use.
-As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that food banks in the city recorded the highest number of visitors in the past 29 years; a rate of increase greater than population growth.
+As recently as 2019, the Hamilton Hunger Report [@hfs2019] noted that food banks in the city recorded the highest number of visitors in the past 29 years; a rate of increase greater than population growth.
 Most troubling, approximately 40% of all visitors were children.
@@
 It is known that urban food environments, within which people make their daily food choices, are essential in influencing eating behaviours and health outcomes, based on factors such as food availability, ease of geographic accessibility and socio-demographic variations [@paez2010relative; @vanderlee2017food; @widener2018spatial].
-However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to food banks.
-Previous work has explored differences in _accessing_ food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource [e.g., @enns2020experiences].
+However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to emergency and community food services.
+Previous work has explored differences in *accessing* food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource [e.g., @enns2020experiences].
 In addition, transportation and locational considerations have been raised as key issues in food bank accessibility in previous qualitative research [@smith2017food].
@@
 The study of place-based geographic accessibility is concerned with capturing the potential to reach destinations of value using the transportation network [@paez2012measuring].
-Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities^[https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with food bank service delivery [@holmes2018nothing].
-However, as with research into the prevalence of food deserts, accessibility to food banks is unlikely to be evenly distributed, and variation throughout a city can be expected due to transportation network characteristics, and the spatial distribution of food bank locations and the population they are meant to serve.
+Indeed, the Government of Canada's recent Food Policy [@foodpolicy2019] has made "access" to healthy food a priority for Canadian communities[^2] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with food bank service delivery [@holmes2018nothing].
+However, as with research into the prevalence of food deserts, accessibility to food banks is unlikely to be evenly distributed, and variation throughout a city can be expected due to transportation network characteristics and the spatial distribution of service locations and the population they are meant to serve.
 Furthermore, policy responses to the COVID-19 pandemic likely have added to the distress of vulnerable households.
@@
 This paper follows reproducible research recommendations [see @brunsdon2020opening], and the research was conducted using open source tools for transportation analysis [@lovelace2021open].
-The code and data necessary to reproduce the analysis are available in a public repository^[https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing].
-
-Food Insecurity and Food Bank Use in Canada
-============
-
-Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake [@enns2020experiences; @kirkpatrick2008food; @tarasuk2009household].
-This nutrient deficiency is associated with major population health concerns, particularly among Canadians at socio-economic disadvantage [@bazerghi2016role].
+The code and data necessary to reproduce the analysis are available in a public repository (see Supplemental Material).
+
+# Food Insecurity and Food Bank Use in Canada
+
+Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake [@kirkpatrick2008food] and poorer physical and mental health outcomes [@ramsey2011food; @seligman2010food; @stuff2004household].
+In this regard, food insecurity is a major population health concern, particularly among Canadians at socio-economic disadvantage [@bazerghi2016role].
 Official government surveys such as the Household Food Security Survey Module (HFSSM), the Canadian Community Health Surveys (CCHS), the Longitudinal and International Study of Adults (LISA), and official classifications determined by Health Canada in relation to socio-demographic variables offer some insight into food insecurity in Canada [@kirkpatrick2008food; @tarasuk2009household; @gundersen2018food].
-Nationally, it has been found that food insecurity impacts approximately 12.3% of Canadian households [@tarasuk2014food]
-
-Food banks - sometimes also referred to as ‘food pantries’ and ‘food shelves’ - originated as a community response to aid those with inadequate food by voluntarily offering them meals and ingredients [@riches2002food; @loopstra2012relationship].
-The scope and objectives of food banks can vary by region and by country, and these organizations can include not only prepared meals and aliments, but also shared spaces to connect in community gardens and community kitchens [@wakefield2013sweet].
-Although in their origin food banks were meant to be provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time they have evolved into a community practice to secure emergency food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
-
+
+Nationally, it has previously been found that food insecurity impacts approximately 12.3% of Canadian households [@tarasuk2014food] using data from the 2011-2012 CCHS.
+Using the same data, Tarasuk et al.
+[-@tarasuk2019geographic] found higher odds of food insecurity amongst households relying on social assistance, those without a university degree or with children under the age of 18, and individuals that lived alone, renters, and those identifying as Aboriginal.
+While surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada [@tarasuk2014food], pre-pandemic research from Ottawa [@enns2020experiences] and Vancouver [@black2020examining] suggests that long-term users tend to be older, have health or mobility challenges, live in large households, and are less likely to have employment income.
+In terms of geography, previous research conducted at the provincial scale using data from the 2011-2012 CCHS found that the prevalence of food insecurity ranged across the country from 11.8% of households in Ontario to 41% of households in Nunavut [@tarasuk2019geographic].
+
+Food banks - sometimes also referred to as 'food pantries' and 'food shelves' - originated as a community response to aid those with inadequate food by voluntarily offering them meals and ingredients [@riches2002food; @loopstra2012relationship].
+Although in their origin food banks were meant to be provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time many have evolved into a community practice to secure food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
 In Canada, the number of food banks has steadily increased in the past few decades [@wakefield2013sweet].
 The largest database of food banks and their use comes from the non-profit association Food Banks Canada (FBC), which conducts an annual assessment through its affiliated members.
-FBC's 2018 Hunger Count report^[https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018_FINAL_EN.pdf.aspx?ext=.pdf] (the most recent available) listed 1,830 member food banks across the country, and found that Canadians visited food banks 1.1 million times in March of 2018.
+FBC's 2018 Hunger Count report [@fbc2018] (the most recent available) listed 1,830 member food banks across the country, and found that Canadians visited food banks 1.1 million times in March of 2018.
 Of those accessing food banks, certain population characteristics tend to be over-represented compared to national totals from the 2016 Canadian Census of Population.
-According to FBC's 2018 data, single-adult households represent 45% of those utilizing food banks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those accessing food bank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
+According to FBC's 2018 data, single-adult households represent 45% of those utilizing food banks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those using food bank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
 In addition, 59% of households accessing food banks list social or disability assistance as their primary source of income.
-Similarly, using data from the 2011-2012 CCHS, Tarasuk et al.
-[-@tarasuk2019geographic] found higher odds of food insecurity amongst households relying on social assistance, those without a university degree or with children under the age of 18, and individuals that lived alone, renters, and those identifying as Aboriginal.
-While surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada [@tarasuk2014food], research from Ottawa [@enns2020experiences] and Vancouver [@black2020examining] suggests that long-term users tend to be older, have health or mobility challenges, live in large households, and are less likely to have employment income.
 
@@
 Pandemic-related food security studies in the US have found a substantial increase in households experiencing food insecurity for the first time, and also in households experiencing more severe food insecurity than before [@niles2020early; @wolfson2020food].
-Most recently in May 2020, Canada recorded 14.7% of its population living in food insecurity in the past 30 days [@statisticscanada2020food].
+Most recently in May 2020, Canada recorded 14.7% of its population living in food insecurity in the past 30 days, up from 10.5% in 2017-2018 [@statisticscanada2020food].
 Considering the negative mental and physical health effects, increases in food insecurity rates due to the onset of COVID-19, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
-Recent data^[https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis_Exec-Sum_EN.pdf] from FBC showed that 52% of member food banks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of the country.
+Recent data from FBC [@fbc2020] showed that 52% of member food banks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of the country.
 The pandemic also created significant staffing issues with 42% of food banks reporting a reduction in volunteers.
@@
 Nevertheless, some of these benefit programs were temporary.
-Although more recent statistics on food bank use in Hamilton in 2020 and 2021 are not yet available, data from the Daily Bread Food Bank in neighbouring Toronto for August 2020 shows visits climbing 51% year-over-year in that city^[https://www.dailybread.ca/wp-content/uploads/2020/11/Whos-Hungry-Report-2020.pdf], which suggests that many households in Hamilton are likely to turn to food bank services to meet their needs.
-
-In terms of geography, previous research conducted at the provincial scale using data from the 2011-2012 CCHS found that the prevalence of food insecurity ranged across the country from 11.8% of households in Ontario to 41% of households in Nunavut [@tarasuk2019geographic].
-However, although previous research has examined the characteristics of individuals and households accessing food banks, the locational or transportation accessibility aspect of food bank access is not well understood.
+Although more recent statistics on food bank use in Hamilton in 2020 and 2021 are not yet available, data from the Daily Bread Food Bank [@dbfb2020] in neighbouring Toronto for August 2020 shows visits climbing 51% year-over-year in that city, which suggests that many households in Hamilton are likely to turn to food bank services to meet their needs.
+
+Beyond traditional conceptualizations of food banks as providers of emergency food assistance, other community food services also play an important role in decreasing food insecurity.
+The scope and objectives of food banks can vary by region and by country, and these organizations can include not only prepared meals and aliments for emergency food supply, but also shared spaces to connect in community gardens and community kitchens [@wakefield2013sweet].
+However, the efficacy of these programs in reducing food insecurity differs by the type of service offered.
+For example, previous qualitative research in the Toronto region has questioned the capacity of community kitchens to improve the food security of low-income households due to their limited scale of operations, and un-subsidized kitchens were found to be particularly inaccessible to to families living in severe poverty [@tarasuk1999qualitative].
+However, other food access options such as no-cost or low-cost meals provided through community meals or congregate dining play an important role in decreasing food insecurity.
+Research in Minnesota found that seniors experiencing food insecurity valued congregate dining for providing affordable meals and a space for social gathering [@oemichen2016investigation].
+Furthermore, because seniors paid for the meal, there was no stigma attached to the use of these services compared to traditional food-purchasing assistance such as the Supplemental Nutrition Assistance Program.
+
+While previous research has examined the characteristics of individuals and households accessing emergency and community food services, the locational or transportation accessibility aspect of food bank access is not well understood.
 A wealth of literature examining the food desert concept suggests that, in addition to socio-economic and demographic factors, location and transportation networks play a key role in a household's accessibility to healthy foods [@paez2010relative; @vanderlee2017food; @widener2018spatial].
@@
 With a population of around 540,000, the City of Hamilton is the fourth largest city in Ontario.
-It has historically been home to major manufacturing industries but de-industrialization that has occurred over the past several decades has led Hamilton to become one of the most highly divided cities in Ontario, with a significant proportion of its residents living at or below Canada’s poverty level [@deluca2012code; @jakar2019turning; @latham2007determinants].
-In 2016, the Hamilton Community Foundation reported that based on the Low-Income Cut-Off, Hamilton recorded a poverty rate of 16.7%, which was well above the average rate of Ontario (13.7%) and the average national rate (12.8%).^[https://www.hamiltoncommunityfoundation.ca/vital-signs/low-income-2018/].
-According to data from Hamilton Food Share^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf], approximately 23,000 individuals accessed food banks in the city in March of 2019.
+It has historically been home to major manufacturing industries but de-industrialization that has occurred over the past several decades has led Hamilton to become one of the most highly divided cities in Ontario, with a significant proportion of its residents living at or below Canada's poverty level [@deluca2012code; @jakar2019turning; @latham2007determinants].
+The Hamilton Community Foundation [@hcf2018] reported that based on the Low-Income Cut-Off, Hamilton recorded a poverty rate of 16.7% in 2016, which was well above the average rate of Ontario (13.7%) and the average national rate (12.8%).
+According to data from Hamilton Food Share [@hfs2019], approximately 23,000 individuals accessed food banks in the city in March of 2019.
 Within this total is 9,125 visits by children (minors up to 18 years old), up from 8,278 the year before.
-Feed Ontario, the province's largest collective of hunger-relief organizations, found that on a per-capita basis, the level of need in the inner core of central Hamilton was second highest in Ontario^[https://feedontario.ca/research/report-hunger-map/].
+Feed Ontario, the province's largest collective of hunger-relief organizations, found that on a per-capita basis, the level of need in the inner core of central Hamilton was second highest in Ontario [@feedontario2019].
 
@@
 
-Data have been prepared for sharing as a data package^[https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing]. The contents of the data package are described next.
+Data have been prepared for sharing as a data package available in the Supplemental Materials. The contents of the data package are described next.
 
@@
 
-The locations of food banks and related food services were obtained from the Hamilton Public Library's Food Access Guide^[http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf].
+The locations of emergency and community food services were obtained from the Hamilton Public Library's Food Access Guide [@hpl2021].
 The guide was updated in April of 2021 to indicate any change affected on the services due to the pandemic.
@@
 Table 1 defines each service type and the number of locations pre- and during the COVID-19 pandemic.
-While some food bank services have a specific a target population, such as prioritizing families with young children aged between 0 and 3 or accepting only those providing proof of low-income status through housing and utility statements, all the food bank services indicated below are designed to accommodate those in need of food at zero to low cost.
+While some food bank services have a specific target population, such as prioritizing families with young children aged between 0 and 3 or accepting only those providing proof of low-income status through housing and utility statements, all the food services indicated below are designed to accommodate those in need of food at zero to low cost.
 With our focus on food banks and food services that offer free or low-cost meals at particular locations, we first removed services such as Meals on Wheels and other food access services such as food box, community kitchens, student nutrition programs, and shopping and transportation.
@@
 On the other hand, two free meal services held on different days at the same location were collapsed into a single service point for the accessibility analysis.
-Additional details on the operations of individual facilities is not publicly available, and with the changes in operations it proved unfeasible to collect it.
+Additional details on the operations of individual facilities is not publicly available and with the changes in operations it proved unfeasible to collect it.
 For this reason, the analysis to follow is of accessibility to the location of food banks and services, but not to specific services (e.g., breakfasts vs.
@@
 
-Travel time tables for three modes (car, transit, walking) were computed using the parcels as the origins and the locations of the food banks as the destinations.
-For routing, the package `r5r` [@pereira2021r5r] was used with a network extract for the City of Hamilton from OpenStreetMaps and the General Transit Feed Specification from Hamilton Street Railway, the local transit operator.
-For routing purposes we used maximum values of 180 min and 10,000 m walking distance: any destination that exceeded these thresholds was ignored.
+Travel time tables for three modes (car, transit, walking) were computed using the parcels as the origins and the locations of the community and emergency food service locations as the destinations.
+For routing, the package `r5r` [@pereira2021r5r] was used with a network extract for the City of Hamilton from OpenStreetMaps and the General Transit Feed Specification (GTFS) files for the Hamilton Street Railway, the local transit operator, as well as for Burlington Transit, which operates some service in the city.
+For transit routing purposes we used maximum travel time values of 300 min and a 2,000 m cap on walking distance: any destination that exceeded these thresholds was ignored.
 The departure time used for routing was 8:00AM on March 30, 2021 to reflect transit service around the morning service peak on a typical Tuesday.
@@
 Figure \ref{fig:foodbanks} shows the location of food banks and services in the City of Hamilton and their status.
-Before the pandemic there were `r nrow(foodbank_locations)` of which `r nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))` (`r round(nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations) * 100, 2)`\%) closed during the pandemic.
+Before the pandemic there were `r nrow(foodbank_locations)` of which `r nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))` (`r round(nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations) * 100, 2)`%) closed during the pandemic.
 As shown in the figure, food services tend to be predominantly located in the central parts of the city.
@@
 # Plot location of foodbanks
-
 # Create a bounding box for an inset
@@
 The general pattern of the levels of service is similar before and during the pandemic, with lower values in the center of the city where low-income households exhibit multimodal trip patterns that favour proximate service locations.
-Three more peripheral facilities towards the south of the city also have lower levels of service, presumably because they are expected to service relatively large suburban/exurban populations generally reliant on automobiles for travel.
+Three more peripheral facilities towards the south of the city have moderate levels of service, presumably because they are expected to service relatively suburban/exurban populations generally reliant on automobiles for travel.
 During the pandemic, however, the levels of service dropped, in some cases quite substantially.
 The pattern of the losses in level of service, moreover, is not uniform.
-The upper pane of Figure \ref{fig:levels-of-service-changes} shows that the three peripheral facilities in the southern suburban/exurban part of the city had low levels of service to begin with, but did not see major declines during the pandemic.
-Further, the inset map shows that the levels of service deteriorated more in the central part of the city.
-However, the loss of level of service was not as large in the core (where most of the food banks/services are found), but instead was more marked in the inner ring around the core, where facilities may have faced greater demand from both central city and suburban populations after the closure of food banks during the pandemic.
+The upper pane of Figure \ref{fig:levels-of-service-changes} shows that the peripheral facilities in the suburban/exurban parts of the city saw major declines during the pandemic as more urban locations closed and demand increased for the remaining locations.
+Further, the inset map shows that the levels of service also deteriorated in the central part of the city.
+However, the loss of level of service was not as large in the core (where most of the food banks/services are found), but instead was more marked in the inner ring around the core, where facilities may have faced greater demand from both central city and suburban populations after the closure of service locations during the pandemic.
 

### 2021-08-27 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd

@@
 The fact that we must rely on a standard created for accessible drinking water in the developing world only serves to highlight the tragedy of food insecurity in an affluent country like Canada.
-More broadly, it points to the absurd need to understand how a bad situation was made worse by the pandemic: in effect, the analysis reveals that disparities in accessibility to food banks and services predated the pandemic, that the pandemic contributed to the deterioration of these services, and that populations already in distress, particularly children, ended up in an even more adverse state.
-How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding what is an *acceptable* or *sufficient* level of service when it comes to food banks.
-Whether this is one facility per 1,000 population or three facilities per 1,000 population, or any other number, we know not.
-But like international human rights standards for accessibility to clean drinking water within 30 minutes (round-trip), this work strongly suggests that adopting better defined standards for access to food banks that move beyond generalized conceptualizations of "access" to consider the geographic component of accessibility (and its interactions with travel behaviour and individual/household socioeconomic and demographic traits) would make a meaningful contribution to understanding food access and "food bank deserts" for food insecure populations.
-
-In summary, while previous work has noted the important role of geography alongside other socio-economic and demographic indicators in household access to healthy food, this is, to the best of our knowledge, the first study that has focused on the geographic component of food bank accessibility.
-Beyond accounting for the inflation of demand that occurs in traditional floating catchment area methods, our application of the balanced floating catchment area approach offers a novel analysis of food bank and food service accessibility that is sensitive to the locations of low-income households, information on their age distributions and typical trip-making behaviour, food bank and food service locations and their operations over time, and the characteristics of the city's multi-modal transportation network.
-Despite the limitations noted above, the findings from this analysis can provide quantitative evidence to help urban planners and community administrators strategically allocate food bank services considering accessibility for different population groups, especially when considering the unequal distribution of impacts from the COVID-19 pandemic.
+More broadly, it points to the absurd need to understand how a bad situation was made worse by the pandemic: in effect, the analysis reveals that disparities in the need for emergency and community food services predated the pandemic, that the pandemic contributed to the deterioration of these services, and that populations already in distress, particularly children, ended up in an even more adverse state.
+How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding _acceptable_ or _sufficient_ level of service when it comes to emergency food services.
+In fact, we would argue that the standard should be that no household faced food insecurity.
+As others have noted [e.g., @poppendieck1999sweet; @men2021food] the root of food insecurity is income poverty.
+There is likely a place for emergency food and community food services.
+In addition to providing food, these services could satisfy social needs by offering a social setting for seniors or by helping to connect households in need with longer term supports.
+From a food security perspective, on the other hand, these services should work only as a short term solution, and not as a semi-permanent feature of life for some of our fellow human beings.
+From a human rights perspective, long-term reliance on emergency food services should be as unacceptable in Canada as lack of clean drinking water within 30 minutes is elsewhere.
+Thus, while our analysis is valuable to map the suffering caused by food insecurity, from a policy perspective maintaining a robust social safety net that includes Employment Insurance and paid sick days are better tools to reduce this suffering than increasing the accessibility of emergency food services for food insecure populations.
 
 In terms of future research, it is important to note that the degree to which our low-income cut-off of CAD\$40,000 reflects food insecurity in the different zones of the study area is not known.
-We also consider all food banks and food service locations equally.
+We also consider all emergency and communal food service locations equally.
 Information on differential capacities at food provider locations is currently not collected, and given the closure of facilities was not possible to obtain.

### 2021-08-27 Accessibility-Foodbanks-Hamilton/Response-to-Reviewers-R1.Rmd
cdh R1 edits

@@
 
-\textcolor{blue}{Words go here.}
+\textcolor{blue}{Thank you for the comment.
+We agree with the Reviewer’s critique regarding the framing of the paper and have adopted the suggestion to reframe the facilities as “emergency and community food services” as they do serve different roles.
+In addition to the terminology, we have also added a paragraph describing the role of food banks as well as other food services in reducing food insecurity in Section 2.}
 
@@
 
-\textcolor{blue}{Words go here.}
+\textcolor{blue}{By considering all facilities equally, we agree that the analysis does conflate the different service types offered.
+We better acknowledge this through the clarified terminology of "emergency and community food services" recommended above.
+However, all of these service types play an important role in food (in)security in the study area and our overarching focus remains on geographic accessibility and adaptation in the food access landscape pre- and during-COVID in Hamilton.
+Beyond the role of the transportation network, a particular interest is in how floating catchment methods can offer insight into how the closure of some of these important services affects demand and “congestion” at those that remain open.}
 
@@
 
-\textcolor{blue}{Words go here.}
+\textcolor{blue}{We agree with the critique and have streamlined the paper's introduction and Section 2 to offer greater clarity on the changing situation in Canada and Hamilton and reduce the comparisons with the United States.}
 
@@
 
-\textcolor{blue}{Words go here.}
+\textcolor{blue}{In response to this comment, we have streamlined both the introduction and concluding sections to remove restated and inapplicable information (e.g.
+the US context information in the introduction).
+Further changes to the conclusion in response to critique 5 are noted below.}
 
@@
 
+\textcolor{blue}{This paragraph has been updated to clarify the pre-covid timeline of some of the papers.}
+
 Pg 2 ln 46 (and also pg 21 ln 33) - were 40% of actual visitors children, or are they extrapolating from households with children?
 It seems unlikely that parents are actually bringing their children along, so "visitors" may be the wrong word ("users"?)
+
+\textcolor{blue}{•	This text has been updated to reflect the Reviewer’s suggestion regarding “users” rather than visitors.}
 
@@
 
+\textcolor{blue}{Thank you for this comment.
+We have re-examined the cited literature and the Kirkpatrick and Tarasuk article “Food Insecurity Is Associated with Nutrient Inadequacies among Canadian Adults and Adolescents’ in the Journal of Nutrition from 2008 is the most direct in this regard.
+This paper is more clearly cited and additional references noting the poorer physical and mental health outcomes associated with food insecurity have been added to clarify the evidence.}
+
 Pg 4 paragraph 2 - not sure that food banks include meal programs/prepared meals (the food BANK wouldn't, although the organization might offer both)
+
+\textcolor{blue}{This section has been revised to focus first on food insecurity and food bank use for emergency food access.
+Furthermore, a paragraph has been added to discuss the role of community food services in reducing food insecurity, such as community kitchens and congregate dining.
+We believe this addition strengthens the paper’s link to food insecurity and thank Reviewer 1 for raising this critique.}
 
@@
 
+\textcolor{blue}{Additional text is now included to note that the 14.7% in May 2020 is up from 10.5% reported in 2017-2018.
+Furthermore, the text about the US and Canada is now more clearly delineated.}
+
 Pg 23 (last page) - would be nice to have a final concluding sentence that restates the contribution of the paper, rather than ending with limitations.
+
+\textcolor{blue}{In addition to the changes noted above, the concluding section has been re-organized to be more impactful with regards to its conclusions.}

### 2021-08-27 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
cdh R1 edits

@@
 # load updated files manually
-load("output-data-files/ttm_transit.RData")
+load("./output-data-files/ttm_transit.RData")
 load("./output-data-files/foodbank_table.RData")
@@
 
-Economic stress caused by the COVID-19 pandemic, rising unemployment rates, and changes in poverty levels have disrupted the food environment and led to higher rates of food insecurity [@niles2020early].
+In the United States, economic stress caused by the COVID-19 pandemic, rising unemployment rates, and changes in poverty levels have disrupted the food environment and led to higher rates of food insecurity [@niles2020early].
 Pandemic-related food security studies in the US have found a substantial increase in households experiencing food insecurity for the first time, and also in households experiencing more severe food insecurity than before [@niles2020early; @wolfson2020food].
-Most recently in May 2020, Canada recorded 14.7% of its population living in food insecurity in the past 30 days, up from 10.5% in 2017-2018 [@statisticscanada2020food].
 Considering the negative mental and physical health effects, increases in food insecurity rates due to the onset of COVID-19, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
+
+Canada has also seen an increase in the number of households living in food insecurity due to the COVID-19 pandemic.
+Survey results from Statistics Canada from May of 2020 suggest that 14.7% of the population was living in food insecurity in the past 30 days, up from 10.5% in 2017-2018 [@statisticscanada2020food].
 Recent data from FBC [@fbc2020] showed that 52% of member food banks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of the country.
@@
 
-Food insecurity is a significant issue for many Canadian households and is associated with reductions in nutrition and poor mental and physical health outcomes.
-While emergency and community food services can provide some relief, the levels of need in the City of Hamilton were on the rise before the onset of the COVID-19 pandemic, with a troubling increase in the number of visits made by children to these services.
-While temporary financial relief measures may have helped in the summer of 2020, the pandemic has in all probability increased food insecurity for many households.
+Food insecurity is a significant issue for many Canadian households and while emergency and community food services can provide some relief, the COVID-19 pandemic has in all probability increased food insecurity for many households.
 To compound matters, the pandemic has also resulted in major disruptions, including to employment, mobility alternatives, and to emergency and community food services.
 In response, this research has sought to better understand accessibility to food banks and food service locations, as well as how the closure of some locations over the pre- and during-COVID time periods affected the potential for low-income households to reach these amenities.
+
+Previous work has noted the important role of geography alongside other socio-economic and demographic indicators in household access to healthy food.
+The present papers is, to the best of our knowledge, among the first studies to focus on the geographic component of accessibility to emergency food services [@allen2021Changes].
 Using the balanced floating catchment area method to account for population demand and congestion effects at service points, we estimated multi-modal accessibility to emergency and community food service locations for low-income households.
@@
 Moreover, our parcel-level analysis presents a disaggregate approach to estimating accessibility based on the locations of residential parcels and dwellings.
-
-Previous work has noted the important role of geography alongside other socio-economic and demographic indicators in household access to healthy food.
-The present papers is, to the best of our knowledge, among the first studies to focus on the geographic component of accessibility to emergency food services [@allen2021Changes].
 Beyond accounting for the inflation of demand that occurs in traditional floating catchment area methods, our application of the balanced floating catchment area approach offers a novel analysis of accessibility to emergency and communal food services that is sensitive to the locations of low-income households, information on their age distributions and typical trip-making behaviour, the locations of services, their operations over time, and the characteristics of the city's multi-modal transportation network.
+
 Our results show that while accessibility levels were lower in the city's more car-oriented suburban and rural areas to begin with, the closure of 14% of the city's emergency and community food service locations during the pandemic resulted in an overall decrease in accessibility across the city.
@@
 Reductions in accessibility in the city's more urban north end, where low-income households conduct higher proportions of trips by transit and walking, emphasize the importance of geographic proximity in the potential to reach service locations for these residents.
-Beyond geography, the results also highlight the differential impact of closures during COVID-19 on population groups, with seniors and children beign the two most impacted groups.
-
-In the absence of information regarding how food insecure households travel to food banks and related services, we examined accessibility to food banks using a 15 minutes (one-way) travel time threshold.
-The fact that we must rely on a standard created for accessible drinking water in the developing world only serves to highlight the tragedy of food insecurity in an affluent country like Canada.
-More broadly, it points to the absurd need to understand how a bad situation was made worse by the pandemic: in effect, the analysis reveals that disparities in the need for emergency and community food services predated the pandemic, that the pandemic contributed to the deterioration of these services, and that populations already in distress, particularly children, ended up in an even more adverse state.
-How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding _acceptable_ or _sufficient_ level of service when it comes to emergency food services.
-In fact, we would argue that the standard should be that no household faced food insecurity.
-As others have noted [e.g., @poppendieck1999sweet; @men2021food] the root of food insecurity is income poverty.
-There is likely a place for emergency food and community food services.
-In addition to providing food, these services could satisfy social needs by offering a social setting for seniors or by helping to connect households in need with longer term supports.
-From a food security perspective, on the other hand, these services should work only as a short term solution, and not as a semi-permanent feature of life for some of our fellow human beings.
-From a human rights perspective, long-term reliance on emergency food services should be as unacceptable in Canada as lack of clean drinking water within 30 minutes is elsewhere.
-Thus, while our analysis is valuable to map the suffering caused by food insecurity, from a policy perspective maintaining a robust social safety net that includes Employment Insurance and paid sick days are better tools to reduce this suffering than increasing the accessibility of emergency food services for food insecure populations.
-
-In terms of future research, it is important to note that the degree to which our low-income cut-off of CAD\$40,000 reflects food insecurity in the different zones of the study area is not known.
+Beyond geography, the results also highlight the differential impact of closures during COVID-19 on population groups, with seniors and children being the two most impacted groups.
+
+It is important to note that the degree to which our low-income cut-off of CAD\$40,000 reflects food insecurity in the different zones of the study area is not known.
 We also consider all emergency and communal food service locations equally.
@@
 
+In the absence of information regarding how food insecure households travel to food banks and related services, we examined accessibility to food banks using a 15 minutes (one-way) travel time threshold.
+The fact that we must rely on a standard created for accessible drinking water in the developing world only serves to highlight the tragedy of food insecurity in an affluent country like Canada.
+More broadly, it points to the absurd need to understand how a bad situation was made worse by the pandemic: in effect, the analysis reveals that disparities in the need for emergency and community food services predated the pandemic, that the pandemic contributed to the deterioration of these services, and that populations already in distress, particularly children, ended up in an even more adverse state.
+How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding _acceptable_ or _sufficient_ level of service when it comes to emergency food services.
+
+While on the one hand this work suggests that inequities in the accessibility to emergency and community food services could be improved through accessibility standards that promote changes in the geographic distribution of service locations and transportation network characteristics, in fact, we would argue that the standard should be that no household faced food insecurity.
+As others have noted [e.g., @poppendieck1999sweet; @men2021food] the root of food insecurity is income poverty and unless it is eliminated, there will continue to be a place for emergency food and community food services.
+In addition to providing food, these services satisfy social needs by offering a social setting for seniors or by helping to connect households in need with longer term supports.
+From a food security perspective, on the other hand, these services should work only as a short term solution, and not as a semi-permanent feature of life for some of our fellow human beings.
+From a human rights perspective, long-term reliance on emergency food services should be as unacceptable in Canada as lack of clean drinking water within 30 minutes is elsewhere.
+Thus, while our analysis is valuable to map the suffering caused by food insecurity, from a policy perspective, maintaining a robust social safety net that includes Employment Insurance and paid sick days are better tools to reduce this suffering than increasing the accessibility of emergency food services for food insecure populations.
+
 # References {#references .unnumbered}

### 2021-08-27 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
cdh R1 edits

@@
 
-Economic stress caused by the COVID-19 pandemic, rising unemployment
-rates, and changes in poverty levels have disrupted the food environment
-and led to higher rates of food insecurity (Niles et al., 2020).
-Pandemic-related food security studies in the US have found a
-substantial increase in households experiencing food insecurity for the
-first time, and also in households experiencing more severe food
+In the United States, economic stress caused by the COVID-19 pandemic,
+rising unemployment rates, and changes in poverty levels have disrupted
+the food environment and led to higher rates of food insecurity (Niles
+et al., 2020). Pandemic-related food security studies in the US have
+found a substantial increase in households experiencing food insecurity
+for the first time, and also in households experiencing more severe food
 insecurity than before (Niles et al., 2020; Wolfson and Leung, 2020).
-Most recently in May 2020, Canada recorded 14.7\% of its population
-living in food insecurity in the past 30 days, up from 10.5\% in
-2017-2018 (Statistics Canada, 2020a). Considering the negative mental
-and physical health effects, increases in food insecurity rates due to
-the onset of COVID-19, signal a change in the food environment with
-potential damages to population health during the course of the pandemic
-and beyond (Niles et al., 2020). Recent data from FBC (FBC, 2020) showed
-that 52\% of member food banks reported an increase in usage in March of
-2020 when initial lockdown restrictions were put in place across much of
-the country. The pandemic also created significant staffing issues with
-42\% of food banks reporting a reduction in volunteers. However, 53\% of
-food banks later reported a decrease in use into the summer of 2020
-which FBC members attributed to emergency financial support programs
-from the federal government. Nevertheless, some of these benefit
-programs were temporary. Although more recent statistics on food bank
-use in Hamilton in 2020 and 2021 are not yet available, data from the
-Daily Bread Food Bank (DBFB, 2020) in neighbouring Toronto for August
-2020 shows visits climbing 51\% year-over-year in that city, which
-suggests that many households in Hamilton are likely to turn to food
-bank services to meet their needs.
+Considering the negative mental and physical health effects, increases
+in food insecurity rates due to the onset of COVID-19, signal a change
+in the food environment with potential damages to population health
+during the course of the pandemic and beyond (Niles et al., 2020).
+
+Canada has also seen an increase in the number of households living in
+food insecurity due to the COVID-19 pandemic. Survey results from
+Statistics Canada from May of 2020 suggest that 14.7\% of the population
+was living in food insecurity in the past 30 days, up from 10.5\% in
+2017-2018 (Statistics Canada, 2020a). Recent data from FBC (FBC, 2020)
+showed that 52\% of member food banks reported an increase in usage in
+March of 2020 when initial lockdown restrictions were put in place
+across much of the country. The pandemic also created significant
+staffing issues with 42\% of food banks reporting a reduction in
+volunteers. However, 53\% of food banks later reported a decrease in use
+into the summer of 2020 which FBC members attributed to emergency
+financial support programs from the federal government. Nevertheless,
+some of these benefit programs were temporary. Although more recent
+statistics on food bank use in Hamilton in 2020 and 2021 are not yet
+available, data from the Daily Bread Food Bank (DBFB, 2020) in
+neighbouring Toronto for August 2020 shows visits climbing 51\%
+year-over-year in that city, which suggests that many households in
+Hamilton are likely to turn to food bank services to meet their needs.
 
@@
 Food insecurity is a significant issue for many Canadian households and
-is associated with reductions in nutrition and poor mental and physical
-health outcomes. While food bank use can provide some relief, the levels
-of need in the City of Hamilton were on the rise before the onset of the
-COVID-19 pandemic, with a troubling increase in the number of visits
-made by children. While temporary financial relief measures may have
-helped reduce food bank use in the summer of 2020, the pandemic has in
-all probability increased food insecurity for many households. To
-compound matters, the pandemic has also resulted in major disruptions,
-including to employment, mobility alternatives, and to food bank and
-food service delivery. In response, this research has sought to better
-understand accessibility to food banks and food service locations, as
-well as how the closure of some locations over the pre- and during-COVID
-time periods affected the potential for low-income households to reach
-these amenities. Using the balanced floating catchment area method to
-account for population demand and congestion effects at service points,
-we estimated multi-modal accessibility to food bank and food service
-locations for low-income households. The weighting of travel time
-estimates by the modal split in different zonal geographies tailors the
-results to patterns of travel behaviour captured in the regional travel
-survey. Moreover, our parcel-level analysis presents a disaggregate
-approach to estimating accessibility based on the locations of
-residential parcels and dwellings.
+while emergency and community food services can provide some relief, the
+COVID-19 pandemic has in all probability increased food insecurity for
+many households. To compound matters, the pandemic has also resulted in
+major disruptions, including to employment, mobility alternatives, and
+to emergency and community food services. In response, this research has
+sought to better understand accessibility to food banks and food service
+locations, as well as how the closure of some locations over the pre-
+and during-COVID time periods affected the potential for low-income
+households to reach these amenities.
+
+Previous work has noted the important role of geography alongside other
+socio-economic and demographic indicators in household access to healthy
+food. The present papers is, to the best of our knowledge, among the
+first studies to focus on the geographic component of accessibility to
+emergency food services (Allen and Farber, 2021). Using the balanced
+floating catchment area method to account for population demand and
+congestion effects at service points, we estimated multi-modal
+accessibility to emergency and community food service locations for
+low-income households. The weighting of travel time estimates by the
+modal split in different zonal geographies tailors the results to
+patterns of travel behaviour captured in the regional travel survey.
+Moreover, our parcel-level analysis presents a disaggregate approach to
+estimating accessibility based on the locations of residential parcels
+and dwellings. Beyond accounting for the inflation of demand that occurs
+in traditional floating catchment area methods, our application of the
+balanced floating catchment area approach offers a novel analysis of
+accessibility to emergency and communal food services that is sensitive
+to the locations of low-income households, information on their age
+distributions and typical trip-making behaviour, the locations of
+services, their operations over time, and the characteristics of the
+city's multi-modal transportation network.
 
@@
 city's more car-oriented suburban and rural areas to begin with, the
-closure of 14\% of the city's food bank and low-cost or free food
-service locations during the pandemic resulted in an overall decrease in
+closure of 14\% of the city's emergency and community food service
+locations during the pandemic resulted in an overall decrease in
 accessibility across the city. However, these effects were not uniform
@@
 floating catchment area method takes into account changes in demand and
-congestion for service providers, the closure of food bank and food
-service locations can reverberate throughout the catchment areas of the
-whole city. For some suburban zones, the closure of a relatively high
-level-of-service location results in the remaining services being spread
-over a larger population. In others, high auto dependence for trips
-leads to decreases in accessibility that accumulate due to the loss of
-several food bank and service locations initially reachable within 15
-minutes by car. Reductions in accessibility in the city's more urban
-north end, where low-income households conduct higher proportions of
-trips by transit and walking, emphasize the importance of geographic
-proximity in the potential to reach food bank and service locations for
-these residents. Beyond geography, the results also highlight the
-differential impact of food banks closing during COVID-19 on population
-groups, with seniors and children most affected.
+congestion for service providers, the closure of some services
+reverberates throughout the catchment areas of the whole city. For some
+suburban zones, the closure of a relatively high level-of-service
+location results in the remaining services being spread over a larger
+population. In others, high auto dependence for trips leads to decreases
+in accessibility that accumulate due to the loss of several locations
+initially reachable within 15 minutes by car. Reductions in
+accessibility in the city's more urban north end, where low-income
+households conduct higher proportions of trips by transit and walking,
+emphasize the importance of geographic proximity in the potential to
+reach service locations for these residents. Beyond geography, the
+results also highlight the differential impact of closures during
+COVID-19 on population groups, with seniors and children being the two
+most impacted groups.
+
+It is important to note that the degree to which our low-income cut-off
+of CAD\$40,000 reflects food insecurity in the different zones of the
+study area is not known. We also consider all emergency and communal
+food service locations equally. Information on differential capacities
+at food provider locations is currently not collected, and given the
+closure of facilities was not possible to obtain. Data on services, such
+as the number of meals served, could be used to refine the analysis in
+the future. Moreover, while the travel survey allows us to model
+multi-modal accessibilities that align with travel behaviour observed in
+the travel survey and capture differences in accessibility by age
+categories, the use of travel survey data for modelling food insecurity
+also has its limitations. Research into the population weighting methods
+used in the TTS note that the survey may under-count the lowest- and
+highest-income households in the survey study region, although the
+magnitude of this under-counting is unknown, and approximately 20\% of
+respondents to the 2016 survey did not report their income (Rose, 2018).
+In that regard, the modal splits of low-income households observed
+through the TTS data may not accurately reflect the travel behaviours of
+food insecure populations, and our estimates might in fact be somewhat
+conservative if those households who do not report income rely more on
+walking and/or transit for their mobility needs.
 
@@
 to the absurd need to understand how a bad situation was made worse by
-the pandemic: in effect, the analysis reveals that disparities in
-accessibility to food banks and services predated the pandemic, that the
-pandemic contributed to the deterioration of these services, and that
-populations already in distress, particularly children, ended up in an
-even more adverse state. How much worse, it is impossible to say, mainly
-because there is also a dearth of information, let alone standards,
-regarding what is an \emph{acceptable} or \emph{sufficient} level of
-service when it comes to food banks. Whether this is one facility per
-1,000 population or three facilities per 1,000 population, or any other
-number, we know not. But like international human rights standards for
-accessibility to clean drinking water within 30 minutes (round-trip),
-this work strongly suggests that adopting better defined standards for
-access to food banks that move beyond generalized conceptualizations of
-``access'' to consider the geographic component of accessibility (and
-its interactions with travel behaviour and individual/household
-socioeconomic and demographic traits) would make a meaningful
-contribution to understanding food access and ``food bank deserts'' for
-food insecure populations.
-
-In summary, while previous work has noted the important role of
-geography alongside other socio-economic and demographic indicators in
-household access to healthy food, this is, to the best of our knowledge,
-the first study that has focused on the geographic component of food
-bank accessibility. Beyond accounting for the inflation of demand that
-occurs in traditional floating catchment area methods, our application
-of the balanced floating catchment area approach offers a novel analysis
-of food bank and food service accessibility that is sensitive to the
-locations of low-income households, information on their age
-distributions and typical trip-making behaviour, food bank and food
-service locations and their operations over time, and the
-characteristics of the city's multi-modal transportation network.
-Despite the limitations noted above, the findings from this analysis can
-provide quantitative evidence to help urban planners and community
-administrators strategically allocate food bank services considering
-accessibility for different population groups, especially when
-considering the unequal distribution of impacts from the COVID-19
-pandemic.
-
-In terms of future research, it is important to note that the degree to
-which our low-income cut-off of CAD\$40,000 reflects food insecurity in
-the different zones of the study area is not known. We also consider all
-food banks and food service locations equally. Information on
-differential capacities at food provider locations is currently not
-collected, and given the closure of facilities was not possible to
-obtain. Data on services, such as the number of meals served, could be
-used to refine the analysis in the future. Moreover, while the travel
-survey allows us to model multi-modal accessibilities that align with
-travel behaviour observed in the travel survey and capture differences
-in accessibility by age categories, the use of travel survey data for
-modelling food insecurity also has its limitations. Research into the
-population weighting methods used in the TTS note that the survey may
-under-count the lowest- and highest-income households in the survey
-study region, although the magnitude of this under-counting is unknown,
-and approximately 20\% of respondents to the 2016 survey did not report
-their income (Rose, 2018). In that regard, the modal splits of
-low-income households observed through the TTS data may not accurately
-reflect the travel behaviours of food insecure populations, and our
-estimates might in fact be somewhat conservative if those households who
-do not report income rely more on walking and/or transit for their
-mobility needs.
+the pandemic: in effect, the analysis reveals that disparities in the
+need for emergency and community food services predated the pandemic,
+that the pandemic contributed to the deterioration of these services,
+and that populations already in distress, particularly children, ended
+up in an even more adverse state. How much worse, it is impossible to
+say, mainly because there is also a dearth of information, let alone
+standards, regarding \emph{acceptable} or \emph{sufficient} level of
+service when it comes to emergency food services.
+
+While on the one hand this work suggests that inequities in the
+accessibility to emergency and community food services could be improved
+through accessibility standards that promote changes in the geographic
+distribution of service locations and transportation network
+characteristics, in fact, we would argue that the standard should be
+that no household faced food insecurity. As others have noted (Men and
+Tarasuk, 2021; e.g., Poppendieck, 1999) the root of food insecurity is
+income poverty and unless it is eliminated, there will continue to be a
+place for emergency food and community food services. In addition to
+providing food, these services satisfy social needs by offering a social
+setting for seniors or by helping to connect households in need with
+longer term supports. From a food security perspective, on the other
+hand, these services should work only as a short term solution, and not
+as a semi-permanent feature of life for some of our fellow human beings.
+From a human rights perspective, long-term reliance on emergency food
+services should be as unacceptable in Canada as lack of clean drinking
+water within 30 minutes is elsewhere. Thus, while our analysis is
+valuable to map the suffering caused by food insecurity, from a policy
+perspective, maintaining a robust social safety net that includes
+Employment Insurance and paid sick days are better tools to reduce this
+suffering than increasing the accessibility of emergency food services
+for food insecure populations.
 
@@
 Canada.
+
+\leavevmode\hypertarget{ref-allen2021Changes}{}%
+Allen, J., Farber, S., 2021. Changes in transit accessibility to food
+banks in toronto during COVID-19. Findings.
+doi:\href{https://doi.org/10.32866/001c.24072}{10.32866/001c.24072}
 
@@
 
+\leavevmode\hypertarget{ref-poppendieck1999sweet}{}%
+Poppendieck, J., 1999. Sweet charity?: Emergency food and the end of
+entitlement. Penguin.
+
 \leavevmode\hypertarget{ref-radke2000spatial}{}%

### 2021-08-27 Accessibility-Foodbanks-Hamilton/Response-to-Reviewers-R1.Rmd
Revised table to show reduction of population serviced during covid, and proofread paper

@@
 \textcolor{blue}{Thank you for the comment.
-We agree with the Reviewer’s critique regarding the framing of the paper and have adopted the suggestion to reframe the facilities as “emergency and community food services” as they do serve different roles.
+We agree with your critique regarding the framing of the paper and have adopted the suggestion to reframe the facilities as “emergency and community food services” as they do serve different roles.
 In addition to the terminology, we have also added a paragraph describing the role of food banks as well as other food services in reducing food insecurity in Section 2.}
@@
 However, all of these service types play an important role in food (in)security in the study area and our overarching focus remains on geographic accessibility and adaptation in the food access landscape pre- and during-COVID in Hamilton.
-Beyond the role of the transportation network, a particular interest is in how floating catchment methods can offer insight into how the closure of some of these important services affects demand and “congestion” at those that remain open.}
+Beyond the role of the transportation network, a particular interest is in how floating catchment methods can offer insight into how the closure of some of these important services affects demand and the level of service at those that remain open.}
 
@@
 \begin{adjustwidth}{1cm}{}
-\textcolor{blue}{The fact that we must rely on a standard created for accessible drinking water in the developing world only serves to highlight the tragedy of food insecurity in an affluent country like Canada.
-More broadly, it points to the absurd need to understand how a bad situation was made worse by the pandemic: in effect, the analysis reveals that disparities in the need for emergency and community food services predated the pandemic, that the pandemic contributed to the deterioration of these services, and that populations already in distress, particularly children, ended up in an even more adverse state.
-How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding acceptable or sufficient level of service when it comes to emergency food services.
-In fact, we would argue that the standard should be that no household faced food insecurity.
-As others have noted [e.g., Poppendieck, 1999; Men and Tarasuk, 2021] the root of food insecurity is income poverty.
-There is likely a place for emergency food and community food services.
-In addition to providing food, these services could satisfy social needs by offering a social setting for seniors or by helping to connect households in need with longer term supports.
+\textcolor{blue}{While on the one hand this work suggests that inequities in the accessibility to emergency and community food services could be improved through accessibility standards that promote changes in the geographic distribution of service locations and transportation network characteristics, in fact, we would argue that the standard should be that no household faced food insecurity.
+As others have noted [e.g., Poppendieck, 1999; Men and Tarasuk, 2021] the root of food insecurity is income poverty and unless it is eliminated, there will continue to be a place for emergency food and community food services.
+In addition to providing food, these services satisfy social needs by offering a social setting for seniors or by helping to connect households in need with longer term supports.
 From a food security perspective, on the other hand, these services should work only as a short term solution, and not as a semi-permanent feature of life for some of our fellow human beings.
 From a human rights perspective, long-term reliance on emergency food services should be as unacceptable in Canada as lack of clean drinking water within 30 minutes is elsewhere.
-Thus, while our analysis is valuable to map the suffering caused by food insecurity, from a policy perspective maintaining a robust social safety net that includes Employment Insurance and paid sick days are better tools to reduce this suffering than increasing the accessibility of emergency food services for food insecure populations.}
+Thus, while our analysis is valuable to map the suffering caused by food insecurity, from a policy perspective, maintaining a robust social safety net that includes Employment Insurance and paid sick days are better tools to reduce this suffering than increasing the accessibility of emergency food services for food insecure populations.}
 \end{adjustwidth}
@@
 
-\textcolor{blue}{•	This text has been updated to reflect the Reviewer’s suggestion regarding “users” rather than visitors.}
+\textcolor{blue}{This text has been updated to reflect the suggestion regarding “users” rather than visitors.}
 
@@
 Furthermore, a paragraph has been added to discuss the role of community food services in reducing food insecurity, such as community kitchens and congregate dining.
-We believe this addition strengthens the paper’s link to food insecurity and thank Reviewer 1 for raising this critique.}
+We believe this addition strengthens the paper’s link to food insecurity and thank the reviewer for raising this critique.}
 
@@
 
-\textcolor{blue}{Additional text is now included to note that the 14.7% in May 2020 is up from 10.5% reported in 2017-2018.
+\textcolor{blue}{Additional text is now included to note that the 14.7\% in May 2020 is up from 10.5\% reported in 2017-2018.
 Furthermore, the text about the US and Canada is now more clearly delineated.}
@@
 \textcolor{blue}{In addition to the changes noted above, the concluding section has been re-organized to be more impactful with regards to its conclusions.}
+
+\textcolor{blue}{We hope that you will find that these revisions address your concerns in a satisfactory manner.}

### 2021-08-27 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Revised table to show reduction of population serviced during covid, and proofread paper

@@
 The impacts of food insecurity during the pandemic are alarming, since diet-related diseases, such as obesity, heart-disease, and diabetes, were already critical public health concerns in Canada prior to COVID-19 [@boucher2017ontario].
-While emergency food services are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least they can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
-As a mid-size city grappling with deindustrialization, Hamilton exhibits high rates of poverty and food bank use.
+While emergency food services are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], in reality provide a resource of last instance to households in precarious situations [@bazerghi2016role].
+As a mid-size city grappling with deindustrialization, Hamilton exhibits high rates of poverty and use of emergency food services.
 As recently as 2019, the Hamilton Hunger Report [@hfs2019] noted that food banks in the city recorded the highest number of visitors in the past 29 years; a rate of increase greater than population growth.
@@
 The study of place-based geographic accessibility is concerned with capturing the potential to reach destinations of value using the transportation network [@paez2012measuring].
-Indeed, the Government of Canada's recent Food Policy [@foodpolicy2019] has made "access" to healthy food a priority for Canadian communities[^2] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with food bank service delivery [@holmes2018nothing].
+Indeed, the Government of Canada's recent Food Policy [@foodpolicy2019] has made "access" to healthy food a priority for Canadian communities and previous research suggests that such accessibility plays a key role in user satisfaction with food bank service delivery [@holmes2018nothing].
 However, as with research into the prevalence of food deserts, accessibility to food banks is unlikely to be evenly distributed, and variation throughout a city can be expected due to transportation network characteristics and the spatial distribution of service locations and the population they are meant to serve.
@@
 
-For this study, we aim to look at how the landscape of food banks and related services (e.g.
+For this study, we aim to look at how the landscape of emergency food and related services (e.g.
 low-cost or free meal service providers) available in Hamilton, Ontario, changed during the pandemic.
-Did the number of open food bank services diminish?
-If so, what was the accessibility to food banks before the pandemic from the perspective of low income households, and how has it changed during the pandemic with respect to geographic access and congestion at remaining sites?
+Did the number of open services diminish?
+If so, what was the accessibility to emergency and community food services before the pandemic from the perspective of low income households, and how has it changed during the pandemic with respect to geographic access and congestion at remaining sites?
 And finally, who are most likely to have been impacted by changes in the accessibility landscape?
-This paper first looks at the distribution of food banks and related services before and during the pandemic.
+This paper first looks at the distribution of emergency and community food services before and during the pandemic.
 Then, we use the balanced floating catchment area approach of Paez et al.
@@
 This paper follows reproducible research recommendations [see @brunsdon2020opening], and the research was conducted using open source tools for transportation analysis [@lovelace2021open].
-The code and data necessary to reproduce the analysis are available in a public repository (see Supplemental Material).
-
-# Food Insecurity and Food Bank Use in Canada
+The data and code needed to reproduce the analysis are available in a public repository (see Supplemental Material).
+
+# Food Insecurity and Emergency and Communal Food Services in Canada
 
@@
 
-Nationally, it has previously been found that food insecurity impacts approximately 12.3% of Canadian households [@tarasuk2014food] using data from the 2011-2012 CCHS.
+Nationally, analysis of the 2011-2012 CCHS has previously revealed that food insecurity impacts approximately 12.3% of Canadian households [@tarasuk2014food].
 Using the same data, Tarasuk et al.
@@
 
-In the United States, economic stress caused by the COVID-19 pandemic, rising unemployment rates, and changes in poverty levels have disrupted the food environment and led to higher rates of food insecurity [@niles2020early].
-Pandemic-related food security studies in the US have found a substantial increase in households experiencing food insecurity for the first time, and also in households experiencing more severe food insecurity than before [@niles2020early; @wolfson2020food].
-Considering the negative mental and physical health effects, increases in food insecurity rates due to the onset of COVID-19, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
-
-Canada has also seen an increase in the number of households living in food insecurity due to the COVID-19 pandemic.
+Canada also saw an increase in the number of households living in food insecurity due to the COVID-19 pandemic.
 Survey results from Statistics Canada from May of 2020 suggest that 14.7% of the population was living in food insecurity in the past 30 days, up from 10.5% in 2017-2018 [@statisticscanada2020food].
@@
 $$
-w_{ij}^\text{st} = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
+w_{ij}^i = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
 $$
@@
 $$
-w_{ji}^\text{st} = \frac{w_{ji}}{\sum_{j=1}^Jw_{ji}}
+w_{ij}^j = \frac{w_{ij}}{\sum_{j=1}^Jw_{ij}}
 $$
@@
 $$
-\sum_{i=1}^nw_{ij}^\text{st}=1
+\sum_{i=1}^nw_{ij}^i=1
 $$
@@
 $$
-\sum_{j=1}^Jw_{ji}^\text{st}=1
+\sum_{j=1}^Jw_{ij}^j=1
 $$
@@
 $$
-\sum_{i=1}^nP_iw_{ij}^\text{st}=P_i
+\sum_{i=1}^nP_iw_{ij}^i=P_i
 $$
@@
 $$
-\sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
+\sum_{j=1}^JL_jw_{ij}^j=L_j
 $$
@@
 
-Data have been prepared for sharing as a data package available in the Supplemental Materials. The contents of the data package are described next.
+Data have been prepared for sharing in the form of an open data product [see @arribas2021open] available in the Supplemental Materials. The contents of the data package are described next.
 
@@
 $$
-w_{ij}=w_{ji}=
+w_{ij}=
 \begin{cases}
@@
 
-Using the 15-minute threshold, we find that the system-wide accessibility (interpreted as a provider-to-population ratio) was `r round(sum(results_pre$accessibility$accessibility), 3)` (food banks/service locations per low income household in the city) before COVID-19, but declined to `r round(sum(results_covid$accessibility$accessibility), 3)` during the pandemic.
-It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`% of food facilities remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`%, suggesting the location of emergency and community food services plays an important role in serving those in need.
+Using the 15-minute threshold, we find that the system-wide accessibility was `r round(sum(results_pre$accessibility$accessibility), 3)` (food banks/service locations per low income household in the city) before COVID-19, but declined to `r round(sum(results_covid$accessibility$accessibility), 3)` during the pandemic.
+It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`\% of facilities remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`\%, suggesting the location of emergency and community food services plays an important role in serving those in need.
 
@@
 
-We find that the accessibility landscape deteriorated substantially during the pandemic, with accessibility dropping on average by almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% mean() %>% abs() %>% round()`%, but with large variations: some zones experienced changes in accessibility of only about `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% max() %>% abs() %>% round()`%, whereas the most affected zone saw a loss of accessibility of almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% min() %>% abs() %>% round()`%.
+We find that the accessibility landscape deteriorated substantially during the pandemic, with accessibility dropping on average by almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% mean() %>% abs() %>% round()`\%, but with large variations: some zones experienced changes in accessibility of only about `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% max() %>% abs() %>% round()`\%, whereas the most affected zone saw a loss of accessibility of almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% min() %>% abs() %>% round()`\%.
 Figure \ref{fig:accessibility-changes-with-local-i} shows the changes in accessibility.
@@
 For the more suburban clusters of zones, the decrease in accessibility is derived from the closure of locations throughout the city reachable by car.
-In the cluster of central suburban zones for example, low-income households in the outer ring of zones that exhibit medium to high decreases in accessibility within this cluster appear to be largely auto-dependent in their tripmaking, which each each exhibiting between `r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(min(p_car)) *100)`-`r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(max(p_car)) *100)`% of their modal split for car trips.
+In the cluster of central suburban zones for example, low-income households in the outer ring of zones that exhibit medium to high decreases in accessibility within this cluster appear to be largely auto-dependent in their tripmaking, which each each exhibiting between `r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(min(p_car)) *100)`-`r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(max(p_car)) *100)`\% of their modal split for car trips.
 This results in the parcels within these zones having a large number of potentially accessible locations in the travel time matrix.
 But by extension, the change in accessibility over the pre- and during-COVID-19 time periods is affected not only by the closure of service locations proximate to the zones, but also the locations in the central city.
-The zone with the greatest decrease in accessibility within this cluster (-0.0009) has a high rate of car trips and connects to the most food bank locations in total as well as those that stayed open or closed.
+The zone with the greatest decrease in accessibility within this cluster (-0.0009) has a high rate of car trips and connects to the most facility locations in total as well as those that stayed open or closed.
 
 In the cluster to the south-west, the decrease in accessibility is predominately driven by the closure of a high level-of-service Community Meals provider.
-However, like the more central suburban zones, low-income households within this cluster are also between 90% to 100% auto-dependent in their tripmaking in the TTS.
+However, like the more central suburban zones, low-income households within this cluster are also between 90% to 100% auto-dependent in their trip-making in the TTS.
 The story is similar for the zone located in the north-west that exhibits the greatest decrease in accessibility.
@@
 Those aged 18 and less appear to have fared worse, and the largest change is in the number of children who were in the third quartile before the pandemic and found themselves in the bottom accessibility class during the pandemic.
+In fact, when we compare the total population serviced before and during the pandemic, we see that a large number of children, adults, and seniors were no longer in the catchment areas of service locations (last row of Table \ref{tab:accessibility-by-age}).
+It is remarkable that despite the loss of population serviced, accessibility still declined for those still within the catchment regions of these services.
+
 These results suggest that, through a combination of the typical modes of transportation of lower income households and the spatial distribution of the population, the food bank/service locations closed had a differential impact that more greatly affected the youngest among the population in low income households.
@@
 ## Analysis
-# Quartiles of accessibility pre-covid:
+# Quantiles of accessibility pre-covid:
 acc_quantiles_pre <- quantile(accessibility_change$`Pre-covid`,
                               c(0, 0.25, 0.5, 0.75, 1))
-# Quartiles of accessibility change:
+# Quantiles of accessibility change:
 acc_quantiles_during <- quantile(accessibility_change$`During covid`,
@@
 
+pre_total <- accessibility %>%
+  filter(timing == "Pre-covid") %>%
+  group_by(timing) %>%
+  summarize(children_pre = sum(children) %>% 
+              round(),
+            adults_pre = sum(adults) %>% 
+              round(),
+            seniors_pre = sum(seniors) %>% 
+              round(),
+            .groups = "drop") %>%
+  rename(accessibility_pre_level = timing) %>%
+  mutate(accessibility_pre_level = "Total population")
+
+summary_pre <- rbind(summary_pre, pre_total)
+
 summary_during <- accessibility %>%
@@
               round(),
-            .groups_during = "drop")
-
-accessibility_by_age <- cbind(summary_pre, summary_during[,2:4]) %>%
+            .groups = "drop")
+
+during_total <- accessibility %>%
+  filter(timing == "During covid") %>%
+  group_by(timing) %>%
+  summarize(children_during = sum(children) %>% 
+              round(),
+            adults_during = sum(adults) %>% 
+              round(),
+            seniors_during = sum(seniors) %>% 
+              round(),
+            .groups = "drop") %>%
+  rename(accessibility_during_level = timing) %>%
+  mutate(accessibility_during_level = "Total population")
+
+summary_during <- rbind(summary_during, during_total)
+
+accessibility_by_age <- cbind(summary_pre, summary_during[,2:4])
+
+accessibility_by_age <- accessibility_by_age %>%  
   mutate(children_change = round((children_during - children_pre)/children_pre * 100, 2),
@@
   add_header_above(c(" " = 1, "Pre-COVID-19" = 3, "During COVID-19" = 3)) %>%
+  row_spec(4, hline_after = TRUE) %>%
   footnote(general = c("Population values have been rounded.", 
@@
 Moreover, our parcel-level analysis presents a disaggregate approach to estimating accessibility based on the locations of residential parcels and dwellings.
-Beyond accounting for the inflation of demand that occurs in traditional floating catchment area methods, our application of the balanced floating catchment area approach offers a novel analysis of accessibility to emergency and communal food services that is sensitive to the locations of low-income households, information on their age distributions and typical trip-making behaviour, the locations of services, their operations over time, and the characteristics of the city's multi-modal transportation network.
+Beyond accounting for the inflation of demand that occurs in traditional floating catchment area methods, our application of the balanced floating catchment area approach offers a novel analysis of accessibility to emergency and community food services that is sensitive to the locations of low-income households, information on their age distributions and typical trip-making behaviour, the locations of services, their operations over time, and the characteristics of the city's multi-modal transportation network.
 
@@
 It is important to note that the degree to which our low-income cut-off of CAD\$40,000 reflects food insecurity in the different zones of the study area is not known.
-We also consider all emergency and communal food service locations equally.
+We also consider all emergency and community food service locations equally.
 Information on differential capacities at food provider locations is currently not collected, and given the closure of facilities was not possible to obtain.

### 2021-08-27 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Revised table to show reduction of population serviced during covid, and proofread paper

@@
 encourage a retrenchment of neoliberal policy (Wakefield et al., 2013),
-at least they can be argued to provide a resource of last instance to
-households in precarious situations (Bazerghi et al., 2016). As a
-mid-size city grappling with deindustrialization, Hamilton exhibits high
-rates of poverty and food bank use. As recently as 2019, the Hamilton
-Hunger Report (HFS, 2019) noted that food banks in the city recorded the
-highest number of visitors in the past 29 years; a rate of increase
-greater than population growth. Most troubling, approximately 40\% of
-all visitors were children.
+in reality provide a resource of last instance to households in
+precarious situations (Bazerghi et al., 2016). As a mid-size city
+grappling with deindustrialization, Hamilton exhibits high rates of
+poverty and use of emergency food services. As recently as 2019, the
+Hamilton Hunger Report (HFS, 2019) noted that food banks in the city
+recorded the highest number of visitors in the past 29 years; a rate of
+increase greater than population growth. Most troubling, approximately
+40\% of all visitors were children.
 
@@
 Canada's recent Food Policy (Agriculture and Canada, 2019) has made
-``access'' to healthy food a priority for Canadian
-communities{[}\^{}2{]} and previous survey research has suggested that
-such accessibility plays a key role in user satisfaction with food bank
-service delivery (Holmes et al., 2018). However, as with research into
-the prevalence of food deserts, accessibility to food banks is unlikely
-to be evenly distributed, and variation throughout a city can be
-expected due to transportation network characteristics and the spatial
-distribution of service locations and the population they are meant to
-serve. Furthermore, policy responses to the COVID-19 pandemic likely
-have added to the distress of vulnerable households. Non-pharmaceutical
-interventions during the pandemic involving restrictions in mobility
-have increased the friction of travel, in particular by transit on which
-low income populations are more reliant (e.g., DeWeese et al., 2020). At
-the same time, the pandemic has created additional stress for the
-operators of food banks through disruptions in the supply chain (e.g.,
-McKay et al., 2021) as well as concerns surrounding the delivery of
-service in safe conditions and possible cancellation of food service
-programs.
-
-For this study, we aim to look at how the landscape of food banks and
-related services (e.g.~low-cost or free meal service providers)
+``access'' to healthy food a priority for Canadian communities and
+previous research suggests that such accessibility plays a key role in
+user satisfaction with food bank service delivery (Holmes et al., 2018).
+However, as with research into the prevalence of food deserts,
+accessibility to food banks is unlikely to be evenly distributed, and
+variation throughout a city can be expected due to transportation
+network characteristics and the spatial distribution of service
+locations and the population they are meant to serve. Furthermore,
+policy responses to the COVID-19 pandemic likely have added to the
+distress of vulnerable households. Non-pharmaceutical interventions
+during the pandemic involving restrictions in mobility have increased
+the friction of travel, in particular by transit on which low income
+populations are more reliant (e.g., DeWeese et al., 2020). At the same
+time, the pandemic has created additional stress for the operators of
+food banks through disruptions in the supply chain (e.g., McKay et al.,
+2021) as well as concerns surrounding the delivery of service in safe
+conditions and possible cancellation of food service programs.
+
+For this study, we aim to look at how the landscape of emergency food
+and related services (e.g.~low-cost or free meal service providers)
 available in Hamilton, Ontario, changed during the pandemic. Did the
-number of open food bank services diminish? If so, what was the
-accessibility to food banks before the pandemic from the perspective of
-low income households, and how has it changed during the pandemic with
-respect to geographic access and congestion at remaining sites? And
-finally, who are most likely to have been impacted by changes in the
-accessibility landscape? This paper first looks at the distribution of
-food banks and related services before and during the pandemic. Then, we
-use the balanced floating catchment area approach of Paez et al. (2019)
-to investigate the accessibility situation. For this, we adopt a fully
-disaggregated approach based on parcel-level data. Socio-economic and
-demographic data are drawn from the latest Census of Canada (2016),
-whereas travel information is from the most recent regional travel
-survey from 2016. This paper follows reproducible research
-recommendations (see Brunsdon and Comber, 2020), and the research was
-conducted using open source tools for transportation analysis (Lovelace,
-2021). The code and data necessary to reproduce the analysis are
-available in a public repository (see Supplemental Material).
-
-\hypertarget{food-insecurity-and-food-bank-use-in-canada}{%
-\section{Food Insecurity and Food Bank Use in
-Canada}\label{food-insecurity-and-food-bank-use-in-canada}}
+number of open services diminish? If so, what was the accessibility to
+emergency and community food services before the pandemic from the
+perspective of low income households, and how has it changed during the
+pandemic with respect to geographic access and congestion at remaining
+sites? And finally, who are most likely to have been impacted by changes
+in the accessibility landscape? This paper first looks at the
+distribution of emergency and community food services before and during
+the pandemic. Then, we use the balanced floating catchment area approach
+of Paez et al. (2019) to investigate the accessibility situation. For
+this, we adopt a fully disaggregated approach based on parcel-level
+data. Socio-economic and demographic data are drawn from the latest
+Census of Canada (2016), whereas travel information is from the most
+recent regional travel survey from 2016. This paper follows reproducible
+research recommendations (see Brunsdon and Comber, 2020), and the
+research was conducted using open source tools for transportation
+analysis (Lovelace, 2021). The data and code needed to reproduce the
+analysis are available in a public repository (see Supplemental
+Material).
+
+\hypertarget{food-insecurity-and-emergency-and-communal-food-services-in-canada}{%
+\section{Food Insecurity and Emergency and Communal Food Services in
+Canada}\label{food-insecurity-and-emergency-and-communal-food-services-in-canada}}
 
@@
 
-Nationally, it has previously been found that food insecurity impacts
-approximately 12.3\% of Canadian households (Tarasuk et al., 2014) using
-data from the 2011-2012 CCHS. Using the same data, Tarasuk et al. (2019)
-found higher odds of food insecurity amongst households relying on
-social assistance, those without a university degree or with children
-under the age of 18, and individuals that lived alone, renters, and
-those identifying as Aboriginal. While surveys revealed that only 20 to
-30 percent of those experiencing food insecurity were found to frequent
+Nationally, analysis of the 2011-2012 CCHS has previously revealed that
+food insecurity impacts approximately 12.3\% of Canadian households
+(Tarasuk et al., 2014). Using the same data, Tarasuk et al. (2019) found
+higher odds of food insecurity amongst households relying on social
+assistance, those without a university degree or with children under the
+age of 18, and individuals that lived alone, renters, and those
+identifying as Aboriginal. While surveys revealed that only 20 to 30
+percent of those experiencing food insecurity were found to frequent
 food banks in Canada (Tarasuk et al., 2014), pre-pandemic research from
@@
 
-In the United States, economic stress caused by the COVID-19 pandemic,
-rising unemployment rates, and changes in poverty levels have disrupted
-the food environment and led to higher rates of food insecurity (Niles
-et al., 2020). Pandemic-related food security studies in the US have
-found a substantial increase in households experiencing food insecurity
-for the first time, and also in households experiencing more severe food
-insecurity than before (Niles et al., 2020; Wolfson and Leung, 2020).
-Considering the negative mental and physical health effects, increases
-in food insecurity rates due to the onset of COVID-19, signal a change
-in the food environment with potential damages to population health
-during the course of the pandemic and beyond (Niles et al., 2020).
-
-Canada has also seen an increase in the number of households living in
-food insecurity due to the COVID-19 pandemic. Survey results from
-Statistics Canada from May of 2020 suggest that 14.7\% of the population
-was living in food insecurity in the past 30 days, up from 10.5\% in
-2017-2018 (Statistics Canada, 2020a). Recent data from FBC (FBC, 2020)
-showed that 52\% of member food banks reported an increase in usage in
-March of 2020 when initial lockdown restrictions were put in place
-across much of the country. The pandemic also created significant
-staffing issues with 42\% of food banks reporting a reduction in
-volunteers. However, 53\% of food banks later reported a decrease in use
-into the summer of 2020 which FBC members attributed to emergency
-financial support programs from the federal government. Nevertheless,
-some of these benefit programs were temporary. Although more recent
-statistics on food bank use in Hamilton in 2020 and 2021 are not yet
-available, data from the Daily Bread Food Bank (DBFB, 2020) in
-neighbouring Toronto for August 2020 shows visits climbing 51\%
-year-over-year in that city, which suggests that many households in
-Hamilton are likely to turn to food bank services to meet their needs.
+Canada also saw an increase in the number of households living in food
+insecurity due to the COVID-19 pandemic. Survey results from Statistics
+Canada from May of 2020 suggest that 14.7\% of the population was living
+in food insecurity in the past 30 days, up from 10.5\% in 2017-2018
+(Statistics Canada, 2020a). Recent data from FBC (FBC, 2020) showed that
+52\% of member food banks reported an increase in usage in March of 2020
+when initial lockdown restrictions were put in place across much of the
+country. The pandemic also created significant staffing issues with 42\%
+of food banks reporting a reduction in volunteers. However, 53\% of food
+banks later reported a decrease in use into the summer of 2020 which FBC
+members attributed to emergency financial support programs from the
+federal government. Nevertheless, some of these benefit programs were
+temporary. Although more recent statistics on food bank use in Hamilton
+in 2020 and 2021 are not yet available, data from the Daily Bread Food
+Bank (DBFB, 2020) in neighbouring Toronto for August 2020 shows visits
+climbing 51\% year-over-year in that city, which suggests that many
+households in Hamilton are likely to turn to food bank services to meet
+their needs.
 
@@
 \[
-w_{ij}^\text{st} = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
+w_{ij}^i = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
 \]
@@
 \[
-w_{ji}^\text{st} = \frac{w_{ji}}{\sum_{j=1}^Jw_{ji}}
+w_{ij}^j = \frac{w_{ij}}{\sum_{j=1}^Jw_{ij}}
 \]
@@
 \[
-\sum_{i=1}^nw_{ij}^\text{st}=1
+\sum_{i=1}^nw_{ij}^i=1
 \]
@@
 \[
-\sum_{j=1}^Jw_{ji}^\text{st}=1
+\sum_{j=1}^Jw_{ij}^j=1
 \]
@@
 \[
-\sum_{i=1}^nP_iw_{ij}^\text{st}=P_i
+\sum_{i=1}^nP_iw_{ij}^i=P_i
 \]
@@
 \[
-\sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
+\sum_{j=1}^JL_jw_{ij}^j=L_j
 \]
@@
 
-Data have been prepared for sharing as a data package available in the
-Supplemental Materials. The contents of the data package are described
-next.
+Data have been prepared for sharing in the form of an open data product
+(see Arribas-Bel et al., 2021) available in the Supplemental Materials.
+The contents of the data package are described next.
 
@@
 \[
-w_{ij}=w_{ji}=
+w_{ij}=
 \begin{cases}
@@
 Using the 15-minute threshold, we find that the system-wide
-accessibility (interpreted as a provider-to-population ratio) was 0.078
-(food banks/service locations per low income household in the city)
-before COVID-19, but declined to 0.048 during the pandemic. It is
-striking that although almost 76\% of food facilities remained in
-operation during the pandemic, there was a loss of accessibility greater
-than 39\%, suggesting the location of emergency and community food
-services plays an important role in serving those in need.
+accessibility was 0.078 (food banks/service locations per low income
+household in the city) before COVID-19, but declined to 0.048 during the
+pandemic. It is striking that although almost 76\% of facilities
+remained in operation during the pandemic, there was a loss of
+accessibility greater than 39\%, suggesting the location of emergency
+and community food services plays an important role in serving those in
+need.
 
@@
 decrease in accessibility within this cluster (-0.0009) has a high rate
-of car trips and connects to the most food bank locations in total as
+of car trips and connects to the most facility locations in total as
 well as those that stayed open or closed.
@@
 low-income households within this cluster are also between 90\% to 100\%
-auto-dependent in their tripmaking in the TTS. The story is similar for
+auto-dependent in their trip-making in the TTS. The story is similar for
 the zone located in the north-west that exhibits the greatest decrease
@@
 pandemic and found themselves in the bottom accessibility class during
-the pandemic. These results suggest that, through a combination of the
-typical modes of transportation of lower income households and the
-spatial distribution of the population, the food bank/service locations
-closed had a differential impact that more greatly affected the youngest
-among the population in low income households.
+the pandemic. In fact, when we compare the total population serviced
+before and during the pandemic, we see that a large number of children,
+adults, and seniors were no longer in the catchment areas of service
+locations (last row of Table \ref{tab:accessibility-by-age}). It is
+remarkable that despite the loss of population serviced, accessibility
+still declined for those still within the catchment regions of these
+services.
+
+These results suggest that, through a combination of the typical modes
+of transportation of lower income households and the spatial
+distribution of the population, the food bank/service locations closed
+had a differential impact that more greatly affected the youngest among
+the population in low income households.
 
@@
 Bottom 25\% & 3,711 & 13,220 & 7,648 & 3,038 (-18.14\%) & 11,811 (-10.66\%) & 6,308 (-17.52\%)\\
+\midrule
+\cellcolor{gray!6}{Total population} & \cellcolor{gray!6}{11,189} & \cellcolor{gray!6}{42,576} & \cellcolor{gray!6}{23,178} & \cellcolor{gray!6}{10,441 (-6.69\%)} & \cellcolor{gray!6}{41,395 (-2.77\%)} & \cellcolor{gray!6}{22,110 (-4.61\%)}\\
 \bottomrule
@@
 balanced floating catchment area approach offers a novel analysis of
-accessibility to emergency and communal food services that is sensitive
+accessibility to emergency and community food services that is sensitive
 to the locations of low-income households, information on their age
@@
 of CAD\$40,000 reflects food insecurity in the different zones of the
-study area is not known. We also consider all emergency and communal
+study area is not known. We also consider all emergency and community
 food service locations equally. Information on differential capacities
@@
 
+\leavevmode\hypertarget{ref-arribas2021open}{}%
+Arribas-Bel, D., Green, M., Rowe, F., Singleton, A., 2021. Open data
+products: A framework for creating valuable analysis-ready data. Journal
+of Geographical Systems.
+doi:\href{https://doi.org/10.1007/s10109-021-00361-7}{10.1007/s10109-021-00361-7}
+
 \leavevmode\hypertarget{ref-bazerghi2016role}{}%
@@
 
-\leavevmode\hypertarget{ref-niles2020early}{}%
-Niles, M.T., Bertmann, F., Belarmino, E.H., Wentworth, T., Biehl, E.,
-Neff, R., 2020. The early food insecurity impacts of COVID-19. Nutrients
-12, 2096.
-
 \leavevmode\hypertarget{ref-oemichen2016investigation}{}%
@@
 
-\leavevmode\hypertarget{ref-wolfson2020food}{}%
-Wolfson, J.A., Leung, C.W., 2020. Food insecurity and COVID-19:
-Disparities in early effects for US adults. Nutrients 12, 1648.
-
 \end{CSLReferences}

### 2021-08-28 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
a few light edits to the text

@@
 We find that accessibility was low to begin with in suburban and exurban parts of the city; furthermore, about 14\% of locations originally available in Hamilton closed during the pandemic, further reducing accessibility.
-The impact of closures on the level of service of the remaining facilities, and on accessibility, was disproportionate, with system-wide losses exceeding 40\%.
+The impact of closures on the level of service of the remaining facilities, and on accessibility, was disproportionate, with system-wide losses exceeding 39\%.
 Those losses were geographically and demographically uneven.
@@
 Food banks - sometimes also referred to as 'food pantries' and 'food shelves' - originated as a community response to aid those with inadequate food by voluntarily offering them meals and ingredients [@riches2002food; @loopstra2012relationship].
-Although in their origin food banks were meant to be provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time many have evolved into a community practice to secure food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
+Although in their origin food banks were meant to provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time many have evolved into a community practice to secure food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
 In Canada, the number of food banks has steadily increased in the past few decades [@wakefield2013sweet].
@@
 
-Canada also saw an increase in the number of households living in food insecurity due to the COVID-19 pandemic.
+The COVID-19 pandemic has led to an increase in the number of households living in food insecurity in Canada.
 Survey results from Statistics Canada from May of 2020 suggest that 14.7% of the population was living in food insecurity in the past 30 days, up from 10.5% in 2017-2018 [@statisticscanada2020food].
@@
 Using the 15-minute threshold, we find that the system-wide accessibility was `r round(sum(results_pre$accessibility$accessibility), 3)` (food banks/service locations per low income household in the city) before COVID-19, but declined to `r round(sum(results_covid$accessibility$accessibility), 3)` during the pandemic.
-It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`\% of facilities remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`\%, suggesting the location of emergency and community food services plays an important role in serving those in need.
+It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`% of facilities remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`%, suggesting the location of emergency and community food services plays an important role in serving those in need.
 
@@
 However, some age groups were affected more.
-As seen in the table, the distribution within quartiles for adults changed only slightly, and in fact there are fewer adults in the bottom three quartiles as some found themselves in the (deteriorated) top quartile of accessibility during the pandemic.
-The story is different for seniors, a greater number of whom are now in the third quartile, due to a combination of seniors facing a worse accessibility situation, and just a few moving from the bottom quartile (pre-pandemic) to the third during the pandemic.
-Those aged 18 and less appear to have fared worse, and the largest change is in the number of children who were in the third quartile before the pandemic and found themselves in the bottom accessibility class during the pandemic.
-In fact, when we compare the total population serviced before and during the pandemic, we see that a large number of children, adults, and seniors were no longer in the catchment areas of service locations (last row of Table \ref{tab:accessibility-by-age}).
+In terms of changes within the quartiles, the largest change for adults appears to be those moving from the first to the second quartile of accessibility during the pandemic.
+The story is generally similar for seniors, a greater number of whom are now in the second and third quartiles due to seniors facing a worse accessibility situation.
+Among those aged 18 and less, the largest change is in the number of children who were in the first and third quartiles before the pandemic and found themselves in the second accessibility class during the pandemic.
+However, when we compare the total population serviced before and during the pandemic, we see that a large number of children, adults, and seniors were no longer in the catchment areas of service locations (last row of Table \ref{tab:accessibility-by-age}).
+This accounts for the loss of population in the fourth quartiles for the different age groups during the pandemic.
 It is remarkable that despite the loss of population serviced, accessibility still declined for those still within the catchment regions of these services.
 
-These results suggest that, through a combination of the typical modes of transportation of lower income households and the spatial distribution of the population, the food bank/service locations closed had a differential impact that more greatly affected the youngest among the population in low income households.
+These results suggest that, through a combination of the typical modes of transportation of lower income households and the spatial distribution of the population, the closure of food bank/service locations had a differential impact that was more greatly affected the youngest and oldest among the population in low income households.
 
@@
 More broadly, it points to the absurd need to understand how a bad situation was made worse by the pandemic: in effect, the analysis reveals that disparities in the need for emergency and community food services predated the pandemic, that the pandemic contributed to the deterioration of these services, and that populations already in distress, particularly children, ended up in an even more adverse state.
-How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding _acceptable_ or _sufficient_ level of service when it comes to emergency food services.
+How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding *acceptable* or *sufficient* level of service when it comes to emergency food services.
 

### 2021-08-28 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
a few light edits to the text

@@
   facilities, and on accessibility, was disproportionate, with
-  system-wide losses exceeding 40\%. Those losses were geographically
+  system-wide losses exceeding 39\%. Those losses were geographically
   and demographically uneven. While every part of the city faced a
@@
 (Loopstra and Tarasuk, 2012; Riches, 2002). Although in their origin
-food banks were meant to be provide a temporary solution to accommodate
+food banks were meant to provide a temporary solution to accommodate
 those in hunger due to job retrenchments and economic downfalls since
@@
 
-Canada also saw an increase in the number of households living in food
-insecurity due to the COVID-19 pandemic. Survey results from Statistics
+The COVID-19 pandemic has led to an increase in the number of households
+living in food insecurity in Canada. Survey results from Statistics
 Canada from May of 2020 suggest that 14.7\% of the population was living
@@
 accessibility to food banks and services in the City of Hamilton.
-However, some age groups were affected more. As seen in the table, the
-distribution within quartiles for adults changed only slightly, and in
-fact there are fewer adults in the bottom three quartiles as some found
-themselves in the (deteriorated) top quartile of accessibility during
-the pandemic. The story is different for seniors, a greater number of
-whom are now in the third quartile, due to a combination of seniors
-facing a worse accessibility situation, and just a few moving from the
-bottom quartile (pre-pandemic) to the third during the pandemic. Those
-aged 18 and less appear to have fared worse, and the largest change is
-in the number of children who were in the third quartile before the
-pandemic and found themselves in the bottom accessibility class during
-the pandemic. In fact, when we compare the total population serviced
-before and during the pandemic, we see that a large number of children,
-adults, and seniors were no longer in the catchment areas of service
-locations (last row of Table \ref{tab:accessibility-by-age}). It is
-remarkable that despite the loss of population serviced, accessibility
-still declined for those still within the catchment regions of these
-services.
+However, some age groups were affected more. In terms of changes within
+the quartiles, the largest change for adults appears to be those moving
+from the first to the second quartile of accessibility during the
+pandemic. The story is generally similar for seniors, a greater number
+of whom are now in the second and third quartiles due to seniors facing
+a worse accessibility situation. Among those aged 18 and less, the
+largest change is in the number of children who were in the first and
+third quartiles before the pandemic and found themselves in the second
+accessibility class during the pandemic. However, when we compare the
+total population serviced before and during the pandemic, we see that a
+large number of children, adults, and seniors were no longer in the
+catchment areas of service locations (last row of Table
+\ref{tab:accessibility-by-age}). This accounts for the loss of
+population in the fourth quartiles for the different age groups during
+the pandemic. It is remarkable that despite the loss of population
+serviced, accessibility still declined for those still within the
+catchment regions of these services.
 
@@
 of transportation of lower income households and the spatial
-distribution of the population, the food bank/service locations closed
-had a differential impact that more greatly affected the youngest among
-the population in low income households.
+distribution of the population, the closure of food bank/service
+locations had a differential impact that was more greatly affected the
+youngest and oldest among the population in low income households.
 

### 2021-08-29 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-V0.Rmd
Edits and file with tracked changes for resubmission to SSM

--- 2021-04-28 README.Rmd
@@
 ---
-output:
-  github_document:
-    pandoc_args: --webtex
-always_allow_html: true
-bibliography: Accessibility-Foodbanks-Hamilton/references.bib
----
-
-<!-- README.md is generated from README.Rmd. Please edit that file -->
-
-```{r, include = FALSE}
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "#>"
-)
-```
-
-# Changes in accessibility to food banks and food services during COVID-19 and implications for low income populations in Hamilton, Ontario
-
-<!-- badges: start -->
-<!-- badges: end -->
-
-Christopher D. Higgins (University of Toronto Scarborough)  
-Antonio Páez (McMaster University)  
-Gyoorie Kim (University of Toronto Mississauga)  
-Jue Wang (University of Toronto Mississauga)
-
-Paper submitted to *Social Science and Medicine*
-
-## Abstract
-
+title: "Changes in accessibility to food banks and food services during COVID-19 and implications for low income populations in Hamilton, Ontario"
+author:
+  - name: Author One
+    email: author1@example.com
+    affiliation: Some University
+  - name: Author Two
+    email: author2@example.com
+    affiliation: Another University
+    footnote: 1
+  - name: Author Three
+    email: author3@example.com
+    affiliation: Some University
+  - name: Author Four
+    email: author4@example.com
+    affiliation: Some University
+address:
+  - code: Some University
+    address: Department, Street, City, State, Zip
+  - code: Another University
+    address: Department, Street, City, State, Zip
+footnote:
+  - code: 1
+    text: "Corresponding Author"
+abstract: |
   In this paper we analyze the changes in accessibility to food banks and related services before and during the COVID-19 pandemic in the City of Hamilton, Ontario.
@@
 
-\newpage
-
 Introduction
@@
 ```{r table-food-bank-info, echo=FALSE, results='asis'}
-foodbank_table %>% kable("html",
+foodbank_table %>% kable("latex",
                          booktabs = TRUE,

### 2021-08-29 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-V0.tex
Edits and file with tracked changes for resubmission to SSM

--- 2021-08-27 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
@@
             pdfauthor={},
-            pdftitle={Changes in accessibility to emergency and community food services during COVID-19 and implications for low income populations in Hamilton, Ontario},
+            pdftitle={Changes in accessibility to food banks and food services during COVID-19 and implications for low income populations in Hamilton, Ontario},
             colorlinks=false,
@@
 
-  \title{Changes in accessibility to emergency and community food
-services during COVID-19 and implications for low income populations in
-Hamilton, Ontario}
+  \title{Changes in accessibility to food banks and food services during
+COVID-19 and implications for low income populations in Hamilton,
+Ontario}
     \author[Some University]{Author One}
@@
   \begin{abstract}
-  In this paper we analyze the changes in accessibility to community and
-  emergency food services before and during the COVID-19 pandemic in the
-  City of Hamilton, Ontario. Many of these food services are the last
-  line of support for households facing food insecurity; as such, their
+  In this paper we analyze the changes in accessibility to food banks
+  and related services before and during the COVID-19 pandemic in the
+  City of Hamilton, Ontario. Food banks and services are the last line
+  of support for households facing food insecurity; as such, their
   relevance cannot be ignored in the midst of the economic upheaval
   caused by the pandemic. Our analysis is based on the application of
-  balanced floating catchment areas and concentrates on households with
+  balanced floating catchment areas, and concentrates on households with
   lower incomes (\textless CAD40,000, approximately the Low Income
@@
 seems to be the case for food insecurity as well (Laborde et al., 2020).
-According to Statistics Canada (2020a), in the early stages of the
-pandemic almost 15\% of individuals reported living in a household that
-faced food insecurity; the risk of food insecurity was substantially
-higher for households with children. The difference between households
-with and without children was significant, and 11.7\% of households with
-children indicated that ``food didn't last and {[}there was{]} no money
-to get more'' sometimes or often, compared to 7.3\% of households
-without children; likewise, 13\% of households with children indicated
-that they ``{[}c{]}ouldn't afford balanced meals'' sometimes or often,
-compared to 8.8\% of households without children. Additionally, Men and
-Tarasuk (2021) report that about 25\% of individuals who experienced job
-insecurity (a relatively common occurrence during the pandemic) also
-experienced food insecurity associated with COVID-related disruptions to
-employment, financial hardship, and use of food charity.
+In the US, for example, it has been estimated that there was an increase
+of more than 30\% in household food insecurity, and more than one third
+of households were discovered to be newly food insecure - meaning they
+did not experience food insecurity before the pandemic (Niles et al.,
+2020). In Canada, Men and Tarasuk (2021) report that about 25\% of
+individuals who experienced job insecurity (a relatively common
+occurrence during the pandemic), also experienced food insecurity.
+Similarly, according to Statistics Canada (2020a), in the early stages
+of the pandemic almost 15\% of individuals reported living in a
+household that faced food insecurity; the risk of food insecurity was
+substantially higher for households with children. The difference
+between households with and without children was significant, and 11.7\%
+of households with children indicated that ``food didn't last and
+{[}there was{]} no money to get more'' sometimes or often, compared to
+7.3\% of households without children; likewise, 13\% of households with
+children indicated that they ``{[}c{]}ouldn't afford balanced meals''
+sometimes or often, compared to 8.8\% of households without children.
 
@@
 were already critical public health concerns in Canada prior to COVID-19
-(Boucher et al., 2017). While emergency food services are not
-necessarily a stable solution to food insecurity and in fact may
-encourage a retrenchment of neoliberal policy (Wakefield et al., 2013),
-at least they can be argued to provide a resource of last instance to
-households in precarious situations (Bazerghi et al., 2016). As a
-mid-size city grappling with deindustrialization, Hamilton exhibits high
-rates of poverty and food bank use. As recently as 2019, the Hamilton
-Hunger Report (HFS, 2019) noted that food banks in the city recorded the
-highest number of visitors in the past 29 years; a rate of increase
-greater than population growth. Most troubling, approximately 40\% of
-all visitors were children.
+(Boucher et al., 2017). While food banks are not necessarily a stable
+solution to food insecurity and in fact may encourage a retrenchment of
+neoliberal policy (Wakefield et al., 2013), at least they can be argued
+to provide a resource of last instance to households in precarious
+situations (Bazerghi et al., 2016). As a mid-size city grappling with
+deindustrialization, Hamilton exhibits high rates of poverty and food
+bank use. As recently as 2019, the Hamilton Hunger Report\footnote{https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf}
+noted that food banks in the city recorded the highest number of
+visitors in the past 29 years; a rate of increase greater than
+population growth. Most troubling, approximately 40\% of all visitors
+were children.
 
@@
 accessibility to healthy food through the ``food desert'' concept, there
-has been little research into accessibility to emergency and community
-food services. Previous work has explored differences in
-\emph{accessing} food banks, such as how some households utilize food
-banks over short periods of time while others regularly utilize food
-banks as longer-term resource (e.g., Enns et al., 2020). In addition,
-transportation and locational considerations have been raised as key
-issues in food bank accessibility in previous qualitative research
-(Smith-Carrier et al., 2017). However, we are not aware of any research
-that has focused on estimating or capturing this geographic component of
-accessibility.
+has been little research into accessibility to food banks. Previous work
+has explored differences in \emph{accessing} food banks, such as how
+some households utilize food banks over short periods of time while
+others regularly utilize food banks as longer-term resource (e.g., Enns
+et al., 2020). In addition, transportation and locational considerations
+have been raised as key issues in food bank accessibility in previous
+qualitative research (Smith-Carrier et al., 2017). However, we are not
+aware of any research that has focused on estimating or capturing this
+geographic component of accessibility.
 
@@
 transportation network (Páez et al., 2012). Indeed, the Government of
-Canada's recent Food Policy (Agriculture and Canada, 2019) has made
-``access'' to healthy food a priority for Canadian
-communities{[}\^{}2{]} and previous survey research has suggested that
-such accessibility plays a key role in user satisfaction with food bank
-service delivery (Holmes et al., 2018). However, as with research into
-the prevalence of food deserts, accessibility to food banks is unlikely
-to be evenly distributed, and variation throughout a city can be
-expected due to transportation network characteristics and the spatial
-distribution of service locations and the population they are meant to
-serve. Furthermore, policy responses to the COVID-19 pandemic likely
-have added to the distress of vulnerable households. Non-pharmaceutical
+Canada's recent Food Policy has made ``access'' to healthy food a
+priority for Canadian communities\footnote{https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042}
+and previous survey research has suggested that such accessibility plays
+a key role in user satisfaction with food bank service delivery (Holmes
+et al., 2018). However, as with research into the prevalence of food
+deserts, accessibility to food banks is unlikely to be evenly
+distributed, and variation throughout a city can be expected due to
+transportation network characteristics, and the spatial distribution of
+food bank locations and the population they are meant to serve.
+Furthermore, policy responses to the COVID-19 pandemic likely have added
+to the distress of vulnerable households. Non-pharmaceutical
 interventions during the pandemic involving restrictions in mobility
@@
 accessibility to food banks before the pandemic from the perspective of
-low income households, and how has it changed during the pandemic with
-respect to geographic access and congestion at remaining sites? And
+low income households, and how it changed during the pandemic? And
 finally, who are most likely to have been impacted by changes in the
@@
 2021). The code and data necessary to reproduce the analysis are
-available in a public repository (see Supplemental Material).
+available in a public repository\footnote{https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j\_M\_BMrXhXFvrV2c?usp=sharing}.
 
@@
 Food insecurity is the inability to acquire and consume an adequate
-amount or good quality food, leading to inadequate nutrient intake
-(Kirkpatrick and Tarasuk, 2008) and poorer physical and mental health
-outcomes (Ramsey et al., 2011; Seligman et al., 2010; Stuff et al.,
-2004). In this regard, food insecurity is a major population health
-concern, particularly among Canadians at socio-economic disadvantage
+amount or good quality food, leading to inadequate nutrient intake (Enns
+et al., 2020; Kirkpatrick and Tarasuk, 2008; Tarasuk and Vogt, 2009).
+This nutrient deficiency is associated with major population health
+concerns, particularly among Canadians at socio-economic disadvantage
 (Bazerghi et al., 2016). Official government surveys such as the
@@
 insecurity in Canada (Gundersen et al., 2018; Kirkpatrick and Tarasuk,
-2008; Tarasuk and Vogt, 2009).
-
-Nationally, it has previously been found that food insecurity impacts
-approximately 12.3\% of Canadian households (Tarasuk et al., 2014) using
-data from the 2011-2012 CCHS. Using the same data, Tarasuk et al. (2019)
-found higher odds of food insecurity amongst households relying on
-social assistance, those without a university degree or with children
-under the age of 18, and individuals that lived alone, renters, and
-those identifying as Aboriginal. While surveys revealed that only 20 to
-30 percent of those experiencing food insecurity were found to frequent
-food banks in Canada (Tarasuk et al., 2014), pre-pandemic research from
-Ottawa (Enns et al., 2020) and Vancouver (Black and Seto, 2020) suggests
-that long-term users tend to be older, have health or mobility
-challenges, live in large households, and are less likely to have
-employment income. In terms of geography, previous research conducted at
-the provincial scale using data from the 2011-2012 CCHS found that the
-prevalence of food insecurity ranged across the country from 11.8\% of
-households in Ontario to 41\% of households in Nunavut (Tarasuk et al.,
-2019).
+2008; Tarasuk and Vogt, 2009). Nationally, it has been found that food
+insecurity impacts approximately 12.3\% of Canadian households (Tarasuk
+et al., 2014)
 
@@
 inadequate food by voluntarily offering them meals and ingredients
-(Loopstra and Tarasuk, 2012; Riches, 2002). Although in their origin
-food banks were meant to be provide a temporary solution to accommodate
-those in hunger due to job retrenchments and economic downfalls since
-the 1980s, over time many have evolved into a community practice to
-secure food supplies for those in need (Loopstra and Tarasuk, 2012;
-Wakefield et al., 2013). In Canada, the number of food banks has
-steadily increased in the past few decades (Wakefield et al., 2013). The
-largest database of food banks and their use comes from the non-profit
-association Food Banks Canada (FBC), which conducts an annual assessment
-through its affiliated members. FBC's 2018 Hunger Count report (FBC,
-2018) (the most recent available) listed 1,830 member food banks across
-the country, and found that Canadians visited food banks 1.1 million
-times in March of 2018. Of those accessing food banks, certain
-population characteristics tend to be over-represented compared to
-national totals from the 2016 Canadian Census of Population. According
-to FBC's 2018 data, single-adult households represent 45\% of those
-utilizing food banks despite making up 28\% of Canada's population, 19\%
-are single-parent households (compared to 10\% nationally), and 35\% of
-those using food bank services are children aged 0-18 even though their
-share of Canada's national population is approximately 20\%. In
+(Loopstra and Tarasuk, 2012; Riches, 2002). The scope and objectives of
+food banks can vary by region and by country, and these organizations
+can include not only prepared meals and aliments, but also shared spaces
+to connect in community gardens and community kitchens (Wakefield et
+al., 2013). Although in their origin food banks were meant to be provide
+a temporary solution to accommodate those in hunger due to job
+retrenchments and economic downfalls since the 1980s, over time they
+have evolved into a community practice to secure emergency food supplies
+for those in need (Loopstra and Tarasuk, 2012; Wakefield et al., 2013).
+
+In Canada, the number of food banks has steadily increased in the past
+few decades (Wakefield et al., 2013). The largest database of food banks
+and their use comes from the non-profit association Food Banks Canada
+(FBC), which conducts an annual assessment through its affiliated
+members. FBC's 2018 Hunger Count report\footnote{https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018\_FINAL\_EN.pdf.aspx?ext=.pdf}
+(the most recent available) listed 1,830 member food banks across the
+country, and found that Canadians visited food banks 1.1 million times
+in March of 2018. Of those accessing food banks, certain population
+characteristics tend to be over-represented compared to national totals
+from the 2016 Canadian Census of Population. According to FBC's 2018
+data, single-adult households represent 45\% of those utilizing food
+banks despite making up 28\% of Canada's population, 19\% are
+single-parent households (compared to 10\% nationally), and 35\% of
+those accessing food bank services are children aged 0-18 even though
+their share of Canada's national population is approximately 20\%. In
 addition, 59\% of households accessing food banks list social or
-disability assistance as their primary source of income.
-
-In the United States, economic stress caused by the COVID-19 pandemic,
-rising unemployment rates, and changes in poverty levels have disrupted
-the food environment and led to higher rates of food insecurity (Niles
-et al., 2020). Pandemic-related food security studies in the US have
-found a substantial increase in households experiencing food insecurity
-for the first time, and also in households experiencing more severe food
+disability assistance as their primary source of income. Similarly,
+using data from the 2011-2012 CCHS, Tarasuk et al. (2019) found higher
+odds of food insecurity amongst households relying on social assistance,
+those without a university degree or with children under the age of 18,
+and individuals that lived alone, renters, and those identifying as
+Aboriginal. While surveys revealed that only 20 to 30 percent of those
+experiencing food insecurity were found to frequent food banks in Canada
+(Tarasuk et al., 2014), research from Ottawa (Enns et al., 2020) and
+Vancouver (Black and Seto, 2020) suggests that long-term users tend to
+be older, have health or mobility challenges, live in large households,
+and are less likely to have employment income.
+
+Economic stress caused by the COVID-19 pandemic, rising unemployment
+rates, and changes in poverty levels have disrupted the food environment
+and led to higher rates of food insecurity (Niles et al., 2020).
+Pandemic-related food security studies in the US have found a
+substantial increase in households experiencing food insecurity for the
+first time, and also in households experiencing more severe food
 insecurity than before (Niles et al., 2020; Wolfson and Leung, 2020).
-Considering the negative mental and physical health effects, increases
-in food insecurity rates due to the onset of COVID-19, signal a change
-in the food environment with potential damages to population health
-during the course of the pandemic and beyond (Niles et al., 2020).
-
-Canada has also seen an increase in the number of households living in
-food insecurity due to the COVID-19 pandemic. Survey results from
-Statistics Canada from May of 2020 suggest that 14.7\% of the population
-was living in food insecurity in the past 30 days, up from 10.5\% in
-2017-2018 (Statistics Canada, 2020a). Recent data from FBC (FBC, 2020)
-showed that 52\% of member food banks reported an increase in usage in
-March of 2020 when initial lockdown restrictions were put in place
-across much of the country. The pandemic also created significant
+Most recently in May 2020, Canada recorded 14.7\% of its population
+living in food insecurity in the past 30 days (Statistics Canada,
+2020a). Considering the negative mental and physical health effects,
+increases in food insecurity rates due to the onset of COVID-19, signal
+a change in the food environment with potential damages to population
+health during the course of the pandemic and beyond (Niles et al.,
+2020). Recent data\footnote{https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report\_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis\_Exec-Sum\_EN.pdf}
+from FBC showed that 52\% of member food banks reported an increase in
+usage in March of 2020 when initial lockdown restrictions were put in
+place across much of the country. The pandemic also created significant
 staffing issues with 42\% of food banks reporting a reduction in
@@
 statistics on food bank use in Hamilton in 2020 and 2021 are not yet
-available, data from the Daily Bread Food Bank (DBFB, 2020) in
-neighbouring Toronto for August 2020 shows visits climbing 51\%
-year-over-year in that city, which suggests that many households in
-Hamilton are likely to turn to food bank services to meet their needs.
-
-Beyond traditional conceptualizations of food banks as providers of
-emergency food assistance, other community food services also play an
-important role in decreasing food insecurity. The scope and objectives
-of food banks can vary by region and by country, and these organizations
-can include not only prepared meals and aliments for emergency food
-supply, but also shared spaces to connect in community gardens and
-community kitchens (Wakefield et al., 2013). However, the efficacy of
-these programs in reducing food insecurity differs by the type of
-service offered. For example, previous qualitative research in the
-Toronto region has questioned the capacity of community kitchens to
-improve the food security of low-income households due to their limited
-scale of operations, and un-subsidized kitchens were found to be
-particularly inaccessible to to families living in severe poverty
-(Tarasuk and Reynolds, 1999). However, other food access options such as
-no-cost or low-cost meals provided through community meals or congregate
-dining play an important role in decreasing food insecurity. Research in
-Minnesota found that seniors experiencing food insecurity valued
-congregate dining for providing affordable meals and a space for social
-gathering (Oemichen and Smith, 2016). Furthermore, because seniors paid
-for the meal, there was no stigma attached to the use of these services
-compared to traditional food-purchasing assistance such as the
-Supplemental Nutrition Assistance Program.
-
-While previous research has examined the characteristics of individuals
-and households accessing emergency and community food services, the
-locational or transportation accessibility aspect of food bank access is
-not well understood. A wealth of literature examining the food desert
-concept suggests that, in addition to socio-economic and demographic
-factors, location and transportation networks play a key role in a
-household's accessibility to healthy foods (Paez et al., 2010; Vanderlee
-and L'Abbé, 2017; Widener, 2018). For food banks specifically, previous
-qualitative research in Ontario by Smith-Carrier et al. (2017) has noted
-that ``transportation can be challenging, particularly if the food bank
-is situated in a remote location'' (p.~32). Particularly, it appears
-that participants experience challenges with the ``inordinate amount of
-time necessary to obtain food, and difficulties associated with
+available, data from the Daily Bread Food Bank in neighbouring Toronto
+for August 2020 shows visits climbing 51\% year-over-year in that
+city\footnote{https://www.dailybread.ca/wp-content/uploads/2020/11/Whos-Hungry-Report-2020.pdf},
+which suggests that many households in Hamilton are likely to turn to
+food bank services to meet their needs.
+
+In terms of geography, previous research conducted at the provincial
+scale using data from the 2011-2012 CCHS found that the prevalence of
+food insecurity ranged across the country from 11.8\% of households in
+Ontario to 41\% of households in Nunavut (Tarasuk et al., 2019).
+However, although previous research has examined the characteristics of
+individuals and households accessing food banks, the locational or
+transportation accessibility aspect of food bank access is not well
+understood. A wealth of literature examining the food desert concept
+suggests that, in addition to socio-economic and demographic factors,
+location and transportation networks play a key role in a household's
+accessibility to healthy foods (Paez et al., 2010; Vanderlee and L'Abbé,
+2017; Widener, 2018). For food banks specifically, previous qualitative
+research in Ontario by Smith-Carrier et al. (2017) has noted that
+``transportation can be challenging, particularly if the food bank is
+situated in a remote location'' (p.~32). Particularly, it appears that
+participants experience challenges with the ``inordinate amount of time
+necessary to obtain food, and difficulties associated with
 transportation'' (p.~39). Users of food banks, according to this
@@
 
-More formally, the first step of this method is as follows:
-
-\[
+More formally, the first step of this method is as follows: \[
 L_j = \frac{S_j}{\sum_{i=1}^nP_iw_{ij}}
@@
 that each population unit can reach, according to the distance-decay
-weight:
-
-\[
+weight: \[
 A_i = \sum_{j=1}^JL_jw_{ji}
@@
 \emph{proportionally}. This is achieved by standardizing the weights as
-follows:
-
-\[
+follows: \[
 w_{ij}^\text{st} = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
-\]
-
-\noindent and:
-
-\[
+\] \noindent and: \[
 w_{ji}^\text{st} = \frac{w_{ji}}{\sum_{j=1}^Jw_{ji}}
@@
 
-The standardized weights satisfy the following conditions:
-
-\[
+The standardized weights satisfy the following conditions: \[
 \sum_{i=1}^nw_{ij}^\text{st}=1
@@
 
-\noindent and:
-
-\[
+\noindent and: \[
 \sum_{j=1}^Jw_{ji}^\text{st}=1
@@
 Since the population is allocated proportionally, its value is
-preserved:
-
-\[
+preserved: \[
 \sum_{i=1}^nP_iw_{ij}^\text{st}=P_i
@@
 
-\noindent as is the level of service:
-
-\[
+\noindent as is the level of service: \[
 \sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
@@
 residents living at or below Canada's poverty level (DeLuca et al.,
-2012; Jakar and Dunn, 2019; Latham and Moffat, 2007). The Hamilton
-Community Foundation (HCF, 2018) reported that based on the Low-Income
-Cut-Off, Hamilton recorded a poverty rate of 16.7\% in 2016, which was
-well above the average rate of Ontario (13.7\%) and the average national
-rate (12.8\%). According to data from Hamilton Food Share (HFS, 2019),
+2012; Jakar and Dunn, 2019; Latham and Moffat, 2007). In 2016, the
+Hamilton Community Foundation reported that based on the Low-Income
+Cut-Off, Hamilton recorded a poverty rate of 16.7\%, which was well
+above the average rate of Ontario (13.7\%) and the average national rate
+(12.8\%).\footnote{https://www.hamiltoncommunityfoundation.ca/vital-signs/low-income-2018/}.
+According to data from Hamilton Food Share\footnote{https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf},
 approximately 23,000 individuals accessed food banks in the city in
@@
 on a per-capita basis, the level of need in the inner core of central
-Hamilton was second highest in Ontario (FO, 2019).
+Hamilton was second highest in Ontario\footnote{https://feedontario.ca/research/report-hunger-map/}.
 
@@
 
-Data have been prepared for sharing as a data package available in the
-Supplemental Materials. The contents of the data package are described
-next.
+Data have been prepared for sharing as a data package\footnote{https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j\_M\_BMrXhXFvrV2c?usp=sharing}.
+The contents of the data package are described next.
 
@@
 
-The locations of emergency and community food services were obtained
-from the Hamilton Public Library's Food Access Guide (HPL, 2021). The
-guide was updated in April of 2021 to indicate any change affected on
-the services due to the pandemic. This includes modified business hours,
-a need to make reservations before frequenting, and locations that have
-completely shut down in consequence. Table 1 defines each service type
-and the number of locations pre- and during the COVID-19 pandemic. While
-some food bank services have a specific target population, such as
-prioritizing families with young children aged between 0 and 3 or
-accepting only those providing proof of low-income status through
-housing and utility statements, all the food services indicated below
-are designed to accommodate those in need of food at zero to low cost.
-With our focus on food banks and food services that offer free or
-low-cost meals at particular locations, we first removed services such
-as Meals on Wheels and other food access services such as food box,
-community kitchens, student nutrition programs, and shopping and
-transportation. With some providers offering different food services at
-the same location (e.g.~food bank with free and community meal
-services), and some of these services closing after the onset of the
-COVID-19 pandemic, we opted to geocode based on the service type. On the
-other hand, two free meal services held on different days at the same
-location were collapsed into a single service point for the
+The locations of food banks and related food services were obtained from
+the Hamilton Public Library's Food Access Guide\footnote{http://foodaccessguide.ca/sites/default/files/partnersites/pdf/foodaccessguide.pdf}.
+The guide was updated in April of 2021 to indicate any change affected
+on the services due to the pandemic. This includes modified business
+hours, a need to make reservations before frequenting, and locations
+that have completely shut down in consequence. Table 1 defines each
+service type and the number of locations pre- and during the COVID-19
+pandemic. While some food bank services have a specific a target
+population, such as prioritizing families with young children aged
+between 0 and 3 or accepting only those providing proof of low-income
+status through housing and utility statements, all the food bank
+services indicated below are designed to accommodate those in need of
+food at zero to low cost. With our focus on food banks and food services
+that offer free or low-cost meals at particular locations, we first
+removed services such as Meals on Wheels and other food access services
+such as food box, community kitchens, student nutrition programs, and
+shopping and transportation. With some providers offering different food
+services at the same location (e.g.~food bank with free and community
+meal services), and some of these services closing after the onset of
+the COVID-19 pandemic, we opted to geocode based on the service type. On
+the other hand, two free meal services held on different days at the
+same location were collapsed into a single service point for the
 accessibility analysis. Additional details on the operations of
-individual facilities is not publicly available and with the changes in
+individual facilities is not publicly available, and with the changes in
 operations it proved unfeasible to collect it. For this reason, the
@@
 Transportation may be provided} & \cellcolor{gray!6}{7} & \cellcolor{gray!6}{2} & \cellcolor{gray!6}{One remaining location reduced hours during COVID}\\
-Community Meals & No-cost programs often run by volunteers that organize suppers, lunches or other get-togethers that give community residents an opportunity to meet one another in a friendly and informal atmosphere while sharing a meal & 11 & 9 & NA\\
-\cellcolor{gray!6}{Food Banks} & \cellcolor{gray!6}{Food Banks and Emergency Food programs provide individuals and families with grocery items free of charge} & \cellcolor{gray!6}{26} & \cellcolor{gray!6}{25} & \cellcolor{gray!6}{One remaining location reduced hours during COVID while 4 others moved to appointment only}\\
+Community Meals & Programs often run by volunteers that organize suppers, lunches or other get-togethers that give community residents an opportunity to meet one another in a friendly and informal atmosphere while sharing a meal & 11 & 9 & NA\\
+\cellcolor{gray!6}{Food Banks} & \cellcolor{gray!6}{Food Banks and Emergency Food programs provide individuals and families with grocery items free of charge} & \cellcolor{gray!6}{27} & \cellcolor{gray!6}{26} & \cellcolor{gray!6}{One remaining location reduced hours during COVID while 4 others moved to appointment only}\\
 Free Meals & Meals are provided free of charge in the community through volunteer labour and donations & 9 & 5 & One remaining location reduced hours during COVID\\
@@
 Travel time tables for three modes (car, transit, walking) were computed
-using the parcels as the origins and the locations of the community and
-emergency food service locations as the destinations. For routing, the
-package \texttt{r5r} (Pereira et al., 2021b) was used with a network
-extract for the City of Hamilton from OpenStreetMaps and the General
-Transit Feed Specification (GTFS) files for the Hamilton Street Railway,
-the local transit operator, as well as for Burlington Transit, which
-operates some service in the city. For transit routing purposes we used
-maximum travel time values of 300 min and a 2,000 m cap on walking
-distance: any destination that exceeded these thresholds was ignored.
-The departure time used for routing was 8:00AM on March 30, 2021 to
-reflect transit service around the morning service peak on a typical
-Tuesday.
+using the parcels as the origins and the locations of the food banks as
+the destinations. For routing, the package \texttt{r5r} (Pereira et al.,
+2021b) was used with a network extract for the City of Hamilton from
+OpenStreetMaps and the General Transit Feed Specification from Hamilton
+Street Railway, the local transit operator. For routing purposes we used
+maximum values of 180 min and 10,000 m walking distance: any destination
+that exceeded these thresholds was ignored. The departure time used for
+routing was 8:00AM on March 30, 2021 to reflect transit service around
+the morning service peak on a typical Tuesday.
 
@@
 travel time \(ett\) from each parcel \(i\) to a food bank or food
-service location \(j\) as follows:
-
-\[
+service location \(j\) as follows: \[
 ett_{ij} = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
@@
 parcel \(i\) to the food bank. In other words, the expected travel time
-reflects the weighted average of travel times to the food bank, with the
-weights given by the expected modal split of trips made by low-income
-households in the TAZ per the TTS data.
+is the weighted sum of travel times to the food bank, with the weights
+given by the expected modal split of trips made by low-income households
+in the TAZ per the TTS data.
 
@@
 
+\begin{verbatim}
+## Warning: `guides(<scale> = FALSE)` is deprecated. Please use `guides(<scale> =
+## "none")` instead.
+
+## Warning: `guides(<scale> = FALSE)` is deprecated. Please use `guides(<scale> =
+## "none")` instead.
+\end{verbatim}
+
 \begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-location-foodbanks-1} \caption{\label{fig:foodbanks}Location of food banks/services and operation status; the dotted box is an inset of the central part of the City of Hamilton}\label{fig:plot-location-foodbanks}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton-V0_files/figure-latex/plot-location-foodbanks-1} \caption{\label{fig:foodbanks}Location of food banks/services and operation status; the dotted box is an inset of the central part of the City of Hamilton}\label{fig:plot-location-foodbanks}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-low-income-households-1} \caption{\label{fig:low-income-households}Number and proportion of households with incomes less than CAD40,000.}\label{fig:plot-low-income-households}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton-V0_files/figure-latex/plot-low-income-households-1} \caption{\label{fig:low-income-households}Number and proportion of households with incomes less than CAD40,000.}\label{fig:plot-low-income-households}
 \end{figure}
@@
 literature, which is mostly silent on the travel patterns of people who
-visit food banks and community food services. For this reason, we opt
-for a simple cumulative opportunities function as follows:
-
-\[
+visit food banks. For this reason, we opt for a simple cumulative
+opportunities function as follows: \[
 w_{ij}=w_{ji}=
@@
 \end{cases}
-\]
-
-\noindent where \(ett_{ij}\) is the multimodal expected travel time as
-described previously, and \(\delta\) is a travel threshold. When the
+\] \noindent where \(ett_{ij}\) is the multimodal expected travel time
+as described previously, and \(\delta\) is a travel threshold. When the
 expected travel time exceeds this threshold, a facility is no longer
@@
 \begin{figure}
-\includegraphics[width=0.6\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-results-sensitivity-analysis-1} \caption{\label{fig:sensitivity-analysis}Accessibility as a function of threshold}\label{fig:plot-results-sensitivity-analysis}
+\includegraphics[width=0.6\linewidth]{Accessibility-Foodbanks-Hamilton-V0_files/figure-latex/plot-results-sensitivity-analysis-1} \caption{\label{fig:sensitivity-analysis}Accessibility as a function of threshold}\label{fig:plot-results-sensitivity-analysis}
 \end{figure}
@@
 Using the 15-minute threshold, we find that the system-wide
-accessibility (interpreted as a provider-to-population ratio) was 0.078
+accessibility (interpreted as a provider-to-population ratio) was 0.065
 (food banks/service locations per low income household in the city)
-before COVID-19, but declined to 0.048 during the pandemic. It is
+before COVID-19, but declined to 0.037 during the pandemic. It is
 striking that although almost 76\% of food facilities remained in
 operation during the pandemic, there was a loss of accessibility greater
-than 39\%, suggesting the location of emergency and community food
-services plays an important role in serving those in need.
+than 43\%, suggesting the location of food banks and related services
+plays an important role in serving those in need.
 
@@
 proximate service locations. Three more peripheral facilities towards
-the south of the city have moderate levels of service, presumably
-because they are expected to service relatively suburban/exurban
+the south of the city also have lower levels of service, presumably
+because they are expected to service relatively large suburban/exurban
 populations generally reliant on automobiles for travel. During the
@@
 is not uniform. The upper pane of Figure
-\ref{fig:levels-of-service-changes} shows that the peripheral facilities
-in the suburban/exurban parts of the city saw major declines during the
-pandemic as more urban locations closed and demand increased for the
-remaining locations. Further, the inset map shows that the levels of
-service also deteriorated in the central part of the city. However, the
-loss of level of service was not as large in the core (where most of the
-food banks/services are found), but instead was more marked in the inner
-ring around the core, where facilities may have faced greater demand
-from both central city and suburban populations after the closure of
-service locations during the pandemic.
+\ref{fig:levels-of-service-changes} shows that the three peripheral
+facilities in the southern suburban/exurban part of the city had low
+levels of service to begin with, but did not see major declines during
+the pandemic. Further, the inset map shows that the levels of service
+deteriorated more in the central part of the city. However, the loss of
+level of service was not as large in the core (where most of the food
+banks/services are found), but instead was more marked in the inner ring
+around the core, where facilities may have faced greater demand from
+both central city and suburban populations after the closure of food
+banks during the pandemic.
+
+\begin{verbatim}
+## Warning: `guides(<scale> = FALSE)` is deprecated. Please use `guides(<scale> =
+## "none")` instead.
+
+## Warning: `guides(<scale> = FALSE)` is deprecated. Please use `guides(<scale> =
+## "none")` instead.
+\end{verbatim}
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-levels-of-service-1} \caption{\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (bottom panel).}\label{fig:plot-levels-of-service}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton-V0_files/figure-latex/plot-levels-of-service-1} \caption{\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (bottom panel).}\label{fig:plot-levels-of-service}
 \end{figure}
 
+\begin{verbatim}
+## Warning: `guides(<scale> = FALSE)` is deprecated. Please use `guides(<scale> =
+## "none")` instead.
+
+## Warning: `guides(<scale> = FALSE)` is deprecated. Please use `guides(<scale> =
+## "none")` instead.
+\end{verbatim}
+
 \begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-levels-of-service-changes-1} \caption{\label{fig:levels-of-service-changes}Changes in levels of service at each facility from pre-COVID-19 to during COVID-19.}\label{fig:plot-levels-of-service-changes}
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton-V0_files/figure-latex/plot-levels-of-service-changes-1} \caption{\label{fig:levels-of-service-changes}Changes in levels of service at each facility from pre-COVID-19 to during COVID-19.}\label{fig:plot-levels-of-service-changes}
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-accessibility-1}
+{\centering \includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton-V0_files/figure-latex/plot-accessibility-1}
 
@@
 during the pandemic, with accessibility dropping on average by almost
-38\%, but with large variations: some zones experienced changes in
-accessibility of only about 8\%, whereas the most affected zone saw a
-loss of accessibility of almost 96\%. Figure
+41\%, but with large variations: some zones experienced changes in
+accessibility of only about 10\%, whereas the most affected zone saw a
+loss of accessibility of almost 71\%. Figure
 \ref{fig:accessibility-changes-with-local-i} shows the changes in
@@
 
+\begin{verbatim}
+## Warning: `guides(<scale> = FALSE)` is deprecated. Please use `guides(<scale> =
+## "none")` instead.
+\end{verbatim}
+
 \begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-accessibility-changes-1}
-
-}
-
-\caption{\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19.}\label{fig:plot-accessibility-changes}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton_files/figure-latex/plot-local-i-1} \caption{\label{fig:accessibility-changes-with-local-i}Changes in accessibility from pre-COVID-19 to during COVID-19.
+\includegraphics[width=1\linewidth]{Accessibility-Foodbanks-Hamilton-V0_files/figure-latex/plot-local-i-1} \caption{\label{fig:accessibility-changes-with-local-i}Changes in accessibility from pre-COVID-19 to during COVID-19.
 Highlighted areas had significantly large changes in accessibility according to Local Moran's I.}\label{fig:plot-local-i}
@@
 of car trips and connects to the most food bank locations in total as
-well as those that stayed open or closed.
+well as those that stayed open or closed. The exception to this is the
+zone in the middle of this cluster, which has a much higher proportion
+of transit use (32\%). This modal split weighting results in fewer food
+bank and service locations accessible from the parcels within this zone
+which, by extension, limits their access to closed locations reachable
+within 15 minutes in our calculations.
 
@@
 in the TTS. Households in some zones take transit more often and one
-zone in particular has 100\% of its trips by walking. For these zones,
-the decrease in accessibility tends to be a product of the closure of
-several inner-city food bank and service locations reachable by multiple
-modes.
+zone in particular has 100\% of its trips by walking while in others,
+car use makes up 100\% of trips. For these zones, the decrease in
+accessibility tends to be a product of the closure of several inner-city
+food bank and service locations.
 
@@
 \midrule
-\cellcolor{gray!6}{Top 25\%} & \cellcolor{gray!6}{2,048} & \cellcolor{gray!6}{7,031} & \cellcolor{gray!6}{3,632} & \cellcolor{gray!6}{1,956 (-4.49\%)} & \cellcolor{gray!6}{6,138 (-12.7\%)} & \cellcolor{gray!6}{3,338 (-8.09\%)}\\
-Second 25\% & 2,190 & 9,631 & 5,023 & 2,365 (7.99\%) & 10,833 (12.48\%) & 5,400 (7.51\%)\\
-\cellcolor{gray!6}{Third 25\%} & \cellcolor{gray!6}{3,239} & \cellcolor{gray!6}{12,694} & \cellcolor{gray!6}{6,875} & \cellcolor{gray!6}{3,082 (-4.85\%)} & \cellcolor{gray!6}{12,612 (-0.65\%)} & \cellcolor{gray!6}{7,064 (2.75\%)}\\
-Bottom 25\% & 3,711 & 13,220 & 7,648 & 3,038 (-18.14\%) & 11,811 (-10.66\%) & 6,308 (-17.52\%)\\
+\cellcolor{gray!6}{Top 25\%} & \cellcolor{gray!6}{2,546} & \cellcolor{gray!6}{10,404} & \cellcolor{gray!6}{5,008} & \cellcolor{gray!6}{2,567 (0.82\%)} & \cellcolor{gray!6}{10,597 (1.86\%)} & \cellcolor{gray!6}{4,941 (-1.34\%)}\\
+Second 25\% & 3,333 & 12,916 & 6,415 & 3,375 (1.26\%) & 12,885 (-0.24\%) & 6,157 (-4.02\%)\\
+\cellcolor{gray!6}{Third 25\%} & \cellcolor{gray!6}{4,381} & \cellcolor{gray!6}{11,559} & \cellcolor{gray!6}{7,067} & \cellcolor{gray!6}{4,062 (-7.28\%)} & \cellcolor{gray!6}{11,412 (-1.27\%)} & \cellcolor{gray!6}{7,678 (8.65\%)}\\
+Bottom 25\% & 3,546 & 12,946 & 6,903 & 3,802 (7.22\%) & 12,897 (-0.38\%) & 6,567 (-4.87\%)\\
 \bottomrule
@@
 Food insecurity is a significant issue for many Canadian households and
-while emergency and community food services can provide some relief, the
-COVID-19 pandemic has in all probability increased food insecurity for
-many households. To compound matters, the pandemic has also resulted in
-major disruptions, including to employment, mobility alternatives, and
-to emergency and community food services. In response, this research has
-sought to better understand accessibility to food banks and food service
-locations, as well as how the closure of some locations over the pre-
-and during-COVID time periods affected the potential for low-income
-households to reach these amenities.
-
-Previous work has noted the important role of geography alongside other
-socio-economic and demographic indicators in household access to healthy
-food. The present papers is, to the best of our knowledge, among the
-first studies to focus on the geographic component of accessibility to
-emergency food services (Allen and Farber, 2021). Using the balanced
-floating catchment area method to account for population demand and
-congestion effects at service points, we estimated multi-modal
-accessibility to emergency and community food service locations for
-low-income households. The weighting of travel time estimates by the
-modal split in different zonal geographies tailors the results to
-patterns of travel behaviour captured in the regional travel survey.
-Moreover, our parcel-level analysis presents a disaggregate approach to
-estimating accessibility based on the locations of residential parcels
-and dwellings. Beyond accounting for the inflation of demand that occurs
-in traditional floating catchment area methods, our application of the
-balanced floating catchment area approach offers a novel analysis of
-accessibility to emergency and communal food services that is sensitive
-to the locations of low-income households, information on their age
-distributions and typical trip-making behaviour, the locations of
-services, their operations over time, and the characteristics of the
-city's multi-modal transportation network.
+is associated with reductions in nutrition and poor mental and physical
+health outcomes. While food bank use can provide some relief, the levels
+of need in the City of Hamilton were on the rise before the onset of the
+COVID-19 pandemic, with a troubling increase in the number of visits
+made by children. While temporary financial relief measures may have
+helped reduce food bank use in the summer of 2020, the pandemic has in
+all probability increased food insecurity for many households. To
+compound matters, the pandemic has also resulted in major disruptions,
+including to employment, mobility alternatives, and to food bank and
+food service delivery. In response, this research has sought to better
+understand accessibility to food banks and food service locations, as
+well as how the closure of some locations over the pre- and during-COVID
+time periods affected the potential for low-income households to reach
+these amenities. Using the balanced floating catchment area method to
+account for population demand and congestion effects at service points,
+we estimated multi-modal accessibility to food bank and food service
+locations for low-income households. The weighting of travel time
+estimates by the modal split in different zonal geographies tailors the
+results to patterns of travel behaviour captured in the regional travel
+survey. Moreover, our parcel-level analysis presents a disaggregate
+approach to estimating accessibility based on the locations of
+residential parcels and dwellings.
 
@@
 city's more car-oriented suburban and rural areas to begin with, the
-closure of 14\% of the city's emergency and community food service
-locations during the pandemic resulted in an overall decrease in
+closure of 14\% of the city's food bank and low-cost or free food
+service locations during the pandemic resulted in an overall decrease in
 accessibility across the city. However, these effects were not uniform
@@
 floating catchment area method takes into account changes in demand and
-congestion for service providers, the closure of some services
-reverberates throughout the catchment areas of the whole city. For some
-suburban zones, the closure of a relatively high level-of-service
-location results in the remaining services being spread over a larger
-population. In others, high auto dependence for trips leads to decreases
-in accessibility that accumulate due to the loss of several locations
-initially reachable within 15 minutes by car. Reductions in
-accessibility in the city's more urban north end, where low-income
-households conduct higher proportions of trips by transit and walking,
-emphasize the importance of geographic proximity in the potential to
-reach service locations for these residents. Beyond geography, the
-results also highlight the differential impact of closures during
-COVID-19 on population groups, with seniors and children being the two
-most impacted groups.
-
-It is important to note that the degree to which our low-income cut-off
-of CAD\$40,000 reflects food insecurity in the different zones of the
-study area is not known. We also consider all emergency and communal
-food service locations equally. Information on differential capacities
-at food provider locations is currently not collected, and given the
-closure of facilities was not possible to obtain. Data on services, such
-as the number of meals served, could be used to refine the analysis in
-the future. Moreover, while the travel survey allows us to model
-multi-modal accessibilities that align with travel behaviour observed in
-the travel survey and capture differences in accessibility by age
-categories, the use of travel survey data for modelling food insecurity
-also has its limitations. Research into the population weighting methods
-used in the TTS note that the survey may under-count the lowest- and
-highest-income households in the survey study region, although the
-magnitude of this under-counting is unknown, and approximately 20\% of
-respondents to the 2016 survey did not report their income (Rose, 2018).
-In that regard, the modal splits of low-income households observed
-through the TTS data may not accurately reflect the travel behaviours of
-food insecure populations, and our estimates might in fact be somewhat
-conservative if those households who do not report income rely more on
-walking and/or transit for their mobility needs.
+congestion for service providers, the closure of food bank and food
+service locations can reverberate throughout the catchment areas of the
+whole city. For some suburban zones, the closure of a relatively high
+level-of-service location results in the remaining services being spread
+over a larger population. In others, high auto dependence for trips
+leads to decreases in accessibility that accumulate due to the loss of
+several food bank and service locations initially reachable within 15
+minutes by car. Reductions in accessibility in the city's more urban
+north end, where low-income households conduct higher proportions of
+trips by transit and walking, emphasize the importance of geographic
+proximity in the potential to reach food bank and service locations for
+these residents. Beyond geography, the results also highlight the
+differential impact of food banks closing during COVID-19 on population
+groups, with seniors and children most affected.
 
@@
 to the absurd need to understand how a bad situation was made worse by
-the pandemic: in effect, the analysis reveals that disparities in the
-need for emergency and community food services predated the pandemic,
-that the pandemic contributed to the deterioration of these services,
-and that populations already in distress, particularly children, ended
-up in an even more adverse state. How much worse, it is impossible to
-say, mainly because there is also a dearth of information, let alone
-standards, regarding \emph{acceptable} or \emph{sufficient} level of
-service when it comes to emergency food services.
-
-While on the one hand this work suggests that inequities in the
-accessibility to emergency and community food services could be improved
-through accessibility standards that promote changes in the geographic
-distribution of service locations and transportation network
-characteristics, in fact, we would argue that the standard should be
-that no household faced food insecurity. As others have noted (Men and
-Tarasuk, 2021; e.g., Poppendieck, 1999) the root of food insecurity is
-income poverty and unless it is eliminated, there will continue to be a
-place for emergency food and community food services. In addition to
-providing food, these services satisfy social needs by offering a social
-setting for seniors or by helping to connect households in need with
-longer term supports. From a food security perspective, on the other
-hand, these services should work only as a short term solution, and not
-as a semi-permanent feature of life for some of our fellow human beings.
-From a human rights perspective, long-term reliance on emergency food
-services should be as unacceptable in Canada as lack of clean drinking
-water within 30 minutes is elsewhere. Thus, while our analysis is
-valuable to map the suffering caused by food insecurity, from a policy
-perspective, maintaining a robust social safety net that includes
-Employment Insurance and paid sick days are better tools to reduce this
-suffering than increasing the accessibility of emergency food services
-for food insecure populations.
+the pandemic: in effect, the analysis reveals that disparities in
+accessibility to food banks and services predated the pandemic, that the
+pandemic contributed to the deterioration of these services, and that
+populations already in distress, particularly children, ended up in an
+even more adverse state. How much worse, it is impossible to say, mainly
+because there is also a dearth of information, let alone standards,
+regarding what is an \emph{acceptable} or \emph{sufficient} level of
+service when it comes to food banks. Whether this is one facility per
+1,000 population or three facilities per 1,000 population, or any other
+number, we know not. But like international human rights standards for
+accessibility to clean drinking water within 30 minutes (round-trip),
+this work strongly suggests that adopting better defined standards for
+access to food banks that move beyond generalized conceptualizations of
+``access'' to consider the geographic component of accessibility (and
+its interactions with travel behaviour and individual/household
+socioeconomic and demographic traits) would make a meaningful
+contribution to understanding food access and ``food bank deserts'' for
+food insecure populations.
+
+In summary, while previous work has noted the important role of
+geography alongside other socio-economic and demographic indicators in
+household access to healthy food, this is, to the best of our knowledge,
+the first study that has focused on the geographic component of food
+bank accessibility. Beyond accounting for the inflation of demand that
+occurs in traditional floating catchment area methods, our application
+of the balanced floating catchment area approach offers a novel analysis
+of food bank and food service accessibility that is sensitive to the
+locations of low-income households, information on their age
+distributions and typical trip-making behaviour, food bank and food
+service locations and their operations over time, and the
+characteristics of the city's multi-modal transportation network.
+Despite the limitations noted above, the findings from this analysis can
+provide quantitative evidence to help urban planners and community
+administrators strategically allocate food bank services considering
+accessibility for different population groups, especially when
+considering the unequal distribution of impacts from the COVID-19
+pandemic.
+
+In terms of future research, it is important to note that the degree to
+which our low-income cut-off of CAD\$40,000 reflects food insecurity in
+the different zones of the study area is not known. We also consider all
+food banks and food service locations equally. Information on
+differential capacities at food provider locations is currently not
+collected, and given the closure of facilities was not possible to
+obtain. Data on services, such as the number of meals served, could be
+used to refine the analysis in the future. Moreover, while the travel
+survey allows us to model multi-modal accessibilities that align with
+travel behaviour observed in the travel survey and capture differences
+in accessibility by age categories, the use of travel survey data for
+modelling food insecurity also has its limitations. Research into the
+population weighting methods used in the TTS note that the survey may
+under-count the lowest- and highest-income households in the survey
+study region, although the magnitude of this under-counting is unknown,
+and approximately 20\% of respondents to the 2016 survey did not report
+their income (Rose, 2018). In that regard, the modal splits of
+low-income households observed through the TTS data may not accurately
+reflect the travel behaviours of food insecure populations, and our
+estimates might in fact be somewhat conservative if those households who
+do not report income rely more on walking and/or transit for their
+mobility needs.
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-foodpolicy2019}{}%
-Agriculture, Canada, A.-F., 2019. Food policy for canada. Government of
-Canada.
-
-\leavevmode\hypertarget{ref-allen2021Changes}{}%
-Allen, J., Farber, S., 2021. Changes in transit accessibility to food
-banks in toronto during COVID-19. Findings.
-doi:\href{https://doi.org/10.32866/001c.24072}{10.32866/001c.24072}
-
 \leavevmode\hypertarget{ref-anselin1995local}{}%
@@
 
-\leavevmode\hypertarget{ref-dbfb2020}{}%
-DBFB, 2020. Who's hungry 2020. Daily Bread Food Bank.
-
 \leavevmode\hypertarget{ref-delamater2013spatial}{}%
@@
 
-\leavevmode\hypertarget{ref-fbc2018}{}%
-FBC, 2018. HungerCount 2018. Food Banks Canada.
-
-\leavevmode\hypertarget{ref-fbc2020}{}%
-FBC, 2020. A snapshot of food banks in canada and the COVID-19 crisis.
-Food Banks Canada.
-
-\leavevmode\hypertarget{ref-feedontario2019}{}%
-FO, 2019. Report: Hunger map. Feed Ontario.
-
 \leavevmode\hypertarget{ref-ghorbanzadeh2021spatial}{}%
@@
 canada. PloS one 13, e0202642.
-
-\leavevmode\hypertarget{ref-hcf2018}{}%
-HCF, 2018. Vital signs: A reflection of hamilton. Hamilton Community
-Foundation.
-
-\leavevmode\hypertarget{ref-hfs2019}{}%
-HFS, 2019. Hamilton hunger report 2019. Hamilton Food Share.
 
@@
 
-\leavevmode\hypertarget{ref-hpl2021}{}%
-HPL, 2021. Food access guide. Hamilton Public Library.
-
 \leavevmode\hypertarget{ref-jakar2019turning}{}%
@@
 12, 2096.
-
-\leavevmode\hypertarget{ref-oemichen2016investigation}{}%
-Oemichen, M., Smith, C., 2016. Investigation of the food choice,
-promoters and barriers to food access issues, and food insecurity among
-low-income, free-living minnesotan seniors. Journal of nutrition
-education and behavior 48, 397--404.
 
@@
 
-\leavevmode\hypertarget{ref-poppendieck1999sweet}{}%
-Poppendieck, J., 1999. Sweet charity?: Emergency food and the end of
-entitlement. Penguin.
-
 \leavevmode\hypertarget{ref-radke2000spatial}{}%
@@
 
-\leavevmode\hypertarget{ref-tarasuk1999qualitative}{}%
-Tarasuk, V., Reynolds, R., 1999. A qualitative study of community
-kitchens as a response to income-related food insecurity. Canadian
-Journal of Dietetic Practice and Research 60, 11.
-
 \leavevmode\hypertarget{ref-tarasuk2009household}{}%

### 2021-08-29 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-abstract.Rmd
Edits and file with tracked changes for resubmission to SSM

@@
 abstract: |
-  In this paper we analyze the changes in accessibility to community and emergency food services before and during the COVID-19 pandemic in the City of Hamilton, Ontario.
+  In this paper we analyze the changes in accessibility to emergency and community food services before and during the COVID-19 pandemic in the City of Hamilton, Ontario.
 Many of these food services are the last line of support for households facing food insecurity; as such, their relevance cannot be ignored in the midst of the economic upheaval caused by the pandemic.

### 2021-08-29 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-abstract.tex
Edits and file with tracked changes for resubmission to SSM

@@
   \begin{abstract}
-  In this paper we analyze the changes in accessibility to food banks
-  and related services before and during the COVID-19 pandemic in the
-  City of Hamilton, Ontario. Food banks and services are the last line
-  of support for households facing food insecurity; as such, their
+  In this paper we analyze the changes in accessibility to emergency and
+  community food services before and during the COVID-19 pandemic in the
+  City of Hamilton, Ontario. Many of these food services are the last
+  line of support for households facing food insecurity; as such, their
   relevance cannot be ignored in the midst of the economic upheaval
   caused by the pandemic. Our analysis is based on the application of
-  balanced floating catchment areas, and concentrates on households with
+  balanced floating catchment areas and concentrates on households with
   lower incomes (\textless CAD40,000, approximately the Low Income

### 2021-08-29 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-title-page.Rmd
Edits and file with tracked changes for resubmission to SSM

@@
 
-**Author 1:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization, Supervision; **Author 2:** Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Visualization; **Author 3:** Conceptualization, Investigation, Data Curation, Writing - Original Draft; **Author 4:** Resources, Writing - Original Draft, Supervision
+**Christopher D.
+Higgins**: Conceptualization $\cdot$ Methodology	$\cdot$ Software $\cdot$ Validation	$\cdot$ Formal analysis	$\cdot$ Investigation	$\cdot$ Data Curation	$\cdot$ Writing - Original Draft $\cdot$ Visualization $\cdot$ Supervision **Antonio Páez**: Conceptualization $\cdot$ Methodology	$\cdot$ Software $\cdot$ Validation	$\cdot$ Formal analysis	$\cdot$ Investigation	$\cdot$ Data Curation	$\cdot$ Writing - Original Draft $\cdot$ Visualization **Gyoorie Kim**: Conceptualization $\cdot$ Investigation	$\cdot$ Data Curation	$\cdot$ Writing - Review & Editing **Jue Wang**: Conceptualization $\cdot$ Resources	$\cdot$ Writing - Review & Editing	$\cdot$ Supervision

### 2021-08-29 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-title-page.tex
Edits and file with tracked changes for resubmission to SSM

@@
 
-\textbf{Author 1:} Conceptualization, Methodology, Software, Validation,
-Formal analysis, Investigation, Data Curation, Writing - Original Draft,
-Visualization, Supervision; \textbf{Author 2:} Conceptualization,
-Methodology, Software, Validation, Formal analysis, Investigation, Data
-Curation, Writing - Original Draft, Visualization; \textbf{Author 3:}
-Conceptualization, Investigation, Data Curation, Writing - Original
-Draft; \textbf{Author 4:} Resources, Writing - Original Draft,
-Supervision
+\textbf{Christopher D. Higgins}: Conceptualization \(\cdot\) Methodology
+\(\cdot\) Software \(\cdot\) Validation \(\cdot\) Formal analysis
+\(\cdot\) Investigation \(\cdot\) Data Curation \(\cdot\) Writing -
+Original Draft \(\cdot\) Visualization \(\cdot\) Supervision
+\textbf{Antonio Páez}: Conceptualization \(\cdot\) Methodology \(\cdot\)
+Software \(\cdot\) Validation \(\cdot\) Formal analysis \(\cdot\)
+Investigation \(\cdot\) Data Curation \(\cdot\) Writing - Original Draft
+\(\cdot\) Visualization \textbf{Gyoorie Kim}: Conceptualization
+\(\cdot\) Investigation \(\cdot\) Data Curation \(\cdot\) Writing -
+Review \& Editing \textbf{Jue Wang}: Conceptualization \(\cdot\)
+Resources \(\cdot\) Writing - Review \& Editing \(\cdot\) Supervision
 

### 2021-08-29 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Edits and file with tracked changes for resubmission to SSM

@@
 This paper follows reproducible research recommendations [see @brunsdon2020opening], and the research was conducted using open source tools for transportation analysis [@lovelace2021open].
-The data and code needed to reproduce the analysis are available in a public repository (see Supplemental Material).
+The data and code needed to reproduce the analysis are available in an (anonymous) [Drive Folder](https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing).
 
@@
 
-Data have been prepared for sharing in the form of an open data product [see @arribas2021open] available in the Supplemental Materials. The contents of the data package are described next.
+Data have been prepared for sharing in the form of an open data product [see @arribas2021open] available in an (anonymous) [Drive Folder](https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing).
+The contents of the data package are described next.
 
@@
 
-These results suggest that, through a combination of the typical modes of transportation of lower income households and the spatial distribution of the population, the closure of food bank/service locations had a differential impact that was more greatly affected the youngest and oldest among the population in low income households.
+These results suggest that, through a combination of the typical modes of transportation of lower income households and the spatial distribution of the population, the closure of emergency and communal food locations had a differential impact that more greatly affected the youngest and oldest among the population in low income households.
 

### 2021-08-29 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Edits and file with tracked changes for resubmission to SSM

@@
   \begin{abstract}
-  In this paper we analyze the changes in accessibility to community and
-  emergency food services before and during the COVID-19 pandemic in the
+  In this paper we analyze the changes in accessibility to emergency and
+  community food services before and during the COVID-19 pandemic in the
   City of Hamilton, Ontario. Many of these food services are the last
@@
 analysis (Lovelace, 2021). The data and code needed to reproduce the
-analysis are available in a public repository (see Supplemental
-Material).
+analysis are available in an (anonymous)
+\href{https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing}{Drive
+Folder}.
 
@@
 Data have been prepared for sharing in the form of an open data product
-(see Arribas-Bel et al., 2021) available in the Supplemental Materials.
-The contents of the data package are described next.
+(see Arribas-Bel et al., 2021) available in an (anonymous)
+\href{https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing}{Drive
+Folder}. The contents of the data package are described next.
 
@@
 of transportation of lower income households and the spatial
-distribution of the population, the closure of food bank/service
-locations had a differential impact that was more greatly affected the
+distribution of the population, the closure of emergency and communal
+food locations had a differential impact that more greatly affected the
 youngest and oldest among the population in low income households.

### 2021-08-29 Accessibility-Foodbanks-Hamilton/CRediT-statement.Rmd
Edits and file with tracked changes for resubmission to SSM

---
title: "CRediT Statement"
author: ""
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
bibliography: references.bib
header-includes:
  - \usepackage{changepage}
---

<!--
Conceptualization $\cdot$ Methodology	$\cdot$ Software $\cdot$ Validation	$\cdot$ Formal analysis	$\cdot$ Investigation	$\cdot$ Resources	$\cdot$ Data Curation	$\cdot$ Writing - Original Draft $\cdot$ Writing - Review & Editing	$\cdot$ Visualization $\cdot$ Supervision $\cdot$ Project administration $\cdot$ Funding 
-->

Christopher D. Higgins: Conceptualization $\cdot$ Methodology	$\cdot$ Software $\cdot$ Validation	$\cdot$ Formal analysis	$\cdot$ Investigation	$\cdot$ Data Curation	$\cdot$ Writing - Original Draft $\cdot$ Visualization $\cdot$ Supervision

Antonio Páez: Conceptualization $\cdot$ Methodology	$\cdot$ Software $\cdot$ Validation	$\cdot$ Formal analysis	$\cdot$ Investigation	$\cdot$ Data Curation	$\cdot$ Writing - Original Draft $\cdot$ Visualization

Gyoorie Kim: Conceptualization $\cdot$ Investigation	$\cdot$ Data Curation	$\cdot$ Writing - Review & Editing

Jue Wang: Conceptualization $\cdot$ Resources	$\cdot$ Writing - Review & Editing	$\cdot$ Supervision

### 2021-08-29 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Recreated data package with new foodbank_table.Rmd and ttm_transit.Rmd; knitted README.Rmd after submission to SSM

@@
 data(foodbank_locations)
-#data(foodbank_table)
+data(foodbank_table)
 data(hamilton_boundary)
@@
 data(ttm_car)
-#data(ttm_transit)
+data(ttm_transit)
 data(ttm_walk)
@@
 # load updated files manually
-load("./output-data-files/ttm_transit.RData")
-load("./output-data-files/foodbank_table.RData")
+#load("./output-data-files/ttm_transit.RData")
+#load("./output-data-files/foodbank_table.RData")
 ```

### 2021-08-29 README.Rmd
Recreated data package with new foodbank_table.Rmd and ttm_transit.Rmd; knitted README.Rmd after submission to SSM

@@
 
-In this paper we analyze the changes in accessibility to food banks and related services before and during the COVID-19 pandemic in the City of Hamilton, Ontario.
-Food banks and services are the last line of support for households facing food insecurity; as such, their relevance cannot be ignored in the midst of the economic upheaval caused by the pandemic.
-Our analysis is based on the application of balanced floating catchment areas, and concentrates on households with lower incomes (<CAD40,000, approximately the Low Income Cutoff Value for a city of Hamilton's size).
+In this paper we analyze the changes in accessibility to emergency and community food services before and during the COVID-19 pandemic in the City of Hamilton, Ontario.
+Many of these food services are the last line of support for households facing food insecurity; as such, their relevance cannot be ignored in the midst of the economic upheaval caused by the pandemic.
+Our analysis is based on the application of balanced floating catchment areas and concentrates on households with lower incomes (<CAD40,000, approximately the Low Income Cutoff Value for a city of Hamilton's size).
 We find that accessibility was low to begin with in suburban and exurban parts of the city; furthermore, about 14\% of locations originally available in Hamilton closed during the pandemic, further reducing accessibility.
@@
 - Food insecurity  
+- Emergency food services  
+- Community food services  
 - Food banks  
 - Accessibility  
+- Balanced Floating Catchment Areas  
 - COVID-19  
@@
 data(ttm_walk)
+
+# load updated files manually
+#oad("./Accessibility-Foodbanks-Hamilton/output-data-files/ttm_transit.RData")
+#load("./Accessibility-Foodbanks-Hamilton/output-data-files/foodbank_table.RData")
 ```
@@
 The total expected travel time is the sum of the expected travel times by mode:
-ttm <- ttm %>%
-  mutate(e_travel_time_car = travel_time_car * p_car,
-         e_travel_time_transit = travel_time_transit * p_transit,
-         e_travel_time_walk = travel_time_walk * p_walk,
-         e_travel_time = e_travel_time_car + e_travel_time_transit + e_travel_time_walk)
-```
-
-\newpage
-
-Introduction
-==========================
+#ttm_2021 <- ttm_2021 %>%
+#  mutate(e_travel_time_car = travel_time_car * p_car,
+#         e_travel_time_transit = travel_time_transit * p_transit,
+#         e_travel_time_walk = travel_time_walk * p_walk,
+#         e_travel_time = e_travel_time_car + e_travel_time_transit + e_travel_time_walk)
+```
+
+```{r calculate-expected-travel-times, include=FALSE}
+ttm_long <- ttm %>% transmute(UID, OBJECTID,
+                                        car_prob = p_car, 
+                                        transit_prob = p_transit, 
+                                        walk_prob = p_walk, 
+                                        car_traveltime = travel_time_car, 
+                                        transit_traveltime = travel_time_transit, 
+                                        walk_traveltime = travel_time_walk) %>% 
+  pivot_longer(cols = !c("UID", "OBJECTID"), 
+               names_to = c("mode", ".value"),
+               names_sep = "_") %>%
+  group_by(UID, OBJECTID) %>%
+  summarize(e_travel_time = weighted.mean(traveltime, prob, na.rm = TRUE))
+# income_less40k = first(income_less40k)) %>% 
+#  filter(!is.nan(e_travel_time))
+
+ttm <- ttm %>% left_join(ttm_long, by = c("UID", "OBJECTID")) %>% 
+  filter(!is.nan(e_travel_time))
+```
+
+# Introduction
 
@@
 Along several other dimensions of stress [e.g., accessibility to health care facilities, @ghorbanzadeh2021spatial; @pereira2021geographic], this seems to be the case for food insecurity as well [@laborde2020poverty].
-In the US, for example, it has been estimated that there was an increase of more than 30\% in household food insecurity, and more than one third of households were discovered to be newly food insecure - meaning they did not experience food insecurity before the pandemic [@niles2020early].
-In Canada, Men and Tarasuk [-@men2021food] report that about 25\% of individuals who experienced job insecurity (a relatively common occurrence during the pandemic), also experienced food insecurity.
-Similarly, according to Statistics Canada [-@statisticscanada2020food], in the early stages of the pandemic almost 15\% of individuals reported living in a household that faced food insecurity; the risk of food insecurity was substantially higher for households with children.
-The difference between households with and without children was significant, and 11.7\% of households with children indicated that "food didn’t last and [there was] no money to get more" sometimes or often, compared to 7.3\% of households without children; likewise, 13\% of households with children indicated that they "[c]ouldn’t afford balanced meals" sometimes or often, compared to 8.8\% of households without children.
+According to Statistics Canada [-@statisticscanada2020food], in the early stages of the pandemic almost 15% of individuals reported living in a household that faced food insecurity; the risk of food insecurity was substantially higher for households with children.
+The difference between households with and without children was significant, and 11.7% of households with children indicated that "food didn't last and [there was] no money to get more" sometimes or often, compared to 7.3% of households without children; likewise, 13% of households with children indicated that they "[c]ouldn't afford balanced meals" sometimes or often, compared to 8.8% of households without children.
+Additionally, Men and Tarasuk [-@men2021food] report that about 25% of individuals who experienced job insecurity (a relatively common occurrence during the pandemic) also experienced food insecurity associated with COVID-related disruptions to employment, financial hardship, and use of food charity.
 
 The impacts of food insecurity during the pandemic are alarming, since diet-related diseases, such as obesity, heart-disease, and diabetes, were already critical public health concerns in Canada prior to COVID-19 [@boucher2017ontario].
-While food banks are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], at least they can be argued to provide a resource of last instance to households in precarious situations [@bazerghi2016role].
-As a mid-size city grappling with deindustrialization, Hamilton exhibits high rates of poverty and food bank use.
-As recently as 2019, the Hamilton Hunger Report^[https://www.hamiltonfoodshare.org/wp-content/uploads/Hamilton-Food-Share-Hunger-Report-2019.pdf] noted that food banks in the city recorded the highest number of visitors in the past 29 years; a rate of increase greater than population growth.
+While emergency food services are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], in reality provide a resource of last instance to households in precarious situations [@bazerghi2016role].
+As a mid-size city grappling with deindustrialization, Hamilton exhibits high rates of poverty and use of emergency food services.
+As recently as 2019, the Hamilton Hunger Report [@hfs2019] noted that food banks in the city recorded the highest number of visitors in the past 29 years; a rate of increase greater than population growth.
 Most troubling, approximately 40% of all visitors were children.
@@
 The study of place-based geographic accessibility is concerned with capturing the potential to reach destinations of value using the transportation network [@paez2012measuring].
-Indeed, the Government of Canada's recent Food Policy has made "access" to healthy food a priority for Canadian communities^[https://www.agr.gc.ca/eng/about-our-department/key-departmental-initiatives/food-policy/the-food-policy-for-canada/?id=1597863791042] and previous survey research has suggested that such accessibility plays a key role in user satisfaction with food bank service delivery [@holmes2018nothing].
-However, as with research into the prevalence of food deserts, accessibility to food banks is unlikely to be evenly distributed, and variation throughout a city can be expected due to transportation network characteristics, and the spatial distribution of food bank locations and the population they are meant to serve.
+Indeed, the Government of Canada's recent Food Policy [@foodpolicy2019] has made "access" to healthy food a priority for Canadian communities and previous research suggests that such accessibility plays a key role in user satisfaction with food bank service delivery [@holmes2018nothing].
+However, as with research into the prevalence of food deserts, accessibility to food banks is unlikely to be evenly distributed, and variation throughout a city can be expected due to transportation network characteristics and the spatial distribution of service locations and the population they are meant to serve.
 Furthermore, policy responses to the COVID-19 pandemic likely have added to the distress of vulnerable households.
@@
 
-For this study, we aim to look at how the landscape of food banks and related services (e.g.
+For this study, we aim to look at how the landscape of emergency food and related services (e.g.
 low-cost or free meal service providers) available in Hamilton, Ontario, changed during the pandemic.
-Did the number of open food bank services diminish?
-If so, what was the accessibility to food banks before the pandemic from the perspective of low income households, and how it changed during the pandemic?
+Did the number of open services diminish?
+If so, what was the accessibility to emergency and community food services before the pandemic from the perspective of low income households, and how has it changed during the pandemic with respect to geographic access and congestion at remaining sites?
 And finally, who are most likely to have been impacted by changes in the accessibility landscape?
-This paper first looks at the distribution of food banks and related services before and during the pandemic.
+This paper first looks at the distribution of emergency and community food services before and during the pandemic.
 Then, we use the balanced floating catchment area approach of Paez et al.
@@
 This paper follows reproducible research recommendations [see @brunsdon2020opening], and the research was conducted using open source tools for transportation analysis [@lovelace2021open].
-The code and data necessary to reproduce the analysis are available in a public repository^[https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing].
-
-Food Insecurity and Food Bank Use in Canada
-============
-
-Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake [@enns2020experiences; @kirkpatrick2008food; @tarasuk2009household].
-This nutrient deficiency is associated with major population health concerns, particularly among Canadians at socio-economic disadvantage [@bazerghi2016role].
+The data and code needed to reproduce the analysis are available in an (anonymous) [Drive Folder](https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing).
+
+# Food Insecurity and Emergency and Communal Food Services in Canada
+
+Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake [@kirkpatrick2008food] and poorer physical and mental health outcomes [@ramsey2011food; @seligman2010food; @stuff2004household].
+In this regard, food insecurity is a major population health concern, particularly among Canadians at socio-economic disadvantage [@bazerghi2016role].
 Official government surveys such as the Household Food Security Survey Module (HFSSM), the Canadian Community Health Surveys (CCHS), the Longitudinal and International Study of Adults (LISA), and official classifications determined by Health Canada in relation to socio-demographic variables offer some insight into food insecurity in Canada [@kirkpatrick2008food; @tarasuk2009household; @gundersen2018food].
-Nationally, it has been found that food insecurity impacts approximately 12.3% of Canadian households [@tarasuk2014food]
-
-Food banks - sometimes also referred to as ‘food pantries’ and ‘food shelves’ - originated as a community response to aid those with inadequate food by voluntarily offering them meals and ingredients [@riches2002food; @loopstra2012relationship].
-The scope and objectives of food banks can vary by region and by country, and these organizations can include not only prepared meals and aliments, but also shared spaces to connect in community gardens and community kitchens [@wakefield2013sweet].
-Although in their origin food banks were meant to be provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time they have evolved into a community practice to secure emergency food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
-
+
+Nationally, analysis of the 2011-2012 CCHS has previously revealed that food insecurity impacts approximately 12.3% of Canadian households [@tarasuk2014food].
+Using the same data, Tarasuk et al.
+[-@tarasuk2019geographic] found higher odds of food insecurity amongst households relying on social assistance, those without a university degree or with children under the age of 18, and individuals that lived alone, renters, and those identifying as Aboriginal.
+While surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada [@tarasuk2014food], pre-pandemic research from Ottawa [@enns2020experiences] and Vancouver [@black2020examining] suggests that long-term users tend to be older, have health or mobility challenges, live in large households, and are less likely to have employment income.
+In terms of geography, previous research conducted at the provincial scale using data from the 2011-2012 CCHS found that the prevalence of food insecurity ranged across the country from 11.8% of households in Ontario to 41% of households in Nunavut [@tarasuk2019geographic].
+
+Food banks - sometimes also referred to as 'food pantries' and 'food shelves' - originated as a community response to aid those with inadequate food by voluntarily offering them meals and ingredients [@riches2002food; @loopstra2012relationship].
+Although in their origin food banks were meant to provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time many have evolved into a community practice to secure food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
 In Canada, the number of food banks has steadily increased in the past few decades [@wakefield2013sweet].
 The largest database of food banks and their use comes from the non-profit association Food Banks Canada (FBC), which conducts an annual assessment through its affiliated members.
-FBC's 2018 Hunger Count report^[https://foodbankscanada.ca/getmedia/241fb659-05f5-44a2-9cef-56f5f51db523/HungerCount-2018_FINAL_EN.pdf.aspx?ext=.pdf] (the most recent available) listed 1,830 member food banks across the country, and found that Canadians visited food banks 1.1 million times in March of 2018.
+FBC's 2018 Hunger Count report [@fbc2018] (the most recent available) listed 1,830 member food banks across the country, and found that Canadians visited food banks 1.1 million times in March of 2018.
 Of those accessing food banks, certain population characteristics tend to be over-represented compared to national totals from the 2016 Canadian Census of Population.
-According to FBC's 2018 data, single-adult households represent 45% of those utilizing food banks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those accessing food bank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
+According to FBC's 2018 data, single-adult households represent 45% of those utilizing food banks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those using food bank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
 In addition, 59% of households accessing food banks list social or disability assistance as their primary source of income.
-Similarly, using data from the 2011-2012 CCHS, Tarasuk et al.
-[-@tarasuk2019geographic] found higher odds of food insecurity amongst households relying on social assistance, those without a university degree or with children under the age of 18, and individuals that lived alone, renters, and those identifying as Aboriginal.
-While surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada [@tarasuk2014food], research from Ottawa [@enns2020experiences] and Vancouver [@black2020examining] suggests that long-term users tend to be older, have health or mobility challenges, live in large households, and are less likely to have employment income.
-
-Economic stress caused by the COVID-19 pandemic, rising unemployment rates, and changes in poverty levels have disrupted the food environment and led to higher rates of food insecurity [@niles2020early].
-Pandemic-related food security studies in the US have found a substantial increase in households experiencing food insecurity for the first time, and also in households experiencing more severe food insecurity than before [@niles2020early; @wolfson2020food].
-Most recently in May 2020, Canada recorded 14.7% of its population living in food insecurity in the past 30 days [@statisticscanada2020food].
-Considering the negative mental and physical health effects, increases in food insecurity rates due to the onset of COVID-19, signal a change in the food environment with potential damages to population health during the course of the pandemic and beyond [@niles2020early].
-Recent data^[https://www.foodbankscanada.ca/FoodBanks/MediaLibrary/COVID-Report_2020/A-Snapshot-of-Food-Banks-in-Canada-and-the-COVID-19-Crisis_Exec-Sum_EN.pdf] from FBC showed that 52% of member food banks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of the country.
+
+The COVID-19 pandemic has led to an increase in the number of households living in food insecurity in Canada.
+Survey results from Statistics Canada from May of 2020 suggest that 14.7% of the population was living in food insecurity in the past 30 days, up from 10.5% in 2017-2018 [@statisticscanada2020food].
+Recent data from FBC [@fbc2020] showed that 52% of member food banks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of the country.
 The pandemic also created significant staffing issues with 42% of food banks reporting a reduction in volunteers.
@@
 [-@paez2019demand].
-In order to avoid the multiple-counting in the summations, the population and the level of service need to be allocated _proportionally_.
+In order to avoid the multiple-counting in the summations, the population and the level of service need to be allocated *proportionally*.
 This is achieved by standardizing the weights as follows:
-$$
-w_{ij}^\text{st} = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
-$$
+
+$$
+w_{ij}^i = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
+$$
+
 \noindent and:
-$$
-w_{ji}^\text{st} = \frac{w_{ji}}{\sum_{j=1}^Jw_{ji}}
+
+$$
+w_{ij}^j = \frac{w_{ij}}{\sum_{j=1}^Jw_{ij}}
 $$
@@
 The standardized weights satisfy the following conditions:
-$$
-\sum_{i=1}^nw_{ij}^\text{st}=1
+
+$$
+\sum_{i=1}^nw_{ij}^i=1
 $$
@@
 \noindent and:
-$$
-\sum_{j=1}^Jw_{ji}^\text{st}=1
+
+$$
+\sum_{j=1}^Jw_{ij}^j=1
 $$
@@
 Since the population is allocated proportionally, its value is preserved:
-$$
-\sum_{i=1}^nP_iw_{ij}^\text{st}=P_i
+
+$$
+\sum_{i=1}^nP_iw_{ij}^i=P_i
 $$
@@
 \noindent as is the level of service:
-$$
-\sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
+
+$$
+\sum_{j=1}^JL_jw_{ij}^j=L_j
 $$
@@
 
-Data have been prepared for sharing as a data package^[https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing]. The contents of the data package are described next.
+Data have been prepared for sharing in the form of an open data product [see @arribas2021open] available in an (anonymous) [Drive Folder](https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing).
+The contents of the data package are described next.
 
@@
 To implement the accessibility calculations, we must select a distance-decay function.
-In this task we find limited support in the literature, which is mostly silent on the travel patterns of people who visit food banks.
+In this task we find limited support in the literature, which is mostly silent on the travel patterns of people who visit food banks and community food services.
 For this reason, we opt for a simple cumulative opportunities function as follows:
-$$
-w_{ij}=w_{ji}=
+
+$$
+w_{ij}=
 \begin{cases}
@@
 
-Using the 15-minute threshold, we find that the system-wide accessibility (interpreted as a provider-to-population ratio) was `r round(sum(results_pre$accessibility$accessibility), 3)` (food banks/service locations per low income household in the city) before COVID-19, but declined to `r round(sum(results_covid$accessibility$accessibility), 3)` during the pandemic.
-It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`\% of food facilities remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`\%, suggesting the location of food banks and related services plays an important role in serving those in need.
+Using the 15-minute threshold, we find that the system-wide accessibility was `r round(sum(results_pre$accessibility$accessibility), 3)` (food banks/service locations per low income household in the city) before COVID-19, but declined to `r round(sum(results_covid$accessibility$accessibility), 3)` during the pandemic.
+It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`% of facilities remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`%, suggesting the location of emergency and community food services plays an important role in serving those in need.
 
@@
 For the more suburban clusters of zones, the decrease in accessibility is derived from the closure of locations throughout the city reachable by car.
-In the cluster of central suburban zones for example, low-income households in the outer ring of zones that exhibit medium to high decreases in accessibility within this cluster appear to be largely auto-dependent in their tripmaking, which each each exhibiting between `r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(min(p_car)) *100)`-`r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(max(p_car)) *100)`\% of their modal split for car trips.
+In the cluster of central suburban zones for example, low-income households in the outer ring of zones that exhibit medium to high decreases in accessibility within this cluster appear to be largely auto-dependent in their tripmaking, which each each exhibiting between `r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(min(p_car)) *100)`-`r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(max(p_car)) *100)`% of their modal split for car trips.
 This results in the parcels within these zones having a large number of potentially accessible locations in the travel time matrix.
 But by extension, the change in accessibility over the pre- and during-COVID-19 time periods is affected not only by the closure of service locations proximate to the zones, but also the locations in the central city.
-The zone with the greatest decrease in accessibility within this cluster (-0.0009) has a high rate of car trips and connects to the most food bank locations in total as well as those that stayed open or closed.
-The exception to this is the zone in the middle of this cluster, which has a much higher proportion of transit use (`r round(modes %>% st_drop_geometry() %>% filter(TAZUID == 5116) %>% summarize(first(p_transit)) *100)`\%).
-This modal split weighting results in fewer food bank and service locations accessible from the parcels within this zone which, by extension, limits their access to closed locations reachable within 15 minutes in our calculations.
+The zone with the greatest decrease in accessibility within this cluster (-0.0009) has a high rate of car trips and connects to the most facility locations in total as well as those that stayed open or closed.
 
 In the cluster to the south-west, the decrease in accessibility is predominately driven by the closure of a high level-of-service Community Meals provider.
-However, like the more central suburban zones, low-income households within this cluster are also between 90\% to 100\% auto-dependent in their tripmaking in the TTS.
+However, like the more central suburban zones, low-income households within this cluster are also between 90% to 100% auto-dependent in their trip-making in the TTS.
 The story is similar for the zone located in the north-west that exhibits the greatest decrease in accessibility.
-Here, low-income households responding to the TTS conducted 100\% of their trips by car and, as a result, dwellings within this zone have access to the second-highest number of food bank and service locations within 15 minutes.
+Here, low-income households responding to the TTS conducted 100% of their trips by car and, as a result, dwellings within this zone have access to the second-highest number of food bank and service locations within 15 minutes.
 However, this also means zonal accessibilities are greatly affected by the number of closures throughout the city.
 Finally, in the city's north end and north-east zones, low-income households exhibit a mixture of tripmaking behaviour in the TTS.
-Households in some zones take transit more often and one zone in particular has 100\% of its trips by walking while in others, car use makes up 100\% of trips.
-For these zones, the decrease in accessibility tends to be a product of the closure of several inner-city food bank and service locations.
+Households in some zones take transit more often and one zone in particular has 100% of its trips by walking.
+For these zones, the decrease in accessibility tends to be a product of the closure of several inner-city food bank and service locations reachable by multiple modes.
 
@@
 Using data on low income individuals by age drawn from the TTS, Table \ref{tab:accessibility-by-age} shows the estimated number of people in each age group by their level of accessibility before and during the pandemic.
-Here, it is important to note that the quartiles are relative: people in the top 25\% of accessibility still have lower accessibility during the pandemic than before.
+Here, it is important to note that the quartiles are relative: people in the top 25% of accessibility still have lower accessibility during the pandemic than before.
 In reality, every population group is worse off during the pandemic in terms of their accessibility to food banks and services in the City of Hamilton.
 However, some age groups were affected more.
-As seen in the table, the distribution within quartiles for adults changed only slightly, and in fact there are fewer adults in the bottom three quartiles as some found themselves in the (deteriorated) top quartile of accessibility during the pandemic.
-The story is different for seniors, a greater number of whom are now in the third quartile, due to a combination of seniors facing a worse accessibility situation, and just a few moving from the bottom quartile (pre-pandemic) to the third during the pandemic.
-Those aged 18 and less appear to have fared worse, and the largest change is in the number of children who were in the third quartile before the pandemic and found themselves in the bottom accessibility class during the pandemic.
-These results suggest that, through a combination of the typical modes of transportation of lower income households and the spatial distribution of the population, the food bank/service locations closed had a differential impact that more greatly affected the youngest among the population in low income households.
+In terms of changes within the quartiles, the largest change for adults appears to be those moving from the first to the second quartile of accessibility during the pandemic.
+The story is generally similar for seniors, a greater number of whom are now in the second and third quartiles due to seniors facing a worse accessibility situation.
+Among those aged 18 and less, the largest change is in the number of children who were in the first and third quartiles before the pandemic and found themselves in the second accessibility class during the pandemic.
+However, when we compare the total population serviced before and during the pandemic, we see that a large number of children, adults, and seniors were no longer in the catchment areas of service locations (last row of Table \ref{tab:accessibility-by-age}).
+This accounts for the loss of population in the fourth quartiles for the different age groups during the pandemic.
+It is remarkable that despite the loss of population serviced, accessibility still declined for those still within the catchment regions of these services.
+
+These results suggest that, through a combination of the typical modes of transportation of lower income households and the spatial distribution of the population, the closure of emergency and communal food locations had a differential impact that more greatly affected the youngest and oldest among the population in low income households.
 
@@
         col.names = c("Accessibility Quartile",
-                      "Children (age <= 18)", 
+                      "Children (age $\\le$ 18)", 
                       "Adults (19-64)", 
-                      "Seniors (age >= 65)", 
-                      "Children (age <= 18)", 
+                      "Seniors (age $\\ge$ 65)", 
+                      "Children (age $\\le$ 18)", 
                       "Adults (19-64)", 
-                      "Seniors (age >= 65)"),
+                      "Seniors (age $\\ge$ 65)"),
         caption = "\\label{tab:accessibility-by-age}Population at each accessibility level by age group among members households with incomes less than CAD40,000.",
         escape = FALSE) %>%
-  kable_styling(bootstrap_options = c("striped")) %>%
+  kable_styling(latex_options = c("striped", "scale_down", "hold_position")) %>%
   add_header_above(c(" " = 1, "Pre-COVID-19" = 3, "During COVID-19" = 3)) %>%
+  row_spec(4, hline_after = TRUE) %>%
   footnote(general = c("Population values have been rounded.", 
@@
             change) %>%
-  kable("latex",
+  kable("html",
         booktabs = TRUE,
@@
 
-Conclusions
-===================
-
-Food insecurity is a significant issue for many Canadian households and is associated with reductions in nutrition and poor mental and physical health outcomes.
-While food bank use can provide some relief, the levels of need in the City of Hamilton were on the rise before the onset of the COVID-19 pandemic, with a troubling increase in the number of visits made by children.
-While temporary financial relief measures may have helped reduce food bank use in the summer of 2020, the pandemic has in all probability increased food insecurity for many households.
-To compound matters, the pandemic has also resulted in major disruptions, including to employment, mobility alternatives, and to food bank and food service delivery.
+# Conclusions
+
+Food insecurity is a significant issue for many Canadian households and while emergency and community food services can provide some relief, the COVID-19 pandemic has in all probability increased food insecurity for many households.
+To compound matters, the pandemic has also resulted in major disruptions, including to employment, mobility alternatives, and to emergency and community food services.
 In response, this research has sought to better understand accessibility to food banks and food service locations, as well as how the closure of some locations over the pre- and during-COVID time periods affected the potential for low-income households to reach these amenities.
-Using the balanced floating catchment area method to account for population demand and congestion effects at service points, we estimated multi-modal accessibility to food bank and food service locations for low-income households.
+
+Previous work has noted the important role of geography alongside other socio-economic and demographic indicators in household access to healthy food.
+The present papers is, to the best of our knowledge, among the first studies to focus on the geographic component of accessibility to emergency food services [@allen2021Changes].
+Using the balanced floating catchment area method to account for population demand and congestion effects at service points, we estimated multi-modal accessibility to emergency and community food service locations for low-income households.
 The weighting of travel time estimates by the modal split in different zonal geographies tailors the results to patterns of travel behaviour captured in the regional travel survey.
 Moreover, our parcel-level analysis presents a disaggregate approach to estimating accessibility based on the locations of residential parcels and dwellings.
-
-Our results show that while accessibility levels were lower in the city's more car-oriented suburban and rural areas to begin with, the closure of 14% of the city's food bank and low-cost or free food service locations during the pandemic resulted in an overall decrease in accessibility across the city.
+Beyond accounting for the inflation of demand that occurs in traditional floating catchment area methods, our application of the balanced floating catchment area approach offers a novel analysis of accessibility to emergency and community food services that is sensitive to the locations of low-income households, information on their age distributions and typical trip-making behaviour, the locations of services, their operations over time, and the characteristics of the city's multi-modal transportation network.
+
+Our results show that while accessibility levels were lower in the city's more car-oriented suburban and rural areas to begin with, the closure of 14% of the city's emergency and community food service locations during the pandemic resulted in an overall decrease in accessibility across the city.
 However, these effects were not uniform over space or for different population groups.
-Since the balanced floating catchment area method takes into account changes in demand and congestion for service providers, the closure of food bank and food service locations can reverberate throughout the catchment areas of the whole city.
+Since the balanced floating catchment area method takes into account changes in demand and congestion for service providers, the closure of some services reverberates throughout the catchment areas of the whole city.
 For some suburban zones, the closure of a relatively high level-of-service location results in the remaining services being spread over a larger population.
-In others, high auto dependence for trips leads to decreases in accessibility that accumulate due to the loss of several food bank and service locations initially reachable within 15 minutes by car.
-Reductions in accessibility in the city's more urban north end, where low-income households conduct higher proportions of trips by transit and walking, emphasize the importance of geographic proximity in the potential to reach food bank and service locations for these residents.
-Beyond geography, the results also highlight the differential impact of food banks closing during COVID-19 on population groups, with seniors and children most affected.
-
-In the absence of information regarding how food insecure households travel to food banks and related services, we examined accessibility to food banks using a 15 minutes (one-way) travel time threshold.
-The fact that we must rely on a standard created for accessible drinking water in the developing world only serves to highlight the tragedy of food insecurity in an affluent country like Canada.
-More broadly, it points to the absurd need to understand how a bad situation was made worse by the pandemic: in effect, the analysis reveals that disparities in accessibility to food banks and services predated the pandemic, that the pandemic contributed to the deterioration of these services, and that populations already in distress, particularly children, ended up in an even more adverse state.
-How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding what is an _acceptable_ or _sufficient_ level of service when it comes to food banks.
-Whether this is one facility per 1,000 population or three facilities per 1,000 population, or any other number, we know not.
-But like international human rights standards for accessibility to clean drinking water within 30 minutes (round-trip), this work strongly suggests that adopting better defined standards for access to food banks that move beyond generalized conceptualizations of "access" to consider the geographic component of accessibility (and its interactions with travel behaviour and individual/household socioeconomic and demographic traits) would make a meaningful contribution to understanding food access and "food bank deserts" for food insecure populations.
-
-In summary, while previous work has noted the important role of geography alongside other socio-economic and demographic indicators in household access to healthy food, this is, to the best of our knowledge, the first study that has focused on the geographic component of food bank accessibility.
-Beyond accounting for the inflation of demand that occurs in traditional floating catchment area methods, our application of the balanced floating catchment area approach offers a novel analysis of food bank and food service accessibility that is sensitive to the locations of low-income households, information on their age distributions and typical trip-making behaviour, food bank and food service locations and their operations over time, and the characteristics of the city's multi-modal transportation network.
-Despite the limitations noted above, the findings from this analysis can provide quantitative evidence to help urban planners and community administrators strategically allocate food bank services considering accessibility for different population groups, especially when considering the unequal distribution of impacts from the COVID-19 pandemic.
-
-In terms of future research, it is important to note that the degree to which our low-income cut-off of CAD\$40,000 reflects food insecurity in the different zones of the study area is not known.
-We also consider all food banks and food service locations equally.
+In others, high auto dependence for trips leads to decreases in accessibility that accumulate due to the loss of several locations initially reachable within 15 minutes by car.
+Reductions in accessibility in the city's more urban north end, where low-income households conduct higher proportions of trips by transit and walking, emphasize the importance of geographic proximity in the potential to reach service locations for these residents.
+Beyond geography, the results also highlight the differential impact of closures during COVID-19 on population groups, with seniors and children being the two most impacted groups.
+
+It is important to note that the degree to which our low-income cut-off of CAD\$40,000 reflects food insecurity in the different zones of the study area is not known.
+We also consider all emergency and community food service locations equally.
 Information on differential capacities at food provider locations is currently not collected, and given the closure of facilities was not possible to obtain.
@@
 
-References {#references .unnumbered}
-==========
+In the absence of information regarding how food insecure households travel to food banks and related services, we examined accessibility to food banks using a 15 minutes (one-way) travel time threshold.
+The fact that we must rely on a standard created for accessible drinking water in the developing world only serves to highlight the tragedy of food insecurity in an affluent country like Canada.
+More broadly, it points to the absurd need to understand how a bad situation was made worse by the pandemic: in effect, the analysis reveals that disparities in the need for emergency and community food services predated the pandemic, that the pandemic contributed to the deterioration of these services, and that populations already in distress, particularly children, ended up in an even more adverse state.
+How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding *acceptable* or *sufficient* level of service when it comes to emergency food services.
+
+While on the one hand this work suggests that inequities in the accessibility to emergency and community food services could be improved through accessibility standards that promote changes in the geographic distribution of service locations and transportation network characteristics, in fact, we would argue that the standard should be that no household faced food insecurity.
+As others have noted [e.g., @poppendieck1999sweet; @men2021food] the root of food insecurity is income poverty and unless it is eliminated, there will continue to be a place for emergency food and community food services.
+In addition to providing food, these services satisfy social needs by offering a social setting for seniors or by helping to connect households in need with longer term supports.
+From a food security perspective, on the other hand, these services should work only as a short term solution, and not as a semi-permanent feature of life for some of our fellow human beings.
+From a human rights perspective, long-term reliance on emergency food services should be as unacceptable in Canada as lack of clean drinking water within 30 minutes is elsewhere.
+Thus, while our analysis is valuable to map the suffering caused by food insecurity, from a policy perspective, maintaining a robust social safety net that includes Employment Insurance and paid sick days are better tools to reduce this suffering than increasing the accessibility of emergency food services for food insecure populations.
+
+# References {#references .unnumbered}

### 2021-08-31 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-title-page.Rmd
R2-Created highlights file, made small edits requested by editorial office on submission

--- 2021-04-28 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-title-page.Rmd
@@
     text: "Corresponding Author"
-abstract: |
-  In this paper we analyze the changes in accessibility to food banks and related services before and during the COVID-19 pandemic in the City of Hamilton, Ontario.
-Food banks and services are the last line of support for households facing food insecurity; as such, their relevance cannot be ignored in the midst of the economic upheaval caused by the pandemic.
-Our analysis is based on the application of balanced floating catchment areas, and concentrates on households with lower incomes (<CAD40,000, approximately the Low Income Cutoff Value for a city of Hamilton's size).
-We find that accessibility was low to begin with in suburban and exurban parts of the city; furthermore, about 14\% of locations originally available in Hamilton closed during the pandemic, further reducing accessibility.
-The impact of closures on the level of service of the remaining facilities, and on accessibility, was disproportionate, with system-wide losses exceeding 40\%.
-Those losses were geographically and demographically uneven.
-While every part of the city faced a reduction in accessibility, inner suburbs fared worse in terms of loss of accessibility.
-As well, children (age $\le 18$) appear to have been impacted the most.
-
 journal: "Social Science and Medicine"

### 2021-08-31 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-title-page.tex
R2-Created highlights file, made small edits requested by editorial office on submission

--- 2021-04-24 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-title-page.tex
@@
   \begin{abstract}
-  In this paper we analyze the changes in accessibility to food banks
-  and related services before and during the COVID-19 pandemic in the
-  City of Hamilton, Ontario. Food banks and services are the last line
-  of support for households facing food insecurity; as such, they have
-  become essential and their relevance cannot be ignored in the midst of
-  the economic upheaval caused by the pandemic. Our analysis is based on
-  the application of balanced floating catchment areas, and concentrates
-  on households with lower incomes (\$\textless CAD40,000, approximately
-  the Low Income Cutoff Value for a city of Hamilton's size).
-  Accessibility was low to begin with in suburban and exurban parts of
-  the city. We find that about 14\% of locations originally available in
-  Hamilton closed during the pandemic; the impact of those closures on
-  the level of service and accessibility was disproportionate, with
-  system-wide losses exceeding 40\%. Those losses were geographically
-  and demographically uneven. While every part of the city faced a
-  reduction in accessibility, inner suburbs fared worse in terms of loss
-  of accessibility. As well, children (age \(\le 18\)) appear to have
-  been impacted the most.
+
   \end{abstract}

### 2021-09-02 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton-tables.Rmd
Created png version of figures and separate file for tables as requested by editorial team on submission of R2

--- 2021-08-29 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
@@
 ---
-title: "Changes in accessibility to emergency and community food services during COVID-19 and implications for low income populations in Hamilton, Ontario"
-author:
-  - name: Author One
-    email: author1@example.com
-    affiliation: Some University
-  - name: Author Two
-    email: author2@example.com
-    affiliation: Another University
-    footnote: 1
-  - name: Author Three
-    email: author3@example.com
-    affiliation: Some University
-  - name: Author Four
-    email: author4@example.com
-    affiliation: Some University
-address:
-  - code: Some University
-    address: Department, Street, City, State, Zip
-  - code: Another University
-    address: Department, Street, City, State, Zip
-footnote:
-  - code: 1
-    text: "Corresponding Author"
-abstract: |
-  In this paper we analyze the changes in accessibility to emergency and community food services before and during the COVID-19 pandemic in the City of Hamilton, Ontario.
-Many of these food services are the last line of support for households facing food insecurity; as such, their relevance cannot be ignored in the midst of the economic upheaval caused by the pandemic.
-Our analysis is based on the application of balanced floating catchment areas and concentrates on households with lower incomes (<CAD40,000, approximately the Low Income Cutoff Value for a city of Hamilton's size).
-We find that accessibility was low to begin with in suburban and exurban parts of the city; furthermore, about 14\% of locations originally available in Hamilton closed during the pandemic, further reducing accessibility.
-The impact of closures on the level of service of the remaining facilities, and on accessibility, was disproportionate, with system-wide losses exceeding 39\%.
-Those losses were geographically and demographically uneven.
-While every part of the city faced a reduction in accessibility, inner suburbs fared worse in terms of loss of accessibility.
-As well, children (age $\le 18$) appear to have been impacted the most.
-journal: "Social Science and Medicine"
-date: "`r Sys.Date()`"
-bibliography: references.bib
-#linenumbers: true
-#numbersections: true
-csl: elsevier-harvard.csl
-output: rticles::elsevier_article
-header-includes:
-    - \usepackage[table]{xcolor}
+title: ""
+output: pdf_document
 ---
@@
 library(tidyverse)
+```
+
+```{r directory-for-figures, include=FALSE}
+ifelse(!dir.exists(file.path("./figures")), dir.create(file.path("./figures")), FALSE)
 ```
@@
 
-# Introduction
-
-Food insecurity is defined as an "inadequate or uncertain access to a sufficient quantity and/or adequate quality of food" due to a household's financial limitations [@enns2020experiences].
-This condition has been associated with reductions in nutritional outcomes [@bhattacharya2004poverty; @kirkpatrick2008food; @olson1999nutrition] and negative physical and mental health impacts in children and adults [@elgar2021relative; @jones2017food; @ramsey2011food; @seligman2010food; @stuff2004household].
-Over at least the past four decades food banks and related services have become an essential line of defense against food insecurity in Canadian communities [@riches2002food; @holmes2018nothing; @black2020examining; @tarasuk2020relationship].
-In this respect, Canada is not unlike numerous other wealthy countries where a systematic dismantling of the welfare state took place in the intervening period [@tarasuk2014food].
-
-The emergence of COVID-19, the worst public health crisis since the 1918 flu pandemic, has exposed important social and economic fault lines, and pre-existing patterns of inequality appear to have been exacerbated.
-Along several other dimensions of stress [e.g., accessibility to health care facilities, @ghorbanzadeh2021spatial; @pereira2021geographic], this seems to be the case for food insecurity as well [@laborde2020poverty].
-According to Statistics Canada [-@statisticscanada2020food], in the early stages of the pandemic almost 15% of individuals reported living in a household that faced food insecurity; the risk of food insecurity was substantially higher for households with children.
-The difference between households with and without children was significant, and 11.7% of households with children indicated that "food didn't last and [there was] no money to get more" sometimes or often, compared to 7.3% of households without children; likewise, 13% of households with children indicated that they "[c]ouldn't afford balanced meals" sometimes or often, compared to 8.8% of households without children.
-Additionally, Men and Tarasuk [-@men2021food] report that about 25% of individuals who experienced job insecurity (a relatively common occurrence during the pandemic) also experienced food insecurity associated with COVID-related disruptions to employment, financial hardship, and use of food charity.
-
-The impacts of food insecurity during the pandemic are alarming, since diet-related diseases, such as obesity, heart-disease, and diabetes, were already critical public health concerns in Canada prior to COVID-19 [@boucher2017ontario].
-While emergency food services are not necessarily a stable solution to food insecurity and in fact may encourage a retrenchment of neoliberal policy [@wakefield2013sweet], in reality provide a resource of last instance to households in precarious situations [@bazerghi2016role].
-As a mid-size city grappling with deindustrialization, Hamilton exhibits high rates of poverty and use of emergency food services.
-As recently as 2019, the Hamilton Hunger Report [@hfs2019] noted that food banks in the city recorded the highest number of visitors in the past 29 years; a rate of increase greater than population growth.
-Most troubling, approximately 40% of all visitors were children.
-
-It is known that urban food environments, within which people make their daily food choices, are essential in influencing eating behaviours and health outcomes, based on factors such as food availability, ease of geographic accessibility and socio-demographic variations [@paez2010relative; @vanderlee2017food; @widener2018spatial].
-However, while there is a wealth of literature that has examined the topic of geographic accessibility to healthy food through the "food desert" concept, there has been little research into accessibility to emergency and community food services.
-Previous work has explored differences in *accessing* food banks, such as how some households utilize food banks over short periods of time while others regularly utilize food banks as longer-term resource [e.g., @enns2020experiences].
-In addition, transportation and locational considerations have been raised as key issues in food bank accessibility in previous qualitative research [@smith2017food].
-However, we are not aware of any research that has focused on estimating or capturing this geographic component of accessibility.
-
-The study of place-based geographic accessibility is concerned with capturing the potential to reach destinations of value using the transportation network [@paez2012measuring].
-Indeed, the Government of Canada's recent Food Policy [@foodpolicy2019] has made "access" to healthy food a priority for Canadian communities and previous research suggests that such accessibility plays a key role in user satisfaction with food bank service delivery [@holmes2018nothing].
-However, as with research into the prevalence of food deserts, accessibility to food banks is unlikely to be evenly distributed, and variation throughout a city can be expected due to transportation network characteristics and the spatial distribution of service locations and the population they are meant to serve.
-Furthermore, policy responses to the COVID-19 pandemic likely have added to the distress of vulnerable households.
-Non-pharmaceutical interventions during the pandemic involving restrictions in mobility have increased the friction of travel, in particular by transit on which low income populations are more reliant [e.g., @deweese2020tale].
-At the same time, the pandemic has created additional stress for the operators of food banks through disruptions in the supply chain [e.g., @mckay2021exploring] as well as concerns surrounding the delivery of service in safe conditions and possible cancellation of food service programs.
-
-For this study, we aim to look at how the landscape of emergency food and related services (e.g.
-low-cost or free meal service providers) available in Hamilton, Ontario, changed during the pandemic.
-Did the number of open services diminish?
-If so, what was the accessibility to emergency and community food services before the pandemic from the perspective of low income households, and how has it changed during the pandemic with respect to geographic access and congestion at remaining sites?
-And finally, who are most likely to have been impacted by changes in the accessibility landscape?
-This paper first looks at the distribution of emergency and community food services before and during the pandemic.
-Then, we use the balanced floating catchment area approach of Paez et al.
-[-@paez2019demand] to investigate the accessibility situation.
-For this, we adopt a fully disaggregated approach based on parcel-level data.
-Socio-economic and demographic data are drawn from the latest Census of Canada (2016), whereas travel information is from the most recent regional travel survey from 2016.
-This paper follows reproducible research recommendations [see @brunsdon2020opening], and the research was conducted using open source tools for transportation analysis [@lovelace2021open].
-The data and code needed to reproduce the analysis are available in an (anonymous) [Drive Folder](https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing).
-
-# Food Insecurity and Emergency and Communal Food Services in Canada
-
-Food insecurity is the inability to acquire and consume an adequate amount or good quality food, leading to inadequate nutrient intake [@kirkpatrick2008food] and poorer physical and mental health outcomes [@ramsey2011food; @seligman2010food; @stuff2004household].
-In this regard, food insecurity is a major population health concern, particularly among Canadians at socio-economic disadvantage [@bazerghi2016role].
-Official government surveys such as the Household Food Security Survey Module (HFSSM), the Canadian Community Health Surveys (CCHS), the Longitudinal and International Study of Adults (LISA), and official classifications determined by Health Canada in relation to socio-demographic variables offer some insight into food insecurity in Canada [@kirkpatrick2008food; @tarasuk2009household; @gundersen2018food].
-
-Nationally, analysis of the 2011-2012 CCHS has previously revealed that food insecurity impacts approximately 12.3% of Canadian households [@tarasuk2014food].
-Using the same data, Tarasuk et al.
-[-@tarasuk2019geographic] found higher odds of food insecurity amongst households relying on social assistance, those without a university degree or with children under the age of 18, and individuals that lived alone, renters, and those identifying as Aboriginal.
-While surveys revealed that only 20 to 30 percent of those experiencing food insecurity were found to frequent food banks in Canada [@tarasuk2014food], pre-pandemic research from Ottawa [@enns2020experiences] and Vancouver [@black2020examining] suggests that long-term users tend to be older, have health or mobility challenges, live in large households, and are less likely to have employment income.
-In terms of geography, previous research conducted at the provincial scale using data from the 2011-2012 CCHS found that the prevalence of food insecurity ranged across the country from 11.8% of households in Ontario to 41% of households in Nunavut [@tarasuk2019geographic].
-
-Food banks - sometimes also referred to as 'food pantries' and 'food shelves' - originated as a community response to aid those with inadequate food by voluntarily offering them meals and ingredients [@riches2002food; @loopstra2012relationship].
-Although in their origin food banks were meant to provide a temporary solution to accommodate those in hunger due to job retrenchments and economic downfalls since the 1980s, over time many have evolved into a community practice to secure food supplies for those in need [@loopstra2012relationship; @wakefield2013sweet].
-In Canada, the number of food banks has steadily increased in the past few decades [@wakefield2013sweet].
-The largest database of food banks and their use comes from the non-profit association Food Banks Canada (FBC), which conducts an annual assessment through its affiliated members.
-FBC's 2018 Hunger Count report [@fbc2018] (the most recent available) listed 1,830 member food banks across the country, and found that Canadians visited food banks 1.1 million times in March of 2018.
-Of those accessing food banks, certain population characteristics tend to be over-represented compared to national totals from the 2016 Canadian Census of Population.
-According to FBC's 2018 data, single-adult households represent 45% of those utilizing food banks despite making up 28% of Canada's population, 19% are single-parent households (compared to 10% nationally), and 35% of those using food bank services are children aged 0-18 even though their share of Canada's national population is approximately 20%.
-In addition, 59% of households accessing food banks list social or disability assistance as their primary source of income.
-
-The COVID-19 pandemic has led to an increase in the number of households living in food insecurity in Canada.
-Survey results from Statistics Canada from May of 2020 suggest that 14.7% of the population was living in food insecurity in the past 30 days, up from 10.5% in 2017-2018 [@statisticscanada2020food].
-Recent data from FBC [@fbc2020] showed that 52% of member food banks reported an increase in usage in March of 2020 when initial lockdown restrictions were put in place across much of the country.
-The pandemic also created significant staffing issues with 42% of food banks reporting a reduction in volunteers.
-However, 53% of food banks later reported a decrease in use into the summer of 2020 which FBC members attributed to emergency financial support programs from the federal government.
-Nevertheless, some of these benefit programs were temporary.
-Although more recent statistics on food bank use in Hamilton in 2020 and 2021 are not yet available, data from the Daily Bread Food Bank [@dbfb2020] in neighbouring Toronto for August 2020 shows visits climbing 51% year-over-year in that city, which suggests that many households in Hamilton are likely to turn to food bank services to meet their needs.
-
-Beyond traditional conceptualizations of food banks as providers of emergency food assistance, other community food services also play an important role in decreasing food insecurity.
-The scope and objectives of food banks can vary by region and by country, and these organizations can include not only prepared meals and aliments for emergency food supply, but also shared spaces to connect in community gardens and community kitchens [@wakefield2013sweet].
-However, the efficacy of these programs in reducing food insecurity differs by the type of service offered.
-For example, previous qualitative research in the Toronto region has questioned the capacity of community kitchens to improve the food security of low-income households due to their limited scale of operations, and un-subsidized kitchens were found to be particularly inaccessible to to families living in severe poverty [@tarasuk1999qualitative].
-However, other food access options such as no-cost or low-cost meals provided through community meals or congregate dining play an important role in decreasing food insecurity.
-Research in Minnesota found that seniors experiencing food insecurity valued congregate dining for providing affordable meals and a space for social gathering [@oemichen2016investigation].
-Furthermore, because seniors paid for the meal, there was no stigma attached to the use of these services compared to traditional food-purchasing assistance such as the Supplemental Nutrition Assistance Program.
-
-While previous research has examined the characteristics of individuals and households accessing emergency and community food services, the locational or transportation accessibility aspect of food bank access is not well understood.
-A wealth of literature examining the food desert concept suggests that, in addition to socio-economic and demographic factors, location and transportation networks play a key role in a household's accessibility to healthy foods [@paez2010relative; @vanderlee2017food; @widener2018spatial].
-For food banks specifically, previous qualitative research in Ontario by Smith-Carrier et al.
-[-@smith2017food] has noted that "transportation can be challenging, particularly if the food bank is situated in a remote location" (p.
-32).
-Particularly, it appears that participants experience challenges with the "inordinate amount of time necessary to obtain food, and difficulties associated with transportation" (p.
-39).
-Users of food banks, according to this research, rely on a variety of modes of transportation to access services.
-Consequently, the location of facilities matters; in the words of an interviewee: "I wish it [the food bank] was a little more centrally located.
-Because if I didn't have a bike I'd have to walk it all the way out there and back.
-I wonder about people who don't" (p.
-39).
-
-To offer greater insight into the role of transportation and location in food bank accessibility, this research examines how geographic accessibility to food banks and food services changed in Hamilton during the COVID-19 pandemic.
-
-# Methods and Materials
-
-## Methods
-
-For the research in this paper we adopt the balanced floating catchment area approach of Paez et al.
-[-@paez2019demand].
-This method for estimating accessibility is a form of the widely-used two-stage floating catchment area method [@radke2000spatial; @luo2003measures].
-Floating catchment areas are used to estimate accessibility when there are potential congestion effects, and operate by calculating first the *demand* for spatially distributed services.
-The demand (usually the number of people who require a service) is used to calculate a level of service.
-In a second step, the level of service is allocated back to the population.
-Demand and level of service are allocated using some form of distance-decay to embody the geographical principle that, given a choice, people prefer to travel less than more when reaching destinations.
-
-More formally, the first step of this method is as follows:
-
-$$
-L_j = \frac{S_j}{\sum_{i=1}^nP_iw_{ij}}
-$$
-
-\noindent where $S_j$ is the level of supply at location $j$, in simplest terms whether a service point is present (i.e., $S_j=1$) or not (i.e., $S_j=0$); $P_i$ is the population at location $i$ that demands the service; and $w_{ij}$ is a weight, typically a function of the distance between locations $i$ and $j$.
-$L_j$ is the level of service at location $j$ and it is the inverse of the number of people that need to be serviced.
-
-The second step in this process is then summing the level of service that each population unit can reach, according to the distance-decay weight:
-
-$$
-A_i = \sum_{j=1}^JL_jw_{ji}
-$$
-
-\noindent where $A_i$ is the accessibility to the service, which is in the same units as the level of service: as the inverse of the population being serviced.
-When the population being serviced is low accessibility is high (i.e., there is little competition for the service), and viceversa.
-
-Floating catchment area methods are prone to overestimation of the population and the level of service due to multiple-counting.
-The population at $P_i$ is allocated to *every* service point $j$ for which $w_{ij}>0$.
-Similarly, the level of service at $LOS_j$ is allocated to *every* population point for which $w_{ji}>0$.
-This inflation effect has been known for several years, and several modifications have been proposed to mitigate it [e.g., @wan2012three; @delamater2013spatial].
-A definitive solution to this issue was presented by Paez et al.
-[-@paez2019demand].
-In order to avoid the multiple-counting in the summations, the population and the level of service need to be allocated *proportionally*.
-This is achieved by standardizing the weights as follows:
-
-$$
-w_{ij}^i = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
-$$
-
-\noindent and:
-
-$$
-w_{ij}^j = \frac{w_{ij}}{\sum_{j=1}^Jw_{ij}}
-$$
-
-The standardized weights satisfy the following conditions:
-
-$$
-\sum_{i=1}^nw_{ij}^i=1
-$$
-
-\noindent and:
-
-$$
-\sum_{j=1}^Jw_{ij}^j=1
-$$
-
-Since the population is allocated proportionally, its value is preserved:
-
-$$
-\sum_{i=1}^nP_iw_{ij}^i=P_i
-$$
-
-\noindent as is the level of service:
-
-$$
-\sum_{j=1}^JL_jw_{ij}^j=L_j
-$$
-
-## Study Area
-
-With a population of around 540,000, the City of Hamilton is the fourth largest city in Ontario.
-It has historically been home to major manufacturing industries but de-industrialization that has occurred over the past several decades has led Hamilton to become one of the most highly divided cities in Ontario, with a significant proportion of its residents living at or below Canada's poverty level [@deluca2012code; @jakar2019turning; @latham2007determinants].
-The Hamilton Community Foundation [@hcf2018] reported that based on the Low-Income Cut-Off, Hamilton recorded a poverty rate of 16.7% in 2016, which was well above the average rate of Ontario (13.7%) and the average national rate (12.8%).
-According to data from Hamilton Food Share [@hfs2019], approximately 23,000 individuals accessed food banks in the city in March of 2019.
-Within this total is 9,125 visits by children (minors up to 18 years old), up from 8,278 the year before.
-Feed Ontario, the province's largest collective of hunger-relief organizations, found that on a per-capita basis, the level of need in the inner core of central Hamilton was second highest in Ontario [@feedontario2019].
-
-Geographically, the "old" City of Hamilton was amalgamated with several of its surrounding municipalities in 2001, with the city now featuring a mix of urban, suburban, exurban, and rural areas.
-Lower-cost housing proximate to the city's industrial north end has traditionally attracted immigrants and less-affluent residents compared to the city's wealthier suburbs.
-However, the decentralization of population from the inner core has led to challenges in transit connectivity to amenities and services and the proportion of auto users compared to transit users remains very high [@behan2008smart; @topalovic2012light].
-In addition, the city is separated geographically by the Niagara Escarpment.
-With sections of rocky cliff that approach 100m in height, the escarpment presents a significant challenge for promoting active travel and transport connections between "mountain" and "lower city" neighbourhoods.
-Taken together, the high level of food need, population locations, and transportation network characteristics combine to inform spatial accessibility to food banks and food services in the city.
-
-## Data
-
-Data have been prepared for sharing in the form of an open data product [see @arribas2021open] available in an (anonymous) [Drive Folder](https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing).
-The contents of the data package are described next.
-
-### Statistics Canada
-
-Population and income statistics for 2016 were retrieved at the level of Dissemination Areas (DAs) using the package `cancensus` [@vonBergmann2021cancensus].
-DAs are the smallest publicly available census geography in Canada.
-Income data corresponds to the count of households by different total income groupings.
-
-### Origins: Residential parcels
-
-We converted all recorded residential land parcels in the City of Hamilton to points on the road network.
-Each point includes information about the number of residential units in the parcel.
-Next, we define low-income households as those having a total income of less than CAD40,000, which is approximately the mid-point of the low income cut-off (LICOs) for families in Canadian cities with populations greater than 500,000 in 2016, to match other Census data [@statisticscanada2020licos].
-We then "populate" each residential unit with the probability of being a low-income household based on the counts of households by income groups in the DA in which the parcel is located.
-While this method assumes a constant probability of low-income household status for all residential units in a DA, the parcel-level analysis affords a high level of spatial disaggregation for the accessibility analysis.
-
-### Destinations: Food Banks and Food Service Locations
-
-The locations of emergency and community food services were obtained from the Hamilton Public Library's Food Access Guide [@hpl2021].
-The guide was updated in April of 2021 to indicate any change affected on the services due to the pandemic.
-This includes modified business hours, a need to make reservations before frequenting, and locations that have completely shut down in consequence.
-Table 1 defines each service type and the number of locations pre- and during the COVID-19 pandemic.
-While some food bank services have a specific target population, such as prioritizing families with young children aged between 0 and 3 or accepting only those providing proof of low-income status through housing and utility statements, all the food services indicated below are designed to accommodate those in need of food at zero to low cost.
-With our focus on food banks and food services that offer free or low-cost meals at particular locations, we first removed services such as Meals on Wheels and other food access services such as food box, community kitchens, student nutrition programs, and shopping and transportation.
-With some providers offering different food services at the same location (e.g.
-food bank with free and community meal services), and some of these services closing after the onset of the COVID-19 pandemic, we opted to geocode based on the service type.
-On the other hand, two free meal services held on different days at the same location were collapsed into a single service point for the accessibility analysis.
-Additional details on the operations of individual facilities is not publicly available and with the changes in operations it proved unfeasible to collect it.
-For this reason, the analysis to follow is of accessibility to the location of food banks and services, but not to specific services (e.g., breakfasts vs.
-food boxes).
-
 ```{r table-food-bank-info, echo=FALSE, results='asis'}
@@
   column_spec(5, width = "15em")
-```
-
-### Routing and travel time tables
-
-Travel time tables for three modes (car, transit, walking) were computed using the parcels as the origins and the locations of the community and emergency food service locations as the destinations.
-For routing, the package `r5r` [@pereira2021r5r] was used with a network extract for the City of Hamilton from OpenStreetMaps and the General Transit Feed Specification (GTFS) files for the Hamilton Street Railway, the local transit operator, as well as for Burlington Transit, which operates some service in the city.
-For transit routing purposes we used maximum travel time values of 300 min and a 2,000 m cap on walking distance: any destination that exceeded these thresholds was ignored.
-The departure time used for routing was 8:00AM on March 30, 2021 to reflect transit service around the morning service peak on a typical Tuesday.
-
-### Transportation Tomorrow Survey
-
-We used the Data Retrieval System of the Transportation Tomorrow Survey (TTS)\footnote{\url{http://dmg.utoronto.ca/}} to download cross-tabulations of: 1) primary mode of travel per trip by income by place of residence; and 2) age by income by place of residence.
-These data are from the 2016 Survey (the most recent available).
-The data are geocoded at the level of Traffic Analysis Zones (TAZ) using the most recent zoning system from 2006 and expansion factors are applied to weight the trips .
-Each parcel point is populated with the proportion of trips by three modes of travel: car (as driver or passenger), transit, and walk.
-
-### Expected Travel Times
-
-Once we obtained travel time tables with population (number of households) and proportion of trips by mode, we calculated the expected travel time $ett$ from each parcel $i$ to a food bank or food service location $j$ as follows:
-
-$$
-ett_{ij} = p^c_i\cdot tt^c_{ij} + p^t_i\cdot tt^t_{ij} + p^w_i\cdot tt^w_{ij}
-$$
-
-\noindent where $p^k_i$ is the proportion of trips by mode $k$ in the TAZ of parcel $i$, and $tt^k_{ij}$ is the travel time from parcel $i$ to the food bank.
-In other words, the expected travel time reflects the weighted average of travel times to the food bank, with the weights given by the expected modal split of trips made by low-income households in the TAZ per the TTS data.
-
-# Results and Discussion
-
-```{r accessibility-threshold-sensitivity, include=FALSE}
-## Analysis
-# Testing different thresholds to check how accessibility changes. Here, the thresholds are varied from 2 min to 30 min in 2 min step increments
-threshold_seq <- seq(from = 2, to = 30, by = 1)
-
-# Initialize a data frame to store results
-accessibility_test <- data.frame(threshold = threshold_seq, 
-                                 accessibility = numeric(length(threshold_seq)))
-
-# Calculate the system-wide accessibility according to each threshold
-for(i in 1:length(threshold_seq)){
-  results_test <-  ttm %>% 
-    transmute(UID, OBJECTID, travel_time = e_travel_time, population = income_less40k, supply = 1) %>%
-    b2sfca(threshold = i )
-  accessibility_test$accessibility[i] <- sum(results_test$accessibility$accessibility)
-}
 ```
@@
 
-Figure \ref{fig:foodbanks} shows the location of food banks and services in the City of Hamilton and their status.
-Before the pandemic there were `r nrow(foodbank_locations)` of which `r nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))` (`r round(nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations) * 100, 2)`%) closed during the pandemic.
-As shown in the figure, food services tend to be predominantly located in the central parts of the city.
-This is not surprising: population density is high there, and it is also the part of the city where lower income households are more numerous in absolute and relative terms (see Figure \ref{fig:low-income-households}).
-Alas, this is also the part of the city where most of the closures during the pandemic happened.
-
-```{r plot-location-foodbanks, echo=FALSE, fig.cap="\\label{fig:foodbanks}Location of food banks/services and operation status; the dotted box is an inset of the central part of the City of Hamilton"}
-## Analysis
-# Plot location of foodbanks
-# Create a bounding box for an inset
-inset_box <- st_bbox(c(xmin = -79.975,
-                       xmax = -79.775, 
-                       ymax = 43.2, 
-                       ymin = 43.3), 
-                     crs = st_crs(4326)) %>%
-  st_as_sfc()
-
-fb <- ggplot() +
-  geom_sf(data = hamilton_boundary,
-          fill = "lightgray") +
-  geom_sf(data = foodbank_locations,
-          aes(size = COVIDClose,
-              shape = COVIDClose)) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_shape_manual(name = "Closed during COVID-19",
-                     values = c("Yes" = 4, "No" = 1)) +
-  scale_size_manual(values = c("Yes" = 3, "No" = 2)) +
-  guides(size = "none") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90),
-        legend.position = "none")
-
-fb_inset <- ggplot() +
-  geom_sf(data = hamilton_boundary %>% 
-            st_transform(4326),
-          fill = "lightgray") +
-  geom_sf(data = foodbank_locations,
-          aes(size = COVIDClose,
-              shape = COVIDClose)) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_shape_manual(name = "Closed during COVID-19",
-                     values = c("Yes" = 4, "No" = 1)) +
-  scale_size_manual(values = c("Yes" = 3, "No" = 2)) +
-  xlim(-79.975, -79.775) +
-  ylim(43.2, 43.3) +
-  guides(size = "none") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90),
-        legend.position = "bottom")
-
-fb / fb_inset
-```
-
-```{r plot-low-income-households, echo=FALSE, fig.cap="\\label{fig:low-income-households}Number and proportion of households with incomes less than CAD40,000."}
-lih <- ggplot() +
-  geom_sf(data = data_da_2016,
-          aes(fill = income_less40k),
-          color = NA) +
-  geom_sf(data = hamilton_boundary,
-          fill = NA) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_fill_viridis_c(name = "Number of households", 
-                       option = "A",
-                       direction = -1,
-                       breaks = c(200, 400, 600)) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "bottom")
-
-p_lih <- ggplot() +
-  geom_sf(data = data_da_2016,
-          aes(fill = p_inc_less40k),
-          color = NA) +
-  geom_sf(data = hamilton_boundary,
-          fill = NA) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_fill_viridis_c(name = "Proportion of households", 
-                       option = "A",
-                       direction = -1,
-                       breaks = c(0.25, 0.50, 0.75)) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "bottom")
-
-lih / p_lih
-```
-
-To implement the accessibility calculations, we must select a distance-decay function.
-In this task we find limited support in the literature, which is mostly silent on the travel patterns of people who visit food banks and community food services.
-For this reason, we opt for a simple cumulative opportunities function as follows:
-
-$$
-w_{ij}=
-\begin{cases}
-1 & \text{ if } ett_{ij}\le \delta\\
-0 & \text{ otherwise}
-\end{cases}
-$$
-
-\noindent where $ett_{ij}$ is the multimodal expected travel time as described previously, and $\delta$ is a travel threshold.
-When the expected travel time exceeds this threshold, a facility is no longer considered accessible.
-Moreover, the weights are standardized for the balanced floating catchment area approach.
-
-Figure \ref{fig:sensitivity-analysis} shows the results of conducting a sensitivity analysis of the system-wide accessibility as we vary the threshold (considering the situation before the pandemic).
-There is a clear pattern whereby more strict values of $\delta$ are associated with higher levels of system-wide accessibility: while increases in accessibility that result from decreases in the travel time window might seem counter-intuitive, this is a result of lower *congestion*, since fewer households are serviced and thus competition for the same resources is more limited.
-System-wide accessibility declines with higher values of $\delta$: as more households are serviced, congestion grows and the level of service declines, although this happens at a declining rate.
-We are not aware of any research that explains how long people are expected to travel for food banks, but we note that in developing countries, accessible sources of drinking water are those that can be reached in less than 30 minutes [round trip, see @world2019progress].
-There is no reason why people in affluent countries should be expected to spend more time travelling for a basic necessity such as food.
-Accordingly, we adopt a 15-minute threshold for the analysis (representing a one-way trip).
-This threshold is also approximately where the rate of change in accessibility slows down.
-
-```{r plot-results-sensitivity-analysis, out.width="60%", echo=FALSE, fig.cap="\\label{fig:sensitivity-analysis}Accessibility as a function of threshold"}
-acc_sensitivity <- ggplot(accessibility_test, 
-                          aes(x = threshold, y = accessibility)) +
-  geom_line() +
-  ylab("Accessibility") +
-  xlab("Threshold (in min)") +
-  theme_minimal()
-
-acc_sensitivity
-```
-
-Using the 15-minute threshold, we find that the system-wide accessibility was `r round(sum(results_pre$accessibility$accessibility), 3)` (food banks/service locations per low income household in the city) before COVID-19, but declined to `r round(sum(results_covid$accessibility$accessibility), 3)` during the pandemic.
-It is striking that although almost `r round((1 - nrow(foodbank_locations %>% filter(COVIDClose == "Yes"))/nrow(foodbank_locations)) * 100)`% of facilities remained in operation during the pandemic, there was a loss of accessibility greater than `r abs(round((sum(results_covid$accessibility$accessibility) - sum(results_pre$accessibility$accessibility))/sum(results_pre$accessibility$accessibility) * 100))`%, suggesting the location of emergency and community food services plays an important role in serving those in need.
-
 ```{r join-zonal-ids-to-accessibility-table, include=FALSE}
@@
 
-Turning to the location of individual facilities, the levels of service offered before and during the pandemic are shown in Figure \ref{fig:levels-of-service}.
-The level of service is functionally the inverse of the number of low-income households in the travel-mode weighted travel time catchment area of the facilities (this is because $S_j=1 \forall j$, i.e., each location represents a "capacity" of 1).
-Higher values mean that a facility is expected to service fewer households.
-Conversely, lower values indicate greater congestion.
-
-The general pattern of the levels of service is similar before and during the pandemic, with lower values in the center of the city where low-income households exhibit multimodal trip patterns that favour proximate service locations.
-Three more peripheral facilities towards the south of the city have moderate levels of service, presumably because they are expected to service relatively suburban/exurban populations generally reliant on automobiles for travel.
-During the pandemic, however, the levels of service dropped, in some cases quite substantially.
-The pattern of the losses in level of service, moreover, is not uniform.
-The upper pane of Figure \ref{fig:levels-of-service-changes} shows that the peripheral facilities in the suburban/exurban parts of the city saw major declines during the pandemic as more urban locations closed and demand increased for the remaining locations.
-Further, the inset map shows that the levels of service also deteriorated in the central part of the city.
-However, the loss of level of service was not as large in the core (where most of the food banks/services are found), but instead was more marked in the inner ring around the core, where facilities may have faced greater demand from both central city and suburban populations after the closure of service locations during the pandemic.
-
-```{r plot-levels-of-service, echo=FALSE, fig.cap="\\label{fig:levels-of-service}Levels of service at each facility pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
-## Analysis
-# Plot levels of service before and during covid
-
-# Pre-covid
-los_pre <- ggplot() +
-  geom_sf(data = hamilton_boundary) +
-  geom_sf(data = los %>%
-            filter(timing == "Pre-covid"),
-          aes(color = los,
-              size = los),
-          alpha = 0.5) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  scale_color_viridis_c(name = "Level of service \npre-COVID-19",
-                        option =  "A", 
-                        direction = -1,
-                        breaks = c(0.0010,
-                                   0.0020,
-                                   0.0040,
-                                   0.0060,
-                                   0.0080),
-                        labels = c("0.0010",
-                                   "0.0020",
-                                   "0.0040",
-                                   "0.0060",
-                                   "0.0080")) +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  guides(size = "none", shape = "none") +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-# During covid
-los_covid <- ggplot() +
-  geom_sf(data = hamilton_boundary) +
-  geom_sf(data = los %>%
-            filter(timing == "During covid"),
-          aes(color = los,
-              size = los),
-          alpha = 0.5) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  scale_color_viridis_c(name = "Level of service \nduring COVID-19",
-                        option =  "A", 
-                        direction = -1,
-                        breaks = c(0.00100,
-                                   0.00200,
-                                   0.00400,
-                                   0.00600,
-                                   0.00800),
-                        labels = c("0.0010",
-                                   "0.0020",
-                                   "0.0040",
-                                   "0.0060",
-                                   "0.0080")) +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  guides(size = "none", shape = "none") +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-los_pre / los_covid
-```
-
-```{r plot-levels-of-service-changes, echo=FALSE, fig.cap="\\label{fig:levels-of-service-changes}Changes in levels of service at each facility from pre-COVID-19 to during COVID-19."}
-# This figure shows the places where the level of service deteriorated most (excluding the places that closed, where it vanished).
-los_change <- los %>%
-  st_drop_geometry() %>%
-  pivot_wider(names_from = timing, 
-              values_from = los, 
-              values_fill = 0) %>%
-  filter(`During covid` > 0) %>%
-  mutate(los_d = `During covid` - `Pre-covid`) %>%
-  left_join(foodbank_locations %>%
-              transmute(OBJECTID = ID,
-                        geometry),
-            by = "OBJECTID") %>%
-  st_as_sf() %>% 
-  left_join(foodbank_locations %>% st_drop_geometry() %>% transmute(OBJECTID = ID, TYPE), by = "OBJECTID")
-
-lc <- ggplot() +
-  geom_sf(data = hamilton_boundary) +
-  geom_sf(data = los_change, 
-          aes(color = los_d,
-              size = los_d),
-          alpha = 0.3) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_color_viridis_c(name = "Change in level of service",
-                        option =  "A", 
-                        direction = 1,
-                        breaks = c(-0.0001,
-                                   -0.0002,
-                                   -0.0004,
-                                   -0.0006,
-                                   -0.0008),
-                        labels = c("-0.0001",
-                                   "-0.0002",
-                                   "-0.0004",
-                                   "-0.0006",
-                                   "-0.0008")) +
-  scale_size(trans = 'reverse') +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  guides(size = "none", shape = "none") +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "none")
-
-lc_inset <- ggplot() +
-  geom_sf(data = hamilton_boundary %>% 
-            st_transform(4326)) +
-  geom_sf(data = los_change, 
-          aes(color = los_d,
-              size = los_d),
-          alpha = 0.7) +
-  geom_sf(data = foodbank_locations,
-          aes(shape = COVIDClose)) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_color_viridis_c(name = "Change in level of service",
-                        option =  "A", 
-                        direction = 1,
-                        breaks = c(-0.0001,
-                                   -0.0002,
-                                   -0.0004,
-                                   -0.0006,
-                                   -0.0008),
-                        labels = c("-0.0001",
-                                   "-0.0002",
-                                   "-0.0004",
-                                   "-0.0006",
-                                   "-0.0008")) +
-  scale_size(name = "Change in level of service",
-             trans = 'reverse',
-             breaks = c(-0.0001,
-                        -0.0002,
-                        -0.0004,
-                        -0.0006,
-                        -0.0008),
-             labels = c("-0.0001",
-                        "-0.0002",
-                        "-0.0004",
-                        "-0.0006",
-                        "-0.0008")) +
-  scale_shape_manual(values = c("Yes" = 4, "No" = 1)) +
-  xlim(-79.975, -79.775) +
-  ylim(43.2, 43.3) +
-  guides(color = guide_legend(title.position = "top"), 
-         #size = FALSE, 
-         shape = "none") +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "bottom")
-
-lc / lc_inset
-```
-
-To further elucidate this issue, we now turn to the results of the accessibility analysis.
-As with the level of service of individual facilities, the general pattern of accessibility before and during the pandemic is similar.
-Figure \ref{fig:accessibility} reveals that, compared with the outer rural zones, the more urban zones of the city generally exhibit higher accessibility to food banks and food service locations.
-However, the pattern is not particularly smooth - this is largely attributable to the weighting of travel times by mode of transportation according to the trip patterns of low-income household respondents captured by the TTS.
-For example, in zones where low-income households make a high proportion of trips by walking, access to food bank locations by walking is afforded a concomitantly high weight in our calculations of travel time compared to transit or car travel.
-From this, highly-accessible locations result from a mix of characteristics: low-income households in locations where travel options that align with zonal modal split are available to connect them to food bank locations with high levels of service within 15 minutes.
-This seems to track with the experience of some users of these services, as reported by Smith-Carrier et al.
-[-@smith2017food].
-
 ```{r aggregate-accessibility-by-taz, include=FALSE}
@@
   st_as_sf()
-```
-
-```{r plot-accessibility, echo=FALSE, fig.align = 'center', fig.cap="\\label{fig:accessibility}Accessibility by traffic analysis zone pre-COVID-19 (top panel) and during COVID-19 (bottom panel)."}
-## Analysis
-# Plot accessibility before and during covid:
-acc_pre <- ggplot() +
-  geom_sf(data = hamilton_boundary,
-          fill = NA) +
-  geom_sf(data = accessibility_taz %>%
-            filter(timing == "Pre-covid"),
-          aes(fill = accessibility)) +
-  scale_fill_viridis_c(name = "Accessibility \npre-COVID-19",
-                       option =  "A", 
-                       direction = -1,
-                       breaks = c(0.0030,
-                                  0.0024,
-                                  0.0018,
-                                  0.0012,
-                                  0.0006),
-                       labels = c("0.0030",
-                                  "0.0024",
-                                  "0.0018",
-                                  "0.0012",
-                                  "0.0006")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-acc_covid <- ggplot() +
-  geom_sf(data = hamilton_boundary,
-          fill = NA) +
-  geom_sf(data = accessibility_taz %>%
-            filter(timing == "During covid"),
-          aes(fill = accessibility)) +
-  scale_fill_viridis_c(name = "Accessibility \nduring COVID-19",
-                       option =  "A", 
-                       direction = -1,
-                       breaks = c(0.0024,
-                                  0.0018,
-                                  0.0012,
-                                  0.0006,
-                                  0.0001),
-                       labels = c("0.0024",
-                                  "0.0018",
-                                  "0.0012",
-                                  "0.0006",
-                                  "0.0001")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "right")
-
-acc_pre / acc_covid
 ```
@@
 ```
-
-We find that the accessibility landscape deteriorated substantially during the pandemic, with accessibility dropping on average by almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% mean() %>% abs() %>% round()`%, but with large variations: some zones experienced changes in accessibility of only about `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% max() %>% abs() %>% round()`%, whereas the most affected zone saw a loss of accessibility of almost `r accessibility_taz_change %>% st_drop_geometry() %>% summarize(d = (.[[6]] - .[[5]])/.[[5]] * 100) %>% pull(d) %>% min() %>% abs() %>% round()`%.
-Figure \ref{fig:accessibility-changes-with-local-i} shows the changes in accessibility.
-Every zone is worse off after the closure of facilities during the pandemic, but some parts of the city seem to have been particularly affected.
-To better highlight these changes, we used a local indicator of spatial autocorrelation [@anselin1995local] to explore the pattern of change in accessibility.
-Twenty-four TAZs are flagged as having significantly large losses of accessibility (at $p\le 0.10$, without correcting for multiple comparisons).
-Those zones are highlighted in the figure, where it can be seen that they form more or less compact neighborhoods.
-Remarkably, the largest significant drops in accessibility are not downtown, but located in two cases in the industrial north of the city, in one case in an inner suburb above the escarpment, and lastly in a more suburban/exurban region in the south-west.
-
-```{r plot-accessibility-changes, echo=FALSE,  fig.align = 'center', fig.cap="\\label{fig:accessibility-changes}Changes in accessibility from pre-COVID-19 to during COVID-19."}
-## Analysis
-# Plot
-acc_change <- ggplot() +
-  #geom_sf(data = hamilton_boundary) +
-  geom_sf(data = accessibility_taz_change,
-          aes(fill = acc_d),
-          color = NA) +
-  geom_sf(data = population_age_income,
-          fill = NA) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_fill_viridis_c(name = "Change in accessibility",
-                       option =  "A", 
-                       direction = 1,
-                       breaks = c(-0.0008,
-                                  -0.0006,
-                                  -0.0004,
-                                  -0.0002),
-                       labels = c("-0.0008",
-                                  "-0.0006",
-                                  "-0.0004",
-                                  "-0.0002")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "none")
-
-acc_change_inset <- ggplot() +
-  geom_sf(data = accessibility_taz_change %>%
-            st_transform(crs = 4326),
-          aes(fill = acc_d)) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_fill_viridis_c(name = "Change in accessibility",
-                       option =  "A", 
-                       direction = 1,
-                       breaks = c(-0.0008,
-                                  -0.0006,
-                                  -0.0004,
-                                  -0.0002),
-                       labels = c("-0.0008",
-                                  "-0.0006",
-                                  "-0.0004",
-                                  "-0.0002")) +
-  guides(fill = guide_legend(title.position = "top")) +
-  xlim(-79.975, -79.775) +
-  ylim(43.2, 43.3) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "bottom")
-
-acc_change / acc_change_inset
-```
-
-```{r weights-matrix, include=FALSE}
-taz_nb <- poly2nb(accessibility_taz_change, queen = TRUE)
-taz_w <- nb2listw(taz_nb)
-```
-
-```{r moran-test-accessibility-changes, include=FALSE}
-moran.test(accessibility_taz_change$acc_d, taz_w)
-```
-
-```{r local-moran-i, include=FALSE}
-local_i <- localmoran(accessibility_taz_change$acc_d,
-                      taz_w, 
-                      alternative = "two.sided",
-                      p.adjust.method = "none") %>% 
-  data.frame() %>%
-  rename(p_val = Pr.z....0.) %>%
-  mutate(significance = ifelse(p_val <= 0.10, "sig", "n.s."),
-         significance = factor(significance,
-                               levels = c("n.s.", "sig")))
-
-summary(local_i)
-```
-
-```{r join-local-i-to-table, include=FALSE}
-accessibility_taz_change$significance <- local_i$significance
-```
-
-```{r plot-local-i, echo=FALSE, fig.cap="\\label{fig:accessibility-changes-with-local-i}Changes in accessibility from pre-COVID-19 to during COVID-19.
-Highlighted areas had significantly large changes in accessibility according to Local Moran's I." }
-ggplot() +
-  geom_sf(data = accessibility_taz_change,
-          aes(fill = acc_d,
-              color = significance,
-              alpha = significance),
-          size = 1) +
-  geom_sf(data = hamilton_boundary,
-          fill = NA) +
-  geom_sf(data = inset_box,
-          linetype = "dotted",
-          color = "red",
-          fill = NA) +
-  scale_fill_viridis_c(name = "Change in accessibility",
-                       option =  "A", 
-                       direction = 1,
-                       breaks = c(-0.0008,
-                                  -0.0006,
-                                  -0.0004,
-                                  -0.0002),
-                       labels = c("-0.0008",
-                                  "-0.0006",
-                                  "-0.0004",
-                                  "-0.0002")) +
-  scale_color_manual(values = c("n.s." = "lightgray", 
-                                "sig" = "black")) +
-  scale_alpha_manual(values = c("n.s." = 0.5, 
-                                "sig" = 1)) +
-  guides(color = "none", 
-         alpha = "none",
-         fill = guide_legend(title.position = "top")) +
-  theme_minimal() +
-  theme(axis.text = element_blank(),
-        legend.position = "bottom")
-```
-
-For the more suburban clusters of zones, the decrease in accessibility is derived from the closure of locations throughout the city reachable by car.
-In the cluster of central suburban zones for example, low-income households in the outer ring of zones that exhibit medium to high decreases in accessibility within this cluster appear to be largely auto-dependent in their tripmaking, which each each exhibiting between `r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(min(p_car)) *100)`-`r round(modes %>% st_drop_geometry() %>% filter(TAZUID %in% c(5143, 5135, 5121, 5110, 5102, 5092, 5063)) %>% summarize(max(p_car)) *100)`% of their modal split for car trips.
-This results in the parcels within these zones having a large number of potentially accessible locations in the travel time matrix.
-But by extension, the change in accessibility over the pre- and during-COVID-19 time periods is affected not only by the closure of service locations proximate to the zones, but also the locations in the central city.
-The zone with the greatest decrease in accessibility within this cluster (-0.0009) has a high rate of car trips and connects to the most facility locations in total as well as those that stayed open or closed.
-
-In the cluster to the south-west, the decrease in accessibility is predominately driven by the closure of a high level-of-service Community Meals provider.
-However, like the more central suburban zones, low-income households within this cluster are also between 90% to 100% auto-dependent in their trip-making in the TTS.
-The story is similar for the zone located in the north-west that exhibits the greatest decrease in accessibility.
-Here, low-income households responding to the TTS conducted 100% of their trips by car and, as a result, dwellings within this zone have access to the second-highest number of food bank and service locations within 15 minutes.
-However, this also means zonal accessibilities are greatly affected by the number of closures throughout the city.
-Finally, in the city's north end and north-east zones, low-income households exhibit a mixture of tripmaking behaviour in the TTS.
-Households in some zones take transit more often and one zone in particular has 100% of its trips by walking.
-For these zones, the decrease in accessibility tends to be a product of the closure of several inner-city food bank and service locations reachable by multiple modes.
-
-Just as the effects of the closures appear to have been uneven in space, they also seem to have had different impacts on various population segments.
-Using data on low income individuals by age drawn from the TTS, Table \ref{tab:accessibility-by-age} shows the estimated number of people in each age group by their level of accessibility before and during the pandemic.
-Here, it is important to note that the quartiles are relative: people in the top 25% of accessibility still have lower accessibility during the pandemic than before.
-In reality, every population group is worse off during the pandemic in terms of their accessibility to food banks and services in the City of Hamilton.
-However, some age groups were affected more.
-In terms of changes within the quartiles, the largest change for adults appears to be those moving from the first to the second quartile of accessibility during the pandemic.
-The story is generally similar for seniors, a greater number of whom are now in the second and third quartiles due to seniors facing a worse accessibility situation.
-Among those aged 18 and less, the largest change is in the number of children who were in the first and third quartiles before the pandemic and found themselves in the second accessibility class during the pandemic.
-However, when we compare the total population serviced before and during the pandemic, we see that a large number of children, adults, and seniors were no longer in the catchment areas of service locations (last row of Table \ref{tab:accessibility-by-age}).
-This accounts for the loss of population in the fourth quartiles for the different age groups during the pandemic.
-It is remarkable that despite the loss of population serviced, accessibility still declined for those still within the catchment regions of these services.
-
-These results suggest that, through a combination of the typical modes of transportation of lower income households and the spatial distribution of the population, the closure of emergency and communal food locations had a differential impact that more greatly affected the youngest and oldest among the population in low income households.
 
@@
 ```
-
-```{r include=FALSE, eval = FALSE}
-## Analysis
-# Calculate accessibility by age group and accessibility level:
-accessibility_by_age <- accessibility %>%
-  filter(timing == "During covid") %>% 
-  group_by(timing, 
-         accessibility_level) %>%
-  summarize(children = sum(children),
-            adults = sum(adults),
-            seniors = sum(seniors),
-            accessibility = sum(accessibility),
-            .groups = "drop")
-# The accessibility level is the magnitude of the accessibility loss: the top 25% faced the smallest losses in accessibility, and the bottom 25% faced the greatest losses in accessibility. 
-```
-
-```{r include=FALSE, eval = FALSE}
-cbind(accessibility_by_age %>% 
-        filter(timing == "Pre-covid") %>% 
-        select(children:accessibility) %>% 
-        rename(accessibility_pre = accessibility),
-      accessibility_by_age %>% 
-        filter(timing == "During covid") %>% 
-        transmute(accessibility_covid = accessibility)) %>%
-  mutate(pct_children = round(children/sum(children) * 100, 1),
-         pct_children = paste0("(", pct_children, "\\%)", sep = ""),
-         pct_adults = round(adults/sum(adults) * 100, 1),
-         pct_adults = paste0("(", pct_adults, "\\%)", sep = ""),
-         pct_seniors = round(seniors/sum(seniors) * 100, 1),
-         pct_seniors = paste0("(", pct_seniors, "\\%)", sep = ""),
-         children = round(children) %>% 
-           prettyNum(big.mark = ","),
-         adults = round(adults) %>% 
-           prettyNum(big.mark = ","),
-         seniors = round(seniors) %>% 
-           prettyNum(big.mark = ","),
-         children = paste0(children, " ", pct_children, sep = ""),
-         adults = paste0(adults, " ", pct_adults, sep = ""),
-         seniors = paste0(seniors, " ", pct_seniors, sep = ""),
-         change = (accessibility_covid - accessibility_pre) %>% 
-           round(5),
-         accessibility_pre = round(accessibility_pre, 5),
-         accessibility_covid = round(accessibility_covid, 5)) %>%
-  transmute(children,
-            adults,
-            seniors,
-            accessibility_pre,
-            accessibility_covid,
-            change) %>%
-  kable("latex",
-        booktabs = TRUE,
-        align = ("cccccc"),
-        col.names = c("Children (age $\\le$ 18)", 
-                      "Adults (19-64)", 
-                      "Seniors (age $\\ge$ 65)", 
-                      "Before COVID-19", 
-                      "During COVID-19", 
-                      "Difference"),
-        caption = "\\label{tab:accessibility-by-age}Accessibility by age group among members households with incomes less than CAD40,000.",
-        escape = FALSE) %>%
-  kable_styling(latex_options = c("striped", "scale_down")) %>%
-  add_header_above(c("Population" = 3, "Accessibility" = 3)) %>%
-  footnote(general = "Population values have been rounded")
-```
-
-# Conclusions
-
-Food insecurity is a significant issue for many Canadian households and while emergency and community food services can provide some relief, the COVID-19 pandemic has in all probability increased food insecurity for many households.
-To compound matters, the pandemic has also resulted in major disruptions, including to employment, mobility alternatives, and to emergency and community food services.
-In response, this research has sought to better understand accessibility to food banks and food service locations, as well as how the closure of some locations over the pre- and during-COVID time periods affected the potential for low-income households to reach these amenities.
-
-Previous work has noted the important role of geography alongside other socio-economic and demographic indicators in household access to healthy food.
-The present papers is, to the best of our knowledge, among the first studies to focus on the geographic component of accessibility to emergency food services [@allen2021Changes].
-Using the balanced floating catchment area method to account for population demand and congestion effects at service points, we estimated multi-modal accessibility to emergency and community food service locations for low-income households.
-The weighting of travel time estimates by the modal split in different zonal geographies tailors the results to patterns of travel behaviour captured in the regional travel survey.
-Moreover, our parcel-level analysis presents a disaggregate approach to estimating accessibility based on the locations of residential parcels and dwellings.
-Beyond accounting for the inflation of demand that occurs in traditional floating catchment area methods, our application of the balanced floating catchment area approach offers a novel analysis of accessibility to emergency and community food services that is sensitive to the locations of low-income households, information on their age distributions and typical trip-making behaviour, the locations of services, their operations over time, and the characteristics of the city's multi-modal transportation network.
-
-Our results show that while accessibility levels were lower in the city's more car-oriented suburban and rural areas to begin with, the closure of 14% of the city's emergency and community food service locations during the pandemic resulted in an overall decrease in accessibility across the city.
-However, these effects were not uniform over space or for different population groups.
-Since the balanced floating catchment area method takes into account changes in demand and congestion for service providers, the closure of some services reverberates throughout the catchment areas of the whole city.
-For some suburban zones, the closure of a relatively high level-of-service location results in the remaining services being spread over a larger population.
-In others, high auto dependence for trips leads to decreases in accessibility that accumulate due to the loss of several locations initially reachable within 15 minutes by car.
-Reductions in accessibility in the city's more urban north end, where low-income households conduct higher proportions of trips by transit and walking, emphasize the importance of geographic proximity in the potential to reach service locations for these residents.
-Beyond geography, the results also highlight the differential impact of closures during COVID-19 on population groups, with seniors and children being the two most impacted groups.
-
-It is important to note that the degree to which our low-income cut-off of CAD\$40,000 reflects food insecurity in the different zones of the study area is not known.
-We also consider all emergency and community food service locations equally.
-Information on differential capacities at food provider locations is currently not collected, and given the closure of facilities was not possible to obtain.
-Data on services, such as the number of meals served, could be used to refine the analysis in the future.
-Moreover, while the travel survey allows us to model multi-modal accessibilities that align with travel behaviour observed in the travel survey and capture differences in accessibility by age categories, the use of travel survey data for modelling food insecurity also has its limitations.
-Research into the population weighting methods used in the TTS note that the survey may under-count the lowest- and highest-income households in the survey study region, although the magnitude of this under-counting is unknown, and approximately 20% of respondents to the 2016 survey did not report their income [@rose2018transportation].
-In that regard, the modal splits of low-income households observed through the TTS data may not accurately reflect the travel behaviours of food insecure populations, and our estimates might in fact be somewhat conservative if those households who do not report income rely more on walking and/or transit for their mobility needs.
-
-In the absence of information regarding how food insecure households travel to food banks and related services, we examined accessibility to food banks using a 15 minutes (one-way) travel time threshold.
-The fact that we must rely on a standard created for accessible drinking water in the developing world only serves to highlight the tragedy of food insecurity in an affluent country like Canada.
-More broadly, it points to the absurd need to understand how a bad situation was made worse by the pandemic: in effect, the analysis reveals that disparities in the need for emergency and community food services predated the pandemic, that the pandemic contributed to the deterioration of these services, and that populations already in distress, particularly children, ended up in an even more adverse state.
-How much worse, it is impossible to say, mainly because there is also a dearth of information, let alone standards, regarding *acceptable* or *sufficient* level of service when it comes to emergency food services.
-
-While on the one hand this work suggests that inequities in the accessibility to emergency and community food services could be improved through accessibility standards that promote changes in the geographic distribution of service locations and transportation network characteristics, in fact, we would argue that the standard should be that no household faced food insecurity.
-As others have noted [e.g., @poppendieck1999sweet; @men2021food] the root of food insecurity is income poverty and unless it is eliminated, there will continue to be a place for emergency food and community food services.
-In addition to providing food, these services satisfy social needs by offering a social setting for seniors or by helping to connect households in need with longer term supports.
-From a food security perspective, on the other hand, these services should work only as a short term solution, and not as a semi-permanent feature of life for some of our fellow human beings.
-From a human rights perspective, long-term reliance on emergency food services should be as unacceptable in Canada as lack of clean drinking water within 30 minutes is elsewhere.
-Thus, while our analysis is valuable to map the suffering caused by food insecurity, from a policy perspective, maintaining a robust social safety net that includes Employment Insurance and paid sick days are better tools to reduce this suffering than increasing the accessibility of emergency food services for food insecure populations.
-
-# References {#references .unnumbered}

### 2021-09-02 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
R2-Created highlights file, made small edits requested by editorial office on submission
Created png version of figures and separate file for tables as requested by editorial team on submission of R2

@@
 
+```{r directory-for-figures, include=FALSE}
+ifelse(!dir.exists(file.path("./figures")), dir.create(file.path("./figures")), FALSE)
+```
+
 ```{r invoke-data, include=FALSE}
@@
 ```
+
+**Keywords**: Food insecurity; Emergency and community food services; Food banks; Accessibility; Balanced Floating Catchment Areas; COVID-19; Hamilton; Ontario; Reproducible research
+
+\newpage
 
@@
 fb / fb_inset
+
+ggsave("./figures/figure-1.png")
 ```
@@
 lih / p_lih
+
+ggsave("./figures/figure-2.png")
 ```
@@
 acc_sensitivity
+
+ggsave("./figures/figure-3.png")
 ```
@@
 los_pre / los_covid
+
+ggsave("./figures/figure-4.png")
 ```
@@
 lc / lc_inset
+
+ggsave("./figures/figure-5.png")
 ```
@@
 acc_pre / acc_covid
+
+ggsave("./figures/figure-6.png")
 ```
@@
 acc_change / acc_change_inset
+
+ggsave("./figures/figure-7.png")
 ```
@@
         legend.position = "bottom")
+
+ggsave("./figures/figure-8.png")
 ```

### 2021-10-09 README.Rmd
Paper accepted for publication in Social Science and Medicine
Title updated in README.Rmd

@@
 
-# Changes in accessibility to food banks and food services during COVID-19 and implications for low income populations in Hamilton, Ontario
+# Changes in accessibility to emergency and community food services during COVID-19 and implications for low income populations in Hamilton, Ontario
 
@@
 
-Paper submitted to *Social Science and Medicine*
+Forthcoming in *Social Science and Medicine*
 

### 2021-10-13 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Updated README after publication in Social Science and Medicine

@@
 This paper follows reproducible research recommendations [see @brunsdon2020opening], and the research was conducted using open source tools for transportation analysis [@lovelace2021open].
-The data and code needed to reproduce the analysis are available in an (anonymous) [Drive Folder](https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing).
+The data and code needed to reproduce the analysis are available in a [public repository](https://github.com/paezha/Accessibility-Food-Banks-Hamilton).
 
@@
 
-Data have been prepared for sharing in the form of an open data product [see @arribas2021open] available in an (anonymous) [Drive Folder](https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing).
+Data have been prepared for sharing in the form of an open data product [see @arribas2021open] available in a [public repository](https://github.com/paezha/Accessibility-Food-Banks-Hamilton).
 The contents of the data package are described next.

### 2021-10-13 README.Rmd
Updated README after publication in Social Science and Medicine

@@
 
-Forthcoming in *Social Science and Medicine*
+Social Science and Medicine (2021) https://doi.org/10.1016/j.socscimed.2021.114442
 
